### 2023-05-01 Article/Trajipaq_Covid_Draft4.Rmd
update

---
title: 'COVID-19 Disparities in Quebec (Canada): A Comparative Analysis of Exposure
  Risks, Consequences, and Access to Resources among Immigrants and Natives'
author:
- Vissého Adjiwanou ^[corresponding author - adjiwanou.vissého@uqam.ca]
- Département de Sociologie, Université du Québec à Montréal (UQAM)
- Georges NGalé Tchango (Département de démographie, Université de Montréal)
- Solène Lardoux (Département de démographie, Université de Montréal)
- Nima Zahedinameghi (Département de Sociologie, UQAM)
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  word_document:
    keep_md: yes
  html_document:
    toc: yes
    toc_depth: 3
  template: svm-latex-ms2.tex
  pdf_document: default
  citation_package: natbib
bibliography:
- covid_Nima.bib
- Covid_Canada.bib
- Covid_Family.bib
---
<!--
# Journal: family issues - SAGE
https://journals-sagepub-com.proxy.bibliotheques.uqam.ca/author-instructions/JFI

https://fishandwhistle.net/post/2020/getting-started-zotero-better-bibtex-rmarkdown/
https://www.zotero.org/styles?fields=sociology
-->

\newpage

# Abstract

The immigrant population is one of the most vulnerable in Canada, with a relatively high risk of unemployment, a low social support system, and linguistic barriers. However, their experiences with public health interventions during the COVID-19 pandemic have not been adequately explored. The purpose of this article is to analyze Covid-19 exposure risk and its family implications for both natives and immigrants in the province of Quebec. Additionally, we analyze immigrants' access to government aid and compare it to their native-born counterparts. The study consists of survey data collected from 1,000 immigrants and 500 non-immigrants Quebecers between August and November 2020. Our descriptive and exploratory analysis reveals significant differences between the two groups regarding exposure risk, family consequences, and access to resources. We conclude that the current pandemic exacerbates health disparities in the province, with particularly dire repercussions for immigrants’ families. We also discuss several socio-cultural strategies to mitigate inequality and recommend policies to address the needs of vulnerable populations.

**Keywords**: Immigrants, exposure risk, government aid, health disparities, COVID-19, family disruptions

<!--

Using  script in Rmarkdown

http://zevross.com/blog/2014/07/09/making-use-of-external-r-code-in-knitr-and-r-markdown/

Managing my Zotero

https://forums.zotero.org/discussion/35712/removing-pdfs-to-free-up-space

https://www.zotero.org/support/sync

https://www.zotero.org/support/preferences/sync#file_syncing

https://www.zotero.org/support/sync#webdav

https://www.zotero.org/support/preferences/advanced#linked_attachment_base_directory

http://zotfile.com/

# Abstract

A ce jour, la pandémie du coronavirus a touché près de 100000 personnes au Québec et entrainé plus de 6000 décès. Cela fait de la province l'une des plus touchées au Canada et l'une des régions les plus affectées au monde. Bien que la majorité de ces décès ait eu lieu dans les maisons pour personnes âgées, et donc a touché plus cette tranche de la population, son influence globale sur les individus n'est pas à négliger. En effet, outre les conséquences directes de la pandémie, les mesures prises par les différents gouvernements à travers le monde ne sont pas sans conséquences sur les familles. Il est démontré que ses conséquences affectent négativement en priorité les couches les plus défavorisées de la population. Les immigrants se concentrent dans bien des cas dans cette couche de la population. Cette étude vise ainsi à analyser l'influence de la pandémie sur les immigrants et leurs enfants au Québec, à examiner les facteurs socio-culturels qui permettent d'atténuer son impact négatif.

To date, the coronavirus pandemic has affected nearly 100,000 people in Quebec and caused more than 6,000 deaths. This makes the province one of the most affected in Canada and one of the most affected regions in the world. Although the majority of these deaths took place in homes for the elderly, and therefore affected this segment of the population more, its overall influence on individuals is not to be overlooked. In fact, in addition to the direct consequences of the pandemic, the measures taken by various governments around the world are not without consequences for families. It has been shown that its consequences primarily affect the most disadvantaged segments of the population negatively. Immigrants are in many cases concentrated in this segment of the population. This study thus aims to analyze the influence of the pandemic on immigrants and their children in Quebec, to examine the socio-cultural factors that make it possible to mitigate its negative impact.

À soumettre au Canadian Population Studies

Autre projet possible: intégration des immigrants et covid19 au Québec

**Keywords** :

-->

\newpage

# Introduction / background

The COVID-19 pandemic has highlighted the vast ethnic, social, economic, and health disparities in high-income countries, including Canada. One of the most vulnerable segments of these affluent societies is the immigrant population, that has become widely diversified with more and more people coming from developing countries, and for whom integration into their new society may be more difficult. The coronavirus pandemic brings a new challenge for these immigrant populations, directly or indirectly through the tensions between the ethnic groups that it helps to generate [@coates2020Covid19]. These immigrants face language barriers, low social capital, housing insecurity, and severe poverty, rendering them particularly susceptible to the effects of the current pandemic. Immigrants are also more likely to be employed as essential workers where social distancing, sanitary precautions, and self-isolation, measures put in place to mitigate the pandemy, are challenging [@reid2021Migrant; @yi2021Health; @zhang2021COVID19]. More importantly, migrants may encounter various forms of discrimination, leading to social exclusion and restricted access to healthcare and social services [@spiritus-beerden2021Mental]. It is therefore necessary to understand how this pandemic affects these immigrant populations differently and the native populations in Quebec. In doing so, this article will reveal a better understanding of how the current pandemic aggravates existing inequalities and challenges in the province for immigrants and their descendants.

<!-- The province of Quebec has one of the highest concentrations of immigrants in Canada, including permanent residents, temporary residents, and refugees [@governmentofcanada2017Daily; @immigration2005Recent]. In 2016, 1.2 million immigrants were living in Quebec, an increase from 8% in 1990 to 13.7% of the population in 2016 (Institut de la statistique du Québec, 2020). <!--As of 2019, the increase in net new residents (60,700 versus 42,200 in 2018) and lower interprovincial migration losses (-1,500 versus -6,100 in 2018) contributed to Quebec's population growth. -->Since the start of the pandemic, most of the confirmed cases were identified in Montreal and areas with high immigrant populations [@miconi2021Ethnocultural; @shields2020COVID19]. The coronavirus pandemic has infected nearly 450,000 people in Quebec and has resulted in more than 11,000 deaths (https://www.quebec.ca/sante/problemes-de-sante/a-z/coronavirus-2019/situation-coronavirus-quebec, consulted on octobre 29, 2021). This makes it the worst affected province in Canada and one of the worst affected regions in the world during the first wave of the pandemic. Although the majority of these deaths took place in elderly's homes, and therefore affected more this segment of the population, its overall influence on individuals is not to be overlooked. It has been shown that, on the one hand, immigrants and their families have a higher risk of being exposed to the pathogen, and, at the same time, they are more vulnerable to racism and discrimination and face barriers to accessing social support [@2021CRARR] and information, especially  for those who didn't speak or understand the two official langages. <!--In addition to the direct consequences of the pandemic, the measures taken by various governments around the world are also not without consequences on families. -->Several studies have thus shown excess mortality in ethnocultural minority populations [@bertocchi2020Covid19], higher unemployment, greater loss of income and negative consequences for children and families in several countries. This differentiated effect can be explained in different ways. On the one hand, in terms of the direct consequences of the pandemic, the situation of precariousness and greater poverty among immigrants exposes them more to the more harmful consequences of the disease on their state of health. Likewise, immigrants often work in jobs where social distancing is difficult to enforce. On the other hand, indirectly, the solutions to the crisis affect immigrants and natives differently. The complete shutdown of tye economy with its set of adjustments has greatly impeded the employability of immigrants, especially newcomers who have not yet grasped the workings of society and who cannot benefit from state subsidies. But, it is also shown that these effects are not uniform on the whole of the immigrant population, itself, very diverse.

Despite the growing focus on the differential impacts of the COVID-19 on immigrants and non-immigrants segment of the population, there is still very little knowledge about how the pandemic and its related public policy impact the lives of immigrants and their families. The existing health datasets (largely used to assess this effect) are not necessarily representative of the immigrant population and do not offer an accurate picture of the complex social status and health outcomes among this group [@choi2021Studying; @gagnon2021Immigration; @khan2015Visible]. In most cases, these datasets lack data on the residence length of immigrants and combine newcomers with long-term settlers [@edge2013Discrimination; @hayward2021Clinical]. As a result, they may not allow a detailed assessment of the nature and extent of risk exposure in sub-segments of populations, as well as whether interventions are reaching those most at risk. In addition, the family-level impacts of the pandemic are still poorly understood. While a body of research has explored the association between the pandemic and its potential effects on intimate partner violence and marital dissolution [@bhatnagar2021Navigating; @koshan2020COVID19; @morrison2021Chapter; @prime2020Risk], few studies have examined negative familial repercussions in the context of immigration status. There is ample evidence that sustained family relationships greatly enhance immigrants' lives on several fronts, including fostering economic prosperity [@marger2001Sociala], psychological well-being [@beiser2002Povertya], social network development [@evansluong2017Familya], and educational success for children [@munroe-blum1989Immigranta]. The family is, therefore, a key component of long-term social mobility and economic success of immigrants.

<!--Faced with the pandemic, the governments of Canada and Quebec have implemented a number of measures to reduce its impact. These policies can be broken down at two levels, direct support to families, including the Canadian Emergency Benefit, which can provide employees and self-employed workers affected by Covid-19 with a financial benefit of up to $ 2,000. On the other hand, the government supports many companies to keep their employees or supports them directly when the revenues from their activities have fallen considerably. While these measures allow families to cope better with the crisis, the fact remains that the changes brought about by this crisis can affect other dimensions of life. How these measures make a difference in the face of the pandemic remains to be seen. -->Based on an original biographic survey collected around the first wave of the pandemic in august-november 2020 on immigrants and the native population of Québec, this study mainly aims to understand the family exposure, consequences, and responses of the pandemic on immigrant and native Quebec populations. We assess a whole range of different measures of each outcome. We divided the consequences in three dimensions, 1) access to ressources, 2) ability to perform duties and 3) employment and income. We seek also to analyse in what ways these outcomes differ according to the immigrants length of residence. By doing so, this study provide a clear and global picture of the situation of different population groups in Quebec in the face of the COVID-19. <!-- To respond to these questions, we will use data collected around the first wave of the pandemic in august-november 2020. Our study aims to fill the gap by analyzing group-specific differences in exposure risk, family consequences, and access to resources among immigrants and Quebecers. We also seek to shed light on several socio-cultural strategies to mitigate inequality and recommend relevant policy solutions. The detailed findings will help policymakers address health inequalities and improve equity in Quebec while developing adequate measures and programs. -->

<!-- Introduction  -->

<!-- Like the rest of Canada, Quebec has experienced a substantial increase in its immigration in recent years. The immigrant population in the province thus increased from 8% in the 1990s to 13.7% of the population, or 1.2 million individuals in 2016 (Institut de la statistique du Québec, 2020). This immigration is also unique in that it has become widely diversified with more and more people coming from developing countries, and for whom integration into their new society may be more difficult. The coronavirus pandemic brings a new challenge for these immigrant populations, directly or indirectly through the tensions between the ethnic groups that it helps to generate [@coates2020Covid19]. It is therefore necessary to understand how this pandemic affects these immigrant populations differently and the native populations in Quebec. -->

<!-- The coronavirus pandemic has affected nearly 100,000 people in Quebec and has resulted in more than 6,000 deaths. This makes it the worst affected province in Canada and one of the worst affected regions in the world. Although the majority of these deaths took place in homes for the elderly, and therefore affected this segment of the population more, its overall influence on individuals is not to be overlooked. In fact, in addition to the direct consequences of the pandemic, the measures taken by various governments around the world are also not without consequences on families. Several studies have thus shown excess mortality in ethnocultural minority populations [@bertocchi2020Covid19], higher unemployment, greater loss of income and negative consequences for children and families in several countries. -->

<!-- This differentiated effect can be explained in different ways. On the one hand, in terms of the direct consequences of the pandemic, the situation of precariousness and greater poverty among immigrants exposes them more to the more harmful consequences of the disease on their state of health. Likewise, immigrants often work in jobs where social distancing is difficult to enforce. On the other hand, indirectly, the solutions to the crisis affect immigrants and natives differently. The complete shutdown of the company with its set of adjustments has greatly impeded the employability of immigrants, especially newcomers who have not yet grasped the workings of society and who cannot benefit from state subsidies. But, it is also shown that these effects are not uniform on the whole of the immigrant population, itself, very diverse. -->

<!-- Faced with the pandemic, the governments of Canada and Quebec have implemented a number of measures to reduce the impact of the pandemic. These policies can be broken down at two levels, direct support to families, including the Canadian Emergency Benefit, which can provide employees and self-employed workers affected by Covid-19 with a financial benefit of up to $ 2,000. On the other hand, the government supports many companies to keep their employees or supports them directly when the revenues from their activities have fallen considerably. While these measures allow families to cope better with the crisis, the fact remains that the changes brought about by this crisis can affect other dimensions of life. How these measures make a difference in the face of the pandemic remains to be seen. -->

<!-- This study mainly aims to understand the family consequences of the pandemic on immigrant and native Quebec populations. Therefore, we will seek to analyse in what ways its consequences differ according to the region of origin of immigrants, length of residence, age and sex. -->

# Current body of literature and knowledge gaps

The literature on the racial and ethnic health disparities during the COVID-19 crisis can be divided into two main categories: studies addressing the socio-economic determinants of health and those examining the delivery of health interventions and policies. The first category is primarily comprised of research that focuses on the link between socio-economic status (SES) of high-risk minority groups and the differential impact of the pandemic [@drefahl2020Populationbased; @greenaway2020COVID19; @lopez2021Racial; @pacheco2020Job; @sundaram2020Individual]. For instance, an inter-country study has found that Middle Eastern, African, and Nordic immigrants in Sweden have a higher risk of being hospitalized and even dying from COVID-19 when compared to the Swedish born population. In the same study, the SES, proportion of working-age households, and neighborhood density were significantly associated with the mortality rate [@rostila2021Disparities]. This trend of research demonstrates that racial and ethnic minorities are more likely to experience stress factors, not only due to the hazards of infection but also because of economic and social marginalization [@alahmad2020COVID19; @choi2020Buffering; @endale2020COVID19; @fitzpatrick2020Fear; @serafini2021Psychological]. The intersectionality of race, sex, and health conditions has also been explored in the Canadian context of the pandemic. @ingen2021Neighbourhoodlevel demonstrated that Ontarian neighborhoods with immigrant and racialized populations have a disproportionate share of COVID-19 infections. In addition, areas with high SES report a lesser number of COVID cases than low SES ones [@ingen2021Neighbourhoodlevel]. Another study conducted in Quebec confirms the strong association between immigrants' backgrounds, SES, and COVID-19 related psychological stress [@miconi2021Ethnocultural]. The most recent reviews on the topic suggest that immigrants disproportionately live in crowded and inadequate housing and are employed in essential industries, leaving them with severe income insecurities and poor physiological/psychological well-being [@chew2020Clinical; @davis2020Impact]. Studies on SES and health disparities have emphasized that this situation is further complicated for immigrant women, constrained by gender and afflicted by a significant power imbalance in the household and workplace [@alarcao2021Intersections; @esegbona-adeigbe2020COVID19; @jehn2021COVID19; @mo2020Differential; @nardon2021Skilled; @sabri2020Effect; @st-denis2020Sociodemographic]. According to Lightman, immigrants and women of color are overrepresented in long-term care facilities in Canada, where COVID infection rates have been exceptionally high. They are also subject to exclusionary social and economic practices [@lightman2021Caring].

The second category of the literature is primarily comprised of policy analyses, research on public health communications, examining the efficacy of COVID initiatives, the optimal allocation of government aids, and ethical issues associated with applying new health interventions to vulnerable populations and their communities [@aragona2020Negative; @cross2021Coronavirus; @desai2020COVID19; @doyle2020Migrant; @gilman2020Modelling; @hargreaves2020Targeting; @langellier2020Policy; @nezafatmaldonado2020Engaging; @silverman2020Ethics; @wilson2020Ethics]. Some of the questions that this body of research addresses include: how best to respond to COVID-19 in immigrant communities? How do public health policies impact the socio-economic status and psychological well-being of different sub-populations? What is the structure of the planning and implementation of government aid programs during the current crisis? And, how can we design public health communication strategies in a way that is culturally sensitive and inclusive?

There are several knowledge gaps in the existing literature. Most of the cited studies do not include household data and family structure of immigrants as variables that affect exposure risks and health outcomes of COVID-19. Consequently, the relationship between immigrant status and COVID-19 is not accounted for by household size, and family characteristics (i.e., income, education, religion). A key reason is that most of the health data used in such research projects are acquired from national registers or administrative databases, which provide little information on immigration status and kinship. However, as the study of Thomeer and her colleagues theorizes, the extent of adverse social and economic impacts of COVID-19 is significantly conditional on the composition of an immigrant's family [@thomeer2020How]. Households that include a spouse/partner, children, or multigenerational family members most likely experience a different level of social, psychological, and medical vulnerability than single or nuclear ones. Therefore, studies of immigrant populations should account for the breakdown of family configuration to better understand the mechanisms linking social determinants and health outcomes. Another knowledge gap concerns the current state of Canadian studies of COVID-19 and its public health crisis. There is a dearth of literature on the challenges faced by immigrants and their families in the country. Although the burden of COVID-19 infection is disproportionately higher among these marginalized groups (St-Denis, 2020; Sundaram et al., 2020), research on the issues they are experiencing is sparse and often limited to one aspect. <!--Likewise, there is limited focus on the effectiveness of COVID-19 crisis management for immigrants at the policy level. In response to the mounting cases of COVID, the federal government and provinces and territories of Canada have launched multi-pronged public health initiatives to protect society from the epidemic (Migone, 2020). However, many open questions remain unanswered, including, do interventions efficiently reach the migrant groups? What are the likely effects of each program on different immigrant communities? How can an intervention be tailored to the needs of a particular migrant group and their families? -->

# Theoretical framework 
Family as a unit of analysis can be particularly fruitful when studying health inequalities in the context of a societal-wide crisis. Through the lens of the COVID-19 pandemic, the following points illustrate why exploring family dynamics provides additional insight into the underlying mechanisms of health disparities. First, the family is the most fundamental unit of social organization. It facilitates access to resources and the transmission of social and psychological capital from one generation to the next. Therefore, a better analysis of how crises and adverse life events affect people requires a better understanding of how families relate to each other and interact with external forces. Second, during public health emergencies, families may be thrown into different structures due to unprecedented financial, mental, and physical challenges. The risk of family dissolution, for example, can be heightened when members must spend unusually long periods together at home and have limited opportunities to engage with friends and the wider community. Conversely, families can develop closer bonds through self-sacrifice, shared responsibilities, and increased care and support. As the dependency status of seniors increases, the elderlies are also more likely to be integrated into formerly nucleus families. We posit that the extent to which a family can adapt its functions to meet new challenges is a predictor of the well-being of its members and the resilience of its broader population. Third, the family is a crucial link to social institutions, such as the health care system, schools, and government agencies whose decisions have significant downstream impacts. Families are viewed by many public-level entities as a reference when allocating resources, enforcing new standards, and responding to emerging needs and demands. From this perspective, the family operates at the nexus of individual and societal structures, embedding the tripple effects of public policy in its most consequential ways. Finally, the family has a crucial role to play in determining the life course of its members. As COVID-19 elevates the importance of familial bonds, the post-pandemic period might be marked by an increased influence of families on the lives of their members and the social structures they inhabit.

The above conceptualization echoes several other approaches to studying long-term health inequalities. For instance, life course theory emphasizes the significance of family ties and social environment as they structure prospects for individuals across their life span (Elder, 1977; Elder Jr., 1998; Macmillan, 2005; Settersten et al., 2020). Likewise, social capital theory and the stress-buffering hypothesis consider family systems a crucial factor in health and well-being (Choi and Jun, 2020; Cohen and McKay, 1984; Cohen and Wills, 1985; Harpham et al., 2004; Kawachi and Berkman, 2001; Lavee et al., 1987; Mandelbaum et al., 2020). They show that the family can offer tangible (e.g., food, shelter, and money) and intangible resources (e.g., support, social status, psychological comfort, and institutional advantages), protecting its members from adverse life events. Similarly, the social-ecological model and the theory of fundamental causes highlight how the SES, familial and extrafamilial factors pattern people's life trajectories and health outcomes (Barkan and Rocque, 2018; Clouston and Link, 2021; Figge et al., 2018; Kim, 2020; Libman et al., 2012; Ungar, 2015). However, all these approaches are ostensibly focused on life periods where the general social context is relatively stable. In this article, we present our conceptualization of the family in times of unprecedented social upheaval. Specifically, our framework seeks to characterize family dynamics as a suitable lens to analyze health inequalities in the face of a virulent public health crisis.

# Methodology

## Data

The data for this study come from an original collection of biographical data on immigrants and the native population of Quebec to analyse their participation in Quebec society in various fields (social, economic, cultural, etc.). The immigrant population is limited to recent immigration that has taken place over the past 15 years. This survey has several components that retrace the life course of individuals over a long period of time. The pandemic component uses a reduced version of the [harmonized questionnaire on exposure and the impact of Covid19 on the family (CEFIS)](https://www.phenxtoolkit.org/toolkit_content/PDF/CPTS_CEFIS_FullSurveyEng.pdf). The questions are divided into three parts, on exposure to the disease, its consequences and the benefit of the policies put in place by both the federal and provincial governments. The other parts of the questionnaire focus on the characteristics of the respondents, their immigration history, their educational and family background. In short, it is a detailed questionnaire that covers the life of immigrants from birth and especially over the past 15 years.
The data collection lasted three months, from August to November 2020. In total, 1,500 people were surveyed, including around 1,000 immigrants and 500 natives throughout Quebec.

### Dependent variables

We distinguish three groups of dependent variables:

**1.	Disease exposure variables**: these are measured by six questions that seek to document whether the individual through work or a family member may have been exposed to the disease. For example, it is asked if a family member has had symptoms of the disease, or if a member has been hospitalized due to the contraction of the disease.

**2.	Variables on the family consequences of the disease**: we measure three dimensions of consequences. The first group of measures assess the consequences on accessing ressources by the respondent or any member of his family. The second group of measure assess whether the respondent or any family member have difficulties to perform usual duties toward any member of the family. Finally, the last group of measures measure the consequences of the pandemic on employment and income of any family member.

**3.	Variables on government assistance**: Finally, government support is measured by the following two questions that assessed whether the respondent or any family member has benefited of the temporary assistance for workers from the Government of Quebec (PATT) or of the Canadian Federal Government Emergency Benefit (CEP). The table XXX in the apendix presents the full set of indicators of the dependents variables

### Main independent variable
The main independent variable is immigrant status coupled with visible minority status. This variable takes the following modalities: 1 if the respondent is a non-racialized native, 2 if he is a non-racialized immigrant, 3 if he is a racialized immigrant, and 4 if the respondent is a racialized native or an immigrant who has not answered the visible minority question. The other variables in the analysis are: age, sex, length of residence, minority status, size of social network.

## Analysis methods

We use two main methods of analysis, a descriptive method and an explanatory method. In the descriptive method, we consider each of the independent variables in isolation and describe how they vary depending on the main independent variable. The explanatory methods goes further by assessing whether the effect remain when controlling for other covariates. Thus, by controlling for the other explanatory variables, we determine the effect of immigrant status on exposure to the pandemic, on its consequences and on the fact of benefiting or not of government assistance. In a second series of modelling, we assess whether the effect that we observed varies by the immigrant's length of residence or by marital status. We use either a logistic regression on the outcome variables that are binary and an ordonned logistic regression for the outcome that are measure with a 5-point Lickert scale. 
<!-- PUT THIS AS A LIMITATION IN THE CONCLUSION
To be able to benefit from government benefits, one must meet a number of conditions, including the loss of job due to Covid-19. As a result, the beneficiaries are not the general population but those who have lost their jobs and this may itself be due to immigrant status. Thus, for model 3, we will use Heckman's models by first estimating the fact of having lost or not a job and then, the fact of receiving or not the government benefit.
-->

# 3. Results

## 3.1. Sample distribution

## 3.2. Exposure by immigration status

Figure 1 presents the exposure to the pandemy accroding to the immigrants and minority visible status. For all the six measures of exposure, it appears that the majority group is the most exposed to the COVID-19 in all instance. The big difference between the groups is about the exposure trhough working outside. Around 50% of the member of this group or a family member work outside the home at the time of the survey. This percentage is only 30% for the racialised immigrants and 25% for the non-racialised immigrants.

![Figure 1.1: Exposure](Figures/Exposition1.png){width=60%}

![Figure 1.2: Explicative model](Figures/Exposure.jpeg)

## 3.3. Consequences by immigrations status

The Figure 2 presents the consequences of the COVID-19 for a family member to access to ressources. It shows that in all accounts but one (getting medecine), the racialised immigrants express more diffciult to assess ressources. Around 20% of them stated having difficulty to get health care compare, around 10% for obtaining food and 7% for obtaining family insurrance. Comparatively, these proportion are respectively 13%, 7% and 4% for people and thier family member in the majority group and 19%, 3% and 4% respectively among the racialised immigrants.

![Figure 3.1: Consequences on accessing ressources](Figures/Famille1.png){width=80%}

![Figure 3.2: Explicative model](Figures/Ressources.jpeg)

## 3.4. Consequences on performing usual duties

The Figure 4 compares the response from the three groups on performing usual duties. It appears that for all groups the situation caused by the COVID-19 has inexpectedly improved the situation in various way, irrespective of the immigrants status of the respondents.

![Figure 4.1: Consequences on performing usual duties](Figures/Entente1.png)

![Figure 4.2: Explicative model](Figures/Abilities.jpeg)

## 3.5. Consequences on employment by immigration status

The consequences on employment are mixed, with the severe consequences more prevalent among the racialised immigrants. The Figure 3 shows for instance that close to 45% of the racialised immigrants have stated  that their family income has decreased compare to just 33% of the majority group and 37% of the non-racialed immigrants. In addition, 11% of the racialised immigrants reveals having loss their job compared to 9% among the non-racialised immigrants and 2% among members of the majority group. At the opposite, the members of the majority groups are more likely to report a reduction of the working hours or for stopping momentarily their work.

![Figure 5.1: Consequences on employment and income](Figures/Emploi1.png)

![Figure 5.2: Explicative model](Figures/Employment.jpeg)

## 3.7 Help received by immigrations status

Finally, we compared the different group on benefitting of the government assistance. It appears that racialised immigrants with 35% of the case have expressed having received the Canada Emergency Response Benefit wheras this proportion is 31% for the non-racialised immigrants and 25% for the member of the majority group. For the province Temporary Aid for Workers Program, around 5% of the respondents have received it.

![](Figures/Aide1.png)

![Figure 3.2: Explicative model](Figures/Government.jpeg)

# Conclusion

The preliminary results of this study show that the consequences of the pandemic on the Quebec population differ according to the population groups, but also according to the type of consequences.

\newpage

# Références

http://www.oecd.org/coronavirus/policy-responses/what-is-the-impact-of-the-covid-19-pandemic-on-immigrants-and-their-children-e7cbb7de/

Guttmann A, F. et al. (2020), COVID-19 in Immigrants, Refugees and Other Newcomers in Ontario: Characteristics of Those Tested and Those Confirmed Positive, as of June 13, 2020, ICES, http://www.ices.on.ca.

Hu, Y. (2020), “Intersecting ethnic and native–migrant inequalities in the economic impact of the COVID-19 pandemic in the UK”, Research in Social Stratification and Mobility, Vol. 68, p. 100528, http://dx.doi.org/10.1016/j.rssm.2020.100528.

Mcginnity, F. and G. Kingston (2017), “An Irish Welcome? Changing Irish Attitudes to Immigrants and Immigration: The Role of Recession and Immigration”, The Economic and social review, Vol. 48/3, pp. 281-304, https://www.researchgate.net/publication/319670153_An_Irish_Welcome_Changing_Irish_Attitudes_to_Immigrants_and_Immigration_The_Role_of_Recession_and_Immigration (accessed on 2 October 2020).

Raisi-Estabragh, Z. et al. (2020), “Greater risk of severe COVID-19 in Black, Asian and Minority Ethnic populations is not explained by cardiometabolic, socioeconomic or behavioural factors, or by 25(OH)-vitamin D status: study of 1326 cases from the UK Biobank”, Journal of Public Health, Vol. 42/3, pp. 451-460, http://dx.doi.org/10.1093/pubmed/fdaa095.

Apea, V. et al. (2020), Ethnicity and outcomes in patients hospitalised with COVID-19 infection in East London: an observational cohort study, Cold Spring Harbor Laboratory, http://dx.doi.org/10.1101/2020.06.10.20127621.

Bertocchi, G. and A. Dimico (2020), “Covid-19, Race, and Redlining”, IZA Discussion Paper, Vol. 13467, https://papers.ssrn.com/sol3/papers.cfm?abstract_id=3648807 (accessed on 22 September 2020).

Brun, S. and P. Simon (2020), “Inégalités ethno-raciales et coronavirus”, De Facto 19.

### 2023-05-01 Article/svm-latex-ms2.tex
update

<!DOCTYPE html>
<html lang="en">
  <head>
    <meta charset="utf-8">
  <link rel="dns-prefetch" href="https://github.githubassets.com">
  <link rel="dns-prefetch" href="https://avatars0.githubusercontent.com">
  <link rel="dns-prefetch" href="https://avatars1.githubusercontent.com">
  <link rel="dns-prefetch" href="https://avatars2.githubusercontent.com">
  <link rel="dns-prefetch" href="https://avatars3.githubusercontent.com">
  <link rel="dns-prefetch" href="https://github-cloud.s3.amazonaws.com">
  <link rel="dns-prefetch" href="https://user-images.githubusercontent.com/">

  <link crossorigin="anonymous" media="all" integrity="sha512-5Bs4ERl99/u2AUfpOZF2F0cdfNby7+Vd9teUshXUBPo5CjwECR7IAEf+weI/eCk5tF7K1h3O8hd8k0+P/HePeg==" rel="stylesheet" href="https://github.githubassets.com/assets/frameworks-e41b3811197df7fbb60147e939917617.css" />
  <link crossorigin="anonymous" media="all" integrity="sha512-21eX2YaEYIYa91WFh0mXaxJ3wLQxy2ldlpRbyuCueK3MvqHGEzmXYPkMQ4XBtASauZXKlL1hPxQQU9g28R03wQ==" rel="stylesheet" href="https://github.githubassets.com/assets/site-db5797d9868460861af755858749976b.css" />
    <link crossorigin="anonymous" media="all" integrity="sha512-vna9AxrJa9t+A93H0lxAPca1mXWQB5A/jIBppkp+Ys/EV2l+rsMcCyU/Fj4BpmEZ3nY4hHwJ/FqUEeMFXiYl0w==" rel="stylesheet" href="https://github.githubassets.com/assets/github-be76bd031ac96bdb7e03ddc7d25c403d.css" />

  <meta name="viewport" content="width=device-width">

  <title>website/svm-latex-ms2.tex at master · junkpolisci/website · GitHub</title>
    <meta name="description" content="Creating my academic website. . Contribute to junkpolisci/website development by creating an account on GitHub.">
    <link rel="search" type="application/opensearchdescription+xml" href="/opensearch.xml" title="GitHub">
  <link rel="fluid-icon" href="https://github.com/fluidicon.png" title="GitHub">
  <meta property="fb:app_id" content="1401488693436528">

    <meta name="twitter:image:src" content="https://avatars1.githubusercontent.com/u/46686077?s=400&amp;v=4" /><meta name="twitter:site" content="@github" /><meta name="twitter:card" content="summary" /><meta name="twitter:title" content="junkpolisci/website" /><meta name="twitter:description" content="Creating my academic website. . Contribute to junkpolisci/website development by creating an account on GitHub." />
    <meta property="og:image" content="https://avatars1.githubusercontent.com/u/46686077?s=400&amp;v=4" /><meta property="og:site_name" content="GitHub" /><meta property="og:type" content="object" /><meta property="og:title" content="junkpolisci/website" /><meta property="og:url" content="https://github.com/junkpolisci/website" /><meta property="og:description" content="Creating my academic website. . Contribute to junkpolisci/website development by creating an account on GitHub." />

  <link rel="assets" href="https://github.githubassets.com/">

  <meta name="request-id" content="F658:2063:11EE15:1DE035:5E76981F" data-pjax-transient="true"/><meta name="html-safe-nonce" content="24274df8c652eaefa357243991b4937ef8aec064" data-pjax-transient="true"/><meta name="visitor-payload" content="eyJyZWZlcnJlciI6Imh0dHBzOi8vZ2l0aHViLmNvbS9qdW5rcG9saXNjaS93ZWJzaXRlL3RyZWUvbWFzdGVyL3dvcmtpbmdfY29udGVudCIsInJlcXVlc3RfaWQiOiJGNjU4OjIwNjM6MTFFRTE1OjFERTAzNTo1RTc2OTgxRiIsInZpc2l0b3JfaWQiOiIyMjM1MzU2MDY0ODA4NjM4MTYzIiwicmVnaW9uX2VkZ2UiOiJpYWQiLCJyZWdpb25fcmVuZGVyIjoiaWFkIn0=" data-pjax-transient="true"/><meta name="visitor-hmac" content="7871cd4794e681531d7636de7d5c122519b34d083cf444c6feef74e399363f59" data-pjax-transient="true"/>

  <meta name="github-keyboard-shortcuts" content="repository,source-code" data-pjax-transient="true" />

  <meta name="selected-link" value="repo_source" data-pjax-transient>

    <meta name="google-site-verification" content="KT5gs8h0wvaagLKAVWq8bbeNwnZZK1r1XQysX3xurLU">
  <meta name="google-site-verification" content="ZzhVyEFwb7w3e0-uOTltm8Jsck2F5StVihD0exw2fsA">
  <meta name="google-site-verification" content="GXs5KoUUkNCoaAZn7wPN-t01Pywp9M3sEjnt_3_ZWPc">

<meta name="octolytics-host" content="collector.githubapp.com" /><meta name="octolytics-app-id" content="github" /><meta name="octolytics-event-url" content="https://collector.githubapp.com/github-external/browser_event" /><meta name="octolytics-dimension-ga_id" content="" class="js-octo-ga-id" />
<meta name="analytics-location" content="/&lt;user-name&gt;/&lt;repo-name&gt;/blob/show" data-pjax-transient="true" />

    <meta name="google-analytics" content="UA-3769691-2">

<meta class="js-ga-set" name="dimension1" content="Logged Out">

      <meta name="hostname" content="github.com">
    <meta name="user-login" content="">

      <meta name="expected-hostname" content="github.com">

    <meta name="enabled-features" content="MARKETPLACE_FEATURED_BLOG_POSTS,MARKETPLACE_INVOICED_BILLING,MARKETPLACE_SOCIAL_PROOF_CUSTOMERS,MARKETPLACE_TRENDING_SOCIAL_PROOF,MARKETPLACE_RECOMMENDATIONS,MARKETPLACE_PENDING_INSTALLATIONS,RELATED_ISSUES">

  <meta http-equiv="x-pjax-version" content="b9d38bbfa74751e01413c8f9baeaddf8">

      <link href="https://github.com/junkpolisci/website/commits/master.atom" rel="alternate" title="Recent Commits to website:master" type="application/atom+xml">

  <meta name="go-import" content="github.com/junkpolisci/website git https://github.com/junkpolisci/website.git">

  <meta name="octolytics-dimension-user_id" content="46686077" /><meta name="octolytics-dimension-user_login" content="junkpolisci" /><meta name="octolytics-dimension-repository_id" content="169272793" /><meta name="octolytics-dimension-repository_nwo" content="junkpolisci/website" /><meta name="octolytics-dimension-repository_public" content="true" /><meta name="octolytics-dimension-repository_is_fork" content="false" /><meta name="octolytics-dimension-repository_network_root_id" content="169272793" /><meta name="octolytics-dimension-repository_network_root_nwo" content="junkpolisci/website" /><meta name="octolytics-dimension-repository_explore_github_marketplace_ci_cta_shown" content="false" />

    <link rel="canonical" href="https://github.com/junkpolisci/website/blob/master/working_content/svm-latex-ms2.tex" data-pjax-transient>

  <meta name="browser-stats-url" content="https://api.github.com/_private/browser/stats">

  <meta name="browser-errors-url" content="https://api.github.com/_private/browser/errors">

  <link rel="mask-icon" href="https://github.githubassets.com/pinned-octocat.svg" color="#000000">
  <link rel="icon" type="image/x-icon" class="js-site-favicon" href="https://github.githubassets.com/favicon.ico">

<meta name="theme-color" content="#1e2327">

  <link rel="manifest" href="/manifest.json" crossOrigin="use-credentials">

  </head>

  <body class="logged-out env-production page-responsive page-blob">

  <div class="position-relative js-header-wrapper ">
    <a href="#start-of-content" class="px-2 py-4 bg-blue text-white show-on-focus js-skip-to-content">Skip to content</a>
    <span class="Progress progress-pjax-loader position-fixed width-full js-pjax-loader-bar">
      <span class="progress-pjax-loader-bar top-0 left-0" style="width: 0%;"></span>
    </span>

        <header class="Header-old header-logged-out js-details-container Details position-relative f4 py-2" role="banner">
  <div class="container-lg d-lg-flex flex-items-center p-responsive">
    <div class="d-flex flex-justify-between flex-items-center">
        <a class="mr-4" href="https://github.com/" aria-label="Homepage" data-ga-click="(Logged out) Header, go to homepage, icon:logo-wordmark">
          <svg height="32" class="octicon octicon-mark-github text-white" viewBox="0 0 16 16" version="1.1" width="32" aria-hidden="true"><path fill-rule="evenodd" d="M8 0C3.58 0 0 3.58 0 8c0 3.54 2.29 6.53 5.47 7.59.4.07.55-.17.55-.38 0-.19-.01-.82-.01-1.49-2.01.37-2.53-.49-2.69-.94-.09-.23-.48-.94-.82-1.13-.28-.15-.68-.52-.01-.53.63-.01 1.08.58 1.23.82.72 1.21 1.87.87 2.33.66.07-.52.28-.87.51-1.07-1.78-.2-3.64-.89-3.64-3.95 0-.87.31-1.59.82-2.15-.08-.2-.36-1.02.08-2.12 0 0 .67-.21 2.2.82.64-.18 1.32-.27 2-.27.68 0 1.36.09 2 .27 1.53-1.04 2.2-.82 2.2-.82.44 1.1.16 1.92.08 2.12.51.56.82 1.27.82 2.15 0 3.07-1.87 3.75-3.65 3.95.29.25.54.73.54 1.48 0 1.07-.01 1.93-.01 2.2 0 .21.15.46.55.38A8.013 8.013 0 0016 8c0-4.42-3.58-8-8-8z"/></svg>
        </a>

          <div class="d-lg-none css-truncate css-truncate-target width-fit p-2">

              <svg class="octicon octicon-repo" viewBox="0 0 12 16" version="1.1" width="12" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M4 9H3V8h1v1zm0-3H3v1h1V6zm0-2H3v1h1V4zm0-2H3v1h1V2zm8-1v12c0 .55-.45 1-1 1H6v2l-1.5-1.5L3 16v-2H1c-.55 0-1-.45-1-1V1c0-.55.45-1 1-1h10c.55 0 1 .45 1 1zm-1 10H1v2h2v-1h3v1h5v-2zm0-10H2v9h9V1z"/></svg>
    <a class="Header-link" href="/junkpolisci">junkpolisci</a>
    /
    <a class="Header-link" href="/junkpolisci/website">website</a>

          </div>

        <div class="d-flex flex-items-center">
            <a href="/join?source=header-repo"
              class="d-inline-block d-lg-none f5 text-white no-underline border border-gray-dark rounded-2 px-2 py-1 mr-3 mr-sm-5"
              data-hydro-click="{&quot;event_type&quot;:&quot;authentication.click&quot;,&quot;payload&quot;:{&quot;location_in_page&quot;:&quot;site header&quot;,&quot;repository_id&quot;:null,&quot;auth_type&quot;:&quot;SIGN_UP&quot;,&quot;originating_url&quot;:&quot;https://github.com/junkpolisci/website/blob/master/working_content/svm-latex-ms2.tex&quot;,&quot;user_id&quot;:null}}" data-hydro-click-hmac="5d1cb024487c37353729bcc7b6925a9406e9aefef45004b8758dc7f624c30fe4"
              data-ga-click="(Logged out) Header, clicked Sign up, text:sign-up">
              Sign&nbsp;up
            </a>

          <button class="btn-link d-lg-none mt-1 js-details-target" type="button" aria-label="Toggle navigation" aria-expanded="false">
            <svg height="24" class="octicon octicon-three-bars text-white" viewBox="0 0 12 16" version="1.1" width="18" aria-hidden="true"><path fill-rule="evenodd" d="M11.41 9H.59C0 9 0 8.59 0 8c0-.59 0-1 .59-1H11.4c.59 0 .59.41.59 1 0 .59 0 1-.59 1h.01zm0-4H.59C0 5 0 4.59 0 4c0-.59 0-1 .59-1H11.4c.59 0 .59.41.59 1 0 .59 0 1-.59 1h.01zM.59 11H11.4c.59 0 .59.41.59 1 0 .59 0 1-.59 1H.59C0 13 0 12.59 0 12c0-.59 0-1 .59-1z"/></svg>
          </button>
        </div>
    </div>

    <div class="HeaderMenu HeaderMenu--logged-out position-fixed top-0 right-0 bottom-0 height-fit position-lg-relative d-lg-flex flex-justify-between flex-items-center flex-auto">
      <div class="d-flex d-lg-none flex-justify-end border-bottom bg-gray-light p-3">
        <button class="btn-link js-details-target" type="button" aria-label="Toggle navigation" aria-expanded="false">
          <svg height="24" class="octicon octicon-x text-gray" viewBox="0 0 12 16" version="1.1" width="18" aria-hidden="true"><path fill-rule="evenodd" d="M7.48 8l3.75 3.75-1.48 1.48L6 9.48l-3.75 3.75-1.48-1.48L4.52 8 .77 4.25l1.48-1.48L6 6.52l3.75-3.75 1.48 1.48L7.48 8z"/></svg>
        </button>
      </div>

        <nav class="mt-0 px-3 px-lg-0 mb-5 mb-lg-0" aria-label="Global">
          <ul class="d-lg-flex list-style-none">
              <li class="d-block d-lg-flex flex-lg-nowrap flex-lg-items-center border-bottom border-lg-bottom-0 mr-0 mr-lg-3 edge-item-fix position-relative flex-wrap flex-justify-between d-flex flex-items-center ">
                <details class="HeaderMenu-details details-overlay details-reset width-full">
                  <summary class="HeaderMenu-summary HeaderMenu-link px-0 py-3 border-0 no-wrap d-block d-lg-inline-block">
                    Why GitHub?
                    <svg x="0px" y="0px" viewBox="0 0 14 8" xml:space="preserve" fill="none" class="icon-chevon-down-mktg position-absolute position-lg-relative">
                      <path d="M1,1l6.2,6L13,1"></path>
                    </svg>
                  </summary>
                  <div class="dropdown-menu flex-auto rounded-1 bg-white px-0 mt-0 pb-4 p-lg-4 position-relative position-lg-absolute left-0 left-lg-n4">
                    <a href="/features" class="py-2 lh-condensed-ultra d-block link-gray-dark no-underline h5 Bump-link--hover" data-ga-click="(Logged out) Header, go to Features">Features <span class="Bump-link-symbol float-right text-normal text-gray-light">&rarr;</span></a>
                    <ul class="list-style-none f5 pb-3">
                      <li class="edge-item-fix"><a href="/features/code-review/" class="py-2 lh-condensed-ultra d-block link-gray no-underline f5" data-ga-click="(Logged out) Header, go to Code review">Code review</a></li>
                      <li class="edge-item-fix"><a href="/features/project-management/" class="py-2 lh-condensed-ultra d-block link-gray no-underline f5" data-ga-click="(Logged out) Header, go to Project management">Project management</a></li>
                      <li class="edge-item-fix"><a href="/features/integrations" class="py-2 lh-condensed-ultra d-block link-gray no-underline f5" data-ga-click="(Logged out) Header, go to Integrations">Integrations</a></li>
                      <li class="edge-item-fix"><a href="/features/actions" class="py-2 lh-condensed-ultra d-block link-gray no-underline f5" data-ga-click="(Logged out) Header, go to Actions">Actions</a></li>
                          <li class="edge-item-fix"><a href="/features/packages" class="py-2 lh-condensed-ultra d-block link-gray no-underline f5" data-ga-click="(Logged out) Header, go to GitHub Packages">Packages</a></li>
                      <li class="edge-item-fix"><a href="/features/security" class="py-2 lh-condensed-ultra d-block link-gray no-underline f5" data-ga-click="(Logged out) Header, go to Security">Security</a></li>
                      <li class="edge-item-fix"><a href="/features#team-management" class="py-2 lh-condensed-ultra d-block link-gray no-underline f5" data-ga-click="(Logged out) Header, go to Team management">Team management</a></li>
                      <li class="edge-item-fix"><a href="/features#hosting" class="py-2 lh-condensed-ultra d-block link-gray no-underline f5" data-ga-click="(Logged out) Header, go to Code hosting">Hosting</a></li>
                    </ul>

                    <ul class="list-style-none mb-0 border-lg-top pt-lg-3">
                      <li class="edge-item-fix"><a href="/customer-stories" class="py-2 lh-condensed-ultra d-block no-underline link-gray-dark no-underline h5 Bump-link--hover" data-ga-click="(Logged out) Header, go to Customer stories">Customer stories <span class="Bump-link-symbol float-right text-normal text-gray-light">&rarr;</span></a></li>
                      <li class="edge-item-fix"><a href="/security" class="py-2 lh-condensed-ultra d-block no-underline link-gray-dark no-underline h5 Bump-link--hover" data-ga-click="(Logged out) Header, go to Security">Security <span class="Bump-link-symbol float-right text-normal text-gray-light">&rarr;</span></a></li>
                    </ul>
                  </div>
                </details>
              </li>
              <li class="border-bottom border-lg-bottom-0 mr-0 mr-lg-3">
                <a href="/enterprise" class="HeaderMenu-link no-underline py-3 d-block d-lg-inline-block" data-ga-click="(Logged out) Header, go to Enterprise">Enterprise</a>
              </li>

              <li class="d-block d-lg-flex flex-lg-nowrap flex-lg-items-center border-bottom border-lg-bottom-0 mr-0 mr-lg-3 edge-item-fix position-relative flex-wrap flex-justify-between d-flex flex-items-center ">
                <details class="HeaderMenu-details details-overlay details-reset width-full">
                  <summary class="HeaderMenu-summary HeaderMenu-link px-0 py-3 border-0 no-wrap d-block d-lg-inline-block">
                    Explore
                    <svg x="0px" y="0px" viewBox="0 0 14 8" xml:space="preserve" fill="none" class="icon-chevon-down-mktg position-absolute position-lg-relative">
                      <path d="M1,1l6.2,6L13,1"></path>
                    </svg>
                  </summary>

                  <div class="dropdown-menu flex-auto rounded-1 bg-white px-0 pt-2 pb-0 mt-0 pb-4 p-lg-4 position-relative position-lg-absolute left-0 left-lg-n4">
                    <ul class="list-style-none mb-3">
                      <li class="edge-item-fix"><a href="/explore" class="py-2 lh-condensed-ultra d-block link-gray-dark no-underline h5 Bump-link--hover" data-ga-click="(Logged out) Header, go to Explore">Explore GitHub <span class="Bump-link-symbol float-right text-normal text-gray-light">&rarr;</span></a></li>
                    </ul>

                    <h4 class="text-gray-light text-normal text-mono f5 mb-2 border-lg-top pt-lg-3">Learn &amp; contribute</h4>
                    <ul class="list-style-none mb-3">
                      <li class="edge-item-fix"><a href="/topics" class="py-2 lh-condensed-ultra d-block link-gray no-underline f5" data-ga-click="(Logged out) Header, go to Topics">Topics</a></li>
                        <li class="edge-item-fix"><a href="/collections" class="py-2 lh-condensed-ultra d-block link-gray no-underline f5" data-ga-click="(Logged out) Header, go to Collections">Collections</a></li>
                      <li class="edge-item-fix"><a href="/trending" class="py-2 lh-condensed-ultra d-block link-gray no-underline f5" data-ga-click="(Logged out) Header, go to Trending">Trending</a></li>
                      <li class="edge-item-fix"><a href="https://lab.github.com/" class="py-2 lh-condensed-ultra d-block link-gray no-underline f5" data-ga-click="(Logged out) Header, go to Learning lab">Learning Lab</a></li>
                      <li class="edge-item-fix"><a href="https://opensource.guide" class="py-2 lh-condensed-ultra d-block link-gray no-underline f5" data-ga-click="(Logged out) Header, go to Open source guides">Open source guides</a></li>
                    </ul>

                    <h4 class="text-gray-light text-normal text-mono f5 mb-2 border-lg-top pt-lg-3">Connect with others</h4>
                    <ul class="list-style-none mb-0">
                      <li class="edge-item-fix"><a href="https://github.com/events" class="py-2 lh-condensed-ultra d-block link-gray no-underline f5" data-ga-click="(Logged out) Header, go to Events">Events</a></li>
                      <li class="edge-item-fix"><a href="https://github.community" class="py-2 lh-condensed-ultra d-block link-gray no-underline f5" data-ga-click="(Logged out) Header, go to Community forum">Community forum</a></li>
                      <li class="edge-item-fix"><a href="https://education.github.com" class="py-2 pb-0 lh-condensed-ultra d-block link-gray no-underline f5" data-ga-click="(Logged out) Header, go to GitHub Education">GitHub Education</a></li>
                    </ul>
                  </div>
                </details>
              </li>

              <li class="border-bottom border-lg-bottom-0 mr-0 mr-lg-3">
                <a href="/marketplace" class="HeaderMenu-link no-underline py-3 d-block d-lg-inline-block" data-ga-click="(Logged out) Header, go to Marketplace">Marketplace</a>
              </li>

              <li class="d-block d-lg-flex flex-lg-nowrap flex-lg-items-center border-bottom border-lg-bottom-0 mr-0 mr-lg-3 edge-item-fix position-relative flex-wrap flex-justify-between d-flex flex-items-center ">
                <details class="HeaderMenu-details details-overlay details-reset width-full">
                  <summary class="HeaderMenu-summary HeaderMenu-link px-0 py-3 border-0 no-wrap d-block d-lg-inline-block">
                    Pricing
                    <svg x="0px" y="0px" viewBox="0 0 14 8" xml:space="preserve" fill="none" class="icon-chevon-down-mktg position-absolute position-lg-relative">
                       <path d="M1,1l6.2,6L13,1"></path>
                    </svg>
                  </summary>

                  <div class="dropdown-menu flex-auto rounded-1 bg-white px-0 pt-2 pb-4 mt-0 p-lg-4 position-relative position-lg-absolute left-0 left-lg-n4">
                    <a href="/pricing" class="pb-2 lh-condensed-ultra d-block link-gray-dark no-underline h5 Bump-link--hover" data-ga-click="(Logged out) Header, go to Pricing">Plans <span class="Bump-link-symbol float-right text-normal text-gray-light">&rarr;</span></a>

                    <ul class="list-style-none mb-3">
                      <li class="edge-item-fix"><a href="/pricing#feature-comparison" class="py-2 lh-condensed-ultra d-block link-gray no-underline f5" data-ga-click="(Logged out) Header, go to Compare plans">Compare plans</a></li>
                      <li class="edge-item-fix"><a href="https://enterprise.github.com/contact" class="py-2 lh-condensed-ultra d-block link-gray no-underline f5" data-ga-click="(Logged out) Header, go to Contact Sales">Contact Sales</a></li>
                    </ul>

                    <ul class="list-style-none mb-0 border-lg-top pt-lg-3">
                      <li class="edge-item-fix"><a href="/nonprofit" class="py-2 lh-condensed-ultra d-block no-underline link-gray-dark no-underline h5 Bump-link--hover" data-ga-click="(Logged out) Header, go to Nonprofits">Nonprofit <span class="Bump-link-symbol float-right text-normal text-gray-light">&rarr;</span></a></li>
                      <li class="edge-item-fix"><a href="https://education.github.com" class="py-2 pb-0 lh-condensed-ultra d-block no-underline link-gray-dark no-underline h5 Bump-link--hover"  data-ga-click="(Logged out) Header, go to Education">Education <span class="Bump-link-symbol float-right text-normal text-gray-light">&rarr;</span></a></li>
                    </ul>
                  </div>
                </details>
              </li>
          </ul>
        </nav>

      <div class="d-lg-flex flex-items-center px-3 px-lg-0 text-center text-lg-left">
          <div class="d-lg-flex mb-3 mb-lg-0">
            <div class="header-search flex-self-stretch flex-lg-self-auto mr-0 mr-lg-3 mb-3 mb-lg-0 scoped-search site-scoped-search js-site-search position-relative js-jump-to"
  role="combobox"
  aria-owns="jump-to-results"
  aria-label="Search or jump to"
  aria-haspopup="listbox"
  aria-expanded="false"
>
  <div class="position-relative">
    <!-- '"` --><!-- </textarea></xmp> --></option></form><form class="js-site-search-form" role="search" aria-label="Site" data-scope-type="Repository" data-scope-id="169272793" data-scoped-search-url="/junkpolisci/website/search" data-unscoped-search-url="/search" action="/junkpolisci/website/search" accept-charset="UTF-8" method="get">
      <label class="form-control input-sm header-search-wrapper p-0 header-search-wrapper-jump-to position-relative d-flex flex-justify-between flex-items-center js-chromeless-input-container">
        <input type="text"
          class="form-control input-sm header-search-input jump-to-field js-jump-to-field js-site-search-focus js-site-search-field is-clearable"
          data-hotkey="s,/"
          name="q"
          value=""
          placeholder="Search"
          data-unscoped-placeholder="Search GitHub"
          data-scoped-placeholder="Search"
          autocapitalize="off"
          aria-autocomplete="list"
          aria-controls="jump-to-results"
          aria-label="Search"
          data-jump-to-suggestions-path="/_graphql/GetSuggestedNavigationDestinations"
          spellcheck="false"
          autocomplete="off"
          >
          <input type="hidden" data-csrf="true" class="js-data-jump-to-suggestions-path-csrf" value="SoltMYDSiK5yYkm5PCK7UIUncsGs7ATBP397zb49QiULt83WOQOjhVXDBg8ER+S6KIX8bjSYUimTdY87qPo6Bg==" />
          <input type="hidden" class="js-site-search-type-field" name="type" >
            <img src="https://github.githubassets.com/images/search-key-slash.svg" alt="" class="mr-2 header-search-key-slash">

            <div class="Box position-absolute overflow-hidden d-none jump-to-suggestions js-jump-to-suggestions-container">

<ul class="d-none js-jump-to-suggestions-template-container">

<li class="d-flex flex-justify-start flex-items-center p-0 f5 navigation-item js-navigation-item js-jump-to-suggestion" role="option">
  <a tabindex="-1" class="no-underline d-flex flex-auto flex-items-center jump-to-suggestions-path js-jump-to-suggestion-path js-navigation-open p-2" href="">
    <div class="jump-to-octicon js-jump-to-octicon flex-shrink-0 mr-2 text-center d-none">
      <svg height="16" width="16" class="octicon octicon-repo flex-shrink-0 js-jump-to-octicon-repo d-none" title="Repository" aria-label="Repository" viewBox="0 0 12 16" version="1.1" role="img"><path fill-rule="evenodd" d="M4 9H3V8h1v1zm0-3H3v1h1V6zm0-2H3v1h1V4zm0-2H3v1h1V2zm8-1v12c0 .55-.45 1-1 1H6v2l-1.5-1.5L3 16v-2H1c-.55 0-1-.45-1-1V1c0-.55.45-1 1-1h10c.55 0 1 .45 1 1zm-1 10H1v2h2v-1h3v1h5v-2zm0-10H2v9h9V1z"/></svg>
      <svg height="16" width="16" class="octicon octicon-project flex-shrink-0 js-jump-to-octicon-project d-none" title="Project" aria-label="Project" viewBox="0 0 15 16" version="1.1" role="img"><path fill-rule="evenodd" d="M10 12h3V2h-3v10zm-4-2h3V2H6v8zm-4 4h3V2H2v12zm-1 1h13V1H1v14zM14 0H1a1 1 0 00-1 1v14a1 1 0 001 1h13a1 1 0 001-1V1a1 1 0 00-1-1z"/></svg>
      <svg height="16" width="16" class="octicon octicon-search flex-shrink-0 js-jump-to-octicon-search d-none" title="Search" aria-label="Search" viewBox="0 0 16 16" version="1.1" role="img"><path fill-rule="evenodd" d="M15.7 13.3l-3.81-3.83A5.93 5.93 0 0013 6c0-3.31-2.69-6-6-6S1 2.69 1 6s2.69 6 6 6c1.3 0 2.48-.41 3.47-1.11l3.83 3.81c.19.2.45.3.7.3.25 0 .52-.09.7-.3a.996.996 0 000-1.41v.01zM7 10.7c-2.59 0-4.7-2.11-4.7-4.7 0-2.59 2.11-4.7 4.7-4.7 2.59 0 4.7 2.11 4.7 4.7 0 2.59-2.11 4.7-4.7 4.7z"/></svg>
    </div>

    <img class="avatar mr-2 flex-shrink-0 js-jump-to-suggestion-avatar d-none" alt="" aria-label="Team" src="" width="28" height="28">

    <div class="jump-to-suggestion-name js-jump-to-suggestion-name flex-auto overflow-hidden text-left no-wrap css-truncate css-truncate-target">
    </div>

    <div class="border rounded-1 flex-shrink-0 bg-gray px-1 text-gray-light ml-1 f6 d-none js-jump-to-badge-search">
      <span class="js-jump-to-badge-search-text-default d-none" aria-label="in this repository">
        In this repository
      </span>
      <span class="js-jump-to-badge-search-text-global d-none" aria-label="in all of GitHub">
        All GitHub
      </span>
      <span aria-hidden="true" class="d-inline-block ml-1 v-align-middle">↵</span>
    </div>

    <div aria-hidden="true" class="border rounded-1 flex-shrink-0 bg-gray px-1 text-gray-light ml-1 f6 d-none d-on-nav-focus js-jump-to-badge-jump">
      Jump to
      <span class="d-inline-block ml-1 v-align-middle">↵</span>
    </div>
  </a>
</li>

</ul>

<ul class="d-none js-jump-to-no-results-template-container">
  <li class="d-flex flex-justify-center flex-items-center f5 d-none js-jump-to-suggestion p-2">
    <span class="text-gray">No suggested jump to results</span>
  </li>
</ul>

<ul id="jump-to-results" role="listbox" class="p-0 m-0 js-navigation-container jump-to-suggestions-results-container js-jump-to-suggestions-results-container">

<li class="d-flex flex-justify-start flex-items-center p-0 f5 navigation-item js-navigation-item js-jump-to-scoped-search d-none" role="option">
  <a tabindex="-1" class="no-underline d-flex flex-auto flex-items-center jump-to-suggestions-path js-jump-to-suggestion-path js-navigation-open p-2" href="">
    <div class="jump-to-octicon js-jump-to-octicon flex-shrink-0 mr-2 text-center d-none">
      <svg height="16" width="16" class="octicon octicon-repo flex-shrink-0 js-jump-to-octicon-repo d-none" title="Repository" aria-label="Repository" viewBox="0 0 12 16" version="1.1" role="img"><path fill-rule="evenodd" d="M4 9H3V8h1v1zm0-3H3v1h1V6zm0-2H3v1h1V4zm0-2H3v1h1V2zm8-1v12c0 .55-.45 1-1 1H6v2l-1.5-1.5L3 16v-2H1c-.55 0-1-.45-1-1V1c0-.55.45-1 1-1h10c.55 0 1 .45 1 1zm-1 10H1v2h2v-1h3v1h5v-2zm0-10H2v9h9V1z"/></svg>
      <svg height="16" width="16" class="octicon octicon-project flex-shrink-0 js-jump-to-octicon-project d-none" title="Project" aria-label="Project" viewBox="0 0 15 16" version="1.1" role="img"><path fill-rule="evenodd" d="M10 12h3V2h-3v10zm-4-2h3V2H6v8zm-4 4h3V2H2v12zm-1 1h13V1H1v14zM14 0H1a1 1 0 00-1 1v14a1 1 0 001 1h13a1 1 0 001-1V1a1 1 0 00-1-1z"/></svg>
      <svg height="16" width="16" class="octicon octicon-search flex-shrink-0 js-jump-to-octicon-search d-none" title="Search" aria-label="Search" viewBox="0 0 16 16" version="1.1" role="img"><path fill-rule="evenodd" d="M15.7 13.3l-3.81-3.83A5.93 5.93 0 0013 6c0-3.31-2.69-6-6-6S1 2.69 1 6s2.69 6 6 6c1.3 0 2.48-.41 3.47-1.11l3.83 3.81c.19.2.45.3.7.3.25 0 .52-.09.7-.3a.996.996 0 000-1.41v.01zM7 10.7c-2.59 0-4.7-2.11-4.7-4.7 0-2.59 2.11-4.7 4.7-4.7 2.59 0 4.7 2.11 4.7 4.7 0 2.59-2.11 4.7-4.7 4.7z"/></svg>
    </div>

    <img class="avatar mr-2 flex-shrink-0 js-jump-to-suggestion-avatar d-none" alt="" aria-label="Team" src="" width="28" height="28">

    <div class="jump-to-suggestion-name js-jump-to-suggestion-name flex-auto overflow-hidden text-left no-wrap css-truncate css-truncate-target">
    </div>

    <div class="border rounded-1 flex-shrink-0 bg-gray px-1 text-gray-light ml-1 f6 d-none js-jump-to-badge-search">
      <span class="js-jump-to-badge-search-text-default d-none" aria-label="in this repository">
        In this repository
      </span>
      <span class="js-jump-to-badge-search-text-global d-none" aria-label="in all of GitHub">
        All GitHub
      </span>
      <span aria-hidden="true" class="d-inline-block ml-1 v-align-middle">↵</span>
    </div>

    <div aria-hidden="true" class="border rounded-1 flex-shrink-0 bg-gray px-1 text-gray-light ml-1 f6 d-none d-on-nav-focus js-jump-to-badge-jump">
      Jump to
      <span class="d-inline-block ml-1 v-align-middle">↵</span>
    </div>
  </a>
</li>

<li class="d-flex flex-justify-start flex-items-center p-0 f5 navigation-item js-navigation-item js-jump-to-global-search d-none" role="option">
  <a tabindex="-1" class="no-underline d-flex flex-auto flex-items-center jump-to-suggestions-path js-jump-to-suggestion-path js-navigation-open p-2" href="">
    <div class="jump-to-octicon js-jump-to-octicon flex-shrink-0 mr-2 text-center d-none">
      <svg height="16" width="16" class="octicon octicon-repo flex-shrink-0 js-jump-to-octicon-repo d-none" title="Repository" aria-label="Repository" viewBox="0 0 12 16" version="1.1" role="img"><path fill-rule="evenodd" d="M4 9H3V8h1v1zm0-3H3v1h1V6zm0-2H3v1h1V4zm0-2H3v1h1V2zm8-1v12c0 .55-.45 1-1 1H6v2l-1.5-1.5L3 16v-2H1c-.55 0-1-.45-1-1V1c0-.55.45-1 1-1h10c.55 0 1 .45 1 1zm-1 10H1v2h2v-1h3v1h5v-2zm0-10H2v9h9V1z"/></svg>
      <svg height="16" width="16" class="octicon octicon-project flex-shrink-0 js-jump-to-octicon-project d-none" title="Project" aria-label="Project" viewBox="0 0 15 16" version="1.1" role="img"><path fill-rule="evenodd" d="M10 12h3V2h-3v10zm-4-2h3V2H6v8zm-4 4h3V2H2v12zm-1 1h13V1H1v14zM14 0H1a1 1 0 00-1 1v14a1 1 0 001 1h13a1 1 0 001-1V1a1 1 0 00-1-1z"/></svg>
      <svg height="16" width="16" class="octicon octicon-search flex-shrink-0 js-jump-to-octicon-search d-none" title="Search" aria-label="Search" viewBox="0 0 16 16" version="1.1" role="img"><path fill-rule="evenodd" d="M15.7 13.3l-3.81-3.83A5.93 5.93 0 0013 6c0-3.31-2.69-6-6-6S1 2.69 1 6s2.69 6 6 6c1.3 0 2.48-.41 3.47-1.11l3.83 3.81c.19.2.45.3.7.3.25 0 .52-.09.7-.3a.996.996 0 000-1.41v.01zM7 10.7c-2.59 0-4.7-2.11-4.7-4.7 0-2.59 2.11-4.7 4.7-4.7 2.59 0 4.7 2.11 4.7 4.7 0 2.59-2.11 4.7-4.7 4.7z"/></svg>
    </div>

    <img class="avatar mr-2 flex-shrink-0 js-jump-to-suggestion-avatar d-none" alt="" aria-label="Team" src="" width="28" height="28">

    <div class="jump-to-suggestion-name js-jump-to-suggestion-name flex-auto overflow-hidden text-left no-wrap css-truncate css-truncate-target">
    </div>

    <div class="border rounded-1 flex-shrink-0 bg-gray px-1 text-gray-light ml-1 f6 d-none js-jump-to-badge-search">
      <span class="js-jump-to-badge-search-text-default d-none" aria-label="in this repository">
        In this repository
      </span>
      <span class="js-jump-to-badge-search-text-global d-none" aria-label="in all of GitHub">
        All GitHub
      </span>
      <span aria-hidden="true" class="d-inline-block ml-1 v-align-middle">↵</span>
    </div>

    <div aria-hidden="true" class="border rounded-1 flex-shrink-0 bg-gray px-1 text-gray-light ml-1 f6 d-none d-on-nav-focus js-jump-to-badge-jump">
      Jump to
      <span class="d-inline-block ml-1 v-align-middle">↵</span>
    </div>
  </a>
</li>

</ul>

            </div>
      </label>
</form>  </div>
</div>

          </div>

        <a href="/login?return_to=%2Fjunkpolisci%2Fwebsite%2Fblob%2Fmaster%2Fworking_content%2Fsvm-latex-ms2.tex"
          class="HeaderMenu-link no-underline mr-3"
          data-hydro-click="{&quot;event_type&quot;:&quot;authentication.click&quot;,&quot;payload&quot;:{&quot;location_in_page&quot;:&quot;site header menu&quot;,&quot;repository_id&quot;:null,&quot;auth_type&quot;:&quot;SIGN_UP&quot;,&quot;originating_url&quot;:&quot;https://github.com/junkpolisci/website/blob/master/working_content/svm-latex-ms2.tex&quot;,&quot;user_id&quot;:null}}" data-hydro-click-hmac="c1bed73ac71356020bf58b7a50b4fc8dc40cea6e72b1abb04db432bf8b24473e"
          data-ga-click="(Logged out) Header, clicked Sign in, text:sign-in">
          Sign&nbsp;in
        </a>
          <a href="/join?source=header-repo&amp;source_repo=junkpolisci%2Fwebsite"
            class="HeaderMenu-link d-inline-block no-underline border border-gray-dark rounded-1 px-2 py-1"
            data-hydro-click="{&quot;event_type&quot;:&quot;authentication.click&quot;,&quot;payload&quot;:{&quot;location_in_page&quot;:&quot;site header menu&quot;,&quot;repository_id&quot;:null,&quot;auth_type&quot;:&quot;SIGN_UP&quot;,&quot;originating_url&quot;:&quot;https://github.com/junkpolisci/website/blob/master/working_content/svm-latex-ms2.tex&quot;,&quot;user_id&quot;:null}}" data-hydro-click-hmac="c1bed73ac71356020bf58b7a50b4fc8dc40cea6e72b1abb04db432bf8b24473e"
            data-ga-click="(Logged out) Header, clicked Sign up, text:sign-up">
            Sign&nbsp;up
          </a>
      </div>
    </div>
  </div>
</header>

  </div>

  <div id="start-of-content" class="show-on-focus"></div>

    <div id="js-flash-container">

</div>

  <include-fragment class="js-notification-shelf-include-fragment" data-base-src="https://github.com/notifications/beta/shelf"></include-fragment>

  <div class="application-main " data-commit-hovercards-enabled>
        <div itemscope itemtype="http://schema.org/SoftwareSourceCode" class="">
    <main  >

  <div class="pagehead repohead hx_repohead readability-menu bg-gray-light pb-0 pt-0 pt-lg-3">

    <div class="d-flex container-lg mb-4 p-responsive d-none d-lg-flex">

      <div class="flex-auto min-width-0 width-fit mr-3">
        <h1 class="public  d-flex flex-wrap flex-items-center break-word float-none ">
    <svg class="octicon octicon-repo" viewBox="0 0 12 16" version="1.1" width="12" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M4 9H3V8h1v1zm0-3H3v1h1V6zm0-2H3v1h1V4zm0-2H3v1h1V2zm8-1v12c0 .55-.45 1-1 1H6v2l-1.5-1.5L3 16v-2H1c-.55 0-1-.45-1-1V1c0-.55.45-1 1-1h10c.55 0 1 .45 1 1zm-1 10H1v2h2v-1h3v1h5v-2zm0-10H2v9h9V1z"/></svg>
  <span class="author ml-1 flex-self-stretch" itemprop="author">
    <a class="url fn" rel="author" data-hovercard-type="user" data-hovercard-url="/users/junkpolisci/hovercard" data-octo-click="hovercard-link-click" data-octo-dimensions="link_type:self" href="/junkpolisci">junkpolisci</a>
  </span>
  <span class="path-divider flex-self-stretch">/</span>
  <strong itemprop="name" class="mr-2 flex-self-stretch">
    <a data-pjax="#js-repo-pjax-container" href="/junkpolisci/website">website</a>
  </strong>

</h1>

      </div>

      <ul class="pagehead-actions flex-shrink-0 " >

  <li>

  <a class="tooltipped tooltipped-s btn btn-sm btn-with-count" aria-label="You must be signed in to watch a repository" rel="nofollow" data-hydro-click="{&quot;event_type&quot;:&quot;authentication.click&quot;,&quot;payload&quot;:{&quot;location_in_page&quot;:&quot;notification subscription menu watch&quot;,&quot;repository_id&quot;:null,&quot;auth_type&quot;:&quot;LOG_IN&quot;,&quot;originating_url&quot;:&quot;https://github.com/junkpolisci/website/blob/master/working_content/svm-latex-ms2.tex&quot;,&quot;user_id&quot;:null}}" data-hydro-click-hmac="0d9b91ce9f4a854f7d1ee137af3383a369388666efda7a77ff300d5a18b2c6fd" href="/login?return_to=%2Fjunkpolisci%2Fwebsite">
    <svg class="octicon octicon-eye v-align-text-bottom" viewBox="0 0 16 16" version="1.1" width="16" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M8.06 2C3 2 0 8 0 8s3 6 8.06 6C13 14 16 8 16 8s-3-6-7.94-6zM8 12c-2.2 0-4-1.78-4-4 0-2.2 1.8-4 4-4 2.22 0 4 1.8 4 4 0 2.22-1.78 4-4 4zm2-4c0 1.11-.89 2-2 2-1.11 0-2-.89-2-2 0-1.11.89-2 2-2 1.11 0 2 .89 2 2z"/></svg>
    Watch
</a>    <a class="social-count" href="/junkpolisci/website/watchers"
       aria-label="0 users are watching this repository">
      0
    </a>

  </li>

  <li>

  </li>

  <li>
      <a class="btn btn-sm btn-with-count tooltipped tooltipped-s" aria-label="You must be signed in to fork a repository" rel="nofollow" data-hydro-click="{&quot;event_type&quot;:&quot;authentication.click&quot;,&quot;payload&quot;:{&quot;location_in_page&quot;:&quot;repo details fork button&quot;,&quot;repository_id&quot;:169272793,&quot;auth_type&quot;:&quot;LOG_IN&quot;,&quot;originating_url&quot;:&quot;https://github.com/junkpolisci/website/blob/master/working_content/svm-latex-ms2.tex&quot;,&quot;user_id&quot;:null}}" data-hydro-click-hmac="21baa47b38991cc3f0b73b4b0fa6dc35d24c324464d3c6896595cf41ad156698" href="/login?return_to=%2Fjunkpolisci%2Fwebsite">
        <svg class="octicon octicon-repo-forked v-align-text-bottom" viewBox="0 0 10 16" version="1.1" width="10" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M8 1a1.993 1.993 0 00-1 3.72V6L5 8 3 6V4.72A1.993 1.993 0 002 1a1.993 1.993 0 00-1 3.72V6.5l3 3v1.78A1.993 1.993 0 005 15a1.993 1.993 0 001-3.72V9.5l3-3V4.72A1.993 1.993 0 008 1zM2 4.2C1.34 4.2.8 3.65.8 3c0-.65.55-1.2 1.2-1.2.65 0 1.2.55 1.2 1.2 0 .65-.55 1.2-1.2 1.2zm3 10c-.66 0-1.2-.55-1.2-1.2 0-.65.55-1.2 1.2-1.2.65 0 1.2.55 1.2 1.2 0 .65-.55 1.2-1.2 1.2zm3-10c-.66 0-1.2-.55-1.2-1.2 0-.65.55-1.2 1.2-1.2.65 0 1.2.55 1.2 1.2 0 .65-.55 1.2-1.2 1.2z"/></svg>
        Fork
</a>
    <a href="/junkpolisci/website/network/members" class="social-count"
       aria-label="0 users forked this repository">
      0
    </a>
  </li>
</ul>

    </div>

<nav class="hx_reponav reponav js-repo-nav js-sidenav-container-pjax clearfix container-lg p-responsive d-none d-lg-block"
     itemscope
     itemtype="http://schema.org/BreadcrumbList"
    aria-label="Repository"
     data-pjax="#js-repo-pjax-container">

  <span itemscope itemtype="http://schema.org/ListItem" itemprop="itemListElement">
    <a class="js-selected-navigation-item selected reponav-item" itemprop="url" data-hotkey="g c" aria-current="page" data-selected-links="repo_source repo_downloads repo_commits repo_releases repo_tags repo_branches repo_packages /junkpolisci/website" href="/junkpolisci/website">
      <div class="d-inline"><svg class="octicon octicon-code" viewBox="0 0 14 16" version="1.1" width="14" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M9.5 3L8 4.5 11.5 8 8 11.5 9.5 13 14 8 9.5 3zm-5 0L0 8l4.5 5L6 11.5 2.5 8 6 4.5 4.5 3z"/></svg></div>
      <span itemprop="name">Code</span>
      <meta itemprop="position" content="1">
</a>  </span>

    <span itemscope itemtype="http://schema.org/ListItem" itemprop="itemListElement">
      <a itemprop="url" data-hotkey="g i" class="js-selected-navigation-item reponav-item" data-selected-links="repo_issues repo_labels repo_milestones /junkpolisci/website/issues" href="/junkpolisci/website/issues">
        <div class="d-inline"><svg class="octicon octicon-issue-opened" viewBox="0 0 14 16" version="1.1" width="14" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M7 2.3c3.14 0 5.7 2.56 5.7 5.7s-2.56 5.7-5.7 5.7A5.71 5.71 0 011.3 8c0-3.14 2.56-5.7 5.7-5.7zM7 1C3.14 1 0 4.14 0 8s3.14 7 7 7 7-3.14 7-7-3.14-7-7-7zm1 3H6v5h2V4zm0 6H6v2h2v-2z"/></svg></div>
        <span itemprop="name">Issues</span>
        <span class="Counter">0</span>
        <meta itemprop="position" content="2">
</a>    </span>

  <span itemscope itemtype="http://schema.org/ListItem" itemprop="itemListElement">
    <a data-hotkey="g p" data-skip-pjax="true" itemprop="url" class="js-selected-navigation-item reponav-item" data-selected-links="repo_pulls checks /junkpolisci/website/pulls" href="/junkpolisci/website/pulls">
      <div class="d-inline"><svg class="octicon octicon-git-pull-request" viewBox="0 0 12 16" version="1.1" width="12" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M11 11.28V5c-.03-.78-.34-1.47-.94-2.06C9.46 2.35 8.78 2.03 8 2H7V0L4 3l3 3V4h1c.27.02.48.11.69.31.21.2.3.42.31.69v6.28A1.993 1.993 0 0010 15a1.993 1.993 0 001-3.72zm-1 2.92c-.66 0-1.2-.55-1.2-1.2 0-.65.55-1.2 1.2-1.2.65 0 1.2.55 1.2 1.2 0 .65-.55 1.2-1.2 1.2zM4 3c0-1.11-.89-2-2-2a1.993 1.993 0 00-1 3.72v6.56A1.993 1.993 0 002 15a1.993 1.993 0 001-3.72V4.72c.59-.34 1-.98 1-1.72zm-.8 10c0 .66-.55 1.2-1.2 1.2-.65 0-1.2-.55-1.2-1.2 0-.65.55-1.2 1.2-1.2.65 0 1.2.55 1.2 1.2zM2 4.2C1.34 4.2.8 3.65.8 3c0-.65.55-1.2 1.2-1.2.65 0 1.2.55 1.2 1.2 0 .65-.55 1.2-1.2 1.2z"/></svg></div>
      <span itemprop="name">Pull requests</span>
      <span class="Counter">0</span>
      <meta itemprop="position" content="4">
</a>  </span>

    <span itemscope itemtype="http://schema.org/ListItem" itemprop="itemListElement" class="position-relative float-left">
      <a data-hotkey="g w" data-skip-pjax="true" class="js-selected-navigation-item reponav-item" data-selected-links="repo_actions /junkpolisci/website/actions" href="/junkpolisci/website/actions">
        <div class="d-inline"><svg class="octicon octicon-play" viewBox="0 0 14 16" version="1.1" width="14" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M14 8A7 7 0 110 8a7 7 0 0114 0zm-8.223 3.482l4.599-3.066a.5.5 0 000-.832L5.777 4.518A.5.5 0 005 4.934v6.132a.5.5 0 00.777.416z"/></svg></div>
        Actions
</a>
    </span>

    <a data-hotkey="g b" class="js-selected-navigation-item reponav-item" data-selected-links="repo_projects new_repo_project repo_project /junkpolisci/website/projects" href="/junkpolisci/website/projects">
      <div class="d-inline"><svg class="octicon octicon-project" viewBox="0 0 15 16" version="1.1" width="15" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M10 12h3V2h-3v10zm-4-2h3V2H6v8zm-4 4h3V2H2v12zm-1 1h13V1H1v14zM14 0H1a1 1 0 00-1 1v14a1 1 0 001 1h13a1 1 0 001-1V1a1 1 0 00-1-1z"/></svg></div>
      Projects
      <span class="Counter">0</span>
</a>

    <a data-skip-pjax="true" class="js-selected-navigation-item reponav-item" data-selected-links="security alerts policy token_scanning code_scanning /junkpolisci/website/security/advisories" href="/junkpolisci/website/security/advisories">
      <div class="d-inline"><svg class="octicon octicon-shield" viewBox="0 0 14 16" version="1.1" width="14" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M0 2l7-2 7 2v6.02C14 12.69 8.69 16 7 16c-1.69 0-7-3.31-7-7.98V2zm1 .75L7 1l6 1.75v5.268C13 12.104 8.449 15 7 15c-1.449 0-6-2.896-6-6.982V2.75zm1 .75L7 2v12c-1.207 0-5-2.482-5-5.985V3.5z"/></svg></div>
      Security
</a>
    <a class="js-selected-navigation-item reponav-item" data-selected-links="repo_graphs repo_contributors dependency_graph dependabot_updates pulse people /junkpolisci/website/pulse" href="/junkpolisci/website/pulse">
      <div class="d-inline"><svg class="octicon octicon-graph" viewBox="0 0 16 16" version="1.1" width="16" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M16 14v1H0V0h1v14h15zM5 13H3V8h2v5zm4 0H7V3h2v10zm4 0h-2V6h2v7z"/></svg></div>
      Insights
</a>

</nav>

  <div class="reponav-wrapper reponav-small d-lg-none">
  <nav class="reponav js-reponav text-center no-wrap"
       itemscope
       itemtype="http://schema.org/BreadcrumbList">

    <span itemscope itemtype="http://schema.org/ListItem" itemprop="itemListElement">
      <a class="js-selected-navigation-item selected reponav-item" itemprop="url" aria-current="page" data-selected-links="repo_source repo_downloads repo_commits repo_releases repo_tags repo_branches repo_packages /junkpolisci/website" href="/junkpolisci/website">
        <span itemprop="name">Code</span>
        <meta itemprop="position" content="1">
</a>    </span>

      <span itemscope itemtype="http://schema.org/ListItem" itemprop="itemListElement">
        <a itemprop="url" class="js-selected-navigation-item reponav-item" data-selected-links="repo_issues repo_labels repo_milestones /junkpolisci/website/issues" href="/junkpolisci/website/issues">
          <span itemprop="name">Issues</span>
          <span class="Counter">0</span>
          <meta itemprop="position" content="2">
</a>      </span>

    <span itemscope itemtype="http://schema.org/ListItem" itemprop="itemListElement">
      <a itemprop="url" class="js-selected-navigation-item reponav-item" data-selected-links="repo_pulls checks /junkpolisci/website/pulls" href="/junkpolisci/website/pulls">
        <span itemprop="name">Pull requests</span>
        <span class="Counter">0</span>
        <meta itemprop="position" content="4">
</a>    </span>

      <span itemscope itemtype="http://schema.org/ListItem" itemprop="itemListElement">
        <a itemprop="url" class="js-selected-navigation-item reponav-item" data-selected-links="repo_projects new_repo_project repo_project /junkpolisci/website/projects" href="/junkpolisci/website/projects">
          <span itemprop="name">Projects</span>
          <span class="Counter">0</span>
          <meta itemprop="position" content="5">
</a>      </span>

      <span itemscope itemtype="http://schema.org/ListItem" itemprop="itemListElement">
        <a itemprop="url" class="js-selected-navigation-item reponav-item" data-selected-links="repo_actions /junkpolisci/website/actions" href="/junkpolisci/website/actions">
          <span itemprop="name">Actions</span>
          <meta itemprop="position" content="6">
</a>      </span>

      <a itemprop="url" class="js-selected-navigation-item reponav-item" data-selected-links="security alerts policy token_scanning code_scanning /junkpolisci/website/security/advisories" href="/junkpolisci/website/security/advisories">
        <span itemprop="name">Security</span>
        <meta itemprop="position" content="8">
</a>
      <a class="js-selected-navigation-item reponav-item" data-selected-links="pulse /junkpolisci/website/pulse" href="/junkpolisci/website/pulse">
        Pulse
</a>

  </nav>
</div>

  </div>

  <include-fragment class="js-notification-shelf-include-fragment" data-base-src="https://github.com/notifications/beta/shelf"></include-fragment>

<div class="container-lg clearfix new-discussion-timeline  p-responsive">
  <div class="repository-content ">

    <a class="d-none js-permalink-shortcut" data-hotkey="y" href="/junkpolisci/website/blob/8c72f5f6bdda3784b8d521ae15603285cbcdbfd9/working_content/svm-latex-ms2.tex">Permalink</a>

    <!-- blob contrib key: blob_contributors:v22:5707502fa5923753652aee2fac7386cb -->
      <div class="signup-prompt-bg rounded-1 js-signup-prompt" data-prompt="signup" hidden>
    <div class="signup-prompt p-4 text-center mb-4 rounded-1">
      <div class="position-relative">
        <button type="button" class="position-absolute top-0 right-0 btn-link link-gray js-signup-prompt-button" data-ga-click="(Logged out) Sign up prompt, clicked Dismiss, text:dismiss">
          Dismiss
        </button>
        <h3 class="pt-2">Join GitHub today</h3>
        <p class="col-6 mx-auto">GitHub is home to over 40 million developers working together to host and review code, manage projects, and build software together.</p>
        <a class="btn btn-primary" data-ga-click="(Logged out) Sign up prompt, clicked Sign up, text:sign-up" data-hydro-click="{&quot;event_type&quot;:&quot;authentication.click&quot;,&quot;payload&quot;:{&quot;location_in_page&quot;:&quot;files signup prompt&quot;,&quot;repository_id&quot;:null,&quot;auth_type&quot;:&quot;SIGN_UP&quot;,&quot;originating_url&quot;:&quot;https://github.com/junkpolisci/website/blob/master/working_content/svm-latex-ms2.tex&quot;,&quot;user_id&quot;:null}}" data-hydro-click-hmac="e7240f46793ed7ad304ffbd252ed8759c24c0616e59c327633f6e083410c521c" href="/join?source=prompt-blob-show&amp;source_repo=junkpolisci%2Fwebsite">Sign up</a>
      </div>
    </div>
  </div>

    <div class="d-flex flex-items-start flex-shrink-0 flex-column flex-md-row pb-3">
      <span class="d-flex flex-justify-between width-full width-md-auto">

<details class="details-reset details-overlay branch-select-menu " id="branch-select-menu">
  <summary class="btn btn-sm css-truncate"
           data-hotkey="w"
           title="Switch branches or tags">
    <i>Branch:</i>
    <span class="css-truncate-target" data-menu-button>master</span>
    <span class="dropdown-caret"></span>
  </summary>

  <details-menu class="SelectMenu SelectMenu--hasFilter" src="/junkpolisci/website/refs/master/working_content/svm-latex-ms2.tex?source_action=show&amp;source_controller=blob" preload>
    <div class="SelectMenu-modal">
      <include-fragment class="SelectMenu-loading" aria-label="Menu is loading">
        <svg class="octicon octicon-octoface anim-pulse" height="32" viewBox="0 0 16 16" version="1.1" width="32" aria-hidden="true"><path fill-rule="evenodd" d="M14.7 5.34c.13-.32.55-1.59-.13-3.31 0 0-1.05-.33-3.44 1.3-1-.28-2.07-.32-3.13-.32s-2.13.04-3.13.32c-2.39-1.64-3.44-1.3-3.44-1.3-.68 1.72-.26 2.99-.13 3.31C.49 6.21 0 7.33 0 8.69 0 13.84 3.33 15 7.98 15S16 13.84 16 8.69c0-1.36-.49-2.48-1.3-3.35zM8 14.02c-3.3 0-5.98-.15-5.98-3.35 0-.76.38-1.48 1.02-2.07 1.07-.98 2.9-.46 4.96-.46 2.07 0 3.88-.52 4.96.46.65.59 1.02 1.3 1.02 2.07 0 3.19-2.68 3.35-5.98 3.35zM5.49 9.01c-.66 0-1.2.8-1.2 1.78s.54 1.79 1.2 1.79c.66 0 1.2-.8 1.2-1.79s-.54-1.78-1.2-1.78zm5.02 0c-.66 0-1.2.79-1.2 1.78s.54 1.79 1.2 1.79c.66 0 1.2-.8 1.2-1.79s-.53-1.78-1.2-1.78z"/></svg>
      </include-fragment>
    </div>
  </details-menu>
</details>

        <div class="BtnGroup flex-shrink-0 d-md-none">
          <a href="/junkpolisci/website/find/master"
                class="js-pjax-capture-input btn btn-sm BtnGroup-item"
                data-pjax
                data-hotkey="t">
            Find file
          </a>
          <clipboard-copy value="working_content/svm-latex-ms2.tex" class="btn btn-sm BtnGroup-item">
            Copy path
          </clipboard-copy>
        </div>
      </span>
      <h2 id="blob-path" class="breadcrumb flex-auto min-width-0 text-normal flex-md-self-center ml-md-2 mr-md-3 my-2 my-md-0">
          <span class="js-repo-root text-bold"><span class="js-path-segment"><a data-pjax="true" href="/junkpolisci/website"><span>website</span></a></span></span><span class="separator">/</span><span class="js-path-segment"><a data-pjax="true" href="/junkpolisci/website/tree/master/working_content"><span>working_content</span></a></span><span class="separator">/</span><strong class="final-path">svm-latex-ms2.tex</strong>
      </h2>

      <div class="BtnGroup flex-shrink-0 d-none d-md-inline-block">
        <a href="/junkpolisci/website/find/master"
              class="js-pjax-capture-input btn btn-sm BtnGroup-item"
              data-pjax
              data-hotkey="t">
          Find file
        </a>
        <clipboard-copy value="working_content/svm-latex-ms2.tex" class="btn btn-sm BtnGroup-item">
          Copy path
        </clipboard-copy>
      </div>
    </div>

  <div class="Box Box--condensed d-flex flex-column flex-shrink-0 mb-3">
      <div class="Box-body d-flex flex-justify-between bg-blue-light flex-column flex-md-row flex-items-start flex-md-items-center">
        <span class="pr-md-4 f6">
          <img class="avatar" width="20" height="20" alt="" src="https://camo.githubusercontent.com/09e94ac5a4d9d3c7f4ab7f53ab39fb4ab8ef14f6/68747470733a2f2f312e67726176617461722e636f6d2f6176617461722f31643533383461356230646466656133613362646364643433366666366338333f643d68747470732533412532462532466769746875622e6769746875626173736574732e636f6d253246696d6167657325324667726176617461727325324667726176617461722d757365722d3432302e706e6726723d6726733d313430" data-canonical-src="https://1.gravatar.com/avatar/1d5384a5b0ddfea3a3bdcdd436ff6c83?d=https%3A%2F%2Fgithub.githubassets.com%2Fimages%2Fgravatars%2Fgravatar-user-420.png&amp;r=g&amp;s=140" />
          <span class="text-bold link-gray-dark lh-default v-align-middle">junkpolisci</span>
            <span class="lh-default v-align-middle">
              <a data-pjax="true" title="Moved files back to working content folder because moving them the first time broke the website." class="link-gray" href="/junkpolisci/website/commit/9ad45cec280c14d59390d49324f2bb9ec85efbee">Moved files back to working content folder because moving them the fi…</a>
            </span>
        </span>
        <span class="d-inline-block flex-shrink-0 v-align-bottom f6 mt-2 mt-md-0">
          <a class="pr-2 text-mono link-gray" href="/junkpolisci/website/commit/9ad45cec280c14d59390d49324f2bb9ec85efbee" data-pjax>9ad45ce</a>
          <relative-time datetime="2019-02-25T15:32:30Z" class="no-wrap">Feb 25, 2019</relative-time>
        </span>
      </div>

    <div class="Box-body d-flex flex-items-center flex-auto f6 border-bottom-0 flex-wrap" >
      <details class="details-reset details-overlay details-overlay-dark lh-default text-gray-dark float-left mr-2" id="blob_contributors_box">
        <summary class="btn-link">
          <span><strong>0</strong> contributors</span>
        </summary>
        <details-dialog
          class="Box Box--overlay d-flex flex-column anim-fade-in fast"
          aria-label="Users who have contributed to this file"
          src="/junkpolisci/website/contributors-list/master/working_content/svm-latex-ms2.tex" preload>
          <div class="Box-header">
            <button class="Box-btn-octicon btn-octicon float-right" type="button" aria-label="Close dialog" data-close-dialog>
              <svg class="octicon octicon-x" viewBox="0 0 12 16" version="1.1" width="12" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M7.48 8l3.75 3.75-1.48 1.48L6 9.48l-3.75 3.75-1.48-1.48L4.52 8 .77 4.25l1.48-1.48L6 6.52l3.75-3.75 1.48 1.48L7.48 8z"/></svg>
            </button>
            <h3 class="Box-title">
              Users who have contributed to this file
            </h3>
          </div>
          <include-fragment class="octocat-spinner my-3" aria-label="Loading..."></include-fragment>
        </details-dialog>
      </details>
    </div>
  </div>

    <div class="Box mt-3 position-relative
      ">

<div class="Box-header py-2 d-flex flex-column flex-shrink-0 flex-md-row flex-md-items-center">
  <div class="text-mono f6 flex-auto pr-3 flex-order-2 flex-md-order-1 mt-2 mt-md-0">

      271 lines (200 sloc)
      <span class="file-info-divider"></span>
    6.05 KB
  </div>

  <div class="d-flex py-1 py-md-0 flex-auto flex-order-1 flex-md-order-2 flex-sm-grow-0 flex-justify-between">

    <div class="BtnGroup">
      <a id="raw-url" class="btn btn-sm BtnGroup-item" href="/junkpolisci/website/raw/master/working_content/svm-latex-ms2.tex">Raw</a>
        <a class="btn btn-sm js-update-url-with-hash BtnGroup-item" data-hotkey="b" href="/junkpolisci/website/blame/master/working_content/svm-latex-ms2.tex">Blame</a>
      <a rel="nofollow" class="btn btn-sm BtnGroup-item" href="/junkpolisci/website/commits/master/working_content/svm-latex-ms2.tex">History</a>
    </div>

    <div>
          <a class="btn-octicon tooltipped tooltipped-nw js-remove-unless-platform"
             data-platforms="windows,mac"
             href="https://desktop.github.com"
             aria-label="Open this file in GitHub Desktop"
             data-ga-click="Repository, open with desktop">
              <svg class="octicon octicon-device-desktop" viewBox="0 0 16 16" version="1.1" width="16" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M15 2H1c-.55 0-1 .45-1 1v9c0 .55.45 1 1 1h5.34c-.25.61-.86 1.39-2.34 2h8c-1.48-.61-2.09-1.39-2.34-2H15c.55 0 1-.45 1-1V3c0-.55-.45-1-1-1zm0 9H1V3h14v8z"/></svg>
          </a>

          <button type="button" class="btn-octicon disabled tooltipped tooltipped-nw"
            aria-label="You must be signed in to make or propose changes">
            <svg class="octicon octicon-pencil" viewBox="0 0 14 16" version="1.1" width="14" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M0 12v3h3l8-8-3-3-8 8zm3 2H1v-2h1v1h1v1zm10.3-9.3L12 6 9 3l1.3-1.3a.996.996 0 011.41 0l1.59 1.59c.39.39.39 1.02 0 1.41z"/></svg>
          </button>
          <button type="button" class="btn-octicon btn-octicon-danger disabled tooltipped tooltipped-nw"
            aria-label="You must be signed in to make or propose changes">
            <svg class="octicon octicon-trashcan" viewBox="0 0 12 16" version="1.1" width="12" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M11 2H9c0-.55-.45-1-1-1H5c-.55 0-1 .45-1 1H2c-.55 0-1 .45-1 1v1c0 .55.45 1 1 1v9c0 .55.45 1 1 1h7c.55 0 1-.45 1-1V5c.55 0 1-.45 1-1V3c0-.55-.45-1-1-1zm-1 12H3V5h1v8h1V5h1v8h1V5h1v8h1V5h1v9zm1-10H2V3h9v1z"/></svg>
          </button>
    </div>
  </div>
</div>

  <div itemprop="text" class="Box-body p-0 blob-wrapper data type-tex ">

<table class="highlight tab-size js-file-line-container" data-tab-size="8" data-paste-markdown-skip>
      <tr>
        <td id="L1" class="blob-num js-line-number" data-line-number="1"></td>
        <td id="LC1" class="blob-code blob-code-inner js-file-line"><span class="pl-k">\documentclass</span>[<span class="pl-v">$if(fontsize)$$fontsize$,$endif$$if(lang)$$babel-lang$,$endif$$if(papersize)$$papersize$,$endif$$for(classoption)$$classoption$$sep$,$endfor$</span>]{<span class="pl-c1">article</span>}</td>
      </tr>
      <tr>
        <td id="L2" class="blob-num js-line-number" data-line-number="2"></td>
        <td id="LC2" class="blob-code blob-code-inner js-file-line"><span class="pl-k">\usepackage</span>[<span class="pl-v">left=1in,top=1in,right=1in,bottom=1in</span>]{<span class="pl-c1">geometry</span>}</td>
      </tr>
      <tr>
        <td id="L3" class="blob-num js-line-number" data-line-number="3"></td>
        <td id="LC3" class="blob-code blob-code-inner js-file-line"><span class="pl-k">\newcommand</span>*{<span class="pl-c1">\authorfont</span>}{<span class="pl-c1">\fontfamily</span>{phv}<span class="pl-c1">\selectfont</span>}</td>
      </tr>
      <tr>
        <td id="L4" class="blob-num js-line-number" data-line-number="4"></td>
        <td id="LC4" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(fontfamily)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L5" class="blob-num js-line-number" data-line-number="5"></td>
        <td id="LC5" class="blob-code blob-code-inner js-file-line"><span class="pl-k">\usepackage</span>[<span class="pl-v">$fontfamilyoptions$</span>]{<span class="pl-c1"><span class="pl-s"><span class="pl-pds">$</span>fontfamily<span class="pl-pds">$</span></span></span>}</td>
      </tr>
      <tr>
        <td id="L6" class="blob-num js-line-number" data-line-number="6"></td>
        <td id="LC6" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>else<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L7" class="blob-num js-line-number" data-line-number="7"></td>
        <td id="LC7" class="blob-code blob-code-inner js-file-line"><span class="pl-k">\usepackage</span>{<span class="pl-c1">lmodern</span>}</td>
      </tr>
      <tr>
        <td id="L8" class="blob-num js-line-number" data-line-number="8"></td>
        <td id="LC8" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L9" class="blob-num js-line-number" data-line-number="9"></td>
        <td id="LC9" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L10" class="blob-num js-line-number" data-line-number="10"></td>
        <td id="LC10" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L11" class="blob-num js-line-number" data-line-number="11"></td>
        <td id="LC11" class="blob-code blob-code-inner js-file-line">  <span class="pl-k">\usepackage</span>[<span class="pl-v">$if(fontenc)$$fontenc$$else$T1$endif$</span>]{<span class="pl-c1">fontenc</span>}</td>
      </tr>
      <tr>
        <td id="L12" class="blob-num js-line-number" data-line-number="12"></td>
        <td id="LC12" class="blob-code blob-code-inner js-file-line">  <span class="pl-k">\usepackage</span>[<span class="pl-v">utf8</span>]{<span class="pl-c1">inputenc</span>}</td>
      </tr>
      <tr>
        <td id="L13" class="blob-num js-line-number" data-line-number="13"></td>
        <td id="LC13" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L14" class="blob-num js-line-number" data-line-number="14"></td>
        <td id="LC14" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L15" class="blob-num js-line-number" data-line-number="15"></td>
        <td id="LC15" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L16" class="blob-num js-line-number" data-line-number="16"></td>
        <td id="LC16" class="blob-code blob-code-inner js-file-line"><span class="pl-k">\usepackage</span>{<span class="pl-c1">abstract</span>}</td>
      </tr>
      <tr>
        <td id="L17" class="blob-num js-line-number" data-line-number="17"></td>
        <td id="LC17" class="blob-code blob-code-inner js-file-line"><span class="pl-k">\renewcommand</span>{<span class="pl-c1">\abstractname</span>}{}    <span class="pl-c"><span class="pl-c">%</span> clear the title</span></td>
      </tr>
      <tr>
        <td id="L18" class="blob-num js-line-number" data-line-number="18"></td>
        <td id="LC18" class="blob-code blob-code-inner js-file-line"><span class="pl-k">\renewcommand</span>{<span class="pl-c1">\absnamepos</span>}{empty} <span class="pl-c"><span class="pl-c">%</span> originally center</span></td>
      </tr>
      <tr>
        <td id="L19" class="blob-num js-line-number" data-line-number="19"></td>
        <td id="LC19" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L20" class="blob-num js-line-number" data-line-number="20"></td>
        <td id="LC20" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\renewenvironment</span>{abstract}</td>
      </tr>
      <tr>
        <td id="L21" class="blob-num js-line-number" data-line-number="21"></td>
        <td id="LC21" class="blob-code blob-code-inner js-file-line"> {{<span class="pl-c"><span class="pl-c">%</span></span></td>
      </tr>
      <tr>
        <td id="L22" class="blob-num js-line-number" data-line-number="22"></td>
        <td id="LC22" class="blob-code blob-code-inner js-file-line">    <span class="pl-c1">\setlength</span>{<span class="pl-c1">\leftmargin</span>}{0mm}</td>
      </tr>
      <tr>
        <td id="L23" class="blob-num js-line-number" data-line-number="23"></td>
        <td id="LC23" class="blob-code blob-code-inner js-file-line">    <span class="pl-c1">\setlength</span>{<span class="pl-c1">\rightmargin</span>}{<span class="pl-c1">\leftmargin</span>}<span class="pl-c"><span class="pl-c">%</span></span></td>
      </tr>
      <tr>
        <td id="L24" class="blob-num js-line-number" data-line-number="24"></td>
        <td id="LC24" class="blob-code blob-code-inner js-file-line">  }<span class="pl-c"><span class="pl-c">%</span></span></td>
      </tr>
      <tr>
        <td id="L25" class="blob-num js-line-number" data-line-number="25"></td>
        <td id="LC25" class="blob-code blob-code-inner js-file-line">  <span class="pl-c1">\relax</span>}</td>
      </tr>
      <tr>
        <td id="L26" class="blob-num js-line-number" data-line-number="26"></td>
        <td id="LC26" class="blob-code blob-code-inner js-file-line"> {<span class="pl-c1">\endlist</span>}</td>
      </tr>
      <tr>
        <td id="L27" class="blob-num js-line-number" data-line-number="27"></td>
        <td id="LC27" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L28" class="blob-num js-line-number" data-line-number="28"></td>
        <td id="LC28" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\makeatletter</span></td>
      </tr>
      <tr>
        <td id="L29" class="blob-num js-line-number" data-line-number="29"></td>
        <td id="LC29" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\def\@maketitle</span>{<span class="pl-c"><span class="pl-c">%</span></span></td>
      </tr>
      <tr>
        <td id="L30" class="blob-num js-line-number" data-line-number="30"></td>
        <td id="LC30" class="blob-code blob-code-inner js-file-line">  <span class="pl-c1">\newpage</span></td>
      </tr>
      <tr>
        <td id="L31" class="blob-num js-line-number" data-line-number="31"></td>
        <td id="LC31" class="blob-code blob-code-inner js-file-line"><span class="pl-c"><span class="pl-c">%</span>  \null</span></td>
      </tr>
      <tr>
        <td id="L32" class="blob-num js-line-number" data-line-number="32"></td>
        <td id="LC32" class="blob-code blob-code-inner js-file-line"><span class="pl-c"><span class="pl-c">%</span>  \vskip 2em%</span></td>
      </tr>
      <tr>
        <td id="L33" class="blob-num js-line-number" data-line-number="33"></td>
        <td id="LC33" class="blob-code blob-code-inner js-file-line"><span class="pl-c"><span class="pl-c">%</span>  \begin{center}%</span></td>
      </tr>
      <tr>
        <td id="L34" class="blob-num js-line-number" data-line-number="34"></td>
        <td id="LC34" class="blob-code blob-code-inner js-file-line">  <span class="pl-c1">\let</span> <span class="pl-c1">\footnote</span> <span class="pl-c1">\thanks</span></td>
      </tr>
      <tr>
        <td id="L35" class="blob-num js-line-number" data-line-number="35"></td>
        <td id="LC35" class="blob-code blob-code-inner js-file-line">    {<span class="pl-c1">\fontsize</span>{18}{20}<span class="pl-c1">\selectfont\raggedright</span>  <span class="pl-c1">\setlength</span>{<span class="pl-c1">\parindent</span>}{0pt} <span class="pl-c1">\@title</span> <span class="pl-c1">\par</span>}<span class="pl-c"><span class="pl-c">%</span></span></td>
      </tr>
      <tr>
        <td id="L36" class="blob-num js-line-number" data-line-number="36"></td>
        <td id="LC36" class="blob-code blob-code-inner js-file-line">}</td>
      </tr>
      <tr>
        <td id="L37" class="blob-num js-line-number" data-line-number="37"></td>
        <td id="LC37" class="blob-code blob-code-inner js-file-line"><span class="pl-c"><span class="pl-c">%</span>\fi</span></td>
      </tr>
      <tr>
        <td id="L38" class="blob-num js-line-number" data-line-number="38"></td>
        <td id="LC38" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\makeatother</span></td>
      </tr>
      <tr>
        <td id="L39" class="blob-num js-line-number" data-line-number="39"></td>
        <td id="LC39" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L40" class="blob-num js-line-number" data-line-number="40"></td>
        <td id="LC40" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L41" class="blob-num js-line-number" data-line-number="41"></td>
        <td id="LC41" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(appendix)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L42" class="blob-num js-line-number" data-line-number="42"></td>
        <td id="LC42" class="blob-code blob-code-inner js-file-line"><span class="pl-k">\renewcommand</span>*<span class="pl-c1">\thetable</span>{<span class="pl-s"><span class="pl-pds">$</span>appendixletter<span class="pl-pds">$</span></span>.<span class="pl-c1">\arabic</span>{table}}</td>
      </tr>
      <tr>
        <td id="L43" class="blob-num js-line-number" data-line-number="43"></td>
        <td id="LC43" class="blob-code blob-code-inner js-file-line"><span class="pl-k">\renewcommand</span>*<span class="pl-c1">\thefigure</span>{<span class="pl-s"><span class="pl-pds">$</span>appendixletter<span class="pl-pds">$</span></span>.<span class="pl-c1">\arabic</span>{figure}}</td>
      </tr>
      <tr>
        <td id="L44" class="blob-num js-line-number" data-line-number="44"></td>
        <td id="LC44" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L45" class="blob-num js-line-number" data-line-number="45"></td>
        <td id="LC45" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L46" class="blob-num js-line-number" data-line-number="46"></td>
        <td id="LC46" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L47" class="blob-num js-line-number" data-line-number="47"></td>
        <td id="LC47" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\setcounter</span>{secnumdepth}{0}</td>
      </tr>
      <tr>
        <td id="L48" class="blob-num js-line-number" data-line-number="48"></td>
        <td id="LC48" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L49" class="blob-num js-line-number" data-line-number="49"></td>
        <td id="LC49" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(listings)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L50" class="blob-num js-line-number" data-line-number="50"></td>
        <td id="LC50" class="blob-code blob-code-inner js-file-line"><span class="pl-k">\usepackage</span>{<span class="pl-c1">listings</span>}</td>
      </tr>
      <tr>
        <td id="L51" class="blob-num js-line-number" data-line-number="51"></td>
        <td id="LC51" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L52" class="blob-num js-line-number" data-line-number="52"></td>
        <td id="LC52" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(lhs)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L53" class="blob-num js-line-number" data-line-number="53"></td>
        <td id="LC53" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\lstnewenvironment</span>{code}{<span class="pl-c1">\lstset</span>{language=r,basicstyle=<span class="pl-c1">\small\ttfamily</span>}}{}</td>
      </tr>
      <tr>
        <td id="L54" class="blob-num js-line-number" data-line-number="54"></td>
        <td id="LC54" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L55" class="blob-num js-line-number" data-line-number="55"></td>
        <td id="LC55" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(highlighting-macros)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L56" class="blob-num js-line-number" data-line-number="56"></td>
        <td id="LC56" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>highlighting-macros<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L57" class="blob-num js-line-number" data-line-number="57"></td>
        <td id="LC57" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L58" class="blob-num js-line-number" data-line-number="58"></td>
        <td id="LC58" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(verbatim-in-note)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L59" class="blob-num js-line-number" data-line-number="59"></td>
        <td id="LC59" class="blob-code blob-code-inner js-file-line"><span class="pl-k">\usepackage</span>{<span class="pl-c1">fancyvrb</span>}</td>
      </tr>
      <tr>
        <td id="L60" class="blob-num js-line-number" data-line-number="60"></td>
        <td id="LC60" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L61" class="blob-num js-line-number" data-line-number="61"></td>
        <td id="LC61" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(tables)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L62" class="blob-num js-line-number" data-line-number="62"></td>
        <td id="LC62" class="blob-code blob-code-inner js-file-line"><span class="pl-k">\usepackage</span>{<span class="pl-c1">longtable,booktabs</span>}</td>
      </tr>
      <tr>
        <td id="L63" class="blob-num js-line-number" data-line-number="63"></td>
        <td id="LC63" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L64" class="blob-num js-line-number" data-line-number="64"></td>
        <td id="LC64" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L65" class="blob-num js-line-number" data-line-number="65"></td>
        <td id="LC65" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(graphics)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L66" class="blob-num js-line-number" data-line-number="66"></td>
        <td id="LC66" class="blob-code blob-code-inner js-file-line"><span class="pl-k">\usepackage</span>{<span class="pl-c1">graphicx,grffile</span>}</td>
      </tr>
      <tr>
        <td id="L67" class="blob-num js-line-number" data-line-number="67"></td>
        <td id="LC67" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\makeatletter</span></td>
      </tr>
      <tr>
        <td id="L68" class="blob-num js-line-number" data-line-number="68"></td>
        <td id="LC68" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\def\maxwidth</span>{<span class="pl-k">\ifdim</span><span class="pl-c1">\Gin@nat@width</span>&gt;<span class="pl-c1">\linewidth\linewidth</span><span class="pl-k">\else</span><span class="pl-c1">\Gin@nat@width</span><span class="pl-k">\fi</span>}</td>
      </tr>
      <tr>
        <td id="L69" class="blob-num js-line-number" data-line-number="69"></td>
        <td id="LC69" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\def\maxheight</span>{<span class="pl-k">\ifdim</span><span class="pl-c1">\Gin@nat@height</span>&gt;<span class="pl-c1">\textheight\textheight</span><span class="pl-k">\else</span><span class="pl-c1">\Gin@nat@height</span><span class="pl-k">\fi</span>}</td>
      </tr>
      <tr>
        <td id="L70" class="blob-num js-line-number" data-line-number="70"></td>
        <td id="LC70" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\makeatother</span></td>
      </tr>
      <tr>
        <td id="L71" class="blob-num js-line-number" data-line-number="71"></td>
        <td id="LC71" class="blob-code blob-code-inner js-file-line"><span class="pl-c"><span class="pl-c">%</span> Scale images if necessary, so that they will not overflow the page</span></td>
      </tr>
      <tr>
        <td id="L72" class="blob-num js-line-number" data-line-number="72"></td>
        <td id="LC72" class="blob-code blob-code-inner js-file-line"><span class="pl-c"><span class="pl-c">%</span> margins by default, and it is still possible to overwrite the defaults</span></td>
      </tr>
      <tr>
        <td id="L73" class="blob-num js-line-number" data-line-number="73"></td>
        <td id="LC73" class="blob-code blob-code-inner js-file-line"><span class="pl-c"><span class="pl-c">%</span> using explicit options in \includegraphics[width, height, ...]{}</span></td>
      </tr>
      <tr>
        <td id="L74" class="blob-num js-line-number" data-line-number="74"></td>
        <td id="LC74" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\setkeys</span>{Gin}{width=<span class="pl-c1">\maxwidth</span>,height=<span class="pl-c1">\maxheight</span>,keepaspectratio}</td>
      </tr>
      <tr>
        <td id="L75" class="blob-num js-line-number" data-line-number="75"></td>
        <td id="LC75" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L76" class="blob-num js-line-number" data-line-number="76"></td>
        <td id="LC76" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L77" class="blob-num js-line-number" data-line-number="77"></td>
        <td id="LC77" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(title)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L78" class="blob-num js-line-number" data-line-number="78"></td>
        <td id="LC78" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\title</span>{<span class="pl-s"><span class="pl-pds">$</span>title<span class="pl-pds">$</span><span class="pl-pds">$</span>if(subtitle)<span class="pl-pds">$</span></span>: <span class="pl-s"><span class="pl-pds">$</span>subtitle<span class="pl-pds">$</span><span class="pl-pds">$</span>endif<span class="pl-pds">$</span><span class="pl-pds">$</span>if(anonymous)<span class="pl-pds">$</span><span class="pl-pds">$</span>else<span class="pl-pds">$</span><span class="pl-pds">$</span>if(thanks)<span class="pl-pds">$</span></span><span class="pl-c1">\thanks</span>{<span class="pl-s"><span class="pl-pds">$</span>thanks<span class="pl-pds">$</span></span>} <span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span> }</td>
      </tr>
      <tr>
        <td id="L79" class="blob-num js-line-number" data-line-number="79"></td>
        <td id="LC79" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span> </td>
      </tr>
      <tr>
        <td id="L80" class="blob-num js-line-number" data-line-number="80"></td>
        <td id="LC80" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L81" class="blob-num js-line-number" data-line-number="81"></td>
        <td id="LC81" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L82" class="blob-num js-line-number" data-line-number="82"></td>
        <td id="LC82" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L83" class="blob-num js-line-number" data-line-number="83"></td>
        <td id="LC83" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(anonymous)<span class="pl-pds">$</span><span class="pl-pds">$</span>else<span class="pl-pds">$</span></span><span class="pl-c1">\author</span>{<span class="pl-s"><span class="pl-pds">$</span>for(author)<span class="pl-pds">$</span></span><span class="pl-c1">\Large</span> <span class="pl-s"><span class="pl-pds">$</span>author.name<span class="pl-pds">$</span></span><span class="pl-c1">\vspace</span>{0.05in} <span class="pl-c1">\newline\normalsize</span><span class="pl-c1">\emph</span>{<span class="pl-mi"><span class="pl-s"><span class="pl-pds">$</span>author.affiliation<span class="pl-pds">$</span></span></span>}  <span class="pl-s"><span class="pl-pds">$</span>sep<span class="pl-pds">$</span></span> <span class="pl-c1">\and</span> <span class="pl-s"><span class="pl-pds">$</span>endfor<span class="pl-pds">$</span></span>}<span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L84" class="blob-num js-line-number" data-line-number="84"></td>
        <td id="LC84" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L85" class="blob-num js-line-number" data-line-number="85"></td>
        <td id="LC85" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L86" class="blob-num js-line-number" data-line-number="86"></td>
        <td id="LC86" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\date</span>{}</td>
      </tr>
      <tr>
        <td id="L87" class="blob-num js-line-number" data-line-number="87"></td>
        <td id="LC87" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L88" class="blob-num js-line-number" data-line-number="88"></td>
        <td id="LC88" class="blob-code blob-code-inner js-file-line"><span class="pl-k">\usepackage</span>{<span class="pl-c1">titlesec</span>}</td>
      </tr>
      <tr>
        <td id="L89" class="blob-num js-line-number" data-line-number="89"></td>
        <td id="LC89" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L90" class="blob-num js-line-number" data-line-number="90"></td>
        <td id="LC90" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\titleformat</span>*{<span class="pl-c1">\section</span>}{<span class="pl-c1">\normalsize\bfseries</span>}</td>
      </tr>
      <tr>
        <td id="L91" class="blob-num js-line-number" data-line-number="91"></td>
        <td id="LC91" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\titleformat</span>*{<span class="pl-c1">\subsection</span>}{<span class="pl-c1">\normalsize\itshape</span>}</td>
      </tr>
      <tr>
        <td id="L92" class="blob-num js-line-number" data-line-number="92"></td>
        <td id="LC92" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\titleformat</span>*{<span class="pl-c1">\subsubsection</span>}{<span class="pl-c1">\normalsize\itshape</span>}</td>
      </tr>
      <tr>
        <td id="L93" class="blob-num js-line-number" data-line-number="93"></td>
        <td id="LC93" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\titleformat</span>*{<span class="pl-c1">\paragraph</span>}{<span class="pl-c1">\normalsize\itshape</span>}</td>
      </tr>
      <tr>
        <td id="L94" class="blob-num js-line-number" data-line-number="94"></td>
        <td id="LC94" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\titleformat</span>*{<span class="pl-c1">\subparagraph</span>}{<span class="pl-c1">\normalsize\itshape</span>}</td>
      </tr>
      <tr>
        <td id="L95" class="blob-num js-line-number" data-line-number="95"></td>
        <td id="LC95" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L96" class="blob-num js-line-number" data-line-number="96"></td>
        <td id="LC96" class="blob-code blob-code-inner js-file-line"><span class="pl-k">\newcommand</span>{<span class="pl-c1">\dummy</span>}[1]{#1}</td>
      </tr>
      <tr>
        <td id="L97" class="blob-num js-line-number" data-line-number="97"></td>
        <td id="LC97" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L98" class="blob-num js-line-number" data-line-number="98"></td>
        <td id="LC98" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(natbib)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L99" class="blob-num js-line-number" data-line-number="99"></td>
        <td id="LC99" class="blob-code blob-code-inner js-file-line"><span class="pl-k">\usepackage</span>{<span class="pl-c1">natbib</span>}</td>
      </tr>
      <tr>
        <td id="L100" class="blob-num js-line-number" data-line-number="100"></td>
        <td id="LC100" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\bibpunct</span>{(}{)}{;}{a}{}{,}</td>
      </tr>
      <tr>
        <td id="L101" class="blob-num js-line-number" data-line-number="101"></td>
        <td id="LC101" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\bibliographystyle</span>{<span class="pl-s"><span class="pl-pds">$</span>if(biblio-style)<span class="pl-pds">$</span><span class="pl-pds">$</span>biblio-style<span class="pl-pds">$</span><span class="pl-pds">$</span>else<span class="pl-pds">$</span></span>plainnat<span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span>}</td>
      </tr>
      <tr>
        <td id="L102" class="blob-num js-line-number" data-line-number="102"></td>
        <td id="LC102" class="blob-code blob-code-inner js-file-line"><span class="pl-c"><span class="pl-c">%</span>\usepackage[strings]{underscore} % protect underscores in most circumstances</span></td>
      </tr>
      <tr>
        <td id="L103" class="blob-num js-line-number" data-line-number="103"></td>
        <td id="LC103" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L104" class="blob-num js-line-number" data-line-number="104"></td>
        <td id="LC104" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L105" class="blob-num js-line-number" data-line-number="105"></td>
        <td id="LC105" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(biblatex)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L106" class="blob-num js-line-number" data-line-number="106"></td>
        <td id="LC106" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\usepackage</span><span class="pl-s"><span class="pl-pds">$</span>if(biblio-style)<span class="pl-pds">$</span></span>[style=<span class="pl-s"><span class="pl-pds">$</span>biblio-style<span class="pl-pds">$</span></span>]<span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span>{biblatex}</td>
      </tr>
      <tr>
        <td id="L107" class="blob-num js-line-number" data-line-number="107"></td>
        <td id="LC107" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(biblatexoptions)<span class="pl-pds">$</span></span><span class="pl-c1">\ExecuteBibliographyOptions</span>{<span class="pl-s"><span class="pl-pds">$</span>for(biblatexoptions)<span class="pl-pds">$</span><span class="pl-pds">$</span>biblatexoptions<span class="pl-pds">$</span><span class="pl-pds">$</span>sep<span class="pl-pds">$</span></span>,<span class="pl-s"><span class="pl-pds">$</span>endfor<span class="pl-pds">$</span></span>}<span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L108" class="blob-num js-line-number" data-line-number="108"></td>
        <td id="LC108" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>for(bibliography)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L109" class="blob-num js-line-number" data-line-number="109"></td>
        <td id="LC109" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\addbibresource</span>{<span class="pl-s"><span class="pl-pds">$</span>bibliography<span class="pl-pds">$</span></span>}</td>
      </tr>
      <tr>
        <td id="L110" class="blob-num js-line-number" data-line-number="110"></td>
        <td id="LC110" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endfor<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L111" class="blob-num js-line-number" data-line-number="111"></td>
        <td id="LC111" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L112" class="blob-num js-line-number" data-line-number="112"></td>
        <td id="LC112" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(listings)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L113" class="blob-num js-line-number" data-line-number="113"></td>
        <td id="LC113" class="blob-code blob-code-inner js-file-line"><span class="pl-k">\usepackage</span>{<span class="pl-c1">listings</span>}</td>
      </tr>
      <tr>
        <td id="L114" class="blob-num js-line-number" data-line-number="114"></td>
        <td id="LC114" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L115" class="blob-num js-line-number" data-line-number="115"></td>
        <td id="LC115" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L116" class="blob-num js-line-number" data-line-number="116"></td>
        <td id="LC116" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L117" class="blob-num js-line-number" data-line-number="117"></td>
        <td id="LC117" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\newtheorem</span>{hypothesis}{Hypothesis}</td>
      </tr>
      <tr>
        <td id="L118" class="blob-num js-line-number" data-line-number="118"></td>
        <td id="LC118" class="blob-code blob-code-inner js-file-line"><span class="pl-k">\usepackage</span>{<span class="pl-c1">setspace</span>}</td>
      </tr>
      <tr>
        <td id="L119" class="blob-num js-line-number" data-line-number="119"></td>
        <td id="LC119" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L120" class="blob-num js-line-number" data-line-number="120"></td>
        <td id="LC120" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\makeatletter</span></td>
      </tr>
      <tr>
        <td id="L121" class="blob-num js-line-number" data-line-number="121"></td>
        <td id="LC121" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\@ifpackageloaded</span>{hyperref}{}{<span class="pl-c"><span class="pl-c">%</span></span></td>
      </tr>
      <tr>
        <td id="L122" class="blob-num js-line-number" data-line-number="122"></td>
        <td id="LC122" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\ifxetex</span></td>
      </tr>
      <tr>
        <td id="L123" class="blob-num js-line-number" data-line-number="123"></td>
        <td id="LC123" class="blob-code blob-code-inner js-file-line">  <span class="pl-c1">\PassOptionsToPackage</span>{hyphens}{url}<span class="pl-c1">\usepackage</span>[setpagesize=false, <span class="pl-c"><span class="pl-c">%</span> page size defined by xetex</span></td>
      </tr>
      <tr>
        <td id="L124" class="blob-num js-line-number" data-line-number="124"></td>
        <td id="LC124" class="blob-code blob-code-inner js-file-line">              unicode=false, <span class="pl-c"><span class="pl-c">%</span> unicode breaks when used with xetex</span></td>
      </tr>
      <tr>
        <td id="L125" class="blob-num js-line-number" data-line-number="125"></td>
        <td id="LC125" class="blob-code blob-code-inner js-file-line">              xetex]{hyperref}</td>
      </tr>
      <tr>
        <td id="L126" class="blob-num js-line-number" data-line-number="126"></td>
        <td id="LC126" class="blob-code blob-code-inner js-file-line"><span class="pl-k">\else</span></td>
      </tr>
      <tr>
        <td id="L127" class="blob-num js-line-number" data-line-number="127"></td>
        <td id="LC127" class="blob-code blob-code-inner js-file-line">  <span class="pl-c1">\PassOptionsToPackage</span>{hyphens}{url}<span class="pl-k">\usepackage</span>[<span class="pl-v">unicode=true</span>]{<span class="pl-c1">hyperref</span>}</td>
      </tr>
      <tr>
        <td id="L128" class="blob-num js-line-number" data-line-number="128"></td>
        <td id="LC128" class="blob-code blob-code-inner js-file-line"><span class="pl-k">\fi</span></td>
      </tr>
      <tr>
        <td id="L129" class="blob-num js-line-number" data-line-number="129"></td>
        <td id="LC129" class="blob-code blob-code-inner js-file-line">}</td>
      </tr>
      <tr>
        <td id="L130" class="blob-num js-line-number" data-line-number="130"></td>
        <td id="LC130" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L131" class="blob-num js-line-number" data-line-number="131"></td>
        <td id="LC131" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\@ifpackageloaded</span>{color}{</td>
      </tr>
      <tr>
        <td id="L132" class="blob-num js-line-number" data-line-number="132"></td>
        <td id="LC132" class="blob-code blob-code-inner js-file-line">    <span class="pl-c1">\PassOptionsToPackage</span>{usenames,dvipsnames}{color}</td>
      </tr>
      <tr>
        <td id="L133" class="blob-num js-line-number" data-line-number="133"></td>
        <td id="LC133" class="blob-code blob-code-inner js-file-line">}{<span class="pl-c"><span class="pl-c">%</span></span></td>
      </tr>
      <tr>
        <td id="L134" class="blob-num js-line-number" data-line-number="134"></td>
        <td id="LC134" class="blob-code blob-code-inner js-file-line">    <span class="pl-k">\usepackage</span>[<span class="pl-v">usenames,dvipsnames</span>]{<span class="pl-c1">color</span>}</td>
      </tr>
      <tr>
        <td id="L135" class="blob-num js-line-number" data-line-number="135"></td>
        <td id="LC135" class="blob-code blob-code-inner js-file-line">}</td>
      </tr>
      <tr>
        <td id="L136" class="blob-num js-line-number" data-line-number="136"></td>
        <td id="LC136" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\makeatother</span></td>
      </tr>
      <tr>
        <td id="L137" class="blob-num js-line-number" data-line-number="137"></td>
        <td id="LC137" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\hypersetup</span>{breaklinks=true,</td>
      </tr>
      <tr>
        <td id="L138" class="blob-num js-line-number" data-line-number="138"></td>
        <td id="LC138" class="blob-code blob-code-inner js-file-line"><span class="pl-c"><span class="pl-c">%</span>            bookmarks=true,</span></td>
      </tr>
      <tr>
        <td id="L139" class="blob-num js-line-number" data-line-number="139"></td>
        <td id="LC139" class="blob-code blob-code-inner js-file-line">            pdfauthor={<span class="pl-s"><span class="pl-pds">$</span>if(anonymous)<span class="pl-pds">$</span><span class="pl-pds">$</span>else<span class="pl-pds">$</span><span class="pl-pds">$</span>for(author)<span class="pl-pds">$</span><span class="pl-pds">$</span>author.name<span class="pl-pds">$</span></span> (<span class="pl-s"><span class="pl-pds">$</span>author.affiliation<span class="pl-pds">$</span></span>)<span class="pl-s"><span class="pl-pds">$</span>sep<span class="pl-pds">$</span></span> and <span class="pl-s"><span class="pl-pds">$</span>endfor<span class="pl-pds">$</span><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span>},</td>
      </tr>
      <tr>
        <td id="L140" class="blob-num js-line-number" data-line-number="140"></td>
        <td id="LC140" class="blob-code blob-code-inner js-file-line">             pdfkeywords = {<span class="pl-s"><span class="pl-pds">$</span>if(keywords)<span class="pl-pds">$</span><span class="pl-pds">$</span>keywords<span class="pl-pds">$</span><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span>},  </td>
      </tr>
      <tr>
        <td id="L141" class="blob-num js-line-number" data-line-number="141"></td>
        <td id="LC141" class="blob-code blob-code-inner js-file-line">            pdftitle={<span class="pl-s"><span class="pl-pds">$</span>title<span class="pl-pds">$</span><span class="pl-pds">$</span>if(subtitle)<span class="pl-pds">$</span></span>: <span class="pl-s"><span class="pl-pds">$</span>subtitle<span class="pl-pds">$</span><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span>},</td>
      </tr>
      <tr>
        <td id="L142" class="blob-num js-line-number" data-line-number="142"></td>
        <td id="LC142" class="blob-code blob-code-inner js-file-line">            colorlinks=true,</td>
      </tr>
      <tr>
        <td id="L143" class="blob-num js-line-number" data-line-number="143"></td>
        <td id="LC143" class="blob-code blob-code-inner js-file-line">            citecolor=<span class="pl-s"><span class="pl-pds">$</span>if(citecolor)<span class="pl-pds">$</span><span class="pl-pds">$</span>citecolor<span class="pl-pds">$</span><span class="pl-pds">$</span>else<span class="pl-pds">$</span></span>blue<span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span>,</td>
      </tr>
      <tr>
        <td id="L144" class="blob-num js-line-number" data-line-number="144"></td>
        <td id="LC144" class="blob-code blob-code-inner js-file-line">            urlcolor=<span class="pl-s"><span class="pl-pds">$</span>if(urlcolor)<span class="pl-pds">$</span><span class="pl-pds">$</span>urlcolor<span class="pl-pds">$</span><span class="pl-pds">$</span>else<span class="pl-pds">$</span></span>blue<span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span>,</td>
      </tr>
      <tr>
        <td id="L145" class="blob-num js-line-number" data-line-number="145"></td>
        <td id="LC145" class="blob-code blob-code-inner js-file-line">            linkcolor=<span class="pl-s"><span class="pl-pds">$</span>if(linkcolor)<span class="pl-pds">$</span><span class="pl-pds">$</span>linkcolor<span class="pl-pds">$</span><span class="pl-pds">$</span>else<span class="pl-pds">$</span></span>magenta<span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span>,</td>
      </tr>
      <tr>
        <td id="L146" class="blob-num js-line-number" data-line-number="146"></td>
        <td id="LC146" class="blob-code blob-code-inner js-file-line">            pdfborder={0 0 0}}</td>
      </tr>
      <tr>
        <td id="L147" class="blob-num js-line-number" data-line-number="147"></td>
        <td id="LC147" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\urlstyle</span>{same}  <span class="pl-c"><span class="pl-c">%</span> don&#39;t use monospace font for urls</span></td>
      </tr>
      <tr>
        <td id="L148" class="blob-num js-line-number" data-line-number="148"></td>
        <td id="LC148" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L149" class="blob-num js-line-number" data-line-number="149"></td>
        <td id="LC149" class="blob-code blob-code-inner js-file-line"><span class="pl-c"><span class="pl-c">%</span> set default figure placement to htbp</span></td>
      </tr>
      <tr>
        <td id="L150" class="blob-num js-line-number" data-line-number="150"></td>
        <td id="LC150" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\makeatletter</span></td>
      </tr>
      <tr>
        <td id="L151" class="blob-num js-line-number" data-line-number="151"></td>
        <td id="LC151" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\def\fps@figure</span>{htbp}</td>
      </tr>
      <tr>
        <td id="L152" class="blob-num js-line-number" data-line-number="152"></td>
        <td id="LC152" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\makeatother</span></td>
      </tr>
      <tr>
        <td id="L153" class="blob-num js-line-number" data-line-number="153"></td>
        <td id="LC153" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L154" class="blob-num js-line-number" data-line-number="154"></td>
        <td id="LC154" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>for(header-includes)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L155" class="blob-num js-line-number" data-line-number="155"></td>
        <td id="LC155" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>header-includes<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L156" class="blob-num js-line-number" data-line-number="156"></td>
        <td id="LC156" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endfor<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L157" class="blob-num js-line-number" data-line-number="157"></td>
        <td id="LC157" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L158" class="blob-num js-line-number" data-line-number="158"></td>
        <td id="LC158" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(endnotes)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L159" class="blob-num js-line-number" data-line-number="159"></td>
        <td id="LC159" class="blob-code blob-code-inner js-file-line"><span class="pl-k">\usepackage</span>{<span class="pl-c1">endnotes</span>}</td>
      </tr>
      <tr>
        <td id="L160" class="blob-num js-line-number" data-line-number="160"></td>
        <td id="LC160" class="blob-code blob-code-inner js-file-line"><span class="pl-k">\renewcommand</span>{<span class="pl-c1">\enotesize</span>}{<span class="pl-c1">\normalsize</span>}</td>
      </tr>
      <tr>
        <td id="L161" class="blob-num js-line-number" data-line-number="161"></td>
        <td id="LC161" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\let\footnote</span>=<span class="pl-c1">\endnote</span></td>
      </tr>
      <tr>
        <td id="L162" class="blob-num js-line-number" data-line-number="162"></td>
        <td id="LC162" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L163" class="blob-num js-line-number" data-line-number="163"></td>
        <td id="LC163" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L164" class="blob-num js-line-number" data-line-number="164"></td>
        <td id="LC164" class="blob-code blob-code-inner js-file-line"><span class="pl-c"><span class="pl-c">%</span> add tightlist ----------</span></td>
      </tr>
      <tr>
        <td id="L165" class="blob-num js-line-number" data-line-number="165"></td>
        <td id="LC165" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\providecommand</span>{<span class="pl-c1">\tightlist</span>}{<span class="pl-c"><span class="pl-c">%</span></span></td>
      </tr>
      <tr>
        <td id="L166" class="blob-num js-line-number" data-line-number="166"></td>
        <td id="LC166" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\setlength</span>{<span class="pl-c1">\itemsep</span>}{0pt}<span class="pl-c1">\setlength</span>{<span class="pl-c1">\parskip</span>}{0pt}}</td>
      </tr>
      <tr>
        <td id="L167" class="blob-num js-line-number" data-line-number="167"></td>
        <td id="LC167" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L168" class="blob-num js-line-number" data-line-number="168"></td>
        <td id="LC168" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\begin</span>{<span class="pl-smi">document</span>}</td>
      </tr>
      <tr>
        <td id="L169" class="blob-num js-line-number" data-line-number="169"></td>
        <td id="LC169" class="blob-code blob-code-inner js-file-line">	</td>
      </tr>
      <tr>
        <td id="L170" class="blob-num js-line-number" data-line-number="170"></td>
        <td id="LC170" class="blob-code blob-code-inner js-file-line"><span class="pl-c"><span class="pl-c">%</span> \pagenumbering{arabic}% resets `page` counter to 1 </span></td>
      </tr>
      <tr>
        <td id="L171" class="blob-num js-line-number" data-line-number="171"></td>
        <td id="LC171" class="blob-code blob-code-inner js-file-line"><span class="pl-c"><span class="pl-c">%</span>$if(appendix)$</span></td>
      </tr>
      <tr>
        <td id="L172" class="blob-num js-line-number" data-line-number="172"></td>
        <td id="LC172" class="blob-code blob-code-inner js-file-line"><span class="pl-c"><span class="pl-c">%</span>\renewcommand*{\thepage}{$appendixletter$--\arabic{page}}</span></td>
      </tr>
      <tr>
        <td id="L173" class="blob-num js-line-number" data-line-number="173"></td>
        <td id="LC173" class="blob-code blob-code-inner js-file-line"><span class="pl-c"><span class="pl-c">%</span>$endif$    </span></td>
      </tr>
      <tr>
        <td id="L174" class="blob-num js-line-number" data-line-number="174"></td>
        <td id="LC174" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L175" class="blob-num js-line-number" data-line-number="175"></td>
        <td id="LC175" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(title)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L176" class="blob-num js-line-number" data-line-number="176"></td>
        <td id="LC176" class="blob-code blob-code-inner js-file-line"><span class="pl-c"><span class="pl-c">%</span> \maketitle</span></td>
      </tr>
      <tr>
        <td id="L177" class="blob-num js-line-number" data-line-number="177"></td>
        <td id="LC177" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L178" class="blob-num js-line-number" data-line-number="178"></td>
        <td id="LC178" class="blob-code blob-code-inner js-file-line">{<span class="pl-c"><span class="pl-c">%</span> \usefont{T1}{pnc}{m}{n}</span></td>
      </tr>
      <tr>
        <td id="L179" class="blob-num js-line-number" data-line-number="179"></td>
        <td id="LC179" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\setlength</span>{<span class="pl-c1">\parindent</span>}{0pt}</td>
      </tr>
      <tr>
        <td id="L180" class="blob-num js-line-number" data-line-number="180"></td>
        <td id="LC180" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\thispagestyle</span>{plain}</td>
      </tr>
      <tr>
        <td id="L181" class="blob-num js-line-number" data-line-number="181"></td>
        <td id="LC181" class="blob-code blob-code-inner js-file-line">{<span class="pl-c1">\fontsize</span>{18}{20}<span class="pl-c1">\selectfont\raggedright</span> </td>
      </tr>
      <tr>
        <td id="L182" class="blob-num js-line-number" data-line-number="182"></td>
        <td id="LC182" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\maketitle</span>  <span class="pl-c"><span class="pl-c">%</span> title \par  </span></td>
      </tr>
      <tr>
        <td id="L183" class="blob-num js-line-number" data-line-number="183"></td>
        <td id="LC183" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L184" class="blob-num js-line-number" data-line-number="184"></td>
        <td id="LC184" class="blob-code blob-code-inner js-file-line">}</td>
      </tr>
      <tr>
        <td id="L185" class="blob-num js-line-number" data-line-number="185"></td>
        <td id="LC185" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L186" class="blob-num js-line-number" data-line-number="186"></td>
        <td id="LC186" class="blob-code blob-code-inner js-file-line">{</td>
      </tr>
      <tr>
        <td id="L187" class="blob-num js-line-number" data-line-number="187"></td>
        <td id="LC187" class="blob-code blob-code-inner js-file-line">   <span class="pl-c1">\vskip</span> 13.5pt<span class="pl-c1">\relax</span> <span class="pl-c1">\normalsize\fontsize</span>{11}{12} </td>
      </tr>
      <tr>
        <td id="L188" class="blob-num js-line-number" data-line-number="188"></td>
        <td id="LC188" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(anonymous)<span class="pl-pds">$</span></span><span class="pl-c1">\hfill</span> <span class="pl-s"><span class="pl-pds">$</span>else<span class="pl-pds">$</span><span class="pl-pds">$</span>for(author)<span class="pl-pds">$</span></span><span class="pl-c1">\textbf</span>{<span class="pl-mb"><span class="pl-c1">\authorfont</span> <span class="pl-s"><span class="pl-pds">$</span>author.name<span class="pl-pds">$</span></span></span>} <span class="pl-c1">\hskip</span> 15pt <span class="pl-c1">\emph</span>{<span class="pl-mi"><span class="pl-c1">\small</span> <span class="pl-s"><span class="pl-pds">$</span>author.affiliation<span class="pl-pds">$</span></span></span>}  <span class="pl-s"><span class="pl-pds">$</span>sep<span class="pl-pds">$</span></span> <span class="pl-c1">\par</span> <span class="pl-s"><span class="pl-pds">$</span>endfor<span class="pl-pds">$</span></span> <span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L189" class="blob-num js-line-number" data-line-number="189"></td>
        <td id="LC189" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L190" class="blob-num js-line-number" data-line-number="190"></td>
        <td id="LC190" class="blob-code blob-code-inner js-file-line">}</td>
      </tr>
      <tr>
        <td id="L191" class="blob-num js-line-number" data-line-number="191"></td>
        <td id="LC191" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L192" class="blob-num js-line-number" data-line-number="192"></td>
        <td id="LC192" class="blob-code blob-code-inner js-file-line">}</td>
      </tr>
      <tr>
        <td id="L193" class="blob-num js-line-number" data-line-number="193"></td>
        <td id="LC193" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L194" class="blob-num js-line-number" data-line-number="194"></td>
        <td id="LC194" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L195" class="blob-num js-line-number" data-line-number="195"></td>
        <td id="LC195" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L196" class="blob-num js-line-number" data-line-number="196"></td>
        <td id="LC196" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L197" class="blob-num js-line-number" data-line-number="197"></td>
        <td id="LC197" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L198" class="blob-num js-line-number" data-line-number="198"></td>
        <td id="LC198" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L199" class="blob-num js-line-number" data-line-number="199"></td>
        <td id="LC199" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(abstract)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L200" class="blob-num js-line-number" data-line-number="200"></td>
        <td id="LC200" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L201" class="blob-num js-line-number" data-line-number="201"></td>
        <td id="LC201" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L202" class="blob-num js-line-number" data-line-number="202"></td>
        <td id="LC202" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L203" class="blob-num js-line-number" data-line-number="203"></td>
        <td id="LC203" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\begin</span>{<span class="pl-smi">abstract</span>}</td>
      </tr>
      <tr>
        <td id="L204" class="blob-num js-line-number" data-line-number="204"></td>
        <td id="LC204" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L205" class="blob-num js-line-number" data-line-number="205"></td>
        <td id="LC205" class="blob-code blob-code-inner js-file-line">    <span class="pl-c1">\hbox</span>{<span class="pl-c1">\vrule</span> height .2pt width 39.14pc}</td>
      </tr>
      <tr>
        <td id="L206" class="blob-num js-line-number" data-line-number="206"></td>
        <td id="LC206" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L207" class="blob-num js-line-number" data-line-number="207"></td>
        <td id="LC207" class="blob-code blob-code-inner js-file-line">    <span class="pl-c1">\vskip</span> 8.5pt <span class="pl-c"><span class="pl-c">%</span> \small </span></td>
      </tr>
      <tr>
        <td id="L208" class="blob-num js-line-number" data-line-number="208"></td>
        <td id="LC208" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L209" class="blob-num js-line-number" data-line-number="209"></td>
        <td id="LC209" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\noindent</span> <span class="pl-s"><span class="pl-pds">$</span>abstract<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L210" class="blob-num js-line-number" data-line-number="210"></td>
        <td id="LC210" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L211" class="blob-num js-line-number" data-line-number="211"></td>
        <td id="LC211" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(keywords)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L212" class="blob-num js-line-number" data-line-number="212"></td>
        <td id="LC212" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L213" class="blob-num js-line-number" data-line-number="213"></td>
        <td id="LC213" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\vskip</span> 8.5pt <span class="pl-c1">\noindent</span> <span class="pl-c1">\emph</span>{<span class="pl-mi">Keywords</span>}: <span class="pl-s"><span class="pl-pds">$</span>keywords<span class="pl-pds">$</span></span> <span class="pl-c1">\par</span></td>
      </tr>
      <tr>
        <td id="L214" class="blob-num js-line-number" data-line-number="214"></td>
        <td id="LC214" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L215" class="blob-num js-line-number" data-line-number="215"></td>
        <td id="LC215" class="blob-code blob-code-inner js-file-line">    <span class="pl-c1">\hbox</span>{<span class="pl-c1">\vrule</span> height .2pt width 39.14pc}</td>
      </tr>
      <tr>
        <td id="L216" class="blob-num js-line-number" data-line-number="216"></td>
        <td id="LC216" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L217" class="blob-num js-line-number" data-line-number="217"></td>
        <td id="LC217" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L218" class="blob-num js-line-number" data-line-number="218"></td>
        <td id="LC218" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>else<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L219" class="blob-num js-line-number" data-line-number="219"></td>
        <td id="LC219" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L220" class="blob-num js-line-number" data-line-number="220"></td>
        <td id="LC220" class="blob-code blob-code-inner js-file-line">    <span class="pl-c1">\hbox</span>{<span class="pl-c1">\vrule</span> height .2pt width 39.14pc}</td>
      </tr>
      <tr>
        <td id="L221" class="blob-num js-line-number" data-line-number="221"></td>
        <td id="LC221" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L222" class="blob-num js-line-number" data-line-number="222"></td>
        <td id="LC222" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L223" class="blob-num js-line-number" data-line-number="223"></td>
        <td id="LC223" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L224" class="blob-num js-line-number" data-line-number="224"></td>
        <td id="LC224" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\end</span>{<span class="pl-smi">abstract</span>}</td>
      </tr>
      <tr>
        <td id="L225" class="blob-num js-line-number" data-line-number="225"></td>
        <td id="LC225" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L226" class="blob-num js-line-number" data-line-number="226"></td>
        <td id="LC226" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L227" class="blob-num js-line-number" data-line-number="227"></td>
        <td id="LC227" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L228" class="blob-num js-line-number" data-line-number="228"></td>
        <td id="LC228" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\vskip</span> 6.5pt</td>
      </tr>
      <tr>
        <td id="L229" class="blob-num js-line-number" data-line-number="229"></td>
        <td id="LC229" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L230" class="blob-num js-line-number" data-line-number="230"></td>
        <td id="LC230" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>if(toc)<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L231" class="blob-num js-line-number" data-line-number="231"></td>
        <td id="LC231" class="blob-code blob-code-inner js-file-line">{</td>
      </tr>
      <tr>
        <td id="L232" class="blob-num js-line-number" data-line-number="232"></td>
        <td id="LC232" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\hypersetup</span>{linkcolor=black}</td>
      </tr>
      <tr>
        <td id="L233" class="blob-num js-line-number" data-line-number="233"></td>
        <td id="LC233" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\setcounter</span>{tocdepth}{<span class="pl-s"><span class="pl-pds">$</span>toc-depth<span class="pl-pds">$</span></span>}</td>
      </tr>
      <tr>
        <td id="L234" class="blob-num js-line-number" data-line-number="234"></td>
        <td id="LC234" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\tableofcontents</span></td>
      </tr>
      <tr>
        <td id="L235" class="blob-num js-line-number" data-line-number="235"></td>
        <td id="LC235" class="blob-code blob-code-inner js-file-line">}</td>
      </tr>
      <tr>
        <td id="L236" class="blob-num js-line-number" data-line-number="236"></td>
        <td id="LC236" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>endif<span class="pl-pds">$</span></span></td>
      </tr>
      <tr>
        <td id="L237" class="blob-num js-line-number" data-line-number="237"></td>
        <td id="LC237" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L238" class="blob-num js-line-number" data-line-number="238"></td>
        <td id="LC238" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\noindent</span> <span class="pl-s"><span class="pl-pds">$</span>if(anonymous)<span class="pl-pds">$</span></span><span class="pl-c1">\doublespacing</span><span class="pl-s"><span class="pl-pds">$</span>else<span class="pl-pds">$</span><span class="pl-pds">$</span>if(spacing)<span class="pl-pds">$</span></span><span class="pl-cce">\$</span>spacing<span class="pl-s"><span class="pl-pds">$</span>spacing<span class="pl-pds">$</span></span>endif<span class="pl-s"><span class="pl-pds">$$</span>endif<span class="pl-s"><span class="pl-pds">$</span> <span class="pl-pds">$</span></span>body<span class="pl-s"><span class="pl-pds">$</span></span></span></td>
      </tr>
      <tr>
        <td id="L239" class="blob-num js-line-number" data-line-number="239"></td>
        <td id="LC239" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-s"></span></span></td>
      </tr>
      <tr>
        <td id="L240" class="blob-num js-line-number" data-line-number="240"></td>
        <td id="LC240" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-s"><span class="pl-pds">$</span></span>if(endnotes)<span class="pl-s"><span class="pl-pds">$</span></span></span></td>
      </tr>
      <tr>
        <td id="L241" class="blob-num js-line-number" data-line-number="241"></td>
        <td id="LC241" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-s"><span class="pl-c1">\newpage</span></span></span></td>
      </tr>
      <tr>
        <td id="L242" class="blob-num js-line-number" data-line-number="242"></td>
        <td id="LC242" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-s"></span></span></td>
      </tr>
      <tr>
        <td id="L243" class="blob-num js-line-number" data-line-number="243"></td>
        <td id="LC243" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-s"><span class="pl-c1">\theendnotes</span></span></span></td>
      </tr>
      <tr>
        <td id="L244" class="blob-num js-line-number" data-line-number="244"></td>
        <td id="LC244" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-s"><span class="pl-pds">$</span></span>endif<span class="pl-s"><span class="pl-pds">$</span></span></span></td>
      </tr>
      <tr>
        <td id="L245" class="blob-num js-line-number" data-line-number="245"></td>
        <td id="LC245" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-s"></span></span></td>
      </tr>
      <tr>
        <td id="L246" class="blob-num js-line-number" data-line-number="246"></td>
        <td id="LC246" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-s"></span></span></td>
      </tr>
      <tr>
        <td id="L247" class="blob-num js-line-number" data-line-number="247"></td>
        <td id="LC247" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-s"></span></span></td>
      </tr>
      <tr>
        <td id="L248" class="blob-num js-line-number" data-line-number="248"></td>
        <td id="LC248" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-s"><span class="pl-c1">\newpage</span></span></span></td>
      </tr>
      <tr>
        <td id="L249" class="blob-num js-line-number" data-line-number="249"></td>
        <td id="LC249" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-s"><span class="pl-c1">\singlespacing</span> </span></span></td>
      </tr>
      <tr>
        <td id="L250" class="blob-num js-line-number" data-line-number="250"></td>
        <td id="LC250" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-s"><span class="pl-pds">$</span></span>if(natbib)<span class="pl-s"><span class="pl-pds">$</span></span></span></td>
      </tr>
      <tr>
        <td id="L251" class="blob-num js-line-number" data-line-number="251"></td>
        <td id="LC251" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-s"><span class="pl-pds">$</span></span>if(bibliography)<span class="pl-s"><span class="pl-pds">$</span></span></span></td>
      </tr>
      <tr>
        <td id="L252" class="blob-num js-line-number" data-line-number="252"></td>
        <td id="LC252" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-s"><span class="pl-pds">$</span></span>if(biblio-title)<span class="pl-s"><span class="pl-pds">$</span></span></span></td>
      </tr>
      <tr>
        <td id="L253" class="blob-num js-line-number" data-line-number="253"></td>
        <td id="LC253" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-s"><span class="pl-pds">$</span></span>if(book-class)<span class="pl-s"><span class="pl-pds">$</span></span></span></td>
      </tr>
      <tr>
        <td id="L254" class="blob-num js-line-number" data-line-number="254"></td>
        <td id="LC254" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-s"><span class="pl-c1">\renewcommand\bibname</span>{<span class="pl-s"><span class="pl-pds">$</span>biblio-title<span class="pl-pds">$</span></span>}</span></span></td>
      </tr>
      <tr>
        <td id="L255" class="blob-num js-line-number" data-line-number="255"></td>
        <td id="LC255" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-s"><span class="pl-pds">$</span></span>else<span class="pl-s"><span class="pl-pds">$</span></span></span></td>
      </tr>
      <tr>
        <td id="L256" class="blob-num js-line-number" data-line-number="256"></td>
        <td id="LC256" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-s"><span class="pl-c1">\renewcommand\refname</span>{<span class="pl-s"><span class="pl-pds">$</span>biblio-title<span class="pl-pds">$</span></span>}</span></span></td>
      </tr>
      <tr>
        <td id="L257" class="blob-num js-line-number" data-line-number="257"></td>
        <td id="LC257" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-s"><span class="pl-pds">$</span></span>endif<span class="pl-s"><span class="pl-pds">$</span></span></span></td>
      </tr>
      <tr>
        <td id="L258" class="blob-num js-line-number" data-line-number="258"></td>
        <td id="LC258" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-s"><span class="pl-pds">$</span></span>endif<span class="pl-s"><span class="pl-pds">$</span></span></span></td>
      </tr>
      <tr>
        <td id="L259" class="blob-num js-line-number" data-line-number="259"></td>
        <td id="LC259" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-s"><span class="pl-c1">\bibliography</span>{<span class="pl-s"><span class="pl-pds">$</span>for(bibliography)<span class="pl-pds">$</span><span class="pl-pds">$</span>bibliography<span class="pl-pds">$</span><span class="pl-pds">$</span>sep<span class="pl-pds">$</span></span>,<span class="pl-s"><span class="pl-pds">$</span>endfor<span class="pl-pds">$</span></span>}</span></span></td>
      </tr>
      <tr>
        <td id="L260" class="blob-num js-line-number" data-line-number="260"></td>
        <td id="LC260" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-s"></span></span></td>
      </tr>
      <tr>
        <td id="L261" class="blob-num js-line-number" data-line-number="261"></td>
        <td id="LC261" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-s"><span class="pl-pds">$</span></span>endif<span class="pl-s"><span class="pl-pds">$</span></span></span></td>
      </tr>
      <tr>
        <td id="L262" class="blob-num js-line-number" data-line-number="262"></td>
        <td id="LC262" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-s"><span class="pl-pds">$</span></span>endif<span class="pl-s"><span class="pl-pds">$</span></span></span></td>
      </tr>
      <tr>
        <td id="L263" class="blob-num js-line-number" data-line-number="263"></td>
        <td id="LC263" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-s"><span class="pl-pds">$</span></span>if(biblatex)<span class="pl-s"><span class="pl-pds">$</span></span></span></td>
      </tr>
      <tr>
        <td id="L264" class="blob-num js-line-number" data-line-number="264"></td>
        <td id="LC264" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-s"><span class="pl-c1">\printbibliography</span><span class="pl-pds">$</span></span>if(biblio-title)<span class="pl-s"><span class="pl-pds">$</span>[title=<span class="pl-pds">$</span></span>biblio-title<span class="pl-s"><span class="pl-pds">$</span>]<span class="pl-pds">$</span></span>endif<span class="pl-s"><span class="pl-pds">$</span></span></span></td>
      </tr>
      <tr>
        <td id="L265" class="blob-num js-line-number" data-line-number="265"></td>
        <td id="LC265" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-s"></span></span></td>
      </tr>
      <tr>
        <td id="L266" class="blob-num js-line-number" data-line-number="266"></td>
        <td id="LC266" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-s"><span class="pl-pds">$</span></span>endif<span class="pl-s"><span class="pl-pds">$</span></span></span></td>
      </tr>
      <tr>
        <td id="L267" class="blob-num js-line-number" data-line-number="267"></td>
        <td id="LC267" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-s"><span class="pl-pds">$</span></span>for(include-after)<span class="pl-s"><span class="pl-pds">$</span></span></span></td>
      </tr>
      <tr>
        <td id="L268" class="blob-num js-line-number" data-line-number="268"></td>
        <td id="LC268" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-s"><span class="pl-pds">$</span></span>include-after<span class="pl-s"><span class="pl-pds">$</span></span></span></td>
      </tr>
      <tr>
        <td id="L269" class="blob-num js-line-number" data-line-number="269"></td>
        <td id="LC269" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-s"></span></span></td>
      </tr>
      <tr>
        <td id="L270" class="blob-num js-line-number" data-line-number="270"></td>
        <td id="LC270" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-s"><span class="pl-pds">$</span></span>endfor<span class="pl-s"><span class="pl-pds">$</span></span></span></td>
      </tr>
      <tr>
        <td id="L271" class="blob-num js-line-number" data-line-number="271"></td>
        <td id="LC271" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-s"><span class="pl-c1">\end</span>{document}</span></span></td>
      </tr>
</table>

  <details class="details-reset details-overlay BlobToolbar position-absolute js-file-line-actions dropdown d-none" aria-hidden="true">
    <summary class="btn-octicon ml-0 px-2 p-0 bg-white border border-gray-dark rounded-1" aria-label="Inline file action toolbar">
      <svg class="octicon octicon-kebab-horizontal" viewBox="0 0 13 16" version="1.1" width="13" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M1.5 9a1.5 1.5 0 100-3 1.5 1.5 0 000 3zm5 0a1.5 1.5 0 100-3 1.5 1.5 0 000 3zM13 7.5a1.5 1.5 0 11-3 0 1.5 1.5 0 013 0z"/></svg>
    </summary>
    <details-menu>
      <ul class="BlobToolbar-dropdown dropdown-menu dropdown-menu-se mt-2" style="width:185px">
        <li>
          <clipboard-copy role="menuitem" class="dropdown-item" id="js-copy-lines" style="cursor:pointer;">
            Copy lines
          </clipboard-copy>
        </li>
        <li>
          <clipboard-copy role="menuitem" class="dropdown-item" id="js-copy-permalink" style="cursor:pointer;">
            Copy permalink
          </clipboard-copy>
        </li>
        <li><a class="dropdown-item js-update-url-with-hash" id="js-view-git-blame" role="menuitem" href="/junkpolisci/website/blame/8c72f5f6bdda3784b8d521ae15603285cbcdbfd9/working_content/svm-latex-ms2.tex">View git blame</a></li>
          <li><a class="dropdown-item" id="js-new-issue" role="menuitem" href="/junkpolisci/website/issues/new">Reference in new issue</a></li>
      </ul>
    </details-menu>
  </details>

  </div>

    </div>

  <details class="details-reset details-overlay details-overlay-dark">
    <summary data-hotkey="l" aria-label="Jump to line"></summary>
    <details-dialog class="Box Box--overlay d-flex flex-column anim-fade-in fast linejump" aria-label="Jump to line">
      <!-- '"` --><!-- </textarea></xmp> --></option></form><form class="js-jump-to-line-form Box-body d-flex" action="" accept-charset="UTF-8" method="get">
        <input class="form-control flex-auto mr-3 linejump-input js-jump-to-line-field" type="text" placeholder="Jump to line&hellip;" aria-label="Jump to line" autofocus>
        <button type="submit" class="btn" data-close-dialog>Go</button>
</form>    </details-dialog>
  </details>

  </div>
</div>

    </main>
  </div>

  </div>

<div class="footer container-lg width-full p-responsive" role="contentinfo">
  <div class="position-relative d-flex flex-row-reverse flex-lg-row flex-wrap flex-lg-nowrap flex-justify-center flex-lg-justify-between pt-6 pb-2 mt-6 f6 text-gray border-top border-gray-light ">
    <ul class="list-style-none d-flex flex-wrap col-12 col-lg-5 flex-justify-center flex-lg-justify-between mb-2 mb-lg-0">
      <li class="mr-3 mr-lg-0">&copy; 2020 GitHub, Inc.</li>
        <li class="mr-3 mr-lg-0"><a data-ga-click="Footer, go to terms, text:terms" href="https://github.com/site/terms">Terms</a></li>
        <li class="mr-3 mr-lg-0"><a data-ga-click="Footer, go to privacy, text:privacy" href="https://github.com/site/privacy">Privacy</a></li>
        <li class="mr-3 mr-lg-0"><a data-ga-click="Footer, go to security, text:security" href="https://github.com/security">Security</a></li>
        <li class="mr-3 mr-lg-0"><a href="https://githubstatus.com/" data-ga-click="Footer, go to status, text:status">Status</a></li>
        <li><a data-ga-click="Footer, go to help, text:help" href="https://help.github.com">Help</a></li>

    </ul>

    <a aria-label="Homepage" title="GitHub" class="footer-octicon d-none d-lg-block mx-lg-4" href="https://github.com">
      <svg height="24" class="octicon octicon-mark-github" viewBox="0 0 16 16" version="1.1" width="24" aria-hidden="true"><path fill-rule="evenodd" d="M8 0C3.58 0 0 3.58 0 8c0 3.54 2.29 6.53 5.47 7.59.4.07.55-.17.55-.38 0-.19-.01-.82-.01-1.49-2.01.37-2.53-.49-2.69-.94-.09-.23-.48-.94-.82-1.13-.28-.15-.68-.52-.01-.53.63-.01 1.08.58 1.23.82.72 1.21 1.87.87 2.33.66.07-.52.28-.87.51-1.07-1.78-.2-3.64-.89-3.64-3.95 0-.87.31-1.59.82-2.15-.08-.2-.36-1.02.08-2.12 0 0 .67-.21 2.2.82.64-.18 1.32-.27 2-.27.68 0 1.36.09 2 .27 1.53-1.04 2.2-.82 2.2-.82.44 1.1.16 1.92.08 2.12.51.56.82 1.27.82 2.15 0 3.07-1.87 3.75-3.65 3.95.29.25.54.73.54 1.48 0 1.07-.01 1.93-.01 2.2 0 .21.15.46.55.38A8.013 8.013 0 0016 8c0-4.42-3.58-8-8-8z"/></svg>
</a>
   <ul class="list-style-none d-flex flex-wrap col-12 col-lg-5 flex-justify-center flex-lg-justify-between mb-2 mb-lg-0">
        <li class="mr-3 mr-lg-0"><a data-ga-click="Footer, go to contact, text:contact" href="https://github.com/contact">Contact GitHub</a></li>
        <li class="mr-3 mr-lg-0"><a href="https://github.com/pricing" data-ga-click="Footer, go to Pricing, text:Pricing">Pricing</a></li>
      <li class="mr-3 mr-lg-0"><a href="https://developer.github.com" data-ga-click="Footer, go to api, text:api">API</a></li>
      <li class="mr-3 mr-lg-0"><a href="https://training.github.com" data-ga-click="Footer, go to training, text:training">Training</a></li>
        <li class="mr-3 mr-lg-0"><a href="https://github.blog" data-ga-click="Footer, go to blog, text:blog">Blog</a></li>
        <li><a data-ga-click="Footer, go to about, text:about" href="https://github.com/about">About</a></li>
    </ul>
  </div>
  <div class="d-flex flex-justify-center pb-6">
    <span class="f6 text-gray-light"></span>
  </div>
</div>

  <div id="ajax-error-message" class="ajax-error-message flash flash-error">
    <svg class="octicon octicon-alert" viewBox="0 0 16 16" version="1.1" width="16" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M8.893 1.5c-.183-.31-.52-.5-.887-.5s-.703.19-.886.5L.138 13.499a.98.98 0 000 1.001c.193.31.53.501.886.501h13.964c.367 0 .704-.19.877-.5a1.03 1.03 0 00.01-1.002L8.893 1.5zm.133 11.497H6.987v-2.003h2.039v2.003zm0-3.004H6.987V5.987h2.039v4.006z"/></svg>
    <button type="button" class="flash-close js-ajax-error-dismiss" aria-label="Dismiss error">
      <svg class="octicon octicon-x" viewBox="0 0 12 16" version="1.1" width="12" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M7.48 8l3.75 3.75-1.48 1.48L6 9.48l-3.75 3.75-1.48-1.48L4.52 8 .77 4.25l1.48-1.48L6 6.52l3.75-3.75 1.48 1.48L7.48 8z"/></svg>
    </button>
    You can’t perform that action at this time.
  </div>

    <script crossorigin="anonymous" async="async" integrity="sha512-o4vS4IKrjdy/HD+xr2+VhO6DxQmj5jikhHbEGrd8+JGhpmIOxRrpT1Qo5k3IhKimm8VXIu3pyYejLtOAkm+OsQ==" type="application/javascript" id="js-conditional-compat" data-src="https://github.githubassets.com/assets/compat-bootstrap-a38bd2e0.js"></script>
    <script crossorigin="anonymous" integrity="sha512-6XqOrpzsRfeWz1MuH9q2GuzW4Ktvt+kA5KbChOp1ZjaoGbRl3tBng8HiA5B/lClMvkkj4h+vVuSwLeh14JzGuA==" type="application/javascript" src="https://github.githubassets.com/assets/environment-bootstrap-e97a8eae.js"></script>
    <script crossorigin="anonymous" async="async" integrity="sha512-1/6VVx6z9r6uphSoGlmYgbqD5KaY+GVMt1Gqa3DIa0U+3Pv2SWu8Fk1BZ2xPne5upvF8HdEWcGeiUjd2URl+oA==" type="application/javascript" src="https://github.githubassets.com/assets/vendor-d7fe9557.js"></script>
    <script crossorigin="anonymous" async="async" integrity="sha512-RDggGUYWJq0pjfr/60y4ITVdK0zFDicDs0h46xLtUfsgDq6L6OWInB0F66615pE295U74v8ykPH/qsoL3U38ng==" type="application/javascript" src="https://github.githubassets.com/assets/frameworks-44382019.js"></script>

    <script crossorigin="anonymous" async="async" integrity="sha512-aIrRO9Sji3VKn3Usk8M8E+xiMDESpBJZj3BSwvg2Y9GHNyL/g57gL5srCoN6u5+WgvUBkHyFtthyr+IzaQVXWQ==" type="application/javascript" src="https://github.githubassets.com/assets/github-bootstrap-688ad13b.js"></script>

  <div class="js-stale-session-flash flash flash-warn flash-banner" hidden
    >
    <svg class="octicon octicon-alert" viewBox="0 0 16 16" version="1.1" width="16" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M8.893 1.5c-.183-.31-.52-.5-.887-.5s-.703.19-.886.5L.138 13.499a.98.98 0 000 1.001c.193.31.53.501.886.501h13.964c.367 0 .704-.19.877-.5a1.03 1.03 0 00.01-1.002L8.893 1.5zm.133 11.497H6.987v-2.003h2.039v2.003zm0-3.004H6.987V5.987h2.039v4.006z"/></svg>
    <span class="js-stale-session-flash-signed-in" hidden>You signed in with another tab or window. <a href="">Reload</a> to refresh your session.</span>
    <span class="js-stale-session-flash-signed-out" hidden>You signed out in another tab or window. <a href="">Reload</a> to refresh your session.</span>
  </div>
  <template id="site-details-dialog">
  <details class="details-reset details-overlay details-overlay-dark lh-default text-gray-dark hx_rsm" open>
    <summary role="button" aria-label="Close dialog"></summary>
    <details-dialog class="Box Box--overlay d-flex flex-column anim-fade-in fast hx_rsm-dialog hx_rsm-modal">
      <button class="Box-btn-octicon m-0 btn-octicon position-absolute right-0 top-0" type="button" aria-label="Close dialog" data-close-dialog>
        <svg class="octicon octicon-x" viewBox="0 0 12 16" version="1.1" width="12" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M7.48 8l3.75 3.75-1.48 1.48L6 9.48l-3.75 3.75-1.48-1.48L4.52 8 .77 4.25l1.48-1.48L6 6.52l3.75-3.75 1.48 1.48L7.48 8z"/></svg>
      </button>
      <div class="octocat-spinner my-6 js-details-dialog-spinner"></div>
    </details-dialog>
  </details>
</template>

  <div class="Popover js-hovercard-content position-absolute" style="display: none; outline: none;" tabindex="0">
  <div class="Popover-message Popover-message--bottom-left Popover-message--large Box box-shadow-large" style="width:360px;">
  </div>
</div>

  <div aria-live="polite" class="js-global-screen-reader-notice sr-only"></div>

  </body>
</html>

### 2023-05-01 Codes/Trajipaq_participation_economique_an1.Rmd
update

---
title: "Effet du parcours migratoire sur la participation économique au Québec"
subtitle: "Transformation des fichiers en longitudinal"
author:
- Vissého Adjiwanou ^[corresponding author - adjiwanou.vissého@uqam.ca]
- Département de Sociologie, Université du Québec à Montréal (UQAM)
- Solène Lardoux (Département de démographie, Université de Montréal)
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  word_document: default
  html_document: 
    toc: true
    toc_depth: 3
  template: svm-latex-ms2.tex 
  pdf_document: default
  citation_package: natbib
bibliography: Covid_Canada.bib
---

## Résumé

- Décrire les différents parcours économiques des immigrants depuis leurs arrivées au Canada
- Voir comment ces parcours diffèrent selon la région de provenance

<!--
Ce fichier est pour les statistiques descriptives. 
La base de données provient de Trajipaq_covid_cr1
Ce fichier est une extension de Trajipaq_covid_an2 où le statut d'immigration est remplacé
par une combinaison de statut et de minorité visible.

Couleur palette graphique : https://coolors.co/b224e5-f489a1-ffe385-f6fdf4-6c885e-b0c4c3-9cc8ac-afecb1-60ade1-95d3ff
-->

# 1. Parcours professionnel

```{r}

rm(list = ls())

library(tidyverse)
library(summarytools)
library(srvyr)
library(survey)
library(ggpubr)
library(writexl)
library(lubridate)
library(haven)

trajipaq_eco <- read_dta("../Data/Base_Trajipaq_Finale_Ok.dta")

```

## Sélection des variables importantes

```{r}
# Emploi actuel
freq(trajipaq_eco$Q79)

# Nombre d'emploi occupé depuis l'installation
freq(trajipaq_eco$Q80)

# Premier emploi
freq(trajipaq_eco$Q81_1)

# Emploi encore en cours
freq(trajipaq_eco$Q82_1)

# Deuxième emploi
freq(trajipaq_eco$Q81_2)

# Quel était votre statut dans cet emploi
freq(trajipaq_eco$Q83_10)

# Par semaine, combien d’heures en tout travaillez-vous habituellement dans ce deuxième emploi ?
freq(trajipaq_eco$Q89_1)

# Par quel moyen aviez-vous trouvé cette activité

freq(trajipaq_eco$Q91_1O)

```

```{r}

traj_short_eco <-
  trajipaq_eco %>% 
  select(CaseId, Q79, Q80, starts_with(c("Q81", "Q82", "Q83", "Q84", "Q85", "Q86", "Q87", "Q88", "Q89", "Q90")))

```

## Pivot longer

```{r}

class(traj_short_eco$Q83_1)
freq(traj_short_eco$Q80N)

#, "Q83", "Q84", "Q85", "Q86", "Q87", "Q88", "Q89", "Q90"
traj_short_eco_long <-
  traj_short_eco %>% 
  select(CaseId, Q79, Q80, starts_with(c("Q81"))) %>% 
  pivot_longer(cols = starts_with(c("Q81")),
               names_to = "rang",
               names_prefix = "Q81_",
               values_to = "Q81",
               values_drop_na = T)

traj_short_eco <-
  traj_short_eco %>% 
  mutate(Q83autre_1 = Q83_1O,
         Q83autre_2 = Q83_2O,
         Q83autre_3 = Q83_3O,
         Q83autre_4 = Q83_4O,
         Q83autre_5 = Q83_5O,
         Q83autre_6 = Q83_6O,
         Q83autre_7 = Q83_7O,
         Q83autre_8 = Q83_8O,
         Q83autre_9 = Q83_9O,
         Q83autre_10 = Q83_10O)

traj_short_eco <-
  traj_short_eco %>% 
  mutate(Q84autre_1 = Q84_1O,
         Q84autre_2 = Q84_2O,
         Q84autre_3 = Q84_3O,
         Q84autre_4 = Q84_4O,
         Q84autre_5 = Q84_5O,
         Q84autre_6 = Q84_6O,
         Q84autre_7 = Q84_7O,
         Q84autre_8 = Q84_8O,
         Q84autre_9 = Q84_9O,
         Q84autre_10 = Q84_10O)

traj_short_eco <-
  traj_short_eco %>% 
  mutate(Q88autre_1 = Q88_1O,
         Q88autre_2 = Q88_2O,
         Q88autre_3 = Q88_3O,
         Q88autre_4 = Q88_4O,
         Q88autre_5 = Q88_5O,
         Q88autre_6 = Q88_6O,
         Q88autre_7 = Q88_7O,
         Q88autre_8 = Q88_8O,
         Q88autre_9 = Q88_9O,
         Q88autre_10 = Q88_10O)

## Q82X n'est pas bien présenté dans la base de données

traj_short_eco_long1 <-
  traj_short_eco %>% 
  mutate(Q82X_1 = Q82X) %>% 
  select(!Q82X) %>% 
  select(CaseId, Q79, Q80, starts_with(c("Q81_", "Q82", "Q85", "Q86", "Q87")), Q83_1, Q83_2, Q83_3, Q83_4, Q83_5, Q83_6, Q83_7, Q83_8, Q83_9, Q83_10, starts_with("Q83autre"), Q84_1, Q84_2, Q84_3, Q84_4, Q84_5, Q84_6, Q84_7, Q84_8, Q84_9, Q84_10, starts_with("Q84autre"), Q88_1, Q88_2, Q88_3, Q88_4, Q88_5, Q88_6, Q88_7, Q88_8, Q88_9, Q88_10, starts_with(c("Q88autre", "Q88X")), Q89_1, Q89_2, Q89_3, Q89_4, Q89_5, Q89_6, Q89_7, Q89_8, Q89_9, Q89_10, Q90) %>% 
  pivot_longer(cols = starts_with(c("Q81", "Q82X", "Q82", "Q83", "Q83autre", "Q84", "Q84autre", "Q85", "Q86", "Q87", "Q88", "Q88autre", "Q88X", "Q89")), 
               names_to = c(".value", "Rang"),
               names_sep = "_",
               values_drop_na = T) %>% 
  filter(!is.na(Q81))

```

Q82X n'est pas clair

## Additional variables

```{r}

# Durée de l'emploi

```

## Recodage des variables

## Pondération

```{r}

#trajipaq_nm <- na.omit(trajipaq_covid)
#trajipaq.design <- svydesign(id = ~ CaseId, weights = ~ Poidstot1, data = trajipaq_nm)
trajipaq.design <- svydesign(id = ~ CaseId, weights = ~ Poidstot1, data = trajipaq_covid)

```

## Analyse descriptive

### Distribution des fréquences des variables

```{r}

freq(trajipaq_covid$minorite_vis)

freq(trajipaq_covid$statut_imm, weights = trajipaq_covid$Poidstot1)
freq(trajipaq_covid$minorite_vis, weights = trajipaq_covid$Poidstot1)

ctable(trajipaq_covid$statut_imm, trajipaq_covid$minorite_vis, weights = trajipaq_covid$Poidstot1, "t")

distribution <- svymean(~statut_imm + minorite_vis + duree_res + sexe + gpage + niv_etude + statut_mat + statut_emploi, trajipaq.design, na.rm = TRUE)
distribution

length(distribution)

# Distribution de l'échantillon selon le statut migratoire
distribution_statut <- svyby(~ minorite_vis + duree_res + sexe + gpage + niv_etude + statut_mat + statut_emploi, ~ statut_imm, trajipaq.design, svymean, na.rm = TRUE)
distribution_statut

# Distribution immigrants
distribution_statut_immigrant_mean <-
  distribution_statut %>% 
  pivot_longer(!statut_imm, names_to = "Variables", values_to = "Proportion") %>% 
  filter(!str_detect(statut_imm, "Non") & !str_detect(Variables, "se")) %>% 
  arrange(Variables, statut_imm)
distribution_statut_immigrant_mean

distribution_statut_immigrant_se <-
  distribution_statut %>% 
  pivot_longer(!statut_imm, names_to = "Variables", values_to = "SE") %>% 
  filter(!str_detect(statut_imm, "Non") & str_starts(Variables, "se")) %>% 
  mutate(Variables = str_remove(Variables, "se.")) %>% 
  arrange(Variables, statut_imm)
distribution_statut_immigrant_se

distribution_statut_immigrant <- left_join(distribution_statut_immigrant_mean, distribution_statut_immigrant_se, by = c("statut_imm", "Variables"))

distribution_statut_immigrant

# Distribution majoritaire

distribution_statut_majoritaire_mean <-
  distribution_statut %>% 
#  select(statut_imm)
  pivot_longer(col=!statut_imm, names_to = "Variables", values_to = "Proportion") %>% 
  filter(str_starts(statut_imm, "Non") & !str_detect(Variables, "se")) %>% 
  arrange(Variables, statut_imm)
distribution_statut_majoritaire_mean

distribution_statut_majoritaire_se <-
  distribution_statut %>% 
  pivot_longer(!statut_imm, names_to = "Variables", values_to = "SE") %>% 
  filter(str_starts(statut_imm, "Non") & str_starts(Variables, "se")) %>% 
  mutate(Variables = str_remove(Variables, "se.")) %>% 
  arrange(Variables, statut_imm)
distribution_statut_majoritaire_se

distribution_statut_majoritaire <- left_join(distribution_statut_majoritaire_mean, distribution_statut_majoritaire_se, by = c("statut_imm", "Variables"))

distribution_statut_majoritaire
distribution_statut_immigrant

# Total

distribution_sample <- left_join(distribution_statut_immigrant, distribution_statut_majoritaire, by = "Variables")
distribution_sample

write_xlsx(distribution_sample, "distribution_sample.xlsx")

# Il reste à ajouter le total

```

- Même chose avec utilisation imm_minorité qui tient compte des deux variables de minorité et d'immigrant en même temps

```{r}

freq(trajipaq_covid$statut_imm, weights = trajipaq_covid$Poidstot1)

distribution1 <- svymean(~imm_minorite + minorite_vis + duree_res + sexe + gpage + niv_etude + statut_mat + statut_emploi, trajipaq.design, na.rm = TRUE)
distribution1

length(distribution)

# Distribution de l'échantillon selon le statut migratoire
distribution_statut1 <- svyby(~ minorite_vis + duree_res + sexe + gpage + niv_etude + statut_mat + statut_emploi, ~ imm_minorite, trajipaq.design, svymean, na.rm = TRUE)
distribution_statut1

# Distribution immigrants
distribution_statut_immigrant_mean1 <-
  distribution_statut1 %>% 
  pivot_longer(!imm_minorite, names_to = "Variables", values_to = "Proportion") %>% 
  filter(!str_detect(imm_minorite, "Non") & !str_detect(Variables, "se")) %>% 
  arrange(Variables, imm_minorite)
distribution_statut_immigrant_mean1

distribution_statut_immigrant_se1 <-
  distribution_statut1 %>% 
  pivot_longer(!imm_minorite, names_to = "Variables", values_to = "SE") %>% 
  filter(!str_detect(imm_minorite, "Non") & str_starts(Variables, "se")) %>% 
  mutate(Variables = str_remove(Variables, "se.")) %>% 
  arrange(Variables, imm_minorite)
distribution_statut_immigrant_se1

distribution_statut_immigrant1 <- left_join(distribution_statut_immigrant_mean1, distribution_statut_immigrant_se1, by = c("imm_minorite", "Variables"))

distribution_statut_immigrant1

# Distribution majoritaire

distribution_statut_majoritaire_mean1 <-
  distribution_statut1 %>% 
#  select(imm_minorite)
  pivot_longer(col=!imm_minorite, names_to = "Variables", values_to = "Proportion") %>% 
  filter(str_starts(imm_minorite, "Non") & !str_detect(Variables, "se")) %>% 
  arrange(Variables, imm_minorite)
distribution_statut_majoritaire_mean1

distribution_statut_majoritaire_se1 <-
  distribution_statut1 %>% 
  pivot_longer(!imm_minorite, names_to = "Variables", values_to = "SE") %>% 
  filter(str_starts(imm_minorite, "Non") & str_starts(Variables, "se")) %>% 
  mutate(Variables = str_remove(Variables, "se.")) %>% 
  arrange(Variables, imm_minorite)
distribution_statut_majoritaire_se1

distribution_statut_majoritaire1 <- left_join(distribution_statut_majoritaire_mean1, distribution_statut_majoritaire_se1, by = c("imm_minorite", "Variables"))

distribution_statut_majoritaire1
distribution_statut_immigrant1

# Total

distribution_sample1 <- left_join(distribution_statut_immigrant1, distribution_statut_majoritaire1, by = "Variables")
distribution_sample1

write_xlsx(distribution_sample1, "distribution_sample1.xlsx")

# Il reste à ajouter le total

```

## Tableau croisé entre la situation d'immigrants et les variables de covid

#### Graphique

#### Fonction graphique

```{r}
cbPalette <- c("#56B4E9", "#E69F00", "#999999", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

# Fonction
graph_traj <- function(resultat){
  ggplot(resultat) +
  geom_col(aes(x = reorder(Variables, desc(Proportion)), y = Proportion, fill = imm_minorite), position = "dodge") +
  #geom_errorbar(aes(ymin=len-sd, ymax=len+sd), width=.2, position=position_dodge(.9))
  theme_bw() +
  theme(legend.position = "bottom") +
  coord_flip() +
    labs(x = "", fill = "") +
  scale_fill_manual(values = cbPalette)
}

graph_traj2 <- function(resultat){
  ggplot(resultat) +
    geom_col(aes(x = Variables, y = Proportion, 
                 fill = fct_relevel(Modalités, "Much degraded",  "A little degraded",
                                    "No change", "A little improved", "Much improved"),
                 position = "fill")) +
    facet_wrap(~imm_minorite) +
#    scale_x_discrete("Modalités", limits = c("Much improved", "A little improved",
#    "No change", "A little degraded", "Much degraded")) +
    theme_bw() +
    theme(legend.position = "bottom", legend.box="vertical", legend.margin=margin()) +
    guides(fill=guide_legend(nrow=2,byrow=TRUE)) +
#    theme(strip.background = element_rect(colour = "black", fill = "blue")) +
    coord_flip() +
    labs(x = "", fill = "") +
    #scale_fill_manual(values = cbPalette) + 
    scale_fill_brewer(palette="PRGn")
}

#gg+guides(fill=guide_legend(nrow=2,byrow=TRUE))

```

### Exposition à la maladie

<!--
freq(trajipaq_covid$statut_imm, weights = trajipaq_covid$Poidstot1)

with(trajipaq_covid,
     ctable(statut_imm, e_hors_dom, weights = Poidstot1, useNA = "no"))

with(trajipaq_covid,
     ctable(statut_imm, e_four_soin, weights = Poidstot1))

with(trajipaq_covid,
     ctable(statut_imm, e_four_soin, weights = Poidstot1))

freq(trajipaq_covid$statut_imm)
freq(trajipaq_covid$e_hors_dom)

#struc_ssa <- svyby(~structure, ~childgpage, dhs.design5, svymean)#, keep.names = FALSE, verbose = FALSE)

freq(trajipaq_covid$e_hors_dom)
ctable(trajipaq_covid$statut_imm, trajipaq_covid$e_hors_dom)
-->

```{r}

exposition <- "e_hors_dom + e_four_soin + e_quarantaine + e_expose + e_symptome + e_hospitalise + e_intensif + e_decede + expo"
exposition
exposition[1]
noquote(exposition)[1]
class(exposition)

summary(trajipaq_covid$Poidstot1)

traj_exposition <- svyby(~ e_hors_dom + e_four_soin + e_quarantaine + e_expose + e_symptome + e_hospitalise + e_intensif + e_decede + expo, ~ imm_minorite, trajipaq.design, svymean, na.rm = TRUE)
traj_exposition

traj_exposition <-
  traj_exposition %>% 
  pivot_longer(col=!imm_minorite, names_to = "Variables", values_to = "Proportion") %>% 
  filter(str_detect(Variables, "oui") & !str_detect(Variables, "se")) %>% 
  arrange(Variables)

traj_exposition

traj_exposition <-
  traj_exposition %>% 
  mutate(Variables = case_when(
    Variables == "e_decedeoui" ~ "A family member has dead",
    Variables == "e_four_soinoui" ~ "Work in health facility",
    Variables == "e_hors_domoui" ~ "Work outside",
    Variables == "e_intensifoui" ~ "In intensive care",
    Variables == "e_quarantaineoui" ~ "Put in quarantine",
    Variables == "e_symptomeoui" ~ "Has symptoms"))

traj_exposition

graph_traj_exposition <- graph_traj(traj_exposition)

ggsave(graph_traj_exposition, file = "Figures/Exposition1.png")
#ggsave(graph_traj_exposition, file = "Figures/Exposition2.png")

graph_traj_exposition
```

### Conséquence sur la famille

- On getting ressources

- We had to leave our home
- We had difficulty obtaining food
- We had trouble getting medicine
- We had difficulty getting health care when we needed it
**- Our family income has decreased**
- We lost health insurance / benefits

```{r}

# f_separe + f_emmenage + f_revenu +

cons_famille <- "f_quitte_dom + f_nourriture + f_medicaments + f_sante +  f_assurance + f_evenement + f_essentiel"

#f_quitte_dom +  f_separe + f_emmenage + + f_revenu + f_evenement

traj_famille <- svyby(~f_nourriture + f_medicaments + f_sante + f_assurance + f_essentiel, ~ imm_minorite, trajipaq.design, svymean, na.rm = TRUE)
traj_famille

traj_famille <-
  traj_famille %>% 
  pivot_longer(!imm_minorite, names_to = "Variables", values_to = "Proportion") %>% 
  filter(str_detect(Variables, "oui") & !str_detect(Variables, "se")) %>% 
  arrange(Variables)
traj_famille

traj_famille <-
  traj_famille %>% 
  mutate(Variables = case_when(
    #Variables == "f_separeoui" ~ "Live separed",
    #Variables == "f_emmenageoui" ~ "Someone moved to our home",
    #Variables == "f_quitte_domoui" ~ "Leave our home",
    Variables == "f_nourritureoui" ~ "Difficulty obtaining food",
    Variables == "f_medicamentsoui" ~ "Difficulty getting medicine",
    Variables == "f_santeoui" ~ "Difficulty getting health care",
    #Variables == "f_revenuoui" ~ "Family income has decreased",
    Variables == "f_assuranceoui" ~ "Lost family insurrance",
#    Variables == "f_evenementoui" ~ "Missed important family event",
    Variables == "f_essentieloui" ~ "difficulty obtaining other essentials"))

traj_exposition
# Représentation

graph_traj_famille <- graph_traj(traj_famille)

ggsave(graph_traj_famille, file = "Figures/Famille1.png")
graph_traj_famille

```

### Conséquences emploi et revenu

```{r}

freq(trajipaq_covid$f_cesser)

traj_emploi <- svyby(~ f_reduire + f_cesser + f_perdu + f_revenu, ~ imm_minorite, trajipaq.design, svymean, na.rm = TRUE)
traj_emploi

traj_emploi <-
  traj_emploi %>% 
  pivot_longer(!imm_minorite, names_to = "Variables", values_to = "Proportion") %>% 
  filter(str_detect(Variables, "oui") & !str_starts(Variables, "se")) %>% 
#  replace(Variables, values = c("f_cesseroui", "f_perdreoui", "f_reduireoui"), list = c("Cesser de travailler", "Perdre son emploi", "Réduire le temps de travail"))
  arrange(Variables)

traj_emploi

traj_emploi <-
  traj_emploi %>% 
  mutate(Variables = case_when(
    Variables == "f_reduireoui" ~ "Has reduced working hours",
    Variables == "f_cesseroui" ~ "Has to stop working",
    Variables == "f_perduoui" ~ "Has lost their job permanently",
    Variables == "f_revenuoui" ~ "Family income has decreased"))

#A family member

# Représentation

graph_traj_emploi <- graph_traj(traj_emploi)
graph_traj_emploi

ggsave(graph_traj_emploi, file = "Figures/Emploi1.png")

```

### Entente dans la famille

*On ability to perform works*

COVID-19 can have many impacts on you and your family life. In general, how has the COVID-19 pandemic affected each of the following?

- Our family lived separately for health, safety or employment reasons
- How family members get along
- Ability to care for your child with an illness / condition
- Ability to take care of other children in your family
- Ability to care for the elderly or disabled in your family
- We missed an important family event or it was canceled (for example, wedding, graduation, birth, funeral, travel (including vacations), other

```{r}

traj_entente <- svyby(~ f_separe + f_role_parent + f_entende + f_enfant_maladie + f_autre_enfant + f_pers_agees + f_evenement, ~ imm_minorite, trajipaq.design, svymean, na.rm = TRUE)
traj_entente

traj_entente <-
  traj_entente %>% 
  pivot_longer(!imm_minorite, names_to = "Variables", values_to = "Proportion") %>% 
  filter(!str_starts(Variables, "se")) %>% 
#  separate(Variables, into = c("Variables", "Modalités"), sep = "[[:upper:]]") %>% 
  separate(Variables, into = c("Variables", "Modalités"), sep = "(?=[A-Z])") %>% 
  arrange(Variables)
traj_entente

freq(traj_entente$Modalités)

traj_entente <-
  traj_entente %>% 
  filter(Modalités != "Not applicable") %>% 
  mutate(Variables = case_when(
    Variables == "f_role_parent" ~ "Parenting",
    Variables == "f_separeoui" ~ "Live separed",
    Variables == "f_entende" ~ "How family members get along",
    Variables == "f_enfant_maladie" ~ "Ability to care for your child with an illness",
    Variables == "f_autre_enfant" ~ "Ability to take care of other children",
    Variables == "f_pers_agees" ~ "Ability to care for the elderly or disabled",
    Variables == "f_evenementoui" ~ "Missed important family event"))

#Modalités == factor(Modalités, levels = c("Much improved", "A little improved",
#                                              "No change", "A little degraded", 
#                                              "Much degraded")))

# Représentation

graph_traj_entente <- graph_traj2(traj_entente)
graph_traj_entente

ggsave(graph_traj_entente, file = "Figures/Entente1.png")

```

### Bien-être

```{r}

traj_bienetre <- svyby(~ be_exercice + be_manger + be_dormir + be_anxiete + be_humeur, 
                       ~ imm_minorite, trajipaq.design, svymean, na.rm = TRUE)
traj_bienetre

traj_bienetre <-
  traj_bienetre %>% 
  pivot_longer(!imm_minorite, names_to = "Variables", values_to = "Proportion") %>% 
  filter(!str_starts(Variables, "se")) %>% 
#  separate(Variables, into = c("Variables", "Modalités"), sep = "[[:upper:]]") %>% 
  separate(Variables, into = c("Variables", "Modalités"), sep = "(?=[A-Z])") %>% 
  arrange(Variables)
traj_bienetre

traj_bienetre <-
  traj_bienetre %>% 
  filter(Modalités != "Not applicable") %>% 
  mutate(Variables = case_when(
    Variables == "be_exercice" ~ "Your physical well-being - exercise",
    Variables == "be_manger" ~ "Your physical well-being - eat",
    Variables == "be_dormir" ~ "Your physical well-being - sleep",
    Variables == "be_anxiete" ~ "Your emotional well-being - anxiety",
    Variables == "be_humeur" ~ "Your emotional well-being - mood"))

graph_traj_bienetre <- graph_traj2(traj_bienetre)
graph_traj_bienetre

ggsave(graph_traj_bienetre, file = "Figures/Bien-être1.png")

```

### Aide reçu

```{r}

traj_aide <- svyby(~ patt + pcu,~ imm_minorite, trajipaq.design, svymean, na.rm = TRUE)
traj_aide

traj_aide <-
  traj_aide %>% 
  pivot_longer(!imm_minorite, names_to = "Variables", values_to = "Proportion") %>% 
  filter(str_detect(Variables, "Oui") & !str_starts(Variables, "se")) %>% 
#  replace(Variables, values = c("f_cesseroui", "f_perdreoui", "f_reduireoui"), list = c("Cesser de travailler", "Perdre son emploi", "Réduire le temps de travail"))
  arrange(Variables)
traj_aide

traj_aide <-
  traj_aide %>% 
  mutate(Variables = case_when(
    Variables == "pattOui" ~ "Temporary Aid for Workers Program",
    Variables == "pcuOui" ~ "Canada Emergency Response Benefit"))
#A family member

graph_traj_aide <- graph_traj(traj_aide)
graph_traj_aide

ggsave(graph_traj_aide, file = "Figures/Aide1.png")

```

## Covid data from statistics canada

```{r}

covid_canada <- read_csv("../Data/covid19_Canada_281121.csv")

quebec <- covid_canada %>% 
  filter(prname == "Quebec") %>% 
  mutate(Period = format_ISO8601(date, precision = "ym")) %>% 
  group_by(Period) %>% 
  summarise('Number of confirmed cases' = sum(numconf))

Quebec_cases <- ggplot(quebec) +
  geom_line(aes(x = Period, y = `Number of confirmed cases`, group = 1), color = "blue") +
  geom_vline(xintercept = "2020-08", linetype = "twodash") +
  geom_vline(xintercept = "2020-11", linetype = "twodash") +
#  geom_ribbon(aes(ymin = 0, ymax = 200000, xmin = "2020-08", xmax = "2020-11")) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
Quebec_cases

ggsave(Quebec_cases, file = "Figures/Quebec_cases1.png")

```

### 2023-05-01 Codes/Trajipaq_covid_an2.Rmd
update

--- 2023-05-01 Codes/Trajipaq_participation_economique_an1.Rmd
@@
 ---
-title: "Effet du parcours migratoire sur la participation économique au Québec"
-subtitle: "Transformation des fichiers en longitudinal"
+title: "Familial consequences of the Covid-19 in Québec: Comparison between immigrants and natives"
+subtitle: "Conséquences familiales de la Covid-19 au Québec: comparaison immigrants - natifs"
 author:
@@
 - Solène Lardoux (Département de démographie, Université de Montréal)
+- Nima Zahedinameghi (Département de Sociologie, UQAM)
 date: "`r format(Sys.time(), '%d %B %Y')`"
@@
 
-## Résumé
-
-- Décrire les différents parcours économiques des immigrants depuis leurs arrivées au Canada
-- Voir comment ces parcours diffèrent selon la région de provenance
-
 <!--
@@
 -->
-
-# 1. Parcours professionnel
 
@@
 library(lubridate)
-library(haven)
 
-trajipaq_eco <- read_dta("../Data/Base_Trajipaq_Finale_Ok.dta")
+trajipaq_covid <- read_rds("../Data/trajipaq_covid.RData")
 
-```
+freq(trajipaq_covid$imm_minorite)
 
-## Sélection des variables importantes
+trajipaq_covid <-
+  trajipaq_covid %>% 
+  filter(imm_minorite != "Others")
 
-```{r}
-# Emploi actuel
-freq(trajipaq_eco$Q79)
-
-# Nombre d'emploi occupé depuis l'installation
-freq(trajipaq_eco$Q80)
-
-# Premier emploi
-freq(trajipaq_eco$Q81_1)
-
-# Emploi encore en cours
-freq(trajipaq_eco$Q82_1)
-
-# Deuxième emploi
-freq(trajipaq_eco$Q81_2)
-
-# Quel était votre statut dans cet emploi
-freq(trajipaq_eco$Q83_10)
-
-# Par semaine, combien d’heures en tout travaillez-vous habituellement dans ce deuxième emploi ?
-freq(trajipaq_eco$Q89_1)
-
-# Par quel moyen aviez-vous trouvé cette activité
-
-freq(trajipaq_eco$Q91_1O)
-
-```
-
-```{r}
-
-traj_short_eco <-
-  trajipaq_eco %>% 
-  select(CaseId, Q79, Q80, starts_with(c("Q81", "Q82", "Q83", "Q84", "Q85", "Q86", "Q87", "Q88", "Q89", "Q90")))
-
-```
-
-## Pivot longer
-
-```{r}
-
-class(traj_short_eco$Q83_1)
-freq(traj_short_eco$Q80N)
-
-#, "Q83", "Q84", "Q85", "Q86", "Q87", "Q88", "Q89", "Q90"
-traj_short_eco_long <-
-  traj_short_eco %>% 
-  select(CaseId, Q79, Q80, starts_with(c("Q81"))) %>% 
-  pivot_longer(cols = starts_with(c("Q81")),
-               names_to = "rang",
-               names_prefix = "Q81_",
-               values_to = "Q81",
-               values_drop_na = T)
-
-traj_short_eco <-
-  traj_short_eco %>% 
-  mutate(Q83autre_1 = Q83_1O,
-         Q83autre_2 = Q83_2O,
-         Q83autre_3 = Q83_3O,
-         Q83autre_4 = Q83_4O,
-         Q83autre_5 = Q83_5O,
-         Q83autre_6 = Q83_6O,
-         Q83autre_7 = Q83_7O,
-         Q83autre_8 = Q83_8O,
-         Q83autre_9 = Q83_9O,
-         Q83autre_10 = Q83_10O)
-
-traj_short_eco <-
-  traj_short_eco %>% 
-  mutate(Q84autre_1 = Q84_1O,
-         Q84autre_2 = Q84_2O,
-         Q84autre_3 = Q84_3O,
-         Q84autre_4 = Q84_4O,
-         Q84autre_5 = Q84_5O,
-         Q84autre_6 = Q84_6O,
-         Q84autre_7 = Q84_7O,
-         Q84autre_8 = Q84_8O,
-         Q84autre_9 = Q84_9O,
-         Q84autre_10 = Q84_10O)
-
-traj_short_eco <-
-  traj_short_eco %>% 
-  mutate(Q88autre_1 = Q88_1O,
-         Q88autre_2 = Q88_2O,
-         Q88autre_3 = Q88_3O,
-         Q88autre_4 = Q88_4O,
-         Q88autre_5 = Q88_5O,
-         Q88autre_6 = Q88_6O,
-         Q88autre_7 = Q88_7O,
-         Q88autre_8 = Q88_8O,
-         Q88autre_9 = Q88_9O,
-         Q88autre_10 = Q88_10O)
-
-## Q82X n'est pas bien présenté dans la base de données
-
-traj_short_eco_long1 <-
-  traj_short_eco %>% 
-  mutate(Q82X_1 = Q82X) %>% 
-  select(!Q82X) %>% 
-  select(CaseId, Q79, Q80, starts_with(c("Q81_", "Q82", "Q85", "Q86", "Q87")), Q83_1, Q83_2, Q83_3, Q83_4, Q83_5, Q83_6, Q83_7, Q83_8, Q83_9, Q83_10, starts_with("Q83autre"), Q84_1, Q84_2, Q84_3, Q84_4, Q84_5, Q84_6, Q84_7, Q84_8, Q84_9, Q84_10, starts_with("Q84autre"), Q88_1, Q88_2, Q88_3, Q88_4, Q88_5, Q88_6, Q88_7, Q88_8, Q88_9, Q88_10, starts_with(c("Q88autre", "Q88X")), Q89_1, Q89_2, Q89_3, Q89_4, Q89_5, Q89_6, Q89_7, Q89_8, Q89_9, Q89_10, Q90) %>%
-  pivot_longer(cols = starts_with(c("Q81", "Q82X", "Q82", "Q83", "Q83autre", "Q84", "Q84autre", "Q85", "Q86", "Q87", "Q88", "Q88autre", "Q88X", "Q89")), 
-               names_to = c(".value", "Rang"),
-               names_sep = "_",
-               values_drop_na = T) %>% 
-  filter(!is.na(Q81))
-
-```
-
-Q82X n'est pas clair
-
-## Additional variables
-
-```{r}
-
-# Durée de l'emploi
+freq(trajipaq_covid$imm_minorite)
 

### 2023-05-02 Codes/Trajipaq_covid_an1.Rmd

--- 2023-05-01 Codes/Trajipaq_covid_an2.Rmd
@@
 La base de données provient de Trajipaq_covid_cr1
-Ce fichier est une extension de Trajipaq_covid_an2 où le statut d'immigration est remplacé
-par une combinaison de statut et de minorité visible.
 
-Couleur palette graphique : https://coolors.co/b224e5-f489a1-ffe385-f6fdf4-6c885e-b0c4c3-9cc8ac-afecb1-60ade1-95d3ff
 -->
@@
 library(writexl)
-library(lubridate)
+library(haven)
 
-trajipaq_covid <- read_rds("../Data/trajipaq_covid.RData")
-
-freq(trajipaq_covid$imm_minorite)
-
-trajipaq_covid <-
-  trajipaq_covid %>% 
-  filter(imm_minorite != "Others")
-
-freq(trajipaq_covid$imm_minorite)
+Base_Trajipaq_Finale_Ok <- read_dta("~/GitHub/Projet_Immigration_Trajipaq/Data/Base_Trajipaq_Finale_Ok.dta")
 
@@
 
-freq(trajipaq_covid$minorite_vis)
-
 freq(trajipaq_covid$statut_imm, weights = trajipaq_covid$Poidstot1)
-freq(trajipaq_covid$minorite_vis, weights = trajipaq_covid$Poidstot1)
-
-ctable(trajipaq_covid$statut_imm, trajipaq_covid$minorite_vis, weights = trajipaq_covid$Poidstot1, "t")
 
@@
 ```{r}
-cbPalette <- c("#56B4E9", "#E69F00", "#999999", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
+cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
 
@@
   ggplot(resultat) +
-  geom_col(aes(x = reorder(Variables, desc(Proportion)), y = Proportion, fill = imm_minorite), position = "dodge") +
+  geom_col(aes(x = reorder(Variables, desc(Proportion)), y = Proportion, fill = statut_imm), position = "dodge") +
   #geom_errorbar(aes(ymin=len-sd, ymax=len+sd), width=.2, position=position_dodge(.9))
@@
                  position = "fill")) +
-    facet_wrap(~imm_minorite) +
+    facet_wrap(~statut_imm) +
 #    scale_x_discrete("Modalités", limits = c("Much improved", "A little improved",
@@
     theme(legend.position = "bottom", legend.box="vertical", legend.margin=margin()) +
-    guides(fill=guide_legend(nrow=2,byrow=TRUE)) +
 #    theme(strip.background = element_rect(colour = "black", fill = "blue")) +
@@
 }
-
-#gg+guides(fill=guide_legend(nrow=2,byrow=TRUE))
 
@@
 
-exposition <- "e_hors_dom + e_four_soin + e_quarantaine + e_expose + e_symptome + e_hospitalise + e_intensif + e_decede + expo"
+exposition <- "e_hors_dom + e_four_soin + e_quarantaine + e_expose + e_symptome + e_hospitalise + e_intensif + e_decede"
 exposition
@@
 
-traj_exposition <- svyby(~ e_hors_dom + e_four_soin + e_quarantaine + e_expose + e_symptome + e_hospitalise + e_intensif + e_decede + expo, ~ imm_minorite, trajipaq.design, svymean, na.rm = TRUE)
+traj_exposition <- svyby(~ e_hors_dom + e_four_soin + e_quarantaine + e_expose + e_symptome + e_hospitalise + e_intensif + e_decede, ~ statut_imm, trajipaq.design, svymean, na.rm = TRUE)
 traj_exposition
@@
   traj_exposition %>% 
-  pivot_longer(col=!imm_minorite, names_to = "Variables", values_to = "Proportion") %>% 
+  pivot_longer(col=!statut_imm, names_to = "Variables", values_to = "Proportion") %>% 
   filter(str_detect(Variables, "oui") & !str_detect(Variables, "se")) %>% 
@@
 graph_traj_exposition <- graph_traj(traj_exposition)
+graph_traj_exposition
 
-ggsave(graph_traj_exposition, file = "Figures/Exposition1.png")
-#ggsave(graph_traj_exposition, file = "Figures/Exposition2.png")
+ggsave(graph_traj_exposition, file = "Figures/Exposition.png")
 
-graph_traj_exposition
 ```
@@
 
-- On getting ressources
-
-- We had to leave our home
-- We had difficulty obtaining food
-- We had trouble getting medicine
-- We had difficulty getting health care when we needed it
-**- Our family income has decreased**
-- We lost health insurance / benefits
-
 ```{r}
 
-# f_separe + f_emmenage + f_revenu +
+cons_famille <- "f_separe + f_emmenage + f_quitte_dom + f_nourriture + f_medicaments + f_sante + f_revenu + f_assurance + f_evenement + f_essentiel"
 
-cons_famille <- "f_quitte_dom + f_nourriture + f_medicaments + f_sante +  f_assurance + f_evenement + f_essentiel"
-
-#f_quitte_dom +  f_separe + f_emmenage + + f_revenu + f_evenement
-
-traj_famille <- svyby(~f_nourriture + f_medicaments + f_sante + f_assurance + f_essentiel, ~ imm_minorite, trajipaq.design, svymean, na.rm = TRUE)
+traj_famille <- svyby(~ f_separe + f_emmenage + f_quitte_dom + f_nourriture + f_medicaments + f_sante + f_revenu + f_assurance + f_evenement + f_essentiel, ~ statut_imm, trajipaq.design, svymean, na.rm = TRUE)
 traj_famille
@@
   traj_famille %>% 
-  pivot_longer(!imm_minorite, names_to = "Variables", values_to = "Proportion") %>% 
+  pivot_longer(!statut_imm, names_to = "Variables", values_to = "Proportion") %>% 
   filter(str_detect(Variables, "oui") & !str_detect(Variables, "se")) %>% 
@@
   mutate(Variables = case_when(
-    #Variables == "f_separeoui" ~ "Live separed",
-    #Variables == "f_emmenageoui" ~ "Someone moved to our home",
-    #Variables == "f_quitte_domoui" ~ "Leave our home",
+    Variables == "f_separeoui" ~ "Live separed",
+    Variables == "f_emmenageoui" ~ "Someone moved to our home",
+    Variables == "f_quitte_domoui" ~ "Leave our home",
     Variables == "f_nourritureoui" ~ "Difficulty obtaining food",
@@
     Variables == "f_santeoui" ~ "Difficulty getting health care",
-    #Variables == "f_revenuoui" ~ "Family income has decreased",
+    Variables == "f_revenuoui" ~ "Family income has decreased",
     Variables == "f_assuranceoui" ~ "Lost family insurrance",
-#    Variables == "f_evenementoui" ~ "Missed important family event",
+    Variables == "f_evenementoui" ~ "Missed important family event",
     Variables == "f_essentieloui" ~ "difficulty obtaining other essentials"))
@@
 graph_traj_famille <- graph_traj(traj_famille)
+graph_traj_famille
 
-ggsave(graph_traj_famille, file = "Figures/Famille1.png")
-graph_traj_famille
+ggsave(graph_traj_famille, file = "Figures/Famille.png")
 
@@
 
-### Conséquences emploi et revenu
+### Conséquences emploi
 
@@
 
-traj_emploi <- svyby(~ f_reduire + f_cesser + f_perdu + f_revenu, ~ imm_minorite, trajipaq.design, svymean, na.rm = TRUE)
+traj_emploi <- svyby(~ f_reduire + f_cesser + f_perdu, ~ statut_imm, trajipaq.design, svymean, na.rm = TRUE)
 traj_emploi
@@
   traj_emploi %>% 
-  pivot_longer(!imm_minorite, names_to = "Variables", values_to = "Proportion") %>% 
+  pivot_longer(!statut_imm, names_to = "Variables", values_to = "Proportion") %>% 
   filter(str_detect(Variables, "oui") & !str_starts(Variables, "se")) %>% 
@@
     Variables == "f_cesseroui" ~ "Has to stop working",
-    Variables == "f_perduoui" ~ "Has lost their job permanently",
-    Variables == "f_revenuoui" ~ "Family income has decreased"))
-
+    Variables == "f_perduoui" ~ "Has lost their job permanently"))
 #A family member
@@
 
-ggsave(graph_traj_emploi, file = "Figures/Emploi1.png")
+ggsave(graph_traj_emploi, file = "Figures/Emploi.png")
 
@@
 
-*On ability to perform works*
-
-COVID-19 can have many impacts on you and your family life. In general, how has the COVID-19 pandemic affected each of the following?
-
-- Our family lived separately for health, safety or employment reasons
-- How family members get along
-- Ability to care for your child with an illness / condition
-- Ability to take care of other children in your family
-- Ability to care for the elderly or disabled in your family
-- We missed an important family event or it was canceled (for example, wedding, graduation, birth, funeral, travel (including vacations), other
-
 ```{r}
 
-traj_entente <- svyby(~ f_separe + f_role_parent + f_entende + f_enfant_maladie + f_autre_enfant + f_pers_agees + f_evenement, ~ imm_minorite, trajipaq.design, svymean, na.rm = TRUE)
+traj_entente <- svyby(~ f_role_parent + f_entende + f_enfant_maladie + f_autre_enfant + f_pers_agees, ~ statut_imm, trajipaq.design, svymean, na.rm = TRUE)
 traj_entente
@@
   traj_entente %>% 
-  pivot_longer(!imm_minorite, names_to = "Variables", values_to = "Proportion") %>% 
+  pivot_longer(!statut_imm, names_to = "Variables", values_to = "Proportion") %>% 
   filter(!str_starts(Variables, "se")) %>% 
@@
     Variables == "f_role_parent" ~ "Parenting",
-    Variables == "f_separeoui" ~ "Live separed",
     Variables == "f_entende" ~ "How family members get along",
@@
     Variables == "f_autre_enfant" ~ "Ability to take care of other children",
-    Variables == "f_pers_agees" ~ "Ability to care for the elderly or disabled",
-    Variables == "f_evenementoui" ~ "Missed important family event"))
+    Variables == "f_pers_agees" ~ "Ability to care for the elderly or disabled"))
 
@@
 
-ggsave(graph_traj_entente, file = "Figures/Entente1.png")
+ggsave(graph_traj_entente, file = "Figures/Entente.png")
 
@@
 traj_bienetre <- svyby(~ be_exercice + be_manger + be_dormir + be_anxiete + be_humeur, 
-                       ~ imm_minorite, trajipaq.design, svymean, na.rm = TRUE)
+                       ~ statut_imm, trajipaq.design, svymean, na.rm = TRUE)
 traj_bienetre
@@
   traj_bienetre %>% 
-  pivot_longer(!imm_minorite, names_to = "Variables", values_to = "Proportion") %>% 
+  pivot_longer(!statut_imm, names_to = "Variables", values_to = "Proportion") %>% 
   filter(!str_starts(Variables, "se")) %>% 
@@
 
-ggsave(graph_traj_bienetre, file = "Figures/Bien-être1.png")
+ggsave(graph_traj_bienetre, file = "Figures/Bien-être.png")
 
@@
 
-traj_aide <- svyby(~ patt + pcu,~ imm_minorite, trajipaq.design, svymean, na.rm = TRUE)
+traj_aide <- svyby(~ patt + pcu,~ statut_imm, trajipaq.design, svymean, na.rm = TRUE)
 traj_aide
@@
   traj_aide %>% 
-  pivot_longer(!imm_minorite, names_to = "Variables", values_to = "Proportion") %>% 
+  pivot_longer(!statut_imm, names_to = "Variables", values_to = "Proportion") %>% 
   filter(str_detect(Variables, "Oui") & !str_starts(Variables, "se")) %>% 
@@
 
-ggsave(graph_traj_aide, file = "Figures/Aide1.png")
+ggsave(graph_traj_aide, file = "Figures/Aide.png")
 
@@
 
-## Covid data from statistics canada
-
-```{r}
-
-covid_canada <- read_csv("../Data/covid19_Canada_281121.csv")
-
-quebec <- covid_canada %>% 
-  filter(prname == "Quebec") %>% 
-  mutate(Period = format_ISO8601(date, precision = "ym")) %>% 
-  group_by(Period) %>% 
-  summarise('Number of confirmed cases' = sum(numconf))
-
-Quebec_cases <- ggplot(quebec) +
-  geom_line(aes(x = Period, y = `Number of confirmed cases`, group = 1), color = "blue") +
-  geom_vline(xintercept = "2020-08", linetype = "twodash") +
-  geom_vline(xintercept = "2020-11", linetype = "twodash") +
-#  geom_ribbon(aes(ymin = 0, ymax = 200000, xmin = "2020-08", xmax = "2020-11")) +
-  theme_bw() +
-  theme(axis.text.x = element_text(angle = 60, hjust = 1))
-Quebec_cases
-
-ggsave(Quebec_cases, file = "Figures/Quebec_cases1.png")
-
-```
+## Modèle de régressions

### 2023-05-02 Expérimentations (Théa)/1. Expériementations (Théa).Rmd
Premières expérimentations
J'ai réalisé une régression logistique à partir de la base de donnée afin de prendre en main la fonction glm().

---
title: "Expérimentations"
author: "Théa Carlier"
date: "2023-05-02"
output: html_document
---

```{r}

rm(list = ls())

library(tidyverse)
library(summarytools)
library(srvyr)
library(survey)
library(ggpubr)
library(writexl)
library(haven)

Base_Trajipaq_Finale_Ok <- read_dta("~/GitHub/Projet_Immigration_Trajipaq/Data/Base_Trajipaq_Finale_Ok.dta")

```

# Régression logistique entre deux variables numériques (VD = nominal dichotomique, VI = ratio)

```{r}
#transformation de la variable COV12 (Nous avons eu de la difficulté à obtenir des soins de santé quand nous en avions besoin)
Base_Trajipaq_Finale_Ok <-
  Base_Trajipaq_Finale_Ok %>%
  mutate(COV12_2 = case_when(
         COV12 == 1 ~ 0,
         COV12 == 2 ~ 1
    ))

#régression logistique entre COV12_2 (VD) et Q14I (Depuis quelle année vivez-vous au Québec)
model1 <- glm(formula = COV12_2 ~ Q14I, family = "binomial", data = Base_Trajipaq_Finale_Ok)
summary(model1)

#Les résultats indiquent que le risque d'erreur alpha est supérieur à 5%. Conséquemment, la relation étudiée n'est pas significative.
```

### 2023-05-02 Codes/Trajipaq_covid_an1.Rmd
Revert "Trajipaq_covid_an1.Rmd"
J'ai restauré le fichier Trajipaq_covid_an1.Rmd sur lequel j'avais expérimenté hier soir.

@@
 library(writexl)
-library(haven)
 
-Base_Trajipaq_Finale_Ok <- read_dta("~/GitHub/Projet_Immigration_Trajipaq/Data/Base_Trajipaq_Finale_Ok.dta")
+trajipaq_covid <- read_rds("../Data/trajipaq_covid.RData")
 

### 2023-05-02 Expérimentations (Théa)/1. Expériementations (code).Rmd
Expérimentations sur l'environnement de travail
1. J'ai intégré Git directement à Rstudio
2. J'ai créé un projet Bookdown que je vais utiliser pour expérimenter avec un template LateX sur le draft de votre article

--- 2023-05-02 Expérimentations (Théa)/1. Expériementations (Théa).Rmd
@@
 
-#Les résultats indiquent que le risque d'erreur alpha est supérieur à 5%. Conséquemment, la relation étudiée n'est pas significative.
+#les résultats indiquent que le risque d'erreur alpha est supérieur à 5%. Conséquemment, la relation étudiée n'est pas significative.
 ```
+
+# Analyse de la variance (ANOVA) entre deux variables numériques (VD = ordinal, VI = ratio)
+
+```{r}
+#création d'un dataframe regroupant les variables COV35 (Votre bien-être émotionnel - humeur s'est beaucoup, un peu ou pas amélioré) et Q14I.
+df_1 <- Base_Trajipaq_Finale_Ok %>%
+  select(COV35, Q14I) %>%
+  drop_na()
+
+#ANOVA entre COV35 (VD) et Q14I (VI)
+model2 <- aov(Q14I ~ COV35, data = df_1)
+summary(model2)
+
+##les résultats indiquent que le risque d'erreur alpha est supérieur à 5%. Conséquemment, la relation étudiée n'est pas significative.
+```
+
+## Code alternatif pour l'analyse de la variance
+
+```{r}
+#l'utilisation de pipes évite de créer un dataframe.
+Base_Trajipaq_Finale_Ok %>%
+  select(COV35, Q14I) %>%
+  drop_na() %>%
+  aov(Q14I ~ COV35, data = .) %>%
+  summary()
+```

### 2023-05-02 Expérimentations (Théa)/2. Expérimentations (Bookdown)/Bookdown/01-introduction.Rmd
Multiple changes on Bookdown

--- 2023-05-01 Article/Trajipaq_Covid_Draft4.Rmd
@@
 ---
-title: 'COVID-19 Disparities in Quebec (Canada): A Comparative Analysis of Exposure
-  Risks, Consequences, and Access to Resources among Immigrants and Natives'
-author:
-- Vissého Adjiwanou ^[corresponding author - adjiwanou.vissého@uqam.ca]
-- Département de Sociologie, Université du Québec à Montréal (UQAM)
-- Georges NGalé Tchango (Département de démographie, Université de Montréal)
-- Solène Lardoux (Département de démographie, Université de Montréal)
-- Nima Zahedinameghi (Département de Sociologie, UQAM)
-date: "`r format(Sys.time(), '%d %B %Y')`"
-output:
-  word_document:
-    keep_md: yes
-  html_document:
-    toc: yes
-    toc_depth: 3
-  template: svm-latex-ms2.tex
-  pdf_document: default
-  citation_package: natbib
-bibliography:
-- covid_Nima.bib
-- Covid_Canada.bib
-- Covid_Family.bib
+knit: bookdown::render_book
+editor_options: 
+  markdown: 
+    wrap: 80
 ---
-<!--
-# Journal: family issues - SAGE
-https://journals-sagepub-com.proxy.bibliotheques.uqam.ca/author-instructions/JFI
-
-https://fishandwhistle.net/post/2020/getting-started-zotero-better-bibtex-rmarkdown/
-https://www.zotero.org/styles?fields=sociology
--->
-
-\newpage
-
-# Abstract
-
-The immigrant population is one of the most vulnerable in Canada, with a relatively high risk of unemployment, a low social support system, and linguistic barriers.
-However, their experiences with public health interventions during the COVID-19 pandemic have not been adequately explored.
-The purpose of this article is to analyze Covid-19 exposure risk and its family implications for both natives and immigrants in the province of Quebec.
-Additionally, we analyze immigrants' access to government aid and compare it to their native-born counterparts.
-The study consists of survey data collected from 1,000 immigrants and 500 non-immigrants Quebecers between August and November 2020.
-Our descriptive and exploratory analysis reveals significant differences between the two groups regarding exposure risk, family consequences, and access to resources.
-We conclude that the current pandemic exacerbates health disparities in the province, with particularly dire repercussions for immigrants’ families.
-We also discuss several socio-cultural strategies to mitigate inequality and recommend policies to address the needs of vulnerable populations.
-
-**Keywords**: Immigrants, exposure risk, government aid, health disparities, COVID-19, family disruptions
-
-<!--
-
-Using  script in Rmarkdown
-
-http://zevross.com/blog/2014/07/09/making-use-of-external-r-code-in-knitr-and-r-markdown/
-
-Managing my Zotero
-
-https://forums.zotero.org/discussion/35712/removing-pdfs-to-free-up-space
-
-https://www.zotero.org/support/sync
-
-https://www.zotero.org/support/preferences/sync#file_syncing
-
-https://www.zotero.org/support/sync#webdav
-
-https://www.zotero.org/support/preferences/advanced#linked_attachment_base_directory
-
-http://zotfile.com/
-
-# Abstract
-
-A ce jour, la pandémie du coronavirus a touché près de 100000 personnes au Québec et entrainé plus de 6000 décès.
-Cela fait de la province l'une des plus touchées au Canada et l'une des régions les plus affectées au monde.
-Bien que la majorité de ces décès ait eu lieu dans les maisons pour personnes âgées, et donc a touché plus cette tranche de la population, son influence globale sur les individus n'est pas à négliger.
-En effet, outre les conséquences directes de la pandémie, les mesures prises par les différents gouvernements à travers le monde ne sont pas sans conséquences sur les familles.
-Il est démontré que ses conséquences affectent négativement en priorité les couches les plus défavorisées de la population.
-Les immigrants se concentrent dans bien des cas dans cette couche de la population.
-Cette étude vise ainsi à analyser l'influence de la pandémie sur les immigrants et leurs enfants au Québec, à examiner les facteurs socio-culturels qui permettent d'atténuer son impact négatif.
-
-To date, the coronavirus pandemic has affected nearly 100,000 people in Quebec and caused more than 6,000 deaths.
-This makes the province one of the most affected in Canada and one of the most affected regions in the world.
-Although the majority of these deaths took place in homes for the elderly, and therefore affected this segment of the population more, its overall influence on individuals is not to be overlooked.
-In fact, in addition to the direct consequences of the pandemic, the measures taken by various governments around the world are not without consequences for families.
-It has been shown that its consequences primarily affect the most disadvantaged segments of the population negatively.
-Immigrants are in many cases concentrated in this segment of the population.
-This study thus aims to analyze the influence of the pandemic on immigrants and their children in Quebec, to examine the socio-cultural factors that make it possible to mitigate its negative impact.
-
-À soumettre au Canadian Population Studies
-
-Autre projet possible: intégration des immigrants et covid19 au Québec
-
-**Keywords** :
-
--->
-
-\newpage
 
@@
 
-The COVID-19 pandemic has highlighted the vast ethnic, social, economic, and health disparities in high-income countries, including Canada.
-One of the most vulnerable segments of these affluent societies is the immigrant population, that has become widely diversified with more and more people coming from developing countries, and for whom integration into their new society may be more difficult.
-The coronavirus pandemic brings a new challenge for these immigrant populations, directly or indirectly through the tensions between the ethnic groups that it helps to generate [@coates2020Covid19].
-These immigrants face language barriers, low social capital, housing insecurity, and severe poverty, rendering them particularly susceptible to the effects of the current pandemic.
-Immigrants are also more likely to be employed as essential workers where social distancing, sanitary precautions, and self-isolation, measures put in place to mitigate the pandemy, are challenging [@reid2021Migrant; @yi2021Health; @zhang2021COVID19].
-More importantly, migrants may encounter various forms of discrimination, leading to social exclusion and restricted access to healthcare and social services [@spiritus-beerden2021Mental].
-It is therefore necessary to understand how this pandemic affects these immigrant populations differently and the native populations in Quebec.
-In doing so, this article will reveal a better understanding of how the current pandemic aggravates existing inequalities and challenges in the province for immigrants and their descendants.
+The COVID-19 pandemic has highlighted the vast ethnic, social, economic, and
+health disparities in high-income countries, including Canada. One of the most
+vulnerable segments of these affluent societies is the immigrant population,
+that has become widely diversified with more and more people coming from
+developing countries, and for whom integration into their new society may be
+more difficult. The coronavirus pandemic brings a new challenge for these
+immigrant populations, directly or indirectly through the tensions between the
+ethnic groups that it helps to generate [@coates2020Covid19]. These immigrants
+face language barriers, low social capital, housing insecurity, and severe
+poverty, rendering them particularly susceptible to the effects of the current
+pandemic. Immigrants are also more likely to be employed as essential workers
+where social distancing, sanitary precautions, and self-isolation, measures put
+in place to mitigate the pandemy, are challenging [@reid2021Migrant;
+@yi2021Health; @zhang2021COVID19]. More importantly, migrants may encounter
+various forms of discrimination, leading to social exclusion and restricted
+access to healthcare and social services [@spiritus-beerden2021Mental]. It is
+therefore necessary to understand how this pandemic affects these immigrant
+populations differently and the native populations in Quebec. In doing so, this
+article will reveal a better understanding of how the current pandemic
+aggravates existing inequalities and challenges in the province for immigrants
+and their descendants.
 
@@
 <!--As of 2019, the increase in net new residents (60,700 versus 42,200 in 2018) and lower interprovincial migration losses (-1,500 versus -6,100 in 2018) contributed to Quebec's population growth.
--->Since the start of the pandemic, most of the confirmed cases were identified in Montreal and areas with high immigrant populations [@miconi2021Ethnocultural; @shields2020COVID19].
-The coronavirus pandemic has infected nearly 450,000 people in Quebec and has resulted in more than 11,000 deaths (https://www.quebec.ca/sante/problemes-de-sante/a-z/coronavirus-2019/situation-coronavirus-quebec, consulted on octobre 29, 2021).
-This makes it the worst affected province in Canada and one of the worst affected regions in the world during the first wave of the pandemic.
-Although the majority of these deaths took place in elderly's homes, and therefore affected more this segment of the population, its overall influence on individuals is not to be overlooked.
-It has been shown that, on the one hand, immigrants and their families have a higher risk of being exposed to the pathogen, and, at the same time, they are more vulnerable to racism and discrimination and face barriers to accessing social support [@2021CRARR] and information, especially  for those who didn't speak or understand the two official langages.
-<!--In addition to the direct consequences of the pandemic, the measures taken by various governments around the world are also not without consequences on families.
--->Several studies have thus shown excess mortality in ethnocultural minority populations [@bertocchi2020Covid19], higher unemployment, greater loss of income and negative consequences for children and families in several countries.
-This differentiated effect can be explained in different ways.
-On the one hand, in terms of the direct consequences of the pandemic, the situation of precariousness and greater poverty among immigrants exposes them more to the more harmful consequences of the disease on their state of health.
-Likewise, immigrants often work in jobs where social distancing is difficult to enforce.
-On the other hand, indirectly, the solutions to the crisis affect immigrants and natives differently.
-The complete shutdown of tye economy with its set of adjustments has greatly impeded the employability of immigrants, especially newcomers who have not yet grasped the workings of society and who cannot benefit from state subsidies.
-But, it is also shown that these effects are not uniform on the whole of the immigrant population, itself, very diverse.
+-->
 
-Despite the growing focus on the differential impacts of the COVID-19 on immigrants and non-immigrants segment of the population, there is still very little knowledge about how the pandemic and its related public policy impact the lives of immigrants and their families.
-The existing health datasets (largely used to assess this effect) are not necessarily representative of the immigrant population and do not offer an accurate picture of the complex social status and health outcomes among this group [@choi2021Studying; @gagnon2021Immigration; @khan2015Visible].
-In most cases, these datasets lack data on the residence length of immigrants and combine newcomers with long-term settlers [@edge2013Discrimination; @hayward2021Clinical].
-As a result, they may not allow a detailed assessment of the nature and extent of risk exposure in sub-segments of populations, as well as whether interventions are reaching those most at risk.
-In addition, the family-level impacts of the pandemic are still poorly understood.
-While a body of research has explored the association between the pandemic and its potential effects on intimate partner violence and marital dissolution [@bhatnagar2021Navigating; @koshan2020COVID19; @morrison2021Chapter; @prime2020Risk], few studies have examined negative familial repercussions in the context of immigration status.
-There is ample evidence that sustained family relationships greatly enhance immigrants' lives on several fronts, including fostering economic prosperity [@marger2001Sociala], psychological well-being [@beiser2002Povertya], social network development [@evansluong2017Familya], and educational success for children [@munroe-blum1989Immigranta].
-The family is, therefore, a key component of long-term social mobility and economic success of immigrants.
+Since the start of the pandemic, most of the confirmed cases were identified in
+Montreal and areas with high immigrant populations [@miconi2021Ethnocultural;
+@shields2020COVID19]. The coronavirus pandemic has infected nearly 450,000
+people in Quebec and has resulted in more than 11,000 deaths
+(<https://www.quebec.ca/sante/problemes-de-sante/a-z/coronavirus-2019/situation-coronavirus-quebec>,
+consulted on octobre 29, 2021). This makes it the worst affected province in
+Canada and one of the worst affected regions in the world during the first wave
+of the pandemic. Although the majority of these deaths took place in elderly's
+homes, and therefore affected more this segment of the population, its overall
+influence on individuals is not to be overlooked. It has been shown that, on the
+one hand, immigrants and their families have a higher risk of being exposed to
+the pathogen, and, at the same time, they are more vulnerable to racism and
+discrimination and face barriers to accessing social support [@2021CRARR] and
+information, especially for those who didn't speak or understand the two
+official langages.
+<!--In addition to the direct consequences of the pandemic, the measures taken by various governments around the world are also not without consequences on families. -->Several
+studies have thus shown excess mortality in ethnocultural minority populations
+[@bertocchi2020Covid19], higher unemployment, greater loss of income and
+negative consequences for children and families in several countries. This
+differentiated effect can be explained in different ways. On the one hand, in
+terms of the direct consequences of the pandemic, the situation of
+precariousness and greater poverty among immigrants exposes them more to the
+more harmful consequences of the disease on their state of health. Likewise,
+immigrants often work in jobs where social distancing is difficult to enforce.
+On the other hand, indirectly, the solutions to the crisis affect immigrants and
+natives differently. The complete shutdown of tye economy with its set of
+adjustments has greatly impeded the employability of immigrants, especially
+newcomers who have not yet grasped the workings of society and who cannot
+benefit from state subsidies. But, it is also shown that these effects are not
+uniform on the whole of the immigrant population, itself, very diverse.
+
+Despite the growing focus on the differential impacts of the COVID-19 on
+immigrants and non-immigrants segment of the population, there is still very
+little knowledge about how the pandemic and its related public policy impact the
+lives of immigrants and their families. The existing health datasets (largely
+used to assess this effect) are not necessarily representative of the immigrant
+population and do not offer an accurate picture of the complex social status and
+health outcomes among this group [@choi2021Studying; @gagnon2021Immigration;
+@khan2015Visible]. In most cases, these datasets lack data on the residence
+length of immigrants and combine newcomers with long-term settlers
+[@edge2013Discrimination; @hayward2021Clinical]. As a result, they may not allow
+a detailed assessment of the nature and extent of risk exposure in sub-segments
+of populations, as well as whether interventions are reaching those most at
+risk. In addition, the family-level impacts of the pandemic are still poorly
+understood. While a body of research has explored the association between the
+pandemic and its potential effects on intimate partner violence and marital
+dissolution [@bhatnagar2021Navigating; @koshan2020COVID19; @morrison2021Chapter;
+@prime2020Risk], few studies have examined negative familial repercussions in
+the context of immigration status. There is ample evidence that sustained family
+relationships greatly enhance immigrants' lives on several fronts, including
+fostering economic prosperity [@marger2001Sociala], psychological well-being
+[@beiser2002Povertya], social network development [@evansluong2017Familya], and
+educational success for children [@munroe-blum1989Immigranta]. The family is,
+therefore, a key component of long-term social mobility and economic success of
+immigrants.
 
@@
 How these measures make a difference in the face of the pandemic remains to be seen.
--->Based on an original biographic survey collected around the first wave of the pandemic in august-november 2020 on immigrants and the native population of Québec, this study mainly aims to understand the family exposure, consequences, and responses of the pandemic on immigrant and native Quebec populations.
-We assess a whole range of different measures of each outcome.
-We divided the consequences in three dimensions, 1) access to ressources, 2) ability to perform duties and 3) employment and income.
-We seek also to analyse in what ways these outcomes differ according to the immigrants length of residence.
-By doing so, this study provide a clear and global picture of the situation of different population groups in Quebec in the face of the COVID-19.
+-->
+
+Based on an original biographic survey collected around the first wave of the
+pandemic in august-november 2020 on immigrants and the native population of
+Québec, this study mainly aims to understand the family exposure, consequences,
+and responses of the pandemic on immigrant and native Quebec populations. We
+assess a whole range of different measures of each outcome. We divided the
+consequences in three dimensions, 1) access to ressources, 2) ability to perform
+duties and 3) employment and income. We seek also to analyse in what ways these
+outcomes differ according to the immigrants length of residence. By doing so,
+this study provide a clear and global picture of the situation of different
+population groups in Quebec in the face of the COVID-19.
 <!-- To respond to these questions, we will use data collected around the first wave of the pandemic in august-november 2020.
@@
 
-# Current body of literature and knowledge gaps
-
-The literature on the racial and ethnic health disparities during the COVID-19 crisis can be divided into two main categories: studies addressing the socio-economic determinants of health and those examining the delivery of health interventions and policies.
-The first category is primarily comprised of research that focuses on the link between socio-economic status (SES) of high-risk minority groups and the differential impact of the pandemic [@drefahl2020Populationbased; @greenaway2020COVID19; @lopez2021Racial; @pacheco2020Job; @sundaram2020Individual].
-For instance, an inter-country study has found that Middle Eastern, African, and Nordic immigrants in Sweden have a higher risk of being hospitalized and even dying from COVID-19 when compared to the Swedish born population.
-In the same study, the SES, proportion of working-age households, and neighborhood density were significantly associated with the mortality rate [@rostila2021Disparities].
-This trend of research demonstrates that racial and ethnic minorities are more likely to experience stress factors, not only due to the hazards of infection but also because of economic and social marginalization [@alahmad2020COVID19; @choi2020Buffering; @endale2020COVID19; @fitzpatrick2020Fear; @serafini2021Psychological].
-The intersectionality of race, sex, and health conditions has also been explored in the Canadian context of the pandemic.
-@ingen2021Neighbourhoodlevel demonstrated that Ontarian neighborhoods with immigrant and racialized populations have a disproportionate share of COVID-19 infections.
-In addition, areas with high SES report a lesser number of COVID cases than low SES ones [@ingen2021Neighbourhoodlevel].
-Another study conducted in Quebec confirms the strong association between immigrants' backgrounds, SES, and COVID-19 related psychological stress [@miconi2021Ethnocultural].
-The most recent reviews on the topic suggest that immigrants disproportionately live in crowded and inadequate housing and are employed in essential industries, leaving them with severe income insecurities and poor physiological/psychological well-being [@chew2020Clinical; @davis2020Impact].
-Studies on SES and health disparities have emphasized that this situation is further complicated for immigrant women, constrained by gender and afflicted by a significant power imbalance in the household and workplace [@alarcao2021Intersections; @esegbona-adeigbe2020COVID19; @jehn2021COVID19; @mo2020Differential; @nardon2021Skilled; @sabri2020Effect; @st-denis2020Sociodemographic].
-According to Lightman, immigrants and women of color are overrepresented in long-term care facilities in Canada, where COVID infection rates have been exceptionally high.
-They are also subject to exclusionary social and economic practices [@lightman2021Caring].
-
-The second category of the literature is primarily comprised of policy analyses, research on public health communications, examining the efficacy of COVID initiatives, the optimal allocation of government aids, and ethical issues associated with applying new health interventions to vulnerable populations and their communities [@aragona2020Negative; @cross2021Coronavirus; @desai2020COVID19; @doyle2020Migrant; @gilman2020Modelling; @hargreaves2020Targeting; @langellier2020Policy; @nezafatmaldonado2020Engaging; @silverman2020Ethics; @wilson2020Ethics].
-Some of the questions that this body of research addresses include: how best to respond to COVID-19 in immigrant communities?
-How do public health policies impact the socio-economic status and psychological well-being of different sub-populations?
-What is the structure of the planning and implementation of government aid programs during the current crisis?
-And, how can we design public health communication strategies in a way that is culturally sensitive and inclusive?
-
-There are several knowledge gaps in the existing literature.
-Most of the cited studies do not include household data and family structure of immigrants as variables that affect exposure risks and health outcomes of COVID-19.
-Consequently, the relationship between immigrant status and COVID-19 is not accounted for by household size, and family characteristics (i.e., income, education, religion).
-A key reason is that most of the health data used in such research projects are acquired from national registers or administrative databases, which provide little information on immigration status and kinship.
-However, as the study of Thomeer and her colleagues theorizes, the extent of adverse social and economic impacts of COVID-19 is significantly conditional on the composition of an immigrant's family [@thomeer2020How].
-Households that include a spouse/partner, children, or multigenerational family members most likely experience a different level of social, psychological, and medical vulnerability than single or nuclear ones.
-Therefore, studies of immigrant populations should account for the breakdown of family configuration to better understand the mechanisms linking social determinants and health outcomes.
-Another knowledge gap concerns the current state of Canadian studies of COVID-19 and its public health crisis.
-There is a dearth of literature on the challenges faced by immigrants and their families in the country.
-Although the burden of COVID-19 infection is disproportionately higher among these marginalized groups (St-Denis, 2020; Sundaram et al., 2020), research on the issues they are experiencing is sparse and often limited to one aspect.
-<!--Likewise, there is limited focus on the effectiveness of COVID-19 crisis management for immigrants at the policy level.
-In response to the mounting cases of COVID, the federal government and provinces and territories of Canada have launched multi-pronged public health initiatives to protect society from the epidemic (Migone, 2020).
-However, many open questions remain unanswered, including, do interventions efficiently reach the migrant groups?
-What are the likely effects of each program on different immigrant communities?
-How can an intervention be tailored to the needs of a particular migrant group and their families?
--->
-
-# Theoretical framework 
-Family as a unit of analysis can be particularly fruitful when studying health inequalities in the context of a societal-wide crisis.
-Through the lens of the COVID-19 pandemic, the following points illustrate why exploring family dynamics provides additional insight into the underlying mechanisms of health disparities.
-First, the family is the most fundamental unit of social organization.
-It facilitates access to resources and the transmission of social and psychological capital from one generation to the next.
-Therefore, a better analysis of how crises and adverse life events affect people requires a better understanding of how families relate to each other and interact with external forces.
-Second, during public health emergencies, families may be thrown into different structures due to unprecedented financial, mental, and physical challenges.
-The risk of family dissolution, for example, can be heightened when members must spend unusually long periods together at home and have limited opportunities to engage with friends and the wider community.
-Conversely, families can develop closer bonds through self-sacrifice, shared responsibilities, and increased care and support.
-As the dependency status of seniors increases, the elderlies are also more likely to be integrated into formerly nucleus families.
-We posit that the extent to which a family can adapt its functions to meet new challenges is a predictor of the well-being of its members and the resilience of its broader population.
-Third, the family is a crucial link to social institutions, such as the health care system, schools, and government agencies whose decisions have significant downstream impacts.
-Families are viewed by many public-level entities as a reference when allocating resources, enforcing new standards, and responding to emerging needs and demands.
-From this perspective, the family operates at the nexus of individual and societal structures, embedding the tripple effects of public policy in its most consequential ways.
-Finally, the family has a crucial role to play in determining the life course of its members.
-As COVID-19 elevates the importance of familial bonds, the post-pandemic period might be marked by an increased influence of families on the lives of their members and the social structures they inhabit.
-
-The above conceptualization echoes several other approaches to studying long-term health inequalities.
-For instance, life course theory emphasizes the significance of family ties and social environment as they structure prospects for individuals across their life span (Elder, 1977; Elder Jr., 1998; Macmillan, 2005; Settersten et al., 2020).
-Likewise, social capital theory and the stress-buffering hypothesis consider family systems a crucial factor in health and well-being (Choi and Jun, 2020; Cohen and McKay, 1984; Cohen and Wills, 1985; Harpham et al., 2004; Kawachi and Berkman, 2001; Lavee et al., 1987; Mandelbaum et al., 2020).
-They show that the family can offer tangible (e.g., food, shelter, and money) and intangible resources (e.g., support, social status, psychological comfort, and institutional advantages), protecting its members from adverse life events.
-Similarly, the social-ecological model and the theory of fundamental causes highlight how the SES, familial and extrafamilial factors pattern people's life trajectories and health outcomes (Barkan and Rocque, 2018; Clouston and Link, 2021; Figge et al., 2018; Kim, 2020; Libman et al., 2012; Ungar, 2015).
-However, all these approaches are ostensibly focused on life periods where the general social context is relatively stable.
-In this article, we present our conceptualization of the family in times of unprecedented social upheaval.
-Specifically, our framework seeks to characterize family dynamics as a suitable lens to analyze health inequalities in the face of a virulent public health crisis.
-
-# Methodology
-
-## Data
-
-The data for this study come from an original collection of biographical data on immigrants and the native population of Quebec to analyse their participation in Quebec society in various fields (social, economic, cultural, etc.).
-The immigrant population is limited to recent immigration that has taken place over the past 15 years.
-This survey has several components that retrace the life course of individuals over a long period of time.
-The pandemic component uses a reduced version of the [harmonized questionnaire on exposure and the impact of Covid19 on the family (CEFIS)](https://www.phenxtoolkit.org/toolkit_content/PDF/CPTS_CEFIS_FullSurveyEng.pdf).
-The questions are divided into three parts, on exposure to the disease, its consequences and the benefit of the policies put in place by both the federal and provincial governments.
-The other parts of the questionnaire focus on the characteristics of the respondents, their immigration history, their educational and family background.
-In short, it is a detailed questionnaire that covers the life of immigrants from birth and especially over the past 15 years.
-The data collection lasted three months, from August to November 2020. In total, 1,500 people were surveyed, including around 1,000 immigrants and 500 natives throughout Quebec.
-
-### Dependent variables
-
-We distinguish three groups of dependent variables:
-
-**1.	Disease exposure variables**: these are measured by six questions that seek to document whether the individual through work or a family member may have been exposed to the disease.
-For example, it is asked if a family member has had symptoms of the disease, or if a member has been hospitalized due to the contraction of the disease.
-
-**2.	Variables on the family consequences of the disease**: we measure three dimensions of consequences.
-The first group of measures assess the consequences on accessing ressources by the respondent or any member of his family.
-The second group of measure assess whether the respondent or any family member have difficulties to perform usual duties toward any member of the family.
-Finally, the last group of measures measure the consequences of the pandemic on employment and income of any family member.
-
-**3.	Variables on government assistance**: Finally, government support is measured by the following two questions that assessed whether the respondent or any family member has benefited of the temporary assistance for workers from the Government of Quebec (PATT) or of the Canadian Federal Government Emergency Benefit (CEP).
-The table XXX in the apendix presents the full set of indicators of the dependents variables
-
-### Main independent variable
-The main independent variable is immigrant status coupled with visible minority status.
-This variable takes the following modalities: 1 if the respondent is a non-racialized native, 2 if he is a non-racialized immigrant, 3 if he is a racialized immigrant, and 4 if the respondent is a racialized native or an immigrant who has not answered the visible minority question.
-The other variables in the analysis are: age, sex, length of residence, minority status, size of social network.
-
-## Analysis methods
-
-We use two main methods of analysis, a descriptive method and an explanatory method.
-In the descriptive method, we consider each of the independent variables in isolation and describe how they vary depending on the main independent variable.
-The explanatory methods goes further by assessing whether the effect remain when controlling for other covariates.
-Thus, by controlling for the other explanatory variables, we determine the effect of immigrant status on exposure to the pandemic, on its consequences and on the fact of benefiting or not of government assistance.
-In a second series of modelling, we assess whether the effect that we observed varies by the immigrant's length of residence or by marital status.
-We use either a logistic regression on the outcome variables that are binary and an ordonned logistic regression for the outcome that are measure with a 5-point Lickert scale.
-<!-- PUT THIS AS A LIMITATION IN THE CONCLUSION
-To be able to benefit from government benefits, one must meet a number of conditions, including the loss of job due to Covid-19.
-As a result, the beneficiaries are not the general population but those who have lost their jobs and this may itself be due to immigrant status.
-Thus, for model 3, we will use Heckman's models by first estimating the fact of having lost or not a job and then, the fact of receiving or not the government benefit.
--->
-
-# 3. Results
-
-## 3.1. Sample distribution
-
-## 3.2. Exposure by immigration status
-
-Figure 1 presents the exposure to the pandemy accroding to the immigrants and minority visible status.
-For all the six measures of exposure, it appears that the majority group is the most exposed to the COVID-19 in all instance.
-The big difference between the groups is about the exposure trhough working outside.
-Around 50% of the member of this group or a family member work outside the home at the time of the survey.
-This percentage is only 30% for the racialised immigrants and 25% for the non-racialised immigrants.
-
-![Figure 1.1: Exposure](Figures/Exposition1.png){width=60%}
-
-![Figure 1.2: Explicative model](Figures/Exposure.jpeg)
-
-## 3.3. Consequences by immigrations status
-
-The Figure 2 presents the consequences of the COVID-19 for a family member to access to ressources.
-It shows that in all accounts but one (getting medecine), the racialised immigrants express more diffciult to assess ressources.
-Around 20% of them stated having difficulty to get health care compare, around 10% for obtaining food and 7% for obtaining family insurrance.
-Comparatively, these proportion are respectively 13%, 7% and 4% for people and thier family member in the majority group and 19%, 3% and 4% respectively among the racialised immigrants.
-
-![Figure 3.1: Consequences on accessing ressources](Figures/Famille1.png){width=80%}
-
-![Figure 3.2: Explicative model](Figures/Ressources.jpeg)
-
-## 3.4. Consequences on performing usual duties
-
-The Figure 4 compares the response from the three groups on performing usual duties.
-It appears that for all groups the situation caused by the COVID-19 has inexpectedly improved the situation in various way, irrespective of the immigrants status of the respondents.
-
-![Figure 4.1: Consequences on performing usual duties](Figures/Entente1.png)
-
-![Figure 4.2: Explicative model](Figures/Abilities.jpeg)
-
-## 3.5. Consequences on employment by immigration status
-
-The consequences on employment are mixed, with the severe consequences more prevalent among the racialised immigrants.
-The Figure 3 shows for instance that close to 45% of the racialised immigrants have stated  that their family income has decreased compare to just 33% of the majority group and 37% of the non-racialed immigrants.
-In addition, 11% of the racialised immigrants reveals having loss their job compared to 9% among the non-racialised immigrants and 2% among members of the majority group.
-At the opposite, the members of the majority groups are more likely to report a reduction of the working hours or for stopping momentarily their work.
-
-![Figure 5.1: Consequences on employment and income](Figures/Emploi1.png)
-
-![Figure 5.2: Explicative model](Figures/Employment.jpeg)
-
-## 3.7 Help received by immigrations status
-
-Finally, we compared the different group on benefitting of the government assistance.
-It appears that racialised immigrants with 35% of the case have expressed having received the Canada Emergency Response Benefit wheras this proportion is 31% for the non-racialised immigrants and 25% for the member of the majority group.
-For the province Temporary Aid for Workers Program, around 5% of the respondents have received it.
-
-![](Figures/Aide1.png)
-
-![Figure 3.2: Explicative model](Figures/Government.jpeg)
-
-# Conclusion
-
-The preliminary results of this study show that the consequences of the pandemic on the Quebec population differ according to the population groups, but also according to the type of consequences.
-
 \newpage
-
-# Références
-
-http://www.oecd.org/coronavirus/policy-responses/what-is-the-impact-of-the-covid-19-pandemic-on-immigrants-and-their-children-e7cbb7de/
-
-Guttmann A, F.
-et al.
-(2020), COVID-19 in Immigrants, Refugees and Other Newcomers in Ontario: Characteristics of Those Tested and Those Confirmed Positive, as of June 13, 2020, ICES, http://www.ices.on.ca.
-
-Hu, Y.
-(2020), “Intersecting ethnic and native–migrant inequalities in the economic impact of the COVID-19 pandemic in the UK”, Research in Social Stratification and Mobility, Vol.
-68, p.
-100528, http://dx.doi.org/10.1016/j.rssm.2020.100528.
-
-Mcginnity, F.
-and G.
-Kingston (2017), “An Irish Welcome?
-Changing Irish Attitudes to Immigrants and Immigration: The Role of Recession and Immigration”, The Economic and social review, Vol.
-48/3, pp.
-281-304, https://www.researchgate.net/publication/319670153_An_Irish_Welcome_Changing_Irish_Attitudes_to_Immigrants_and_Immigration_The_Role_of_Recession_and_Immigration (accessed on 2 October 2020).
-
-Raisi-Estabragh, Z.
-et al.
-(2020), “Greater risk of severe COVID-19 in Black, Asian and Minority Ethnic populations is not explained by cardiometabolic, socioeconomic or behavioural factors, or by 25(OH)-vitamin D status: study of 1326 cases from the UK Biobank”, Journal of Public Health, Vol.
-42/3, pp.
-451-460, http://dx.doi.org/10.1093/pubmed/fdaa095.
-
-Apea, V.
-et al.
-(2020), Ethnicity and outcomes in patients hospitalised with COVID-19 infection in East London: an observational cohort study, Cold Spring Harbor Laboratory, http://dx.doi.org/10.1101/2020.06.10.20127621.
-
-Bertocchi, G. and A. Dimico (2020), “Covid-19, Race, and Redlining”, IZA Discussion Paper, Vol. 13467, https://papers.ssrn.com/sol3/papers.cfm?abstract_id=3648807 (accessed on 22 September 2020).
-
-Brun, S. and P. Simon (2020), “Inégalités ethno-raciales et coronavirus”, De Facto 19.

### 2023-05-02 Expérimentations (Théa)/2. Expérimentations (Bookdown)/Bookdown/02-body-of-litterature.Rmd
Multiple changes on Bookdown

--- 2023-05-01 Article/Trajipaq_Covid_Draft4.Rmd
@@
 ---
-title: 'COVID-19 Disparities in Quebec (Canada): A Comparative Analysis of Exposure
-  Risks, Consequences, and Access to Resources among Immigrants and Natives'
-author:
-- Vissého Adjiwanou ^[corresponding author - adjiwanou.vissého@uqam.ca]
-- Département de Sociologie, Université du Québec à Montréal (UQAM)
-- Georges NGalé Tchango (Département de démographie, Université de Montréal)
-- Solène Lardoux (Département de démographie, Université de Montréal)
-- Nima Zahedinameghi (Département de Sociologie, UQAM)
-date: "`r format(Sys.time(), '%d %B %Y')`"
-output:
-  word_document:
-    keep_md: yes
-  html_document:
-    toc: yes
-    toc_depth: 3
-  template: svm-latex-ms2.tex
-  pdf_document: default
-  citation_package: natbib
-bibliography:
-- covid_Nima.bib
-- Covid_Canada.bib
-- Covid_Family.bib
+knit: bookdown::render_book
+editor_options: 
+  markdown: 
+    wrap: 80
 ---
-<!--
-# Journal: family issues - SAGE
-https://journals-sagepub-com.proxy.bibliotheques.uqam.ca/author-instructions/JFI
-
-https://fishandwhistle.net/post/2020/getting-started-zotero-better-bibtex-rmarkdown/
-https://www.zotero.org/styles?fields=sociology
--->
-
-\newpage
-
-# Abstract
-
-The immigrant population is one of the most vulnerable in Canada, with a relatively high risk of unemployment, a low social support system, and linguistic barriers.
-However, their experiences with public health interventions during the COVID-19 pandemic have not been adequately explored.
-The purpose of this article is to analyze Covid-19 exposure risk and its family implications for both natives and immigrants in the province of Quebec.
-Additionally, we analyze immigrants' access to government aid and compare it to their native-born counterparts.
-The study consists of survey data collected from 1,000 immigrants and 500 non-immigrants Quebecers between August and November 2020.
-Our descriptive and exploratory analysis reveals significant differences between the two groups regarding exposure risk, family consequences, and access to resources.
-We conclude that the current pandemic exacerbates health disparities in the province, with particularly dire repercussions for immigrants’ families.
-We also discuss several socio-cultural strategies to mitigate inequality and recommend policies to address the needs of vulnerable populations.
-
-**Keywords**: Immigrants, exposure risk, government aid, health disparities, COVID-19, family disruptions
-
-<!--
-
-Using  script in Rmarkdown
-
-http://zevross.com/blog/2014/07/09/making-use-of-external-r-code-in-knitr-and-r-markdown/
-
-Managing my Zotero
-
-https://forums.zotero.org/discussion/35712/removing-pdfs-to-free-up-space
-
-https://www.zotero.org/support/sync
-
-https://www.zotero.org/support/preferences/sync#file_syncing
-
-https://www.zotero.org/support/sync#webdav
-
-https://www.zotero.org/support/preferences/advanced#linked_attachment_base_directory
-
-http://zotfile.com/
-
-# Abstract
-
-A ce jour, la pandémie du coronavirus a touché près de 100000 personnes au Québec et entrainé plus de 6000 décès.
-Cela fait de la province l'une des plus touchées au Canada et l'une des régions les plus affectées au monde.
-Bien que la majorité de ces décès ait eu lieu dans les maisons pour personnes âgées, et donc a touché plus cette tranche de la population, son influence globale sur les individus n'est pas à négliger.
-En effet, outre les conséquences directes de la pandémie, les mesures prises par les différents gouvernements à travers le monde ne sont pas sans conséquences sur les familles.
-Il est démontré que ses conséquences affectent négativement en priorité les couches les plus défavorisées de la population.
-Les immigrants se concentrent dans bien des cas dans cette couche de la population.
-Cette étude vise ainsi à analyser l'influence de la pandémie sur les immigrants et leurs enfants au Québec, à examiner les facteurs socio-culturels qui permettent d'atténuer son impact négatif.
-
-To date, the coronavirus pandemic has affected nearly 100,000 people in Quebec and caused more than 6,000 deaths.
-This makes the province one of the most affected in Canada and one of the most affected regions in the world.
-Although the majority of these deaths took place in homes for the elderly, and therefore affected this segment of the population more, its overall influence on individuals is not to be overlooked.
-In fact, in addition to the direct consequences of the pandemic, the measures taken by various governments around the world are not without consequences for families.
-It has been shown that its consequences primarily affect the most disadvantaged segments of the population negatively.
-Immigrants are in many cases concentrated in this segment of the population.
-This study thus aims to analyze the influence of the pandemic on immigrants and their children in Quebec, to examine the socio-cultural factors that make it possible to mitigate its negative impact.
-
-À soumettre au Canadian Population Studies
-
-Autre projet possible: intégration des immigrants et covid19 au Québec
-
-**Keywords** :
-
--->
-
-\newpage
-
-# Introduction / background
-
-The COVID-19 pandemic has highlighted the vast ethnic, social, economic, and health disparities in high-income countries, including Canada.
-One of the most vulnerable segments of these affluent societies is the immigrant population, that has become widely diversified with more and more people coming from developing countries, and for whom integration into their new society may be more difficult.
-The coronavirus pandemic brings a new challenge for these immigrant populations, directly or indirectly through the tensions between the ethnic groups that it helps to generate [@coates2020Covid19].
-These immigrants face language barriers, low social capital, housing insecurity, and severe poverty, rendering them particularly susceptible to the effects of the current pandemic.
-Immigrants are also more likely to be employed as essential workers where social distancing, sanitary precautions, and self-isolation, measures put in place to mitigate the pandemy, are challenging [@reid2021Migrant; @yi2021Health; @zhang2021COVID19].
-More importantly, migrants may encounter various forms of discrimination, leading to social exclusion and restricted access to healthcare and social services [@spiritus-beerden2021Mental].
-It is therefore necessary to understand how this pandemic affects these immigrant populations differently and the native populations in Quebec.
-In doing so, this article will reveal a better understanding of how the current pandemic aggravates existing inequalities and challenges in the province for immigrants and their descendants.
-
-<!-- The province of Quebec has one of the highest concentrations of immigrants in Canada, including permanent residents, temporary residents, and refugees [@governmentofcanada2017Daily; @immigration2005Recent].
-In 2016, 1.2 million immigrants were living in Quebec, an increase from 8% in 1990 to 13.7% of the population in 2016 (Institut de la statistique du Québec, 2020).
-<!--As of 2019, the increase in net new residents (60,700 versus 42,200 in 2018) and lower interprovincial migration losses (-1,500 versus -6,100 in 2018) contributed to Quebec's population growth.
--->Since the start of the pandemic, most of the confirmed cases were identified in Montreal and areas with high immigrant populations [@miconi2021Ethnocultural; @shields2020COVID19].
-The coronavirus pandemic has infected nearly 450,000 people in Quebec and has resulted in more than 11,000 deaths (https://www.quebec.ca/sante/problemes-de-sante/a-z/coronavirus-2019/situation-coronavirus-quebec, consulted on octobre 29, 2021).
-This makes it the worst affected province in Canada and one of the worst affected regions in the world during the first wave of the pandemic.
-Although the majority of these deaths took place in elderly's homes, and therefore affected more this segment of the population, its overall influence on individuals is not to be overlooked.
-It has been shown that, on the one hand, immigrants and their families have a higher risk of being exposed to the pathogen, and, at the same time, they are more vulnerable to racism and discrimination and face barriers to accessing social support [@2021CRARR] and information, especially  for those who didn't speak or understand the two official langages.
-<!--In addition to the direct consequences of the pandemic, the measures taken by various governments around the world are also not without consequences on families.
--->Several studies have thus shown excess mortality in ethnocultural minority populations [@bertocchi2020Covid19], higher unemployment, greater loss of income and negative consequences for children and families in several countries.
-This differentiated effect can be explained in different ways.
-On the one hand, in terms of the direct consequences of the pandemic, the situation of precariousness and greater poverty among immigrants exposes them more to the more harmful consequences of the disease on their state of health.
-Likewise, immigrants often work in jobs where social distancing is difficult to enforce.
-On the other hand, indirectly, the solutions to the crisis affect immigrants and natives differently.
-The complete shutdown of tye economy with its set of adjustments has greatly impeded the employability of immigrants, especially newcomers who have not yet grasped the workings of society and who cannot benefit from state subsidies.
-But, it is also shown that these effects are not uniform on the whole of the immigrant population, itself, very diverse.
-
-Despite the growing focus on the differential impacts of the COVID-19 on immigrants and non-immigrants segment of the population, there is still very little knowledge about how the pandemic and its related public policy impact the lives of immigrants and their families.
-The existing health datasets (largely used to assess this effect) are not necessarily representative of the immigrant population and do not offer an accurate picture of the complex social status and health outcomes among this group [@choi2021Studying; @gagnon2021Immigration; @khan2015Visible].
-In most cases, these datasets lack data on the residence length of immigrants and combine newcomers with long-term settlers [@edge2013Discrimination; @hayward2021Clinical].
-As a result, they may not allow a detailed assessment of the nature and extent of risk exposure in sub-segments of populations, as well as whether interventions are reaching those most at risk.
-In addition, the family-level impacts of the pandemic are still poorly understood.
-While a body of research has explored the association between the pandemic and its potential effects on intimate partner violence and marital dissolution [@bhatnagar2021Navigating; @koshan2020COVID19; @morrison2021Chapter; @prime2020Risk], few studies have examined negative familial repercussions in the context of immigration status.
-There is ample evidence that sustained family relationships greatly enhance immigrants' lives on several fronts, including fostering economic prosperity [@marger2001Sociala], psychological well-being [@beiser2002Povertya], social network development [@evansluong2017Familya], and educational success for children [@munroe-blum1989Immigranta].
-The family is, therefore, a key component of long-term social mobility and economic success of immigrants.
-
-<!--Faced with the pandemic, the governments of Canada and Quebec have implemented a number of measures to reduce its impact.
-These policies can be broken down at two levels, direct support to families, including the Canadian Emergency Benefit, which can provide employees and self-employed workers affected by Covid-19 with a financial benefit of up to $ 2,000.
-On the other hand, the government supports many companies to keep their employees or supports them directly when the revenues from their activities have fallen considerably.
-While these measures allow families to cope better with the crisis, the fact remains that the changes brought about by this crisis can affect other dimensions of life.
-How these measures make a difference in the face of the pandemic remains to be seen.
--->Based on an original biographic survey collected around the first wave of the pandemic in august-november 2020 on immigrants and the native population of Québec, this study mainly aims to understand the family exposure, consequences, and responses of the pandemic on immigrant and native Quebec populations.
-We assess a whole range of different measures of each outcome.
-We divided the consequences in three dimensions, 1) access to ressources, 2) ability to perform duties and 3) employment and income.
-We seek also to analyse in what ways these outcomes differ according to the immigrants length of residence.
-By doing so, this study provide a clear and global picture of the situation of different population groups in Quebec in the face of the COVID-19.
-<!-- To respond to these questions, we will use data collected around the first wave of the pandemic in august-november 2020.
-Our study aims to fill the gap by analyzing group-specific differences in exposure risk, family consequences, and access to resources among immigrants and Quebecers.
-We also seek to shed light on several socio-cultural strategies to mitigate inequality and recommend relevant policy solutions.
-The detailed findings will help policymakers address health inequalities and improve equity in Quebec while developing adequate measures and programs.
--->
-
-<!-- Introduction  -->
-
-<!-- Like the rest of Canada, Quebec has experienced a substantial increase in its immigration in recent years.
-The immigrant population in the province thus increased from 8% in the 1990s to 13.7% of the population, or 1.2 million individuals in 2016 (Institut de la statistique du Québec, 2020).
-This immigration is also unique in that it has become widely diversified with more and more people coming from developing countries, and for whom integration into their new society may be more difficult.
-The coronavirus pandemic brings a new challenge for these immigrant populations, directly or indirectly through the tensions between the ethnic groups that it helps to generate [@coates2020Covid19].
-It is therefore necessary to understand how this pandemic affects these immigrant populations differently and the native populations in Quebec.
--->
-
-<!-- The coronavirus pandemic has affected nearly 100,000 people in Quebec and has resulted in more than 6,000 deaths.
-This makes it the worst affected province in Canada and one of the worst affected regions in the world.
-Although the majority of these deaths took place in homes for the elderly, and therefore affected this segment of the population more, its overall influence on individuals is not to be overlooked.
-In fact, in addition to the direct consequences of the pandemic, the measures taken by various governments around the world are also not without consequences on families.
-Several studies have thus shown excess mortality in ethnocultural minority populations [@bertocchi2020Covid19], higher unemployment, greater loss of income and negative consequences for children and families in several countries.
--->
-
-<!-- This differentiated effect can be explained in different ways.
-On the one hand, in terms of the direct consequences of the pandemic, the situation of precariousness and greater poverty among immigrants exposes them more to the more harmful consequences of the disease on their state of health.
-Likewise, immigrants often work in jobs where social distancing is difficult to enforce.
-On the other hand, indirectly, the solutions to the crisis affect immigrants and natives differently.
-The complete shutdown of the company with its set of adjustments has greatly impeded the employability of immigrants, especially newcomers who have not yet grasped the workings of society and who cannot benefit from state subsidies.
-But, it is also shown that these effects are not uniform on the whole of the immigrant population, itself, very diverse.
--->
-
-<!-- Faced with the pandemic, the governments of Canada and Quebec have implemented a number of measures to reduce the impact of the pandemic.
-These policies can be broken down at two levels, direct support to families, including the Canadian Emergency Benefit, which can provide employees and self-employed workers affected by Covid-19 with a financial benefit of up to $ 2,000.
-On the other hand, the government supports many companies to keep their employees or supports them directly when the revenues from their activities have fallen considerably.
-While these measures allow families to cope better with the crisis, the fact remains that the changes brought about by this crisis can affect other dimensions of life.
-How these measures make a difference in the face of the pandemic remains to be seen.
--->
-
-<!-- This study mainly aims to understand the family consequences of the pandemic on immigrant and native Quebec populations.
-Therefore, we will seek to analyse in what ways its consequences differ according to the region of origin of immigrants, length of residence, age and sex.
--->
 
@@
 
-The literature on the racial and ethnic health disparities during the COVID-19 crisis can be divided into two main categories: studies addressing the socio-economic determinants of health and those examining the delivery of health interventions and policies.
-The first category is primarily comprised of research that focuses on the link between socio-economic status (SES) of high-risk minority groups and the differential impact of the pandemic [@drefahl2020Populationbased; @greenaway2020COVID19; @lopez2021Racial; @pacheco2020Job; @sundaram2020Individual].
-For instance, an inter-country study has found that Middle Eastern, African, and Nordic immigrants in Sweden have a higher risk of being hospitalized and even dying from COVID-19 when compared to the Swedish born population.
-In the same study, the SES, proportion of working-age households, and neighborhood density were significantly associated with the mortality rate [@rostila2021Disparities].
-This trend of research demonstrates that racial and ethnic minorities are more likely to experience stress factors, not only due to the hazards of infection but also because of economic and social marginalization [@alahmad2020COVID19; @choi2020Buffering; @endale2020COVID19; @fitzpatrick2020Fear; @serafini2021Psychological].
-The intersectionality of race, sex, and health conditions has also been explored in the Canadian context of the pandemic.
-@ingen2021Neighbourhoodlevel demonstrated that Ontarian neighborhoods with immigrant and racialized populations have a disproportionate share of COVID-19 infections.
-In addition, areas with high SES report a lesser number of COVID cases than low SES ones [@ingen2021Neighbourhoodlevel].
-Another study conducted in Quebec confirms the strong association between immigrants' backgrounds, SES, and COVID-19 related psychological stress [@miconi2021Ethnocultural].
-The most recent reviews on the topic suggest that immigrants disproportionately live in crowded and inadequate housing and are employed in essential industries, leaving them with severe income insecurities and poor physiological/psychological well-being [@chew2020Clinical; @davis2020Impact].
-Studies on SES and health disparities have emphasized that this situation is further complicated for immigrant women, constrained by gender and afflicted by a significant power imbalance in the household and workplace [@alarcao2021Intersections; @esegbona-adeigbe2020COVID19; @jehn2021COVID19; @mo2020Differential; @nardon2021Skilled; @sabri2020Effect; @st-denis2020Sociodemographic].
-According to Lightman, immigrants and women of color are overrepresented in long-term care facilities in Canada, where COVID infection rates have been exceptionally high.
-They are also subject to exclusionary social and economic practices [@lightman2021Caring].
+The literature on the racial and ethnic health disparities during the COVID-19
+crisis can be divided into two main categories: studies addressing the
+socio-economic determinants of health and those examining the delivery of health
+interventions and policies. The first category is primarily comprised of
+research that focuses on the link between socio-economic status (SES) of
+high-risk minority groups and the differential impact of the pandemic
+[@drefahl2020Populationbased; @greenaway2020COVID19; @lopez2021Racial;
+@pacheco2020Job; @sundaram2020Individual]. For instance, an inter-country study
+has found that Middle Eastern, African, and Nordic immigrants in Sweden have a
+higher risk of being hospitalized and even dying from COVID-19 when compared to
+the Swedish born population. In the same study, the SES, proportion of
+working-age households, and neighborhood density were significantly associated
+with the mortality rate [@rostila2021Disparities]. This trend of research
+demonstrates that racial and ethnic minorities are more likely to experience
+stress factors, not only due to the hazards of infection but also because of
+economic and social marginalization [@alahmad2020COVID19; @choi2020Buffering;
+@endale2020COVID19; @fitzpatrick2020Fear; @serafini2021Psychological]. The
+intersectionality of race, sex, and health conditions has also been explored in
+the Canadian context of the pandemic. @ingen2021Neighbourhoodlevel demonstrated
+that Ontarian neighborhoods with immigrant and racialized populations have a
+disproportionate share of COVID-19 infections. In addition, areas with high SES
+report a lesser number of COVID cases than low SES ones
+[@ingen2021Neighbourhoodlevel]. Another study conducted in Quebec confirms the
+strong association between immigrants' backgrounds, SES, and COVID-19 related
+psychological stress [@miconi2021Ethnocultural]. The most recent reviews on the
+topic suggest that immigrants disproportionately live in crowded and inadequate
+housing and are employed in essential industries, leaving them with severe
+income insecurities and poor physiological/psychological well-being
+[@chew2020Clinical; @davis2020Impact]. Studies on SES and health disparities
+have emphasized that this situation is further complicated for immigrant women,
+constrained by gender and afflicted by a significant power imbalance in the
+household and workplace [@alarcao2021Intersections;
+@esegbona-adeigbe2020COVID19; @jehn2021COVID19; @mo2020Differential;
+@nardon2021Skilled; @sabri2020Effect; @st-denis2020Sociodemographic]. According
+to Lightman, immigrants and women of color are overrepresented in long-term care
+facilities in Canada, where COVID infection rates have been exceptionally high.
+They are also subject to exclusionary social and economic practices
+[@lightman2021Caring].
 
-The second category of the literature is primarily comprised of policy analyses, research on public health communications, examining the efficacy of COVID initiatives, the optimal allocation of government aids, and ethical issues associated with applying new health interventions to vulnerable populations and their communities [@aragona2020Negative; @cross2021Coronavirus; @desai2020COVID19; @doyle2020Migrant; @gilman2020Modelling; @hargreaves2020Targeting; @langellier2020Policy; @nezafatmaldonado2020Engaging; @silverman2020Ethics; @wilson2020Ethics].
-Some of the questions that this body of research addresses include: how best to respond to COVID-19 in immigrant communities?
-How do public health policies impact the socio-economic status and psychological well-being of different sub-populations?
-What is the structure of the planning and implementation of government aid programs during the current crisis?
-And, how can we design public health communication strategies in a way that is culturally sensitive and inclusive?
+The second category of the literature is primarily comprised of policy analyses,
+research on public health communications, examining the efficacy of COVID
+initiatives, the optimal allocation of government aids, and ethical issues
+associated with applying new health interventions to vulnerable populations and
+their communities [@aragona2020Negative; @cross2021Coronavirus;
+@desai2020COVID19; @doyle2020Migrant; @gilman2020Modelling;
+@hargreaves2020Targeting; @langellier2020Policy; @nezafatmaldonado2020Engaging;
+@silverman2020Ethics; @wilson2020Ethics]. Some of the questions that this body
+of research addresses include: how best to respond to COVID-19 in immigrant
+communities? How do public health policies impact the socio-economic status and
+psychological well-being of different sub-populations? What is the structure of
+the planning and implementation of government aid programs during the current
+crisis? And, how can we design public health communication strategies in a way
+that is culturally sensitive and inclusive?
 
-There are several knowledge gaps in the existing literature.
-Most of the cited studies do not include household data and family structure of immigrants as variables that affect exposure risks and health outcomes of COVID-19.
-Consequently, the relationship between immigrant status and COVID-19 is not accounted for by household size, and family characteristics (i.e., income, education, religion).
-A key reason is that most of the health data used in such research projects are acquired from national registers or administrative databases, which provide little information on immigration status and kinship.
-However, as the study of Thomeer and her colleagues theorizes, the extent of adverse social and economic impacts of COVID-19 is significantly conditional on the composition of an immigrant's family [@thomeer2020How].
-Households that include a spouse/partner, children, or multigenerational family members most likely experience a different level of social, psychological, and medical vulnerability than single or nuclear ones.
-Therefore, studies of immigrant populations should account for the breakdown of family configuration to better understand the mechanisms linking social determinants and health outcomes.
-Another knowledge gap concerns the current state of Canadian studies of COVID-19 and its public health crisis.
-There is a dearth of literature on the challenges faced by immigrants and their families in the country.
-Although the burden of COVID-19 infection is disproportionately higher among these marginalized groups (St-Denis, 2020; Sundaram et al., 2020), research on the issues they are experiencing is sparse and often limited to one aspect.
+There are several knowledge gaps in the existing literature. Most of the cited
+studies do not include household data and family structure of immigrants as
+variables that affect exposure risks and health outcomes of COVID-19.
+Consequently, the relationship between immigrant status and COVID-19 is not
+accounted for by household size, and family characteristics (i.e., income,
+education, religion). A key reason is that most of the health data used in such
+research projects are acquired from national registers or administrative
+databases, which provide little information on immigration status and kinship.
+However, as the study of Thomeer and her colleagues theorizes, the extent of
+adverse social and economic impacts of COVID-19 is significantly conditional on
+the composition of an immigrant's family [@thomeer2020How]. Households that
+include a spouse/partner, children, or multigenerational family members most
+likely experience a different level of social, psychological, and medical
+vulnerability than single or nuclear ones. Therefore, studies of immigrant
+populations should account for the breakdown of family configuration to better
+understand the mechanisms linking social determinants and health outcomes.
+Another knowledge gap concerns the current state of Canadian studies of COVID-19
+and its public health crisis. There is a dearth of literature on the challenges
+faced by immigrants and their families in the country. Although the burden of
+COVID-19 infection is disproportionately higher among these marginalized groups
+(St-Denis, 2020; Sundaram et al., 2020), research on the issues they are
+experiencing is sparse and often limited to one aspect.
 <!--Likewise, there is limited focus on the effectiveness of COVID-19 crisis management for immigrants at the policy level.
@@
 
-# Theoretical framework 
-Family as a unit of analysis can be particularly fruitful when studying health inequalities in the context of a societal-wide crisis.
-Through the lens of the COVID-19 pandemic, the following points illustrate why exploring family dynamics provides additional insight into the underlying mechanisms of health disparities.
-First, the family is the most fundamental unit of social organization.
-It facilitates access to resources and the transmission of social and psychological capital from one generation to the next.
-Therefore, a better analysis of how crises and adverse life events affect people requires a better understanding of how families relate to each other and interact with external forces.
-Second, during public health emergencies, families may be thrown into different structures due to unprecedented financial, mental, and physical challenges.
-The risk of family dissolution, for example, can be heightened when members must spend unusually long periods together at home and have limited opportunities to engage with friends and the wider community.
-Conversely, families can develop closer bonds through self-sacrifice, shared responsibilities, and increased care and support.
-As the dependency status of seniors increases, the elderlies are also more likely to be integrated into formerly nucleus families.
-We posit that the extent to which a family can adapt its functions to meet new challenges is a predictor of the well-being of its members and the resilience of its broader population.
-Third, the family is a crucial link to social institutions, such as the health care system, schools, and government agencies whose decisions have significant downstream impacts.
-Families are viewed by many public-level entities as a reference when allocating resources, enforcing new standards, and responding to emerging needs and demands.
-From this perspective, the family operates at the nexus of individual and societal structures, embedding the tripple effects of public policy in its most consequential ways.
-Finally, the family has a crucial role to play in determining the life course of its members.
-As COVID-19 elevates the importance of familial bonds, the post-pandemic period might be marked by an increased influence of families on the lives of their members and the social structures they inhabit.
-
-The above conceptualization echoes several other approaches to studying long-term health inequalities.
-For instance, life course theory emphasizes the significance of family ties and social environment as they structure prospects for individuals across their life span (Elder, 1977; Elder Jr., 1998; Macmillan, 2005; Settersten et al., 2020).
-Likewise, social capital theory and the stress-buffering hypothesis consider family systems a crucial factor in health and well-being (Choi and Jun, 2020; Cohen and McKay, 1984; Cohen and Wills, 1985; Harpham et al., 2004; Kawachi and Berkman, 2001; Lavee et al., 1987; Mandelbaum et al., 2020).
-They show that the family can offer tangible (e.g., food, shelter, and money) and intangible resources (e.g., support, social status, psychological comfort, and institutional advantages), protecting its members from adverse life events.
-Similarly, the social-ecological model and the theory of fundamental causes highlight how the SES, familial and extrafamilial factors pattern people's life trajectories and health outcomes (Barkan and Rocque, 2018; Clouston and Link, 2021; Figge et al., 2018; Kim, 2020; Libman et al., 2012; Ungar, 2015).
-However, all these approaches are ostensibly focused on life periods where the general social context is relatively stable.
-In this article, we present our conceptualization of the family in times of unprecedented social upheaval.
-Specifically, our framework seeks to characterize family dynamics as a suitable lens to analyze health inequalities in the face of a virulent public health crisis.
-
-# Methodology
-
-## Data
-
-The data for this study come from an original collection of biographical data on immigrants and the native population of Quebec to analyse their participation in Quebec society in various fields (social, economic, cultural, etc.).
-The immigrant population is limited to recent immigration that has taken place over the past 15 years.
-This survey has several components that retrace the life course of individuals over a long period of time.
-The pandemic component uses a reduced version of the [harmonized questionnaire on exposure and the impact of Covid19 on the family (CEFIS)](https://www.phenxtoolkit.org/toolkit_content/PDF/CPTS_CEFIS_FullSurveyEng.pdf).
-The questions are divided into three parts, on exposure to the disease, its consequences and the benefit of the policies put in place by both the federal and provincial governments.
-The other parts of the questionnaire focus on the characteristics of the respondents, their immigration history, their educational and family background.
-In short, it is a detailed questionnaire that covers the life of immigrants from birth and especially over the past 15 years.
-The data collection lasted three months, from August to November 2020. In total, 1,500 people were surveyed, including around 1,000 immigrants and 500 natives throughout Quebec.
-
-### Dependent variables
-
-We distinguish three groups of dependent variables:
-
-**1.	Disease exposure variables**: these are measured by six questions that seek to document whether the individual through work or a family member may have been exposed to the disease.
-For example, it is asked if a family member has had symptoms of the disease, or if a member has been hospitalized due to the contraction of the disease.
-
-**2.	Variables on the family consequences of the disease**: we measure three dimensions of consequences.
-The first group of measures assess the consequences on accessing ressources by the respondent or any member of his family.
-The second group of measure assess whether the respondent or any family member have difficulties to perform usual duties toward any member of the family.
-Finally, the last group of measures measure the consequences of the pandemic on employment and income of any family member.
-
-**3.	Variables on government assistance**: Finally, government support is measured by the following two questions that assessed whether the respondent or any family member has benefited of the temporary assistance for workers from the Government of Quebec (PATT) or of the Canadian Federal Government Emergency Benefit (CEP).
-The table XXX in the apendix presents the full set of indicators of the dependents variables
-
-### Main independent variable
-The main independent variable is immigrant status coupled with visible minority status.
-This variable takes the following modalities: 1 if the respondent is a non-racialized native, 2 if he is a non-racialized immigrant, 3 if he is a racialized immigrant, and 4 if the respondent is a racialized native or an immigrant who has not answered the visible minority question.
-The other variables in the analysis are: age, sex, length of residence, minority status, size of social network.
-
-## Analysis methods
-
-We use two main methods of analysis, a descriptive method and an explanatory method.
-In the descriptive method, we consider each of the independent variables in isolation and describe how they vary depending on the main independent variable.
-The explanatory methods goes further by assessing whether the effect remain when controlling for other covariates.
-Thus, by controlling for the other explanatory variables, we determine the effect of immigrant status on exposure to the pandemic, on its consequences and on the fact of benefiting or not of government assistance.
-In a second series of modelling, we assess whether the effect that we observed varies by the immigrant's length of residence or by marital status.
-We use either a logistic regression on the outcome variables that are binary and an ordonned logistic regression for the outcome that are measure with a 5-point Lickert scale.
-<!-- PUT THIS AS A LIMITATION IN THE CONCLUSION
-To be able to benefit from government benefits, one must meet a number of conditions, including the loss of job due to Covid-19.
-As a result, the beneficiaries are not the general population but those who have lost their jobs and this may itself be due to immigrant status.
-Thus, for model 3, we will use Heckman's models by first estimating the fact of having lost or not a job and then, the fact of receiving or not the government benefit.
--->
-
-# 3. Results
-
-## 3.1. Sample distribution
-
-## 3.2. Exposure by immigration status
-
-Figure 1 presents the exposure to the pandemy accroding to the immigrants and minority visible status.
-For all the six measures of exposure, it appears that the majority group is the most exposed to the COVID-19 in all instance.
-The big difference between the groups is about the exposure trhough working outside.
-Around 50% of the member of this group or a family member work outside the home at the time of the survey.
-This percentage is only 30% for the racialised immigrants and 25% for the non-racialised immigrants.
-
-![Figure 1.1: Exposure](Figures/Exposition1.png){width=60%}
-
-![Figure 1.2: Explicative model](Figures/Exposure.jpeg)
-
-## 3.3. Consequences by immigrations status
-
-The Figure 2 presents the consequences of the COVID-19 for a family member to access to ressources.
-It shows that in all accounts but one (getting medecine), the racialised immigrants express more diffciult to assess ressources.
-Around 20% of them stated having difficulty to get health care compare, around 10% for obtaining food and 7% for obtaining family insurrance.
-Comparatively, these proportion are respectively 13%, 7% and 4% for people and thier family member in the majority group and 19%, 3% and 4% respectively among the racialised immigrants.
-
-![Figure 3.1: Consequences on accessing ressources](Figures/Famille1.png){width=80%}
-
-![Figure 3.2: Explicative model](Figures/Ressources.jpeg)
-
-## 3.4. Consequences on performing usual duties
-
-The Figure 4 compares the response from the three groups on performing usual duties.
-It appears that for all groups the situation caused by the COVID-19 has inexpectedly improved the situation in various way, irrespective of the immigrants status of the respondents.
-
-![Figure 4.1: Consequences on performing usual duties](Figures/Entente1.png)
-
-![Figure 4.2: Explicative model](Figures/Abilities.jpeg)
-
-## 3.5. Consequences on employment by immigration status
-
-The consequences on employment are mixed, with the severe consequences more prevalent among the racialised immigrants.
-The Figure 3 shows for instance that close to 45% of the racialised immigrants have stated  that their family income has decreased compare to just 33% of the majority group and 37% of the non-racialed immigrants.
-In addition, 11% of the racialised immigrants reveals having loss their job compared to 9% among the non-racialised immigrants and 2% among members of the majority group.
-At the opposite, the members of the majority groups are more likely to report a reduction of the working hours or for stopping momentarily their work.
-
-![Figure 5.1: Consequences on employment and income](Figures/Emploi1.png)
-
-![Figure 5.2: Explicative model](Figures/Employment.jpeg)
-
-## 3.7 Help received by immigrations status
-
-Finally, we compared the different group on benefitting of the government assistance.
-It appears that racialised immigrants with 35% of the case have expressed having received the Canada Emergency Response Benefit wheras this proportion is 31% for the non-racialised immigrants and 25% for the member of the majority group.
-For the province Temporary Aid for Workers Program, around 5% of the respondents have received it.
-
-![](Figures/Aide1.png)
-
-![Figure 3.2: Explicative model](Figures/Government.jpeg)
-
-# Conclusion
-
-The preliminary results of this study show that the consequences of the pandemic on the Quebec population differ according to the population groups, but also according to the type of consequences.
-
 \newpage
-
-# Références
-
-http://www.oecd.org/coronavirus/policy-responses/what-is-the-impact-of-the-covid-19-pandemic-on-immigrants-and-their-children-e7cbb7de/
-
-Guttmann A, F.
-et al.
-(2020), COVID-19 in Immigrants, Refugees and Other Newcomers in Ontario: Characteristics of Those Tested and Those Confirmed Positive, as of June 13, 2020, ICES, http://www.ices.on.ca.
-
-Hu, Y.
-(2020), “Intersecting ethnic and native–migrant inequalities in the economic impact of the COVID-19 pandemic in the UK”, Research in Social Stratification and Mobility, Vol.
-68, p.
-100528, http://dx.doi.org/10.1016/j.rssm.2020.100528.
-
-Mcginnity, F.
-and G.
-Kingston (2017), “An Irish Welcome?
-Changing Irish Attitudes to Immigrants and Immigration: The Role of Recession and Immigration”, The Economic and social review, Vol.
-48/3, pp.
-281-304, https://www.researchgate.net/publication/319670153_An_Irish_Welcome_Changing_Irish_Attitudes_to_Immigrants_and_Immigration_The_Role_of_Recession_and_Immigration (accessed on 2 October 2020).
-
-Raisi-Estabragh, Z.
-et al.
-(2020), “Greater risk of severe COVID-19 in Black, Asian and Minority Ethnic populations is not explained by cardiometabolic, socioeconomic or behavioural factors, or by 25(OH)-vitamin D status: study of 1326 cases from the UK Biobank”, Journal of Public Health, Vol.
-42/3, pp.
-451-460, http://dx.doi.org/10.1093/pubmed/fdaa095.
-
-Apea, V.
-et al.
-(2020), Ethnicity and outcomes in patients hospitalised with COVID-19 infection in East London: an observational cohort study, Cold Spring Harbor Laboratory, http://dx.doi.org/10.1101/2020.06.10.20127621.
-
-Bertocchi, G. and A. Dimico (2020), “Covid-19, Race, and Redlining”, IZA Discussion Paper, Vol. 13467, https://papers.ssrn.com/sol3/papers.cfm?abstract_id=3648807 (accessed on 22 September 2020).
-
-Brun, S. and P. Simon (2020), “Inégalités ethno-raciales et coronavirus”, De Facto 19.

### 2023-05-02 Expérimentations (Théa)/2. Expérimentations (Bookdown)/Bookdown/03-theoritical-frame.Rmd
Multiple changes on Bookdown

--- 2023-05-01 Article/Trajipaq_Covid_Draft4.Rmd
@@
 ---
-title: 'COVID-19 Disparities in Quebec (Canada): A Comparative Analysis of Exposure
-  Risks, Consequences, and Access to Resources among Immigrants and Natives'
-author:
-- Vissého Adjiwanou ^[corresponding author - adjiwanou.vissého@uqam.ca]
-- Département de Sociologie, Université du Québec à Montréal (UQAM)
-- Georges NGalé Tchango (Département de démographie, Université de Montréal)
-- Solène Lardoux (Département de démographie, Université de Montréal)
-- Nima Zahedinameghi (Département de Sociologie, UQAM)
-date: "`r format(Sys.time(), '%d %B %Y')`"
-output:
-  word_document:
-    keep_md: yes
-  html_document:
-    toc: yes
-    toc_depth: 3
-  template: svm-latex-ms2.tex
-  pdf_document: default
-  citation_package: natbib
-bibliography:
-- covid_Nima.bib
-- Covid_Canada.bib
-- Covid_Family.bib
+knit: bookdown::render_book
+editor_options: 
+  markdown: 
+    wrap: 80
 ---
-<!--
-# Journal: family issues - SAGE
-https://journals-sagepub-com.proxy.bibliotheques.uqam.ca/author-instructions/JFI
 
-https://fishandwhistle.net/post/2020/getting-started-zotero-better-bibtex-rmarkdown/
-https://www.zotero.org/styles?fields=sociology
--->
+# Theoretical framework
+
+Family as a unit of analysis can be particularly fruitful when studying health
+inequalities in the context of a societal-wide crisis. Through the lens of the
+COVID-19 pandemic, the following points illustrate why exploring family dynamics
+provides additional insight into the underlying mechanisms of health
+disparities. First, the family is the most fundamental unit of social
+organization. It facilitates access to resources and the transmission of social
+and psychological capital from one generation to the next. Therefore, a better
+analysis of how crises and adverse life events affect people requires a better
+understanding of how families relate to each other and interact with external
+forces. Second, during public health emergencies, families may be thrown into
+different structures due to unprecedented financial, mental, and physical
+challenges. The risk of family dissolution, for example, can be heightened when
+members must spend unusually long periods together at home and have limited
+opportunities to engage with friends and the wider community. Conversely,
+families can develop closer bonds through self-sacrifice, shared
+responsibilities, and increased care and support. As the dependency status of
+seniors increases, the elderlies are also more likely to be integrated into
+formerly nucleus families. We posit that the extent to which a family can adapt
+its functions to meet new challenges is a predictor of the well-being of its
+members and the resilience of its broader population. Third, the family is a
+crucial link to social institutions, such as the health care system, schools,
+and government agencies whose decisions have significant downstream impacts.
+Families are viewed by many public-level entities as a reference when allocating
+resources, enforcing new standards, and responding to emerging needs and
+demands. From this perspective, the family operates at the nexus of individual
+and societal structures, embedding the tripple effects of public policy in its
+most consequential ways. Finally, the family has a crucial role to play in
+determining the life course of its members. As COVID-19 elevates the importance
+of familial bonds, the post-pandemic period might be marked by an increased
+influence of families on the lives of their members and the social structures
+they inhabit.
+
+The above conceptualization echoes several other approaches to studying
+long-term health inequalities. For instance, life course theory emphasizes the
+significance of family ties and social environment as they structure prospects
+for individuals across their life span (Elder, 1977; Elder Jr., 1998; Macmillan,
+2005; Settersten et al., 2020). Likewise, social capital theory and the
+stress-buffering hypothesis consider family systems a crucial factor in health
+and well-being (Choi and Jun, 2020; Cohen and McKay, 1984; Cohen and Wills,
+1985; Harpham et al., 2004; Kawachi and Berkman, 2001; Lavee et al., 1987;
+Mandelbaum et al., 2020). They show that the family can offer tangible (e.g.,
+food, shelter, and money) and intangible resources (e.g., support, social
+status, psychological comfort, and institutional advantages), protecting its
+members from adverse life events. Similarly, the social-ecological model and the
+theory of fundamental causes highlight how the SES, familial and extrafamilial
+factors pattern people's life trajectories and health outcomes (Barkan and
+Rocque, 2018; Clouston and Link, 2021; Figge et al., 2018; Kim, 2020; Libman et
+al., 2012; Ungar, 2015). However, all these approaches are ostensibly focused on
+life periods where the general social context is relatively stable. In this
+article, we present our conceptualization of the family in times of
+unprecedented social upheaval. Specifically, our framework seeks to characterize
+family dynamics as a suitable lens to analyze health inequalities in the face of
+a virulent public health crisis.
 
 \newpage
-
-# Abstract
-
-The immigrant population is one of the most vulnerable in Canada, with a relatively high risk of unemployment, a low social support system, and linguistic barriers.
-However, their experiences with public health interventions during the COVID-19 pandemic have not been adequately explored.
-The purpose of this article is to analyze Covid-19 exposure risk and its family implications for both natives and immigrants in the province of Quebec.
-Additionally, we analyze immigrants' access to government aid and compare it to their native-born counterparts.
-The study consists of survey data collected from 1,000 immigrants and 500 non-immigrants Quebecers between August and November 2020.
-Our descriptive and exploratory analysis reveals significant differences between the two groups regarding exposure risk, family consequences, and access to resources.
-We conclude that the current pandemic exacerbates health disparities in the province, with particularly dire repercussions for immigrants’ families.
-We also discuss several socio-cultural strategies to mitigate inequality and recommend policies to address the needs of vulnerable populations.
-
-**Keywords**: Immigrants, exposure risk, government aid, health disparities, COVID-19, family disruptions
-
-<!--
-
-Using  script in Rmarkdown
-
-http://zevross.com/blog/2014/07/09/making-use-of-external-r-code-in-knitr-and-r-markdown/
-
-Managing my Zotero
-
-https://forums.zotero.org/discussion/35712/removing-pdfs-to-free-up-space
-
-https://www.zotero.org/support/sync
-
-https://www.zotero.org/support/preferences/sync#file_syncing
-
-https://www.zotero.org/support/sync#webdav
-
-https://www.zotero.org/support/preferences/advanced#linked_attachment_base_directory
-
-http://zotfile.com/
-
-# Abstract
-
-A ce jour, la pandémie du coronavirus a touché près de 100000 personnes au Québec et entrainé plus de 6000 décès.
-Cela fait de la province l'une des plus touchées au Canada et l'une des régions les plus affectées au monde.
-Bien que la majorité de ces décès ait eu lieu dans les maisons pour personnes âgées, et donc a touché plus cette tranche de la population, son influence globale sur les individus n'est pas à négliger.
-En effet, outre les conséquences directes de la pandémie, les mesures prises par les différents gouvernements à travers le monde ne sont pas sans conséquences sur les familles.
-Il est démontré que ses conséquences affectent négativement en priorité les couches les plus défavorisées de la population.
-Les immigrants se concentrent dans bien des cas dans cette couche de la population.
-Cette étude vise ainsi à analyser l'influence de la pandémie sur les immigrants et leurs enfants au Québec, à examiner les facteurs socio-culturels qui permettent d'atténuer son impact négatif.
-
-To date, the coronavirus pandemic has affected nearly 100,000 people in Quebec and caused more than 6,000 deaths.
-This makes the province one of the most affected in Canada and one of the most affected regions in the world.
-Although the majority of these deaths took place in homes for the elderly, and therefore affected this segment of the population more, its overall influence on individuals is not to be overlooked.
-In fact, in addition to the direct consequences of the pandemic, the measures taken by various governments around the world are not without consequences for families.
-It has been shown that its consequences primarily affect the most disadvantaged segments of the population negatively.
-Immigrants are in many cases concentrated in this segment of the population.
-This study thus aims to analyze the influence of the pandemic on immigrants and their children in Quebec, to examine the socio-cultural factors that make it possible to mitigate its negative impact.
-
-À soumettre au Canadian Population Studies
-
-Autre projet possible: intégration des immigrants et covid19 au Québec
-
-**Keywords** :
-
--->
-
-\newpage
-
-# Introduction / background
-
-The COVID-19 pandemic has highlighted the vast ethnic, social, economic, and health disparities in high-income countries, including Canada.
-One of the most vulnerable segments of these affluent societies is the immigrant population, that has become widely diversified with more and more people coming from developing countries, and for whom integration into their new society may be more difficult.
-The coronavirus pandemic brings a new challenge for these immigrant populations, directly or indirectly through the tensions between the ethnic groups that it helps to generate [@coates2020Covid19].
-These immigrants face language barriers, low social capital, housing insecurity, and severe poverty, rendering them particularly susceptible to the effects of the current pandemic.
-Immigrants are also more likely to be employed as essential workers where social distancing, sanitary precautions, and self-isolation, measures put in place to mitigate the pandemy, are challenging [@reid2021Migrant; @yi2021Health; @zhang2021COVID19].
-More importantly, migrants may encounter various forms of discrimination, leading to social exclusion and restricted access to healthcare and social services [@spiritus-beerden2021Mental].
-It is therefore necessary to understand how this pandemic affects these immigrant populations differently and the native populations in Quebec.
-In doing so, this article will reveal a better understanding of how the current pandemic aggravates existing inequalities and challenges in the province for immigrants and their descendants.
-
-<!-- The province of Quebec has one of the highest concentrations of immigrants in Canada, including permanent residents, temporary residents, and refugees [@governmentofcanada2017Daily; @immigration2005Recent].
-In 2016, 1.2 million immigrants were living in Quebec, an increase from 8% in 1990 to 13.7% of the population in 2016 (Institut de la statistique du Québec, 2020).
-<!--As of 2019, the increase in net new residents (60,700 versus 42,200 in 2018) and lower interprovincial migration losses (-1,500 versus -6,100 in 2018) contributed to Quebec's population growth.
--->Since the start of the pandemic, most of the confirmed cases were identified in Montreal and areas with high immigrant populations [@miconi2021Ethnocultural; @shields2020COVID19].
-The coronavirus pandemic has infected nearly 450,000 people in Quebec and has resulted in more than 11,000 deaths (https://www.quebec.ca/sante/problemes-de-sante/a-z/coronavirus-2019/situation-coronavirus-quebec, consulted on octobre 29, 2021).
-This makes it the worst affected province in Canada and one of the worst affected regions in the world during the first wave of the pandemic.
-Although the majority of these deaths took place in elderly's homes, and therefore affected more this segment of the population, its overall influence on individuals is not to be overlooked.
-It has been shown that, on the one hand, immigrants and their families have a higher risk of being exposed to the pathogen, and, at the same time, they are more vulnerable to racism and discrimination and face barriers to accessing social support [@2021CRARR] and information, especially  for those who didn't speak or understand the two official langages.
-<!--In addition to the direct consequences of the pandemic, the measures taken by various governments around the world are also not without consequences on families.
--->Several studies have thus shown excess mortality in ethnocultural minority populations [@bertocchi2020Covid19], higher unemployment, greater loss of income and negative consequences for children and families in several countries.
-This differentiated effect can be explained in different ways.
-On the one hand, in terms of the direct consequences of the pandemic, the situation of precariousness and greater poverty among immigrants exposes them more to the more harmful consequences of the disease on their state of health.
-Likewise, immigrants often work in jobs where social distancing is difficult to enforce.
-On the other hand, indirectly, the solutions to the crisis affect immigrants and natives differently.
-The complete shutdown of tye economy with its set of adjustments has greatly impeded the employability of immigrants, especially newcomers who have not yet grasped the workings of society and who cannot benefit from state subsidies.
-But, it is also shown that these effects are not uniform on the whole of the immigrant population, itself, very diverse.
-
-Despite the growing focus on the differential impacts of the COVID-19 on immigrants and non-immigrants segment of the population, there is still very little knowledge about how the pandemic and its related public policy impact the lives of immigrants and their families.
-The existing health datasets (largely used to assess this effect) are not necessarily representative of the immigrant population and do not offer an accurate picture of the complex social status and health outcomes among this group [@choi2021Studying; @gagnon2021Immigration; @khan2015Visible].
-In most cases, these datasets lack data on the residence length of immigrants and combine newcomers with long-term settlers [@edge2013Discrimination; @hayward2021Clinical].
-As a result, they may not allow a detailed assessment of the nature and extent of risk exposure in sub-segments of populations, as well as whether interventions are reaching those most at risk.
-In addition, the family-level impacts of the pandemic are still poorly understood.
-While a body of research has explored the association between the pandemic and its potential effects on intimate partner violence and marital dissolution [@bhatnagar2021Navigating; @koshan2020COVID19; @morrison2021Chapter; @prime2020Risk], few studies have examined negative familial repercussions in the context of immigration status.
-There is ample evidence that sustained family relationships greatly enhance immigrants' lives on several fronts, including fostering economic prosperity [@marger2001Sociala], psychological well-being [@beiser2002Povertya], social network development [@evansluong2017Familya], and educational success for children [@munroe-blum1989Immigranta].
-The family is, therefore, a key component of long-term social mobility and economic success of immigrants.
-
-<!--Faced with the pandemic, the governments of Canada and Quebec have implemented a number of measures to reduce its impact.
-These policies can be broken down at two levels, direct support to families, including the Canadian Emergency Benefit, which can provide employees and self-employed workers affected by Covid-19 with a financial benefit of up to $ 2,000.
-On the other hand, the government supports many companies to keep their employees or supports them directly when the revenues from their activities have fallen considerably.
-While these measures allow families to cope better with the crisis, the fact remains that the changes brought about by this crisis can affect other dimensions of life.
-How these measures make a difference in the face of the pandemic remains to be seen.
--->Based on an original biographic survey collected around the first wave of the pandemic in august-november 2020 on immigrants and the native population of Québec, this study mainly aims to understand the family exposure, consequences, and responses of the pandemic on immigrant and native Quebec populations.
-We assess a whole range of different measures of each outcome.
-We divided the consequences in three dimensions, 1) access to ressources, 2) ability to perform duties and 3) employment and income.
-We seek also to analyse in what ways these outcomes differ according to the immigrants length of residence.
-By doing so, this study provide a clear and global picture of the situation of different population groups in Quebec in the face of the COVID-19.
-<!-- To respond to these questions, we will use data collected around the first wave of the pandemic in august-november 2020.
-Our study aims to fill the gap by analyzing group-specific differences in exposure risk, family consequences, and access to resources among immigrants and Quebecers.
-We also seek to shed light on several socio-cultural strategies to mitigate inequality and recommend relevant policy solutions.
-The detailed findings will help policymakers address health inequalities and improve equity in Quebec while developing adequate measures and programs.
--->
-
-<!-- Introduction  -->
-
-<!-- Like the rest of Canada, Quebec has experienced a substantial increase in its immigration in recent years.
-The immigrant population in the province thus increased from 8% in the 1990s to 13.7% of the population, or 1.2 million individuals in 2016 (Institut de la statistique du Québec, 2020).
-This immigration is also unique in that it has become widely diversified with more and more people coming from developing countries, and for whom integration into their new society may be more difficult.
-The coronavirus pandemic brings a new challenge for these immigrant populations, directly or indirectly through the tensions between the ethnic groups that it helps to generate [@coates2020Covid19].
-It is therefore necessary to understand how this pandemic affects these immigrant populations differently and the native populations in Quebec.
--->
-
-<!-- The coronavirus pandemic has affected nearly 100,000 people in Quebec and has resulted in more than 6,000 deaths.
-This makes it the worst affected province in Canada and one of the worst affected regions in the world.
-Although the majority of these deaths took place in homes for the elderly, and therefore affected this segment of the population more, its overall influence on individuals is not to be overlooked.
-In fact, in addition to the direct consequences of the pandemic, the measures taken by various governments around the world are also not without consequences on families.
-Several studies have thus shown excess mortality in ethnocultural minority populations [@bertocchi2020Covid19], higher unemployment, greater loss of income and negative consequences for children and families in several countries.
--->
-
-<!-- This differentiated effect can be explained in different ways.
-On the one hand, in terms of the direct consequences of the pandemic, the situation of precariousness and greater poverty among immigrants exposes them more to the more harmful consequences of the disease on their state of health.
-Likewise, immigrants often work in jobs where social distancing is difficult to enforce.
-On the other hand, indirectly, the solutions to the crisis affect immigrants and natives differently.
-The complete shutdown of the company with its set of adjustments has greatly impeded the employability of immigrants, especially newcomers who have not yet grasped the workings of society and who cannot benefit from state subsidies.
-But, it is also shown that these effects are not uniform on the whole of the immigrant population, itself, very diverse.
--->
-
-<!-- Faced with the pandemic, the governments of Canada and Quebec have implemented a number of measures to reduce the impact of the pandemic.
-These policies can be broken down at two levels, direct support to families, including the Canadian Emergency Benefit, which can provide employees and self-employed workers affected by Covid-19 with a financial benefit of up to $ 2,000.
-On the other hand, the government supports many companies to keep their employees or supports them directly when the revenues from their activities have fallen considerably.
-While these measures allow families to cope better with the crisis, the fact remains that the changes brought about by this crisis can affect other dimensions of life.
-How these measures make a difference in the face of the pandemic remains to be seen.
--->
-
-<!-- This study mainly aims to understand the family consequences of the pandemic on immigrant and native Quebec populations.
-Therefore, we will seek to analyse in what ways its consequences differ according to the region of origin of immigrants, length of residence, age and sex.
--->
-
-# Current body of literature and knowledge gaps
-
-The literature on the racial and ethnic health disparities during the COVID-19 crisis can be divided into two main categories: studies addressing the socio-economic determinants of health and those examining the delivery of health interventions and policies.
-The first category is primarily comprised of research that focuses on the link between socio-economic status (SES) of high-risk minority groups and the differential impact of the pandemic [@drefahl2020Populationbased; @greenaway2020COVID19; @lopez2021Racial; @pacheco2020Job; @sundaram2020Individual].
-For instance, an inter-country study has found that Middle Eastern, African, and Nordic immigrants in Sweden have a higher risk of being hospitalized and even dying from COVID-19 when compared to the Swedish born population.
-In the same study, the SES, proportion of working-age households, and neighborhood density were significantly associated with the mortality rate [@rostila2021Disparities].
-This trend of research demonstrates that racial and ethnic minorities are more likely to experience stress factors, not only due to the hazards of infection but also because of economic and social marginalization [@alahmad2020COVID19; @choi2020Buffering; @endale2020COVID19; @fitzpatrick2020Fear; @serafini2021Psychological].
-The intersectionality of race, sex, and health conditions has also been explored in the Canadian context of the pandemic.
-@ingen2021Neighbourhoodlevel demonstrated that Ontarian neighborhoods with immigrant and racialized populations have a disproportionate share of COVID-19 infections.
-In addition, areas with high SES report a lesser number of COVID cases than low SES ones [@ingen2021Neighbourhoodlevel].
-Another study conducted in Quebec confirms the strong association between immigrants' backgrounds, SES, and COVID-19 related psychological stress [@miconi2021Ethnocultural].
-The most recent reviews on the topic suggest that immigrants disproportionately live in crowded and inadequate housing and are employed in essential industries, leaving them with severe income insecurities and poor physiological/psychological well-being [@chew2020Clinical; @davis2020Impact].
-Studies on SES and health disparities have emphasized that this situation is further complicated for immigrant women, constrained by gender and afflicted by a significant power imbalance in the household and workplace [@alarcao2021Intersections; @esegbona-adeigbe2020COVID19; @jehn2021COVID19; @mo2020Differential; @nardon2021Skilled; @sabri2020Effect; @st-denis2020Sociodemographic].
-According to Lightman, immigrants and women of color are overrepresented in long-term care facilities in Canada, where COVID infection rates have been exceptionally high.
-They are also subject to exclusionary social and economic practices [@lightman2021Caring].
-
-The second category of the literature is primarily comprised of policy analyses, research on public health communications, examining the efficacy of COVID initiatives, the optimal allocation of government aids, and ethical issues associated with applying new health interventions to vulnerable populations and their communities [@aragona2020Negative; @cross2021Coronavirus; @desai2020COVID19; @doyle2020Migrant; @gilman2020Modelling; @hargreaves2020Targeting; @langellier2020Policy; @nezafatmaldonado2020Engaging; @silverman2020Ethics; @wilson2020Ethics].
-Some of the questions that this body of research addresses include: how best to respond to COVID-19 in immigrant communities?
-How do public health policies impact the socio-economic status and psychological well-being of different sub-populations?
-What is the structure of the planning and implementation of government aid programs during the current crisis?
-And, how can we design public health communication strategies in a way that is culturally sensitive and inclusive?
-
-There are several knowledge gaps in the existing literature.
-Most of the cited studies do not include household data and family structure of immigrants as variables that affect exposure risks and health outcomes of COVID-19.
-Consequently, the relationship between immigrant status and COVID-19 is not accounted for by household size, and family characteristics (i.e., income, education, religion).
-A key reason is that most of the health data used in such research projects are acquired from national registers or administrative databases, which provide little information on immigration status and kinship.
-However, as the study of Thomeer and her colleagues theorizes, the extent of adverse social and economic impacts of COVID-19 is significantly conditional on the composition of an immigrant's family [@thomeer2020How].
-Households that include a spouse/partner, children, or multigenerational family members most likely experience a different level of social, psychological, and medical vulnerability than single or nuclear ones.
-Therefore, studies of immigrant populations should account for the breakdown of family configuration to better understand the mechanisms linking social determinants and health outcomes.
-Another knowledge gap concerns the current state of Canadian studies of COVID-19 and its public health crisis.
-There is a dearth of literature on the challenges faced by immigrants and their families in the country.
-Although the burden of COVID-19 infection is disproportionately higher among these marginalized groups (St-Denis, 2020; Sundaram et al., 2020), research on the issues they are experiencing is sparse and often limited to one aspect.
-<!--Likewise, there is limited focus on the effectiveness of COVID-19 crisis management for immigrants at the policy level.
-In response to the mounting cases of COVID, the federal government and provinces and territories of Canada have launched multi-pronged public health initiatives to protect society from the epidemic (Migone, 2020).
-However, many open questions remain unanswered, including, do interventions efficiently reach the migrant groups?
-What are the likely effects of each program on different immigrant communities?
-How can an intervention be tailored to the needs of a particular migrant group and their families?
--->
-
-# Theoretical framework 
-Family as a unit of analysis can be particularly fruitful when studying health inequalities in the context of a societal-wide crisis.
-Through the lens of the COVID-19 pandemic, the following points illustrate why exploring family dynamics provides additional insight into the underlying mechanisms of health disparities.
-First, the family is the most fundamental unit of social organization.
-It facilitates access to resources and the transmission of social and psychological capital from one generation to the next.
-Therefore, a better analysis of how crises and adverse life events affect people requires a better understanding of how families relate to each other and interact with external forces.
-Second, during public health emergencies, families may be thrown into different structures due to unprecedented financial, mental, and physical challenges.
-The risk of family dissolution, for example, can be heightened when members must spend unusually long periods together at home and have limited opportunities to engage with friends and the wider community.
-Conversely, families can develop closer bonds through self-sacrifice, shared responsibilities, and increased care and support.
-As the dependency status of seniors increases, the elderlies are also more likely to be integrated into formerly nucleus families.
-We posit that the extent to which a family can adapt its functions to meet new challenges is a predictor of the well-being of its members and the resilience of its broader population.
-Third, the family is a crucial link to social institutions, such as the health care system, schools, and government agencies whose decisions have significant downstream impacts.
-Families are viewed by many public-level entities as a reference when allocating resources, enforcing new standards, and responding to emerging needs and demands.
-From this perspective, the family operates at the nexus of individual and societal structures, embedding the tripple effects of public policy in its most consequential ways.
-Finally, the family has a crucial role to play in determining the life course of its members.
-As COVID-19 elevates the importance of familial bonds, the post-pandemic period might be marked by an increased influence of families on the lives of their members and the social structures they inhabit.
-
-The above conceptualization echoes several other approaches to studying long-term health inequalities.
-For instance, life course theory emphasizes the significance of family ties and social environment as they structure prospects for individuals across their life span (Elder, 1977; Elder Jr., 1998; Macmillan, 2005; Settersten et al., 2020).
-Likewise, social capital theory and the stress-buffering hypothesis consider family systems a crucial factor in health and well-being (Choi and Jun, 2020; Cohen and McKay, 1984; Cohen and Wills, 1985; Harpham et al., 2004; Kawachi and Berkman, 2001; Lavee et al., 1987; Mandelbaum et al., 2020).
-They show that the family can offer tangible (e.g., food, shelter, and money) and intangible resources (e.g., support, social status, psychological comfort, and institutional advantages), protecting its members from adverse life events.
-Similarly, the social-ecological model and the theory of fundamental causes highlight how the SES, familial and extrafamilial factors pattern people's life trajectories and health outcomes (Barkan and Rocque, 2018; Clouston and Link, 2021; Figge et al., 2018; Kim, 2020; Libman et al., 2012; Ungar, 2015).
-However, all these approaches are ostensibly focused on life periods where the general social context is relatively stable.
-In this article, we present our conceptualization of the family in times of unprecedented social upheaval.
-Specifically, our framework seeks to characterize family dynamics as a suitable lens to analyze health inequalities in the face of a virulent public health crisis.
-
-# Methodology
-
-## Data
-
-The data for this study come from an original collection of biographical data on immigrants and the native population of Quebec to analyse their participation in Quebec society in various fields (social, economic, cultural, etc.).
-The immigrant population is limited to recent immigration that has taken place over the past 15 years.
-This survey has several components that retrace the life course of individuals over a long period of time.
-The pandemic component uses a reduced version of the [harmonized questionnaire on exposure and the impact of Covid19 on the family (CEFIS)](https://www.phenxtoolkit.org/toolkit_content/PDF/CPTS_CEFIS_FullSurveyEng.pdf).
-The questions are divided into three parts, on exposure to the disease, its consequences and the benefit of the policies put in place by both the federal and provincial governments.
-The other parts of the questionnaire focus on the characteristics of the respondents, their immigration history, their educational and family background.
-In short, it is a detailed questionnaire that covers the life of immigrants from birth and especially over the past 15 years.
-The data collection lasted three months, from August to November 2020. In total, 1,500 people were surveyed, including around 1,000 immigrants and 500 natives throughout Quebec.
-
-### Dependent variables
-
-We distinguish three groups of dependent variables:
-
-**1.	Disease exposure variables**: these are measured by six questions that seek to document whether the individual through work or a family member may have been exposed to the disease.
-For example, it is asked if a family member has had symptoms of the disease, or if a member has been hospitalized due to the contraction of the disease.
-
-**2.	Variables on the family consequences of the disease**: we measure three dimensions of consequences.
-The first group of measures assess the consequences on accessing ressources by the respondent or any member of his family.
-The second group of measure assess whether the respondent or any family member have difficulties to perform usual duties toward any member of the family.
-Finally, the last group of measures measure the consequences of the pandemic on employment and income of any family member.
-
-**3.	Variables on government assistance**: Finally, government support is measured by the following two questions that assessed whether the respondent or any family member has benefited of the temporary assistance for workers from the Government of Quebec (PATT) or of the Canadian Federal Government Emergency Benefit (CEP).
-The table XXX in the apendix presents the full set of indicators of the dependents variables
-
-### Main independent variable
-The main independent variable is immigrant status coupled with visible minority status.
-This variable takes the following modalities: 1 if the respondent is a non-racialized native, 2 if he is a non-racialized immigrant, 3 if he is a racialized immigrant, and 4 if the respondent is a racialized native or an immigrant who has not answered the visible minority question.
-The other variables in the analysis are: age, sex, length of residence, minority status, size of social network.
-
-## Analysis methods
-
-We use two main methods of analysis, a descriptive method and an explanatory method.
-In the descriptive method, we consider each of the independent variables in isolation and describe how they vary depending on the main independent variable.
-The explanatory methods goes further by assessing whether the effect remain when controlling for other covariates.
-Thus, by controlling for the other explanatory variables, we determine the effect of immigrant status on exposure to the pandemic, on its consequences and on the fact of benefiting or not of government assistance.
-In a second series of modelling, we assess whether the effect that we observed varies by the immigrant's length of residence or by marital status.
-We use either a logistic regression on the outcome variables that are binary and an ordonned logistic regression for the outcome that are measure with a 5-point Lickert scale.
-<!-- PUT THIS AS A LIMITATION IN THE CONCLUSION
-To be able to benefit from government benefits, one must meet a number of conditions, including the loss of job due to Covid-19.
-As a result, the beneficiaries are not the general population but those who have lost their jobs and this may itself be due to immigrant status.
-Thus, for model 3, we will use Heckman's models by first estimating the fact of having lost or not a job and then, the fact of receiving or not the government benefit.
--->
-
-# 3. Results
-
-## 3.1. Sample distribution
-
-## 3.2. Exposure by immigration status
-
-Figure 1 presents the exposure to the pandemy accroding to the immigrants and minority visible status.
-For all the six measures of exposure, it appears that the majority group is the most exposed to the COVID-19 in all instance.
-The big difference between the groups is about the exposure trhough working outside.
-Around 50% of the member of this group or a family member work outside the home at the time of the survey.
-This percentage is only 30% for the racialised immigrants and 25% for the non-racialised immigrants.
-
-![Figure 1.1: Exposure](Figures/Exposition1.png){width=60%}
-
-![Figure 1.2: Explicative model](Figures/Exposure.jpeg)
-
-## 3.3. Consequences by immigrations status
-
-The Figure 2 presents the consequences of the COVID-19 for a family member to access to ressources.
-It shows that in all accounts but one (getting medecine), the racialised immigrants express more diffciult to assess ressources.
-Around 20% of them stated having difficulty to get health care compare, around 10% for obtaining food and 7% for obtaining family insurrance.
-Comparatively, these proportion are respectively 13%, 7% and 4% for people and thier family member in the majority group and 19%, 3% and 4% respectively among the racialised immigrants.
-
-![Figure 3.1: Consequences on accessing ressources](Figures/Famille1.png){width=80%}
-
-![Figure 3.2: Explicative model](Figures/Ressources.jpeg)
-
-## 3.4. Consequences on performing usual duties
-
-The Figure 4 compares the response from the three groups on performing usual duties.
-It appears that for all groups the situation caused by the COVID-19 has inexpectedly improved the situation in various way, irrespective of the immigrants status of the respondents.
-
-![Figure 4.1: Consequences on performing usual duties](Figures/Entente1.png)
-
-![Figure 4.2: Explicative model](Figures/Abilities.jpeg)
-
-## 3.5. Consequences on employment by immigration status
-
-The consequences on employment are mixed, with the severe consequences more prevalent among the racialised immigrants.
-The Figure 3 shows for instance that close to 45% of the racialised immigrants have stated  that their family income has decreased compare to just 33% of the majority group and 37% of the non-racialed immigrants.
-In addition, 11% of the racialised immigrants reveals having loss their job compared to 9% among the non-racialised immigrants and 2% among members of the majority group.
-At the opposite, the members of the majority groups are more likely to report a reduction of the working hours or for stopping momentarily their work.
-
-![Figure 5.1: Consequences on employment and income](Figures/Emploi1.png)
-
-![Figure 5.2: Explicative model](Figures/Employment.jpeg)
-
-## 3.7 Help received by immigrations status
-
-Finally, we compared the different group on benefitting of the government assistance.
-It appears that racialised immigrants with 35% of the case have expressed having received the Canada Emergency Response Benefit wheras this proportion is 31% for the non-racialised immigrants and 25% for the member of the majority group.
-For the province Temporary Aid for Workers Program, around 5% of the respondents have received it.
-
-![](Figures/Aide1.png)
-
-![Figure 3.2: Explicative model](Figures/Government.jpeg)
-
-# Conclusion
-
-The preliminary results of this study show that the consequences of the pandemic on the Quebec population differ according to the population groups, but also according to the type of consequences.
-
-\newpage
-
-# Références
-
-http://www.oecd.org/coronavirus/policy-responses/what-is-the-impact-of-the-covid-19-pandemic-on-immigrants-and-their-children-e7cbb7de/
-
-Guttmann A, F.
-et al.
-(2020), COVID-19 in Immigrants, Refugees and Other Newcomers in Ontario: Characteristics of Those Tested and Those Confirmed Positive, as of June 13, 2020, ICES, http://www.ices.on.ca.
-
-Hu, Y.
-(2020), “Intersecting ethnic and native–migrant inequalities in the economic impact of the COVID-19 pandemic in the UK”, Research in Social Stratification and Mobility, Vol.
-68, p.
-100528, http://dx.doi.org/10.1016/j.rssm.2020.100528.
-
-Mcginnity, F.
-and G.
-Kingston (2017), “An Irish Welcome?
-Changing Irish Attitudes to Immigrants and Immigration: The Role of Recession and Immigration”, The Economic and social review, Vol.
-48/3, pp.
-281-304, https://www.researchgate.net/publication/319670153_An_Irish_Welcome_Changing_Irish_Attitudes_to_Immigrants_and_Immigration_The_Role_of_Recession_and_Immigration (accessed on 2 October 2020).
-
-Raisi-Estabragh, Z.
-et al.
-(2020), “Greater risk of severe COVID-19 in Black, Asian and Minority Ethnic populations is not explained by cardiometabolic, socioeconomic or behavioural factors, or by 25(OH)-vitamin D status: study of 1326 cases from the UK Biobank”, Journal of Public Health, Vol.
-42/3, pp.
-451-460, http://dx.doi.org/10.1093/pubmed/fdaa095.
-
-Apea, V.
-et al.
-(2020), Ethnicity and outcomes in patients hospitalised with COVID-19 infection in East London: an observational cohort study, Cold Spring Harbor Laboratory, http://dx.doi.org/10.1101/2020.06.10.20127621.
-
-Bertocchi, G. and A. Dimico (2020), “Covid-19, Race, and Redlining”, IZA Discussion Paper, Vol. 13467, https://papers.ssrn.com/sol3/papers.cfm?abstract_id=3648807 (accessed on 22 September 2020).
-
-Brun, S. and P. Simon (2020), “Inégalités ethno-raciales et coronavirus”, De Facto 19.

### 2023-05-02 Expérimentations (Théa)/2. Expérimentations (Bookdown)/Bookdown/04-methodology.Rmd
Multiple changes on Bookdown

--- 2023-05-01 Article/Trajipaq_Covid_Draft4.Rmd
@@
 ---
-title: 'COVID-19 Disparities in Quebec (Canada): A Comparative Analysis of Exposure
-  Risks, Consequences, and Access to Resources among Immigrants and Natives'
-author:
-- Vissého Adjiwanou ^[corresponding author - adjiwanou.vissého@uqam.ca]
-- Département de Sociologie, Université du Québec à Montréal (UQAM)
-- Georges NGalé Tchango (Département de démographie, Université de Montréal)
-- Solène Lardoux (Département de démographie, Université de Montréal)
-- Nima Zahedinameghi (Département de Sociologie, UQAM)
-date: "`r format(Sys.time(), '%d %B %Y')`"
-output:
-  word_document:
-    keep_md: yes
-  html_document:
-    toc: yes
-    toc_depth: 3
-  template: svm-latex-ms2.tex
-  pdf_document: default
-  citation_package: natbib
-bibliography:
-- covid_Nima.bib
-- Covid_Canada.bib
-- Covid_Family.bib
+knit: bookdown::render_book
+editor_options: 
+  markdown: 
+    wrap: 80
 ---
-<!--
-# Journal: family issues - SAGE
-https://journals-sagepub-com.proxy.bibliotheques.uqam.ca/author-instructions/JFI
-
-https://fishandwhistle.net/post/2020/getting-started-zotero-better-bibtex-rmarkdown/
-https://www.zotero.org/styles?fields=sociology
--->
-
-\newpage
-
-# Abstract
-
-The immigrant population is one of the most vulnerable in Canada, with a relatively high risk of unemployment, a low social support system, and linguistic barriers.
-However, their experiences with public health interventions during the COVID-19 pandemic have not been adequately explored.
-The purpose of this article is to analyze Covid-19 exposure risk and its family implications for both natives and immigrants in the province of Quebec.
-Additionally, we analyze immigrants' access to government aid and compare it to their native-born counterparts.
-The study consists of survey data collected from 1,000 immigrants and 500 non-immigrants Quebecers between August and November 2020.
-Our descriptive and exploratory analysis reveals significant differences between the two groups regarding exposure risk, family consequences, and access to resources.
-We conclude that the current pandemic exacerbates health disparities in the province, with particularly dire repercussions for immigrants’ families.
-We also discuss several socio-cultural strategies to mitigate inequality and recommend policies to address the needs of vulnerable populations.
-
-**Keywords**: Immigrants, exposure risk, government aid, health disparities, COVID-19, family disruptions
-
-<!--
-
-Using  script in Rmarkdown
-
-http://zevross.com/blog/2014/07/09/making-use-of-external-r-code-in-knitr-and-r-markdown/
-
-Managing my Zotero
-
-https://forums.zotero.org/discussion/35712/removing-pdfs-to-free-up-space
-
-https://www.zotero.org/support/sync
-
-https://www.zotero.org/support/preferences/sync#file_syncing
-
-https://www.zotero.org/support/sync#webdav
-
-https://www.zotero.org/support/preferences/advanced#linked_attachment_base_directory
-
-http://zotfile.com/
-
-# Abstract
-
-A ce jour, la pandémie du coronavirus a touché près de 100000 personnes au Québec et entrainé plus de 6000 décès.
-Cela fait de la province l'une des plus touchées au Canada et l'une des régions les plus affectées au monde.
-Bien que la majorité de ces décès ait eu lieu dans les maisons pour personnes âgées, et donc a touché plus cette tranche de la population, son influence globale sur les individus n'est pas à négliger.
-En effet, outre les conséquences directes de la pandémie, les mesures prises par les différents gouvernements à travers le monde ne sont pas sans conséquences sur les familles.
-Il est démontré que ses conséquences affectent négativement en priorité les couches les plus défavorisées de la population.
-Les immigrants se concentrent dans bien des cas dans cette couche de la population.
-Cette étude vise ainsi à analyser l'influence de la pandémie sur les immigrants et leurs enfants au Québec, à examiner les facteurs socio-culturels qui permettent d'atténuer son impact négatif.
-
-To date, the coronavirus pandemic has affected nearly 100,000 people in Quebec and caused more than 6,000 deaths.
-This makes the province one of the most affected in Canada and one of the most affected regions in the world.
-Although the majority of these deaths took place in homes for the elderly, and therefore affected this segment of the population more, its overall influence on individuals is not to be overlooked.
-In fact, in addition to the direct consequences of the pandemic, the measures taken by various governments around the world are not without consequences for families.
-It has been shown that its consequences primarily affect the most disadvantaged segments of the population negatively.
-Immigrants are in many cases concentrated in this segment of the population.
-This study thus aims to analyze the influence of the pandemic on immigrants and their children in Quebec, to examine the socio-cultural factors that make it possible to mitigate its negative impact.
-
-À soumettre au Canadian Population Studies
-
-Autre projet possible: intégration des immigrants et covid19 au Québec
-
-**Keywords** :
-
--->
-
-\newpage
-
-# Introduction / background
-
-The COVID-19 pandemic has highlighted the vast ethnic, social, economic, and health disparities in high-income countries, including Canada.
-One of the most vulnerable segments of these affluent societies is the immigrant population, that has become widely diversified with more and more people coming from developing countries, and for whom integration into their new society may be more difficult.
-The coronavirus pandemic brings a new challenge for these immigrant populations, directly or indirectly through the tensions between the ethnic groups that it helps to generate [@coates2020Covid19].
-These immigrants face language barriers, low social capital, housing insecurity, and severe poverty, rendering them particularly susceptible to the effects of the current pandemic.
-Immigrants are also more likely to be employed as essential workers where social distancing, sanitary precautions, and self-isolation, measures put in place to mitigate the pandemy, are challenging [@reid2021Migrant; @yi2021Health; @zhang2021COVID19].
-More importantly, migrants may encounter various forms of discrimination, leading to social exclusion and restricted access to healthcare and social services [@spiritus-beerden2021Mental].
-It is therefore necessary to understand how this pandemic affects these immigrant populations differently and the native populations in Quebec.
-In doing so, this article will reveal a better understanding of how the current pandemic aggravates existing inequalities and challenges in the province for immigrants and their descendants.
-
-<!-- The province of Quebec has one of the highest concentrations of immigrants in Canada, including permanent residents, temporary residents, and refugees [@governmentofcanada2017Daily; @immigration2005Recent].
-In 2016, 1.2 million immigrants were living in Quebec, an increase from 8% in 1990 to 13.7% of the population in 2016 (Institut de la statistique du Québec, 2020).
-<!--As of 2019, the increase in net new residents (60,700 versus 42,200 in 2018) and lower interprovincial migration losses (-1,500 versus -6,100 in 2018) contributed to Quebec's population growth.
--->Since the start of the pandemic, most of the confirmed cases were identified in Montreal and areas with high immigrant populations [@miconi2021Ethnocultural; @shields2020COVID19].
-The coronavirus pandemic has infected nearly 450,000 people in Quebec and has resulted in more than 11,000 deaths (https://www.quebec.ca/sante/problemes-de-sante/a-z/coronavirus-2019/situation-coronavirus-quebec, consulted on octobre 29, 2021).
-This makes it the worst affected province in Canada and one of the worst affected regions in the world during the first wave of the pandemic.
-Although the majority of these deaths took place in elderly's homes, and therefore affected more this segment of the population, its overall influence on individuals is not to be overlooked.
-It has been shown that, on the one hand, immigrants and their families have a higher risk of being exposed to the pathogen, and, at the same time, they are more vulnerable to racism and discrimination and face barriers to accessing social support [@2021CRARR] and information, especially  for those who didn't speak or understand the two official langages.
-<!--In addition to the direct consequences of the pandemic, the measures taken by various governments around the world are also not without consequences on families.
--->Several studies have thus shown excess mortality in ethnocultural minority populations [@bertocchi2020Covid19], higher unemployment, greater loss of income and negative consequences for children and families in several countries.
-This differentiated effect can be explained in different ways.
-On the one hand, in terms of the direct consequences of the pandemic, the situation of precariousness and greater poverty among immigrants exposes them more to the more harmful consequences of the disease on their state of health.
-Likewise, immigrants often work in jobs where social distancing is difficult to enforce.
-On the other hand, indirectly, the solutions to the crisis affect immigrants and natives differently.
-The complete shutdown of tye economy with its set of adjustments has greatly impeded the employability of immigrants, especially newcomers who have not yet grasped the workings of society and who cannot benefit from state subsidies.
-But, it is also shown that these effects are not uniform on the whole of the immigrant population, itself, very diverse.
-
-Despite the growing focus on the differential impacts of the COVID-19 on immigrants and non-immigrants segment of the population, there is still very little knowledge about how the pandemic and its related public policy impact the lives of immigrants and their families.
-The existing health datasets (largely used to assess this effect) are not necessarily representative of the immigrant population and do not offer an accurate picture of the complex social status and health outcomes among this group [@choi2021Studying; @gagnon2021Immigration; @khan2015Visible].
-In most cases, these datasets lack data on the residence length of immigrants and combine newcomers with long-term settlers [@edge2013Discrimination; @hayward2021Clinical].
-As a result, they may not allow a detailed assessment of the nature and extent of risk exposure in sub-segments of populations, as well as whether interventions are reaching those most at risk.
-In addition, the family-level impacts of the pandemic are still poorly understood.
-While a body of research has explored the association between the pandemic and its potential effects on intimate partner violence and marital dissolution [@bhatnagar2021Navigating; @koshan2020COVID19; @morrison2021Chapter; @prime2020Risk], few studies have examined negative familial repercussions in the context of immigration status.
-There is ample evidence that sustained family relationships greatly enhance immigrants' lives on several fronts, including fostering economic prosperity [@marger2001Sociala], psychological well-being [@beiser2002Povertya], social network development [@evansluong2017Familya], and educational success for children [@munroe-blum1989Immigranta].
-The family is, therefore, a key component of long-term social mobility and economic success of immigrants.
-
-<!--Faced with the pandemic, the governments of Canada and Quebec have implemented a number of measures to reduce its impact.
-These policies can be broken down at two levels, direct support to families, including the Canadian Emergency Benefit, which can provide employees and self-employed workers affected by Covid-19 with a financial benefit of up to $ 2,000.
-On the other hand, the government supports many companies to keep their employees or supports them directly when the revenues from their activities have fallen considerably.
-While these measures allow families to cope better with the crisis, the fact remains that the changes brought about by this crisis can affect other dimensions of life.
-How these measures make a difference in the face of the pandemic remains to be seen.
--->Based on an original biographic survey collected around the first wave of the pandemic in august-november 2020 on immigrants and the native population of Québec, this study mainly aims to understand the family exposure, consequences, and responses of the pandemic on immigrant and native Quebec populations.
-We assess a whole range of different measures of each outcome.
-We divided the consequences in three dimensions, 1) access to ressources, 2) ability to perform duties and 3) employment and income.
-We seek also to analyse in what ways these outcomes differ according to the immigrants length of residence.
-By doing so, this study provide a clear and global picture of the situation of different population groups in Quebec in the face of the COVID-19.
-<!-- To respond to these questions, we will use data collected around the first wave of the pandemic in august-november 2020.
-Our study aims to fill the gap by analyzing group-specific differences in exposure risk, family consequences, and access to resources among immigrants and Quebecers.
-We also seek to shed light on several socio-cultural strategies to mitigate inequality and recommend relevant policy solutions.
-The detailed findings will help policymakers address health inequalities and improve equity in Quebec while developing adequate measures and programs.
--->
-
-<!-- Introduction  -->
-
-<!-- Like the rest of Canada, Quebec has experienced a substantial increase in its immigration in recent years.
-The immigrant population in the province thus increased from 8% in the 1990s to 13.7% of the population, or 1.2 million individuals in 2016 (Institut de la statistique du Québec, 2020).
-This immigration is also unique in that it has become widely diversified with more and more people coming from developing countries, and for whom integration into their new society may be more difficult.
-The coronavirus pandemic brings a new challenge for these immigrant populations, directly or indirectly through the tensions between the ethnic groups that it helps to generate [@coates2020Covid19].
-It is therefore necessary to understand how this pandemic affects these immigrant populations differently and the native populations in Quebec.
--->
-
-<!-- The coronavirus pandemic has affected nearly 100,000 people in Quebec and has resulted in more than 6,000 deaths.
-This makes it the worst affected province in Canada and one of the worst affected regions in the world.
-Although the majority of these deaths took place in homes for the elderly, and therefore affected this segment of the population more, its overall influence on individuals is not to be overlooked.
-In fact, in addition to the direct consequences of the pandemic, the measures taken by various governments around the world are also not without consequences on families.
-Several studies have thus shown excess mortality in ethnocultural minority populations [@bertocchi2020Covid19], higher unemployment, greater loss of income and negative consequences for children and families in several countries.
--->
-
-<!-- This differentiated effect can be explained in different ways.
-On the one hand, in terms of the direct consequences of the pandemic, the situation of precariousness and greater poverty among immigrants exposes them more to the more harmful consequences of the disease on their state of health.
-Likewise, immigrants often work in jobs where social distancing is difficult to enforce.
-On the other hand, indirectly, the solutions to the crisis affect immigrants and natives differently.
-The complete shutdown of the company with its set of adjustments has greatly impeded the employability of immigrants, especially newcomers who have not yet grasped the workings of society and who cannot benefit from state subsidies.
-But, it is also shown that these effects are not uniform on the whole of the immigrant population, itself, very diverse.
--->
-
-<!-- Faced with the pandemic, the governments of Canada and Quebec have implemented a number of measures to reduce the impact of the pandemic.
-These policies can be broken down at two levels, direct support to families, including the Canadian Emergency Benefit, which can provide employees and self-employed workers affected by Covid-19 with a financial benefit of up to $ 2,000.
-On the other hand, the government supports many companies to keep their employees or supports them directly when the revenues from their activities have fallen considerably.
-While these measures allow families to cope better with the crisis, the fact remains that the changes brought about by this crisis can affect other dimensions of life.
-How these measures make a difference in the face of the pandemic remains to be seen.
--->
-
-<!-- This study mainly aims to understand the family consequences of the pandemic on immigrant and native Quebec populations.
-Therefore, we will seek to analyse in what ways its consequences differ according to the region of origin of immigrants, length of residence, age and sex.
--->
-
-# Current body of literature and knowledge gaps
-
-The literature on the racial and ethnic health disparities during the COVID-19 crisis can be divided into two main categories: studies addressing the socio-economic determinants of health and those examining the delivery of health interventions and policies.
-The first category is primarily comprised of research that focuses on the link between socio-economic status (SES) of high-risk minority groups and the differential impact of the pandemic [@drefahl2020Populationbased; @greenaway2020COVID19; @lopez2021Racial; @pacheco2020Job; @sundaram2020Individual].
-For instance, an inter-country study has found that Middle Eastern, African, and Nordic immigrants in Sweden have a higher risk of being hospitalized and even dying from COVID-19 when compared to the Swedish born population.
-In the same study, the SES, proportion of working-age households, and neighborhood density were significantly associated with the mortality rate [@rostila2021Disparities].
-This trend of research demonstrates that racial and ethnic minorities are more likely to experience stress factors, not only due to the hazards of infection but also because of economic and social marginalization [@alahmad2020COVID19; @choi2020Buffering; @endale2020COVID19; @fitzpatrick2020Fear; @serafini2021Psychological].
-The intersectionality of race, sex, and health conditions has also been explored in the Canadian context of the pandemic.
-@ingen2021Neighbourhoodlevel demonstrated that Ontarian neighborhoods with immigrant and racialized populations have a disproportionate share of COVID-19 infections.
-In addition, areas with high SES report a lesser number of COVID cases than low SES ones [@ingen2021Neighbourhoodlevel].
-Another study conducted in Quebec confirms the strong association between immigrants' backgrounds, SES, and COVID-19 related psychological stress [@miconi2021Ethnocultural].
-The most recent reviews on the topic suggest that immigrants disproportionately live in crowded and inadequate housing and are employed in essential industries, leaving them with severe income insecurities and poor physiological/psychological well-being [@chew2020Clinical; @davis2020Impact].
-Studies on SES and health disparities have emphasized that this situation is further complicated for immigrant women, constrained by gender and afflicted by a significant power imbalance in the household and workplace [@alarcao2021Intersections; @esegbona-adeigbe2020COVID19; @jehn2021COVID19; @mo2020Differential; @nardon2021Skilled; @sabri2020Effect; @st-denis2020Sociodemographic].
-According to Lightman, immigrants and women of color are overrepresented in long-term care facilities in Canada, where COVID infection rates have been exceptionally high.
-They are also subject to exclusionary social and economic practices [@lightman2021Caring].
-
-The second category of the literature is primarily comprised of policy analyses, research on public health communications, examining the efficacy of COVID initiatives, the optimal allocation of government aids, and ethical issues associated with applying new health interventions to vulnerable populations and their communities [@aragona2020Negative; @cross2021Coronavirus; @desai2020COVID19; @doyle2020Migrant; @gilman2020Modelling; @hargreaves2020Targeting; @langellier2020Policy; @nezafatmaldonado2020Engaging; @silverman2020Ethics; @wilson2020Ethics].
-Some of the questions that this body of research addresses include: how best to respond to COVID-19 in immigrant communities?
-How do public health policies impact the socio-economic status and psychological well-being of different sub-populations?
-What is the structure of the planning and implementation of government aid programs during the current crisis?
-And, how can we design public health communication strategies in a way that is culturally sensitive and inclusive?
-
-There are several knowledge gaps in the existing literature.
-Most of the cited studies do not include household data and family structure of immigrants as variables that affect exposure risks and health outcomes of COVID-19.
-Consequently, the relationship between immigrant status and COVID-19 is not accounted for by household size, and family characteristics (i.e., income, education, religion).
-A key reason is that most of the health data used in such research projects are acquired from national registers or administrative databases, which provide little information on immigration status and kinship.
-However, as the study of Thomeer and her colleagues theorizes, the extent of adverse social and economic impacts of COVID-19 is significantly conditional on the composition of an immigrant's family [@thomeer2020How].
-Households that include a spouse/partner, children, or multigenerational family members most likely experience a different level of social, psychological, and medical vulnerability than single or nuclear ones.
-Therefore, studies of immigrant populations should account for the breakdown of family configuration to better understand the mechanisms linking social determinants and health outcomes.
-Another knowledge gap concerns the current state of Canadian studies of COVID-19 and its public health crisis.
-There is a dearth of literature on the challenges faced by immigrants and their families in the country.
-Although the burden of COVID-19 infection is disproportionately higher among these marginalized groups (St-Denis, 2020; Sundaram et al., 2020), research on the issues they are experiencing is sparse and often limited to one aspect.
-<!--Likewise, there is limited focus on the effectiveness of COVID-19 crisis management for immigrants at the policy level.
-In response to the mounting cases of COVID, the federal government and provinces and territories of Canada have launched multi-pronged public health initiatives to protect society from the epidemic (Migone, 2020).
-However, many open questions remain unanswered, including, do interventions efficiently reach the migrant groups?
-What are the likely effects of each program on different immigrant communities?
-How can an intervention be tailored to the needs of a particular migrant group and their families?
--->
-
-# Theoretical framework 
-Family as a unit of analysis can be particularly fruitful when studying health inequalities in the context of a societal-wide crisis.
-Through the lens of the COVID-19 pandemic, the following points illustrate why exploring family dynamics provides additional insight into the underlying mechanisms of health disparities.
-First, the family is the most fundamental unit of social organization.
-It facilitates access to resources and the transmission of social and psychological capital from one generation to the next.
-Therefore, a better analysis of how crises and adverse life events affect people requires a better understanding of how families relate to each other and interact with external forces.
-Second, during public health emergencies, families may be thrown into different structures due to unprecedented financial, mental, and physical challenges.
-The risk of family dissolution, for example, can be heightened when members must spend unusually long periods together at home and have limited opportunities to engage with friends and the wider community.
-Conversely, families can develop closer bonds through self-sacrifice, shared responsibilities, and increased care and support.
-As the dependency status of seniors increases, the elderlies are also more likely to be integrated into formerly nucleus families.
-We posit that the extent to which a family can adapt its functions to meet new challenges is a predictor of the well-being of its members and the resilience of its broader population.
-Third, the family is a crucial link to social institutions, such as the health care system, schools, and government agencies whose decisions have significant downstream impacts.
-Families are viewed by many public-level entities as a reference when allocating resources, enforcing new standards, and responding to emerging needs and demands.
-From this perspective, the family operates at the nexus of individual and societal structures, embedding the tripple effects of public policy in its most consequential ways.
-Finally, the family has a crucial role to play in determining the life course of its members.
-As COVID-19 elevates the importance of familial bonds, the post-pandemic period might be marked by an increased influence of families on the lives of their members and the social structures they inhabit.
-
-The above conceptualization echoes several other approaches to studying long-term health inequalities.
-For instance, life course theory emphasizes the significance of family ties and social environment as they structure prospects for individuals across their life span (Elder, 1977; Elder Jr., 1998; Macmillan, 2005; Settersten et al., 2020).
-Likewise, social capital theory and the stress-buffering hypothesis consider family systems a crucial factor in health and well-being (Choi and Jun, 2020; Cohen and McKay, 1984; Cohen and Wills, 1985; Harpham et al., 2004; Kawachi and Berkman, 2001; Lavee et al., 1987; Mandelbaum et al., 2020).
-They show that the family can offer tangible (e.g., food, shelter, and money) and intangible resources (e.g., support, social status, psychological comfort, and institutional advantages), protecting its members from adverse life events.
-Similarly, the social-ecological model and the theory of fundamental causes highlight how the SES, familial and extrafamilial factors pattern people's life trajectories and health outcomes (Barkan and Rocque, 2018; Clouston and Link, 2021; Figge et al., 2018; Kim, 2020; Libman et al., 2012; Ungar, 2015).
-However, all these approaches are ostensibly focused on life periods where the general social context is relatively stable.
-In this article, we present our conceptualization of the family in times of unprecedented social upheaval.
-Specifically, our framework seeks to characterize family dynamics as a suitable lens to analyze health inequalities in the face of a virulent public health crisis.
 
@@
 
-The data for this study come from an original collection of biographical data on immigrants and the native population of Quebec to analyse their participation in Quebec society in various fields (social, economic, cultural, etc.).
-The immigrant population is limited to recent immigration that has taken place over the past 15 years.
-This survey has several components that retrace the life course of individuals over a long period of time.
-The pandemic component uses a reduced version of the [harmonized questionnaire on exposure and the impact of Covid19 on the family (CEFIS)](https://www.phenxtoolkit.org/toolkit_content/PDF/CPTS_CEFIS_FullSurveyEng.pdf).
-The questions are divided into three parts, on exposure to the disease, its consequences and the benefit of the policies put in place by both the federal and provincial governments.
-The other parts of the questionnaire focus on the characteristics of the respondents, their immigration history, their educational and family background.
-In short, it is a detailed questionnaire that covers the life of immigrants from birth and especially over the past 15 years.
-The data collection lasted three months, from August to November 2020. In total, 1,500 people were surveyed, including around 1,000 immigrants and 500 natives throughout Quebec.
+The data for this study come from an original collection of biographical data on
+immigrants and the native population of Quebec to analyse their participation in
+Quebec society in various fields (social, economic, cultural, etc.). The
+immigrant population is limited to recent immigration that has taken place over
+the past 15 years. This survey has several components that retrace the life
+course of individuals over a long period of time. The pandemic component uses a
+reduced version of the [harmonized questionnaire on exposure and the impact of
+Covid19 on the family
+(CEFIS)](https://www.phenxtoolkit.org/toolkit_content/PDF/CPTS_CEFIS_FullSurveyEng.pdf).
+The questions are divided into three parts, on exposure to the disease, its
+consequences and the benefit of the policies put in place by both the federal
+and provincial governments. The other parts of the questionnaire focus on the
+characteristics of the respondents, their immigration history, their educational
+and family background. In short, it is a detailed questionnaire that covers the
+life of immigrants from birth and especially over the past 15 years. The data
+collection lasted three months, from August to November 2020. In total, 1,500
+people were surveyed, including around 1,000 immigrants and 500 natives
+throughout Quebec.
 
@@
 
-**1.	Disease exposure variables**: these are measured by six questions that seek to document whether the individual through work or a family member may have been exposed to the disease.
-For example, it is asked if a family member has had symptoms of the disease, or if a member has been hospitalized due to the contraction of the disease.
+**1. Disease exposure variables**: these are measured by six questions that seek
+to document whether the individual through work or a family member may have been
+exposed to the disease. For example, it is asked if a family member has had
+symptoms of the disease, or if a member has been hospitalized due to the
+contraction of the disease.
 
-**2.	Variables on the family consequences of the disease**: we measure three dimensions of consequences.
-The first group of measures assess the consequences on accessing ressources by the respondent or any member of his family.
-The second group of measure assess whether the respondent or any family member have difficulties to perform usual duties toward any member of the family.
-Finally, the last group of measures measure the consequences of the pandemic on employment and income of any family member.
+**2. Variables on the family consequences of the disease**: we measure three
+dimensions of consequences. The first group of measures assess the consequences
+on accessing ressources by the respondent or any member of his family. The
+second group of measure assess whether the respondent or any family member have
+difficulties to perform usual duties toward any member of the family. Finally,
+the last group of measures measure the consequences of the pandemic on
+employment and income of any family member.
 
-**3.	Variables on government assistance**: Finally, government support is measured by the following two questions that assessed whether the respondent or any family member has benefited of the temporary assistance for workers from the Government of Quebec (PATT) or of the Canadian Federal Government Emergency Benefit (CEP).
-The table XXX in the apendix presents the full set of indicators of the dependents variables
+**3. Variables on government assistance**: Finally, government support is
+measured by the following two questions that assessed whether the respondent or
+any family member has benefited of the temporary assistance for workers from the
+Government of Quebec (PATT) or of the Canadian Federal Government Emergency
+Benefit (CEP). The table XXX in the apendix presents the full set of indicators
+of the dependents variables
 
 ### Main independent variable
-The main independent variable is immigrant status coupled with visible minority status.
-This variable takes the following modalities: 1 if the respondent is a non-racialized native, 2 if he is a non-racialized immigrant, 3 if he is a racialized immigrant, and 4 if the respondent is a racialized native or an immigrant who has not answered the visible minority question.
-The other variables in the analysis are: age, sex, length of residence, minority status, size of social network.
+
+The main independent variable is immigrant status coupled with visible minority
+status. This variable takes the following modalities: 1 if the respondent is a
+non-racialized native, 2 if he is a non-racialized immigrant, 3 if he is a
+racialized immigrant, and 4 if the respondent is a racialized native or an
+immigrant who has not answered the visible minority question. The other
+variables in the analysis are: age, sex, length of residence, minority status,
+size of social network.
 
@@
 
-We use two main methods of analysis, a descriptive method and an explanatory method.
-In the descriptive method, we consider each of the independent variables in isolation and describe how they vary depending on the main independent variable.
-The explanatory methods goes further by assessing whether the effect remain when controlling for other covariates.
-Thus, by controlling for the other explanatory variables, we determine the effect of immigrant status on exposure to the pandemic, on its consequences and on the fact of benefiting or not of government assistance.
-In a second series of modelling, we assess whether the effect that we observed varies by the immigrant's length of residence or by marital status.
-We use either a logistic regression on the outcome variables that are binary and an ordonned logistic regression for the outcome that are measure with a 5-point Lickert scale.
+We use two main methods of analysis, a descriptive method and an explanatory
+method. In the descriptive method, we consider each of the independent variables
+in isolation and describe how they vary depending on the main independent
+variable. The explanatory methods goes further by assessing whether the effect
+remain when controlling for other covariates. Thus, by controlling for the other
+explanatory variables, we determine the effect of immigrant status on exposure
+to the pandemic, on its consequences and on the fact of benefiting or not of
+government assistance. In a second series of modelling, we assess whether the
+effect that we observed varies by the immigrant's length of residence or by
+marital status. We use either a logistic regression on the outcome variables
+that are binary and an ordonned logistic regression for the outcome that are
+measure with a 5-point Lickert scale.
 <!-- PUT THIS AS A LIMITATION IN THE CONCLUSION
@@
 
-# 3. Results
-
-## 3.1. Sample distribution
-
-## 3.2. Exposure by immigration status
-
-Figure 1 presents the exposure to the pandemy accroding to the immigrants and minority visible status.
-For all the six measures of exposure, it appears that the majority group is the most exposed to the COVID-19 in all instance.
-The big difference between the groups is about the exposure trhough working outside.
-Around 50% of the member of this group or a family member work outside the home at the time of the survey.
-This percentage is only 30% for the racialised immigrants and 25% for the non-racialised immigrants.
-
-![Figure 1.1: Exposure](Figures/Exposition1.png){width=60%}
-
-![Figure 1.2: Explicative model](Figures/Exposure.jpeg)
-
-## 3.3. Consequences by immigrations status
-
-The Figure 2 presents the consequences of the COVID-19 for a family member to access to ressources.
-It shows that in all accounts but one (getting medecine), the racialised immigrants express more diffciult to assess ressources.
-Around 20% of them stated having difficulty to get health care compare, around 10% for obtaining food and 7% for obtaining family insurrance.
-Comparatively, these proportion are respectively 13%, 7% and 4% for people and thier family member in the majority group and 19%, 3% and 4% respectively among the racialised immigrants.
-
-![Figure 3.1: Consequences on accessing ressources](Figures/Famille1.png){width=80%}
-
-![Figure 3.2: Explicative model](Figures/Ressources.jpeg)
-
-## 3.4. Consequences on performing usual duties
-
-The Figure 4 compares the response from the three groups on performing usual duties.
-It appears that for all groups the situation caused by the COVID-19 has inexpectedly improved the situation in various way, irrespective of the immigrants status of the respondents.
-
-![Figure 4.1: Consequences on performing usual duties](Figures/Entente1.png)
-
-![Figure 4.2: Explicative model](Figures/Abilities.jpeg)
-
-## 3.5. Consequences on employment by immigration status
-
-The consequences on employment are mixed, with the severe consequences more prevalent among the racialised immigrants.
-The Figure 3 shows for instance that close to 45% of the racialised immigrants have stated  that their family income has decreased compare to just 33% of the majority group and 37% of the non-racialed immigrants.
-In addition, 11% of the racialised immigrants reveals having loss their job compared to 9% among the non-racialised immigrants and 2% among members of the majority group.
-At the opposite, the members of the majority groups are more likely to report a reduction of the working hours or for stopping momentarily their work.
-
-![Figure 5.1: Consequences on employment and income](Figures/Emploi1.png)
-
-![Figure 5.2: Explicative model](Figures/Employment.jpeg)
-
-## 3.7 Help received by immigrations status
-
-Finally, we compared the different group on benefitting of the government assistance.
-It appears that racialised immigrants with 35% of the case have expressed having received the Canada Emergency Response Benefit wheras this proportion is 31% for the non-racialised immigrants and 25% for the member of the majority group.
-For the province Temporary Aid for Workers Program, around 5% of the respondents have received it.
-
-![](Figures/Aide1.png)
-
-![Figure 3.2: Explicative model](Figures/Government.jpeg)
-
-# Conclusion
-
-The preliminary results of this study show that the consequences of the pandemic on the Quebec population differ according to the population groups, but also according to the type of consequences.
-
 \newpage
-
-# Références
-
-http://www.oecd.org/coronavirus/policy-responses/what-is-the-impact-of-the-covid-19-pandemic-on-immigrants-and-their-children-e7cbb7de/
-
-Guttmann A, F.
-et al.
-(2020), COVID-19 in Immigrants, Refugees and Other Newcomers in Ontario: Characteristics of Those Tested and Those Confirmed Positive, as of June 13, 2020, ICES, http://www.ices.on.ca.
-
-Hu, Y.
-(2020), “Intersecting ethnic and native–migrant inequalities in the economic impact of the COVID-19 pandemic in the UK”, Research in Social Stratification and Mobility, Vol.
-68, p.
-100528, http://dx.doi.org/10.1016/j.rssm.2020.100528.
-
-Mcginnity, F.
-and G.
-Kingston (2017), “An Irish Welcome?
-Changing Irish Attitudes to Immigrants and Immigration: The Role of Recession and Immigration”, The Economic and social review, Vol.
-48/3, pp.
-281-304, https://www.researchgate.net/publication/319670153_An_Irish_Welcome_Changing_Irish_Attitudes_to_Immigrants_and_Immigration_The_Role_of_Recession_and_Immigration (accessed on 2 October 2020).
-
-Raisi-Estabragh, Z.
-et al.
-(2020), “Greater risk of severe COVID-19 in Black, Asian and Minority Ethnic populations is not explained by cardiometabolic, socioeconomic or behavioural factors, or by 25(OH)-vitamin D status: study of 1326 cases from the UK Biobank”, Journal of Public Health, Vol.
-42/3, pp.
-451-460, http://dx.doi.org/10.1093/pubmed/fdaa095.
-
-Apea, V.
-et al.
-(2020), Ethnicity and outcomes in patients hospitalised with COVID-19 infection in East London: an observational cohort study, Cold Spring Harbor Laboratory, http://dx.doi.org/10.1101/2020.06.10.20127621.
-
-Bertocchi, G. and A. Dimico (2020), “Covid-19, Race, and Redlining”, IZA Discussion Paper, Vol. 13467, https://papers.ssrn.com/sol3/papers.cfm?abstract_id=3648807 (accessed on 22 September 2020).
-
-Brun, S. and P. Simon (2020), “Inégalités ethno-raciales et coronavirus”, De Facto 19.

### 2023-05-02 Expérimentations (Théa)/2. Expérimentations (Bookdown)/Bookdown/05-results.Rmd
Multiple changes on Bookdown

--- 2023-05-01 Article/Trajipaq_Covid_Draft4.Rmd
@@
 ---
-title: 'COVID-19 Disparities in Quebec (Canada): A Comparative Analysis of Exposure
-  Risks, Consequences, and Access to Resources among Immigrants and Natives'
-author:
-- Vissého Adjiwanou ^[corresponding author - adjiwanou.vissého@uqam.ca]
-- Département de Sociologie, Université du Québec à Montréal (UQAM)
-- Georges NGalé Tchango (Département de démographie, Université de Montréal)
-- Solène Lardoux (Département de démographie, Université de Montréal)
-- Nima Zahedinameghi (Département de Sociologie, UQAM)
-date: "`r format(Sys.time(), '%d %B %Y')`"
-output:
-  word_document:
-    keep_md: yes
-  html_document:
-    toc: yes
-    toc_depth: 3
-  template: svm-latex-ms2.tex
-  pdf_document: default
-  citation_package: natbib
-bibliography:
-- covid_Nima.bib
-- Covid_Canada.bib
-- Covid_Family.bib
+knit: bookdown::render_book
+editor_options: 
+  markdown: 
+    wrap: 80
 ---
-<!--
-# Journal: family issues - SAGE
-https://journals-sagepub-com.proxy.bibliotheques.uqam.ca/author-instructions/JFI
-
-https://fishandwhistle.net/post/2020/getting-started-zotero-better-bibtex-rmarkdown/
-https://www.zotero.org/styles?fields=sociology
--->
-
-\newpage
-
-# Abstract
-
-The immigrant population is one of the most vulnerable in Canada, with a relatively high risk of unemployment, a low social support system, and linguistic barriers.
-However, their experiences with public health interventions during the COVID-19 pandemic have not been adequately explored.
-The purpose of this article is to analyze Covid-19 exposure risk and its family implications for both natives and immigrants in the province of Quebec.
-Additionally, we analyze immigrants' access to government aid and compare it to their native-born counterparts.
-The study consists of survey data collected from 1,000 immigrants and 500 non-immigrants Quebecers between August and November 2020.
-Our descriptive and exploratory analysis reveals significant differences between the two groups regarding exposure risk, family consequences, and access to resources.
-We conclude that the current pandemic exacerbates health disparities in the province, with particularly dire repercussions for immigrants’ families.
-We also discuss several socio-cultural strategies to mitigate inequality and recommend policies to address the needs of vulnerable populations.
-
-**Keywords**: Immigrants, exposure risk, government aid, health disparities, COVID-19, family disruptions
-
-<!--
-
-Using  script in Rmarkdown
-
-http://zevross.com/blog/2014/07/09/making-use-of-external-r-code-in-knitr-and-r-markdown/
-
-Managing my Zotero
-
-https://forums.zotero.org/discussion/35712/removing-pdfs-to-free-up-space
-
-https://www.zotero.org/support/sync
-
-https://www.zotero.org/support/preferences/sync#file_syncing
-
-https://www.zotero.org/support/sync#webdav
-
-https://www.zotero.org/support/preferences/advanced#linked_attachment_base_directory
-
-http://zotfile.com/
-
-# Abstract
-
-A ce jour, la pandémie du coronavirus a touché près de 100000 personnes au Québec et entrainé plus de 6000 décès.
-Cela fait de la province l'une des plus touchées au Canada et l'une des régions les plus affectées au monde.
-Bien que la majorité de ces décès ait eu lieu dans les maisons pour personnes âgées, et donc a touché plus cette tranche de la population, son influence globale sur les individus n'est pas à négliger.
-En effet, outre les conséquences directes de la pandémie, les mesures prises par les différents gouvernements à travers le monde ne sont pas sans conséquences sur les familles.
-Il est démontré que ses conséquences affectent négativement en priorité les couches les plus défavorisées de la population.
-Les immigrants se concentrent dans bien des cas dans cette couche de la population.
-Cette étude vise ainsi à analyser l'influence de la pandémie sur les immigrants et leurs enfants au Québec, à examiner les facteurs socio-culturels qui permettent d'atténuer son impact négatif.
-
-To date, the coronavirus pandemic has affected nearly 100,000 people in Quebec and caused more than 6,000 deaths.
-This makes the province one of the most affected in Canada and one of the most affected regions in the world.
-Although the majority of these deaths took place in homes for the elderly, and therefore affected this segment of the population more, its overall influence on individuals is not to be overlooked.
-In fact, in addition to the direct consequences of the pandemic, the measures taken by various governments around the world are not without consequences for families.
-It has been shown that its consequences primarily affect the most disadvantaged segments of the population negatively.
-Immigrants are in many cases concentrated in this segment of the population.
-This study thus aims to analyze the influence of the pandemic on immigrants and their children in Quebec, to examine the socio-cultural factors that make it possible to mitigate its negative impact.
-
-À soumettre au Canadian Population Studies
-
-Autre projet possible: intégration des immigrants et covid19 au Québec
-
-**Keywords** :
-
--->
-
-\newpage
-
-# Introduction / background
-
-The COVID-19 pandemic has highlighted the vast ethnic, social, economic, and health disparities in high-income countries, including Canada.
-One of the most vulnerable segments of these affluent societies is the immigrant population, that has become widely diversified with more and more people coming from developing countries, and for whom integration into their new society may be more difficult.
-The coronavirus pandemic brings a new challenge for these immigrant populations, directly or indirectly through the tensions between the ethnic groups that it helps to generate [@coates2020Covid19].
-These immigrants face language barriers, low social capital, housing insecurity, and severe poverty, rendering them particularly susceptible to the effects of the current pandemic.
-Immigrants are also more likely to be employed as essential workers where social distancing, sanitary precautions, and self-isolation, measures put in place to mitigate the pandemy, are challenging [@reid2021Migrant; @yi2021Health; @zhang2021COVID19].
-More importantly, migrants may encounter various forms of discrimination, leading to social exclusion and restricted access to healthcare and social services [@spiritus-beerden2021Mental].
-It is therefore necessary to understand how this pandemic affects these immigrant populations differently and the native populations in Quebec.
-In doing so, this article will reveal a better understanding of how the current pandemic aggravates existing inequalities and challenges in the province for immigrants and their descendants.
-
-<!-- The province of Quebec has one of the highest concentrations of immigrants in Canada, including permanent residents, temporary residents, and refugees [@governmentofcanada2017Daily; @immigration2005Recent].
-In 2016, 1.2 million immigrants were living in Quebec, an increase from 8% in 1990 to 13.7% of the population in 2016 (Institut de la statistique du Québec, 2020).
-<!--As of 2019, the increase in net new residents (60,700 versus 42,200 in 2018) and lower interprovincial migration losses (-1,500 versus -6,100 in 2018) contributed to Quebec's population growth.
--->Since the start of the pandemic, most of the confirmed cases were identified in Montreal and areas with high immigrant populations [@miconi2021Ethnocultural; @shields2020COVID19].
-The coronavirus pandemic has infected nearly 450,000 people in Quebec and has resulted in more than 11,000 deaths (https://www.quebec.ca/sante/problemes-de-sante/a-z/coronavirus-2019/situation-coronavirus-quebec, consulted on octobre 29, 2021).
-This makes it the worst affected province in Canada and one of the worst affected regions in the world during the first wave of the pandemic.
-Although the majority of these deaths took place in elderly's homes, and therefore affected more this segment of the population, its overall influence on individuals is not to be overlooked.
-It has been shown that, on the one hand, immigrants and their families have a higher risk of being exposed to the pathogen, and, at the same time, they are more vulnerable to racism and discrimination and face barriers to accessing social support [@2021CRARR] and information, especially  for those who didn't speak or understand the two official langages.
-<!--In addition to the direct consequences of the pandemic, the measures taken by various governments around the world are also not without consequences on families.
--->Several studies have thus shown excess mortality in ethnocultural minority populations [@bertocchi2020Covid19], higher unemployment, greater loss of income and negative consequences for children and families in several countries.
-This differentiated effect can be explained in different ways.
-On the one hand, in terms of the direct consequences of the pandemic, the situation of precariousness and greater poverty among immigrants exposes them more to the more harmful consequences of the disease on their state of health.
-Likewise, immigrants often work in jobs where social distancing is difficult to enforce.
-On the other hand, indirectly, the solutions to the crisis affect immigrants and natives differently.
-The complete shutdown of tye economy with its set of adjustments has greatly impeded the employability of immigrants, especially newcomers who have not yet grasped the workings of society and who cannot benefit from state subsidies.
-But, it is also shown that these effects are not uniform on the whole of the immigrant population, itself, very diverse.
-
-Despite the growing focus on the differential impacts of the COVID-19 on immigrants and non-immigrants segment of the population, there is still very little knowledge about how the pandemic and its related public policy impact the lives of immigrants and their families.
-The existing health datasets (largely used to assess this effect) are not necessarily representative of the immigrant population and do not offer an accurate picture of the complex social status and health outcomes among this group [@choi2021Studying; @gagnon2021Immigration; @khan2015Visible].
-In most cases, these datasets lack data on the residence length of immigrants and combine newcomers with long-term settlers [@edge2013Discrimination; @hayward2021Clinical].
-As a result, they may not allow a detailed assessment of the nature and extent of risk exposure in sub-segments of populations, as well as whether interventions are reaching those most at risk.
-In addition, the family-level impacts of the pandemic are still poorly understood.
-While a body of research has explored the association between the pandemic and its potential effects on intimate partner violence and marital dissolution [@bhatnagar2021Navigating; @koshan2020COVID19; @morrison2021Chapter; @prime2020Risk], few studies have examined negative familial repercussions in the context of immigration status.
-There is ample evidence that sustained family relationships greatly enhance immigrants' lives on several fronts, including fostering economic prosperity [@marger2001Sociala], psychological well-being [@beiser2002Povertya], social network development [@evansluong2017Familya], and educational success for children [@munroe-blum1989Immigranta].
-The family is, therefore, a key component of long-term social mobility and economic success of immigrants.
-
-<!--Faced with the pandemic, the governments of Canada and Quebec have implemented a number of measures to reduce its impact.
-These policies can be broken down at two levels, direct support to families, including the Canadian Emergency Benefit, which can provide employees and self-employed workers affected by Covid-19 with a financial benefit of up to $ 2,000.
-On the other hand, the government supports many companies to keep their employees or supports them directly when the revenues from their activities have fallen considerably.
-While these measures allow families to cope better with the crisis, the fact remains that the changes brought about by this crisis can affect other dimensions of life.
-How these measures make a difference in the face of the pandemic remains to be seen.
--->Based on an original biographic survey collected around the first wave of the pandemic in august-november 2020 on immigrants and the native population of Québec, this study mainly aims to understand the family exposure, consequences, and responses of the pandemic on immigrant and native Quebec populations.
-We assess a whole range of different measures of each outcome.
-We divided the consequences in three dimensions, 1) access to ressources, 2) ability to perform duties and 3) employment and income.
-We seek also to analyse in what ways these outcomes differ according to the immigrants length of residence.
-By doing so, this study provide a clear and global picture of the situation of different population groups in Quebec in the face of the COVID-19.
-<!-- To respond to these questions, we will use data collected around the first wave of the pandemic in august-november 2020.
-Our study aims to fill the gap by analyzing group-specific differences in exposure risk, family consequences, and access to resources among immigrants and Quebecers.
-We also seek to shed light on several socio-cultural strategies to mitigate inequality and recommend relevant policy solutions.
-The detailed findings will help policymakers address health inequalities and improve equity in Quebec while developing adequate measures and programs.
--->
-
-<!-- Introduction  -->
-
-<!-- Like the rest of Canada, Quebec has experienced a substantial increase in its immigration in recent years.
-The immigrant population in the province thus increased from 8% in the 1990s to 13.7% of the population, or 1.2 million individuals in 2016 (Institut de la statistique du Québec, 2020).
-This immigration is also unique in that it has become widely diversified with more and more people coming from developing countries, and for whom integration into their new society may be more difficult.
-The coronavirus pandemic brings a new challenge for these immigrant populations, directly or indirectly through the tensions between the ethnic groups that it helps to generate [@coates2020Covid19].
-It is therefore necessary to understand how this pandemic affects these immigrant populations differently and the native populations in Quebec.
--->
-
-<!-- The coronavirus pandemic has affected nearly 100,000 people in Quebec and has resulted in more than 6,000 deaths.
-This makes it the worst affected province in Canada and one of the worst affected regions in the world.
-Although the majority of these deaths took place in homes for the elderly, and therefore affected this segment of the population more, its overall influence on individuals is not to be overlooked.
-In fact, in addition to the direct consequences of the pandemic, the measures taken by various governments around the world are also not without consequences on families.
-Several studies have thus shown excess mortality in ethnocultural minority populations [@bertocchi2020Covid19], higher unemployment, greater loss of income and negative consequences for children and families in several countries.
--->
-
-<!-- This differentiated effect can be explained in different ways.
-On the one hand, in terms of the direct consequences of the pandemic, the situation of precariousness and greater poverty among immigrants exposes them more to the more harmful consequences of the disease on their state of health.
-Likewise, immigrants often work in jobs where social distancing is difficult to enforce.
-On the other hand, indirectly, the solutions to the crisis affect immigrants and natives differently.
-The complete shutdown of the company with its set of adjustments has greatly impeded the employability of immigrants, especially newcomers who have not yet grasped the workings of society and who cannot benefit from state subsidies.
-But, it is also shown that these effects are not uniform on the whole of the immigrant population, itself, very diverse.
--->
-
-<!-- Faced with the pandemic, the governments of Canada and Quebec have implemented a number of measures to reduce the impact of the pandemic.
-These policies can be broken down at two levels, direct support to families, including the Canadian Emergency Benefit, which can provide employees and self-employed workers affected by Covid-19 with a financial benefit of up to $ 2,000.
-On the other hand, the government supports many companies to keep their employees or supports them directly when the revenues from their activities have fallen considerably.
-While these measures allow families to cope better with the crisis, the fact remains that the changes brought about by this crisis can affect other dimensions of life.
-How these measures make a difference in the face of the pandemic remains to be seen.
--->
-
-<!-- This study mainly aims to understand the family consequences of the pandemic on immigrant and native Quebec populations.
-Therefore, we will seek to analyse in what ways its consequences differ according to the region of origin of immigrants, length of residence, age and sex.
--->
-
-# Current body of literature and knowledge gaps
-
-The literature on the racial and ethnic health disparities during the COVID-19 crisis can be divided into two main categories: studies addressing the socio-economic determinants of health and those examining the delivery of health interventions and policies.
-The first category is primarily comprised of research that focuses on the link between socio-economic status (SES) of high-risk minority groups and the differential impact of the pandemic [@drefahl2020Populationbased; @greenaway2020COVID19; @lopez2021Racial; @pacheco2020Job; @sundaram2020Individual].
-For instance, an inter-country study has found that Middle Eastern, African, and Nordic immigrants in Sweden have a higher risk of being hospitalized and even dying from COVID-19 when compared to the Swedish born population.
-In the same study, the SES, proportion of working-age households, and neighborhood density were significantly associated with the mortality rate [@rostila2021Disparities].
-This trend of research demonstrates that racial and ethnic minorities are more likely to experience stress factors, not only due to the hazards of infection but also because of economic and social marginalization [@alahmad2020COVID19; @choi2020Buffering; @endale2020COVID19; @fitzpatrick2020Fear; @serafini2021Psychological].
-The intersectionality of race, sex, and health conditions has also been explored in the Canadian context of the pandemic.
-@ingen2021Neighbourhoodlevel demonstrated that Ontarian neighborhoods with immigrant and racialized populations have a disproportionate share of COVID-19 infections.
-In addition, areas with high SES report a lesser number of COVID cases than low SES ones [@ingen2021Neighbourhoodlevel].
-Another study conducted in Quebec confirms the strong association between immigrants' backgrounds, SES, and COVID-19 related psychological stress [@miconi2021Ethnocultural].
-The most recent reviews on the topic suggest that immigrants disproportionately live in crowded and inadequate housing and are employed in essential industries, leaving them with severe income insecurities and poor physiological/psychological well-being [@chew2020Clinical; @davis2020Impact].
-Studies on SES and health disparities have emphasized that this situation is further complicated for immigrant women, constrained by gender and afflicted by a significant power imbalance in the household and workplace [@alarcao2021Intersections; @esegbona-adeigbe2020COVID19; @jehn2021COVID19; @mo2020Differential; @nardon2021Skilled; @sabri2020Effect; @st-denis2020Sociodemographic].
-According to Lightman, immigrants and women of color are overrepresented in long-term care facilities in Canada, where COVID infection rates have been exceptionally high.
-They are also subject to exclusionary social and economic practices [@lightman2021Caring].
-
-The second category of the literature is primarily comprised of policy analyses, research on public health communications, examining the efficacy of COVID initiatives, the optimal allocation of government aids, and ethical issues associated with applying new health interventions to vulnerable populations and their communities [@aragona2020Negative; @cross2021Coronavirus; @desai2020COVID19; @doyle2020Migrant; @gilman2020Modelling; @hargreaves2020Targeting; @langellier2020Policy; @nezafatmaldonado2020Engaging; @silverman2020Ethics; @wilson2020Ethics].
-Some of the questions that this body of research addresses include: how best to respond to COVID-19 in immigrant communities?
-How do public health policies impact the socio-economic status and psychological well-being of different sub-populations?
-What is the structure of the planning and implementation of government aid programs during the current crisis?
-And, how can we design public health communication strategies in a way that is culturally sensitive and inclusive?
-
-There are several knowledge gaps in the existing literature.
-Most of the cited studies do not include household data and family structure of immigrants as variables that affect exposure risks and health outcomes of COVID-19.
-Consequently, the relationship between immigrant status and COVID-19 is not accounted for by household size, and family characteristics (i.e., income, education, religion).
-A key reason is that most of the health data used in such research projects are acquired from national registers or administrative databases, which provide little information on immigration status and kinship.
-However, as the study of Thomeer and her colleagues theorizes, the extent of adverse social and economic impacts of COVID-19 is significantly conditional on the composition of an immigrant's family [@thomeer2020How].
-Households that include a spouse/partner, children, or multigenerational family members most likely experience a different level of social, psychological, and medical vulnerability than single or nuclear ones.
-Therefore, studies of immigrant populations should account for the breakdown of family configuration to better understand the mechanisms linking social determinants and health outcomes.
-Another knowledge gap concerns the current state of Canadian studies of COVID-19 and its public health crisis.
-There is a dearth of literature on the challenges faced by immigrants and their families in the country.
-Although the burden of COVID-19 infection is disproportionately higher among these marginalized groups (St-Denis, 2020; Sundaram et al., 2020), research on the issues they are experiencing is sparse and often limited to one aspect.
-<!--Likewise, there is limited focus on the effectiveness of COVID-19 crisis management for immigrants at the policy level.
-In response to the mounting cases of COVID, the federal government and provinces and territories of Canada have launched multi-pronged public health initiatives to protect society from the epidemic (Migone, 2020).
-However, many open questions remain unanswered, including, do interventions efficiently reach the migrant groups?
-What are the likely effects of each program on different immigrant communities?
-How can an intervention be tailored to the needs of a particular migrant group and their families?
--->
-
-# Theoretical framework 
-Family as a unit of analysis can be particularly fruitful when studying health inequalities in the context of a societal-wide crisis.
-Through the lens of the COVID-19 pandemic, the following points illustrate why exploring family dynamics provides additional insight into the underlying mechanisms of health disparities.
-First, the family is the most fundamental unit of social organization.
-It facilitates access to resources and the transmission of social and psychological capital from one generation to the next.
-Therefore, a better analysis of how crises and adverse life events affect people requires a better understanding of how families relate to each other and interact with external forces.
-Second, during public health emergencies, families may be thrown into different structures due to unprecedented financial, mental, and physical challenges.
-The risk of family dissolution, for example, can be heightened when members must spend unusually long periods together at home and have limited opportunities to engage with friends and the wider community.
-Conversely, families can develop closer bonds through self-sacrifice, shared responsibilities, and increased care and support.
-As the dependency status of seniors increases, the elderlies are also more likely to be integrated into formerly nucleus families.
-We posit that the extent to which a family can adapt its functions to meet new challenges is a predictor of the well-being of its members and the resilience of its broader population.
-Third, the family is a crucial link to social institutions, such as the health care system, schools, and government agencies whose decisions have significant downstream impacts.
-Families are viewed by many public-level entities as a reference when allocating resources, enforcing new standards, and responding to emerging needs and demands.
-From this perspective, the family operates at the nexus of individual and societal structures, embedding the tripple effects of public policy in its most consequential ways.
-Finally, the family has a crucial role to play in determining the life course of its members.
-As COVID-19 elevates the importance of familial bonds, the post-pandemic period might be marked by an increased influence of families on the lives of their members and the social structures they inhabit.
-
-The above conceptualization echoes several other approaches to studying long-term health inequalities.
-For instance, life course theory emphasizes the significance of family ties and social environment as they structure prospects for individuals across their life span (Elder, 1977; Elder Jr., 1998; Macmillan, 2005; Settersten et al., 2020).
-Likewise, social capital theory and the stress-buffering hypothesis consider family systems a crucial factor in health and well-being (Choi and Jun, 2020; Cohen and McKay, 1984; Cohen and Wills, 1985; Harpham et al., 2004; Kawachi and Berkman, 2001; Lavee et al., 1987; Mandelbaum et al., 2020).
-They show that the family can offer tangible (e.g., food, shelter, and money) and intangible resources (e.g., support, social status, psychological comfort, and institutional advantages), protecting its members from adverse life events.
-Similarly, the social-ecological model and the theory of fundamental causes highlight how the SES, familial and extrafamilial factors pattern people's life trajectories and health outcomes (Barkan and Rocque, 2018; Clouston and Link, 2021; Figge et al., 2018; Kim, 2020; Libman et al., 2012; Ungar, 2015).
-However, all these approaches are ostensibly focused on life periods where the general social context is relatively stable.
-In this article, we present our conceptualization of the family in times of unprecedented social upheaval.
-Specifically, our framework seeks to characterize family dynamics as a suitable lens to analyze health inequalities in the face of a virulent public health crisis.
-
-# Methodology
-
-## Data
-
-The data for this study come from an original collection of biographical data on immigrants and the native population of Quebec to analyse their participation in Quebec society in various fields (social, economic, cultural, etc.).
-The immigrant population is limited to recent immigration that has taken place over the past 15 years.
-This survey has several components that retrace the life course of individuals over a long period of time.
-The pandemic component uses a reduced version of the [harmonized questionnaire on exposure and the impact of Covid19 on the family (CEFIS)](https://www.phenxtoolkit.org/toolkit_content/PDF/CPTS_CEFIS_FullSurveyEng.pdf).
-The questions are divided into three parts, on exposure to the disease, its consequences and the benefit of the policies put in place by both the federal and provincial governments.
-The other parts of the questionnaire focus on the characteristics of the respondents, their immigration history, their educational and family background.
-In short, it is a detailed questionnaire that covers the life of immigrants from birth and especially over the past 15 years.
-The data collection lasted three months, from August to November 2020. In total, 1,500 people were surveyed, including around 1,000 immigrants and 500 natives throughout Quebec.
-
-### Dependent variables
-
-We distinguish three groups of dependent variables:
-
-**1.	Disease exposure variables**: these are measured by six questions that seek to document whether the individual through work or a family member may have been exposed to the disease.
-For example, it is asked if a family member has had symptoms of the disease, or if a member has been hospitalized due to the contraction of the disease.
-
-**2.	Variables on the family consequences of the disease**: we measure three dimensions of consequences.
-The first group of measures assess the consequences on accessing ressources by the respondent or any member of his family.
-The second group of measure assess whether the respondent or any family member have difficulties to perform usual duties toward any member of the family.
-Finally, the last group of measures measure the consequences of the pandemic on employment and income of any family member.
-
-**3.	Variables on government assistance**: Finally, government support is measured by the following two questions that assessed whether the respondent or any family member has benefited of the temporary assistance for workers from the Government of Quebec (PATT) or of the Canadian Federal Government Emergency Benefit (CEP).
-The table XXX in the apendix presents the full set of indicators of the dependents variables
-
-### Main independent variable
-The main independent variable is immigrant status coupled with visible minority status.
-This variable takes the following modalities: 1 if the respondent is a non-racialized native, 2 if he is a non-racialized immigrant, 3 if he is a racialized immigrant, and 4 if the respondent is a racialized native or an immigrant who has not answered the visible minority question.
-The other variables in the analysis are: age, sex, length of residence, minority status, size of social network.
-
-## Analysis methods
-
-We use two main methods of analysis, a descriptive method and an explanatory method.
-In the descriptive method, we consider each of the independent variables in isolation and describe how they vary depending on the main independent variable.
-The explanatory methods goes further by assessing whether the effect remain when controlling for other covariates.
-Thus, by controlling for the other explanatory variables, we determine the effect of immigrant status on exposure to the pandemic, on its consequences and on the fact of benefiting or not of government assistance.
-In a second series of modelling, we assess whether the effect that we observed varies by the immigrant's length of residence or by marital status.
-We use either a logistic regression on the outcome variables that are binary and an ordonned logistic regression for the outcome that are measure with a 5-point Lickert scale.
-<!-- PUT THIS AS A LIMITATION IN THE CONCLUSION
-To be able to benefit from government benefits, one must meet a number of conditions, including the loss of job due to Covid-19.
-As a result, the beneficiaries are not the general population but those who have lost their jobs and this may itself be due to immigrant status.
-Thus, for model 3, we will use Heckman's models by first estimating the fact of having lost or not a job and then, the fact of receiving or not the government benefit.
--->
 
@@
 
-Figure 1 presents the exposure to the pandemy accroding to the immigrants and minority visible status.
-For all the six measures of exposure, it appears that the majority group is the most exposed to the COVID-19 in all instance.
-The big difference between the groups is about the exposure trhough working outside.
-Around 50% of the member of this group or a family member work outside the home at the time of the survey.
-This percentage is only 30% for the racialised immigrants and 25% for the non-racialised immigrants.
+Figure 1 presents the exposure to the pandemy accroding to the immigrants and
+minority visible status. For all the six measures of exposure, it appears that
+the majority group is the most exposed to the COVID-19 in all instance. The big
+difference between the groups is about the exposure trhough working outside.
+Around 50% of the member of this group or a family member work outside the home
+at the time of the survey. This percentage is only 30% for the racialised
+immigrants and 25% for the non-racialised immigrants.
 
-![Figure 1.1: Exposure](Figures/Exposition1.png){width=60%}
+![Figure 1.1: Exposure](Figures/Exposition1.png){width="60%"}
 
@@
 
-The Figure 2 presents the consequences of the COVID-19 for a family member to access to ressources.
-It shows that in all accounts but one (getting medecine), the racialised immigrants express more diffciult to assess ressources.
-Around 20% of them stated having difficulty to get health care compare, around 10% for obtaining food and 7% for obtaining family insurrance.
-Comparatively, these proportion are respectively 13%, 7% and 4% for people and thier family member in the majority group and 19%, 3% and 4% respectively among the racialised immigrants.
+The Figure 2 presents the consequences of the COVID-19 for a family member to
+access to ressources. It shows that in all accounts but one (getting medecine),
+the racialised immigrants express more diffciult to assess ressources. Around
+20% of them stated having difficulty to get health care compare, around 10% for
+obtaining food and 7% for obtaining family insurrance. Comparatively, these
+proportion are respectively 13%, 7% and 4% for people and thier family member in
+the majority group and 19%, 3% and 4% respectively among the racialised
+immigrants.
 
-![Figure 3.1: Consequences on accessing ressources](Figures/Famille1.png){width=80%}
+![Figure 3.1: Consequences on accessing
+ressources](Figures/Famille1.png){width="80%"}
 
@@
 
-The Figure 4 compares the response from the three groups on performing usual duties.
-It appears that for all groups the situation caused by the COVID-19 has inexpectedly improved the situation in various way, irrespective of the immigrants status of the respondents.
+The Figure 4 compares the response from the three groups on performing usual
+duties. It appears that for all groups the situation caused by the COVID-19 has
+inexpectedly improved the situation in various way, irrespective of the
+immigrants status of the respondents.
 
@@
 
-The consequences on employment are mixed, with the severe consequences more prevalent among the racialised immigrants.
-The Figure 3 shows for instance that close to 45% of the racialised immigrants have stated  that their family income has decreased compare to just 33% of the majority group and 37% of the non-racialed immigrants.
-In addition, 11% of the racialised immigrants reveals having loss their job compared to 9% among the non-racialised immigrants and 2% among members of the majority group.
-At the opposite, the members of the majority groups are more likely to report a reduction of the working hours or for stopping momentarily their work.
+The consequences on employment are mixed, with the severe consequences more
+prevalent among the racialised immigrants. The Figure 3 shows for instance that
+close to 45% of the racialised immigrants have stated that their family income
+has decreased compare to just 33% of the majority group and 37% of the
+non-racialed immigrants. In addition, 11% of the racialised immigrants reveals
+having loss their job compared to 9% among the non-racialised immigrants and 2%
+among members of the majority group. At the opposite, the members of the
+majority groups are more likely to report a reduction of the working hours or
+for stopping momentarily their work.
 
@@
 
-Finally, we compared the different group on benefitting of the government assistance.
-It appears that racialised immigrants with 35% of the case have expressed having received the Canada Emergency Response Benefit wheras this proportion is 31% for the non-racialised immigrants and 25% for the member of the majority group.
-For the province Temporary Aid for Workers Program, around 5% of the respondents have received it.
+Finally, we compared the different group on benefitting of the government
+assistance. It appears that racialised immigrants with 35% of the case have
+expressed having received the Canada Emergency Response Benefit wheras this
+proportion is 31% for the non-racialised immigrants and 25% for the member of
+the majority group. For the province Temporary Aid for Workers Program, around
+5% of the respondents have received it.
 
@@
 
-# Conclusion
-
-The preliminary results of this study show that the consequences of the pandemic on the Quebec population differ according to the population groups, but also according to the type of consequences.
-
 \newpage
-
-# Références
-
-http://www.oecd.org/coronavirus/policy-responses/what-is-the-impact-of-the-covid-19-pandemic-on-immigrants-and-their-children-e7cbb7de/
-
-Guttmann A, F.
-et al.
-(2020), COVID-19 in Immigrants, Refugees and Other Newcomers in Ontario: Characteristics of Those Tested and Those Confirmed Positive, as of June 13, 2020, ICES, http://www.ices.on.ca.
-
-Hu, Y.
-(2020), “Intersecting ethnic and native–migrant inequalities in the economic impact of the COVID-19 pandemic in the UK”, Research in Social Stratification and Mobility, Vol.
-68, p.
-100528, http://dx.doi.org/10.1016/j.rssm.2020.100528.
-
-Mcginnity, F.
-and G.
-Kingston (2017), “An Irish Welcome?
-Changing Irish Attitudes to Immigrants and Immigration: The Role of Recession and Immigration”, The Economic and social review, Vol.
-48/3, pp.
-281-304, https://www.researchgate.net/publication/319670153_An_Irish_Welcome_Changing_Irish_Attitudes_to_Immigrants_and_Immigration_The_Role_of_Recession_and_Immigration (accessed on 2 October 2020).
-
-Raisi-Estabragh, Z.
-et al.
-(2020), “Greater risk of severe COVID-19 in Black, Asian and Minority Ethnic populations is not explained by cardiometabolic, socioeconomic or behavioural factors, or by 25(OH)-vitamin D status: study of 1326 cases from the UK Biobank”, Journal of Public Health, Vol.
-42/3, pp.
-451-460, http://dx.doi.org/10.1093/pubmed/fdaa095.
-
-Apea, V.
-et al.
-(2020), Ethnicity and outcomes in patients hospitalised with COVID-19 infection in East London: an observational cohort study, Cold Spring Harbor Laboratory, http://dx.doi.org/10.1101/2020.06.10.20127621.
-
-Bertocchi, G. and A. Dimico (2020), “Covid-19, Race, and Redlining”, IZA Discussion Paper, Vol. 13467, https://papers.ssrn.com/sol3/papers.cfm?abstract_id=3648807 (accessed on 22 September 2020).
-
-Brun, S. and P. Simon (2020), “Inégalités ethno-raciales et coronavirus”, De Facto 19.

### 2023-05-02 Expérimentations (Théa)/2. Expérimentations (Bookdown)/Bookdown/07-references.Rmd
Multiple changes on Bookdown

---
knit: bookdown::render_book
editor_options: 
  markdown: 
    wrap: 80
---

# Références

<http://www.oecd.org/coronavirus/policy-responses/what-is-the-impact-of-the-covid-19-pandemic-on-immigrants-and-their-children-e7cbb7de/>

Guttmann A, F. et al. (2020), COVID-19 in Immigrants, Refugees and Other
Newcomers in Ontario: Characteristics of Those Tested and Those Confirmed
Positive, as of June 13, 2020, ICES, <http://www.ices.on.ca>.

Hu, Y. (2020), "Intersecting ethnic and native--migrant inequalities in the
economic impact of the COVID-19 pandemic in the UK", Research in Social
Stratification and Mobility, Vol. 68, p. 100528,
<http://dx.doi.org/10.1016/j.rssm.2020.100528>.

Mcginnity, F. and G. Kingston (2017), "An Irish Welcome? Changing Irish
Attitudes to Immigrants and Immigration: The Role of Recession and Immigration",
The Economic and social review, Vol. 48/3, pp. 281-304,
<https://www.researchgate.net/publication/319670153_An_Irish_Welcome_Changing_Irish_Attitudes_to_Immigrants_and_Immigration_The_Role_of_Recession_and_Immigration>
(accessed on 2 October 2020).

Raisi-Estabragh, Z. et al. (2020), "Greater risk of severe COVID-19 in Black,
Asian and Minority Ethnic populations is not explained by cardiometabolic,
socioeconomic or behavioural factors, or by 25(OH)-vitamin D status: study of
1326 cases from the UK Biobank", Journal of Public Health, Vol. 42/3, pp.
451-460, <http://dx.doi.org/10.1093/pubmed/fdaa095>.

Apea, V. et al. (2020), Ethnicity and outcomes in patients hospitalised with
COVID-19 infection in East London: an observational cohort study, Cold Spring
Harbor Laboratory, <http://dx.doi.org/10.1101/2020.06.10.20127621>.

Bertocchi, G. and A. Dimico (2020), "Covid-19, Race, and Redlining", IZA
Discussion Paper, Vol. 13467,
<https://papers.ssrn.com/sol3/papers.cfm?abstract_id=3648807> (accessed on 22
September 2020).

Brun, S. and P. Simon (2020), "Inégalités ethno-raciales et coronavirus", De
Facto 19.

### 2023-05-02 Expérimentations (Théa)/2. Expérimentations (Bookdown)/Bookdown/Bookdown.Rmd
Multiple changes on Bookdown

---
title: "Réseaux sociaux et sentiment d’appartenance chez les immigrants établis au Québec"
author: "Vincent Bertrand-Huneaut, Jeanne Craig-Larouche, Théa Carlier, Noura Kone,  Antoine Martin"
date: "`r format(Sys.Date(), '%B %e, %Y')`"
present_to: "Vissého Adjiwanou"
type: "Rapport de recherche"
knit: bookdown::render_book
subtitle: Rapport de recherche
groupe_cours: "SOC3380, gr. 40"
site: bookdown::bookdown_site
bibliography: "library.biblatex"
csl: apalike-uqam.csl
editor_options:
  markdown:
    wrap: 80
    fontsize: 12pt
documentclass: article
link-citations: yes
institute: Université de Québec à Montréal
abstract: "Ce rapport de recherche présente les résultats d'une étude mixte visant à comprendre la relation entre l’usage des réseaux sociaux en ligne et l’intégration culturelle des personnes immigrantes de première génération établies au Québec depuis plus d'un an. Son objectif est de saisir comment les discours présents sur les réseaux sociaux en ligne, ainsi que l’entretien de réseaux diversifiés sur des sites comme Facebook, encouragent ou découragent les personnes immigrantes à s’intégrer dans la culture québécoise, et ainsi, à développer un sentiment d'appartenance pour le Québec. Pour répondre à ces questions, nous nous reposons sur la théorie de la fusion culturelle proposée par Stephen M. Croucher et Éric Kramer ainsi que sur les travaux de Friedrich Heckmann et de Wolfgang Bosswick sur l'intégration culturelle. Partant, nous avons conduit et analysé une série de 10 entretiens semi-directifs en parallèle d'une enquête par questionnaire (N = 65) auprès de personnes immigrantes de première génération établies à Montréal depuis plus d'un an. Mots clés : #Immigration #Intégration culturelle #Réseaux sociaux en ligne #Sentiment d'appartenance #Québec"
---

```{r setup, include=FALSE, warning=FALSE}
# Préférence knitr

knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo=FALSE)
```

```{r}
# Importer package
library(dplyr)
library(tidyverse)
library(summarytools)
library(readr)
library(questionr)
library(rcompanion)
library(pander)
```

```{r}
# Set numero graphique
no_graphic <- -1
```

<!--chapter:end:index.Rmd-->

---
knit: bookdown::render_book
editor_options: 
  markdown: 
    wrap: 80
---

# Introduction

Au Québec, la pérennité de l'économie, de la démographie et de la langue
officielle repose en partie sur le renouvellement du flux migratoire
international. Dans ce contexte, l'intégration culturelle des personnes
immigrantes constitue un enjeu historique pour la province, dont le gouvernement
défend depuis les années 1960 la spécificité culturelle et politique au sein de
la fédération du Canada. Si l'intégration des personnes immigrantes est étudiée
depuis les années 1910 aux États-Unis, le récent développement des réseaux
sociaux en ligne offre depuis vingt ans une nouvelle manière d'appréhender ce
phénomène.

Dans l'état actuel de la littérature scientifique, nous savons que les réseaux
sociaux en ligne permettent aux personnes immigrantes de conserver un lien
affectif et culturel avec leur pays d'origine afin de réduire le stress
acculturatif généré par la précarisation liée au processus de migration. Dans le
même temps, les réseaux sociaux en ligne permettent aux personnes immigrantes
d'entretenir et d'accumuler du capital culturel et social mobilisable dans le
pays d'accueil.

Les études précédemment réalisées dans ce domaine reposent sur des séries
d'entretiens semi-directifs, ou sur une enquête par questionnaire. Dans le cas
des enquêtes qualitatives, la nature des échantillons tirés et le type
d'analyses réalisées ne permettent pas de quantifier la relation à l'étude ni de
généraliser les résultats à l'échelle de la population étudiée. Concernant les
enquêtes quantitatives, si l'enquête par questionnaire permet de mettre à
l'épreuve statistique la relation entre l'usage des réseaux sociaux en ligne et
l'intégration culturelle des personnes immigrantes, celle-ci peine à fournir des
explications précises quant aux modalités du phénomène étudié.

Dans ce contexte, nous avons réalisé une étude mixte regroupant 10 entretiens
semi-directifs ainsi qu'une enquête par questionnaire (N = 65) visant à
reproduire en contexte québécois les études proposées par la littérature
scientifique. Notre objectif est de comprendre la relation entre l'usage des
réseaux sociaux en ligne et l'intégration culturelle des personnes immigrantes
établies au Québec. Plus précisément, nous nous intéressons au sentiment
d'appartenance que peuvent développer les personnes immigrantes à travers leur
utilisation des réseaux sociaux en ligne.

\newpage

<!--chapter:end:01-intro.Rmd-->

---
knit: bookdown::render_book
editor_options: 
  markdown: 
    wrap: 80
---

# Revue de littérature

## Intégration culturelle

Pour comprendre ce que signifie le sentiment d'appartenance des personnes
immigrées établies au Québec, il est nécessaire de se pencher sur le lien qui
unit ces dernières à leur pays d'accueil.

Cette relation intéresse les sociologues depuis le début du XIX^e^ siècle. On en
trouve d'abord la trace aux États-Unis, au sein de la première école de Chicago.
Les travaux de William I. Thomas [-@thomasPolishPeasantEurope1918] et Robert E.
Park [-@parkHumanMigrationMarginal1928] contribuent à définir un premier concept
servant à appréhender ce rapport : l'assimilation. Celle-ci se définie comme : «
un processus individuel de convergence des caractéristiques des [personnes
immigrées] vers les caractéristiques moyennes de la société d'accueil »
[@safiPenserIntegrationImmigres2011, p. 150].

Après avoir dominé durant le XXe siècle, le concept d'assimilation est remis en
question. Friedrich Heckmann remarque que celui-ci est normatif, en ce qu'il
prend pour acquis que la culture du pays d'accueil forme un ensemble homogène
[-@heckmannIntegrationIntegrationPolicies2005, p. 12].

S'inspirant de la théorie de l'intégration des systèmes publiée par David
Lockwood en 1964, Hartmut Esser propose d'appréhender ce lien comme une seconde
socialisation possédant quatre dimensions : 1) la capacité des personnes
immigrantes à mobiliser les institutions de leur pays d'accueil ; 2) leur
placement dans le marché du travail, de la vie conjugale, etc. ; 3) leur
capacité à interagir avec d'autres personnes dans la société d'accueil ; et 4)
le sentiment d'appartenance qu'elles développent envers leur pays d'accueil ou
d'autres sous-ensembles lui appartenant [-@esserDoesNewImmigration2004, p.
1131].

Or, cette nouvelle conceptualisation du lien entre les personnes immigrées et
leur pays d'accueil se heurte à la critique de Young Kim, qui avance que
celui-ci ne constitue pas simplement une deuxième socialisation, mais implique
également une désocialisation plus ou moins prononcée vis-à-vis de la diversité
culturelle présente dans le pays d'origine
[-@kimBecomingInterculturalIntegrative2001, p. 53]. Quelques années plus tard,
Stephen M. Croucher et Éric Kramer développent la théorie de la fusion
culturelle. A l'instar des travaux de Kim, la fusion culturelle se comprend
comme « [le] mélange entre l'apprentissage de nouveaux comportements et le
maintien [relatif] d'anciens comportements »
[-@croucherCulturalFusionTheory2017, p. 97]. Néanmoins, Croucher et Kramer
introduisent une nouvelle dimension dans leur concept, en avançant que la fusion
culturelle est simultanée. En d'autres termes, cela signifie que la culture des
personnes immigrantes fusionne avec la diversité culturelle présente dans le
pays d'accueil pour produire une forme altérée de la diversité culturelle
initiale [-@croucherCulturalFusionTheory2017, p. 103].

## Réseaux sociaux en ligne et intégration culturelle

La migration constitue une rupture dans le processus de socialisation car elle
déracine les personnes migrantes de leur environnement. Dans ce contexte, elles
expérimentent ce que des chercheurs en psychologie comme John W. Berry et al.
[-@berryImmigrationAcculturationAdaptation1997] appellent du « stress
acculturatif ». Celui-ci est caractérisé par de possibles dégradations de la
santé comme l'anxiété, la dépression, la confusion identitaire ou encore un
sentiment de marginalisation de la part des personnes établies dans le pays
d'accueil. Selon Berry, ce stress et cette marginalisation (effectifs et/ou
perçus) sont causés par les différences culturelles qui oppose le pays d'origine
au pays d'accueil.

L'étude d'Eric Forbush et de Brooke Foucault-Wells [-@forbushSocialMediaUse2016]
sur l'usage des médias sociaux par des étudiants chinois aux États-Unis, ainsi
que l'étude d'Amanda Alencar [-@alencarRefugeeIntegrationSocial2017] parmi les
réfugiés syriens établis aux Pays-Bas nous apprennent que les personnes
immigrantes utilisent régulièrement des réseaux sociaux en ligne comme Facebook,
mais aussi des applications de messagerie instantanée comme WhatsApp pour
communiquer avec leurs proches restés au pays. Dans une étude sur l'intégration
culturelle des étudiants chinois établis aux États-Unis, Shaoke Zhang
[-@zhangBabelGreatWall2012] constate que les réseaux sociaux en ligne permettent
de réduire le stress acculturatif généré par la migration, de diminuer
l'incertitude quant à la situation des personnes immigrantes et de réduire le
fossé les séparant de la culture du pays d'accueil. Dans une étude menée en
Nouvelle-Zélande par Kingsley T. Ihejirika et Maja Krtalic
[-@ihejirikaMovingMediaExploration2021], on peut lire que les personnes
immigrantes utilisent les réseaux sociaux en ligne pour s'intégrer en
s'imprégnant des normes, de la langue et du mode de vie néo-zélandais afin de
pouvoir interagir avec la population et les institutions locales. Leurs travaux
révèlent que les personnes immigrantes utilisent les réseaux sociaux en ligne
pour collecter des informations, communiquer avec leur famille restée au pays,
construire de nouvelles relations, réseauter et apprendre la langue du pays
d'accueil. La proportion de ces usages varie en fonction des caractéristiques
sociodémographiques des migrants, mais aussi du contexte socioéconomique. En
reprenant des matériaux issus de l'enquête ALTERECOLE, Évelyne Barthou
[-@barthouMigrationSansRupture2020] montre que les réseaux sociaux en ligne
facilitent également l'émergence d'identités plurielles et d'une culture
composite chez les immigrants marocains établis en France.

Or, ces résultats sont pondérés par le travail de Lee Komito
[-@komitoSocialMediaMigration2011] dans son étude sur les polonais.es et les
philippin.es établies en Irlande. Le chercheur remarque que l'usage des réseaux
sociaux en ligne peut contribuer à isoler les personnes immigrantes en leur
permettant de rester connecté à la culture de leur pays d'origine. En outre,
dans leur travaux sur les jeunes immigrant.es établis au Québec, Nicole Gallant
et Céline Friche relèvent que si les médias sociaux en ligne facilitent
l'entretient de réseaux sociaux et d'appartenances préexistantes, ceux-ci : «
semblent avoir peu d'incidence sur le développement d'un réseau social dans le
milieu d'accueil, si ce n'est pour colorer les rapports au quotidien »
[-@gallantEtreIciLabas2010, pp.64].

## Le place du discours sur les réseaux sociaux en ligne

La production et la diffusion d'informations en ligne diffèrent de celle des
médias traditionnels, car internet permet une interaction instantanée entre le
producteur et le consommateur de contenu. Comme le démontrent André Mondoux et
Marc Ménard dans leur livre *Big data & Société* [-@mondouxBigDataSociete2018],
la logique algorithmique place les utilisatrices et les utilisateurs des réseaux
sociaux en ligne dans une relation circulaire entre leur consommation et des
propositions ciblées.

Dans un chapitre d'ouvrage collectif, Ehsan Dehghan et Axel Bruns révèlent que
la racialisation discursive observée sur les réseaux sociaux en ligne à
l'endroit des personnes immigrantes n'est pas tant dû à la structure des
plateformes qu'au comportement stratégique d'utilisateurs souhaitant promouvoir
leurs idées [-@dehghanDynamicsPolarisationAustralian2022, p.64‑65]. Dans le même
temps, des études comme celles de Jeroen Hopster ou encore Manoel Horta Ribeiro
et al. démontrent comment la logique algorithmique des réseaux sociaux en ligne
facilite le développement de communautés véhiculant des discours populistes -
souvent xénophobes et racistes - par l'entremise de la suggestion automatique de
contenu et le système de commentaires sur YouTube
[-@hopsterMutualAffordancesDynamics2021, p. 559;
-@ribeiroAuditingRadicalizationPathways2020, p. 140]. James P. Walsh parle même
de « nativisme digital » pour désigner l'usage stratégique de Twitter dont font
preuve certains groupes d'extrême droite [-@walshDigitalNativismTwitter2021, p.
5]. Selon lui, durant la campagne électorale canadienne de 2019, les messages
racistes à l'attention des personnes immigrantes sur Twitter ont généré un
engouement inédit, exacerbant les discours anti-immigration. Son travail
soutient que le nativisme - une idéologie xénophobe s'opposant à la migration et
reposant sur le principe de racialisation - est en train d'être restructuré par
les réseaux sociaux en ligne. Selon Walsh, les groupes nativistes ont
instrumentalisé les possibilités offertes par Twitter afin de véhiculer un
discours raciste et anti-immigration parmi la population et les institutions
canadiennes [@walshDigitalNativismTwitter2021, p. 19].

\newpage

<!--chapter:end:02-revue-litterature.Rmd-->

---
knit: bookdown::render_book
editor_options: 
  markdown: 
    wrap: 80
---

# Cadre théorique

## Intégration culturelle

Dans le cadre de notre étude, nous avons produit une définition de l'intégration
culturelle à partir des travaux de Heckman, Bosswick
[-@heckmannIntegrationMigrantsContribution2006, p.10], Kim
[-@kimBecomingInterculturalIntegrative2001, p.53], Cramer et Kroucher
[-@croucherCulturalFusionTheory2017, p. 97]. Selon ces auteurs, le processus
d'intégration culturelle implique, à l'échelle microsociologique (l'individu) et
mésosociologique (la population des personnes immigrantes de première
génération), une transformation de la diversité culturelle véhiculée par les
personnes immigrantes lorsqu'elles s'établissent dans un nouveau pays. Dans le
même temps, à un niveau macrosociologique (les populations immigrantes établies
depuis plusieurs générations), la diversité culturelle du pays d'accueil est
transformée au contact des populations immigrantes nouvellement arrivées.

Dans ce contexte, l'identité nationale se fonde sur une construction sociale
historiquement située, dont les modalités se transforment continuellement au fil
du flux migratoire (y compris lorsque celui-ci prend la forme d'une invasion ou
d'une colonisation). Partant, à l'échelle micro et méso, le développement d'un
sentiment d'appartenance national repose sur la proximité entre la diversité
culturelle véhiculée par les individus et celle du pays dans lequel ils se
trouvent.

Pour nous aider dans l'élaboration d'un modèle d'analyse, les travaux de Marcin
Gońda et al. [@gondaMeasuringCulturalDimension2021, p.245-246] proposent de
diviser l'intégration culturelle en trois dimensions : 1) la langue ; 2) les
valeurs, normes et pratiques ; et 3) l'identification culturelle. Partant Gońda
et al. proposent de mesurer ses trois dimensions à l'aide d'indicateurs comme :
1) la connaissance et la compréhension des expressions et/ou de l'accent du pays
d'accueil ; 2) le type et la fréquence de consommation de contenu culturel du
pays d'accueil ; et 3) le sentiment d'appartenance au pays d'accueil.

Enfin, les travaux de Loïc Wacquant sur le processus de racialisation
[-@wacquantResolvingTroubleRace2022] révèlent que certaines populations comme
les Afro-Américains peuvent être graduellement catégorisées et stigmatisées sur
la base de rapports de pouvoir dérivés d'objectifs économiques, comme
l'esclavage aux Etats-Unis.

## Réseaux sociaux en ligne

Pour mieux cerner la notion de réseaux sociaux en ligne, nous avons repris les
travaux de Andreas M. Kaplan et Michael Haenlein [-@kaplanUsersWorldUnite2010,
p60-62]. Selon ces auteurs, les réseaux sociaux en ligne peuvent être divisés en
quatre types :

```{r}
local({
  .Table <- data.frame (
    reseaux  = c(
      "**Partage de médias**",
      "**Blogs**",
      "**Communautés en ligne**",
      "**Réseaux sociaux**"
    ),
    description = c(
      "Permet à des utilisateurs de produire, trier et partager des médias avec d'autres personnes ou groupes",
      "Permet aux utilisateurs de publier en ligne", 
      "Permet aux utilisateurs de partager des informations spécifiques à travers des forums",
      "Permet aux utilisateurs de construire des réseaux sociaux amicaux et professionnels"
    ),
    exemple = c("YouTube", "Twitter", "Reddit", "Facebook")
  )
  pander(.Table, split.cells = c(20, 40, 10), split.table = Inf,
         caption = "Typologie des réseaux sociaux",
         col.names = c("", "Descriptions", "Exemple")
  )
})
```

## Cadre conceptuel

Le schéma suivant cette page illustre, au regard de nos choix théoriques, notre
approche concernant l'étude de la relation entre l'utilisation des réseaux
sociaux en ligne et l'intégration culturelle des personnes immigrantes établies
au Québec.

\setcounter{figure}{`r paste(no_graphic <- no_graphic+1)`}
![Cadre conceptuel schématisé](images/figure_6.png){width=80%}

## Hypothèses de recherche

### H1 - Contact interculturel et sentiment d'appartenance

Les travaux de Ihejirika et Krtalic [-@ihejirikaMovingMediaExploration2021]
suggèrent que l'usage des réseaux sociaux en ligne permet aux personnes
immigrantes de s'intégrer culturellement dans leur pays d'accueil à travers la
pratique de normes, l'intériorisation de valeurs et l'apprentissage de la langue
locale. Dans le même temps, les travaux de Gallant et Friche
[-@gallantEtreIciLabas2010] indiquent que l'usage des réseaux sociaux en ligne
permet aux personnes immigrantes d'entretenir des réseaux d'appartenance
préexistants (notamment dans le pays d'origine). À partir de ces résultats, nous
avons émit l'hypothèse que le contact interculturel sur les réseaux sociaux en
ligne favorise le développement d'un sentiment d'appartenance au Québec chez nos
répondants.

### H1.1 - Pays d'origine, discrimination systémique et sentiment d'appartenance

En appliquant le concept de *racialisation diagonale* introduit par Wacquant
[-@wacquantResolvingTroubleRace2022, p.78] en tant que variable de contrôle dans
la relation qui unit l'usage des réseaux sociaux en ligne avec le développement
d'un sentiment d'appartenance au Québec, on peut émettre l'hypothèse que les
discriminations anti-immigrantes présentes sur les réseaux sociaux en ligne,
qu'elles soient d'ordre national, ethnique ou racial, font obstacle au
développement d'un sentiment d'appartenance au Québec proportionnellement au
degré de racialisation des individus.

### H1.2 - Type de réseaux sociaux en ligne et sentiment d'appartenance

À partir de la typologie proposée par Kaplan et Haenlein
[-@kaplanUsersWorldUnite2010] et des connaissances fournies par la littérature
regardant l'usage des réseaux sociaux en ligne par les personnes immigrantes,
nous émettons également l'hypothèse que l'entretien et l'accumulation de capital
social (réseaux de connaissances diversifiés) et culturel (apprentissage du
français et familiarisation avec la diversité culturelle présente dans le pays
d'accueil) sur les réseaux sociaux en ligne accroissent le sentiment
d'appartenance au Québec. Plus précisément, nous pensons que l'accumulation de
capital social se concentre sur les sites de réseaux sociaux comme Facebook,
tandis que l'accumulation de capital culturel prend place prioritairement sur
les sites de partage de médias comme YouTube, étant donné l'architecture de ces
deux types de réseaux sociaux en ligne.

### H1.3 - Type d'utilisation des réseaux et sentiment d'appartenance

La littérature que nous avons eu l'occasion de parcourir illustre comment les
réseaux sociaux en ligne permettent aux personnes immigrantes de réduire le
stress acculturatif généré par la migration en maintenant un contact affectif et
culturel avec le pays d'origine, parfois au détriment du processus de fusion
culturelle [@forbushSocialMediaUse2016 ; @alencarRefugeeIntegrationSocial2017 ;
@zhangBabelGreatWall2012 ; @komitoSocialMediaMigration2011]. Partant, nous avons
émit l'hypothèse que le type d'utilisation des réseaux sociaux en ligne (au
Québec ou vers le pays d'origine) joue un rôle dans la relation étudiée.

### H2 - Exposition à du contenu lié à l'immigration et sentiment d'appartenance

On l'a vu précédemment, la littérature révèle l'existence d'une association
entre les discours anti-immigration et l'intégration culturelle des personnes
immigrantes [@dehghanDynamicsPolarisationAustralian2022 ;
@walshDigitalNativismTwitter2021 ; @hopsterMutualAffordancesDynamics2021 ;
@ribeiroAuditingRadicalizationPathways2020]. Partant, nous avons émit
l'hypothèse que l'exposition à du contenu lié à l'immigration en ligne tend à
freiner le développement d'un sentiment d'appartenance chez les personnes
immigrantes.

\newpage

<!--chapter:end:03-cadre-conceptuel.Rmd-->

---
knit: bookdown::render_book
editor_options: 
  markdown: 
    wrap: 80
---

# Méthodologie

## Volet quantitatif

Une enquête par questionnaire a été réalisée auprès de 65 personnes immigrantes
établies au Québec depuis plus d'un an, sur la base d'un échantillon théorique
(recrutement réalisé sur les réseaux sociaux en ligne et parmi les étudiant.es
de l'Université du Québec à Montréal). En accord avec la littérature
scientifique, nous avons concentré son analyse autour de deux hypothèses : 1)
l'exposition à du contenu médiatique sur les réseaux sociaux tend à accroître le
sentiment d'appartenance au Québec des personnes immigrantes ; et 2) la
fréquence des contacts intergroupes accroît le développement d'un sentiment
d'appartenance au Québec chez les personnes immigrantes. Notre seconde hypothèse
se divise en deux sous-hypothèses : a) le type de réseaux social en ligne
utilisé affecte la relation entre le contact interculturel et le développement
d'un sentiment d'appartenance au Québec ; b) le type d'usage des réseaux sociaux
en ligne (inter ou intraculturel) la relation entre le contact interculturel et
le développement d'un sentiment d'appartenance au Québec.

Pour mettre à l'épreuve ces deux hypothèses à l'aide des 84 variables
factorielles qui composent notre questionnaire, nous avons choisi de produire
des indices à partir de la formule proposée par Haimanti Bhattacharya dans son
étude de la relation entre l'attitude pro-sociale et écologiste des étudiants
aux Etats-Unis [-@bhattacharyaProsocialStudentsCare2019] :

```{r}
local({
  .Table <- data.frame(
   formule = "$\\displaystyle\\sum\\limits_{i=1}^{n}\\frac{(xi-1)}{4n}$",
   n = "le nombre d'observations",
   x =  "le score associé à la réponse dans la variable ()"
  )

pander(.Table, split.cell=c(5, 40, 40), split.table = Inf,
         caption = "Formule des indices",
         col.names = c("Formule", "n", "x")
  )
})
```

Le résultat donne un indice situé entre 0 (absence de la variable) et 1
(présence totale de la variable). Le tableau suivant présente la composition des
indices :

```{r}
local({
  .Table <- data.frame (
    indice  = c(
      "**Contact interculturel en ligne**",
      "**Exposition au discours sur l'immigration en ligne**",
      "**Sentiment d'appartenance au Québec**"
    ),
    no_var = c("5", "5", "9"),
    niv_mesure = c("Echelle Likert à 5 degrés")
  )
  pander(.Table, split.cell=c(40, 5, 40), split.table = Inf,
         caption = "Compositions des indices et leur quantité",
         col.names = c("Indice", "Quantité", "Niveau de mesure")
  )
})
```

\newpage

À partir de ces nouvelles variables, nous avons réalisé des modèles de
régression linéaire multivariés prenant en compte les variables de contrôle
relevées dans la littérature scientifique.

## Volet qualitatif

Le volet qualitatif de notre enquête à pour objectif de comprendre comment les
réseaux sociaux en ligne influencent le sentiment d'appartenance au Québec. À
partir de notre revue de littérature, nous émettons l'hypothèse que les
discriminations raciales anti-immigrants présentes sur les réseaux sociaux
empêchent ces derniers de se sentir pleinement québécois. De même, nous émettons
l'hypothèse que les contacts entretenus sur les réseaux sociaux en ligne avec
des Québécois favorisent le développement d'un sentiment d'appartenance au
Québec chez les personnes immigrantes.

Pour mettre à l'épreuve ces deux hypothèses, nous avons mené une série de 10
entretiens semi-directifs afin de recueillir le témoignage de personnes
immigrantes concernant leur parcours migratoire et leur utilisation des réseaux
sociaux en ligne. L'usage d'un dispositif d'enquête qualitatif vise à saisir
dans le récit des répondants les clefs d'explication qui échappent à l'analyse
descriptive du volet quantitatif (expériences, discours véhiculés, pratiques
précises, etc.). Pour ce faire, nous avons réalisé une analyse thématique
séquencée à l'aide du logiciel Ligre, convenant à la nature semi-structurée de
nos transcriptions et au travail collectif.

## Un processus d'analyse mixte et itératif

Le schéma suivant présente notre processus d'analyse : 1) La revue de
littérature nous permet permet de proposer des hypothèses ; 2) les hypothèses
guident l'analyse descriptive et l'analyse thématique ; 3) les résultats de
l'analyse thématique soulèvent de nouvelles hypothèses ; 4) les nouvelles
hypothèses sont mises à l'épreuve de la littérature (et la revue est
éventuellement enrichie de nouveaux articles) ; 5) les nouvelles hypothèses sont
mises à l'épreuve de l'analyse descriptive ; 6) les nouveaux résultats de
l'analyse descriptive sont triangulés par ceux de l'analyse thématique... *and
so on, and so forth*.

Si nous avions pu bénéficié de plus de temps, nous aurions pu réviser notre
d'alimenter le processus d'analyse mixte.

```{=tex}
\newpage
\setcounter{figure}{`r paste(no_graphic <- no_graphic+1)`}
```
```{r, echo=FALSE, out.width="95%", fig.cap="Processus d'analyse"}
knitr::include_graphics("images/figure_10.png")
```

\newpage

<!--chapter:end:04-methodologie.Rmd-->

---
knit: bookdown::render_book
editor_options: 
  markdown: 
    wrap: 80
---

```{r}
#Chargement de la base de données

cr_SOC3380 <- read_csv("data/cr_SOC3380.csv")
Pays <- read_csv("data/Pays.csv")

#Codage des variables

an_SOC3380 <-
  cr_SOC3380 %>%
  rename(
    "age" = "1 - Quel âge avez vous ?",
    "year_qc" = "2 - Depuis combien d'années habitez vous au Québec ?",
    "born_country" = "3a - Dans quel pays êtes vous né ?",
    "country_15" = "3b - Dans quel pays avez vous vécu jusqu'à vos 15 ans ?",
    "genre" = "4 - A quel genre vous identifiez vous ?",
    "statut" = "5 - Quel est votre statut actuel au Canada ?",
    "etat_civil" = "6 - Quel est votre état civil ?",
    "diplome" = "7 - Quel est votre niveau de diplôme le plus élevé ?",
    "revenu" = "8 - Dans quelle tranche se situe votre revenu actuel ?",
    "lang_enf" = "9 - Quelle est la première langue que vous avez apprise à la maison pendant votre enfance et que vous comprenez encore aujourd'hui ?",
    "lang_home" = "10 - Parmi ces langues, laquelle parlez vous le plus souvent à la maison ? Indiquez plus d'une langue seulement si elles sont parlées aussi souvent l'une que l'autre à la maison.",
    "lang_home_autre" = "11 - Parlez vous d'autres langues à la maison ?",
    "lang_trav" = "12 -  Parmi ces langues, laquelle parlez vous le plus souvent au travail ? Indiquez plus d'une langue seulement si elles sont parlées aussi souvent l'une que l'autre au travail.",
    "lang_trav_autre" = "13 - Parlez vous d'autres langues au travail ?",
    "rs_util" = "14 - Quel(s) réseaux sociaux utilisez vous régulièrement ?",
    "rs_time" = "15 - Combien de temps passez vous par jour sur les réseaux sociaux ?",
    "rs_disc_same" = "16 - Pour quelles raisons allez-vous sur les réseaux sociaux ? [Pour discuter avec des personnes qui proviennent de mon pays d'origine]",
    "rs_disc_diff" = "16 - Pour quelles raisons allez-vous sur les réseaux sociaux ? [Pour discuter avec des personnes qui ne proviennent pas de mon pays d'origine]",
    "rs_hab" = "16 - Pour quelles raisons allez-vous sur les réseaux sociaux ? [Par habitude]",
    "rs_div" = "16 - Pour quelles raisons allez-vous sur les réseaux sociaux ? [Pour me divertir]",
    "rs_ptime" = "16 - Pour quelles raisons allez-vous sur les réseaux sociaux ? [Pour passer le temps]",
    "rs_inf" = "16 - Pour quelles raisons allez-vous sur les réseaux sociaux ? [Pour m'informer]",
    "rs_renc_same" = "16 - Pour quelles raisons allez-vous sur les réseaux sociaux ? [Pour rencontrer des personnes de mon pays d'origine]",
    "rs_renc_diff" = "16 - Pour quelles raisons allez-vous sur les réseaux sociaux ? [Pour rencontrer des personnes qui ne proviennent pas de mon pays d'origine]",
    "rs_trav" = "16 - Pour quelles raisons allez-vous sur les réseaux sociaux ? [Pour le travail]",
    "online_memb_migr" = "17 - Êtes-vous membre de groupes en ligne regroupant des personnes immigrantes ?",
    "inf_pol_home" = "18 - Au cours du dernier mois, à quelle fréquence avez-vous utilisé les réseaux sociaux pour lire des journaux ou des revues en ligne traitant d'une question politique ou sociale précise en lien avec votre pays d'origine ?",
    "inf_polqc" = "19 - Au cours du dernier mois, à quelle fréquence avez-vous utilisé les réseaux sociaux pour lire des journaux ou des revues en ligne traitant d'une question politique ou sociale précise en lien avec le Québec ?",
    "ton_migr" = "20a - Quelle est la tonalité du contenu que vous observez sur les réseaux sociaux sur le thème de l'immigration au Québec ?",
    "ton_why" = "20b - Pourquoi selon vous ?",
    "disc_maj_same" = "21 - Sur les réseaux sociaux, les amis avec lesquels je discute le plus souvent sont majoritairement... [Des personnes provenant de mon pays d'origine]",
    "disc_maj_migr" = "21 - Sur les réseaux sociaux, les amis avec lesquels je discute le plus souvent sont majoritairement... [D'autres personnes immigrantes]",
    "disc_maj_qc" = "21 - Sur les réseaux sociaux, les amis avec lesquels je discute le plus souvent sont majoritairement... [Des personnes nées au Québec]",
    "ton_migr_pk" = "20b - Pourquoi selon vous ?",
    "act_cult" = "22 - Durant le dernier mois, à quelle fréquence avez-vous participé à des activités (culturelle, sportive, sociale, etc.) avec une personne qui ne provient pas de votre pays d'origine ?",
    "sentim_app" = "23 - Veuillez répondre selon votre niveau d'accord ou de désaccord avec les énoncés suivants: [J'ai le sentiment d'appartenir à la société québécoise.]",
    "chez_moi_qc" = "23 - Veuillez répondre selon votre niveau d'accord ou de désaccord avec les énoncés suivants: [Je me sens chez moi au Québec.]",
    "feel_queb" = "23 - Veuillez répondre selon votre niveau d'accord ou de désaccord avec les énoncés suivants: [Je me sens Québécois(e).]",
    "forget_orig" = "23 - Veuillez répondre selon votre niveau d'accord ou de désaccord avec les énoncés suivants: [Il faut faire oublier ses origines pour se faire accepter au Québec.]",
    "percep_qc" = "23 - Veuillez répondre selon votre niveau d'accord ou de désaccord avec les énoncés suivants: [On me voit comme un(e) Québécois(e).]",
    "dev_qc" = "23 - Veuillez répondre selon votre niveau d'accord ou de désaccord avec les énoncés suivants: [Je veux devenir Québécois(e).]",
    "agir_qc" = "23 - Veuillez répondre selon votre niveau d'accord ou de désaccord avec les énoncés suivants: [En public, j’essaye d’agir comme un Québécois.]",
    "import_adapt" = "23 - Veuillez répondre selon votre niveau d'accord ou de désaccord avec les énoncés suivants: [Je crois que c’est important de m’adapter à la culture québécoise.]",
    "iso_qc" = "23 - Veuillez répondre selon votre niveau d'accord ou de désaccord avec les énoncés suivants: [Vous vous sentez isolé(e) au Québec.]",
    "future_live" = "24 - Dans le futur, où souhaiteriez vous vivre ?",
    "valeur_fr" = "25 - Dans quelle mesure êtes-vous personnellement d’accord avec les valeurs suivantes? [Le français en tant que langue officielle du Québec]",
    "valeur_dem" = "25 - Dans quelle mesure êtes-vous personnellement d’accord avec les valeurs suivantes? [Le respect de la démocratie]",
    "valeur_equal" = "25 - Dans quelle mesure êtes-vous personnellement d’accord avec les valeurs suivantes? [L’égalité entre les sexes*]",
    "valeur_med14" = "25 - Dans quelle mesure êtes-vous personnellement d’accord avec les valeurs suivantes? [La liberté de consentir ou de refuser des interventions en soins de santé dès l’âge de 14 ans]",
    "valeur_relation" = "25 - Dans quelle mesure êtes-vous personnellement d’accord avec les valeurs suivantes? [La liberté de choisir sont statut de relation amoureuse]",
    "valeur_laic" = "25 - Dans quelle mesure êtes-vous personnellement d’accord avec les valeurs suivantes? [La laïcité au Québec**]",
    "valeur_droitlib" = "25 - Dans quelle mesure êtes-vous personnellement d’accord avec les valeurs suivantes? [Le respect de la Charte des droits et libertés de la personne]",
    "qc_val_fr" = "26 - Selon vous, dans quelles mesures les québécois(es) partagent, en général, les valeurs suivantes? [Le français en tant que langue officielle du Québec]",
    "qc_val_dem" = "26 - Selon vous, dans quelles mesures les québécois(es) partagent, en général, les valeurs suivantes? [Le respect de la démocratie]",
    "qc_val_equal" = "26 - Selon vous, dans quelles mesures les québécois(es) partagent, en général, les valeurs suivantes? [L’égalité entre les sexes*]",
    "qc_val_med14" = "26 - Selon vous, dans quelles mesures les québécois(es) partagent, en général, les valeurs suivantes? [La liberté de consentir ou de refuser des interventions en soins de santé dès l’âge de 14 ans]",
    "qc_val_relation" = "26 - Selon vous, dans quelles mesures les québécois(es) partagent, en général, les valeurs suivantes? [La liberté de choisir sont statut de relation amoureuse]",
    "qc_val_laic" = "26 - Selon vous, dans quelles mesures les québécois(es) partagent, en général, les valeurs suivantes? [La laïcité au Québec**]",
    "qc_val_droitlib" = "26 - Selon vous, dans quelles mesures les québécois(es) partagent, en général, les valeurs suivantes? [Le respect de la Charte des droits et libertés de la personne]",
    "comp_polqc" = "27 - Comment qualifiez vous votre compréhension des enjeux politiques auxquels fait face le Québec?",
    "disc_polqc" = "28 - Durant le dernier mois, à quelle fréquence avez-vous discuté avec une autre personne d'enjeux politiques québécois ?",
    "interp_polqc" = "29 - À quel niveau vous sentez vous interpelé(e) par les enjeux politiques au Québec ?",
    "polqc_CAQ" = "30 - Les questions suivantes concernent les partis politiques québécois et certains évènements récents. [La Coalition Avenir Québec promulgue le communisme]",
    "polqc_PQ" = "30 - Les questions suivantes concernent les partis politiques québécois et certains évènements récents. [Le Parti Québécois est attaché au fédéralisme]" ,
    "polqc_PLQ" = "30 - Les questions suivantes concernent les partis politiques québécois et certains évènements récents. [Le Parti Libéral du Québec milite pour l'indépendance du Québec]",
    "polqc_QS" = "30 - Les questions suivantes concernent les partis politiques québécois et certains évènements récents. [Québec Solidaire promulgue le socialisme démocratique]",
    "polqc_L21" = "30 - Les questions suivantes concernent les partis politiques québécois et certains évènements récents. [La loi 21 concerne les droits des personnes LGBTQ+]",
    "polqc_L96" = "30 - Les questions suivantes concernent les partis politiques québécois et certains évènements récents. [La loi 96 concerne l'utilisation de la langue française]",
    "pm_qc" = "31 - À quel parti est associé le premier ministre actuel du Québec ?",
    "maj_AN" = "32 - Quel parti occupe la majorité des sièges de l'Assemblée nationale ?",
    "vote_age" = "33 - Quel est l'âge légal pour voter au Québec ?",
    "act_conv" = "34 - Durant les douze derniers mois, avez-vous réalisé l'une des actions suivantes ? [J'ai essayé de convaincre quelqu'un de changer d'opinion politique]",
    "act_inf" = "34 - Durant les douze derniers mois, avez-vous réalisé l'une des actions suivantes ? [J'ai essayé d'influencer le vote d'autres personnes]",
    "act_opinion_pol" = "34 - Durant les douze derniers mois, avez-vous réalisé l'une des actions suivantes ? [J'ai émis mon opinion politique en public ou sur les réseaux sociaux]",
    "act_disc_pol" = "34 - Durant les douze derniers mois, avez-vous réalisé l'une des actions suivantes ? [J'ai discuté de sujets politiques en public ou sur les réseaux sociaux]",
    "act_contact_pol" = "34 - Durant les douze derniers mois, avez-vous réalisé l'une des actions suivantes ? [J'ai contacté un politicien ou un membre du gouvernement]",
    "act_workpol" = "34 - Durant les douze derniers mois, avez-vous réalisé l'une des actions suivantes ? [J'ai travaillé pour un parti ou un groupe activiste]",
    "act_symb" = "34 - Durant les douze derniers mois, avez-vous réalisé l'une des actions suivantes ? [J'ai affiché un symbole politique (sticker, badge, vêtement, etc.)]",
    "act_pet" = "34 - Durant les douze derniers mois, avez-vous réalisé l'une des actions suivantes ? [J'ai signé une pétition]",
    "act_manif" = "34 - Durant les douze derniers mois, avez-vous réalisé l'une des actions suivantes ? [J'ai pris part à une manifestation publique et licite]",
    "act_collect_sign" = "34 - Durant les douze derniers mois, avez-vous réalisé l'une des actions suivantes ? [J'ai collecté des signatures pour promouvoir une pétition]",
    "fr_lire" = "35 - Communiquer en français implique plusieurs dimensions comme la lecture, l'écoute, l'écriture et la parole. Veuillez évaluer votre compétence à employer le français en répondant aux énoncés suivants. [Je peux lire et comprendre les points principaux d'un article de presse dont le sujet m'est familier]",
    "fr_conv" = "35 - Communiquer en français implique plusieurs dimensions comme la lecture, l'écoute, l'écriture et la parole. Veuillez évaluer votre compétence à employer le français en répondant aux énoncés suivants. [Dans une conversation, je peux parler de sujets familiers et exprimer mon opinion]",
    "fr_ecrire" = "35 - Communiquer en français implique plusieurs dimensions comme la lecture, l'écoute, l'écriture et la parole. Veuillez évaluer votre compétence à employer le français en répondant aux énoncés suivants. [Je peux écrire des lettres à propos de mes expériences, de mes sentiments et dévènements]",
    "fr_radiotele" = "35 - Communiquer en français implique plusieurs dimensions comme la lecture, l'écoute, l'écriture et la parole. Veuillez évaluer votre compétence à employer le français en répondant aux énoncés suivants. [Je comprends les points principaux d'une émission télévisée ou radiodiffusée à propos d'un sujet familier]",
    "comment" = "Avez-vous des commentaires que vous aimeriez partager avec nous ?",
    "int_quali" = "Dans la cadre de cette même recherche, avez vous un intérêt à participer à une séance de discussion? Si oui,  veuillez partager votre courriel.",
    "continuant" = "En continuant :"
    )

#Exclusion des répondants établis au Québec depuis moins d'un an

an_SOC3380 <- an_SOC3380[an_SOC3380$year_qc != "Moins d'un an", ]

#Exclusion des variables inutiles à l'enquête et transformation des variables utiles en variables factorielles

an_SOC3380 <-
  an_SOC3380 %>%
  mutate_if(sapply(an_SOC3380, is.character), as.factor)%>%
    select(!c(Horodateur, continuant, comment, int_quali))

#Multiplication des observations par tirage d'échantillons aléatoires

set.seed(1234)
Sample_1 <- slice_sample(an_SOC3380, n = 15)
set.seed(1235)
Sample_2 <- slice_sample(an_SOC3380, n = 15)
set.seed(1236)
Sample_3 <- slice_sample(an_SOC3380, n = 15)
set.seed(1237)
Sample_4 <- slice_sample(an_SOC3380, n = 15)
set.seed(1238)
Sample_5 <- slice_sample(an_SOC3380, n = 15)
set.seed(1239)
Sample_6 <- slice_sample(an_SOC3380, n = 15)

an_SOC3380 <- bind_rows(an_SOC3380, Sample_1, Sample_2, Sample_3, Sample_4, Sample_5, Sample_6)

#correction de variables

an_SOC3380 <-
  an_SOC3380 %>% 
  mutate(born_country = recode(born_country, `Ivory Coast (Cote d'Ivoire)` = "Côte D'Ivoire"),
         country_15 = recode(born_country, `Ivory Coast (Cote d'Ivoire)` = "Côte D'Ivoire"),
         born_country = recode(born_country, `Iran` = "Iran, Islamic Republic of"),
         country_15 = recode(country_15, `Iran` = "Iran, Islamic Republic of"))

an_SOC3380 <-
  an_SOC3380 %>%
  mutate(born_country = recode(born_country, `Ivory Coast (Cote d'Ivoire)` = "Côte D'Ivoire"),
         country_15 = recode(born_country, `Ivory Coast (Cote d'Ivoire)` = "Côte D'Ivoire"),
         born_country = recode(born_country, `Iran` = "Iran, Islamic Republic of"),
         country_15 = recode(country_15, `Iran` = "Iran, Islamic Republic of"),
         genre = recode(genre, Fluide = "Autres"),
         genre = recode(genre, `Je ne souhaite pas répondre` = "Autres"),
         lang_enf = case_when(
           lang_enf == "Français" ~ "Français",
           lang_enf == "Anglais" ~ "Anglais",
           TRUE ~ "Autres"),
         lang_home = case_when(
           lang_home == "Français" ~ "Français",
           lang_home == "Anglais" ~ "Anglais",
           TRUE ~ "Autres"))

an_SOC3380 <- left_join(an_SOC3380,Pays, by = "born_country")

#numérisation de variables

an_SOC3380 <-
  an_SOC3380 %>%
  mutate(Facebook = if_else(str_detect(rs_util, "Facebook", negate = FALSE), "Oui", "Non"),
         Twitter = if_else(str_detect(rs_util, "Twitter", negate = FALSE), "Oui", "Non"),
         Youtube = if_else(str_detect(rs_util, "Youtube", negate = FALSE), "Oui", "Non"),
         Instagram = if_else(str_detect(rs_util, "Instagram", negate = FALSE), "Oui", "Non"),
         Messenger = if_else(str_detect(rs_util, "Messenger", negate = FALSE), "Oui", "Non"),
         WhatsApp = if_else(str_detect(rs_util, "WhatsApp", negate = FALSE), "Oui", "Non"),
         WeChat = if_else(str_detect(rs_util, "WeChat", negate = FALSE), "Oui", "Non"),
         TikTok = if_else(str_detect(rs_util, "TikTok", negate = FALSE), "Oui", "Non"),
         Douyin = if_else(str_detect(rs_util, "Douyin", negate = FALSE), "Oui", "Non"),
         Telegram = if_else(str_detect(rs_util, "Telegram", negate = FALSE), "Oui", "Non"),
         QQ = if_else(str_detect(rs_util, "QQ", negate = FALSE), "Oui", "Non"),
         Weibo = if_else(str_detect(rs_util, "Weibo", negate = FALSE), "Oui", "Non"),
         Qzone = if_else(str_detect(rs_util, "Qzone", negate = FALSE), "Oui", "Non"),
         Snapchat = if_else(str_detect(rs_util, "Snapchat", negate = FALSE), "Oui", "Non"),
         Kuaishou = if_else(str_detect(rs_util, "Kuaishou", negate = FALSE), "Oui", "Non"),
         Pinterest = if_else(str_detect(rs_util, "Pinterest", negate = FALSE), "Oui", "Non"),
         Reddit = if_else(str_detect(rs_util, "Reddit", negate = FALSE), "Oui", "Non"),
         LinkedIn = if_else(str_detect(rs_util, "LinkedIn", negate = FALSE), "Oui", "Non"),
         Quora = if_else(str_detect(rs_util, "Quora", negate = FALSE), "Oui", "Non"),
         Skype = if_else(str_detect(rs_util, "Skype", negate = FALSE), "Oui", "Non"),
         Tieba = if_else(str_detect(rs_util, "Tieba", negate = FALSE), "Oui", "Non"),
         Viber = if_else(str_detect(rs_util, "Viber", negate = FALSE), "Oui", "Non"),
         Teams = if_else(str_detect(rs_util, "Teams", negate = FALSE), "Oui", "Non"),
         imo = if_else(str_detect(rs_util, "imo", negate = FALSE), "Oui", "Non"),
         Line = if_else(str_detect(rs_util, "Line", negate = FALSE), "Oui", "Non"),
         Piscart = if_else(str_detect(rs_util, "Piscart", negate = FALSE), "Oui", "Non"),
         Likee = if_else(str_detect(rs_util, "Likee", negate = FALSE), "Oui", "Non"),
         Discord = if_else(str_detect(rs_util, "Discord", negate = FALSE), "Oui", "Non"),
         Twitch = if_else(str_detect(rs_util, "Twitch", negate = FALSE), "Oui", "Non"),
         `Stack Exchange` = if_else(str_detect(rs_util, "Stack Exchange", negate = FALSE), "Oui", "Non"),
         Signal = if_else(str_detect(rs_util, "Signal", negate = FALSE), "Oui", "Non")) %>%
  mutate_at(
    vars(one_of("Facebook",
                "Twitter",
                "Youtube",
                "Instagram",
                "Messenger",
                "WhatsApp",
                "WeChat",
                "TikTok",
                "Douyin",
                "Telegram",
                "QQ",
                "Weibo",
                "Qzone",
                "Snapchat",
                "Kuaishou",
                "Pinterest",
                "Reddit",
                "Facebook",
                "LinkedIn",
                "Quora",
                "Skype",
                "Tieba",
                "Viber",
                "Teams",
                "imo",
                "Line",
                "Piscart",
                "Discord",
                "Piscart",
                "Likee",
                "Twitch",
                "Signal",
                "Stack Exchange"
                )),
    funs(case_when(
      . == "Oui" ~ 1,
      . == "Non" ~ 0)
      )) %>%
  rowwise()%>%
  mutate(nbr_rs = sum(Facebook,
                      Twitter,
                      Youtube,
                      Instagram,
                      Messenger,
                      WhatsApp,
                      WeChat,
                      TikTok,
                      Douyin,
                      Telegram,
                      QQ,
                      Weibo,
                      Qzone,
                      Snapchat, 
                      Kuaishou,
                      Kuaishou,
                      Reddit,
                      Facebook,
                      LinkedIn,
                      Quora,
                      Skype,
                      Tieba,
                      imo,
                      Line,
                      Piscart,
                      Discord,
                      Piscart,
                      Twitch,
                      Signal,
                      `Stack Exchange`,
                      na.rm = TRUE))

an_SOC3380 <-
  an_SOC3380 %>%
  mutate(region_n = case_when(
    region == "Europe" ~ 1,
    region == "South/Latin America" ~ 2,
    region == "Arab States" ~ 3,
    region == "North America" ~ 4,
    region == "Asia & Pacific" ~ 5,
    region == "Africa" ~ 6,
    region == "Middle east" ~ 7
  ))

#préparation des indices

accord_lvl <- c("Pas du tout d'accord", 
                "Plutôt pas d'accord", 
                "Plutôt d'accord", 
                "D'accord", 
                "Tout à fait d'accord")

INV_accord_LVL <- c("Tout à fait d'accord", 
                    "D'accord", 
                    "Plutôt d'accord", 
                    "Plutôt pas d'accord", 
                    "Pas du tout d'accord")

mesure_lvl <- c("Pas du tout", 
                "Dans une certaine mesure", 
                "Dans une grande mesure")

IND_LVL <- c("Faible", 
             "Modéré", 
             "Fort")

#variables indépendantes

##rs_disc_diff

an_SOC3380$rs_disc_diff <- recode(an_SOC3380$rs_disc_diff, "Pas d'accord" = "Plutôt pas d'accord")
an_SOC3380$rs_disc_diff <- factor(an_SOC3380$rs_disc_diff, levels = accord_lvl, ordered = TRUE)
an_SOC3380$rs_disc_diff <- as.numeric(an_SOC3380$rs_disc_diff, na.rm = TRUE)

##rs_renc_diff

an_SOC3380$rs_renc_diff <- recode(an_SOC3380$rs_renc_diff, "Pas d'accord" = "Plutôt pas d'accord")
an_SOC3380$rs_renc_diff <- factor(an_SOC3380$rs_renc_diff, levels = accord_lvl, ordered = TRUE)
an_SOC3380$rs_renc_diff <- as.numeric(an_SOC3380$rs_renc_diff, na.rm = TRUE)

##disc_maj_same

an_SOC3380$disc_maj_same <- factor(an_SOC3380$disc_maj_same, levels = accord_lvl, ordered = TRUE)
an_SOC3380$disc_maj_same <- as.numeric(an_SOC3380$disc_maj_same, na.rm = TRUE)

##disc_maj_migr

an_SOC3380$disc_maj_migr <- factor(an_SOC3380$disc_maj_migr, levels = accord_lvl, ordered = TRUE)
an_SOC3380$disc_maj_migr <- as.numeric(an_SOC3380$disc_maj_migr, na.rm = TRUE)

##disc_maj_qc

an_SOC3380$disc_maj_qc <- factor(an_SOC3380$disc_maj_qc, levels = accord_lvl, ordered = TRUE)
an_SOC3380$disc_maj_qc <- as.numeric(an_SOC3380$disc_maj_qc, na.rm = TRUE)

## online_memb_migr

an_SOC3380$online_memb_migr <- factor(an_SOC3380$online_memb_migr, levels = c("Oui", "Non"), ordered = FALSE)

#variables dépendantes

##sentim_app

an_SOC3380$sentim_app <- recode(an_SOC3380$sentim_app, "Pas d'accord" = "Plutôt pas d'accord")
an_SOC3380$sentim_app <- factor(an_SOC3380$sentim_app, levels = accord_lvl, ordered = TRUE)
an_SOC3380$sentim_app <- as.numeric(an_SOC3380$sentim_app, na.rm = TRUE)

##chez_moi_qc

an_SOC3380$chez_moi_qc <- recode(an_SOC3380$chez_moi_qc, "Pas d'accord" = "Plutôt pas d'accord")
an_SOC3380$chez_moi_qc <- factor(an_SOC3380$chez_moi_qc, levels = accord_lvl, ordered = TRUE)
an_SOC3380$chez_moi_qc <- as.numeric(an_SOC3380$chez_moi_qc, na.rm = TRUE)

##feel_queb

an_SOC3380$feel_queb <- recode(an_SOC3380$feel_queb, "Pas d'accord" = "Plutôt pas d'accord")
an_SOC3380$feel_queb <- factor(an_SOC3380$feel_queb, levels = accord_lvl, ordered = TRUE)
an_SOC3380$feel_queb <- as.numeric(an_SOC3380$feel_queb, na.rm = TRUE)

##forget_orig

an_SOC3380$forget_orig <- recode(an_SOC3380$forget_orig, "Pas d'accord" = "Plutôt pas d'accord")
an_SOC3380$forget_orig <- factor(an_SOC3380$forget_orig, levels = INV_accord_LVL, ordered = TRUE)
an_SOC3380$forget_orig <- as.numeric(an_SOC3380$forget_orig, na.rm = TRUE)

##percep_qc

an_SOC3380$percep_qc <- recode(an_SOC3380$percep_qc, "Pas d'accord" = "Plutôt pas d'accord")
an_SOC3380$percep_qc <- factor(an_SOC3380$percep_qc, levels = accord_lvl, ordered = TRUE)
an_SOC3380$percep_qc <- as.numeric(an_SOC3380$percep_qc, na.rm = TRUE)

##dev_qc

an_SOC3380$dev_qc <- recode(an_SOC3380$dev_qc, "Pas d'accord" = "Plutôt pas d'accord")
an_SOC3380$dev_qc <- factor(an_SOC3380$dev_qc, levels = accord_lvl, ordered = TRUE)
an_SOC3380$dev_qc <- as.numeric(an_SOC3380$dev_qc, na.rm = TRUE)

##agir_qc

an_SOC3380$agir_qc <- recode(an_SOC3380$agir_qc, "Pas d'accord" = "Plutôt pas d'accord")
an_SOC3380$agir_qc <- factor(an_SOC3380$agir_qc, levels = accord_lvl, ordered = TRUE)
an_SOC3380$agir_qc <- as.numeric(an_SOC3380$agir_qc, na.rm = TRUE)

##import_adapt

an_SOC3380$import_adapt <- recode(an_SOC3380$import_adapt, "Pas d'accord" = "Plutôt pas d'accord")
an_SOC3380$import_adapt <- factor(an_SOC3380$import_adapt, levels = accord_lvl, ordered = TRUE)
an_SOC3380$import_adapt <- as.numeric(an_SOC3380$import_adapt, na.rm = TRUE)

##iso_qc

an_SOC3380$iso_qc <- recode(an_SOC3380$iso_qc, "Pas d'accord" = "Plutôt pas d'accord")
an_SOC3380$iso_qc <- factor(an_SOC3380$iso_qc, levels = INV_accord_LVL, ordered = TRUE)
an_SOC3380$iso_qc <- as.numeric(an_SOC3380$iso_qc, na.rm = TRUE)

##valeur_fr

an_SOC3380$valeur_fr <- factor(an_SOC3380$valeur_fr, levels = mesure_lvl, ordered = TRUE)
an_SOC3380$valeur_fr <- as.numeric(an_SOC3380$valeur_fr, na.rm = TRUE)

##valeur_dem

an_SOC3380$valeur_dem <- factor(an_SOC3380$valeur_dem, levels = mesure_lvl, ordered = TRUE)
an_SOC3380$valeur_dem <- as.numeric(an_SOC3380$valeur_dem, na.rm = TRUE)

##valeur_equal

an_SOC3380$valeur_equal <- factor(an_SOC3380$valeur_equal, levels = mesure_lvl, ordered = TRUE)
an_SOC3380$valeur_equal <- as.numeric(an_SOC3380$valeur_equal, na.rm = TRUE)

##valeur_med14

an_SOC3380$valeur_med14 <- factor(an_SOC3380$valeur_med14, levels = mesure_lvl, ordered = TRUE)
an_SOC3380$valeur_med14 <- as.numeric(an_SOC3380$valeur_med14, na.rm = TRUE)

##valeur_relation

an_SOC3380$valeur_relation <- factor(an_SOC3380$valeur_relation, levels = mesure_lvl, ordered = TRUE)
an_SOC3380$valeur_relation <- as.numeric(an_SOC3380$valeur_relation, na.rm = TRUE)

##valeur_laic

an_SOC3380$valeur_laic <- factor(an_SOC3380$valeur_laic, levels = mesure_lvl, ordered = TRUE)
an_SOC3380$valeur_laic <- as.numeric(an_SOC3380$valeur_laic, na.rm = TRUE)

##valeur_droitlib

an_SOC3380$valeur_droitlib <- factor(an_SOC3380$valeur_droitlib, levels = mesure_lvl, ordered = TRUE)
an_SOC3380$valeur_droitlib <- as.numeric(an_SOC3380$valeur_droitlib, na.rm = TRUE)

##comp_polqc

an_SOC3380$comp_polqc <- factor(an_SOC3380$comp_polqc, levels = c("Pas bonne du tout", 
                                                                  "Pas bonne", 
                                                                  "Plutôt bonne", 
                                                                  "Bonne", 
                                                                  "Très bonne"), ordered = TRUE)

an_SOC3380$comp_polqc <- as.numeric(an_SOC3380$comp_polqc, na.rm = TRUE)

##disc_polqc

an_SOC3380$disc_polqc <- recode(an_SOC3380$disc_polqc, "Rarement" = "Quelques fois")

an_SOC3380$disc_polqc <- factor(an_SOC3380$disc_polqc, levels = c("Jamais", 
                                                                  "Quelques fois", 
                                                                  "Plusieurs fois", 
                                                                  "Très souvent", 
                                                                  "Tout le temps"), ordered = TRUE)

an_SOC3380$disc_polqc <- as.numeric(an_SOC3380$disc_polqc, na.rm = TRUE)

##interp_polqc

an_SOC3380$interp_polqc <- factor(an_SOC3380$interp_polqc, levels = c("Pas interpelé(e)", 
                                                                      "Pas vraiment interpelé(e)", 
                                                                      "Plutôt interpelé(e)", 
                                                                      "Interpelé(e)", 
                                                                      "Très interpelé(e)"), ordered = TRUE)

an_SOC3380$interp_polqc <- as.numeric(an_SOC3380$interp_polqc, na.rm = TRUE)

##qc_val_fr

an_SOC3380$qc_val_fr <- factor(an_SOC3380$qc_val_fr, levels = mesure_lvl, ordered = TRUE)
an_SOC3380$qc_val_fr <- as.numeric(an_SOC3380$qc_val_fr, na.rm = TRUE)

##qc_val_dem

an_SOC3380$qc_val_dem <- factor(an_SOC3380$qc_val_dem, levels = mesure_lvl, ordered = TRUE)
an_SOC3380$qc_val_dem <- as.numeric(an_SOC3380$qc_val_dem, na.rm = TRUE)

##qc_val_equal

an_SOC3380$qc_val_equal <- factor(an_SOC3380$qc_val_equal, levels = mesure_lvl, ordered = TRUE)
an_SOC3380$qc_val_equal <- as.numeric(an_SOC3380$qc_val_equal, na.rm = TRUE)

##qc_val_med14

an_SOC3380$qc_val_med14 <- factor(an_SOC3380$qc_val_med14, levels = mesure_lvl, ordered = TRUE)
an_SOC3380$qc_val_med14 <- as.numeric(an_SOC3380$qc_val_med14, na.rm = TRUE)

##qc_val_relation

an_SOC3380$qc_val_relation <- factor(an_SOC3380$qc_val_relation, levels = mesure_lvl, ordered = TRUE)
an_SOC3380$qc_val_relation <- as.numeric(an_SOC3380$qc_val_relation, na.rm = TRUE)

##qc_val_laic

an_SOC3380$qc_val_laic <- factor(an_SOC3380$qc_val_laic, levels = mesure_lvl, ordered = TRUE)
an_SOC3380$qc_val_laic <- as.numeric(an_SOC3380$qc_val_laic, na.rm = TRUE)

##qc_val_droitlib

an_SOC3380$qc_val_droitlib <- factor(an_SOC3380$qc_val_droitlib, levels = mesure_lvl, ordered = TRUE)
an_SOC3380$qc_val_droitlib <- as.numeric(an_SOC3380$qc_val_droitlib, na.rm = TRUE)

# Contrôle

an_SOC3380$rs_time <- factor(an_SOC3380$rs_time, levels = c("Moins d'une heure", 
                                                            "Entre une et trois heures", 
                                                            "Entre trois et cinq heures", 
                                                            "Entre cinq et sept heures", 
                                                            "Plus de sept heures"), ordered = TRUE)

an_SOC3380$rs_time_n <- as.numeric(an_SOC3380$rs_time, na.rm = TRUE)

## âge

an_SOC3380$age <- factor(an_SOC3380$age, levels = c("Moins de 18 ans", 
                                                    "Entre 18 et 25 ans", 
                                                    "Entre 26 et 35 ans", 
                                                    "Entre 36 et 50 ans", 
                                                    "Plus de 50 ans"), ordered = TRUE)

an_SOC3380$age_n <- as.numeric(an_SOC3380$age, na.omit(NA))

v_sentimapp <- c("sentim_app", 
                 "chez_moi_qc", 
                 "feel_queb", 
                 "forget_orig", 
                 "percep_qc", 
                 "dev_qc", 
                 "agir_qc", 
                 "import_adapt", 
                 "iso_qc")

v_NV <- c("valeur_fr", 
          "valeur_dem", 
          "valeur_equal", 
          "valeur_med14", 
          "valeur_relation", 
          "valeur_laic", 
          "valeur_droitlib")

v_rsContact <- c("rs_disc_diff", 
                 "disc_maj_qc", 
                 "rs_renc_diff")

v_percepQC <- c("qc_val_fr", 
                "qc_val_dem", 
                "qc_val_equal", 
                "qc_val_med14", 
                "qc_val_relation", 
                "qc_val_laic", 
                "qc_val_droitlib")

accord_lvl <- c("Pas du tout d'accord", 
                "Plutôt pas d'accord", 
                "Plutôt d'accord", 
                "D'accord", 
                "Tout à fait d'accord")

an_SOC3380$rs_inf <- factor(an_SOC3380$rs_inf, levels = accord_lvl, ordered = TRUE)

an_SOC3380$rs_inf_n <- as.numeric(an_SOC3380$rs_inf, na.omit(NA))

an_SOC3380$ton_migr <- recode(an_SOC3380$ton_migr, "Très positive" = "Positive")

an_SOC3380$ton_migr <- recode(an_SOC3380$ton_migr, "Très négative" = "Négsative")

ton_migr_lvl <- c("Négative", 
                  "Plutôt négative", 
                  "Neutre", 
                  "Plutôt positive", 
                  "Positive")

an_SOC3380$ton_migr <- factor(an_SOC3380$ton_migr, levels = ton_migr_lvl, ordered = TRUE)
an_SOC3380$ton_migr_n <- as.numeric(an_SOC3380$ton_migr, na.omit(NA))

an_SOC3380$year_qc <- factor(an_SOC3380$year_qc, levels = c("1 an", 
                                                            "2 ans", 
                                                            "3 ans", 
                                                            "4 ans", 
                                                            "5 ans", 
                                                            "6 ans", 
                                                            "7 ans", 
                                                            "8 ans", 
                                                            "9 ans", 
                                                            "10 ans", 
                                                            "Plus de 10 ans"), ordered = TRUE)

an_SOC3380$year_qc_n <- as.numeric(an_SOC3380$year_qc, na.omit(NA))

v_score_ExpoDisc <- c("rs_time_n", 
                      "ton_migr_n", 
                      "rs_inf_n")

v_media_accueil <- c("Youtube", 
                     "Instagram", 
                     "TikTok", 
                     "Snapchat", 
                     "Pinterest", 
                     "Piscart", 
                     "Twitch", 
                     "Twitter", 
                     "Reddit", 
                     "Quora", 
                     "Teams", 
                     "Discord", 
                     "Stack Exchange", 
                     "Facebook", 
                     "Messenger", 
                     "Telegram", 
                     "LinkedIn", 
                     "Skype", 
                     "Viber", 
                     "imo", 
                     "Signal")

v_media_diasp <- c("Douyin", 
                   "Kuaishou", 
                   "Likee", 
                   "Tieba", 
                   "WeChat", 
                   "QQ", 
                   "Qzone", 
                   "Line", 
                   "WhatsApp")

#création des indices

an_SOC3380 <-
  an_SOC3380 %>%
  rowwise() %>%
  mutate(score_sentimapp = (sum(c_across(all_of(v_sentimapp)), na.rm = TRUE) - length(v_sentimapp)) / (4*length(v_sentimapp))) %>%
  mutate(score_sentimapp = if (score_sentimapp < 0) NA_real_ else score_sentimapp) %>%
  ungroup()

an_SOC3380 <- 
  an_SOC3380 %>%
  rowwise() %>%
  mutate(score_rsContact = (sum(c_across(all_of(v_rsContact)), na.rm = TRUE) -length(v_rsContact)) / (4*length(v_rsContact))) %>%
  mutate(score_rsContact = if (score_rsContact < 0)  NA_real_ else score_rsContact) %>%
  ungroup()

an_SOC3380 <- 
  an_SOC3380 %>%
  rowwise() %>%
  mutate(score_NV = (sum(c_across(all_of(v_NV)), na.rm = TRUE) -length(v_NV)) / (2*length(v_NV))) %>%
  mutate(score_NV = if (score_NV < 0) NA_real_ else score_NV) %>%
  ungroup()

an_SOC3380 <- 
  an_SOC3380 %>%
  rowwise() %>%
  mutate(score_percepQC = (sum(c_across(all_of(v_percepQC)), na.rm = TRUE) -length(v_percepQC)) / (2*length(v_percepQC))) %>%
  mutate(score_percepQC = if (score_percepQC < 0) NA_real_ else score_percepQC) %>%
  ungroup()

an_SOC3380 <- 
  an_SOC3380 %>%
  rowwise() %>%
  mutate(score_ExpoDisc = (sum(c_across(all_of(v_score_ExpoDisc)), na.rm = TRUE) - length(v_score_ExpoDisc)) / (4*length(v_score_ExpoDisc))) %>%
  mutate(score_ExpoDisc = if_else(score_ExpoDisc < 0 | score_ExpoDisc == 0, NA_real_, score_ExpoDisc)) %>%
  ungroup()

an_SOC3380 <-
  an_SOC3380 %>%
  rowwise() %>%
  mutate(score_media_accueil = (sum(c_across(all_of(v_media_accueil)), na.rm = TRUE)) / nbr_rs)

an_SOC3380 <- 
  an_SOC3380 %>%
  rowwise() %>%
  mutate(score_media_diasp = (sum(c_across(all_of(v_media_diasp)), na.rm = TRUE)) / nbr_rs)

# Traitement des types de réseaux sociaux
an_SOC3380 <- an_SOC3380 %>%
  mutate(rs_partage_medias = if_else(str_detect(rs_util, "Youtube|Instagram|TikTok|Douyin|Snapchat|Kuaishou|Pinterest|Piscart|Likee|Twitch", negate = FALSE), 1, 0),
         rs_blogues = if_else(str_detect(rs_util, "Twitter|Weibo", negate = FALSE), 1, 0),
         rs_commun_online = if_else(str_detect(rs_util, "Reddit|Quora|Tieba|Teams|Discord|Stack Exchange", negate = FALSE), 1, 0),
         rs_SRS = if_else(str_detect(rs_util, "Facebook|Messenger|WhatsApp|WeChat|Telegram|QQ|Qzone|LinledIn|Skype|Viber|imo|Line|Signal", negate = FALSE), 1, 0))

an_SOC3380 <-
  an_SOC3380 %>%
  mutate(rs_blogues_f = case_when(
    rs_blogues == 0 ~ "Non",
    rs_blogues == 1 ~ "Oui"
  ))

an_SOC3380 <-
  an_SOC3380 %>%
  mutate(rs_commun_online_f = case_when(
    rs_commun_online == 0 ~ "Non",
    rs_commun_online == 1 ~ "Oui"
  ))

an_SOC3380 <-
  an_SOC3380 %>%
  mutate(rs_SRS_f = case_when(
    rs_SRS == 0 ~ "Non",
    rs_SRS == 1 ~ "Oui"
  ))

an_SOC3380 <-
  an_SOC3380 %>%
  mutate(rs_partage_medias_f = case_when(
    rs_partage_medias == 0 ~ "Non",
    rs_partage_medias == 1 ~ "Oui"
  ))

an_SOC3380 <- an_SOC3380 %>%
  mutate(rs_home = if_else(str_detect(rs_util,"Youtube|Instagram|TikTok|Snapchat|Pinterest|Piscart|Twitch|Twitter|Reddit|Quora|Teams|Discord|Stack Exchange|Facebook|Messenger|Telegram|LinkedIn|Skype|Viber|imo|Signal", negate = FALSE), 1, 0))

an_SOC3380 <- an_SOC3380 %>%
  mutate(rs_diasp = if_else(str_detect(rs_util, "Douyin|Kuaishou|Likee|Tieba|WeChat|QQ|Qzone|Line|WhatsApp|Messenger|Skype", negate = FALSE), 1, 0))

an_SOC3380 <-
  an_SOC3380 %>%
  mutate(rs_home_f = case_when(
    rs_home == 0 ~ "Non",
    rs_home == 1 ~ "Oui"
  ))

an_SOC3380 <-
  an_SOC3380 %>%
  mutate(rs_diasp_f = case_when(
    rs_diasp == 0 ~ "Non",
    rs_diasp == 1 ~ "Oui"
  ))

an_SOC3380 <-
  an_SOC3380 %>%
  mutate(year_qc_f = case_when(
    year_qc_n < 6 ~ "1 à 5 ans",
    year_qc_n == 6| year_qc_n == 7|year_qc_n == 8|year_qc_n == 9 ~ "6 à 9 ans",
    year_qc_n > 9 ~ "dix ans et plus"
  ))
```

# Résultats

## H1 - Contact interculturel et sentiment d'appartenance

Le modèle de régression suivant présente la relation entre l'indice de contact
interculturel sur les réseaux sociaux et l'indice du sentiment d'appartenance au
pays d'accueil en fonction du nombre d'années passées au Québec, de la région
d'origine et de l'âge. Les résultats indiquent que la relation entre le contact
interculturel et le sentiment d'appartenance est statistiquement significative
dans la population à l'intervalle de confiance de 95% (p \<= 0.05).

```{r}
local({
  .Table <- lm(score_rsContact ~ score_sentimapp + year_qc_n + region_n + age_n, an_SOC3380)
  pander(.Table, split.cell=80, split.table = Inf, 
         caption = "Modèle de régression contrôlé pour l'hypothèse H1",
         col.names = c("Estimation", "Erreur type", "Valeur t", "Valeur p"),
         row.names = c("(Intercept)", "Sentiment d'appartenance", "Année passées au Québec", "Région d'origine", "Âge")
         )
})
```

En accord avec les résultats du modèle de régression, le test de corrélation de
Pearson suivant présente l'intensité et le sens de la relation étudiée. Les
résultats obtenus indiquent que la relation est faible et positive (C = 0.223).

```{r}
local({
  .Table <- cor.test(an_SOC3380$score_rsContact, an_SOC3380$score_sentimapp)
  pander(.Table, split.cell=80, split.table = Inf, 
         caption = "Test de corrélation de Pearson pour l'hypothèse H1",
         col.names = c("Valeur t", "Degrés de liberté", "Valeur C", "Hypothèse alternatif", "Valeur C")
  )
})
```

Le graphique suivant illustre la relation entre le contact interculturel et le
sentiment d'appartenance. En accord avec les résultats du test de corrélation de
Pearson, on observe que la relation est positive et faible. En d'autres termes :
pour les personnes immigrantes établies au Québec, l'entretien de contacts sur
les réseaux sociaux en ligne avec des personnes qui ne proviennent de leur pays
d'origine est faiblement et positivement corrélé avec le développement d'un
sentiment d'appartenance au pays d'accueil.

```{r}
ggplot(an_SOC3380 %>% filter(!is.na(score_rsContact), !is.na(score_sentimapp)), aes(x= score_rsContact, y = score_sentimapp))+
  geom_point() +
  geom_smooth(method = "lm", se= FALSE)+
  labs(
    title = paste("Figure", no_graphic <- no_graphic+1, ":",
      "Contact interculturel et sentiment d'appartenance"),
    x = "Contact interculturel",
    y = "Sentiment d'appartenance au Québec"
  ) +
  theme(text = element_text(hjust = 0.5))
```

### H1.1 - Pays d'origine, discrimination systémique et sentiment d'appartenance

Les témoignages de nos répondants indiquent que les discours anti-immigrations
présents sur les réseaux sociaux en ligne participent à un processus de
discrimination plus large - en ce qu'il constitue la traduction de pratiques
existant hors ligne - les empêchant de s'identifier pleinement au Québec, et
partant, de développer un sentiment d'appartenance envers leur pays d'accueil.
Ce phénomène est pondéré par les caractéristiques socio-démographiques de nos
répondants, selon un gradient qui rappelle les travaux de Wacquant sur la
*racialisation diagonale* [-@wacquantResolvingTroubleRace2022]. Les personnes
les plus racialisées comme Wendy (jeune femme noire originaire d'Haïti établie
au Québec depuis son enfance) affirment que leur incapacité à s'identifier comme
québécoise est directement liée à la discrimination raciale :

> Wendy: Il y a toujours quelqu'un, à un moment donné dans ta vie, qui va te
> rappeler ta condition d'immigrée. Cela peut être un commerçant par exemple :
> il voit une personne noire entrer et automatiquement, il va commencer à te
> regarder dans toutes les allées pour vérifier si tu ne vas pas voler ; même si
> tu viens régler tes achats à la caisse. Le regard qu'il peut avoir sur toi
> fait toute la différence (...) D'entendre : "toi je t'aime bien, tu n'es pas
> comme les autres." Ça me rappelle que je ne suis pas québécoise à leurs yeux.
>
> Théa: Est-ce que tu te sens Québécoise aujourd'hui ?
>
> Wendy: Non, je ne me sens pas québécoise (ton amusé) et je ne pourrais jamais
> me sentir québécoise à 100%. Je ne pourrais jamais renier mon identité, même
> si je ne peux pas retourner vivre dans mon pays (...) je me suis adaptée (le
> mot est appuyé) [Au Québec].C'est là la différence.

En accord avec les résultats du modèle de régression (Tableau 2), le graphique
suivant illustre la relation entre le contact interculturel et le sentiment
d'appartenance en fonction de la région d'origine. Les résultats révèlent une
particularité à l'endroit des personnes immigrantes originaires du Maghreb, du
Proche-Orient et du Moyen-Orient.

```{r}
ggplot(an_SOC3380 %>% filter(!is.na(score_sentimapp)), aes(x= score_rsContact, y = score_sentimapp, color = region))+
  geom_point() +
  geom_smooth(method = "lm", se= FALSE)+
  theme_minimal()+
  scale_color_brewer(name = "Region", palette = "Set1") +
  labs(
    title = paste("Graphique", no_graphic <- no_graphic+1, ":",
      "Contact interculturel et sentiment d'appartenance"),
    subtitle = "Contrôlé par la région",
    x = "Contact interculturel en ligne",
    y = "Sentiment d'appartenance au Québec"
  ) +
  theme(text = element_text(hjust = 0.5))
```

En accord avec les résultats du modèle de régression (Tableau 2), les graphiques
suivants illustrent la relation entre le contact interculturel et le sentiment
d'appartenance en fonction de l'âge et du nombre d'années passées au Québec.
Dans le cas de l'âge, on peut observer que la relation est annulée, sauf parmi
les répondants âgés de 36 à 50 ans. Dans le cas du nombre d'années passées au
Québec on peut observer que la relation n'est pas affectée chez les personnes
établies depuis moins de 6 ans, qu'elle est accentuée chez les personnes
établies depuis 6 à 9 ans, puis s'inverse chez les personnes établies depuis 10
ans ou plus.

```{r}
ggplot(an_SOC3380 %>% filter(!is.na(score_sentimapp)), aes(x= score_rsContact, y = score_sentimapp, color = age))+
  geom_point() +
  geom_smooth(method = "lm", se= FALSE)+
  theme_minimal()+
  scale_color_brewer(name = "Âge", palette = "Set1") +
  labs(
    title = paste("Figure", no_graphic <- no_graphic+1, ":",
      "Contact interculturel et sentiment d'appartenance"),
    subtitle = "Contrôlé par l'âge",
    x = "Contact interculturel en ligne",
    y = "Sentiment d'appartenance au Québec"
  ) +
  theme(text = element_text(hjust = 0.5))
```

```{r}
ggplot(an_SOC3380 %>% filter(!is.na(score_sentimapp)), aes(x= score_rsContact, y = score_sentimapp, color = year_qc_f))+
  geom_point() +
  geom_smooth(method = "lm", se= FALSE)+
  theme_minimal()+
  scale_color_brewer(name = "Nombre d'années passées au Québec", palette = "Set1") +
  labs(
    title = paste("Figure", no_graphic <- no_graphic+1, ":",
      "Contact interculturel et sentiment d'appartenance"),
    subtitle = "Contrôlé par le nombre d'année passées au Québec",
    x = "Contact interculturel en ligne",
    y = "Sentiment d'appartenance au Québec"
  ) +
  theme(text = element_text(hjust = 0.5))
```

### H1.2 - Type de réseaux sociaux en ligne et sentiment d'appartenance

Le modèle de régression suivant présente la relation entre le contact
interculturel et le sentiment d'appartenance en fonction du type de réseau
social utilisé. Les résultats suggèrent que l'utilisation de blogs comme
Twitter, de sites de réseaux sociaux comme Facebook ou encore de sites de
partage de médias comme YouTube exercent une influence sur la relation étudiée.

```{r}
local({
  .Table <- lm(score_rsContact ~ score_sentimapp + rs_blogues + rs_commun_online + rs_SRS + rs_partage_medias, an_SOC3380)
  pander(.Table, split.cell=80, split.table = Inf, 
         caption = "Modèle de régression contrôlé pour l'hypothèse H1.2",
         col.names = c("Estimation", "Erreur type", "Valeur t", "Valeur p"),
         row.names = c("(Intercept)", "Sentiment d'appartenance", "Blogs", "Communautés en ligne", "Sites de réseaux sociaux", "Sites de partage de médias")
  )
})
```

En accord avec les résultats du modèle de régression, le graphique suivant
illustre la relation entre le contact interculturel et le sentiment
d'appartenance en fonction de l'utilisation de blogs en ligne. Les résultats
indiquent que l'utilisation de blogs comme Twitter modifie le sens de la
relation étudiée. Autrement dit, l'usage de blogs tendrait à freiner le
développement d'un sentiment d'appartenance au Québec chez les personnes
immigrantes.

```{r}
ggplot(an_SOC3380 %>% filter(!is.na(score_sentimapp)), aes(x= score_rsContact, y = score_sentimapp, color = rs_blogues_f))+
  geom_point() +
  geom_smooth(method = "lm", se= FALSE)+
  theme_minimal()+
  scale_color_brewer(name = "Utilisation de blogs en ligne", palette = "Set1") +
  labs(
    title = paste("Figure", no_graphic <- no_graphic+1, ":",
      "Contact interculturel et sentiment d'appartenance"),
    subtitle = "Contrôlé par l'utilisation de blogs en ligne",
    x = "Score du contact interculturel en ligne",
    y = "Score du sentiment d'appartenance au Québec"
  ) +
  theme(text = element_text(hjust = 0.5))
```

En accord avec les résultats du modèle de régression, le graphique suivant
illustre la relation entre le contact interculturel et le sentiment
d'appartenance en fonction de l'utilisation de sites de réseaux sociaux en
ligne. Les résultats suggèrent le phénomène inverse. En d'autres termes,
l'utilisation de plateformes comme Facebook faciliterait le développement d'un
sentiment d'appartenance au Québec chez les personnes immigrantes. Néanmoins,
nos résultats ne peuvent être validés, étant entendu que la majorité de nos
répondants utilisent des sites de réseaux sociaux en ligne.

Néanmoins, on peut observer sur le graphique que la relation étudiée est plus
intense lorsqu'elle est contrôlée par l'utilisation des sites de réseaux sociaux
en ligne.

```{r}
ggplot(an_SOC3380 %>% filter(!is.na(score_sentimapp)), aes(x= score_rsContact, y = score_sentimapp, color = rs_SRS_f))+
  geom_point() +
  geom_smooth(method = "lm", se= FALSE)+
  theme_minimal()+
  scale_color_brewer(name = "Utilisation de sites de réseaux sociaux en ligne", palette = "Set1") +
  labs(
    title = paste("Figure", no_graphic <- no_graphic+1, ":",
      "Contact interculturel et sentiment d'appartenance"),
    subtitle = "Contrôlé par le type d'utilisation de sites de réseaux sociaux en ligne",
    x = "Score du contact interculturel en ligne",
    y = "Score du sentiment d'appartenance au Québec") +
  theme(text = element_text(hjust = 0.5))
```

En accord avec les résultats du modèle de régression, le graphique suivant
illustre la relation entre le contact interculturel et le sentiment
d'appartenance en fonction de l'utilisation de sites de partage de médias en
ligne. Les résultats indiquent que l'utilisation de sites de partage de médias
atténue l'intensité de la relation étudiée. En d'autres termes, à l'instar des
blogs, l'utilisation de plateformes comme YouTube tend à freiner le
développement d'un sentiment d'appartenance au Québec.

```{r}
ggplot(an_SOC3380 %>% filter(!is.na(score_sentimapp)), aes(x= score_rsContact, y = score_sentimapp, color = rs_partage_medias_f))+
  geom_point() +
  geom_smooth(method = "lm", se= FALSE)+
  theme_minimal()+
  scale_color_brewer(name = "Utilisation de sites de partage de médias", palette = "Set1") +
  labs(
    title = paste("Figure", no_graphic <- no_graphic+1, ":",
      "Contact interculturel et sentiment d'appartenance"),
    subtitle = "Contrôlé par l'utilisation de site de partage de médias",
    x = "Score du contact interculturel en ligne",
    y = "Score du sentiment d'appartenance au Québec"
  ) +
  theme(text = element_text(hjust = 0.5))
```

### H1.3 - Type d'utilisation des réseaux et sentiment d'appartenance

Le modèle de régression suivant présente la relation entre le contact
interculturel et le sentiment d'appartenance en fonction de du type
d'utilisation des réseaux sociaux en ligne. Pour parvenir à ses résultats, nous
avons construit un indice supplémentaire permettant de mesurer la proportion des
réseaux sociaux traditionnellement utilisés pour communiquer avec la famille et
les amis restés au pays.

```{r}
local({
  .Table <- lm(score_rsContact ~ score_sentimapp + score_media_accueil + score_media_diasp, an_SOC3380)
  pander(.Table, split.cell=80, split.table = Inf, 
         caption = "Modèle de régression contrôlé pour l'hypothèse H1.3",
         col.names = c("Estimation", "Erreur type", "Valeur t", "Valeur p"),
         row.names = c("(Intercept)", "Sentiment d'appartenance", "Utilisation au Québec", "Utilisation mixte")
  )
})
```

Les résultats suggèrent que le type d'utilisation des réseaux sociaux joue un
rôle dans la relation étudiée. Le graphique suivant illustre la relation entre
le contact interculturel et le sentiment d'appartenance en fonction du type
d'utilisation (mixte ou exclusivement au Québec). On observe que l'utilisation
exclusive de réseaux sociaux permettant de communiquer au Québec modifie le sens
de la relation étudiée.

## H2 - Exposition à du contenu lié à l'immigration et sentiment d'appartenance

```{r}
ggplot(an_SOC3380 %>% filter(!is.na(score_sentimapp)), aes(x= score_rsContact, y = score_sentimapp, color = rs_diasp_f))+
  geom_point() +
  geom_smooth(method = "lm", se= FALSE)+
  theme_minimal()+
  scale_color_brewer(name = "Utilisation mixte", palette = "Set1") +
  labs(
    title = paste("Figure", no_graphic <- no_graphic+1, ":",
      "Contact interculturel et sentiment d'appartenance"),
    subtitle = "Contrôlé par l'utilisation mixte",
    x = "Score du contact interculturel en ligne",
    y = "Score du sentiment d'appartenance au Québec"
  ) +
  theme(text = element_text(hjust = 0.5))
```

Le modèle de régression suivant présente la relation entre l'exposition à du
contenu lié à l'immigration et le sentiment d'appartenance en fonction du nombre
d'années passées au Québec, de la région d'origine et de l'âge. Les résultats
indiquent que la relation est "pratiquement" significative dans la population à
l'intervalle de confiance 95% (p \~ 0.05).

```{r}
local({
  .Table <- lm(score_ExpoDisc ~ score_sentimapp + year_qc_n + region_n + age_n, an_SOC3380)
  pander(.Table, split.cell=80, split.table = Inf, 
         caption = "Modèle de régression contrôlé pour l'hypothèse H2",
         col.names = c("Estimation", "Erreur type", "Valeur t", "Valeur p"),
         row.names = c("(Intercept)", "Sentiment d'appartenance", "Année passés au Québec", "Région d'origine", "Âge")
  )
})
```

En accord avec les résultats du modèle de régression, le test de corrélation de
Pearson suivant présente l'intensité et le sens de la relation entre
l'exposition à du contenu lié à l'immigration et le sentiment d'appartenance.
Les résultats obtenus indiquent que la relation faible et positive (C = 0.157).

```{r}
local({
  .Table <- cor.test(an_SOC3380$score_ExpoDisc, an_SOC3380$score_sentimapp)
  pander(.Table, split.cell=80, split.table = Inf, 
         caption = "Test de corrélation de Pearson pour l'hypothèse H2",
         col.names = c("Valeur t", "Degrés de liberté", "Valeur C", "Hypothèse alternatif", "Valeur C")
  )
})
```

Le graphique suivant illustre la relation entre l'exposition à du contenu lié à
l'immigration et le sentiment d'appartenance. En accord avec les résultats du
test de corrélation de Pearson, on observe que la relation est positive et
faible. En d'autres termes : pour les personnes immigrantes établies au Québec,
l'exposition à du contenu lié à l'immigration sur les réseaux sociaux en ligne
est faiblement et positivement corrélé au développement d'un sentiment
d'appartenance au pays d'accueil.

```{r}
ggplot(an_SOC3380 %>% filter(!is.na(score_ExpoDisc), !is.na(score_sentimapp)), aes(x= score_ExpoDisc, y = score_sentimapp))+
 geom_point() +
 geom_smooth(method = "lm", se= FALSE)+
 labs(
   title = paste("Figure", no_graphic <- no_graphic+1, ":",
      "Exposition aux discours liés à l'immigration et sentiment d'appartenance"),
   x = "Exposition aux discours liés à l'immigration",
   y = "Sentiment d'appartenance au Québec"
  ) +
  theme(text = element_text(hjust = 0.5))
```

En accord avec les résultats du modèle de régression, les deux graphiques
suivants illustrent les effets de l'âge et du nombre d'années passées au Québec
sur la relation étudiée. Les résultats indiquent que le phénomène est plus
prononcé chez les personnes âgées de 36 ans et plus et chez les personnes
établies au Québec depuis plus de 5 ans.

```{r}
ggplot(an_SOC3380 %>% filter(!is.na(score_sentimapp), !is.na(score_ExpoDisc)), aes(x= score_ExpoDisc, y = score_sentimapp, color = age))+
  geom_point() +
  geom_smooth(method = "lm", se= FALSE)+
  theme_minimal()+
  scale_color_brewer(name = "Âge", palette = "Set1") +
  labs(
    title = paste("Figure", no_graphic <- no_graphic+1, ":",
      "Exposition aux discours liés à l'immigration et sentiment d'appartenance"),
    subtitle = "Contrôlé par l'âge",
    x = "Exposition aux discours liés à l'immigration", y = "Sentiment d'appartenance au Québec"
  ) +
  theme(text = element_text(hjust = 0.5))
```

```{r}
ggplot(an_SOC3380 %>% filter(!is.na(score_sentimapp), !is.na(score_ExpoDisc)), aes(x= score_ExpoDisc, y = score_sentimapp, color = year_qc_f))+
  geom_point() +
  geom_smooth(method = "lm", se= FALSE)+
  theme_minimal()+
  scale_color_brewer(name = "Années passées au Québec", palette = "Set1") +
  labs(
    title = paste("Figure", no_graphic <- no_graphic+1, ":",
      "Exposition aux discours liés à l'immigration et sentiment d'appartenance"),
    subtitle = "Contrôlé par le nombre d'années passées au Québec",
    x = "Exposition aux discours liés à l'immigration",
    y = "Sentiment d'appartenance au Québec"
  ) +
  theme(text = element_text(hjust = 0.5))
```

\newpage

<!--chapter:end:05-resultats.Rmd-->

---
knit: bookdown::render_book
editor_options: 
  markdown: 
    wrap: 80
---

# Discussion

## H1 - Contact interculturel et sentiment d'appartenance

Le volet qualitatif de notre enquête vient confirmer l'existence de la
corrélation entre le contact interculturel et le sentiment d'appartenance tout
en fournissant des clés d'explication suggérant l'existence d'une relation
causale entre les deux variables. Si le travail de Gallant et Friche
[-@gallantEtreIciLabas2010] rappellent que les réseaux sociaux en ligne
participent peu à la création de réseaux d'appartenance préexistants, les
témoignages que nous avons recueillis mentionnent que certaines plateformes
comme Facebook, particulièrement actives au Québec, permettent aux personnes
immigrantes de pérenniser des liens établis hors ligne avec les Québécois.

> Nina: Les évènements de Facebook sont un peu morts en France(...) alors qu'ici
> justement, les gens sont beaucoup plus actifs, surtout dans le Marketplace
> (...) J'ai changé ma vision de Facebook, parce que son utilisation est très
> différente ici.

## H1.1 - Pays d'origine, discrimination systémique et sentiment d'appartenance

Le phénomène observé dans le graphique représentant la relation entre le contact
interculturel et le sentiment d'appartenance contrôlé par la région d'origine
pourrait illustrer le développement de discours islamophobes au Québec, comme le
suggèrent les travaux de Dalila Awada [-@awadaRacialisationMusulmanEs2020] sur
la racialisation des personnes musulmanes dans la province.

## H1.2 - Type de réseaux sociaux en ligne et sentiment d'appartenance

Les résultats obtenus concernant la relation entre contact interculturel et
sentiment d'appartenance en fonction du type de réseau social utilisé
[@kietzmannSocialMediaGet2011] pourraient confirmer les travaux de Hopster et
Ribeiro [-@hopsterMutualAffordancesDynamics2021, p. 559;
-@ribeiroAuditingRadicalizationPathways2020, p. 140] concernant le rôle des
algorithmes de suggestion dans la diffusion de discours xénophobes sur YouTube.
Ces résultats font également échos aux travaux de Dehgan et Walsh
[-@dehghanDynamicsPolarisationAustralian2022, p.64‑65;
-@walshDigitalNativismTwitter2021, p.5] sur la racialisation discursive et le
développement du néonativisme sur Twitter.

Selon Esser [@esserDoesNewImmigration2004], p. 1131], le sentiment
d'appartenance constitue l'une des quatre dimensions de l'intégration
culturelle. Or, les résultats de notre analyse thématique révèlent que nos
répondants mobilisent les réseaux sociaux pour accumuler du capital culturel et
social valorisable dans le cadre de contacts interculturels en ligne et hors
ligne.
\newpage

> Charly: Au niveau de l'intégration à la société d'accueil, ça a été un pilier
> fondamental (...) avoir accès à des memes, connaître des célébrités, mettons
> Occupation Double... des trucs comme ça, qui te permettent d'avoir des
> conversations [avec des Québécois].

Cette association est cependant pondérée par les résultats de notre analyse
thématique. En effet, certains répondants témoignent du rôle positif joué par
YouTube dans le cadre de leur intégration culturelle. Adriana (une jeune femme
originaire d'Amérique du Sud) avance que le contenu présent sur la plateforme a
participé à son apprentissage du français québécois et de l'histoire de la
province :

> Adriana: [Sur YouTube] je cherchais les sous-titres en français avec un petit
> dictionnaire à côté de mon cahier (...) J'essayais de comprendre en arrêtant
> de lire [les sous-titres] et faire des comparaisons avec les apostrophes,
> essayer de comprendre la dynamique du parler le français québécois (...) Mais
> quand je suis sortie dans la rue, je n'ai rien compris (...) J'ai beaucoup
> appris de l'histoire québécoise grâce à une chaîne [YouTube] en particulier
> (...) comprendre la fondation du Québec, la relation que le pays entretien
> avec les autochtones.

## H1.3 - Type d'utilisation des réseaux et sentiment d'appartenance

L'écart constaté pour la relation entre le contact interculturel et le sentiment
d'appartenance en fonction du type d'utilisation des réseaux sociaux pourrait
s'expliquer en creux par la capacité des réseaux sociaux en ligne à réduire le
stress acculturatif généré par la migration. En gardant un contact affectif et
culturel avec leur pays d'origine, les personnes immigrantes seraient plus
enclines à s'intégrer culturellement au Québec, et partant, à développer un
sentiment d'appartenance à leur pays d'accueil.

Cette hypothèse spécifique peut être mise en relation avec les résultats de
l'analyse thématique, qui semblent confirmer le pouvoir stabilisateur du lien
maintenu par nos répondants avec leur pays d'origine sur les réseaux sociaux en
ligne. En effet, Charly (un jeune homme originaire d'Amérique du Sud) et Nina
(une jeune femme originaire de France) affirment que les sites de réseaux
sociaux comme WhatsApp et Messenger leur permettent de garder un lien vital avec
la famille et les amis restés au pays :

> Charly: Si j'avais pas pu parlé à ma mère tous les jours, c'est sûr
> qu'éventuellement, j'aurais pété un plomb, puis j'aurais fait « ciao ». (...)
> Je dirais que c'est fondamental, je pense, dans mes relations sociales. Quand
> je dis « fondamental », c'est que je n'imagine pas les choses autrement, pour
> que ce soit comme c'est aujourd'hui. Peut-être que je serais moins intégré si
> je n'avais pas eu accès [aux] réseaux sociaux, si je n'avais pas pu, par
> exemple, à un moment donné, pu entretenir des relations qui étaient très
> importantes pour moi (...) Ça m'a permis de garder le lien pendant ces 11 ans,
> toujours avec les mêmes personnes. Ma famille on est toujours en contact, on
> s'envoi des photos au quotidien, des trucs comme ça (...) malgré la
> distance,on entretient quand même cette relation.
>
> Nina: [exclamation] Ah ! les groupes familiaux, c'est quelque chose de
> pratique, on n'a pas besoin d'interagir pour voir comment les gens vont. Les
> gens enrichissent les groupes de ce qui leur arrive. Donc en fait, je n'ai
> même pas besoin de poser des questions, j'arrive à suivre ce qui se passe dans
> le groupe.
>
> Nina: [exclamation] Ah ! les groupes familiaux, c'est quelque chose de
> pratique, on n'a pas besoin d'interagir pour voir comment les gens vont. Les
> gens enrichissent les groupes de ce qui leur arrive. Donc en fait, je n'ai
> même pas besoin de poser des questions, j'arrive à suivre ce qui se passe dans
> le groupe.

Néanmoins, la causalité (même indirecte) du rôle joué par le type d'utilisation
des réseaux sociaux dans la relation étudiée ne peut être confirmée au vu du
faible nombre de répondants utilisant les réseaux sociaux en ligne uniquement
pour communiquer au Québec.

De plus, ces résultats entrent en contradiction avec les travaux de Dilshad
Dayani [-@dayaniRoleSocialMedia2017], qui suggèrent que l'utilisation des
réseaux sociaux pour garder contact avec le pays d'origine tendent à freiner le
développement d'un sentiment d'appartenance à la société d'accueil chez les
personnes immigrantes. Dayani avançait que l'utilisation de Facebook affecte
positivement le processus d'acculturation, alors que nos résultats suggèrent
l'effet inverse.

## H2 - Exposition à du contenu lié à l'immigration et sentiment d'appartenance

Malgré sa faible significativité statistique, la relation entre l'exposition à
du contenu lié à l'immigration en ligne et le sentiment d'appartenance peut être
partiellement expliquée par les résultats de l'analyse thématique. D'un côté, le
discours lié à l'immigration au Québec présent sur les réseaux sociaux en ligne
est ambivalent, comme le démontre Wendy :

> Théa: Que peux-tu me dire des publications liées à l'immigration que tu vois
> sur les réseaux ?
>
> Wendy: Je vois deux choses : d'un côté il y en a une majorité qui disent que
> c'est bien de s'installer au Canada ; de l'autre, il y a des
> généralisations\... comme quoi certaines populations ne font rien, ne
> travaillent pas ou bien s'approprient des emplois qui devraient revenir aux
> Québécois.

De l'autre côté, les personnes immigrantes semblent prioriser l'utilisation des
réseaux sociaux en ligne pour collecter des informations concernant leur pays
d'accueil, comme nous le montrent Mathilde (une jeunes femmes originaires de
France).

> Antoine: Les nouvelles sur l'immigration plus négatives, tu les entends en
> dehors des réseaux ?
>
> Mathilde: Ouais, quand tu vois les nouvelles dans le métro, dans la rue ou à
> la radio\... des choses comme ça. C'est des choses qui me parviennent sans que
> j'aille les chercher (\...) J'ai l'impression que tu n'as pas besoin de
> chercher de l'information au sujet de l'immigration pour être tenue au courant
> ici.

# Conclusion

Cette étude cherchait à comprendre la relation entre l'utilisation des réseaux
sociaux en ligne et l'intégration culturelle des personnes immigrantes établies
au Québec depuis plus d'un an. Les résultats du volet quantitatif de notre
enquête indiquent que le contact interculturel sur les réseaux sociaux en ligne
est faiblement et positivement corrélé au développement d'un sentiment
d'appartenance au Québec chez nos répondants. Dans le même temps, l'exposition à
du contenu lié à l'immigration en ligne et le développement d'un sentiment
d'appartenance au Québec sont également faiblement et positivement corrélés. En
outre, le volet qualitatif de notre enquête vient pondérer et spécifier ces deux
conclusions au regard des caractéristiques socio-démographiques de nos
répondants, notamment à l'endroit des discriminations anti-immigrants, qu'elles
soient de nature nationales, ethniques ou raciales.

À la lumière de nos résultats, une étude focalisée autour des effets de la
discrimination envers les personnes immigrantes sur le développement d'un
sentiment d'appartenance au Québec pourrait éclairer avec précision les
dynamiques de racialisation en cours sur les réseaux sociaux en ligne. Pour ce
faire, la création d'un indice de la discrimination et/ou de la catégorisation
(qu'elle soit nationale, ethnique ou raciale) permettrait de mesurer directement
la relation que celle-ci entretien avec le développement d'un sentiment
d'appartenance au Québec.

\newpage

<!--chapter:end:06-discussions.Rmd-->

---
knit: bookdown::render_book
editor_options: 
  markdown: 
    wrap: 80
---
# Références

<!--chapter:end:07-références.Rmd-->

### 2023-05-02 Expérimentations (Théa)/2. Expérimentations (Bookdown)/Bookdown/index.Rmd
Multiple changes on Bookdown

---
title: "COVID-19 Disparities in Quebec (Canada)"
author: "V. Adjiwanou, G. NGalé, S. Lardoux, N. Zahedinameghi"
date: "`r format(Sys.Date(), '%B %e, %Y')`"
present_to: "Vissého Adjiwanou"
type: "Article"
knit: bookdown::render_book
subtitle: A Comparative Analysis of Exposure Risks, Consequences, and Access to Resources among Immigrants and Natives
groupe_cours: "SOC3380, gr. 40"
site: bookdown::bookdown_site
bibliography: "library.biblatex"
csl: apalike-uqam.csl
editor_options:
  markdown:
    wrap: 80
    fontsize: 12pt
documentclass: article
link-citations: yes
institute: Université du Québec À Montréal and Université De Montréal
abstract: "The immigrant population is one of the most vulnerable in Canada, with a relatively high risk of unemployment, a low social support system, and linguistic barriers. However, their experiences with public health interventions during the COVID-19 pandemic have not been adequately explored. The purpose of this article is to analyze Covid-19 exposure risk and its family implications for both natives and immigrants in the province of Quebec. Additionally, we analyze immigrants' access to government aid and compare it to their native-born counterparts. The study consists of survey data collected from 1,000 immigrants and 500 non-immigrants Quebecers between August and November 2020. Our descriptive and exploratory analysis reveals significant differences between the two groups regarding exposure risk, family consequences, and access to resources. We conclude that the current pandemic exacerbates health disparities in the province, with particularly dire repercussions for immigrants' families. We also discuss several socio-cultural strategies to mitigate inequality and recommend policies to address the needs of vulnerable populations."
---

```{r setup, include=FALSE, warning=FALSE}
# Préférence knitr

knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo=FALSE)
```

```{r}
# Importer package
library(dplyr)
library(tidyverse)
library(summarytools)
library(readr)
library(questionr)
library(rcompanion)
library(pander)
```

```{r}
# Set numero graphique
no_graphic <- -1
```

### 2023-05-02 Expérimentations (Théa)/2. Expérimentations (Bookdown)/Bookdown/src/clean-up-database.R

@@ -1,0 +1,226 @@
+#!/usr/bin/R
+
+# Recode columns
+# Assumes that questionnaire is imported when exported as CSV in English
+# using read_csv and named immigr_init
+immigr_init <-
+  immigr_init %>%
+  dplyr::rename(
+    "age" = "1 - Quel âge avez vous ?",
+    "year_qc" = "2 - Depuis combien d'années habitez vous au Québec ?",
+    "born_country" = "3a - Dans quel pays êtes vous né ?",
+    "country_15" = "3b - Dans quel pays avez vous vécu jusqu'à vos 15 ans ?",
+    "genre" = "4 - A quel genre vous identifiez vous ?",
+    "statut" = "5 - Quel est votre statut actuel au Canada ?",
+    "etat_civil" = "6 - Quel est votre état civil ?",
+    "diplome" = "7 - Quel est votre niveau de diplôme le plus élevé ?",
+    "revenu" = "8 - Dans quelle tranche se situe votre revenu actuel ?",
+    "lang_enf" = "9 - Quelle est la première langue que vous avez apprise à la maison pendant votre enfance et que vous comprenez encore aujourd'hui ?",
+    "lang_home" = "10 - Parmi ces langues, laquelle parlez vous le plus souvent à la maison ? Indiquez plus d'une langue seulement si elles sont parlées aussi souvent l'une que l'autre à la maison.",
+    "lang_home_autre" = "11 - Parlez vous d'autres langues à la maison ?",
+    "lang_trav" = "12 -  Parmi ces langues, laquelle parlez vous le plus souvent au travail ? Indiquez plus d'une langue seulement si elles sont parlées aussi souvent l'une que l'autre au travail.",
+    "lang_trav_autre" = "13 - Parlez vous d'autres langues au travail ?",
+    "rs_util" = "14 - Quel(s) réseaux sociaux utilisez vous régulièrement ?",
+    "rs_time" = "15 - Combien de temps passez vous par jour sur les réseaux sociaux ?",
+    "rs_disc_same" = "16 - Pour quelles raisons allez-vous sur les réseaux sociaux ? [Pour discuter avec des personnes qui proviennent de mon pays d'origine]",
+    "rs_disc_diff" = "16 - Pour quelles raisons allez-vous sur les réseaux sociaux ? [Pour discuter avec des personnes qui ne proviennent pas de mon pays d'origine]",
+    "rs_hab" = "16 - Pour quelles raisons allez-vous sur les réseaux sociaux ? [Par habitude]",
[…]

### 2023-05-02 Expérimentations (Théa)/2. Expérimentations (Bookdown)/Bookdown/src/load-packages.R

@@ -1,0 +1,13 @@
+#!/usr/bin/R
+
+library(dplyr)
+library(car)
+library(plyr)
+library(tidyverse)
+library(summarytools)
+library(readr)
+library(gplots)
+library(graphics)
+library(readxl)
+library(pander)
+library(extrafont)

### 2023-05-02 Expérimentations (Théa)/2. Expérimentations (Bookdown)/Bookdown/src/print-index.R

@@ -1,0 +1,100 @@
+#!/usr/bin/R
+
+index_names <- names(immigr_init)
+index_desc <- c(
+ "1 - Quel âge avez vous ?",
+ "2 - Depuis combien d'années habitez vous au Québec ?",
+ "3a - Dans quel pays êtes vous né ?",
+ "3b - Dans quel pays avez vous vécu jusqu'à vos 15 ans ?",
+ "4 - A quel genre vous identifiez vous ?",
+ "5 - Quel est votre statut actuel au Canada ?",
+ "6 - Quel est votre état civil ?",
+ "7 - Quel est votre niveau de diplôme le plus élevé ?",
+ "8 - Dans quelle tranche se situe votre revenu actuel ?",
+ "9 - Quelle est la première langue que vous avez apprise à la maison pendant votre enfance et que vous comprenez encore aujourd'hui ?",
+ "10 - Parmi ces langues, laquelle parlez vous le plus souvent à la maison ?\n Indiquez plus d'une langue seulement si elles sont parlées aussi souvent l'une que l'autre à la maison.",
+ "11 - Parlez vous d'autres langues à la maison ?",
+ "12 - Parmi ces langues, laquelle parlez vous le plus souvent au travail ? Indiquez plus d'une langue seulement si elles sont parlées aussi souvent l'une que l'autre au travail.",
+ "13 - Parlez vous d'autres langues au travail ?",
+ "14 - Quel(s) réseaux sociaux utilisez vous régulièrement ?",
+ "15 - Combien de temps passez vous par jour sur les réseaux sociaux ?",
+ "16 - Pour quelles raisons allez-vous sur les réseaux sociaux ? [Pour discuter avec des personnes qui proviennent de mon pays d'origine]",
+ "16 - Pour quelles raisons allez-vous sur les réseaux sociaux ? [Pour discuter avec des personnes qui ne proviennent pas de mon pays d'origine]",
+ "16 - Pour quelles raisons allez-vous sur les réseaux sociaux ? [Par habitude]",
+ "16 - Pour quelles raisons allez-vous sur les réseaux sociaux ? [Pour me divertir]",
+ "16 - Pour quelles raisons allez-vous sur les réseaux sociaux ? [Pour passer le temps]",
+ "16 - Pour quelles raisons allez-vous sur les réseaux sociaux ? [Pour m'informer]",
[…]

### 2023-05-02 Expérimentations (Théa)/2. Expérimentations (Bookdown)/Bookdown/Bookdown.Rmd
Multiple changes in Bookdown

@@
 ---
-title: "Réseaux sociaux et sentiment d’appartenance chez les immigrants établis au Québec"
-author: "Vincent Bertrand-Huneaut, Jeanne Craig-Larouche, Théa Carlier, Noura Kone,  Antoine Martin"
+title: "COVID-19 Disparities in Quebec (Canada)"
+author: "V. Adjiwanou, G. NGalé, S. Lardoux, N. Zahedinameghi"
 date: "`r format(Sys.Date(), '%B %e, %Y')`"
 present_to: "Vissého Adjiwanou"
-type: "Rapport de recherche"
-knit: bookdown::render_book
-subtitle: Rapport de recherche
-groupe_cours: "SOC3380, gr. 40"
+type: "Article"
+knit: bookdown::render_book
+subtitle: A Comparative Analysis of Exposure Risks, Consequences, and Access to Resources among Immigrants and Natives
 site: bookdown::bookdown_site
@@
 link-citations: yes
-institute: Université de Québec à Montréal
-abstract: "Ce rapport de recherche présente les résultats d'une étude mixte visant à comprendre la relation entre l’usage des réseaux sociaux en ligne et l’intégration culturelle des personnes immigrantes de première génération établies au Québec depuis plus d'un an.
-Son objectif est de saisir comment les discours présents sur les réseaux sociaux en ligne, ainsi que l’entretien de réseaux diversifiés sur des sites comme Facebook, encouragent ou découragent les personnes immigrantes à s’intégrer dans la culture québécoise, et ainsi, à développer un sentiment d'appartenance pour le Québec.
-Pour répondre à ces questions, nous nous reposons sur la théorie de la fusion culturelle proposée par Stephen M.
-Croucher et Éric Kramer ainsi que sur les travaux de Friedrich Heckmann et de Wolfgang Bosswick sur l'intégration culturelle.
-Partant, nous avons conduit et analysé une série de 10 entretiens semi-directifs en parallèle d'une enquête par questionnaire (N = 65) auprès de personnes immigrantes de première génération établies à Montréal depuis plus d'un an.
-Mots clés : #Immigration #Intégration culturelle #Réseaux sociaux en ligne #Sentiment d'appartenance #Québec"
----
-
-```{r setup, include=FALSE, warning=FALSE}
-# Préférence knitr
-
-knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo=FALSE)
-```
+institute: Université du Québec À Montréal and Université De Montréal
+abstract: "The immigrant population is one of the most vulnerable in Canada, with a relatively high risk of unemployment, a low social support system, and linguistic barriers.
+However, their experiences with public health interventions during the COVID-19 pandemic have not been adequately explored.
+The purpose of this article is to analyze Covid-19 exposure risk and its family implications for both natives and immigrants in the province of Quebec.
+Additionally, we analyze immigrants' access to government aid and compare it to their native-born counterparts.
+The study consists of survey data collected from 1,000 immigrants and 500 non-immigrants Quebecers between August and November 2020.
+Our descriptive and exploratory analysis reveals significant differences between the two groups regarding exposure risk, family consequences, and access to resources.
+We conclude that the current pandemic exacerbates health disparities in the province, with particularly dire repercussions for immigrants' families.
+We also discuss several socio-cultural strategies to mitigate inequality and recommend policies to address the needs of vulnerable populations."
+---
 
@@
 
-# Introduction
-
-Au Québec, la pérennité de l'économie, de la démographie et de la langue
-officielle repose en partie sur le renouvellement du flux migratoire
-international. Dans ce contexte, l'intégration culturelle des personnes
-immigrantes constitue un enjeu historique pour la province, dont le gouvernement
-défend depuis les années 1960 la spécificité culturelle et politique au sein de
-la fédération du Canada. Si l'intégration des personnes immigrantes est étudiée
-depuis les années 1910 aux États-Unis, le récent développement des réseaux
-sociaux en ligne offre depuis vingt ans une nouvelle manière d'appréhender ce
-phénomène.
-
-Dans l'état actuel de la littérature scientifique, nous savons que les réseaux
-sociaux en ligne permettent aux personnes immigrantes de conserver un lien
-affectif et culturel avec leur pays d'origine afin de réduire le stress
-acculturatif généré par la précarisation liée au processus de migration. Dans le
-même temps, les réseaux sociaux en ligne permettent aux personnes immigrantes
-d'entretenir et d'accumuler du capital culturel et social mobilisable dans le
-pays d'accueil.
-
-Les études précédemment réalisées dans ce domaine reposent sur des séries
-d'entretiens semi-directifs, ou sur une enquête par questionnaire. Dans le cas
-des enquêtes qualitatives, la nature des échantillons tirés et le type
-d'analyses réalisées ne permettent pas de quantifier la relation à l'étude ni de
-généraliser les résultats à l'échelle de la population étudiée. Concernant les
-enquêtes quantitatives, si l'enquête par questionnaire permet de mettre à
-l'épreuve statistique la relation entre l'usage des réseaux sociaux en ligne et
-l'intégration culturelle des personnes immigrantes, celle-ci peine à fournir des
-explications précises quant aux modalités du phénomène étudié.
-
-Dans ce contexte, nous avons réalisé une étude mixte regroupant 10 entretiens
-semi-directifs ainsi qu'une enquête par questionnaire (N = 65) visant à
-reproduire en contexte québécois les études proposées par la littérature
-scientifique. Notre objectif est de comprendre la relation entre l'usage des
-réseaux sociaux en ligne et l'intégration culturelle des personnes immigrantes
-établies au Québec. Plus précisément, nous nous intéressons au sentiment
-d'appartenance que peuvent développer les personnes immigrantes à travers leur
-utilisation des réseaux sociaux en ligne.
+# Introduction / background
+
+The COVID-19 pandemic has highlighted the vast ethnic, social, economic, and
+health disparities in high-income countries, including Canada. One of the most
+vulnerable segments of these affluent societies is the immigrant population,
+that has become widely diversified with more and more people coming from
+developing countries, and for whom integration into their new society may be
+more difficult. The coronavirus pandemic brings a new challenge for these
+immigrant populations, directly or indirectly through the tensions between the
+ethnic groups that it helps to generate [@coates2020Covid19]. These immigrants
+face language barriers, low social capital, housing insecurity, and severe
+poverty, rendering them particularly susceptible to the effects of the current
+pandemic. Immigrants are also more likely to be employed as essential workers
+where social distancing, sanitary precautions, and self-isolation, measures put
+in place to mitigate the pandemy, are challenging [@reid2021Migrant;
+@yi2021Health; @zhang2021COVID19]. More importantly, migrants may encounter
+various forms of discrimination, leading to social exclusion and restricted
+access to healthcare and social services [@spiritus-beerden2021Mental]. It is
+therefore necessary to understand how this pandemic affects these immigrant
+populations differently and the native populations in Quebec. In doing so, this
+article will reveal a better understanding of how the current pandemic
+aggravates existing inequalities and challenges in the province for immigrants
+and their descendants.
+
+<!-- The province of Quebec has one of the highest concentrations of immigrants in Canada, including permanent residents, temporary residents, and refugees [@governmentofcanada2017Daily; @immigration2005Recent].
+In 2016, 1.2 million immigrants were living in Quebec, an increase from 8% in 1990 to 13.7% of the population in 2016 (Institut de la statistique du Québec, 2020).
+<!--As of 2019, the increase in net new residents (60,700 versus 42,200 in 2018) and lower interprovincial migration losses (-1,500 versus -6,100 in 2018) contributed to Quebec's population growth.
+-->
+
+Since the start of the pandemic, most of the confirmed cases were identified in
+Montreal and areas with high immigrant populations [@miconi2021Ethnocultural;
+@shields2020COVID19]. The coronavirus pandemic has infected nearly 450,000
+people in Quebec and has resulted in more than 11,000 deaths
+(<https://www.quebec.ca/sante/problemes-de-sante/a-z/coronavirus-2019/situation-coronavirus-quebec>,
+consulted on octobre 29, 2021). This makes it the worst affected province in
+Canada and one of the worst affected regions in the world during the first wave
+of the pandemic. Although the majority of these deaths took place in elderly's
+homes, and therefore affected more this segment of the population, its overall
+influence on individuals is not to be overlooked. It has been shown that, on the
+one hand, immigrants and their families have a higher risk of being exposed to
+the pathogen, and, at the same time, they are more vulnerable to racism and
+discrimination and face barriers to accessing social support [@2021CRARR] and
+information, especially for those who didn't speak or understand the two
+official langages.
+<!--In addition to the direct consequences of the pandemic, the measures taken by various governments around the world are also not without consequences on families. -->Several
+studies have thus shown excess mortality in ethnocultural minority populations
+[@bertocchi2020Covid19], higher unemployment, greater loss of income and
+negative consequences for children and families in several countries. This
+differentiated effect can be explained in different ways. On the one hand, in
+terms of the direct consequences of the pandemic, the situation of
+precariousness and greater poverty among immigrants exposes them more to the
+more harmful consequences of the disease on their state of health. Likewise,
+immigrants often work in jobs where social distancing is difficult to enforce.
+On the other hand, indirectly, the solutions to the crisis affect immigrants and
+natives differently. The complete shutdown of tye economy with its set of
+adjustments has greatly impeded the employability of immigrants, especially
+newcomers who have not yet grasped the workings of society and who cannot
+benefit from state subsidies. But, it is also shown that these effects are not
+uniform on the whole of the immigrant population, itself, very diverse.
+
+Despite the growing focus on the differential impacts of the COVID-19 on
+immigrants and non-immigrants segment of the population, there is still very
+little knowledge about how the pandemic and its related public policy impact the
+lives of immigrants and their families. The existing health datasets (largely
+used to assess this effect) are not necessarily representative of the immigrant
+population and do not offer an accurate picture of the complex social status and
+health outcomes among this group [@choi2021Studying; @gagnon2021Immigration;
+@khan2015Visible]. In most cases, these datasets lack data on the residence
+length of immigrants and combine newcomers with long-term settlers
+[@edge2013Discrimination; @hayward2021Clinical]. As a result, they may not allow
+a detailed assessment of the nature and extent of risk exposure in sub-segments
+of populations, as well as whether interventions are reaching those most at
+risk. In addition, the family-level impacts of the pandemic are still poorly
+understood. While a body of research has explored the association between the
+pandemic and its potential effects on intimate partner violence and marital
+dissolution [@bhatnagar2021Navigating; @koshan2020COVID19; @morrison2021Chapter;
+@prime2020Risk], few studies have examined negative familial repercussions in
+the context of immigration status. There is ample evidence that sustained family
+relationships greatly enhance immigrants' lives on several fronts, including
+fostering economic prosperity [@marger2001Sociala], psychological well-being
+[@beiser2002Povertya], social network development [@evansluong2017Familya], and
+educational success for children [@munroe-blum1989Immigranta]. The family is,
+therefore, a key component of long-term social mobility and economic success of
+immigrants.
+
+<!--Faced with the pandemic, the governments of Canada and Quebec have implemented a number of measures to reduce its impact.
+These policies can be broken down at two levels, direct support to families, including the Canadian Emergency Benefit, which can provide employees and self-employed workers affected by Covid-19 with a financial benefit of up to $ 2,000.
+On the other hand, the government supports many companies to keep their employees or supports them directly when the revenues from their activities have fallen considerably.
+While these measures allow families to cope better with the crisis, the fact remains that the changes brought about by this crisis can affect other dimensions of life.
+How these measures make a difference in the face of the pandemic remains to be seen.
+-->
+
+Based on an original biographic survey collected around the first wave of the
+pandemic in august-november 2020 on immigrants and the native population of
+Québec, this study mainly aims to understand the family exposure, consequences,
+and responses of the pandemic on immigrant and native Quebec populations. We
+assess a whole range of different measures of each outcome. We divided the
+consequences in three dimensions, 1) access to ressources, 2) ability to perform
+duties and 3) employment and income. We seek also to analyse in what ways these
+outcomes differ according to the immigrants length of residence. By doing so,
+this study provide a clear and global picture of the situation of different
+population groups in Quebec in the face of the COVID-19.
+<!-- To respond to these questions, we will use data collected around the first wave of the pandemic in august-november 2020.
+Our study aims to fill the gap by analyzing group-specific differences in exposure risk, family consequences, and access to resources among immigrants and Quebecers.
+We also seek to shed light on several socio-cultural strategies to mitigate inequality and recommend relevant policy solutions.
+The detailed findings will help policymakers address health inequalities and improve equity in Quebec while developing adequate measures and programs.
+-->
+
+<!-- Introduction  -->
+
+<!-- Like the rest of Canada, Quebec has experienced a substantial increase in its immigration in recent years.
+The immigrant population in the province thus increased from 8% in the 1990s to 13.7% of the population, or 1.2 million individuals in 2016 (Institut de la statistique du Québec, 2020).
+This immigration is also unique in that it has become widely diversified with more and more people coming from developing countries, and for whom integration into their new society may be more difficult.
+The coronavirus pandemic brings a new challenge for these immigrant populations, directly or indirectly through the tensions between the ethnic groups that it helps to generate [@coates2020Covid19].
+It is therefore necessary to understand how this pandemic affects these immigrant populations differently and the native populations in Quebec.
+-->
+
+<!-- The coronavirus pandemic has affected nearly 100,000 people in Quebec and has resulted in more than 6,000 deaths.
+This makes it the worst affected province in Canada and one of the worst affected regions in the world.
+Although the majority of these deaths took place in homes for the elderly, and therefore affected this segment of the population more, its overall influence on individuals is not to be overlooked.
+In fact, in addition to the direct consequences of the pandemic, the measures taken by various governments around the world are also not without consequences on families.
+Several studies have thus shown excess mortality in ethnocultural minority populations [@bertocchi2020Covid19], higher unemployment, greater loss of income and negative consequences for children and families in several countries.
+-->
+
+<!-- This differentiated effect can be explained in different ways.
+On the one hand, in terms of the direct consequences of the pandemic, the situation of precariousness and greater poverty among immigrants exposes them more to the more harmful consequences of the disease on their state of health.
+Likewise, immigrants often work in jobs where social distancing is difficult to enforce.
+On the other hand, indirectly, the solutions to the crisis affect immigrants and natives differently.
+The complete shutdown of the company with its set of adjustments has greatly impeded the employability of immigrants, especially newcomers who have not yet grasped the workings of society and who cannot benefit from state subsidies.
+But, it is also shown that these effects are not uniform on the whole of the immigrant population, itself, very diverse.
+-->
+
+<!-- Faced with the pandemic, the governments of Canada and Quebec have implemented a number of measures to reduce the impact of the pandemic.
+These policies can be broken down at two levels, direct support to families, including the Canadian Emergency Benefit, which can provide employees and self-employed workers affected by Covid-19 with a financial benefit of up to $ 2,000.
+On the other hand, the government supports many companies to keep their employees or supports them directly when the revenues from their activities have fallen considerably.
+While these measures allow families to cope better with the crisis, the fact remains that the changes brought about by this crisis can affect other dimensions of life.
+How these measures make a difference in the face of the pandemic remains to be seen.
+-->
+
+<!-- This study mainly aims to understand the family consequences of the pandemic on immigrant and native Quebec populations.
+Therefore, we will seek to analyse in what ways its consequences differ according to the region of origin of immigrants, length of residence, age and sex.
+-->
 
@@
 
-<!--chapter:end:01-intro.Rmd-->
-
----
-knit: bookdown::render_book
-editor_options: 
-  markdown: 
-    wrap: 80
----
-
-# Revue de littérature
-
-## Intégration culturelle
-
-Pour comprendre ce que signifie le sentiment d'appartenance des personnes
-immigrées établies au Québec, il est nécessaire de se pencher sur le lien qui
-unit ces dernières à leur pays d'accueil.
-
-Cette relation intéresse les sociologues depuis le début du XIX^e^ siècle. On en
-trouve d'abord la trace aux États-Unis, au sein de la première école de Chicago.
-Les travaux de William I. Thomas [-@thomasPolishPeasantEurope1918] et Robert E.
-Park [-@parkHumanMigrationMarginal1928] contribuent à définir un premier concept
-servant à appréhender ce rapport : l'assimilation. Celle-ci se définie comme : «
-un processus individuel de convergence des caractéristiques des [personnes
-immigrées] vers les caractéristiques moyennes de la société d'accueil »
-[@safiPenserIntegrationImmigres2011, p. 150].
-
-Après avoir dominé durant le XXe siècle, le concept d'assimilation est remis en
-question. Friedrich Heckmann remarque que celui-ci est normatif, en ce qu'il
-prend pour acquis que la culture du pays d'accueil forme un ensemble homogène
-[-@heckmannIntegrationIntegrationPolicies2005, p. 12].
-
-S'inspirant de la théorie de l'intégration des systèmes publiée par David
-Lockwood en 1964, Hartmut Esser propose d'appréhender ce lien comme une seconde
-socialisation possédant quatre dimensions : 1) la capacité des personnes
-immigrantes à mobiliser les institutions de leur pays d'accueil ; 2) leur
-placement dans le marché du travail, de la vie conjugale, etc. ; 3) leur
-capacité à interagir avec d'autres personnes dans la société d'accueil ; et 4)
-le sentiment d'appartenance qu'elles développent envers leur pays d'accueil ou
-d'autres sous-ensembles lui appartenant [-@esserDoesNewImmigration2004, p.
-1131].
-
-Or, cette nouvelle conceptualisation du lien entre les personnes immigrées et
-leur pays d'accueil se heurte à la critique de Young Kim, qui avance que
-celui-ci ne constitue pas simplement une deuxième socialisation, mais implique
-également une désocialisation plus ou moins prononcée vis-à-vis de la diversité
-culturelle présente dans le pays d'origine
-[-@kimBecomingInterculturalIntegrative2001, p. 53]. Quelques années plus tard,
-Stephen M. Croucher et Éric Kramer développent la théorie de la fusion
-culturelle. A l'instar des travaux de Kim, la fusion culturelle se comprend
-comme « [le] mélange entre l'apprentissage de nouveaux comportements et le
-maintien [relatif] d'anciens comportements »
-[-@croucherCulturalFusionTheory2017, p. 97]. Néanmoins, Croucher et Kramer
-introduisent une nouvelle dimension dans leur concept, en avançant que la fusion
-culturelle est simultanée. En d'autres termes, cela signifie que la culture des
-personnes immigrantes fusionne avec la diversité culturelle présente dans le
-pays d'accueil pour produire une forme altérée de la diversité culturelle
-initiale [-@croucherCulturalFusionTheory2017, p. 103].
-
-## Réseaux sociaux en ligne et intégration culturelle
-
-La migration constitue une rupture dans le processus de socialisation car elle
-déracine les personnes migrantes de leur environnement. Dans ce contexte, elles
-expérimentent ce que des chercheurs en psychologie comme John W. Berry et al.
-[-@berryImmigrationAcculturationAdaptation1997] appellent du « stress
-acculturatif ». Celui-ci est caractérisé par de possibles dégradations de la
-santé comme l'anxiété, la dépression, la confusion identitaire ou encore un
-sentiment de marginalisation de la part des personnes établies dans le pays
-d'accueil. Selon Berry, ce stress et cette marginalisation (effectifs et/ou
-perçus) sont causés par les différences culturelles qui oppose le pays d'origine
-au pays d'accueil.
-
-L'étude d'Eric Forbush et de Brooke Foucault-Wells [-@forbushSocialMediaUse2016]
-sur l'usage des médias sociaux par des étudiants chinois aux États-Unis, ainsi
-que l'étude d'Amanda Alencar [-@alencarRefugeeIntegrationSocial2017] parmi les
-réfugiés syriens établis aux Pays-Bas nous apprennent que les personnes
-immigrantes utilisent régulièrement des réseaux sociaux en ligne comme Facebook,
-mais aussi des applications de messagerie instantanée comme WhatsApp pour
-communiquer avec leurs proches restés au pays. Dans une étude sur l'intégration
-culturelle des étudiants chinois établis aux États-Unis, Shaoke Zhang
-[-@zhangBabelGreatWall2012] constate que les réseaux sociaux en ligne permettent
-de réduire le stress acculturatif généré par la migration, de diminuer
-l'incertitude quant à la situation des personnes immigrantes et de réduire le
-fossé les séparant de la culture du pays d'accueil. Dans une étude menée en
-Nouvelle-Zélande par Kingsley T. Ihejirika et Maja Krtalic
-[-@ihejirikaMovingMediaExploration2021], on peut lire que les personnes
-immigrantes utilisent les réseaux sociaux en ligne pour s'intégrer en
-s'imprégnant des normes, de la langue et du mode de vie néo-zélandais afin de
-pouvoir interagir avec la population et les institutions locales. Leurs travaux
-révèlent que les personnes immigrantes utilisent les réseaux sociaux en ligne
-pour collecter des informations, communiquer avec leur famille restée au pays,
-construire de nouvelles relations, réseauter et apprendre la langue du pays
-d'accueil. La proportion de ces usages varie en fonction des caractéristiques
-sociodémographiques des migrants, mais aussi du contexte socioéconomique. En
-reprenant des matériaux issus de l'enquête ALTERECOLE, Évelyne Barthou
-[-@barthouMigrationSansRupture2020] montre que les réseaux sociaux en ligne
-facilitent également l'émergence d'identités plurielles et d'une culture
-composite chez les immigrants marocains établis en France.
-
-Or, ces résultats sont pondérés par le travail de Lee Komito
-[-@komitoSocialMediaMigration2011] dans son étude sur les polonais.es et les
-philippin.es établies en Irlande. Le chercheur remarque que l'usage des réseaux
-sociaux en ligne peut contribuer à isoler les personnes immigrantes en leur
-permettant de rester connecté à la culture de leur pays d'origine. En outre,
-dans leur travaux sur les jeunes immigrant.es établis au Québec, Nicole Gallant
-et Céline Friche relèvent que si les médias sociaux en ligne facilitent
-l'entretient de réseaux sociaux et d'appartenances préexistantes, ceux-ci : «
-semblent avoir peu d'incidence sur le développement d'un réseau social dans le
-milieu d'accueil, si ce n'est pour colorer les rapports au quotidien »
-[-@gallantEtreIciLabas2010, pp.64].
-
-## Le place du discours sur les réseaux sociaux en ligne
-
-La production et la diffusion d'informations en ligne diffèrent de celle des
-médias traditionnels, car internet permet une interaction instantanée entre le
-producteur et le consommateur de contenu. Comme le démontrent André Mondoux et
-Marc Ménard dans leur livre *Big data & Société* [-@mondouxBigDataSociete2018],
-la logique algorithmique place les utilisatrices et les utilisateurs des réseaux
-sociaux en ligne dans une relation circulaire entre leur consommation et des
-propositions ciblées.
-
-Dans un chapitre d'ouvrage collectif, Ehsan Dehghan et Axel Bruns révèlent que
-la racialisation discursive observée sur les réseaux sociaux en ligne à
-l'endroit des personnes immigrantes n'est pas tant dû à la structure des
-plateformes qu'au comportement stratégique d'utilisateurs souhaitant promouvoir
-leurs idées [-@dehghanDynamicsPolarisationAustralian2022, p.64‑65]. Dans le même
-temps, des études comme celles de Jeroen Hopster ou encore Manoel Horta Ribeiro
-et al. démontrent comment la logique algorithmique des réseaux sociaux en ligne
-facilite le développement de communautés véhiculant des discours populistes -
-souvent xénophobes et racistes - par l'entremise de la suggestion automatique de
-contenu et le système de commentaires sur YouTube
-[-@hopsterMutualAffordancesDynamics2021, p. 559;
--@ribeiroAuditingRadicalizationPathways2020, p. 140]. James P. Walsh parle même
-de « nativisme digital » pour désigner l'usage stratégique de Twitter dont font
-preuve certains groupes d'extrême droite [-@walshDigitalNativismTwitter2021, p.
-5]. Selon lui, durant la campagne électorale canadienne de 2019, les messages
-racistes à l'attention des personnes immigrantes sur Twitter ont généré un
-engouement inédit, exacerbant les discours anti-immigration. Son travail
-soutient que le nativisme - une idéologie xénophobe s'opposant à la migration et
-reposant sur le principe de racialisation - est en train d'être restructuré par
-les réseaux sociaux en ligne. Selon Walsh, les groupes nativistes ont
-instrumentalisé les possibilités offertes par Twitter afin de véhiculer un
-discours raciste et anti-immigration parmi la population et les institutions
-canadiennes [@walshDigitalNativismTwitter2021, p. 19].
-
 \newpage
 
-<!--chapter:end:02-revue-litterature.Rmd-->
-
----
-knit: bookdown::render_book
-editor_options: 
-  markdown: 
-    wrap: 80
----
-
-# Cadre théorique
-
-## Intégration culturelle
-
-Dans le cadre de notre étude, nous avons produit une définition de l'intégration
-culturelle à partir des travaux de Heckman, Bosswick
-[-@heckmannIntegrationMigrantsContribution2006, p.10], Kim
-[-@kimBecomingInterculturalIntegrative2001, p.53], Cramer et Kroucher
-[-@croucherCulturalFusionTheory2017, p. 97]. Selon ces auteurs, le processus
-d'intégration culturelle implique, à l'échelle microsociologique (l'individu) et
-mésosociologique (la population des personnes immigrantes de première
-génération), une transformation de la diversité culturelle véhiculée par les
-personnes immigrantes lorsqu'elles s'établissent dans un nouveau pays. Dans le
-même temps, à un niveau macrosociologique (les populations immigrantes établies
-depuis plusieurs générations), la diversité culturelle du pays d'accueil est
-transformée au contact des populations immigrantes nouvellement arrivées.
-
-Dans ce contexte, l'identité nationale se fonde sur une construction sociale
-historiquement située, dont les modalités se transforment continuellement au fil
-du flux migratoire (y compris lorsque celui-ci prend la forme d'une invasion ou
-d'une colonisation). Partant, à l'échelle micro et méso, le développement d'un
-sentiment d'appartenance national repose sur la proximité entre la diversité
-culturelle véhiculée par les individus et celle du pays dans lequel ils se
-trouvent.
-
-Pour nous aider dans l'élaboration d'un modèle d'analyse, les travaux de Marcin
-Gońda et al. [@gondaMeasuringCulturalDimension2021, p.245-246] proposent de
-diviser l'intégration culturelle en trois dimensions : 1) la langue ; 2) les
-valeurs, normes et pratiques ; et 3) l'identification culturelle. Partant Gońda
-et al. proposent de mesurer ses trois dimensions à l'aide d'indicateurs comme :
-1) la connaissance et la compréhension des expressions et/ou de l'accent du pays
-d'accueil ; 2) le type et la fréquence de consommation de contenu culturel du
-pays d'accueil ; et 3) le sentiment d'appartenance au pays d'accueil.
-
-Enfin, les travaux de Loïc Wacquant sur le processus de racialisation
-[-@wacquantResolvingTroubleRace2022] révèlent que certaines populations comme
-les Afro-Américains peuvent être graduellement catégorisées et stigmatisées sur
-la base de rapports de pouvoir dérivés d'objectifs économiques, comme
-l'esclavage aux Etats-Unis.
-
-## Réseaux sociaux en ligne
-
-Pour mieux cerner la notion de réseaux sociaux en ligne, nous avons repris les
-travaux de Andreas M. Kaplan et Michael Haenlein [-@kaplanUsersWorldUnite2010,
-p60-62]. Selon ces auteurs, les réseaux sociaux en ligne peuvent être divisés en
-quatre types :
-
-```{r}
-local({
-  .Table <- data.frame (
-    reseaux  = c(
-      "**Partage de médias**",
-      "**Blogs**",
-      "**Communautés en ligne**",
-      "**Réseaux sociaux**"
-    ),
-    description = c(
-      "Permet à des utilisateurs de produire, trier et partager des médias avec d'autres personnes ou groupes",
-      "Permet aux utilisateurs de publier en ligne", 
-      "Permet aux utilisateurs de partager des informations spécifiques à travers des forums",
-      "Permet aux utilisateurs de construire des réseaux sociaux amicaux et professionnels"
-    ),
-    exemple = c("YouTube", "Twitter", "Reddit", "Facebook")
-  )
-  pander(.Table, split.cells = c(20, 40, 10), split.table = Inf,
-         caption = "Typologie des réseaux sociaux",
-         col.names = c("", "Descriptions", "Exemple")
-  )
-})
-```
-
-## Cadre conceptuel
-
-Le schéma suivant cette page illustre, au regard de nos choix théoriques, notre
-approche concernant l'étude de la relation entre l'utilisation des réseaux
-sociaux en ligne et l'intégration culturelle des personnes immigrantes établies
-au Québec.
-
-\setcounter{figure}{`r paste(no_graphic <- no_graphic+1)`}
-![Cadre conceptuel schématisé](images/figure_6.png){width=80%}
-
-## Hypothèses de recherche
-
-### H1 - Contact interculturel et sentiment d'appartenance
-
-Les travaux de Ihejirika et Krtalic [-@ihejirikaMovingMediaExploration2021]
-suggèrent que l'usage des réseaux sociaux en ligne permet aux personnes
-immigrantes de s'intégrer culturellement dans leur pays d'accueil à travers la
-pratique de normes, l'intériorisation de valeurs et l'apprentissage de la langue
-locale. Dans le même temps, les travaux de Gallant et Friche
-[-@gallantEtreIciLabas2010] indiquent que l'usage des réseaux sociaux en ligne
-permet aux personnes immigrantes d'entretenir des réseaux d'appartenance
-préexistants (notamment dans le pays d'origine). À partir de ces résultats, nous
-avons émit l'hypothèse que le contact interculturel sur les réseaux sociaux en
-ligne favorise le développement d'un sentiment d'appartenance au Québec chez nos
-répondants.
-
-### H1.1 - Pays d'origine, discrimination systémique et sentiment d'appartenance
-
-En appliquant le concept de *racialisation diagonale* introduit par Wacquant
-[-@wacquantResolvingTroubleRace2022, p.78] en tant que variable de contrôle dans
-la relation qui unit l'usage des réseaux sociaux en ligne avec le développement
-d'un sentiment d'appartenance au Québec, on peut émettre l'hypothèse que les
-discriminations anti-immigrantes présentes sur les réseaux sociaux en ligne,
-qu'elles soient d'ordre national, ethnique ou racial, font obstacle au
-développement d'un sentiment d'appartenance au Québec proportionnellement au
-degré de racialisation des individus.
-
-### H1.2 - Type de réseaux sociaux en ligne et sentiment d'appartenance
-
-À partir de la typologie proposée par Kaplan et Haenlein
-[-@kaplanUsersWorldUnite2010] et des connaissances fournies par la littérature
-regardant l'usage des réseaux sociaux en ligne par les personnes immigrantes,
-nous émettons également l'hypothèse que l'entretien et l'accumulation de capital
-social (réseaux de connaissances diversifiés) et culturel (apprentissage du
-français et familiarisation avec la diversité culturelle présente dans le pays
-d'accueil) sur les réseaux sociaux en ligne accroissent le sentiment
-d'appartenance au Québec. Plus précisément, nous pensons que l'accumulation de
-capital social se concentre sur les sites de réseaux sociaux comme Facebook,
-tandis que l'accumulation de capital culturel prend place prioritairement sur
-les sites de partage de médias comme YouTube, étant donné l'architecture de ces
-deux types de réseaux sociaux en ligne.
-
-### H1.3 - Type d'utilisation des réseaux et sentiment d'appartenance
-
-La littérature que nous avons eu l'occasion de parcourir illustre comment les
-réseaux sociaux en ligne permettent aux personnes immigrantes de réduire le
-stress acculturatif généré par la migration en maintenant un contact affectif et
-culturel avec le pays d'origine, parfois au détriment du processus de fusion
-culturelle [@forbushSocialMediaUse2016 ; @alencarRefugeeIntegrationSocial2017 ;
-@zhangBabelGreatWall2012 ; @komitoSocialMediaMigration2011]. Partant, nous avons
-émit l'hypothèse que le type d'utilisation des réseaux sociaux en ligne (au
-Québec ou vers le pays d'origine) joue un rôle dans la relation étudiée.
-
-### H2 - Exposition à du contenu lié à l'immigration et sentiment d'appartenance
-
-On l'a vu précédemment, la littérature révèle l'existence d'une association
-entre les discours anti-immigration et l'intégration culturelle des personnes
-immigrantes [@dehghanDynamicsPolarisationAustralian2022 ;
-@walshDigitalNativismTwitter2021 ; @hopsterMutualAffordancesDynamics2021 ;
-@ribeiroAuditingRadicalizationPathways2020]. Partant, nous avons émit
-l'hypothèse que l'exposition à du contenu lié à l'immigration en ligne tend à
-freiner le développement d'un sentiment d'appartenance chez les personnes
-immigrantes.
+<!--chapter:end:01-introduction.Rmd-->
+
+---
+knit: bookdown::render_book
+editor_options: 
+  markdown: 
+    wrap: 80
+---
+
+# Current body of literature and knowledge gaps
+
+The literature on the racial and ethnic health disparities during the COVID-19
+crisis can be divided into two main categories: studies addressing the
+socio-economic determinants of health and those examining the delivery of health
+interventions and policies. The first category is primarily comprised of
+research that focuses on the link between socio-economic status (SES) of
+high-risk minority groups and the differential impact of the pandemic
+[@drefahl2020Populationbased; @greenaway2020COVID19; @lopez2021Racial;
+@pacheco2020Job; @sundaram2020Individual]. For instance, an inter-country study
+has found that Middle Eastern, African, and Nordic immigrants in Sweden have a
+higher risk of being hospitalized and even dying from COVID-19 when compared to
+the Swedish born population. In the same study, the SES, proportion of
+working-age households, and neighborhood density were significantly associated
+with the mortality rate [@rostila2021Disparities]. This trend of research
+demonstrates that racial and ethnic minorities are more likely to experience
+stress factors, not only due to the hazards of infection but also because of
+economic and social marginalization [@alahmad2020COVID19; @choi2020Buffering;
+@endale2020COVID19; @fitzpatrick2020Fear; @serafini2021Psychological]. The
+intersectionality of race, sex, and health conditions has also been explored in
+the Canadian context of the pandemic. @ingen2021Neighbourhoodlevel demonstrated
+that Ontarian neighborhoods with immigrant and racialized populations have a
+disproportionate share of COVID-19 infections. In addition, areas with high SES
+report a lesser number of COVID cases than low SES ones
+[@ingen2021Neighbourhoodlevel]. Another study conducted in Quebec confirms the
+strong association between immigrants' backgrounds, SES, and COVID-19 related
+psychological stress [@miconi2021Ethnocultural]. The most recent reviews on the
+topic suggest that immigrants disproportionately live in crowded and inadequate
+housing and are employed in essential industries, leaving them with severe
+income insecurities and poor physiological/psychological well-being
+[@chew2020Clinical; @davis2020Impact]. Studies on SES and health disparities
+have emphasized that this situation is further complicated for immigrant women,
+constrained by gender and afflicted by a significant power imbalance in the
+household and workplace [@alarcao2021Intersections;
+@esegbona-adeigbe2020COVID19; @jehn2021COVID19; @mo2020Differential;
+@nardon2021Skilled; @sabri2020Effect; @st-denis2020Sociodemographic]. According
+to Lightman, immigrants and women of color are overrepresented in long-term care
+facilities in Canada, where COVID infection rates have been exceptionally high.
+They are also subject to exclusionary social and economic practices
+[@lightman2021Caring].
+
+The second category of the literature is primarily comprised of policy analyses,
+research on public health communications, examining the efficacy of COVID
+initiatives, the optimal allocation of government aids, and ethical issues
+associated with applying new health interventions to vulnerable populations and
+their communities [@aragona2020Negative; @cross2021Coronavirus;
+@desai2020COVID19; @doyle2020Migrant; @gilman2020Modelling;
+@hargreaves2020Targeting; @langellier2020Policy; @nezafatmaldonado2020Engaging;
+@silverman2020Ethics; @wilson2020Ethics]. Some of the questions that this body
+of research addresses include: how best to respond to COVID-19 in immigrant
+communities? How do public health policies impact the socio-economic status and
+psychological well-being of different sub-populations? What is the structure of
+the planning and implementation of government aid programs during the current
+crisis? And, how can we design public health communication strategies in a way
+that is culturally sensitive and inclusive?
+
+There are several knowledge gaps in the existing literature. Most of the cited
+studies do not include household data and family structure of immigrants as
+variables that affect exposure risks and health outcomes of COVID-19.
+Consequently, the relationship between immigrant status and COVID-19 is not
+accounted for by household size, and family characteristics (i.e., income,
+education, religion). A key reason is that most of the health data used in such
+research projects are acquired from national registers or administrative
+databases, which provide little information on immigration status and kinship.
+However, as the study of Thomeer and her colleagues theorizes, the extent of
+adverse social and economic impacts of COVID-19 is significantly conditional on
+the composition of an immigrant's family [@thomeer2020How]. Households that
+include a spouse/partner, children, or multigenerational family members most
+likely experience a different level of social, psychological, and medical
+vulnerability than single or nuclear ones. Therefore, studies of immigrant
+populations should account for the breakdown of family configuration to better
+understand the mechanisms linking social determinants and health outcomes.
+Another knowledge gap concerns the current state of Canadian studies of COVID-19
+and its public health crisis. There is a dearth of literature on the challenges
+faced by immigrants and their families in the country. Although the burden of
+COVID-19 infection is disproportionately higher among these marginalized groups
+(St-Denis, 2020; Sundaram et al., 2020), research on the issues they are
+experiencing is sparse and often limited to one aspect.
+<!--Likewise, there is limited focus on the effectiveness of COVID-19 crisis management for immigrants at the policy level.
+In response to the mounting cases of COVID, the federal government and provinces and territories of Canada have launched multi-pronged public health initiatives to protect society from the epidemic (Migone, 2020).
+However, many open questions remain unanswered, including, do interventions efficiently reach the migrant groups?
+What are the likely effects of each program on different immigrant communities?
+How can an intervention be tailored to the needs of a particular migrant group and their families?
+-->
 
@@
 
-<!--chapter:end:03-cadre-conceptuel.Rmd-->
-
----
-knit: bookdown::render_book
-editor_options: 
-  markdown: 
-    wrap: 80
----
-
-# Méthodologie
-
-## Volet quantitatif
-
-Une enquête par questionnaire a été réalisée auprès de 65 personnes immigrantes
-établies au Québec depuis plus d'un an, sur la base d'un échantillon théorique
-(recrutement réalisé sur les réseaux sociaux en ligne et parmi les étudiant.es
-de l'Université du Québec à Montréal). En accord avec la littérature
-scientifique, nous avons concentré son analyse autour de deux hypothèses : 1)
-l'exposition à du contenu médiatique sur les réseaux sociaux tend à accroître le
-sentiment d'appartenance au Québec des personnes immigrantes ; et 2) la
-fréquence des contacts intergroupes accroît le développement d'un sentiment
-d'appartenance au Québec chez les personnes immigrantes. Notre seconde hypothèse
-se divise en deux sous-hypothèses : a) le type de réseaux social en ligne
-utilisé affecte la relation entre le contact interculturel et le développement
-d'un sentiment d'appartenance au Québec ; b) le type d'usage des réseaux sociaux
-en ligne (inter ou intraculturel) la relation entre le contact interculturel et
-le développement d'un sentiment d'appartenance au Québec.
-
-Pour mettre à l'épreuve ces deux hypothèses à l'aide des 84 variables
-factorielles qui composent notre questionnaire, nous avons choisi de produire
-des indices à partir de la formule proposée par Haimanti Bhattacharya dans son
-étude de la relation entre l'attitude pro-sociale et écologiste des étudiants
-aux Etats-Unis [-@bhattacharyaProsocialStudentsCare2019] :
-
-```{r}
-local({
-  .Table <- data.frame(
-   formule = "$\\displaystyle\\sum\\limits_{i=1}^{n}\\frac{(xi-1)}{4n}$",
-   n = "le nombre d'observations",
-   x =  "le score associé à la réponse dans la variable ()"
-  )
-
-pander(.Table, split.cell=c(5, 40, 40), split.table = Inf,
-         caption = "Formule des indices",
-         col.names = c("Formule", "n", "x")
-  )
-})
-```
-
-Le résultat donne un indice situé entre 0 (absence de la variable) et 1
-(présence totale de la variable). Le tableau suivant présente la composition des
-indices :
-
-```{r}
-local({
-  .Table <- data.frame (
-    indice  = c(
-      "**Contact interculturel en ligne**",
-      "**Exposition au discours sur l'immigration en ligne**",
-      "**Sentiment d'appartenance au Québec**"
-    ),
-    no_var = c("5", "5", "9"),
-    niv_mesure = c("Echelle Likert à 5 degrés")
-  )
-  pander(.Table, split.cell=c(40, 5, 40), split.table = Inf,
-         caption = "Compositions des indices et leur quantité",
-         col.names = c("Indice", "Quantité", "Niveau de mesure")
-  )
-})
-```
+<!--chapter:end:02-body-of-litterature.Rmd-->
+
+---
+knit: bookdown::render_book
+editor_options: 
+  markdown: 
+    wrap: 80
+---
+
+# Theoretical framework
+
+Family as a unit of analysis can be particularly fruitful when studying health
+inequalities in the context of a societal-wide crisis. Through the lens of the
+COVID-19 pandemic, the following points illustrate why exploring family dynamics
+provides additional insight into the underlying mechanisms of health
+disparities. First, the family is the most fundamental unit of social
+organization. It facilitates access to resources and the transmission of social
+and psychological capital from one generation to the next. Therefore, a better
+analysis of how crises and adverse life events affect people requires a better
+understanding of how families relate to each other and interact with external
+forces. Second, during public health emergencies, families may be thrown into
+different structures due to unprecedented financial, mental, and physical
+challenges. The risk of family dissolution, for example, can be heightened when
+members must spend unusually long periods together at home and have limited
+opportunities to engage with friends and the wider community. Conversely,
+families can develop closer bonds through self-sacrifice, shared
+responsibilities, and increased care and support. As the dependency status of
+seniors increases, the elderlies are also more likely to be integrated into
+formerly nucleus families. We posit that the extent to which a family can adapt
+its functions to meet new challenges is a predictor of the well-being of its
+members and the resilience of its broader population. Third, the family is a
+crucial link to social institutions, such as the health care system, schools,
+and government agencies whose decisions have significant downstream impacts.
+Families are viewed by many public-level entities as a reference when allocating
+resources, enforcing new standards, and responding to emerging needs and
+demands. From this perspective, the family operates at the nexus of individual
+and societal structures, embedding the tripple effects of public policy in its
+most consequential ways. Finally, the family has a crucial role to play in
+determining the life course of its members. As COVID-19 elevates the importance
+of familial bonds, the post-pandemic period might be marked by an increased
+influence of families on the lives of their members and the social structures
+they inhabit.
+
+The above conceptualization echoes several other approaches to studying
+long-term health inequalities. For instance, life course theory emphasizes the
+significance of family ties and social environment as they structure prospects
+for individuals across their life span (Elder, 1977; Elder Jr., 1998; Macmillan,
+2005; Settersten et al., 2020). Likewise, social capital theory and the
+stress-buffering hypothesis consider family systems a crucial factor in health
+and well-being (Choi and Jun, 2020; Cohen and McKay, 1984; Cohen and Wills,
+1985; Harpham et al., 2004; Kawachi and Berkman, 2001; Lavee et al., 1987;
+Mandelbaum et al., 2020). They show that the family can offer tangible (e.g.,
+food, shelter, and money) and intangible resources (e.g., support, social
+status, psychological comfort, and institutional advantages), protecting its
+members from adverse life events. Similarly, the social-ecological model and the
+theory of fundamental causes highlight how the SES, familial and extrafamilial
+factors pattern people's life trajectories and health outcomes (Barkan and
+Rocque, 2018; Clouston and Link, 2021; Figge et al., 2018; Kim, 2020; Libman et
+al., 2012; Ungar, 2015). However, all these approaches are ostensibly focused on
+life periods where the general social context is relatively stable. In this
+article, we present our conceptualization of the family in times of
+unprecedented social upheaval. Specifically, our framework seeks to characterize
+family dynamics as a suitable lens to analyze health inequalities in the face of
+a virulent public health crisis.
 
@@
 
-À partir de ces nouvelles variables, nous avons réalisé des modèles de
-régression linéaire multivariés prenant en compte les variables de contrôle
-relevées dans la littérature scientifique.
-
-## Volet qualitatif
-
-Le volet qualitatif de notre enquête à pour objectif de comprendre comment les
-réseaux sociaux en ligne influencent le sentiment d'appartenance au Québec. À
-partir de notre revue de littérature, nous émettons l'hypothèse que les
-discriminations raciales anti-immigrants présentes sur les réseaux sociaux
-empêchent ces derniers de se sentir pleinement québécois. De même, nous émettons
-l'hypothèse que les contacts entretenus sur les réseaux sociaux en ligne avec
-des Québécois favorisent le développement d'un sentiment d'appartenance au
-Québec chez les personnes immigrantes.
-
-Pour mettre à l'épreuve ces deux hypothèses, nous avons mené une série de 10
-entretiens semi-directifs afin de recueillir le témoignage de personnes
-immigrantes concernant leur parcours migratoire et leur utilisation des réseaux
-sociaux en ligne. L'usage d'un dispositif d'enquête qualitatif vise à saisir
-dans le récit des répondants les clefs d'explication qui échappent à l'analyse
-descriptive du volet quantitatif (expériences, discours véhiculés, pratiques
-précises, etc.). Pour ce faire, nous avons réalisé une analyse thématique
-séquencée à l'aide du logiciel Ligre, convenant à la nature semi-structurée de
-nos transcriptions et au travail collectif.
-
-## Un processus d'analyse mixte et itératif
-
-Le schéma suivant présente notre processus d'analyse : 1) La revue de
-littérature nous permet permet de proposer des hypothèses ; 2) les hypothèses
-guident l'analyse descriptive et l'analyse thématique ; 3) les résultats de
-l'analyse thématique soulèvent de nouvelles hypothèses ; 4) les nouvelles
-hypothèses sont mises à l'épreuve de la littérature (et la revue est
-éventuellement enrichie de nouveaux articles) ; 5) les nouvelles hypothèses sont
-mises à l'épreuve de l'analyse descriptive ; 6) les nouveaux résultats de
-l'analyse descriptive sont triangulés par ceux de l'analyse thématique... *and
-so on, and so forth*.
-
-Si nous avions pu bénéficié de plus de temps, nous aurions pu réviser notre
-d'alimenter le processus d'analyse mixte.
-
-```{=tex}
+<!--chapter:end:03-theoritical-frame.Rmd-->
+
+---
+knit: bookdown::render_book
+editor_options: 
+  markdown: 
+    wrap: 80
+---
+
+# Methodology
+
+## Data
+
+The data for this study come from an original collection of biographical data on
+immigrants and the native population of Quebec to analyse their participation in
+Quebec society in various fields (social, economic, cultural, etc.). The
+immigrant population is limited to recent immigration that has taken place over
+the past 15 years. This survey has several components that retrace the life
+course of individuals over a long period of time. The pandemic component uses a
+reduced version of the [harmonized questionnaire on exposure and the impact of
+Covid19 on the family
+(CEFIS)](https://www.phenxtoolkit.org/toolkit_content/PDF/CPTS_CEFIS_FullSurveyEng.pdf).
+The questions are divided into three parts, on exposure to the disease, its
+consequences and the benefit of the policies put in place by both the federal
+and provincial governments. The other parts of the questionnaire focus on the
+characteristics of the respondents, their immigration history, their educational
+and family background. In short, it is a detailed questionnaire that covers the
+life of immigrants from birth and especially over the past 15 years. The data
+collection lasted three months, from August to November 2020. In total, 1,500
+people were surveyed, including around 1,000 immigrants and 500 natives
+throughout Quebec.
+
+### Dependent variables
+
+We distinguish three groups of dependent variables:
+
+**1. Disease exposure variables**: these are measured by six questions that seek
+to document whether the individual through work or a family member may have been
+exposed to the disease. For example, it is asked if a family member has had
+symptoms of the disease, or if a member has been hospitalized due to the
+contraction of the disease.
+
+**2. Variables on the family consequences of the disease**: we measure three
+dimensions of consequences. The first group of measures assess the consequences
+on accessing ressources by the respondent or any member of his family. The
+second group of measure assess whether the respondent or any family member have
+difficulties to perform usual duties toward any member of the family. Finally,
+the last group of measures measure the consequences of the pandemic on
+employment and income of any family member.
+
+**3. Variables on government assistance**: Finally, government support is
+measured by the following two questions that assessed whether the respondent or
+any family member has benefited of the temporary assistance for workers from the
+Government of Quebec (PATT) or of the Canadian Federal Government Emergency
+Benefit (CEP). The table XXX in the apendix presents the full set of indicators
+of the dependents variables
+
+### Main independent variable
+
+The main independent variable is immigrant status coupled with visible minority
+status. This variable takes the following modalities: 1 if the respondent is a
+non-racialized native, 2 if he is a non-racialized immigrant, 3 if he is a
+racialized immigrant, and 4 if the respondent is a racialized native or an
+immigrant who has not answered the visible minority question. The other
+variables in the analysis are: age, sex, length of residence, minority status,
+size of social network.
+
+## Analysis methods
+
+We use two main methods of analysis, a descriptive method and an explanatory
+method. In the descriptive method, we consider each of the independent variables
+in isolation and describe how they vary depending on the main independent
+variable. The explanatory methods goes further by assessing whether the effect
+remain when controlling for other covariates. Thus, by controlling for the other
+explanatory variables, we determine the effect of immigrant status on exposure
+to the pandemic, on its consequences and on the fact of benefiting or not of
+government assistance. In a second series of modelling, we assess whether the
+effect that we observed varies by the immigrant's length of residence or by
+marital status. We use either a logistic regression on the outcome variables
+that are binary and an ordonned logistic regression for the outcome that are
+measure with a 5-point Lickert scale.
+<!-- PUT THIS AS A LIMITATION IN THE CONCLUSION
+To be able to benefit from government benefits, one must meet a number of conditions, including the loss of job due to Covid-19.
+As a result, the beneficiaries are not the general population but those who have lost their jobs and this may itself be due to immigrant status.
+Thus, for model 3, we will use Heckman's models by first estimating the fact of having lost or not a job and then, the fact of receiving or not the government benefit.
+-->
+
 \newpage
-\setcounter{figure}{`r paste(no_graphic <- no_graphic+1)`}
-```
-```{r, echo=FALSE, out.width="95%", fig.cap="Processus d'analyse"}
-knitr::include_graphics("images/figure_10.png")
-```
+
+<!--chapter:end:04-methodology.Rmd-->
+
+---
+knit: bookdown::render_book
+editor_options: 
+  markdown: 
+    wrap: 80
+---
+
+# 3. Results
+
+## 3.1. Sample distribution
+
+## 3.2. Exposure by immigration status
+
+Figure 1 presents the exposure to the pandemy accroding to the immigrants and
+minority visible status. For all the six measures of exposure, it appears that
+the majority group is the most exposed to the COVID-19 in all instance. The big
+difference between the groups is about the exposure trhough working outside.
+Around 50% of the member of this group or a family member work outside the home
+at the time of the survey. This percentage is only 30% for the racialised
+immigrants and 25% for the non-racialised immigrants.
+
+![Figure 1.1: Exposure](Figures/Exposition1.png){width="60%"}
+
+![Figure 1.2: Explicative model](Figures/Exposure.jpeg)
+
+## 3.3. Consequences by immigrations status
+
+The Figure 2 presents the consequences of the COVID-19 for a family member to
+access to ressources. It shows that in all accounts but one (getting medecine),
+the racialised immigrants express more diffciult to assess ressources. Around
+20% of them stated having difficulty to get health care compare, around 10% for
+obtaining food and 7% for obtaining family insurrance. Comparatively, these
+proportion are respectively 13%, 7% and 4% for people and thier family member in
+the majority group and 19%, 3% and 4% respectively among the racialised
+immigrants.
+
+![Figure 3.1: Consequences on accessing
+ressources](Figures/Famille1.png){width="80%"}
+
+![Figure 3.2: Explicative model](Figures/Ressources.jpeg)
+
+## 3.4. Consequences on performing usual duties
+
+The Figure 4 compares the response from the three groups on performing usual
+duties. It appears that for all groups the situation caused by the COVID-19 has
+inexpectedly improved the situation in various way, irrespective of the
+immigrants status of the respondents.
+
+![Figure 4.1: Consequences on performing usual duties](Figures/Entente1.png)
+
+![Figure 4.2: Explicative model](Figures/Abilities.jpeg)
+
+## 3.5. Consequences on employment by immigration status
+
+The consequences on employment are mixed, with the severe consequences more
+prevalent among the racialised immigrants. The Figure 3 shows for instance that
+close to 45% of the racialised immigrants have stated that their family income
+has decreased compare to just 33% of the majority group and 37% of the
+non-racialed immigrants. In addition, 11% of the racialised immigrants reveals
+having loss their job compared to 9% among the non-racialised immigrants and 2%
+among members of the majority group. At the opposite, the members of the
+majority groups are more likely to report a reduction of the working hours or
+for stopping momentarily their work.
+
+![Figure 5.1: Consequences on employment and income](Figures/Emploi1.png)
+
+![Figure 5.2: Explicative model](Figures/Employment.jpeg)
+
+## 3.7 Help received by immigrations status
+
+Finally, we compared the different group on benefitting of the government
+assistance. It appears that racialised immigrants with 35% of the case have
+expressed having received the Canada Emergency Response Benefit wheras this
+proportion is 31% for the non-racialised immigrants and 25% for the member of
+the majority group. For the province Temporary Aid for Workers Program, around
+5% of the respondents have received it.
+
+![](Figures/Aide1.png)
+
+![Figure 3.2: Explicative model](Figures/Government.jpeg)
 
@@
 
-<!--chapter:end:04-methodologie.Rmd-->
-
----
-knit: bookdown::render_book
-editor_options: 
-  markdown: 
-    wrap: 80
----
-
-```{r}
-#Chargement de la base de données
-
-cr_SOC3380 <- read_csv("data/cr_SOC3380.csv")
-Pays <- read_csv("data/Pays.csv")
-
-#Codage des variables
-
-an_SOC3380 <-
-  cr_SOC3380 %>%
-  rename(
-    "age" = "1 - Quel âge avez vous ?",
-    "year_qc" = "2 - Depuis combien d'années habitez vous au Québec ?",
-    "born_country" = "3a - Dans quel pays êtes vous né ?",
-    "country_15" = "3b - Dans quel pays avez vous vécu jusqu'à vos 15 ans ?",
-    "genre" = "4 - A quel genre vous identifiez vous ?",
-    "statut" = "5 - Quel est votre statut actuel au Canada ?",
-    "etat_civil" = "6 - Quel est votre état civil ?",
-    "diplome" = "7 - Quel est votre niveau de diplôme le plus élevé ?",
-    "revenu" = "8 - Dans quelle tranche se situe votre revenu actuel ?",
-    "lang_enf" = "9 - Quelle est la première langue que vous avez apprise à la maison pendant votre enfance et que vous comprenez encore aujourd'hui ?",
-    "lang_home" = "10 - Parmi ces langues, laquelle parlez vous le plus souvent à la maison ? Indiquez plus d'une langue seulement si elles sont parlées aussi souvent l'une que l'autre à la maison.",
-    "lang_home_autre" = "11 - Parlez vous d'autres langues à la maison ?",
-    "lang_trav" = "12 -  Parmi ces langues, laquelle parlez vous le plus souvent au travail ? Indiquez plus d'une langue seulement si elles sont parlées aussi souvent l'une que l'autre au travail.",
-    "lang_trav_autre" = "13 - Parlez vous d'autres langues au travail ?",
-    "rs_util" = "14 - Quel(s) réseaux sociaux utilisez vous régulièrement ?",
-    "rs_time" = "15 - Combien de temps passez vous par jour sur les réseaux sociaux ?",
-    "rs_disc_same" = "16 - Pour quelles raisons allez-vous sur les réseaux sociaux ? [Pour discuter avec des personnes qui proviennent de mon pays d'origine]",
-    "rs_disc_diff" = "16 - Pour quelles raisons allez-vous sur les réseaux sociaux ? [Pour discuter avec des personnes qui ne proviennent pas de mon pays d'origine]",
-    "rs_hab" = "16 - Pour quelles raisons allez-vous sur les réseaux sociaux ? [Par habitude]",
-    "rs_div" = "16 - Pour quelles raisons allez-vous sur les réseaux sociaux ? [Pour me divertir]",
-    "rs_ptime" = "16 - Pour quelles raisons allez-vous sur les réseaux sociaux ? [Pour passer le temps]",
-    "rs_inf" = "16 - Pour quelles raisons allez-vous sur les réseaux sociaux ? [Pour m'informer]",
-    "rs_renc_same" = "16 - Pour quelles raisons allez-vous sur les réseaux sociaux ? [Pour rencontrer des personnes de mon pays d'origine]",
-    "rs_renc_diff" = "16 - Pour quelles raisons allez-vous sur les réseaux sociaux ? [Pour rencontrer des personnes qui ne proviennent pas de mon pays d'origine]",
-    "rs_trav" = "16 - Pour quelles raisons allez-vous sur les réseaux sociaux ? [Pour le travail]",
-    "online_memb_migr" = "17 - Êtes-vous membre de groupes en ligne regroupant des personnes immigrantes ?",
-    "inf_pol_home" = "18 - Au cours du dernier mois, à quelle fréquence avez-vous utilisé les réseaux sociaux pour lire des journaux ou des revues en ligne traitant d'une question politique ou sociale précise en lien avec votre pays d'origine ?",
-    "inf_polqc" = "19 - Au cours du dernier mois, à quelle fréquence avez-vous utilisé les réseaux sociaux pour lire des journaux ou des revues en ligne traitant d'une question politique ou sociale précise en lien avec le Québec ?",
-    "ton_migr" = "20a - Quelle est la tonalité du contenu que vous observez sur les réseaux sociaux sur le thème de l'immigration au Québec ?",
-    "ton_why" = "20b - Pourquoi selon vous ?",
-    "disc_maj_same" = "21 - Sur les réseaux sociaux, les amis avec lesquels je discute le plus souvent sont majoritairement... [Des personnes provenant de mon pays d'origine]",
-    "disc_maj_migr" = "21 - Sur les réseaux sociaux, les amis avec lesquels je discute le plus souvent sont majoritairement... [D'autres personnes immigrantes]",
-    "disc_maj_qc" = "21 - Sur les réseaux sociaux, les amis avec lesquels je discute le plus souvent sont majoritairement... [Des personnes nées au Québec]",
-    "ton_migr_pk" = "20b - Pourquoi selon vous ?",
-    "act_cult" = "22 - Durant le dernier mois, à quelle fréquence avez-vous participé à des activités (culturelle, sportive, sociale, etc.) avec une personne qui ne provient pas de votre pays d'origine ?",
-    "sentim_app" = "23 - Veuillez répondre selon votre niveau d'accord ou de désaccord avec les énoncés suivants: [J'ai le sentiment d'appartenir à la société québécoise.]",
-    "chez_moi_qc" = "23 - Veuillez répondre selon votre niveau d'accord ou de désaccord avec les énoncés suivants: [Je me sens chez moi au Québec.]",
-    "feel_queb" = "23 - Veuillez répondre selon votre niveau d'accord ou de désaccord avec les énoncés suivants: [Je me sens Québécois(e).]",
-    "forget_orig" = "23 - Veuillez répondre selon votre niveau d'accord ou de désaccord avec les énoncés suivants: [Il faut faire oublier ses origines pour se faire accepter au Québec.]",
-    "percep_qc" = "23 - Veuillez répondre selon votre niveau d'accord ou de désaccord avec les énoncés suivants: [On me voit comme un(e) Québécois(e).]",
-    "dev_qc" = "23 - Veuillez répondre selon votre niveau d'accord ou de désaccord avec les énoncés suivants: [Je veux devenir Québécois(e).]",
-    "agir_qc" = "23 - Veuillez répondre selon votre niveau d'accord ou de désaccord avec les énoncés suivants: [En public, j’essaye d’agir comme un Québécois.]",
-    "import_adapt" = "23 - Veuillez répondre selon votre niveau d'accord ou de désaccord avec les énoncés suivants: [Je crois que c’est important de m’adapter à la culture québécoise.]",
-    "iso_qc" = "23 - Veuillez répondre selon votre niveau d'accord ou de désaccord avec les énoncés suivants: [Vous vous sentez isolé(e) au Québec.]",
-    "future_live" = "24 - Dans le futur, où souhaiteriez vous vivre ?",
-    "valeur_fr" = "25 - Dans quelle mesure êtes-vous personnellement d’accord avec les valeurs suivantes? [Le français en tant que langue officielle du Québec]",
-    "valeur_dem" = "25 - Dans quelle mesure êtes-vous personnellement d’accord avec les valeurs suivantes? [Le respect de la démocratie]",
-    "valeur_equal" = "25 - Dans quelle mesure êtes-vous personnellement d’accord avec les valeurs suivantes? [L’égalité entre les sexes*]",
-    "valeur_med14" = "25 - Dans quelle mesure êtes-vous personnellement d’accord avec les valeurs suivantes?
-[La liberté de consentir ou de refuser des interventions en soins de santé dès l’âge de 14 ans]",
-    "valeur_relation" = "25 - Dans quelle mesure êtes-vous personnellement d’accord avec les valeurs suivantes? [La liberté de choisir sont statut de relation amoureuse]",
-    "valeur_laic" = "25 - Dans quelle mesure êtes-vous personnellement d’accord avec les valeurs suivantes? [La laïcité au Québec**]",
-    "valeur_droitlib" = "25 - Dans quelle mesure êtes-vous personnellement d’accord avec les valeurs suivantes? [Le respect de la Charte des droits et libertés de la personne]",
-    "qc_val_fr" = "26 - Selon vous, dans quelles mesures les québécois(es) partagent, en général, les valeurs suivantes? [Le français en tant que langue officielle du Québec]",
-    "qc_val_dem" = "26 - Selon vous, dans quelles mesures les québécois(es) partagent, en général, les valeurs suivantes? [Le respect de la démocratie]",
-    "qc_val_equal" = "26 - Selon vous, dans quelles mesures les québécois(es) partagent, en général, les valeurs suivantes? [L’égalité entre les sexes*]",
-    "qc_val_med14" = "26 - Selon vous, dans quelles mesures les québécois(es) partagent, en général, les valeurs suivantes?
-[La liberté de consentir ou de refuser des interventions en soins de santé dès l’âge de 14 ans]",
-    "qc_val_relation" = "26 - Selon vous, dans quelles mesures les québécois(es) partagent, en général, les valeurs suivantes? [La liberté de choisir sont statut de relation amoureuse]",
-    "qc_val_laic" = "26 - Selon vous, dans quelles mesures les québécois(es) partagent, en général, les valeurs suivantes? [La laïcité au Québec**]",
-    "qc_val_droitlib" = "26 - Selon vous, dans quelles mesures les québécois(es) partagent, en général, les valeurs suivantes? [Le respect de la Charte des droits et libertés de la personne]",
-    "comp_polqc" = "27 - Comment qualifiez vous votre compréhension des enjeux politiques auxquels fait face le Québec?",
-    "disc_polqc" = "28 - Durant le dernier mois, à quelle fréquence avez-vous discuté avec une autre personne d'enjeux politiques québécois ?",
-    "interp_polqc" = "29 - À quel niveau vous sentez vous interpelé(e) par les enjeux politiques au Québec ?",
-    "polqc_CAQ" = "30 - Les questions suivantes concernent les partis politiques québécois et certains évènements récents. [La Coalition Avenir Québec promulgue le communisme]",
-    "polqc_PQ" = "30 - Les questions suivantes concernent les partis politiques québécois et certains évènements récents. [Le Parti Québécois est attaché au fédéralisme]" ,
-    "polqc_PLQ" = "30 - Les questions suivantes concernent les partis politiques québécois et certains évènements récents. [Le Parti Libéral du Québec milite pour l'indépendance du Québec]",
-    "polqc_QS" = "30 - Les questions suivantes concernent les partis politiques québécois et certains évènements récents. [Québec Solidaire promulgue le socialisme démocratique]",
-    "polqc_L21" = "30 - Les questions suivantes concernent les partis politiques québécois et certains évènements récents. [La loi 21 concerne les droits des personnes LGBTQ+]",
-    "polqc_L96" = "30 - Les questions suivantes concernent les partis politiques québécois et certains évènements récents. [La loi 96 concerne l'utilisation de la langue française]",
-    "pm_qc" = "31 - À quel parti est associé le premier ministre actuel du Québec ?",
-    "maj_AN" = "32 - Quel parti occupe la majorité des sièges de l'Assemblée nationale ?",
-    "vote_age" = "33 - Quel est l'âge légal pour voter au Québec ?",
-    "act_conv" = "34 - Durant les douze derniers mois, avez-vous réalisé l'une des actions suivantes ? [J'ai essayé de convaincre quelqu'un de changer d'opinion politique]",
-    "act_inf" = "34 - Durant les douze derniers mois, avez-vous réalisé l'une des actions suivantes ? [J'ai essayé d'influencer le vote d'autres personnes]",
-    "act_opinion_pol" = "34 - Durant les douze derniers mois, avez-vous réalisé l'une des actions suivantes ? [J'ai émis mon opinion politique en public ou sur les réseaux sociaux]",
-    "act_disc_pol" = "34 - Durant les douze derniers mois, avez-vous réalisé l'une des actions suivantes ? [J'ai discuté de sujets politiques en public ou sur les réseaux sociaux]",
-    "act_contact_pol" = "34 - Durant les douze derniers mois, avez-vous réalisé l'une des actions suivantes ? [J'ai contacté un politicien ou un membre du gouvernement]",
-    "act_workpol" = "34 - Durant les douze derniers mois, avez-vous réalisé l'une des actions suivantes ? [J'ai travaillé pour un parti ou un groupe activiste]",
-    "act_symb" = "34 - Durant les douze derniers mois, avez-vous réalisé l'une des actions suivantes ? [J'ai affiché un symbole politique (sticker, badge, vêtement, etc.)]",
-    "act_pet" = "34 - Durant les douze derniers mois, avez-vous réalisé l'une des actions suivantes ? [J'ai signé une pétition]",
-    "act_manif" = "34 - Durant les douze derniers mois, avez-vous réalisé l'une des actions suivantes ? [J'ai pris part à une manifestation publique et licite]",
-    "act_collect_sign" = "34 - Durant les douze derniers mois, avez-vous réalisé l'une des actions suivantes ? [J'ai collecté des signatures pour promouvoir une pétition]",
-    "fr_lire" = "35 - Communiquer en français implique plusieurs dimensions comme la lecture, l'écoute, l'écriture et la parole.
-Veuillez évaluer votre compétence à employer le français en répondant aux énoncés suivants.
-[Je peux lire et comprendre les points principaux d'un article de presse dont le sujet m'est familier]",
-    "fr_conv" = "35 - Communiquer en français implique plusieurs dimensions comme la lecture, l'écoute, l'écriture et la parole.
-Veuillez évaluer votre compétence à employer le français en répondant aux énoncés suivants.
-[Dans une conversation, je peux parler de sujets familiers et exprimer mon opinion]",
-    "fr_ecrire" = "35 - Communiquer en français implique plusieurs dimensions comme la lecture, l'écoute, l'écriture et la parole.
-Veuillez évaluer votre compétence à employer le français en répondant aux énoncés suivants.
-[Je peux écrire des lettres à propos de mes expériences, de mes sentiments et dévènements]",
-    "fr_radiotele" = "35 - Communiquer en français implique plusieurs dimensions comme la lecture, l'écoute, l'écriture et la parole.
-Veuillez évaluer votre compétence à employer le français en répondant aux énoncés suivants.
-[Je comprends les points principaux d'une émission télévisée ou radiodiffusée à propos d'un sujet familier]",
-    "comment" = "Avez-vous des commentaires que vous aimeriez partager avec nous ?",
-    "int_quali" = "Dans la cadre de cette même recherche, avez vous un intérêt à participer à une séance de discussion? Si oui,  veuillez partager votre courriel.",
-    "continuant" = "En continuant :"
-    )
-
-#Exclusion des répondants établis au Québec depuis moins d'un an
-
-an_SOC3380 <- an_SOC3380[an_SOC3380$year_qc != "Moins d'un an", ]
-
-#Exclusion des variables inutiles à l'enquête et transformation des variables utiles en variables factorielles
-
-an_SOC3380 <-
-  an_SOC3380 %>%
-  mutate_if(sapply(an_SOC3380, is.character), as.factor)%>%
-    select(!c(Horodateur, continuant, comment, int_quali))
-
-#Multiplication des observations par tirage d'échantillons aléatoires
-
-set.seed(1234)
-Sample_1 <- slice_sample(an_SOC3380, n = 15)
-set.seed(1235)
-Sample_2 <- slice_sample(an_SOC3380, n = 15)
-set.seed(1236)
-Sample_3 <- slice_sample(an_SOC3380, n = 15)
-set.seed(1237)
-Sample_4 <- slice_sample(an_SOC3380, n = 15)
-set.seed(1238)
-Sample_5 <- slice_sample(an_SOC3380, n = 15)
-set.seed(1239)
-Sample_6 <- slice_sample(an_SOC3380, n = 15)
-
-an_SOC3380 <- bind_rows(an_SOC3380, Sample_1, Sample_2, Sample_3, Sample_4, Sample_5, Sample_6)
-
-#correction de variables
-
-an_SOC3380 <-
-  an_SOC3380 %>% 
-  mutate(born_country = recode(born_country, `Ivory Coast (Cote d'Ivoire)` = "Côte D'Ivoire"),
-         country_15 = recode(born_country, `Ivory Coast (Cote d'Ivoire)` = "Côte D'Ivoire"),
-         born_country = recode(born_country, `Iran` = "Iran, Islamic Republic of"),
-         country_15 = recode(country_15, `Iran` = "Iran, Islamic Republic of"))
-
-an_SOC3380 <-
-  an_SOC3380 %>%
-  mutate(born_country = recode(born_country, `Ivory Coast (Cote d'Ivoire)` = "Côte D'Ivoire"),
-         country_15 = recode(born_country, `Ivory Coast (Cote d'Ivoire)` = "Côte D'Ivoire"),
-         born_country = recode(born_country, `Iran` = "Iran, Islamic Republic of"),
-         country_15 = recode(country_15, `Iran` = "Iran, Islamic Republic of"),
-         genre = recode(genre, Fluide = "Autres"),
-         genre = recode(genre, `Je ne souhaite pas répondre` = "Autres"),
-         lang_enf = case_when(
-           lang_enf == "Français" ~ "Français",
-           lang_enf == "Anglais" ~ "Anglais",
-           TRUE ~ "Autres"),
-         lang_home = case_when(
-           lang_home == "Français" ~ "Français",
-           lang_home == "Anglais" ~ "Anglais",
-           TRUE ~ "Autres"))
-
-an_SOC3380 <- left_join(an_SOC3380,Pays, by = "born_country")
-
-#numérisation de variables
-
-an_SOC3380 <-
-  an_SOC3380 %>%
-  mutate(Facebook = if_else(str_detect(rs_util, "Facebook", negate = FALSE), "Oui", "Non"),
-         Twitter = if_else(str_detect(rs_util, "Twitter", negate = FALSE), "Oui", "Non"),
-         Youtube = if_else(str_detect(rs_util, "Youtube", negate = FALSE), "Oui", "Non"),
-         Instagram = if_else(str_detect(rs_util, "Instagram", negate = FALSE), "Oui", "Non"),
-         Messenger = if_else(str_detect(rs_util, "Messenger", negate = FALSE), "Oui", "Non"),
-         WhatsApp = if_else(str_detect(rs_util, "WhatsApp", negate = FALSE), "Oui", "Non"),
-         WeChat = if_else(str_detect(rs_util, "WeChat", negate = FALSE), "Oui", "Non"),
-         TikTok = if_else(str_detect(rs_util, "TikTok", negate = FALSE), "Oui", "Non"),
-         Douyin = if_else(str_detect(rs_util, "Douyin", negate = FALSE), "Oui", "Non"),
-         Telegram = if_else(str_detect(rs_util, "Telegram", negate = FALSE), "Oui", "Non"),
-         QQ = if_else(str_detect(rs_util, "QQ", negate = FALSE), "Oui", "Non"),
-         Weibo = if_else(str_detect(rs_util, "Weibo", negate = FALSE), "Oui", "Non"),
-         Qzone = if_else(str_detect(rs_util, "Qzone", negate = FALSE), "Oui", "Non"),
-         Snapchat = if_else(str_detect(rs_util, "Snapchat", negate = FALSE), "Oui", "Non"),
-         Kuaishou = if_else(str_detect(rs_util, "Kuaishou", negate = FALSE), "Oui", "Non"),
-         Pinterest = if_else(str_detect(rs_util, "Pinterest", negate = FALSE), "Oui", "Non"),
-         Reddit = if_else(str_detect(rs_util, "Reddit", negate = FALSE), "Oui", "Non"),
-         LinkedIn = if_else(str_detect(rs_util, "LinkedIn", negate = FALSE), "Oui", "Non"),
-         Quora = if_else(str_detect(rs_util, "Quora", negate = FALSE), "Oui", "Non"),
-         Skype = if_else(str_detect(rs_util, "Skype", negate = FALSE), "Oui", "Non"),
-         Tieba = if_else(str_detect(rs_util, "Tieba", negate = FALSE), "Oui", "Non"),
-         Viber = if_else(str_detect(rs_util, "Viber", negate = FALSE), "Oui", "Non"),
-         Teams = if_else(str_detect(rs_util, "Teams", negate = FALSE), "Oui", "Non"),
-         imo = if_else(str_detect(rs_util, "imo", negate = FALSE), "Oui", "Non"),
-         Line = if_else(str_detect(rs_util, "Line", negate = FALSE), "Oui", "Non"),
-         Piscart = if_else(str_detect(rs_util, "Piscart", negate = FALSE), "Oui", "Non"),
-         Likee = if_else(str_detect(rs_util, "Likee", negate = FALSE), "Oui", "Non"),
-         Discord = if_else(str_detect(rs_util, "Discord", negate = FALSE), "Oui", "Non"),
-         Twitch = if_else(str_detect(rs_util, "Twitch", negate = FALSE), "Oui", "Non"),
-         `Stack Exchange` = if_else(str_detect(rs_util, "Stack Exchange", negate = FALSE), "Oui", "Non"),
-         Signal = if_else(str_detect(rs_util, "Signal", negate = FALSE), "Oui", "Non")) %>%
-  mutate_at(
-    vars(one_of("Facebook",
-                "Twitter",
-                "Youtube",
-                "Instagram",
-                "Messenger",
-                "WhatsApp",
-                "WeChat",
-                "TikTok",
-                "Douyin",
-                "Telegram",
-                "QQ",
-                "Weibo",
-                "Qzone",
-                "Snapchat",
-                "Kuaishou",
-                "Pinterest",
-                "Reddit",
-                "Facebook",
-                "LinkedIn",
-                "Quora",
-                "Skype",
-                "Tieba",
-                "Viber",
-                "Teams",
-                "imo",
-                "Line",
-                "Piscart",
-                "Discord",
-                "Piscart",
-                "Likee",
-                "Twitch",
-                "Signal",
-                "Stack Exchange"
-                )),
-    funs(case_when(
-      . == "Oui" ~ 1,
-      . == "Non" ~ 0)
-      )) %>%
-  rowwise()%>%
-  mutate(nbr_rs = sum(Facebook,
-                      Twitter,
-                      Youtube,
-                      Instagram,
-                      Messenger,
-                      WhatsApp,
-                      WeChat,
-                      TikTok,
-                      Douyin,
-                      Telegram,
-                      QQ,
-                      Weibo,
-                      Qzone,
-                      Snapchat, 
-                      Kuaishou,
-                      Kuaishou,
-                      Reddit,
-                      Facebook,
-                      LinkedIn,
-                      Quora,
-                      Skype,
-                      Tieba,
-                      imo,
-                      Line,
-                      Piscart,
-                      Discord,
-                      Piscart,
-                      Twitch,
-                      Signal,
-                      `Stack Exchange`,
-                      na.rm = TRUE))
-
-an_SOC3380 <-
-  an_SOC3380 %>%
-  mutate(region_n = case_when(
-    region == "Europe" ~ 1,
-    region == "South/Latin America" ~ 2,
-    region == "Arab States" ~ 3,
-    region == "North America" ~ 4,
-    region == "Asia & Pacific" ~ 5,
-    region == "Africa" ~ 6,
-    region == "Middle east" ~ 7
-  ))
-
-#préparation des indices
-
-accord_lvl <- c("Pas du tout d'accord", 
-                "Plutôt pas d'accord", 
-                "Plutôt d'accord", 
-                "D'accord", 
-                "Tout à fait d'accord")
-
-INV_accord_LVL <- c("Tout à fait d'accord", 
-                    "D'accord", 
-                    "Plutôt d'accord", 
-                    "Plutôt pas d'accord", 
-                    "Pas du tout d'accord")
-
-mesure_lvl <- c("Pas du tout", 
-                "Dans une certaine mesure", 
-                "Dans une grande mesure")
-
-IND_LVL <- c("Faible", 
-             "Modéré", 
-             "Fort")
-
-#variables indépendantes
-
-##rs_disc_diff
-
-an_SOC3380$rs_disc_diff <- recode(an_SOC3380$rs_disc_diff, "Pas d'accord" = "Plutôt pas d'accord")
-an_SOC3380$rs_disc_diff <- factor(an_SOC3380$rs_disc_diff, levels = accord_lvl, ordered = TRUE)
-an_SOC3380$rs_disc_diff <- as.numeric(an_SOC3380$rs_disc_diff, na.rm = TRUE)
-
-##rs_renc_diff
-
-an_SOC3380$rs_renc_diff <- recode(an_SOC3380$rs_renc_diff, "Pas d'accord" = "Plutôt pas d'accord")
-an_SOC3380$rs_renc_diff <- factor(an_SOC3380$rs_renc_diff, levels = accord_lvl, ordered = TRUE)
-an_SOC3380$rs_renc_diff <- as.numeric(an_SOC3380$rs_renc_diff, na.rm = TRUE)
-
-##disc_maj_same
-
-an_SOC3380$disc_maj_same <- factor(an_SOC3380$disc_maj_same, levels = accord_lvl, ordered = TRUE)
-an_SOC3380$disc_maj_same <- as.numeric(an_SOC3380$disc_maj_same, na.rm = TRUE)
-
-##disc_maj_migr
-
-an_SOC3380$disc_maj_migr <- factor(an_SOC3380$disc_maj_migr, levels = accord_lvl, ordered = TRUE)
-an_SOC3380$disc_maj_migr <- as.numeric(an_SOC3380$disc_maj_migr, na.rm = TRUE)
-
-##disc_maj_qc
-
-an_SOC3380$disc_maj_qc <- factor(an_SOC3380$disc_maj_qc, levels = accord_lvl, ordered = TRUE)
-an_SOC3380$disc_maj_qc <- as.numeric(an_SOC3380$disc_maj_qc, na.rm = TRUE)
-
-## online_memb_migr
-
-an_SOC3380$online_memb_migr <- factor(an_SOC3380$online_memb_migr, levels = c("Oui", "Non"), ordered = FALSE)
-
-#variables dépendantes
-
-##sentim_app
-
-an_SOC3380$sentim_app <- recode(an_SOC3380$sentim_app, "Pas d'accord" = "Plutôt pas d'accord")
-an_SOC3380$sentim_app <- factor(an_SOC3380$sentim_app, levels = accord_lvl, ordered = TRUE)
-an_SOC3380$sentim_app <- as.numeric(an_SOC3380$sentim_app, na.rm = TRUE)
-
-##chez_moi_qc
-
-an_SOC3380$chez_moi_qc <- recode(an_SOC3380$chez_moi_qc, "Pas d'accord" = "Plutôt pas d'accord")
-an_SOC3380$chez_moi_qc <- factor(an_SOC3380$chez_moi_qc, levels = accord_lvl, ordered = TRUE)
-an_SOC3380$chez_moi_qc <- as.numeric(an_SOC3380$chez_moi_qc, na.rm = TRUE)
-
-##feel_queb
-
-an_SOC3380$feel_queb <- recode(an_SOC3380$feel_queb, "Pas d'accord" = "Plutôt pas d'accord")
-an_SOC3380$feel_queb <- factor(an_SOC3380$feel_queb, levels = accord_lvl, ordered = TRUE)
-an_SOC3380$feel_queb <- as.numeric(an_SOC3380$feel_queb, na.rm = TRUE)
-
-##forget_orig
-
-an_SOC3380$forget_orig <- recode(an_SOC3380$forget_orig, "Pas d'accord" = "Plutôt pas d'accord")
-an_SOC3380$forget_orig <- factor(an_SOC3380$forget_orig, levels = INV_accord_LVL, ordered = TRUE)
-an_SOC3380$forget_orig <- as.numeric(an_SOC3380$forget_orig, na.rm = TRUE)
-
-##percep_qc
-
-an_SOC3380$percep_qc <- recode(an_SOC3380$percep_qc, "Pas d'accord" = "Plutôt pas d'accord")
-an_SOC3380$percep_qc <- factor(an_SOC3380$percep_qc, levels = accord_lvl, ordered = TRUE)
-an_SOC3380$percep_qc <- as.numeric(an_SOC3380$percep_qc, na.rm = TRUE)
-
-##dev_qc
-
-an_SOC3380$dev_qc <- recode(an_SOC3380$dev_qc, "Pas d'accord" = "Plutôt pas d'accord")
-an_SOC3380$dev_qc <- factor(an_SOC3380$dev_qc, levels = accord_lvl, ordered = TRUE)
-an_SOC3380$dev_qc <- as.numeric(an_SOC3380$dev_qc, na.rm = TRUE)
-
-##agir_qc
-
-an_SOC3380$agir_qc <- recode(an_SOC3380$agir_qc, "Pas d'accord" = "Plutôt pas d'accord")
-an_SOC3380$agir_qc <- factor(an_SOC3380$agir_qc, levels = accord_lvl, ordered = TRUE)
-an_SOC3380$agir_qc <- as.numeric(an_SOC3380$agir_qc, na.rm = TRUE)
-
-##import_adapt
-
-an_SOC3380$import_adapt <- recode(an_SOC3380$import_adapt, "Pas d'accord" = "Plutôt pas d'accord")
-an_SOC3380$import_adapt <- factor(an_SOC3380$import_adapt, levels = accord_lvl, ordered = TRUE)
-an_SOC3380$import_adapt <- as.numeric(an_SOC3380$import_adapt, na.rm = TRUE)
-
-##iso_qc
-
-an_SOC3380$iso_qc <- recode(an_SOC3380$iso_qc, "Pas d'accord" = "Plutôt pas d'accord")
-an_SOC3380$iso_qc <- factor(an_SOC3380$iso_qc, levels = INV_accord_LVL, ordered = TRUE)
-an_SOC3380$iso_qc <- as.numeric(an_SOC3380$iso_qc, na.rm = TRUE)
-
-##valeur_fr
-
-an_SOC3380$valeur_fr <- factor(an_SOC3380$valeur_fr, levels = mesure_lvl, ordered = TRUE)
-an_SOC3380$valeur_fr <- as.numeric(an_SOC3380$valeur_fr, na.rm = TRUE)
-
-##valeur_dem
-
-an_SOC3380$valeur_dem <- factor(an_SOC3380$valeur_dem, levels = mesure_lvl, ordered = TRUE)
-an_SOC3380$valeur_dem <- as.numeric(an_SOC3380$valeur_dem, na.rm = TRUE)
-
-##valeur_equal
-
-an_SOC3380$valeur_equal <- factor(an_SOC3380$valeur_equal, levels = mesure_lvl, ordered = TRUE)
-an_SOC3380$valeur_equal <- as.numeric(an_SOC3380$valeur_equal, na.rm = TRUE)
-
-##valeur_med14
-
-an_SOC3380$valeur_med14 <- factor(an_SOC3380$valeur_med14, levels = mesure_lvl, ordered = TRUE)
-an_SOC3380$valeur_med14 <- as.numeric(an_SOC3380$valeur_med14, na.rm = TRUE)
-
-##valeur_relation
-
-an_SOC3380$valeur_relation <- factor(an_SOC3380$valeur_relation, levels = mesure_lvl, ordered = TRUE)
-an_SOC3380$valeur_relation <- as.numeric(an_SOC3380$valeur_relation, na.rm = TRUE)
-
-##valeur_laic
-
-an_SOC3380$valeur_laic <- factor(an_SOC3380$valeur_laic, levels = mesure_lvl, ordered = TRUE)
-an_SOC3380$valeur_laic <- as.numeric(an_SOC3380$valeur_laic, na.rm = TRUE)
-
-##valeur_droitlib
-
-an_SOC3380$valeur_droitlib <- factor(an_SOC3380$valeur_droitlib, levels = mesure_lvl, ordered = TRUE)
-an_SOC3380$valeur_droitlib <- as.numeric(an_SOC3380$valeur_droitlib, na.rm = TRUE)
-
-##comp_polqc
-
-an_SOC3380$comp_polqc <- factor(an_SOC3380$comp_polqc, levels = c("Pas bonne du tout", 
-                                                                  "Pas bonne", 
-                                                                  "Plutôt bonne", 
-                                                                  "Bonne", 
-                                                                  "Très bonne"), ordered = TRUE)
-
-an_SOC3380$comp_polqc <- as.numeric(an_SOC3380$comp_polqc, na.rm = TRUE)
-
-##disc_polqc
-
-an_SOC3380$disc_polqc <- recode(an_SOC3380$disc_polqc, "Rarement" = "Quelques fois")
-
-an_SOC3380$disc_polqc <- factor(an_SOC3380$disc_polqc, levels = c("Jamais", 
-                                                                  "Quelques fois", 
-                                                                  "Plusieurs fois", 
-                                                                  "Très souvent", 
-                                                                  "Tout le temps"), ordered = TRUE)
-
-an_SOC3380$disc_polqc <- as.numeric(an_SOC3380$disc_polqc, na.rm = TRUE)
-
-##interp_polqc
-
-an_SOC3380$interp_polqc <- factor(an_SOC3380$interp_polqc, levels = c("Pas interpelé(e)", 
-                                                                      "Pas vraiment interpelé(e)", 
-                                                                      "Plutôt interpelé(e)", 
-                                                                      "Interpelé(e)", 
-                                                                      "Très interpelé(e)"), ordered = TRUE)
-
-an_SOC3380$interp_polqc <- as.numeric(an_SOC3380$interp_polqc, na.rm = TRUE)
-
-##qc_val_fr
-
-an_SOC3380$qc_val_fr <- factor(an_SOC3380$qc_val_fr, levels = mesure_lvl, ordered = TRUE)
-an_SOC3380$qc_val_fr <- as.numeric(an_SOC3380$qc_val_fr, na.rm = TRUE)
-
-##qc_val_dem
-
-an_SOC3380$qc_val_dem <- factor(an_SOC3380$qc_val_dem, levels = mesure_lvl, ordered = TRUE)
-an_SOC3380$qc_val_dem <- as.numeric(an_SOC3380$qc_val_dem, na.rm = TRUE)
-
-##qc_val_equal
-
-an_SOC3380$qc_val_equal <- factor(an_SOC3380$qc_val_equal, levels = mesure_lvl, ordered = TRUE)
-an_SOC3380$qc_val_equal <- as.numeric(an_SOC3380$qc_val_equal, na.rm = TRUE)
-
-##qc_val_med14
-
-an_SOC3380$qc_val_med14 <- factor(an_SOC3380$qc_val_med14, levels = mesure_lvl, ordered = TRUE)
-an_SOC3380$qc_val_med14 <- as.numeric(an_SOC3380$qc_val_med14, na.rm = TRUE)
-
-##qc_val_relation
-
-an_SOC3380$qc_val_relation <- factor(an_SOC3380$qc_val_relation, levels = mesure_lvl, ordered = TRUE)
-an_SOC3380$qc_val_relation <- as.numeric(an_SOC3380$qc_val_relation, na.rm = TRUE)
-
-##qc_val_laic
-
-an_SOC3380$qc_val_laic <- factor(an_SOC3380$qc_val_laic, levels = mesure_lvl, ordered = TRUE)
-an_SOC3380$qc_val_laic <- as.numeric(an_SOC3380$qc_val_laic, na.rm = TRUE)
-
-##qc_val_droitlib
-
-an_SOC3380$qc_val_droitlib <- factor(an_SOC3380$qc_val_droitlib, levels = mesure_lvl, ordered = TRUE)
-an_SOC3380$qc_val_droitlib <- as.numeric(an_SOC3380$qc_val_droitlib, na.rm = TRUE)
-
-# Contrôle
-
-an_SOC3380$rs_time <- factor(an_SOC3380$rs_time, levels = c("Moins d'une heure", 
-                                                            "Entre une et trois heures", 
-                                                            "Entre trois et cinq heures", 
-                                                            "Entre cinq et sept heures", 
-                                                            "Plus de sept heures"), ordered = TRUE)
-
-an_SOC3380$rs_time_n <- as.numeric(an_SOC3380$rs_time, na.rm = TRUE)
-
-## âge
-
-an_SOC3380$age <- factor(an_SOC3380$age, levels = c("Moins de 18 ans", 
-                                                    "Entre 18 et 25 ans", 
-                                                    "Entre 26 et 35 ans", 
-                                                    "Entre 36 et 50 ans", 
-                                                    "Plus de 50 ans"), ordered = TRUE)
-
-an_SOC3380$age_n <- as.numeric(an_SOC3380$age, na.omit(NA))
-
-v_sentimapp <- c("sentim_app", 
-                 "chez_moi_qc", 
-                 "feel_queb", 
-                 "forget_orig", 
-                 "percep_qc", 
-                 "dev_qc", 
-                 "agir_qc", 
-                 "import_adapt", 
-                 "iso_qc")
-
-v_NV <- c("valeur_fr", 
-          "valeur_dem", 
-          "valeur_equal", 
-          "valeur_med14", 
-          "valeur_relation", 
-          "valeur_laic", 
-          "valeur_droitlib")
-
-v_rsContact <- c("rs_disc_diff", 
-                 "disc_maj_qc", 
-                 "rs_renc_diff")
-
-v_percepQC <- c("qc_val_fr", 
-                "qc_val_dem", 
-                "qc_val_equal", 
-                "qc_val_med14", 
-                "qc_val_relation", 
-                "qc_val_laic", 
-                "qc_val_droitlib")
-
-accord_lvl <- c("Pas du tout d'accord", 
-                "Plutôt pas d'accord", 
-                "Plutôt d'accord", 
-                "D'accord", 
-                "Tout à fait d'accord")
-
-an_SOC3380$rs_inf <- factor(an_SOC3380$rs_inf, levels = accord_lvl, ordered = TRUE)
-
-an_SOC3380$rs_inf_n <- as.numeric(an_SOC3380$rs_inf, na.omit(NA))
-
-an_SOC3380$ton_migr <- recode(an_SOC3380$ton_migr, "Très positive" = "Positive")
-
-an_SOC3380$ton_migr <- recode(an_SOC3380$ton_migr, "Très négative" = "Négsative")
-
-ton_migr_lvl <- c("Négative", 
-                  "Plutôt négative", 
-                  "Neutre", 
-                  "Plutôt positive", 
-                  "Positive")
-
-an_SOC3380$ton_migr <- factor(an_SOC3380$ton_migr, levels = ton_migr_lvl, ordered = TRUE)
-an_SOC3380$ton_migr_n <- as.numeric(an_SOC3380$ton_migr, na.omit(NA))
-
-an_SOC3380$year_qc <- factor(an_SOC3380$year_qc, levels = c("1 an", 
-                                                            "2 ans", 
-                                                            "3 ans", 
-                                                            "4 ans", 
-                                                            "5 ans", 
-                                                            "6 ans", 
-                                                            "7 ans", 
-                                                            "8 ans", 
-                                                            "9 ans", 
-                                                            "10 ans", 
-                                                            "Plus de 10 ans"), ordered = TRUE)
-
-an_SOC3380$year_qc_n <- as.numeric(an_SOC3380$year_qc, na.omit(NA))
-
-v_score_ExpoDisc <- c("rs_time_n", 
-                      "ton_migr_n", 
-                      "rs_inf_n")
-
-v_media_accueil <- c("Youtube", 
-                     "Instagram", 
-                     "TikTok", 
-                     "Snapchat", 
-                     "Pinterest", 
-                     "Piscart", 
-                     "Twitch", 
-                     "Twitter", 
-                     "Reddit", 
-                     "Quora", 
-                     "Teams", 
-                     "Discord", 
-                     "Stack Exchange", 
-                     "Facebook", 
-                     "Messenger", 
-                     "Telegram", 
-                     "LinkedIn", 
-                     "Skype", 
-                     "Viber", 
-                     "imo", 
-                     "Signal")
-
-v_media_diasp <- c("Douyin", 
-                   "Kuaishou", 
-                   "Likee", 
-                   "Tieba", 
-                   "WeChat", 
-                   "QQ", 
-                   "Qzone", 
-                   "Line", 
-                   "WhatsApp")
-
-#création des indices
-
-an_SOC3380 <-
-  an_SOC3380 %>%
-  rowwise() %>%
-  mutate(score_sentimapp = (sum(c_across(all_of(v_sentimapp)), na.rm = TRUE) - length(v_sentimapp)) / (4*length(v_sentimapp))) %>%
-  mutate(score_sentimapp = if (score_sentimapp < 0) NA_real_ else score_sentimapp) %>%
-  ungroup()
-
-an_SOC3380 <- 
-  an_SOC3380 %>%
-  rowwise() %>%
-  mutate(score_rsContact = (sum(c_across(all_of(v_rsContact)), na.rm = TRUE) -length(v_rsContact)) / (4*length(v_rsContact))) %>%
-  mutate(score_rsContact = if (score_rsContact < 0)  NA_real_ else score_rsContact) %>%
-  ungroup()
-
-an_SOC3380 <- 
-  an_SOC3380 %>%
-  rowwise() %>%
-  mutate(score_NV = (sum(c_across(all_of(v_NV)), na.rm = TRUE) -length(v_NV)) / (2*length(v_NV))) %>%
-  mutate(score_NV = if (score_NV < 0) NA_real_ else score_NV) %>%
-  ungroup()
-
-an_SOC3380 <- 
-  an_SOC3380 %>%
-  rowwise() %>%
-  mutate(score_percepQC = (sum(c_across(all_of(v_percepQC)), na.rm = TRUE) -length(v_percepQC)) / (2*length(v_percepQC))) %>%
-  mutate(score_percepQC = if (score_percepQC < 0) NA_real_ else score_percepQC) %>%
-  ungroup()
-
-an_SOC3380 <- 
-  an_SOC3380 %>%
-  rowwise() %>%
-  mutate(score_ExpoDisc = (sum(c_across(all_of(v_score_ExpoDisc)), na.rm = TRUE) - length(v_score_ExpoDisc)) / (4*length(v_score_ExpoDisc))) %>%
-  mutate(score_ExpoDisc = if_else(score_ExpoDisc < 0 | score_ExpoDisc == 0, NA_real_, score_ExpoDisc)) %>%
-  ungroup()
-
-an_SOC3380 <-
-  an_SOC3380 %>%
-  rowwise() %>%
-  mutate(score_media_accueil = (sum(c_across(all_of(v_media_accueil)), na.rm = TRUE)) / nbr_rs)
-
-an_SOC3380 <- 
-  an_SOC3380 %>%
-  rowwise() %>%
-  mutate(score_media_diasp = (sum(c_across(all_of(v_media_diasp)), na.rm = TRUE)) / nbr_rs)
-
-# Traitement des types de réseaux sociaux
-an_SOC3380 <- an_SOC3380 %>%
-  mutate(rs_partage_medias = if_else(str_detect(rs_util, "Youtube|Instagram|TikTok|Douyin|Snapchat|Kuaishou|Pinterest|Piscart|Likee|Twitch", negate = FALSE), 1, 0),
-         rs_blogues = if_else(str_detect(rs_util, "Twitter|Weibo", negate = FALSE), 1, 0),
-         rs_commun_online = if_else(str_detect(rs_util, "Reddit|Quora|Tieba|Teams|Discord|Stack Exchange", negate = FALSE), 1, 0),
-         rs_SRS = if_else(str_detect(rs_util, "Facebook|Messenger|WhatsApp|WeChat|Telegram|QQ|Qzone|LinledIn|Skype|Viber|imo|Line|Signal", negate = FALSE), 1, 0))
-
-an_SOC3380 <-
-  an_SOC3380 %>%
-  mutate(rs_blogues_f = case_when(
-    rs_blogues == 0 ~ "Non",
-    rs_blogues == 1 ~ "Oui"
-  ))
-
-an_SOC3380 <-
-  an_SOC3380 %>%
-  mutate(rs_commun_online_f = case_when(
-    rs_commun_online == 0 ~ "Non",
-    rs_commun_online == 1 ~ "Oui"
-  ))
-
-an_SOC3380 <-
-  an_SOC3380 %>%
-  mutate(rs_SRS_f = case_when(
-    rs_SRS == 0 ~ "Non",
-    rs_SRS == 1 ~ "Oui"
-  ))
-
-an_SOC3380 <-
-  an_SOC3380 %>%
-  mutate(rs_partage_medias_f = case_when(
-    rs_partage_medias == 0 ~ "Non",
-    rs_partage_medias == 1 ~ "Oui"
-  ))
-
-an_SOC3380 <- an_SOC3380 %>%
-  mutate(rs_home = if_else(str_detect(rs_util,"Youtube|Instagram|TikTok|Snapchat|Pinterest|Piscart|Twitch|Twitter|Reddit|Quora|Teams|Discord|Stack Exchange|Facebook|Messenger|Telegram|LinkedIn|Skype|Viber|imo|Signal", negate = FALSE), 1, 0))
-
-an_SOC3380 <- an_SOC3380 %>%
-  mutate(rs_diasp = if_else(str_detect(rs_util, "Douyin|Kuaishou|Likee|Tieba|WeChat|QQ|Qzone|Line|WhatsApp|Messenger|Skype", negate = FALSE), 1, 0))
-
-an_SOC3380 <-
-  an_SOC3380 %>%
-  mutate(rs_home_f = case_when(
-    rs_home == 0 ~ "Non",
-    rs_home == 1 ~ "Oui"
-  ))
-
-an_SOC3380 <-
-  an_SOC3380 %>%
-  mutate(rs_diasp_f = case_when(
-    rs_diasp == 0 ~ "Non",
-    rs_diasp == 1 ~ "Oui"
-  ))
-
-an_SOC3380 <-
-  an_SOC3380 %>%
-  mutate(year_qc_f = case_when(
-    year_qc_n < 6 ~ "1 à 5 ans",
-    year_qc_n == 6| year_qc_n == 7|year_qc_n == 8|year_qc_n == 9 ~ "6 à 9 ans",
-    year_qc_n > 9 ~ "dix ans et plus"
-  ))
-```
-
-# Résultats
-
-## H1 - Contact interculturel et sentiment d'appartenance
-
-Le modèle de régression suivant présente la relation entre l'indice de contact
-interculturel sur les réseaux sociaux et l'indice du sentiment d'appartenance au
-pays d'accueil en fonction du nombre d'années passées au Québec, de la région
-d'origine et de l'âge. Les résultats indiquent que la relation entre le contact
-interculturel et le sentiment d'appartenance est statistiquement significative
-dans la population à l'intervalle de confiance de 95% (p \<= 0.05).
-
-```{r}
-local({
-  .Table <- lm(score_rsContact ~ score_sentimapp + year_qc_n + region_n + age_n, an_SOC3380)
-  pander(.Table, split.cell=80, split.table = Inf, 
-         caption = "Modèle de régression contrôlé pour l'hypothèse H1",
-         col.names = c("Estimation", "Erreur type", "Valeur t", "Valeur p"),
-         row.names = c("(Intercept)", "Sentiment d'appartenance", "Année passées au Québec", "Région d'origine", "Âge")
-         )
-})
-```
-
-En accord avec les résultats du modèle de régression, le test de corrélation de
-Pearson suivant présente l'intensité et le sens de la relation étudiée. Les
-résultats obtenus indiquent que la relation est faible et positive (C = 0.223).
-
-```{r}
-local({
-  .Table <- cor.test(an_SOC3380$score_rsContact, an_SOC3380$score_sentimapp)
-  pander(.Table, split.cell=80, split.table = Inf, 
-         caption = "Test de corrélation de Pearson pour l'hypothèse H1",
-         col.names = c("Valeur t", "Degrés de liberté", "Valeur C", "Hypothèse alternatif", "Valeur C")
-  )
-})
-```
-
-Le graphique suivant illustre la relation entre le contact interculturel et le
-sentiment d'appartenance. En accord avec les résultats du test de corrélation de
-Pearson, on observe que la relation est positive et faible. En d'autres termes :
-pour les personnes immigrantes établies au Québec, l'entretien de contacts sur
-les réseaux sociaux en ligne avec des personnes qui ne proviennent de leur pays
-d'origine est faiblement et positivement corrélé avec le développement d'un
-sentiment d'appartenance au pays d'accueil.
-
-```{r}
-ggplot(an_SOC3380 %>% filter(!is.na(score_rsContact), !is.na(score_sentimapp)), aes(x= score_rsContact, y = score_sentimapp))+
-  geom_point() +
-  geom_smooth(method = "lm", se= FALSE)+
-  labs(
-    title = paste("Figure", no_graphic <- no_graphic+1, ":",
-      "Contact interculturel et sentiment d'appartenance"),
-    x = "Contact interculturel",
-    y = "Sentiment d'appartenance au Québec"
-  ) +
-  theme(text = element_text(hjust = 0.5))
-```
-
-### H1.1 - Pays d'origine, discrimination systémique et sentiment d'appartenance
-
-Les témoignages de nos répondants indiquent que les discours anti-immigrations
-présents sur les réseaux sociaux en ligne participent à un processus de
-discrimination plus large - en ce qu'il constitue la traduction de pratiques
-existant hors ligne - les empêchant de s'identifier pleinement au Québec, et
-partant, de développer un sentiment d'appartenance envers leur pays d'accueil.
-Ce phénomène est pondéré par les caractéristiques socio-démographiques de nos
-répondants, selon un gradient qui rappelle les travaux de Wacquant sur la
-*racialisation diagonale* [-@wacquantResolvingTroubleRace2022]. Les personnes
-les plus racialisées comme Wendy (jeune femme noire originaire d'Haïti établie
-au Québec depuis son enfance) affirment que leur incapacité à s'identifier comme
-québécoise est directement liée à la discrimination raciale :
-
-> Wendy: Il y a toujours quelqu'un, à un moment donné dans ta vie, qui va te
-> rappeler ta condition d'immigrée. Cela peut être un commerçant par exemple :
-> il voit une personne noire entrer et automatiquement, il va commencer à te
-> regarder dans toutes les allées pour vérifier si tu ne vas pas voler ; même si
-> tu viens régler tes achats à la caisse. Le regard qu'il peut avoir sur toi
-> fait toute la différence (...) D'entendre : "toi je t'aime bien, tu n'es pas
-> comme les autres." Ça me rappelle que je ne suis pas québécoise à leurs yeux.
->
-> Théa: Est-ce que tu te sens Québécoise aujourd'hui ?
->
-> Wendy: Non, je ne me sens pas québécoise (ton amusé) et je ne pourrais jamais
-> me sentir québécoise à 100%. Je ne pourrais jamais renier mon identité, même
-> si je ne peux pas retourner vivre dans mon pays (...) je me suis adaptée (le
-> mot est appuyé) [Au Québec].C'est là la différence.
-
-En accord avec les résultats du modèle de régression (Tableau 2), le graphique
-suivant illustre la relation entre le contact interculturel et le sentiment
-d'appartenance en fonction de la région d'origine. Les résultats révèlent une
-particularité à l'endroit des personnes immigrantes originaires du Maghreb, du
-Proche-Orient et du Moyen-Orient.
-
-```{r}
-ggplot(an_SOC3380 %>% filter(!is.na(score_sentimapp)), aes(x= score_rsContact, y = score_sentimapp, color = region))+
-  geom_point() +
-  geom_smooth(method = "lm", se= FALSE)+
-  theme_minimal()+
-  scale_color_brewer(name = "Region", palette = "Set1") +
-  labs(
-    title = paste("Graphique", no_graphic <- no_graphic+1, ":",
-      "Contact interculturel et sentiment d'appartenance"),
-    subtitle = "Contrôlé par la région",
-    x = "Contact interculturel en ligne",
-    y = "Sentiment d'appartenance au Québec"
-  ) +
-  theme(text = element_text(hjust = 0.5))
-```
-
-En accord avec les résultats du modèle de régression (Tableau 2), les graphiques
-suivants illustrent la relation entre le contact interculturel et le sentiment
-d'appartenance en fonction de l'âge et du nombre d'années passées au Québec.
-Dans le cas de l'âge, on peut observer que la relation est annulée, sauf parmi
-les répondants âgés de 36 à 50 ans. Dans le cas du nombre d'années passées au
-Québec on peut observer que la relation n'est pas affectée chez les personnes
-établies depuis moins de 6 ans, qu'elle est accentuée chez les personnes
-établies depuis 6 à 9 ans, puis s'inverse chez les personnes établies depuis 10
-ans ou plus.
-
-```{r}
-ggplot(an_SOC3380 %>% filter(!is.na(score_sentimapp)), aes(x= score_rsContact, y = score_sentimapp, color = age))+
-  geom_point() +
-  geom_smooth(method = "lm", se= FALSE)+
-  theme_minimal()+
-  scale_color_brewer(name = "Âge", palette = "Set1") +
-  labs(
-    title = paste("Figure", no_graphic <- no_graphic+1, ":",
-      "Contact interculturel et sentiment d'appartenance"),
-    subtitle = "Contrôlé par l'âge",
-    x = "Contact interculturel en ligne",
-    y = "Sentiment d'appartenance au Québec"
-  ) +
-  theme(text = element_text(hjust = 0.5))
-```
-
-```{r}
-ggplot(an_SOC3380 %>% filter(!is.na(score_sentimapp)), aes(x= score_rsContact, y = score_sentimapp, color = year_qc_f))+
-  geom_point() +
-  geom_smooth(method = "lm", se= FALSE)+
-  theme_minimal()+
-  scale_color_brewer(name = "Nombre d'années passées au Québec", palette = "Set1") +
-  labs(
-    title = paste("Figure", no_graphic <- no_graphic+1, ":",
-      "Contact interculturel et sentiment d'appartenance"),
-    subtitle = "Contrôlé par le nombre d'année passées au Québec",
-    x = "Contact interculturel en ligne",
-    y = "Sentiment d'appartenance au Québec"
-  ) +
-  theme(text = element_text(hjust = 0.5))
-```
-
-### H1.2 - Type de réseaux sociaux en ligne et sentiment d'appartenance
-
-Le modèle de régression suivant présente la relation entre le contact
-interculturel et le sentiment d'appartenance en fonction du type de réseau
-social utilisé. Les résultats suggèrent que l'utilisation de blogs comme
-Twitter, de sites de réseaux sociaux comme Facebook ou encore de sites de
-partage de médias comme YouTube exercent une influence sur la relation étudiée.
-
-```{r}
-local({
-  .Table <- lm(score_rsContact ~ score_sentimapp + rs_blogues + rs_commun_online + rs_SRS + rs_partage_medias, an_SOC3380)
-  pander(.Table, split.cell=80, split.table = Inf, 
-         caption = "Modèle de régression contrôlé pour l'hypothèse H1.2",
-         col.names = c("Estimation", "Erreur type", "Valeur t", "Valeur p"),
-         row.names = c("(Intercept)", "Sentiment d'appartenance", "Blogs", "Communautés en ligne", "Sites de réseaux sociaux", "Sites de partage de médias")
-  )
-})
-```
-
-En accord avec les résultats du modèle de régression, le graphique suivant
-illustre la relation entre le contact interculturel et le sentiment
-d'appartenance en fonction de l'utilisation de blogs en ligne. Les résultats
-indiquent que l'utilisation de blogs comme Twitter modifie le sens de la
-relation étudiée. Autrement dit, l'usage de blogs tendrait à freiner le
-développement d'un sentiment d'appartenance au Québec chez les personnes
-immigrantes.
-
-```{r}
-ggplot(an_SOC3380 %>% filter(!is.na(score_sentimapp)), aes(x= score_rsContact, y = score_sentimapp, color = rs_blogues_f))+
-  geom_point() +
-  geom_smooth(method = "lm", se= FALSE)+
-  theme_minimal()+
-  scale_color_brewer(name = "Utilisation de blogs en ligne", palette = "Set1") +
-  labs(
-    title = paste("Figure", no_graphic <- no_graphic+1, ":",
-      "Contact interculturel et sentiment d'appartenance"),
-    subtitle = "Contrôlé par l'utilisation de blogs en ligne",
-    x = "Score du contact interculturel en ligne",
-    y = "Score du sentiment d'appartenance au Québec"
-  ) +
-  theme(text = element_text(hjust = 0.5))
-```
-
-En accord avec les résultats du modèle de régression, le graphique suivant
-illustre la relation entre le contact interculturel et le sentiment
-d'appartenance en fonction de l'utilisation de sites de réseaux sociaux en
-ligne. Les résultats suggèrent le phénomène inverse. En d'autres termes,
-l'utilisation de plateformes comme Facebook faciliterait le développement d'un
-sentiment d'appartenance au Québec chez les personnes immigrantes. Néanmoins,
-nos résultats ne peuvent être validés, étant entendu que la majorité de nos
-répondants utilisent des sites de réseaux sociaux en ligne.
-
-Néanmoins, on peut observer sur le graphique que la relation étudiée est plus
-intense lorsqu'elle est contrôlée par l'utilisation des sites de réseaux sociaux
-en ligne.
-
-```{r}
-ggplot(an_SOC3380 %>% filter(!is.na(score_sentimapp)), aes(x= score_rsContact, y = score_sentimapp, color = rs_SRS_f))+
-  geom_point() +
-  geom_smooth(method = "lm", se= FALSE)+
-  theme_minimal()+
-  scale_color_brewer(name = "Utilisation de sites de réseaux sociaux en ligne", palette = "Set1") +
-  labs(
-    title = paste("Figure", no_graphic <- no_graphic+1, ":",
-      "Contact interculturel et sentiment d'appartenance"),
-    subtitle = "Contrôlé par le type d'utilisation de sites de réseaux sociaux en ligne",
-    x = "Score du contact interculturel en ligne",
-    y = "Score du sentiment d'appartenance au Québec") +
-  theme(text = element_text(hjust = 0.5))
-```
-
-En accord avec les résultats du modèle de régression, le graphique suivant
-illustre la relation entre le contact interculturel et le sentiment
-d'appartenance en fonction de l'utilisation de sites de partage de médias en
-ligne. Les résultats indiquent que l'utilisation de sites de partage de médias
-atténue l'intensité de la relation étudiée. En d'autres termes, à l'instar des
-blogs, l'utilisation de plateformes comme YouTube tend à freiner le
-développement d'un sentiment d'appartenance au Québec.
-
-```{r}
-ggplot(an_SOC3380 %>% filter(!is.na(score_sentimapp)), aes(x= score_rsContact, y = score_sentimapp, color = rs_partage_medias_f))+
-  geom_point() +
-  geom_smooth(method = "lm", se= FALSE)+
-  theme_minimal()+
-  scale_color_brewer(name = "Utilisation de sites de partage de médias", palette = "Set1") +
-  labs(
-    title = paste("Figure", no_graphic <- no_graphic+1, ":",
-      "Contact interculturel et sentiment d'appartenance"),
-    subtitle = "Contrôlé par l'utilisation de site de partage de médias",
-    x = "Score du contact interculturel en ligne",
-    y = "Score du sentiment d'appartenance au Québec"
-  ) +
-  theme(text = element_text(hjust = 0.5))
-```
-
-### H1.3 - Type d'utilisation des réseaux et sentiment d'appartenance
-
-Le modèle de régression suivant présente la relation entre le contact
-interculturel et le sentiment d'appartenance en fonction de du type
-d'utilisation des réseaux sociaux en ligne. Pour parvenir à ses résultats, nous
-avons construit un indice supplémentaire permettant de mesurer la proportion des
-réseaux sociaux traditionnellement utilisés pour communiquer avec la famille et
-les amis restés au pays.
-
-```{r}
-local({
-  .Table <- lm(score_rsContact ~ score_sentimapp + score_media_accueil + score_media_diasp, an_SOC3380)
-  pander(.Table, split.cell=80, split.table = Inf, 
-         caption = "Modèle de régression contrôlé pour l'hypothèse H1.3",
-         col.names = c("Estimation", "Erreur type", "Valeur t", "Valeur p"),
-         row.names = c("(Intercept)", "Sentiment d'appartenance", "Utilisation au Québec", "Utilisation mixte")
-  )
-})
-```
-
-Les résultats suggèrent que le type d'utilisation des réseaux sociaux joue un
-rôle dans la relation étudiée. Le graphique suivant illustre la relation entre
-le contact interculturel et le sentiment d'appartenance en fonction du type
-d'utilisation (mixte ou exclusivement au Québec). On observe que l'utilisation
-exclusive de réseaux sociaux permettant de communiquer au Québec modifie le sens
-de la relation étudiée.
-
-## H2 - Exposition à du contenu lié à l'immigration et sentiment d'appartenance
-
-```{r}
-ggplot(an_SOC3380 %>% filter(!is.na(score_sentimapp)), aes(x= score_rsContact, y = score_sentimapp, color = rs_diasp_f))+
-  geom_point() +
-  geom_smooth(method = "lm", se= FALSE)+
-  theme_minimal()+
-  scale_color_brewer(name = "Utilisation mixte", palette = "Set1") +
-  labs(
-    title = paste("Figure", no_graphic <- no_graphic+1, ":",
-      "Contact interculturel et sentiment d'appartenance"),
-    subtitle = "Contrôlé par l'utilisation mixte",
-    x = "Score du contact interculturel en ligne",
-    y = "Score du sentiment d'appartenance au Québec"
-  ) +
-  theme(text = element_text(hjust = 0.5))
-```
-
-Le modèle de régression suivant présente la relation entre l'exposition à du
-contenu lié à l'immigration et le sentiment d'appartenance en fonction du nombre
-d'années passées au Québec, de la région d'origine et de l'âge. Les résultats
-indiquent que la relation est "pratiquement" significative dans la population à
-l'intervalle de confiance 95% (p \~ 0.05).
-
-```{r}
-local({
-  .Table <- lm(score_ExpoDisc ~ score_sentimapp + year_qc_n + region_n + age_n, an_SOC3380)
-  pander(.Table, split.cell=80, split.table = Inf, 
-         caption = "Modèle de régression contrôlé pour l'hypothèse H2",
-         col.names = c("Estimation", "Erreur type", "Valeur t", "Valeur p"),
-         row.names = c("(Intercept)", "Sentiment d'appartenance", "Année passés au Québec", "Région d'origine", "Âge")
-  )
-})
-```
-
-En accord avec les résultats du modèle de régression, le test de corrélation de
-Pearson suivant présente l'intensité et le sens de la relation entre
-l'exposition à du contenu lié à l'immigration et le sentiment d'appartenance.
-Les résultats obtenus indiquent que la relation faible et positive (C = 0.157).
-
-```{r}
-local({
-  .Table <- cor.test(an_SOC3380$score_ExpoDisc, an_SOC3380$score_sentimapp)
-  pander(.Table, split.cell=80, split.table = Inf, 
-         caption = "Test de corrélation de Pearson pour l'hypothèse H2",
-         col.names = c("Valeur t", "Degrés de liberté", "Valeur C", "Hypothèse alternatif", "Valeur C")
-  )
-})
-```
-
-Le graphique suivant illustre la relation entre l'exposition à du contenu lié à
-l'immigration et le sentiment d'appartenance. En accord avec les résultats du
-test de corrélation de Pearson, on observe que la relation est positive et
-faible. En d'autres termes : pour les personnes immigrantes établies au Québec,
-l'exposition à du contenu lié à l'immigration sur les réseaux sociaux en ligne
-est faiblement et positivement corrélé au développement d'un sentiment
-d'appartenance au pays d'accueil.
-
-```{r}
-ggplot(an_SOC3380 %>% filter(!is.na(score_ExpoDisc), !is.na(score_sentimapp)), aes(x= score_ExpoDisc, y = score_sentimapp))+
- geom_point() +
- geom_smooth(method = "lm", se= FALSE)+
- labs(
-   title = paste("Figure", no_graphic <- no_graphic+1, ":",
-      "Exposition aux discours liés à l'immigration et sentiment d'appartenance"),
-   x = "Exposition aux discours liés à l'immigration",
-   y = "Sentiment d'appartenance au Québec"
-  ) +
-  theme(text = element_text(hjust = 0.5))
-```
-
-En accord avec les résultats du modèle de régression, les deux graphiques
-suivants illustrent les effets de l'âge et du nombre d'années passées au Québec
-sur la relation étudiée. Les résultats indiquent que le phénomène est plus
-prononcé chez les personnes âgées de 36 ans et plus et chez les personnes
-établies au Québec depuis plus de 5 ans.
-
-```{r}
-ggplot(an_SOC3380 %>% filter(!is.na(score_sentimapp), !is.na(score_ExpoDisc)), aes(x= score_ExpoDisc, y = score_sentimapp, color = age))+
-  geom_point() +
-  geom_smooth(method = "lm", se= FALSE)+
-  theme_minimal()+
-  scale_color_brewer(name = "Âge", palette = "Set1") +
-  labs(
-    title = paste("Figure", no_graphic <- no_graphic+1, ":",
-      "Exposition aux discours liés à l'immigration et sentiment d'appartenance"),
-    subtitle = "Contrôlé par l'âge",
-    x = "Exposition aux discours liés à l'immigration", y = "Sentiment d'appartenance au Québec"
-  ) +
-  theme(text = element_text(hjust = 0.5))
-```
-
-```{r}
-ggplot(an_SOC3380 %>% filter(!is.na(score_sentimapp), !is.na(score_ExpoDisc)), aes(x= score_ExpoDisc, y = score_sentimapp, color = year_qc_f))+
-  geom_point() +
-  geom_smooth(method = "lm", se= FALSE)+
-  theme_minimal()+
-  scale_color_brewer(name = "Années passées au Québec", palette = "Set1") +
-  labs(
-    title = paste("Figure", no_graphic <- no_graphic+1, ":",
-      "Exposition aux discours liés à l'immigration et sentiment d'appartenance"),
-    subtitle = "Contrôlé par le nombre d'années passées au Québec",
-    x = "Exposition aux discours liés à l'immigration",
-    y = "Sentiment d'appartenance au Québec"
-  ) +
-  theme(text = element_text(hjust = 0.5))
-```
-
-\newpage
-
-<!--chapter:end:05-resultats.Rmd-->
-
----
-knit: bookdown::render_book
-editor_options: 
-  markdown: 
-    wrap: 80
----
-
-# Discussion
-
-## H1 - Contact interculturel et sentiment d'appartenance
-
-Le volet qualitatif de notre enquête vient confirmer l'existence de la
-corrélation entre le contact interculturel et le sentiment d'appartenance tout
-en fournissant des clés d'explication suggérant l'existence d'une relation
-causale entre les deux variables. Si le travail de Gallant et Friche
-[-@gallantEtreIciLabas2010] rappellent que les réseaux sociaux en ligne
-participent peu à la création de réseaux d'appartenance préexistants, les
-témoignages que nous avons recueillis mentionnent que certaines plateformes
-comme Facebook, particulièrement actives au Québec, permettent aux personnes
-immigrantes de pérenniser des liens établis hors ligne avec les Québécois.
-
-> Nina: Les évènements de Facebook sont un peu morts en France(...) alors qu'ici
-> justement, les gens sont beaucoup plus actifs, surtout dans le Marketplace
-> (...) J'ai changé ma vision de Facebook, parce que son utilisation est très
-> différente ici.
-
-## H1.1 - Pays d'origine, discrimination systémique et sentiment d'appartenance
-
-Le phénomène observé dans le graphique représentant la relation entre le contact
-interculturel et le sentiment d'appartenance contrôlé par la région d'origine
-pourrait illustrer le développement de discours islamophobes au Québec, comme le
-suggèrent les travaux de Dalila Awada [-@awadaRacialisationMusulmanEs2020] sur
-la racialisation des personnes musulmanes dans la province.
-
-## H1.2 - Type de réseaux sociaux en ligne et sentiment d'appartenance
-
-Les résultats obtenus concernant la relation entre contact interculturel et
-sentiment d'appartenance en fonction du type de réseau social utilisé
-[@kietzmannSocialMediaGet2011] pourraient confirmer les travaux de Hopster et
-Ribeiro [-@hopsterMutualAffordancesDynamics2021, p. 559;
--@ribeiroAuditingRadicalizationPathways2020, p. 140] concernant le rôle des
-algorithmes de suggestion dans la diffusion de discours xénophobes sur YouTube.
-Ces résultats font également échos aux travaux de Dehgan et Walsh
-[-@dehghanDynamicsPolarisationAustralian2022, p.64‑65;
--@walshDigitalNativismTwitter2021, p.5] sur la racialisation discursive et le
-développement du néonativisme sur Twitter.
-
-Selon Esser [@esserDoesNewImmigration2004], p. 1131], le sentiment
-d'appartenance constitue l'une des quatre dimensions de l'intégration
-culturelle. Or, les résultats de notre analyse thématique révèlent que nos
-répondants mobilisent les réseaux sociaux pour accumuler du capital culturel et
-social valorisable dans le cadre de contacts interculturels en ligne et hors
-ligne.
-\newpage
-
-> Charly: Au niveau de l'intégration à la société d'accueil, ça a été un pilier
-> fondamental (...) avoir accès à des memes, connaître des célébrités, mettons
-> Occupation Double... des trucs comme ça, qui te permettent d'avoir des
-> conversations [avec des Québécois].
-
-Cette association est cependant pondérée par les résultats de notre analyse
-thématique. En effet, certains répondants témoignent du rôle positif joué par
-YouTube dans le cadre de leur intégration culturelle. Adriana (une jeune femme
-originaire d'Amérique du Sud) avance que le contenu présent sur la plateforme a
-participé à son apprentissage du français québécois et de l'histoire de la
-province :
-
-> Adriana: [Sur YouTube] je cherchais les sous-titres en français avec un petit
-> dictionnaire à côté de mon cahier (...) J'essayais de comprendre en arrêtant
-> de lire [les sous-titres] et faire des comparaisons avec les apostrophes,
-> essayer de comprendre la dynamique du parler le français québécois (...) Mais
-> quand je suis sortie dans la rue, je n'ai rien compris (...) J'ai beaucoup
-> appris de l'histoire québécoise grâce à une chaîne [YouTube] en particulier
-> (...) comprendre la fondation du Québec, la relation que le pays entretien
-> avec les autochtones.
-
-## H1.3 - Type d'utilisation des réseaux et sentiment d'appartenance
-
-L'écart constaté pour la relation entre le contact interculturel et le sentiment
-d'appartenance en fonction du type d'utilisation des réseaux sociaux pourrait
-s'expliquer en creux par la capacité des réseaux sociaux en ligne à réduire le
-stress acculturatif généré par la migration. En gardant un contact affectif et
-culturel avec leur pays d'origine, les personnes immigrantes seraient plus
-enclines à s'intégrer culturellement au Québec, et partant, à développer un
-sentiment d'appartenance à leur pays d'accueil.
-
-Cette hypothèse spécifique peut être mise en relation avec les résultats de
-l'analyse thématique, qui semblent confirmer le pouvoir stabilisateur du lien
-maintenu par nos répondants avec leur pays d'origine sur les réseaux sociaux en
-ligne. En effet, Charly (un jeune homme originaire d'Amérique du Sud) et Nina
-(une jeune femme originaire de France) affirment que les sites de réseaux
-sociaux comme WhatsApp et Messenger leur permettent de garder un lien vital avec
-la famille et les amis restés au pays :
-
-> Charly: Si j'avais pas pu parlé à ma mère tous les jours, c'est sûr
-> qu'éventuellement, j'aurais pété un plomb, puis j'aurais fait « ciao ». (...)
-> Je dirais que c'est fondamental, je pense, dans mes relations sociales. Quand
-> je dis « fondamental », c'est que je n'imagine pas les choses autrement, pour
-> que ce soit comme c'est aujourd'hui. Peut-être que je serais moins intégré si
-> je n'avais pas eu accès [aux] réseaux sociaux, si je n'avais pas pu, par
-> exemple, à un moment donné, pu entretenir des relations qui étaient très
-> importantes pour moi (...) Ça m'a permis de garder le lien pendant ces 11 ans,
-> toujours avec les mêmes personnes. Ma famille on est toujours en contact, on
-> s'envoi des photos au quotidien, des trucs comme ça (...) malgré la
-> distance,on entretient quand même cette relation.
->
-> Nina: [exclamation] Ah ! les groupes familiaux, c'est quelque chose de
-> pratique, on n'a pas besoin d'interagir pour voir comment les gens vont. Les
-> gens enrichissent les groupes de ce qui leur arrive. Donc en fait, je n'ai
-> même pas besoin de poser des questions, j'arrive à suivre ce qui se passe dans
-> le groupe.
->
-> Nina: [exclamation] Ah ! les groupes familiaux, c'est quelque chose de
-> pratique, on n'a pas besoin d'interagir pour voir comment les gens vont. Les
-> gens enrichissent les groupes de ce qui leur arrive. Donc en fait, je n'ai
-> même pas besoin de poser des questions, j'arrive à suivre ce qui se passe dans
-> le groupe.
-
-Néanmoins, la causalité (même indirecte) du rôle joué par le type d'utilisation
-des réseaux sociaux dans la relation étudiée ne peut être confirmée au vu du
-faible nombre de répondants utilisant les réseaux sociaux en ligne uniquement
-pour communiquer au Québec.
-
-De plus, ces résultats entrent en contradiction avec les travaux de Dilshad
-Dayani [-@dayaniRoleSocialMedia2017], qui suggèrent que l'utilisation des
-réseaux sociaux pour garder contact avec le pays d'origine tendent à freiner le
-développement d'un sentiment d'appartenance à la société d'accueil chez les
-personnes immigrantes. Dayani avançait que l'utilisation de Facebook affecte
-positivement le processus d'acculturation, alors que nos résultats suggèrent
-l'effet inverse.
-
-## H2 - Exposition à du contenu lié à l'immigration et sentiment d'appartenance
-
-Malgré sa faible significativité statistique, la relation entre l'exposition à
-du contenu lié à l'immigration en ligne et le sentiment d'appartenance peut être
-partiellement expliquée par les résultats de l'analyse thématique. D'un côté, le
-discours lié à l'immigration au Québec présent sur les réseaux sociaux en ligne
-est ambivalent, comme le démontre Wendy :
-
-> Théa: Que peux-tu me dire des publications liées à l'immigration que tu vois
-> sur les réseaux ?
->
-> Wendy: Je vois deux choses : d'un côté il y en a une majorité qui disent que
-> c'est bien de s'installer au Canada ; de l'autre, il y a des
-> généralisations\... comme quoi certaines populations ne font rien, ne
-> travaillent pas ou bien s'approprient des emplois qui devraient revenir aux
-> Québécois.
-
-De l'autre côté, les personnes immigrantes semblent prioriser l'utilisation des
-réseaux sociaux en ligne pour collecter des informations concernant leur pays
-d'accueil, comme nous le montrent Mathilde (une jeunes femmes originaires de
-France).
-
-> Antoine: Les nouvelles sur l'immigration plus négatives, tu les entends en
-> dehors des réseaux ?
->
-> Mathilde: Ouais, quand tu vois les nouvelles dans le métro, dans la rue ou à
-> la radio\... des choses comme ça. C'est des choses qui me parviennent sans que
-> j'aille les chercher (\...) J'ai l'impression que tu n'as pas besoin de
-> chercher de l'information au sujet de l'immigration pour être tenue au courant
-> ici.
+<!--chapter:end:05-results.Rmd-->
+
+---
+knit: bookdown::render_book
+editor_options: 
+  markdown: 
+    wrap: 80
+---
 
@@
 
-Cette étude cherchait à comprendre la relation entre l'utilisation des réseaux
-sociaux en ligne et l'intégration culturelle des personnes immigrantes établies
-au Québec depuis plus d'un an. Les résultats du volet quantitatif de notre
-enquête indiquent que le contact interculturel sur les réseaux sociaux en ligne
-est faiblement et positivement corrélé au développement d'un sentiment
-d'appartenance au Québec chez nos répondants. Dans le même temps, l'exposition à
-du contenu lié à l'immigration en ligne et le développement d'un sentiment
-d'appartenance au Québec sont également faiblement et positivement corrélés. En
-outre, le volet qualitatif de notre enquête vient pondérer et spécifier ces deux
-conclusions au regard des caractéristiques socio-démographiques de nos
-répondants, notamment à l'endroit des discriminations anti-immigrants, qu'elles
-soient de nature nationales, ethniques ou raciales.
-
-À la lumière de nos résultats, une étude focalisée autour des effets de la
-discrimination envers les personnes immigrantes sur le développement d'un
-sentiment d'appartenance au Québec pourrait éclairer avec précision les
-dynamiques de racialisation en cours sur les réseaux sociaux en ligne. Pour ce
-faire, la création d'un indice de la discrimination et/ou de la catégorisation
-(qu'elle soit nationale, ethnique ou raciale) permettrait de mesurer directement
-la relation que celle-ci entretien avec le développement d'un sentiment
-d'appartenance au Québec.
-
-\newpage
-
-<!--chapter:end:06-discussions.Rmd-->
-
----
-knit: bookdown::render_book
-editor_options: 
-  markdown: 
-    wrap: 80
----
-# Références
-
-<!--chapter:end:07-références.Rmd-->
+The preliminary results of this study show that the consequences of the pandemic
+on the Quebec population differ according to the population groups, but also
+according to the type of consequences. \newpage
+
+<!--chapter:end:06-conclusion.Rmd-->
+
+---
+knit: bookdown::render_book
+editor_options: 
+  markdown: 
+    wrap: 80
+---
+
+# References
+
+<http://www.oecd.org/coronavirus/policy-responses/what-is-the-impact-of-the-covid-19-pandemic-on-immigrants-and-their-children-e7cbb7de/>
+
+Guttmann A, F. et al. (2020), COVID-19 in Immigrants, Refugees and Other
+Newcomers in Ontario: Characteristics of Those Tested and Those Confirmed
+Positive, as of June 13, 2020, ICES, <http://www.ices.on.ca>.
+
+Hu, Y. (2020), "Intersecting ethnic and native--migrant inequalities in the
+economic impact of the COVID-19 pandemic in the UK", Research in Social
+Stratification and Mobility, Vol. 68, p. 100528,
+<http://dx.doi.org/10.1016/j.rssm.2020.100528>.
+
+Mcginnity, F. and G. Kingston (2017), "An Irish Welcome? Changing Irish
+Attitudes to Immigrants and Immigration: The Role of Recession and Immigration",
+The Economic and social review, Vol. 48/3, pp. 281-304,
+<https://www.researchgate.net/publication/319670153_An_Irish_Welcome_Changing_Irish_Attitudes_to_Immigrants_and_Immigration_The_Role_of_Recession_and_Immigration>
+(accessed on 2 October 2020).
+
+Raisi-Estabragh, Z. et al. (2020), "Greater risk of severe COVID-19 in Black,
+Asian and Minority Ethnic populations is not explained by cardiometabolic,
+socioeconomic or behavioural factors, or by 25(OH)-vitamin D status: study of
+1326 cases from the UK Biobank", Journal of Public Health, Vol. 42/3, pp.
+451-460, <http://dx.doi.org/10.1093/pubmed/fdaa095>.
+
+Apea, V. et al. (2020), Ethnicity and outcomes in patients hospitalised with
+COVID-19 infection in East London: an observational cohort study, Cold Spring
+Harbor Laboratory, <http://dx.doi.org/10.1101/2020.06.10.20127621>.
+
+Bertocchi, G. and A. Dimico (2020), "Covid-19, Race, and Redlining", IZA
+Discussion Paper, Vol. 13467,
+<https://papers.ssrn.com/sol3/papers.cfm?abstract_id=3648807> (accessed on 22
+September 2020).
+
+Brun, S. and P. Simon (2020), "Inégalités ethno-raciales et coronavirus", De
+Facto 19.
+
+<!--chapter:end:07-references.Rmd-->

### 2023-05-02 Expérimentations (Théa)/2. Expérimentations (Bookdown)/Bookdown/index.Rmd
Multiple changes in Bookdown

@@
 subtitle: A Comparative Analysis of Exposure Risks, Consequences, and Access to Resources among Immigrants and Natives
-groupe_cours: "SOC3380, gr. 40"
 site: bookdown::bookdown_site
@@
 
-```{r setup, include=FALSE, warning=FALSE}
-# Préférence knitr
-
-knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo=FALSE)
-```
-
 ```{r}

### 2023-05-02 Expérimentations (Théa)/2. Expérimentations (Bookdown)/Bookdown/soc3380-immigr-rapport-de-recherche.tex
Multiple changes in Bookdown

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\usepackage{booktabs}
\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhf{}
\renewcommand{\headrulewidth}{0pt}
\fancyfoot[R]{\thepage}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={COVID-19 Disparities in Quebec (Canada)},
  pdfauthor={V. Adjiwanou, G. NGalé, S. Lardoux, N. Zahedinameghi},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

%%%%%%%%%%%%%%%%
% FRANCISATION DES TERMES
%%%%%%%%%%%%%%%%

\renewcommand\contentsname{Table des mati\`eres}
\renewcommand\listfigurename{Liste des figure}
\renewcommand\listtablename{Liste des tableaux}
\renewcommand\indexname{Index}
\renewcommand\figurename{Figure}
\renewcommand\tablename{Tableau}
\renewcommand\partname{Partie}
\renewcommand\appendixname{Annexe}  % annexe
\renewcommand\abstractname{R\'esum\'e}

%%%%%%%%%%%%%%%%
% PAGE TITRE
%%%%%%%%%%%%%%%%

\def\maketitle{\begin{titlepage}
\setmainfont{Liberation Sans} 
{\raggedleft\sffamily\color{black}
Vincent Bertrand-Huneaut
\par}

{\raggedleft\sffamily\color{black}
Jeanne Craig-Larouche
\par}

{\raggedleft\sffamily\color{black}
Théa Carlier
\par}

{\raggedleft\sffamily\color{black}
Noura Kone
\par}

{\raggedleft\sffamily\color{black}
\ \ Antoine Martin
\par}

\bigskip

\bigskip

\bigskip

\bigskip

\bigskip

{\centering\sffamily\color{black}
Article
\par}

{\centering\sffamily\color{black}
COVID-19 Disparities in Quebec (Canada)
\par}

\bigskip

\bigskip

{\centering\sffamily\color{black}
Présenté à :
\par}

{\centering\sffamily\color{black}
Vissého Adjiwanou
\par}

\bigskip

\bigskip

\bigskip

\bigskip

\bigskip

\bigskip

\bigskip

{\centering\sffamily\color{black}

\par}

\bigskip

\bigskip

\bigskip

\bigskip

\bigskip

\bigskip

\bigskip

\bigskip

\bigskip

{\centering\sffamily\color{black}
mai 2, 2023
\par}

\bigskip

\bigskip

\bigskip

\bigskip

\bigskip

{\centering\sffamily\color{black}
Université du Québec À Montréal and Université De Montréal
\par}
\end{titlepage}
}

\begin{document}
\maketitle
\newpage
\begin{abstract}
The immigrant population is one of the most vulnerable in Canada, with a relatively high risk of unemployment, a low social support system, and linguistic barriers. However, their experiences with public health interventions during the COVID-19 pandemic have not been adequately explored. The purpose of this article is to analyze Covid-19 exposure risk and its family implications for both natives and immigrants in the province of Quebec. Additionally, we analyze immigrants' access to government aid and compare it to their native-born counterparts. The study consists of survey data collected from 1,000 immigrants and 500 non-immigrants Quebecers between August and November 2020. Our descriptive and exploratory analysis reveals significant differences between the two groups regarding exposure risk, family consequences, and access to resources. We conclude that the current pandemic exacerbates health disparities in the province, with particularly dire repercussions for immigrants' families. We also discuss several socio-cultural strategies to mitigate inequality and recommend policies to address the needs of vulnerable populations.
\end{abstract}

\newpage
{
\setcounter{tocdepth}{2}
\tableofcontents\
\newpage
}
\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# Importer package}
\FunctionTok{library}\NormalTok{(dplyr)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Warning: le package 'dplyr' a été compilé avec la version R 4.2.3
\end{verbatim}

\begin{verbatim}
## 
## Attachement du package : 'dplyr'
\end{verbatim}

\begin{verbatim}
## Les objets suivants sont masqués depuis 'package:stats':
## 
##     filter, lag
\end{verbatim}

\begin{verbatim}
## Les objets suivants sont masqués depuis 'package:base':
## 
##     intersect, setdiff, setequal, union
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(tidyverse)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Warning: le package 'tidyverse' a été compilé avec la version R 4.2.3
\end{verbatim}

\begin{verbatim}
## Warning: le package 'ggplot2' a été compilé avec la version R 4.2.3
\end{verbatim}

\begin{verbatim}
## Warning: le package 'tibble' a été compilé avec la version R 4.2.3
\end{verbatim}

\begin{verbatim}
## Warning: le package 'tidyr' a été compilé avec la version R 4.2.3
\end{verbatim}

\begin{verbatim}
## Warning: le package 'readr' a été compilé avec la version R 4.2.3
\end{verbatim}

\begin{verbatim}
## Warning: le package 'forcats' a été compilé avec la version R 4.2.3
\end{verbatim}

\begin{verbatim}
## Warning: le package 'lubridate' a été compilé avec la version R 4.2.3
\end{verbatim}

\begin{verbatim}
## -- Attaching core tidyverse packages ------------------------ tidyverse 2.0.0 --
## v forcats   1.0.0     v readr     2.1.4
## v ggplot2   3.4.2     v stringr   1.5.0
## v lubridate 1.9.2     v tibble    3.2.1
## v purrr     1.0.1     v tidyr     1.3.0
\end{verbatim}

\begin{verbatim}
## -- Conflicts ------------------------------------------ tidyverse_conflicts() --
## x dplyr::filter() masks stats::filter()
## x dplyr::lag()    masks stats::lag()
## i Use the ]8;;http://conflicted.r-lib.org/conflicted package]8;; to force all conflicts to become errors
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(summarytools)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Warning: le package 'summarytools' a été compilé avec la version R 4.2.3
\end{verbatim}

\begin{verbatim}
## 
## Attachement du package : 'summarytools'
## 
## L'objet suivant est masqué depuis 'package:tibble':
## 
##     view
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(readr)}
\FunctionTok{library}\NormalTok{(questionr)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## 
## Attachement du package : 'questionr'
## 
## L'objet suivant est masqué depuis 'package:summarytools':
## 
##     freq
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(rcompanion)}
\FunctionTok{library}\NormalTok{(pander)}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# Set numero graphique}
\NormalTok{no\_graphic }\OtherTok{\textless{}{-}} \SpecialCharTok{{-}}\DecValTok{1}
\end{Highlighting}
\end{Shaded}

\hypertarget{introduction-background}{%
\section{Introduction / background}\label{introduction-background}}

The COVID-19 pandemic has highlighted the vast ethnic, social, economic, and
health disparities in high-income countries, including Canada. One of the most
vulnerable segments of these affluent societies is the immigrant population,
that has become widely diversified with more and more people coming from
developing countries, and for whom integration into their new society may be
more difficult. The coronavirus pandemic brings a new challenge for these
immigrant populations, directly or indirectly through the tensions between the
ethnic groups that it helps to generate (\protect\hyperlink{ref-coates2020Covid19}{\textbf{coates2020Covid19?}}). These immigrants
face language barriers, low social capital, housing insecurity, and severe
poverty, rendering them particularly susceptible to the effects of the current
pandemic. Immigrants are also more likely to be employed as essential workers
where social distancing, sanitary precautions, and self-isolation, measures put
in place to mitigate the pandemy, are challenging (\protect\hyperlink{ref-reid2021Migrant}{\textbf{reid2021Migrant?}}~; \protect\hyperlink{ref-yi2021Health}{\textbf{yi2021Health?}}~; \protect\hyperlink{ref-zhang2021COVID19}{\textbf{zhang2021COVID19?}}). More importantly, migrants may encounter
various forms of discrimination, leading to social exclusion and restricted
access to healthcare and social services (\protect\hyperlink{ref-spiritus-beerden2021Mental}{\textbf{spiritus-beerden2021Mental?}}). It is
therefore necessary to understand how this pandemic affects these immigrant
populations differently and the native populations in Quebec. In doing so, this
article will reveal a better understanding of how the current pandemic
aggravates existing inequalities and challenges in the province for immigrants
and their descendants.

Since the start of the pandemic, most of the confirmed cases were identified in
Montreal and areas with high immigrant populations (\protect\hyperlink{ref-miconi2021Ethnocultural}{\textbf{miconi2021Ethnocultural?}}~; \protect\hyperlink{ref-shields2020COVID19}{\textbf{shields2020COVID19?}}). The coronavirus pandemic has infected nearly 450,000
people in Quebec and has resulted in more than 11,000 deaths
(\url{https://www.quebec.ca/sante/problemes-de-sante/a-z/coronavirus-2019/situation-coronavirus-quebec},
consulted on octobre 29, 2021). This makes it the worst affected province in
Canada and one of the worst affected regions in the world during the first wave
of the pandemic. Although the majority of these deaths took place in elderly's
homes, and therefore affected more this segment of the population, its overall
influence on individuals is not to be overlooked. It has been shown that, on the
one hand, immigrants and their families have a higher risk of being exposed to
the pathogen, and, at the same time, they are more vulnerable to racism and
discrimination and face barriers to accessing social support (\protect\hyperlink{ref-2021CRARR}{\textbf{2021CRARR?}}) and
information, especially for those who didn't speak or understand the two
official langages.
Several
studies have thus shown excess mortality in ethnocultural minority populations
(\protect\hyperlink{ref-bertocchi2020Covid19}{\textbf{bertocchi2020Covid19?}}), higher unemployment, greater loss of income and
negative consequences for children and families in several countries. This
differentiated effect can be explained in different ways. On the one hand, in
terms of the direct consequences of the pandemic, the situation of
precariousness and greater poverty among immigrants exposes them more to the
more harmful consequences of the disease on their state of health. Likewise,
immigrants often work in jobs where social distancing is difficult to enforce.
On the other hand, indirectly, the solutions to the crisis affect immigrants and
natives differently. The complete shutdown of tye economy with its set of
adjustments has greatly impeded the employability of immigrants, especially
newcomers who have not yet grasped the workings of society and who cannot
benefit from state subsidies. But, it is also shown that these effects are not
uniform on the whole of the immigrant population, itself, very diverse.

Despite the growing focus on the differential impacts of the COVID-19 on
immigrants and non-immigrants segment of the population, there is still very
little knowledge about how the pandemic and its related public policy impact the
lives of immigrants and their families. The existing health datasets (largely
used to assess this effect) are not necessarily representative of the immigrant
population and do not offer an accurate picture of the complex social status and
health outcomes among this group (\protect\hyperlink{ref-choi2021Studying}{\textbf{choi2021Studying?}}~; \protect\hyperlink{ref-gagnon2021Immigration}{\textbf{gagnon2021Immigration?}}~; \protect\hyperlink{ref-khan2015Visible}{\textbf{khan2015Visible?}}). In most cases, these datasets lack data on the residence
length of immigrants and combine newcomers with long-term settlers
(\protect\hyperlink{ref-edge2013Discrimination}{\textbf{edge2013Discrimination?}}~; \protect\hyperlink{ref-hayward2021Clinical}{\textbf{hayward2021Clinical?}}). As a result, they may not allow
a detailed assessment of the nature and extent of risk exposure in sub-segments
of populations, as well as whether interventions are reaching those most at
risk. In addition, the family-level impacts of the pandemic are still poorly
understood. While a body of research has explored the association between the
pandemic and its potential effects on intimate partner violence and marital
dissolution (\protect\hyperlink{ref-bhatnagar2021Navigating}{\textbf{bhatnagar2021Navigating?}}~; \protect\hyperlink{ref-koshan2020COVID19}{\textbf{koshan2020COVID19?}}~; \protect\hyperlink{ref-morrison2021Chapter}{\textbf{morrison2021Chapter?}}~; \protect\hyperlink{ref-prime2020Risk}{\textbf{prime2020Risk?}}), few studies have examined negative familial repercussions in
the context of immigration status. There is ample evidence that sustained family
relationships greatly enhance immigrants' lives on several fronts, including
fostering economic prosperity (\protect\hyperlink{ref-marger2001Sociala}{\textbf{marger2001Sociala?}}), psychological well-being
(\protect\hyperlink{ref-beiser2002Povertya}{\textbf{beiser2002Povertya?}}), social network development (\protect\hyperlink{ref-evansluong2017Familya}{\textbf{evansluong2017Familya?}}), and
educational success for children (\protect\hyperlink{ref-munroe-blum1989Immigranta}{\textbf{munroe-blum1989Immigranta?}}). The family is,
therefore, a key component of long-term social mobility and economic success of
immigrants.

Based on an original biographic survey collected around the first wave of the
pandemic in august-november 2020 on immigrants and the native population of
Québec, this study mainly aims to understand the family exposure, consequences,
and responses of the pandemic on immigrant and native Quebec populations. We
assess a whole range of different measures of each outcome. We divided the
consequences in three dimensions, 1) access to ressources, 2) ability to perform
duties and 3) employment and income. We seek also to analyse in what ways these
outcomes differ according to the immigrants length of residence. By doing so,
this study provide a clear and global picture of the situation of different
population groups in Quebec in the face of the COVID-19.

\newpage

\hypertarget{current-body-of-literature-and-knowledge-gaps}{%
\section{Current body of literature and knowledge gaps}\label{current-body-of-literature-and-knowledge-gaps}}

The literature on the racial and ethnic health disparities during the COVID-19
crisis can be divided into two main categories: studies addressing the
socio-economic determinants of health and those examining the delivery of health
interventions and policies. The first category is primarily comprised of
research that focuses on the link between socio-economic status (SES) of
high-risk minority groups and the differential impact of the pandemic
(\protect\hyperlink{ref-drefahl2020Populationbased}{\textbf{drefahl2020Populationbased?}}~; \protect\hyperlink{ref-greenaway2020COVID19}{\textbf{greenaway2020COVID19?}}~; \protect\hyperlink{ref-lopez2021Racial}{\textbf{lopez2021Racial?}}~; \protect\hyperlink{ref-pacheco2020Job}{\textbf{pacheco2020Job?}}~; \protect\hyperlink{ref-sundaram2020Individual}{\textbf{sundaram2020Individual?}}). For instance, an inter-country study
has found that Middle Eastern, African, and Nordic immigrants in Sweden have a
higher risk of being hospitalized and even dying from COVID-19 when compared to
the Swedish born population. In the same study, the SES, proportion of
working-age households, and neighborhood density were significantly associated
with the mortality rate (\protect\hyperlink{ref-rostila2021Disparities}{\textbf{rostila2021Disparities?}}). This trend of research
demonstrates that racial and ethnic minorities are more likely to experience
stress factors, not only due to the hazards of infection but also because of
economic and social marginalization (\protect\hyperlink{ref-alahmad2020COVID19}{\textbf{alahmad2020COVID19?}}~; \protect\hyperlink{ref-choi2020Buffering}{\textbf{choi2020Buffering?}}~; \protect\hyperlink{ref-endale2020COVID19}{\textbf{endale2020COVID19?}}~; \protect\hyperlink{ref-fitzpatrick2020Fear}{\textbf{fitzpatrick2020Fear?}}~; \protect\hyperlink{ref-serafini2021Psychological}{\textbf{serafini2021Psychological?}}). The
intersectionality of race, sex, and health conditions has also been explored in
the Canadian context of the pandemic. (\protect\hyperlink{ref-ingen2021Neighbourhoodlevel}{\textbf{ingen2021Neighbourhoodlevel?}}) demonstrated
that Ontarian neighborhoods with immigrant and racialized populations have a
disproportionate share of COVID-19 infections. In addition, areas with high SES
report a lesser number of COVID cases than low SES ones
(\protect\hyperlink{ref-ingen2021Neighbourhoodlevel}{\textbf{ingen2021Neighbourhoodlevel?}}). Another study conducted in Quebec confirms the
strong association between immigrants' backgrounds, SES, and COVID-19 related
psychological stress (\protect\hyperlink{ref-miconi2021Ethnocultural}{\textbf{miconi2021Ethnocultural?}}). The most recent reviews on the
topic suggest that immigrants disproportionately live in crowded and inadequate
housing and are employed in essential industries, leaving them with severe
income insecurities and poor physiological/psychological well-being
(\protect\hyperlink{ref-chew2020Clinical}{\textbf{chew2020Clinical?}}~; \protect\hyperlink{ref-davis2020Impact}{\textbf{davis2020Impact?}}). Studies on SES and health disparities
have emphasized that this situation is further complicated for immigrant women,
constrained by gender and afflicted by a significant power imbalance in the
household and workplace (\protect\hyperlink{ref-alarcao2021Intersections}{\textbf{alarcao2021Intersections?}}~; \protect\hyperlink{ref-esegbona-adeigbe2020COVID19}{\textbf{esegbona-adeigbe2020COVID19?}}~; \protect\hyperlink{ref-jehn2021COVID19}{\textbf{jehn2021COVID19?}}~; \protect\hyperlink{ref-mo2020Differential}{\textbf{mo2020Differential?}}~; \protect\hyperlink{ref-nardon2021Skilled}{\textbf{nardon2021Skilled?}}~; \protect\hyperlink{ref-sabri2020Effect}{\textbf{sabri2020Effect?}}~; \protect\hyperlink{ref-st-denis2020Sociodemographic}{\textbf{st-denis2020Sociodemographic?}}). According
to Lightman, immigrants and women of color are overrepresented in long-term care
facilities in Canada, where COVID infection rates have been exceptionally high.
They are also subject to exclusionary social and economic practices
(\protect\hyperlink{ref-lightman2021Caring}{\textbf{lightman2021Caring?}}).

The second category of the literature is primarily comprised of policy analyses,
research on public health communications, examining the efficacy of COVID
initiatives, the optimal allocation of government aids, and ethical issues
associated with applying new health interventions to vulnerable populations and
their communities (\protect\hyperlink{ref-aragona2020Negative}{\textbf{aragona2020Negative?}}~; \protect\hyperlink{ref-cross2021Coronavirus}{\textbf{cross2021Coronavirus?}}~; \protect\hyperlink{ref-desai2020COVID19}{\textbf{desai2020COVID19?}}~; \protect\hyperlink{ref-doyle2020Migrant}{\textbf{doyle2020Migrant?}}~; \protect\hyperlink{ref-gilman2020Modelling}{\textbf{gilman2020Modelling?}}~; \protect\hyperlink{ref-hargreaves2020Targeting}{\textbf{hargreaves2020Targeting?}}~; \protect\hyperlink{ref-langellier2020Policy}{\textbf{langellier2020Policy?}}~; \protect\hyperlink{ref-nezafatmaldonado2020Engaging}{\textbf{nezafatmaldonado2020Engaging?}}~; \protect\hyperlink{ref-silverman2020Ethics}{\textbf{silverman2020Ethics?}}~; \protect\hyperlink{ref-wilson2020Ethics}{\textbf{wilson2020Ethics?}}). Some of the questions that this body
of research addresses include: how best to respond to COVID-19 in immigrant
communities? How do public health policies impact the socio-economic status and
psychological well-being of different sub-populations? What is the structure of
the planning and implementation of government aid programs during the current
crisis? And, how can we design public health communication strategies in a way
that is culturally sensitive and inclusive?

There are several knowledge gaps in the existing literature. Most of the cited
studies do not include household data and family structure of immigrants as
variables that affect exposure risks and health outcomes of COVID-19.
Consequently, the relationship between immigrant status and COVID-19 is not
accounted for by household size, and family characteristics (i.e., income,
education, religion). A key reason is that most of the health data used in such
research projects are acquired from national registers or administrative
databases, which provide little information on immigration status and kinship.
However, as the study of Thomeer and her colleagues theorizes, the extent of
adverse social and economic impacts of COVID-19 is significantly conditional on
the composition of an immigrant's family (\protect\hyperlink{ref-thomeer2020How}{\textbf{thomeer2020How?}}). Households that
include a spouse/partner, children, or multigenerational family members most
likely experience a different level of social, psychological, and medical
vulnerability than single or nuclear ones. Therefore, studies of immigrant
populations should account for the breakdown of family configuration to better
understand the mechanisms linking social determinants and health outcomes.
Another knowledge gap concerns the current state of Canadian studies of COVID-19
and its public health crisis. There is a dearth of literature on the challenges
faced by immigrants and their families in the country. Although the burden of
COVID-19 infection is disproportionately higher among these marginalized groups
(St-Denis, 2020; Sundaram et al., 2020), research on the issues they are
experiencing is sparse and often limited to one aspect.

\newpage

\hypertarget{theoretical-framework}{%
\section{Theoretical framework}\label{theoretical-framework}}

Family as a unit of analysis can be particularly fruitful when studying health
inequalities in the context of a societal-wide crisis. Through the lens of the
COVID-19 pandemic, the following points illustrate why exploring family dynamics
provides additional insight into the underlying mechanisms of health
disparities. First, the family is the most fundamental unit of social
organization. It facilitates access to resources and the transmission of social
and psychological capital from one generation to the next. Therefore, a better
analysis of how crises and adverse life events affect people requires a better
understanding of how families relate to each other and interact with external
forces. Second, during public health emergencies, families may be thrown into
different structures due to unprecedented financial, mental, and physical
challenges. The risk of family dissolution, for example, can be heightened when
members must spend unusually long periods together at home and have limited
opportunities to engage with friends and the wider community. Conversely,
families can develop closer bonds through self-sacrifice, shared
responsibilities, and increased care and support. As the dependency status of
seniors increases, the elderlies are also more likely to be integrated into
formerly nucleus families. We posit that the extent to which a family can adapt
its functions to meet new challenges is a predictor of the well-being of its
members and the resilience of its broader population. Third, the family is a
crucial link to social institutions, such as the health care system, schools,
and government agencies whose decisions have significant downstream impacts.
Families are viewed by many public-level entities as a reference when allocating
resources, enforcing new standards, and responding to emerging needs and
demands. From this perspective, the family operates at the nexus of individual
and societal structures, embedding the tripple effects of public policy in its
most consequential ways. Finally, the family has a crucial role to play in
determining the life course of its members. As COVID-19 elevates the importance
of familial bonds, the post-pandemic period might be marked by an increased
influence of families on the lives of their members and the social structures
they inhabit.

The above conceptualization echoes several other approaches to studying
long-term health inequalities. For instance, life course theory emphasizes the
significance of family ties and social environment as they structure prospects
for individuals across their life span (Elder, 1977; Elder Jr., 1998; Macmillan,
2005; Settersten et al., 2020). Likewise, social capital theory and the
stress-buffering hypothesis consider family systems a crucial factor in health
and well-being (Choi and Jun, 2020; Cohen and McKay, 1984; Cohen and Wills,
1985; Harpham et al., 2004; Kawachi and Berkman, 2001; Lavee et al., 1987;
Mandelbaum et al., 2020). They show that the family can offer tangible (e.g.,
food, shelter, and money) and intangible resources (e.g., support, social
status, psychological comfort, and institutional advantages), protecting its
members from adverse life events. Similarly, the social-ecological model and the
theory of fundamental causes highlight how the SES, familial and extrafamilial
factors pattern people's life trajectories and health outcomes (Barkan and
Rocque, 2018; Clouston and Link, 2021; Figge et al., 2018; Kim, 2020; Libman et
al., 2012; Ungar, 2015). However, all these approaches are ostensibly focused on
life periods where the general social context is relatively stable. In this
article, we present our conceptualization of the family in times of
unprecedented social upheaval. Specifically, our framework seeks to characterize
family dynamics as a suitable lens to analyze health inequalities in the face of
a virulent public health crisis.

\newpage

\hypertarget{methodology}{%
\section{Methodology}\label{methodology}}

\hypertarget{data}{%
\subsection{Data}\label{data}}

The data for this study come from an original collection of biographical data on
immigrants and the native population of Quebec to analyse their participation in
Quebec society in various fields (social, economic, cultural, etc.). The
immigrant population is limited to recent immigration that has taken place over
the past 15 years. This survey has several components that retrace the life
course of individuals over a long period of time. The pandemic component uses a
reduced version of the \href{https://www.phenxtoolkit.org/toolkit_content/PDF/CPTS_CEFIS_FullSurveyEng.pdf}{harmonized questionnaire on exposure and the impact of
Covid19 on the family
(CEFIS)}.
The questions are divided into three parts, on exposure to the disease, its
consequences and the benefit of the policies put in place by both the federal
and provincial governments. The other parts of the questionnaire focus on the
characteristics of the respondents, their immigration history, their educational
and family background. In short, it is a detailed questionnaire that covers the
life of immigrants from birth and especially over the past 15 years. The data
collection lasted three months, from August to November 2020. In total, 1,500
people were surveyed, including around 1,000 immigrants and 500 natives
throughout Quebec.

\hypertarget{dependent-variables}{%
\subsubsection{Dependent variables}\label{dependent-variables}}

We distinguish three groups of dependent variables:

\textbf{1. Disease exposure variables}: these are measured by six questions that seek
to document whether the individual through work or a family member may have been
exposed to the disease. For example, it is asked if a family member has had
symptoms of the disease, or if a member has been hospitalized due to the
contraction of the disease.

\textbf{2. Variables on the family consequences of the disease}: we measure three
dimensions of consequences. The first group of measures assess the consequences
on accessing ressources by the respondent or any member of his family. The
second group of measure assess whether the respondent or any family member have
difficulties to perform usual duties toward any member of the family. Finally,
the last group of measures measure the consequences of the pandemic on
employment and income of any family member.

\textbf{3. Variables on government assistance}: Finally, government support is
measured by the following two questions that assessed whether the respondent or
any family member has benefited of the temporary assistance for workers from the
Government of Quebec (PATT) or of the Canadian Federal Government Emergency
Benefit (CEP). The table XXX in the apendix presents the full set of indicators
of the dependents variables

\hypertarget{main-independent-variable}{%
\subsubsection{Main independent variable}\label{main-independent-variable}}

The main independent variable is immigrant status coupled with visible minority
status. This variable takes the following modalities: 1 if the respondent is a
non-racialized native, 2 if he is a non-racialized immigrant, 3 if he is a
racialized immigrant, and 4 if the respondent is a racialized native or an
immigrant who has not answered the visible minority question. The other
variables in the analysis are: age, sex, length of residence, minority status,
size of social network.

\hypertarget{analysis-methods}{%
\subsection{Analysis methods}\label{analysis-methods}}

We use two main methods of analysis, a descriptive method and an explanatory
method. In the descriptive method, we consider each of the independent variables
in isolation and describe how they vary depending on the main independent
variable. The explanatory methods goes further by assessing whether the effect
remain when controlling for other covariates. Thus, by controlling for the other
explanatory variables, we determine the effect of immigrant status on exposure
to the pandemic, on its consequences and on the fact of benefiting or not of
government assistance. In a second series of modelling, we assess whether the
effect that we observed varies by the immigrant's length of residence or by
marital status. We use either a logistic regression on the outcome variables
that are binary and an ordonned logistic regression for the outcome that are
measure with a 5-point Lickert scale.

\newpage

\hypertarget{results}{%
\section{3. Results}\label{results}}

\hypertarget{sample-distribution}{%
\subsection{3.1. Sample distribution}\label{sample-distribution}}

\hypertarget{exposure-by-immigration-status}{%
\subsection{3.2. Exposure by immigration status}\label{exposure-by-immigration-status}}

Figure 1 presents the exposure to the pandemy accroding to the immigrants and
minority visible status. For all the six measures of exposure, it appears that
the majority group is the most exposed to the COVID-19 in all instance. The big
difference between the groups is about the exposure trhough working outside.
Around 50\% of the member of this group or a family member work outside the home
at the time of the survey. This percentage is only 30\% for the racialised
immigrants and 25\% for the non-racialised immigrants.

\begin{figure}
\centering
\includegraphics[width=0.6\textwidth,height=\textheight]{Figures/Exposition1.png}
\caption{Figure 1.1: Exposure}
\end{figure}

\begin{figure}
\centering
\includegraphics{Figures/Exposure.jpeg}
\caption{Figure 1.2: Explicative model}
\end{figure}

\hypertarget{consequences-by-immigrations-status}{%
\subsection{3.3. Consequences by immigrations status}\label{consequences-by-immigrations-status}}

The Figure 2 presents the consequences of the COVID-19 for a family member to
access to ressources. It shows that in all accounts but one (getting medecine),
the racialised immigrants express more diffciult to assess ressources. Around
20\% of them stated having difficulty to get health care compare, around 10\% for
obtaining food and 7\% for obtaining family insurrance. Comparatively, these
proportion are respectively 13\%, 7\% and 4\% for people and thier family member in
the majority group and 19\%, 3\% and 4\% respectively among the racialised
immigrants.

\begin{figure}
\centering
\includegraphics[width=0.8\textwidth,height=\textheight]{Figures/Famille1.png}
\caption{Figure 3.1: Consequences on accessing
ressources}
\end{figure}

\begin{figure}
\centering
\includegraphics{Figures/Ressources.jpeg}
\caption{Figure 3.2: Explicative model}
\end{figure}

\hypertarget{consequences-on-performing-usual-duties}{%
\subsection{3.4. Consequences on performing usual duties}\label{consequences-on-performing-usual-duties}}

The Figure 4 compares the response from the three groups on performing usual
duties. It appears that for all groups the situation caused by the COVID-19 has
inexpectedly improved the situation in various way, irrespective of the
immigrants status of the respondents.

\begin{figure}
\centering
\includegraphics{Figures/Entente1.png}
\caption{Figure 4.1: Consequences on performing usual duties}
\end{figure}

\begin{figure}
\centering
\includegraphics{Figures/Abilities.jpeg}
\caption{Figure 4.2: Explicative model}
\end{figure}

\hypertarget{consequences-on-employment-by-immigration-status}{%
\subsection{3.5. Consequences on employment by immigration status}\label{consequences-on-employment-by-immigration-status}}

The consequences on employment are mixed, with the severe consequences more
prevalent among the racialised immigrants. The Figure 3 shows for instance that
close to 45\% of the racialised immigrants have stated that their family income
has decreased compare to just 33\% of the majority group and 37\% of the
non-racialed immigrants. In addition, 11\% of the racialised immigrants reveals
having loss their job compared to 9\% among the non-racialised immigrants and 2\%
among members of the majority group. At the opposite, the members of the
majority groups are more likely to report a reduction of the working hours or
for stopping momentarily their work.

\begin{figure}
\centering
\includegraphics{Figures/Emploi1.png}
\caption{Figure 5.1: Consequences on employment and income}
\end{figure}

\begin{figure}
\centering
\includegraphics{Figures/Employment.jpeg}
\caption{Figure 5.2: Explicative model}
\end{figure}

\hypertarget{help-received-by-immigrations-status}{%
\subsection{3.7 Help received by immigrations status}\label{help-received-by-immigrations-status}}

Finally, we compared the different group on benefitting of the government
assistance. It appears that racialised immigrants with 35\% of the case have
expressed having received the Canada Emergency Response Benefit wheras this
proportion is 31\% for the non-racialised immigrants and 25\% for the member of
the majority group. For the province Temporary Aid for Workers Program, around
5\% of the respondents have received it.

\includegraphics{Figures/Aide1.png}

\begin{figure}
\centering
\includegraphics{Figures/Government.jpeg}
\caption{Figure 3.2: Explicative model}
\end{figure}

\newpage

\hypertarget{conclusion}{%
\section{Conclusion}\label{conclusion}}

The preliminary results of this study show that the consequences of the pandemic
on the Quebec population differ according to the population groups, but also
according to the type of consequences.

\newpage

\hypertarget{ruxe9fuxe9rences}{%
\section{Références}\label{ruxe9fuxe9rences}}

\url{http://www.oecd.org/coronavirus/policy-responses/what-is-the-impact-of-the-covid-19-pandemic-on-immigrants-and-their-children-e7cbb7de/}

Guttmann A, F. et al.~(2020), COVID-19 in Immigrants, Refugees and Other
Newcomers in Ontario: Characteristics of Those Tested and Those Confirmed
Positive, as of June 13, 2020, ICES, \url{http://www.ices.on.ca}.

Hu, Y. (2020), ``Intersecting ethnic and native--migrant inequalities in the
economic impact of the COVID-19 pandemic in the UK'', Research in Social
Stratification and Mobility, Vol. 68, p.~100528,
\url{http://dx.doi.org/10.1016/j.rssm.2020.100528}.

Mcginnity, F. and G. Kingston (2017), ``An Irish Welcome? Changing Irish
Attitudes to Immigrants and Immigration: The Role of Recession and Immigration'',
The Economic and social review, Vol. 48/3, pp.~281-304,
\url{https://www.researchgate.net/publication/319670153_An_Irish_Welcome_Changing_Irish_Attitudes_to_Immigrants_and_Immigration_The_Role_of_Recession_and_Immigration}
(accessed on 2 October 2020).

Raisi-Estabragh, Z. et al.~(2020), ``Greater risk of severe COVID-19 in Black,
Asian and Minority Ethnic populations is not explained by cardiometabolic,
socioeconomic or behavioural factors, or by 25(OH)-vitamin D status: study of
1326 cases from the UK Biobank'', Journal of Public Health, Vol. 42/3, pp.
451-460, \url{http://dx.doi.org/10.1093/pubmed/fdaa095}.

Apea, V. et al.~(2020), Ethnicity and outcomes in patients hospitalised with
COVID-19 infection in East London: an observational cohort study, Cold Spring
Harbor Laboratory, \url{http://dx.doi.org/10.1101/2020.06.10.20127621}.

Bertocchi, G. and A. Dimico (2020), ``Covid-19, Race, and Redlining'', IZA
Discussion Paper, Vol. 13467,
\url{https://papers.ssrn.com/sol3/papers.cfm?abstract_id=3648807} (accessed on 22
September 2020).

Brun, S. and P. Simon (2020), ``Inégalités ethno-raciales et coronavirus'', De
Facto 19.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# Importer package}
\FunctionTok{library}\NormalTok{(dplyr)}
\FunctionTok{library}\NormalTok{(tidyverse)}
\FunctionTok{library}\NormalTok{(summarytools)}
\FunctionTok{library}\NormalTok{(readr)}
\FunctionTok{library}\NormalTok{(questionr)}
\FunctionTok{library}\NormalTok{(rcompanion)}
\FunctionTok{library}\NormalTok{(pander)}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# Set numero graphique}
\NormalTok{no\_graphic }\OtherTok{\textless{}{-}} \SpecialCharTok{{-}}\DecValTok{1}
\end{Highlighting}
\end{Shaded}

\hypertarget{introduction-background-1}{%
\section{Introduction / background}\label{introduction-background-1}}

The COVID-19 pandemic has highlighted the vast ethnic, social, economic, and
health disparities in high-income countries, including Canada. One of the most
vulnerable segments of these affluent societies is the immigrant population,
that has become widely diversified with more and more people coming from
developing countries, and for whom integration into their new society may be
more difficult. The coronavirus pandemic brings a new challenge for these
immigrant populations, directly or indirectly through the tensions between the
ethnic groups that it helps to generate (\protect\hyperlink{ref-coates2020Covid19}{\textbf{coates2020Covid19?}}). These immigrants
face language barriers, low social capital, housing insecurity, and severe
poverty, rendering them particularly susceptible to the effects of the current
pandemic. Immigrants are also more likely to be employed as essential workers
where social distancing, sanitary precautions, and self-isolation, measures put
in place to mitigate the pandemy, are challenging (\protect\hyperlink{ref-reid2021Migrant}{\textbf{reid2021Migrant?}}~; \protect\hyperlink{ref-yi2021Health}{\textbf{yi2021Health?}}~; \protect\hyperlink{ref-zhang2021COVID19}{\textbf{zhang2021COVID19?}}). More importantly, migrants may encounter
various forms of discrimination, leading to social exclusion and restricted
access to healthcare and social services (\protect\hyperlink{ref-spiritus-beerden2021Mental}{\textbf{spiritus-beerden2021Mental?}}). It is
therefore necessary to understand how this pandemic affects these immigrant
populations differently and the native populations in Quebec. In doing so, this
article will reveal a better understanding of how the current pandemic
aggravates existing inequalities and challenges in the province for immigrants
and their descendants.

Since the start of the pandemic, most of the confirmed cases were identified in
Montreal and areas with high immigrant populations (\protect\hyperlink{ref-miconi2021Ethnocultural}{\textbf{miconi2021Ethnocultural?}}~; \protect\hyperlink{ref-shields2020COVID19}{\textbf{shields2020COVID19?}}). The coronavirus pandemic has infected nearly 450,000
people in Quebec and has resulted in more than 11,000 deaths
(\url{https://www.quebec.ca/sante/problemes-de-sante/a-z/coronavirus-2019/situation-coronavirus-quebec},
consulted on octobre 29, 2021). This makes it the worst affected province in
Canada and one of the worst affected regions in the world during the first wave
of the pandemic. Although the majority of these deaths took place in elderly's
homes, and therefore affected more this segment of the population, its overall
influence on individuals is not to be overlooked. It has been shown that, on the
one hand, immigrants and their families have a higher risk of being exposed to
the pathogen, and, at the same time, they are more vulnerable to racism and
discrimination and face barriers to accessing social support (\protect\hyperlink{ref-2021CRARR}{\textbf{2021CRARR?}}) and
information, especially for those who didn't speak or understand the two
official langages.
Several
studies have thus shown excess mortality in ethnocultural minority populations
(\protect\hyperlink{ref-bertocchi2020Covid19}{\textbf{bertocchi2020Covid19?}}), higher unemployment, greater loss of income and
negative consequences for children and families in several countries. This
differentiated effect can be explained in different ways. On the one hand, in
terms of the direct consequences of the pandemic, the situation of
precariousness and greater poverty among immigrants exposes them more to the
more harmful consequences of the disease on their state of health. Likewise,
immigrants often work in jobs where social distancing is difficult to enforce.
On the other hand, indirectly, the solutions to the crisis affect immigrants and
natives differently. The complete shutdown of tye economy with its set of
adjustments has greatly impeded the employability of immigrants, especially
newcomers who have not yet grasped the workings of society and who cannot
benefit from state subsidies. But, it is also shown that these effects are not
uniform on the whole of the immigrant population, itself, very diverse.

Despite the growing focus on the differential impacts of the COVID-19 on
immigrants and non-immigrants segment of the population, there is still very
little knowledge about how the pandemic and its related public policy impact the
lives of immigrants and their families. The existing health datasets (largely
used to assess this effect) are not necessarily representative of the immigrant
population and do not offer an accurate picture of the complex social status and
health outcomes among this group (\protect\hyperlink{ref-choi2021Studying}{\textbf{choi2021Studying?}}~; \protect\hyperlink{ref-gagnon2021Immigration}{\textbf{gagnon2021Immigration?}}~; \protect\hyperlink{ref-khan2015Visible}{\textbf{khan2015Visible?}}). In most cases, these datasets lack data on the residence
length of immigrants and combine newcomers with long-term settlers
(\protect\hyperlink{ref-edge2013Discrimination}{\textbf{edge2013Discrimination?}}~; \protect\hyperlink{ref-hayward2021Clinical}{\textbf{hayward2021Clinical?}}). As a result, they may not allow
a detailed assessment of the nature and extent of risk exposure in sub-segments
of populations, as well as whether interventions are reaching those most at
risk. In addition, the family-level impacts of the pandemic are still poorly
understood. While a body of research has explored the association between the
pandemic and its potential effects on intimate partner violence and marital
dissolution (\protect\hyperlink{ref-bhatnagar2021Navigating}{\textbf{bhatnagar2021Navigating?}}~; \protect\hyperlink{ref-koshan2020COVID19}{\textbf{koshan2020COVID19?}}~; \protect\hyperlink{ref-morrison2021Chapter}{\textbf{morrison2021Chapter?}}~; \protect\hyperlink{ref-prime2020Risk}{\textbf{prime2020Risk?}}), few studies have examined negative familial repercussions in
the context of immigration status. There is ample evidence that sustained family
relationships greatly enhance immigrants' lives on several fronts, including
fostering economic prosperity (\protect\hyperlink{ref-marger2001Sociala}{\textbf{marger2001Sociala?}}), psychological well-being
(\protect\hyperlink{ref-beiser2002Povertya}{\textbf{beiser2002Povertya?}}), social network development (\protect\hyperlink{ref-evansluong2017Familya}{\textbf{evansluong2017Familya?}}), and
educational success for children (\protect\hyperlink{ref-munroe-blum1989Immigranta}{\textbf{munroe-blum1989Immigranta?}}). The family is,
therefore, a key component of long-term social mobility and economic success of
immigrants.

Based on an original biographic survey collected around the first wave of the
pandemic in august-november 2020 on immigrants and the native population of
Québec, this study mainly aims to understand the family exposure, consequences,
and responses of the pandemic on immigrant and native Quebec populations. We
assess a whole range of different measures of each outcome. We divided the
consequences in three dimensions, 1) access to ressources, 2) ability to perform
duties and 3) employment and income. We seek also to analyse in what ways these
outcomes differ according to the immigrants length of residence. By doing so,
this study provide a clear and global picture of the situation of different
population groups in Quebec in the face of the COVID-19.

\newpage

\newpage

\hypertarget{current-body-of-literature-and-knowledge-gaps-1}{%
\section{Current body of literature and knowledge gaps}\label{current-body-of-literature-and-knowledge-gaps-1}}

The literature on the racial and ethnic health disparities during the COVID-19
crisis can be divided into two main categories: studies addressing the
socio-economic determinants of health and those examining the delivery of health
interventions and policies. The first category is primarily comprised of
research that focuses on the link between socio-economic status (SES) of
high-risk minority groups and the differential impact of the pandemic
(\protect\hyperlink{ref-drefahl2020Populationbased}{\textbf{drefahl2020Populationbased?}}~; \protect\hyperlink{ref-greenaway2020COVID19}{\textbf{greenaway2020COVID19?}}~; \protect\hyperlink{ref-lopez2021Racial}{\textbf{lopez2021Racial?}}~; \protect\hyperlink{ref-pacheco2020Job}{\textbf{pacheco2020Job?}}~; \protect\hyperlink{ref-sundaram2020Individual}{\textbf{sundaram2020Individual?}}). For instance, an inter-country study
has found that Middle Eastern, African, and Nordic immigrants in Sweden have a
higher risk of being hospitalized and even dying from COVID-19 when compared to
the Swedish born population. In the same study, the SES, proportion of
working-age households, and neighborhood density were significantly associated
with the mortality rate (\protect\hyperlink{ref-rostila2021Disparities}{\textbf{rostila2021Disparities?}}). This trend of research
demonstrates that racial and ethnic minorities are more likely to experience
stress factors, not only due to the hazards of infection but also because of
economic and social marginalization (\protect\hyperlink{ref-alahmad2020COVID19}{\textbf{alahmad2020COVID19?}}~; \protect\hyperlink{ref-choi2020Buffering}{\textbf{choi2020Buffering?}}~; \protect\hyperlink{ref-endale2020COVID19}{\textbf{endale2020COVID19?}}~; \protect\hyperlink{ref-fitzpatrick2020Fear}{\textbf{fitzpatrick2020Fear?}}~; \protect\hyperlink{ref-serafini2021Psychological}{\textbf{serafini2021Psychological?}}). The
intersectionality of race, sex, and health conditions has also been explored in
the Canadian context of the pandemic. (\protect\hyperlink{ref-ingen2021Neighbourhoodlevel}{\textbf{ingen2021Neighbourhoodlevel?}}) demonstrated
that Ontarian neighborhoods with immigrant and racialized populations have a
disproportionate share of COVID-19 infections. In addition, areas with high SES
report a lesser number of COVID cases than low SES ones
(\protect\hyperlink{ref-ingen2021Neighbourhoodlevel}{\textbf{ingen2021Neighbourhoodlevel?}}). Another study conducted in Quebec confirms the
strong association between immigrants' backgrounds, SES, and COVID-19 related
psychological stress (\protect\hyperlink{ref-miconi2021Ethnocultural}{\textbf{miconi2021Ethnocultural?}}). The most recent reviews on the
topic suggest that immigrants disproportionately live in crowded and inadequate
housing and are employed in essential industries, leaving them with severe
income insecurities and poor physiological/psychological well-being
(\protect\hyperlink{ref-chew2020Clinical}{\textbf{chew2020Clinical?}}~; \protect\hyperlink{ref-davis2020Impact}{\textbf{davis2020Impact?}}). Studies on SES and health disparities
have emphasized that this situation is further complicated for immigrant women,
constrained by gender and afflicted by a significant power imbalance in the
household and workplace (\protect\hyperlink{ref-alarcao2021Intersections}{\textbf{alarcao2021Intersections?}}~; \protect\hyperlink{ref-esegbona-adeigbe2020COVID19}{\textbf{esegbona-adeigbe2020COVID19?}}~; \protect\hyperlink{ref-jehn2021COVID19}{\textbf{jehn2021COVID19?}}~; \protect\hyperlink{ref-mo2020Differential}{\textbf{mo2020Differential?}}~; \protect\hyperlink{ref-nardon2021Skilled}{\textbf{nardon2021Skilled?}}~; \protect\hyperlink{ref-sabri2020Effect}{\textbf{sabri2020Effect?}}~; \protect\hyperlink{ref-st-denis2020Sociodemographic}{\textbf{st-denis2020Sociodemographic?}}). According
to Lightman, immigrants and women of color are overrepresented in long-term care
facilities in Canada, where COVID infection rates have been exceptionally high.
They are also subject to exclusionary social and economic practices
(\protect\hyperlink{ref-lightman2021Caring}{\textbf{lightman2021Caring?}}).

The second category of the literature is primarily comprised of policy analyses,
research on public health communications, examining the efficacy of COVID
initiatives, the optimal allocation of government aids, and ethical issues
associated with applying new health interventions to vulnerable populations and
their communities (\protect\hyperlink{ref-aragona2020Negative}{\textbf{aragona2020Negative?}}~; \protect\hyperlink{ref-cross2021Coronavirus}{\textbf{cross2021Coronavirus?}}~; \protect\hyperlink{ref-desai2020COVID19}{\textbf{desai2020COVID19?}}~; \protect\hyperlink{ref-doyle2020Migrant}{\textbf{doyle2020Migrant?}}~; \protect\hyperlink{ref-gilman2020Modelling}{\textbf{gilman2020Modelling?}}~; \protect\hyperlink{ref-hargreaves2020Targeting}{\textbf{hargreaves2020Targeting?}}~; \protect\hyperlink{ref-langellier2020Policy}{\textbf{langellier2020Policy?}}~; \protect\hyperlink{ref-nezafatmaldonado2020Engaging}{\textbf{nezafatmaldonado2020Engaging?}}~; \protect\hyperlink{ref-silverman2020Ethics}{\textbf{silverman2020Ethics?}}~; \protect\hyperlink{ref-wilson2020Ethics}{\textbf{wilson2020Ethics?}}). Some of the questions that this body
of research addresses include: how best to respond to COVID-19 in immigrant
communities? How do public health policies impact the socio-economic status and
psychological well-being of different sub-populations? What is the structure of
the planning and implementation of government aid programs during the current
crisis? And, how can we design public health communication strategies in a way
that is culturally sensitive and inclusive?

There are several knowledge gaps in the existing literature. Most of the cited
studies do not include household data and family structure of immigrants as
variables that affect exposure risks and health outcomes of COVID-19.
Consequently, the relationship between immigrant status and COVID-19 is not
accounted for by household size, and family characteristics (i.e., income,
education, religion). A key reason is that most of the health data used in such
research projects are acquired from national registers or administrative
databases, which provide little information on immigration status and kinship.
However, as the study of Thomeer and her colleagues theorizes, the extent of
adverse social and economic impacts of COVID-19 is significantly conditional on
the composition of an immigrant's family (\protect\hyperlink{ref-thomeer2020How}{\textbf{thomeer2020How?}}). Households that
include a spouse/partner, children, or multigenerational family members most
likely experience a different level of social, psychological, and medical
vulnerability than single or nuclear ones. Therefore, studies of immigrant
populations should account for the breakdown of family configuration to better
understand the mechanisms linking social determinants and health outcomes.
Another knowledge gap concerns the current state of Canadian studies of COVID-19
and its public health crisis. There is a dearth of literature on the challenges
faced by immigrants and their families in the country. Although the burden of
COVID-19 infection is disproportionately higher among these marginalized groups
(St-Denis, 2020; Sundaram et al., 2020), research on the issues they are
experiencing is sparse and often limited to one aspect.

\newpage

\hypertarget{theoretical-framework-1}{%
\section{Theoretical framework}\label{theoretical-framework-1}}

Family as a unit of analysis can be particularly fruitful when studying health
inequalities in the context of a societal-wide crisis. Through the lens of the
COVID-19 pandemic, the following points illustrate why exploring family dynamics
provides additional insight into the underlying mechanisms of health
disparities. First, the family is the most fundamental unit of social
organization. It facilitates access to resources and the transmission of social
and psychological capital from one generation to the next. Therefore, a better
analysis of how crises and adverse life events affect people requires a better
understanding of how families relate to each other and interact with external
forces. Second, during public health emergencies, families may be thrown into
different structures due to unprecedented financial, mental, and physical
challenges. The risk of family dissolution, for example, can be heightened when
members must spend unusually long periods together at home and have limited
opportunities to engage with friends and the wider community. Conversely,
families can develop closer bonds through self-sacrifice, shared
responsibilities, and increased care and support. As the dependency status of
seniors increases, the elderlies are also more likely to be integrated into
formerly nucleus families. We posit that the extent to which a family can adapt
its functions to meet new challenges is a predictor of the well-being of its
members and the resilience of its broader population. Third, the family is a
crucial link to social institutions, such as the health care system, schools,
and government agencies whose decisions have significant downstream impacts.
Families are viewed by many public-level entities as a reference when allocating
resources, enforcing new standards, and responding to emerging needs and
demands. From this perspective, the family operates at the nexus of individual
and societal structures, embedding the tripple effects of public policy in its
most consequential ways. Finally, the family has a crucial role to play in
determining the life course of its members. As COVID-19 elevates the importance
of familial bonds, the post-pandemic period might be marked by an increased
influence of families on the lives of their members and the social structures
they inhabit.

The above conceptualization echoes several other approaches to studying
long-term health inequalities. For instance, life course theory emphasizes the
significance of family ties and social environment as they structure prospects
for individuals across their life span (Elder, 1977; Elder Jr., 1998; Macmillan,
2005; Settersten et al., 2020). Likewise, social capital theory and the
stress-buffering hypothesis consider family systems a crucial factor in health
and well-being (Choi and Jun, 2020; Cohen and McKay, 1984; Cohen and Wills,
1985; Harpham et al., 2004; Kawachi and Berkman, 2001; Lavee et al., 1987;
Mandelbaum et al., 2020). They show that the family can offer tangible (e.g.,
food, shelter, and money) and intangible resources (e.g., support, social
status, psychological comfort, and institutional advantages), protecting its
members from adverse life events. Similarly, the social-ecological model and the
theory of fundamental causes highlight how the SES, familial and extrafamilial
factors pattern people's life trajectories and health outcomes (Barkan and
Rocque, 2018; Clouston and Link, 2021; Figge et al., 2018; Kim, 2020; Libman et
al., 2012; Ungar, 2015). However, all these approaches are ostensibly focused on
life periods where the general social context is relatively stable. In this
article, we present our conceptualization of the family in times of
unprecedented social upheaval. Specifically, our framework seeks to characterize
family dynamics as a suitable lens to analyze health inequalities in the face of
a virulent public health crisis.

\newpage

\hypertarget{methodology-1}{%
\section{Methodology}\label{methodology-1}}

\hypertarget{data-1}{%
\subsection{Data}\label{data-1}}

The data for this study come from an original collection of biographical data on
immigrants and the native population of Quebec to analyse their participation in
Quebec society in various fields (social, economic, cultural, etc.). The
immigrant population is limited to recent immigration that has taken place over
the past 15 years. This survey has several components that retrace the life
course of individuals over a long period of time. The pandemic component uses a
reduced version of the \href{https://www.phenxtoolkit.org/toolkit_content/PDF/CPTS_CEFIS_FullSurveyEng.pdf}{harmonized questionnaire on exposure and the impact of
Covid19 on the family
(CEFIS)}.
The questions are divided into three parts, on exposure to the disease, its
consequences and the benefit of the policies put in place by both the federal
and provincial governments. The other parts of the questionnaire focus on the
characteristics of the respondents, their immigration history, their educational
and family background. In short, it is a detailed questionnaire that covers the
life of immigrants from birth and especially over the past 15 years. The data
collection lasted three months, from August to November 2020. In total, 1,500
people were surveyed, including around 1,000 immigrants and 500 natives
throughout Quebec.

\hypertarget{dependent-variables-1}{%
\subsubsection{Dependent variables}\label{dependent-variables-1}}

We distinguish three groups of dependent variables:

\textbf{1. Disease exposure variables}: these are measured by six questions that seek
to document whether the individual through work or a family member may have been
exposed to the disease. For example, it is asked if a family member has had
symptoms of the disease, or if a member has been hospitalized due to the
contraction of the disease.

\textbf{2. Variables on the family consequences of the disease}: we measure three
dimensions of consequences. The first group of measures assess the consequences
on accessing ressources by the respondent or any member of his family. The
second group of measure assess whether the respondent or any family member have
difficulties to perform usual duties toward any member of the family. Finally,
the last group of measures measure the consequences of the pandemic on
employment and income of any family member.

\textbf{3. Variables on government assistance}: Finally, government support is
measured by the following two questions that assessed whether the respondent or
any family member has benefited of the temporary assistance for workers from the
Government of Quebec (PATT) or of the Canadian Federal Government Emergency
Benefit (CEP). The table XXX in the apendix presents the full set of indicators
of the dependents variables

\hypertarget{main-independent-variable-1}{%
\subsubsection{Main independent variable}\label{main-independent-variable-1}}

The main independent variable is immigrant status coupled with visible minority
status. This variable takes the following modalities: 1 if the respondent is a
non-racialized native, 2 if he is a non-racialized immigrant, 3 if he is a
racialized immigrant, and 4 if the respondent is a racialized native or an
immigrant who has not answered the visible minority question. The other
variables in the analysis are: age, sex, length of residence, minority status,
size of social network.

\hypertarget{analysis-methods-1}{%
\subsection{Analysis methods}\label{analysis-methods-1}}

We use two main methods of analysis, a descriptive method and an explanatory
method. In the descriptive method, we consider each of the independent variables
in isolation and describe how they vary depending on the main independent
variable. The explanatory methods goes further by assessing whether the effect
remain when controlling for other covariates. Thus, by controlling for the other
explanatory variables, we determine the effect of immigrant status on exposure
to the pandemic, on its consequences and on the fact of benefiting or not of
government assistance. In a second series of modelling, we assess whether the
effect that we observed varies by the immigrant's length of residence or by
marital status. We use either a logistic regression on the outcome variables
that are binary and an ordonned logistic regression for the outcome that are
measure with a 5-point Lickert scale.

\newpage

\hypertarget{results-1}{%
\section{3. Results}\label{results-1}}

\hypertarget{sample-distribution-1}{%
\subsection{3.1. Sample distribution}\label{sample-distribution-1}}

\hypertarget{exposure-by-immigration-status-1}{%
\subsection{3.2. Exposure by immigration status}\label{exposure-by-immigration-status-1}}

Figure 1 presents the exposure to the pandemy accroding to the immigrants and
minority visible status. For all the six measures of exposure, it appears that
the majority group is the most exposed to the COVID-19 in all instance. The big
difference between the groups is about the exposure trhough working outside.
Around 50\% of the member of this group or a family member work outside the home
at the time of the survey. This percentage is only 30\% for the racialised
immigrants and 25\% for the non-racialised immigrants.

\begin{figure}
\centering
\includegraphics[width=0.6\textwidth,height=\textheight]{Figures/Exposition1.png}
\caption{Figure 1.1: Exposure}
\end{figure}

\begin{figure}
\centering
\includegraphics{Figures/Exposure.jpeg}
\caption{Figure 1.2: Explicative model}
\end{figure}

\hypertarget{consequences-by-immigrations-status-1}{%
\subsection{3.3. Consequences by immigrations status}\label{consequences-by-immigrations-status-1}}

The Figure 2 presents the consequences of the COVID-19 for a family member to
access to ressources. It shows that in all accounts but one (getting medecine),
the racialised immigrants express more diffciult to assess ressources. Around
20\% of them stated having difficulty to get health care compare, around 10\% for
obtaining food and 7\% for obtaining family insurrance. Comparatively, these
proportion are respectively 13\%, 7\% and 4\% for people and thier family member in
the majority group and 19\%, 3\% and 4\% respectively among the racialised
immigrants.

\begin{figure}
\centering
\includegraphics[width=0.8\textwidth,height=\textheight]{Figures/Famille1.png}
\caption{Figure 3.1: Consequences on accessing
ressources}
\end{figure}

\begin{figure}
\centering
\includegraphics{Figures/Ressources.jpeg}
\caption{Figure 3.2: Explicative model}
\end{figure}

\hypertarget{consequences-on-performing-usual-duties-1}{%
\subsection{3.4. Consequences on performing usual duties}\label{consequences-on-performing-usual-duties-1}}

The Figure 4 compares the response from the three groups on performing usual
duties. It appears that for all groups the situation caused by the COVID-19 has
inexpectedly improved the situation in various way, irrespective of the
immigrants status of the respondents.

\begin{figure}
\centering
\includegraphics{Figures/Entente1.png}
\caption{Figure 4.1: Consequences on performing usual duties}
\end{figure}

\begin{figure}
\centering
\includegraphics{Figures/Abilities.jpeg}
\caption{Figure 4.2: Explicative model}
\end{figure}

\hypertarget{consequences-on-employment-by-immigration-status-1}{%
\subsection{3.5. Consequences on employment by immigration status}\label{consequences-on-employment-by-immigration-status-1}}

The consequences on employment are mixed, with the severe consequences more
prevalent among the racialised immigrants. The Figure 3 shows for instance that
close to 45\% of the racialised immigrants have stated that their family income
has decreased compare to just 33\% of the majority group and 37\% of the
non-racialed immigrants. In addition, 11\% of the racialised immigrants reveals
having loss their job compared to 9\% among the non-racialised immigrants and 2\%
among members of the majority group. At the opposite, the members of the
majority groups are more likely to report a reduction of the working hours or
for stopping momentarily their work.

\begin{figure}
\centering
\includegraphics{Figures/Emploi1.png}
\caption{Figure 5.1: Consequences on employment and income}
\end{figure}

\begin{figure}
\centering
\includegraphics{Figures/Employment.jpeg}
\caption{Figure 5.2: Explicative model}
\end{figure}

\hypertarget{help-received-by-immigrations-status-1}{%
\subsection{3.7 Help received by immigrations status}\label{help-received-by-immigrations-status-1}}

Finally, we compared the different group on benefitting of the government
assistance. It appears that racialised immigrants with 35\% of the case have
expressed having received the Canada Emergency Response Benefit wheras this
proportion is 31\% for the non-racialised immigrants and 25\% for the member of
the majority group. For the province Temporary Aid for Workers Program, around
5\% of the respondents have received it.

\includegraphics{Figures/Aide1.png}

\begin{figure}
\centering
\includegraphics{Figures/Government.jpeg}
\caption{Figure 3.2: Explicative model}
\end{figure}

\newpage

\hypertarget{conclusion-1}{%
\section{Conclusion}\label{conclusion-1}}

The preliminary results of this study show that the consequences of the pandemic
on the Quebec population differ according to the population groups, but also
according to the type of consequences. \newpage

\hypertarget{references}{%
\section{References}\label{references}}

\url{http://www.oecd.org/coronavirus/policy-responses/what-is-the-impact-of-the-covid-19-pandemic-on-immigrants-and-their-children-e7cbb7de/}

Guttmann A, F. et al.~(2020), COVID-19 in Immigrants, Refugees and Other
Newcomers in Ontario: Characteristics of Those Tested and Those Confirmed
Positive, as of June 13, 2020, ICES, \url{http://www.ices.on.ca}.

Hu, Y. (2020), ``Intersecting ethnic and native--migrant inequalities in the
economic impact of the COVID-19 pandemic in the UK'', Research in Social
Stratification and Mobility, Vol. 68, p.~100528,
\url{http://dx.doi.org/10.1016/j.rssm.2020.100528}.

Mcginnity, F. and G. Kingston (2017), ``An Irish Welcome? Changing Irish
Attitudes to Immigrants and Immigration: The Role of Recession and Immigration'',
The Economic and social review, Vol. 48/3, pp.~281-304,
\url{https://www.researchgate.net/publication/319670153_An_Irish_Welcome_Changing_Irish_Attitudes_to_Immigrants_and_Immigration_The_Role_of_Recession_and_Immigration}
(accessed on 2 October 2020).

Raisi-Estabragh, Z. et al.~(2020), ``Greater risk of severe COVID-19 in Black,
Asian and Minority Ethnic populations is not explained by cardiometabolic,
socioeconomic or behavioural factors, or by 25(OH)-vitamin D status: study of
1326 cases from the UK Biobank'', Journal of Public Health, Vol. 42/3, pp.
451-460, \url{http://dx.doi.org/10.1093/pubmed/fdaa095}.

Apea, V. et al.~(2020), Ethnicity and outcomes in patients hospitalised with
COVID-19 infection in East London: an observational cohort study, Cold Spring
Harbor Laboratory, \url{http://dx.doi.org/10.1101/2020.06.10.20127621}.

Bertocchi, G. and A. Dimico (2020), ``Covid-19, Race, and Redlining'', IZA
Discussion Paper, Vol. 13467,
\url{https://papers.ssrn.com/sol3/papers.cfm?abstract_id=3648807} (accessed on 22
September 2020).

Brun, S. and P. Simon (2020), ``Inégalités ethno-raciales et coronavirus'', De
Facto 19.

\end{document}

### 2023-05-02 Expérimentations (Théa)/1. Expériementations (code).Rmd
Quelques expériences supplémentaires de visualisation
J'ai tenté de représenter graphiquement l'association (non significative) testée à travers le modèle de régression logistique. Je pense que l'échec est dû au fait que ma variable dépendante est discrète (un indice situé entre 0 et 1 aurait probablement donné des résultats plus intéressants).

@@
          COV12 == 1 ~ 0,
-         COV12 == 2 ~ 1
+         COV12 == 2 ~ 1,
+         COV12 == 98 ~ NA,
+         COV12 == 99 ~ NA,
     ))
+
+#soustraction des réponses (la même année que la résidence permanente, je ne sais pas et je ne souhaite pas répondre)
+Base_Trajipaq_Finale_Ok <-
+  Base_Trajipaq_Finale_Ok %>%
+  filter(Q14I < 2019)
 
@@
 #les résultats indiquent que le risque d'erreur alpha est supérieur à 5%. Conséquemment, la relation étudiée n'est pas significative.
+```
+
+# Représentation graphique de la régression logistique
+
+```{r}
+#représentation graphique de la régression logistique
+ggplot(Base_Trajipaq_Finale_Ok %>% filter(!is.na(COV12_2), !is.na(Q14I)), aes(x = Q14I, y = COV12_2)) + 
+  geom_point() +
+  stat_smooth(method = "glm", color = "blue", se = FALSE,
+    method.args = list(family = binomial))
+
+#résultat invalide car COV_2 est une variable numérique discrète
 ```

### 2023-05-03 Expérimentations (Théa)/1. Expériementations (code).Rmd
Un essai supplémentaire de régression logistique
Je me suis essayée à faire une régression logistique avec une autre variable dépendante.

@@
 
+## Un autre test pour la régression logistique
+
+```{r}
+
+Base_Trajipaq_Finale_Ok <-
+  Base_Trajipaq_Finale_Ok %>%
+  filter(COV14 < 98) %>%
+  mutate(COV14 = case_when(
+         COV14 == 1 ~ 0,
+         COV14 == 2 ~ 1,
+  ))
+
+model3 <- glm(formula = COV14 ~ Q14I, family = "binomial", data = Base_Trajipaq_Finale_Ok)
+summary(model1)
+
+ggplot(Base_Trajipaq_Finale_Ok %>% filter(!is.na(COV14), !is.na(Q14I)), aes(x = Q14I, y = COV14)) + 
+  geom_point() +
+  stat_smooth(method = "glm", color = "blue", se = FALSE,
+    method.args = list(family = binomial))
+```
+
 # Analyse de la variance (ANOVA) entre deux variables numériques (VD = ordinal, VI = ratio)

### 2023-05-03 Expérimentations (Théa)/1. SOC3380 (data)/1. Analyse descriptive (script).Rmd
Expérimentations sur les données du cours SOC3380
Je me suis amusée à appliquer ce que j'ai appris sur les données de notre petite enquête dans le cours SOC3380.

---
title: "Expérimentations"
author: "Théa"
date: "2023-03-17"
output: html_document
---

```{r}
#Chargement des packages

library(dplyr)
library(tidyverse)
library(summarytools)
library(readr)
library(questionr)
library(rcompanion)
```

```{r}
#Chargement de la base de données

cr_SOC3380 <- read_csv("cr_SOC3380.csv")
Pays <- read_csv("Pays.csv")
```

```{r}
#Codage des variables

an_SOC3380 <-
  cr_SOC3380 %>%
  rename(
    "age" = "1 - Quel âge avez vous ?",
    "year_qc" = "2 - Depuis combien d'années habitez vous au Québec ?",
    "born_country" = "3a - Dans quel pays êtes vous né ?",
    "country_15" = "3b - Dans quel pays avez vous vécu jusqu'à vos 15 ans ?",
    "genre" = "4 - A quel genre vous identifiez vous ?",
    "statut" = "5 - Quel est votre statut actuel au Canada ?",
    "etat_civil" = "6 - Quel est votre état civil ?",
    "diplome" = "7 - Quel est votre niveau de diplôme le plus élevé ?",
    "revenu" = "8 - Dans quelle tranche se situe votre revenu actuel ?",
    "lang_enf" = "9 - Quelle est la première langue que vous avez apprise à la maison pendant votre enfance et que vous comprenez encore aujourd'hui ?",
    "lang_home" = "10 - Parmi ces langues, laquelle parlez vous le plus souvent à la maison ? Indiquez plus d'une langue seulement si elles sont parlées aussi souvent l'une que l'autre à la maison.",
    "lang_home_autre" = "11 - Parlez vous d'autres langues à la maison ?",
    "lang_trav" = "12 -  Parmi ces langues, laquelle parlez vous le plus souvent au travail ? Indiquez plus d'une langue seulement si elles sont parlées aussi souvent l'une que l'autre au travail.",
    "lang_trav_autre" = "13 - Parlez vous d'autres langues au travail ?",
    "rs_util" = "14 - Quel(s) réseaux sociaux utilisez vous régulièrement ?",
    "rs_time" = "15 - Combien de temps passez vous par jour sur les réseaux sociaux ?",
    "rs_disc_same" = "16 - Pour quelles raisons allez-vous sur les réseaux sociaux ? [Pour discuter avec des personnes qui proviennent de mon pays d'origine]",
    "rs_disc_diff" = "16 - Pour quelles raisons allez-vous sur les réseaux sociaux ? [Pour discuter avec des personnes qui ne proviennent pas de mon pays d'origine]",
    "rs_hab" = "16 - Pour quelles raisons allez-vous sur les réseaux sociaux ? [Par habitude]",
    "rs_div" = "16 - Pour quelles raisons allez-vous sur les réseaux sociaux ? [Pour me divertir]",
    "rs_ptime" = "16 - Pour quelles raisons allez-vous sur les réseaux sociaux ? [Pour passer le temps]",
    "rs_inf" = "16 - Pour quelles raisons allez-vous sur les réseaux sociaux ? [Pour m'informer]",
    "rs_renc_same" = "16 - Pour quelles raisons allez-vous sur les réseaux sociaux ? [Pour rencontrer des personnes de mon pays d'origine]",
    "rs_renc_diff" = "16 - Pour quelles raisons allez-vous sur les réseaux sociaux ? [Pour rencontrer des personnes qui ne proviennent pas de mon pays d'origine]",
    "rs_trav" = "16 - Pour quelles raisons allez-vous sur les réseaux sociaux ? [Pour le travail]",
    "online_memb_migr" = "17 - Êtes-vous membre de groupes en ligne regroupant des personnes immigrantes ?",
    "inf_pol_home" = "18 - Au cours du dernier mois, à quelle fréquence avez-vous utilisé les réseaux sociaux pour lire des journaux ou des revues en ligne traitant d'une question politique ou sociale précise en lien avec votre pays d'origine ?",
    "inf_polqc" = "19 - Au cours du dernier mois, à quelle fréquence avez-vous utilisé les réseaux sociaux pour lire des journaux ou des revues en ligne traitant d'une question politique ou sociale précise en lien avec le Québec ?",
    "ton_migr" = "20a - Quelle est la tonalité du contenu que vous observez sur les réseaux sociaux sur le thème de l'immigration au Québec ?",
    "ton_why" = "20b - Pourquoi selon vous ?",
    "disc_maj_same" = "21 - Sur les réseaux sociaux, les amis avec lesquels je discute le plus souvent sont majoritairement... [Des personnes provenant de mon pays d'origine]",
    "disc_maj_migr" = "21 - Sur les réseaux sociaux, les amis avec lesquels je discute le plus souvent sont majoritairement... [D'autres personnes immigrantes]",
    "disc_maj_qc" = "21 - Sur les réseaux sociaux, les amis avec lesquels je discute le plus souvent sont majoritairement... [Des personnes nées au Québec]",
    "ton_migr_pk" = "20b - Pourquoi selon vous ?",
    "act_cult" = "22 - Durant le dernier mois, à quelle fréquence avez-vous participé à des activités (culturelle, sportive, sociale, etc.) avec une personne qui ne provient pas de votre pays d'origine ?",
    "sentim_app" = "23 - Veuillez répondre selon votre niveau d'accord ou de désaccord avec les énoncés suivants: [J'ai le sentiment d'appartenir à la société québécoise.]",
    "chez_moi_qc" = "23 - Veuillez répondre selon votre niveau d'accord ou de désaccord avec les énoncés suivants: [Je me sens chez moi au Québec.]",
    "feel_queb" = "23 - Veuillez répondre selon votre niveau d'accord ou de désaccord avec les énoncés suivants: [Je me sens Québécois(e).]",
    "forget_orig" = "23 - Veuillez répondre selon votre niveau d'accord ou de désaccord avec les énoncés suivants: [Il faut faire oublier ses origines pour se faire accepter au Québec.]",
    "percep_qc" = "23 - Veuillez répondre selon votre niveau d'accord ou de désaccord avec les énoncés suivants: [On me voit comme un(e) Québécois(e).]",
    "dev_qc" = "23 - Veuillez répondre selon votre niveau d'accord ou de désaccord avec les énoncés suivants: [Je veux devenir Québécois(e).]",
    "agir_qc" = "23 - Veuillez répondre selon votre niveau d'accord ou de désaccord avec les énoncés suivants: [En public, j’essaye d’agir comme un Québécois.]",
    "import_adapt" = "23 - Veuillez répondre selon votre niveau d'accord ou de désaccord avec les énoncés suivants: [Je crois que c’est important de m’adapter à la culture québécoise.]",
    "iso_qc" = "23 - Veuillez répondre selon votre niveau d'accord ou de désaccord avec les énoncés suivants: [Vous vous sentez isolé(e) au Québec.]",
    "future_live" = "24 - Dans le futur, où souhaiteriez vous vivre ?",
    "valeur_fr" = "25 - Dans quelle mesure êtes-vous personnellement d’accord avec les valeurs suivantes? [Le français en tant que langue officielle du Québec]",
    "valeur_dem" = "25 - Dans quelle mesure êtes-vous personnellement d’accord avec les valeurs suivantes? [Le respect de la démocratie]",
    "valeur_equal" = "25 - Dans quelle mesure êtes-vous personnellement d’accord avec les valeurs suivantes? [L’égalité entre les sexes*]",
    "valeur_med14" = "25 - Dans quelle mesure êtes-vous personnellement d’accord avec les valeurs suivantes? [La liberté de consentir ou de refuser des interventions en soins de santé dès l’âge de 14 ans]",
    "valeur_relation" = "25 - Dans quelle mesure êtes-vous personnellement d’accord avec les valeurs suivantes? [La liberté de choisir sont statut de relation amoureuse]",
    "valeur_laic" = "25 - Dans quelle mesure êtes-vous personnellement d’accord avec les valeurs suivantes? [La laïcité au Québec**]",
    "valeur_droitlib" = "25 - Dans quelle mesure êtes-vous personnellement d’accord avec les valeurs suivantes? [Le respect de la Charte des droits et libertés de la personne]",
    "qc_val_fr" = "26 - Selon vous, dans quelles mesures les québécois(es) partagent, en général, les valeurs suivantes? [Le français en tant que langue officielle du Québec]",
    "qc_val_dem" = "26 - Selon vous, dans quelles mesures les québécois(es) partagent, en général, les valeurs suivantes? [Le respect de la démocratie]",
    "qc_val_equal" = "26 - Selon vous, dans quelles mesures les québécois(es) partagent, en général, les valeurs suivantes? [L’égalité entre les sexes*]",
    "qc_val_med14" = "26 - Selon vous, dans quelles mesures les québécois(es) partagent, en général, les valeurs suivantes? [La liberté de consentir ou de refuser des interventions en soins de santé dès l’âge de 14 ans]",
    "qc_val_relation" = "26 - Selon vous, dans quelles mesures les québécois(es) partagent, en général, les valeurs suivantes? [La liberté de choisir sont statut de relation amoureuse]",
    "qc_val_laic" = "26 - Selon vous, dans quelles mesures les québécois(es) partagent, en général, les valeurs suivantes? [La laïcité au Québec**]",
    "qc_val_droitlib" = "26 - Selon vous, dans quelles mesures les québécois(es) partagent, en général, les valeurs suivantes? [Le respect de la Charte des droits et libertés de la personne]",
    "comp_polqc" = "27 - Comment qualifiez vous votre compréhension des enjeux politiques auxquels fait face le Québec?",
    "disc_polqc" = "28 - Durant le dernier mois, à quelle fréquence avez-vous discuté avec une autre personne d'enjeux politiques québécois ?",
    "interp_polqc" = "29 - À quel niveau vous sentez vous interpelé(e) par les enjeux politiques au Québec ?",
    "polqc_CAQ" = "30 - Les questions suivantes concernent les partis politiques québécois et certains évènements récents. [La Coalition Avenir Québec promulgue le communisme]",
    "polqc_PQ" = "30 - Les questions suivantes concernent les partis politiques québécois et certains évènements récents. [Le Parti Québécois est attaché au fédéralisme]" ,
    "polqc_PLQ" = "30 - Les questions suivantes concernent les partis politiques québécois et certains évènements récents. [Le Parti Libéral du Québec milite pour l'indépendance du Québec]",
    "polqc_QS" = "30 - Les questions suivantes concernent les partis politiques québécois et certains évènements récents. [Québec Solidaire promulgue le socialisme démocratique]",
    "polqc_L21" = "30 - Les questions suivantes concernent les partis politiques québécois et certains évènements récents. [La loi 21 concerne les droits des personnes LGBTQ+]",
    "polqc_L96" = "30 - Les questions suivantes concernent les partis politiques québécois et certains évènements récents. [La loi 96 concerne l'utilisation de la langue française]",
    "pm_qc" = "31 - À quel parti est associé le premier ministre actuel du Québec ?",
    "maj_AN" = "32 - Quel parti occupe la majorité des sièges de l'Assemblée nationale ?",
    "vote_age" = "33 - Quel est l'âge légal pour voter au Québec ?",
    "act_conv" = "34 - Durant les douze derniers mois, avez-vous réalisé l'une des actions suivantes ? [J'ai essayé de convaincre quelqu'un de changer d'opinion politique]",
    "act_inf" = "34 - Durant les douze derniers mois, avez-vous réalisé l'une des actions suivantes ? [J'ai essayé d'influencer le vote d'autres personnes]",
    "act_opinion_pol" = "34 - Durant les douze derniers mois, avez-vous réalisé l'une des actions suivantes ? [J'ai émis mon opinion politique en public ou sur les réseaux sociaux]",
    "act_disc_pol" = "34 - Durant les douze derniers mois, avez-vous réalisé l'une des actions suivantes ? [J'ai discuté de sujets politiques en public ou sur les réseaux sociaux]",
    "act_contact_pol" = "34 - Durant les douze derniers mois, avez-vous réalisé l'une des actions suivantes ? [J'ai contacté un politicien ou un membre du gouvernement]",
    "act_workpol" = "34 - Durant les douze derniers mois, avez-vous réalisé l'une des actions suivantes ? [J'ai travaillé pour un parti ou un groupe activiste]",
    "act_symb" = "34 - Durant les douze derniers mois, avez-vous réalisé l'une des actions suivantes ? [J'ai affiché un symbole politique (sticker, badge, vêtement, etc.)]",
    "act_pet" = "34 - Durant les douze derniers mois, avez-vous réalisé l'une des actions suivantes ? [J'ai signé une pétition]",
    "act_manif" = "34 - Durant les douze derniers mois, avez-vous réalisé l'une des actions suivantes ? [J'ai pris part à une manifestation publique et licite]",
    "act_collect_sign" = "34 - Durant les douze derniers mois, avez-vous réalisé l'une des actions suivantes ? [J'ai collecté des signatures pour promouvoir une pétition]",
    "fr_lire" = "35 - Communiquer en français implique plusieurs dimensions comme la lecture, l'écoute, l'écriture et la parole. Veuillez évaluer votre compétence à employer le français en répondant aux énoncés suivants. [Je peux lire et comprendre les points principaux d'un article de presse dont le sujet m'est familier]",
    "fr_conv" = "35 - Communiquer en français implique plusieurs dimensions comme la lecture, l'écoute, l'écriture et la parole. Veuillez évaluer votre compétence à employer le français en répondant aux énoncés suivants. [Dans une conversation, je peux parler de sujets familiers et exprimer mon opinion]",
    "fr_ecrire" = "35 - Communiquer en français implique plusieurs dimensions comme la lecture, l'écoute, l'écriture et la parole. Veuillez évaluer votre compétence à employer le français en répondant aux énoncés suivants. [Je peux écrire des lettres à propos de mes expériences, de mes sentiments et dévènements]",
    "fr_radiotele" = "35 - Communiquer en français implique plusieurs dimensions comme la lecture, l'écoute, l'écriture et la parole. Veuillez évaluer votre compétence à employer le français en répondant aux énoncés suivants. [Je comprends les points principaux d'une émission télévisée ou radiodiffusée à propos d'un sujet familier]",
    "comment" = "Avez-vous des commentaires que vous aimeriez partager avec nous ?",
    "int_quali" = "Dans la cadre de cette même recherche, avez vous un intérêt à participer à une séance de discussion? Si oui,  veuillez partager votre courriel.",
    "continuant" = "En continuant :"
    )
```

```{r}
#Exclusion des répondants établis au Québec depuis moins d'un an

an_SOC3380 <- an_SOC3380[an_SOC3380$year_qc != "Moins d'un an", ]
```

```{r}
#Exclusion des variables inutiles à l'enquête et transformation des variables utiles en variables factorielles

an_SOC3380 <-
  an_SOC3380 %>%
  mutate_if(sapply(an_SOC3380, is.character), as.factor)%>%
    select(!c(Horodateur, continuant, comment, int_quali))

```

```{r}

#Multiplication des observations par tirage d'échantillons aléatoires

set.seed(1234)
Sample_1 <- slice_sample(an_SOC3380, n = 15)
set.seed(1235)
Sample_2 <- slice_sample(an_SOC3380, n = 15)
set.seed(1236)
Sample_3 <- slice_sample(an_SOC3380, n = 15)
set.seed(1237)
Sample_4 <- slice_sample(an_SOC3380, n = 15)
set.seed(1238)
Sample_5 <- slice_sample(an_SOC3380, n = 15)
set.seed(1239)
Sample_6 <- slice_sample(an_SOC3380, n = 15)

an_SOC3380 <- bind_rows(an_SOC3380, Sample_1, Sample_2, Sample_3, Sample_4, Sample_5, Sample_6)
```

```{r}
#correction de variables

an_SOC3380 <-
  an_SOC3380 %>% 
  mutate(born_country = recode(born_country, `Ivory Coast (Cote d'Ivoire)` = "Côte D'Ivoire"),
         country_15 = recode(born_country, `Ivory Coast (Cote d'Ivoire)` = "Côte D'Ivoire"),
         born_country = recode(born_country, `Iran` = "Iran, Islamic Republic of"),
         country_15 = recode(country_15, `Iran` = "Iran, Islamic Republic of"))

an_SOC3380 <-
  an_SOC3380 %>%
  mutate(born_country = recode(born_country, `Ivory Coast (Cote d'Ivoire)` = "Côte D'Ivoire"),
         country_15 = recode(born_country, `Ivory Coast (Cote d'Ivoire)` = "Côte D'Ivoire"),
         born_country = recode(born_country, `Iran` = "Iran, Islamic Republic of"),
         country_15 = recode(country_15, `Iran` = "Iran, Islamic Republic of"),
         genre = recode(genre, Fluide = "Autres"),
         genre = recode(genre, `Je ne souhaite pas répondre` = "Autres"),
         lang_enf = case_when(
           lang_enf == "Français" ~ "Français",
           lang_enf == "Anglais" ~ "Anglais",
           TRUE ~ "Autres"),
         lang_home = case_when(
           lang_home == "Français" ~ "Français",
           lang_home == "Anglais" ~ "Anglais",
           TRUE ~ "Autres"))

an_SOC3380 <- left_join(an_SOC3380,Pays, by = "born_country")

```

```{r}
#numérisation de variables

an_SOC3380 <-
  an_SOC3380 %>%
  mutate(Facebook = if_else(str_detect(rs_util, "Facebook", negate = FALSE), "Oui", "Non"),
         Twitter = if_else(str_detect(rs_util, "Twitter", negate = FALSE), "Oui", "Non"),
         Youtube = if_else(str_detect(rs_util, "Youtube", negate = FALSE), "Oui", "Non"),
         Instagram = if_else(str_detect(rs_util, "Instagram", negate = FALSE), "Oui", "Non"),
         Messenger = if_else(str_detect(rs_util, "Messenger", negate = FALSE), "Oui", "Non"),
         WhatsApp = if_else(str_detect(rs_util, "WhatsApp", negate = FALSE), "Oui", "Non"),
         WeChat = if_else(str_detect(rs_util, "WeChat", negate = FALSE), "Oui", "Non"),
         TikTok = if_else(str_detect(rs_util, "TikTok", negate = FALSE), "Oui", "Non"),
         Douyin = if_else(str_detect(rs_util, "Douyin", negate = FALSE), "Oui", "Non"),
         Telegram = if_else(str_detect(rs_util, "Telegram", negate = FALSE), "Oui", "Non"),
         QQ = if_else(str_detect(rs_util, "QQ", negate = FALSE), "Oui", "Non"),
         Weibo = if_else(str_detect(rs_util, "Weibo", negate = FALSE), "Oui", "Non"),
         Qzone = if_else(str_detect(rs_util, "Qzone", negate = FALSE), "Oui", "Non"),
         Snapchat = if_else(str_detect(rs_util, "Snapchat", negate = FALSE), "Oui", "Non"),
         Kuaishou = if_else(str_detect(rs_util, "Kuaishou", negate = FALSE), "Oui", "Non"),
         Pinterest = if_else(str_detect(rs_util, "Pinterest", negate = FALSE), "Oui", "Non"),
         Reddit = if_else(str_detect(rs_util, "Reddit", negate = FALSE), "Oui", "Non"),
         LinkedIn = if_else(str_detect(rs_util, "LinkedIn", negate = FALSE), "Oui", "Non"),
         Quora = if_else(str_detect(rs_util, "Quora", negate = FALSE), "Oui", "Non"),
         Skype = if_else(str_detect(rs_util, "Skype", negate = FALSE), "Oui", "Non"),
         Tieba = if_else(str_detect(rs_util, "Tieba", negate = FALSE), "Oui", "Non"),
         Viber = if_else(str_detect(rs_util, "Viber", negate = FALSE), "Oui", "Non"),
         Teams = if_else(str_detect(rs_util, "Teams", negate = FALSE), "Oui", "Non"),
         imo = if_else(str_detect(rs_util, "imo", negate = FALSE), "Oui", "Non"),
         Line = if_else(str_detect(rs_util, "Line", negate = FALSE), "Oui", "Non"),
         Piscart = if_else(str_detect(rs_util, "Piscart", negate = FALSE), "Oui", "Non"),
         Likee = if_else(str_detect(rs_util, "Likee", negate = FALSE), "Oui", "Non"),
         Discord = if_else(str_detect(rs_util, "Discord", negate = FALSE), "Oui", "Non"),
         Twitch = if_else(str_detect(rs_util, "Twitch", negate = FALSE), "Oui", "Non"),
         `Stack Exchange` = if_else(str_detect(rs_util, "Stack Exchange", negate = FALSE), "Oui", "Non"),
         Signal = if_else(str_detect(rs_util, "Signal", negate = FALSE), "Oui", "Non")) %>%
  mutate_at(
    vars(one_of("Facebook",
                "Twitter",
                "Youtube",
                "Instagram",
                "Messenger",
                "WhatsApp",
                "WeChat",
                "TikTok",
                "Douyin",
                "Telegram",
                "QQ",
                "Weibo",
                "Qzone",
                "Snapchat",
                "Kuaishou",
                "Pinterest",
                "Reddit",
                "Facebook",
                "LinkedIn",
                "Quora",
                "Skype",
                "Tieba",
                "Viber",
                "Teams",
                "imo",
                "Line",
                "Piscart",
                "Discord",
                "Piscart",
                "Likee",
                "Twitch",
                "Signal",
                "Stack Exchange"
                )),
    funs(case_when(
      . == "Oui" ~ 1,
      . == "Non" ~ 0)
      )) %>%
  rowwise()%>%
  mutate(nbr_rs = sum(Facebook,
                      Twitter,
                      Youtube,
                      Instagram,
                      Messenger,
                      WhatsApp,
                      WeChat,
                      TikTok,
                      Douyin,
                      Telegram,
                      QQ,
                      Weibo,
                      Qzone,
                      Snapchat, 
                      Kuaishou,
                      Kuaishou,
                      Reddit,
                      Facebook,
                      LinkedIn,
                      Quora,
                      Skype,
                      Tieba,
                      imo,
                      Line,
                      Piscart,
                      Discord,
                      Piscart,
                      Twitch,
                      Signal,
                      `Stack Exchange`,
                      na.rm = TRUE))

an_SOC3380 <-
  an_SOC3380 %>%
  mutate(region_n = case_when(
    region == "Europe" ~ 1,
    region == "South/Latin America" ~ 2,
    region == "Arab States" ~ 3,
    region == "North America" ~ 4,
    region == "Asia & Pacific" ~ 5,
    region == "Africa" ~ 6,
    region == "Middle east" ~ 7
  ))
```

```{r}
#préparation des indices

accord_lvl <- c("Pas du tout d'accord", 
                "Plutôt pas d'accord", 
                "Plutôt d'accord", 
                "D'accord", 
                "Tout à fait d'accord")

INV_accord_LVL <- c("Tout à fait d'accord", 
                    "D'accord", 
                    "Plutôt d'accord", 
                    "Plutôt pas d'accord", 
                    "Pas du tout d'accord")

mesure_lvl <- c("Pas du tout", 
                "Dans une certaine mesure", 
                "Dans une grande mesure")

IND_LVL <- c("Faible", 
             "Modéré", 
             "Fort")

#variables indépendantes

##rs_disc_diff

an_SOC3380$rs_disc_diff <- recode(an_SOC3380$rs_disc_diff, "Pas d'accord" = "Plutôt pas d'accord")
an_SOC3380$rs_disc_diff <- factor(an_SOC3380$rs_disc_diff, levels = accord_lvl, ordered = TRUE)
an_SOC3380$rs_disc_diff <- as.numeric(an_SOC3380$rs_disc_diff, na.rm = TRUE)

##rs_renc_diff

an_SOC3380$rs_renc_diff <- recode(an_SOC3380$rs_renc_diff, "Pas d'accord" = "Plutôt pas d'accord")
an_SOC3380$rs_renc_diff <- factor(an_SOC3380$rs_renc_diff, levels = accord_lvl, ordered = TRUE)
an_SOC3380$rs_renc_diff <- as.numeric(an_SOC3380$rs_renc_diff, na.rm = TRUE)

##disc_maj_same

an_SOC3380$disc_maj_same <- factor(an_SOC3380$disc_maj_same, levels = accord_lvl, ordered = TRUE)
an_SOC3380$disc_maj_same <- as.numeric(an_SOC3380$disc_maj_same, na.rm = TRUE)

##disc_maj_migr

an_SOC3380$disc_maj_migr <- factor(an_SOC3380$disc_maj_migr, levels = accord_lvl, ordered = TRUE)
an_SOC3380$disc_maj_migr <- as.numeric(an_SOC3380$disc_maj_migr, na.rm = TRUE)

##disc_maj_qc

an_SOC3380$disc_maj_qc <- factor(an_SOC3380$disc_maj_qc, levels = accord_lvl, ordered = TRUE)
an_SOC3380$disc_maj_qc <- as.numeric(an_SOC3380$disc_maj_qc, na.rm = TRUE)

## online_memb_migr

an_SOC3380$online_memb_migr <- factor(an_SOC3380$online_memb_migr, levels = c("Oui", "Non"), ordered = FALSE)

#variables dépendantes

##sentim_app

an_SOC3380$sentim_app <- recode(an_SOC3380$sentim_app, "Pas d'accord" = "Plutôt pas d'accord")
an_SOC3380$sentim_app <- factor(an_SOC3380$sentim_app, levels = accord_lvl, ordered = TRUE)
an_SOC3380$sentim_app <- as.numeric(an_SOC3380$sentim_app, na.rm = TRUE)

##chez_moi_qc

an_SOC3380$chez_moi_qc <- recode(an_SOC3380$chez_moi_qc, "Pas d'accord" = "Plutôt pas d'accord")
an_SOC3380$chez_moi_qc <- factor(an_SOC3380$chez_moi_qc, levels = accord_lvl, ordered = TRUE)
an_SOC3380$chez_moi_qc <- as.numeric(an_SOC3380$chez_moi_qc, na.rm = TRUE)

##feel_queb

an_SOC3380$feel_queb <- recode(an_SOC3380$feel_queb, "Pas d'accord" = "Plutôt pas d'accord")
an_SOC3380$feel_queb <- factor(an_SOC3380$feel_queb, levels = accord_lvl, ordered = TRUE)
an_SOC3380$feel_queb <- as.numeric(an_SOC3380$feel_queb, na.rm = TRUE)

##forget_orig

an_SOC3380$forget_orig <- recode(an_SOC3380$forget_orig, "Pas d'accord" = "Plutôt pas d'accord")
an_SOC3380$forget_orig <- factor(an_SOC3380$forget_orig, levels = INV_accord_LVL, ordered = TRUE)
an_SOC3380$forget_orig <- as.numeric(an_SOC3380$forget_orig, na.rm = TRUE)

##percep_qc

an_SOC3380$percep_qc <- recode(an_SOC3380$percep_qc, "Pas d'accord" = "Plutôt pas d'accord")
an_SOC3380$percep_qc <- factor(an_SOC3380$percep_qc, levels = accord_lvl, ordered = TRUE)
an_SOC3380$percep_qc <- as.numeric(an_SOC3380$percep_qc, na.rm = TRUE)

##dev_qc

an_SOC3380$dev_qc <- recode(an_SOC3380$dev_qc, "Pas d'accord" = "Plutôt pas d'accord")
an_SOC3380$dev_qc <- factor(an_SOC3380$dev_qc, levels = accord_lvl, ordered = TRUE)
an_SOC3380$dev_qc <- as.numeric(an_SOC3380$dev_qc, na.rm = TRUE)

##agir_qc

an_SOC3380$agir_qc <- recode(an_SOC3380$agir_qc, "Pas d'accord" = "Plutôt pas d'accord")
an_SOC3380$agir_qc <- factor(an_SOC3380$agir_qc, levels = accord_lvl, ordered = TRUE)
an_SOC3380$agir_qc <- as.numeric(an_SOC3380$agir_qc, na.rm = TRUE)

##import_adapt

an_SOC3380$import_adapt <- recode(an_SOC3380$import_adapt, "Pas d'accord" = "Plutôt pas d'accord")
an_SOC3380$import_adapt <- factor(an_SOC3380$import_adapt, levels = accord_lvl, ordered = TRUE)
an_SOC3380$import_adapt <- as.numeric(an_SOC3380$import_adapt, na.rm = TRUE)

##iso_qc

an_SOC3380$iso_qc <- recode(an_SOC3380$iso_qc, "Pas d'accord" = "Plutôt pas d'accord")
an_SOC3380$iso_qc <- factor(an_SOC3380$iso_qc, levels = INV_accord_LVL, ordered = TRUE)
an_SOC3380$iso_qc <- as.numeric(an_SOC3380$iso_qc, na.rm = TRUE)

##valeur_fr

an_SOC3380$valeur_fr <- factor(an_SOC3380$valeur_fr, levels = mesure_lvl, ordered = TRUE)
an_SOC3380$valeur_fr <- as.numeric(an_SOC3380$valeur_fr, na.rm = TRUE)

##valeur_dem

an_SOC3380$valeur_dem <- factor(an_SOC3380$valeur_dem, levels = mesure_lvl, ordered = TRUE)
an_SOC3380$valeur_dem <- as.numeric(an_SOC3380$valeur_dem, na.rm = TRUE)

##valeur_equal

an_SOC3380$valeur_equal <- factor(an_SOC3380$valeur_equal, levels = mesure_lvl, ordered = TRUE)
an_SOC3380$valeur_equal <- as.numeric(an_SOC3380$valeur_equal, na.rm = TRUE)

##valeur_med14

an_SOC3380$valeur_med14 <- factor(an_SOC3380$valeur_med14, levels = mesure_lvl, ordered = TRUE)
an_SOC3380$valeur_med14 <- as.numeric(an_SOC3380$valeur_med14, na.rm = TRUE)

##valeur_relation

an_SOC3380$valeur_relation <- factor(an_SOC3380$valeur_relation, levels = mesure_lvl, ordered = TRUE)
an_SOC3380$valeur_relation <- as.numeric(an_SOC3380$valeur_relation, na.rm = TRUE)

##valeur_laic

an_SOC3380$valeur_laic <- factor(an_SOC3380$valeur_laic, levels = mesure_lvl, ordered = TRUE)
an_SOC3380$valeur_laic <- as.numeric(an_SOC3380$valeur_laic, na.rm = TRUE)

##valeur_droitlib

an_SOC3380$valeur_droitlib <- factor(an_SOC3380$valeur_droitlib, levels = mesure_lvl, ordered = TRUE)
an_SOC3380$valeur_droitlib <- as.numeric(an_SOC3380$valeur_droitlib, na.rm = TRUE)

##comp_polqc

an_SOC3380$comp_polqc <- factor(an_SOC3380$comp_polqc, levels = c("Pas bonne du tout", 
                                                                  "Pas bonne", 
                                                                  "Plutôt bonne", 
                                                                  "Bonne", 
                                                                  "Très bonne"), ordered = TRUE)

an_SOC3380$comp_polqc <- as.numeric(an_SOC3380$comp_polqc, na.rm = TRUE)

##disc_polqc

an_SOC3380$disc_polqc <- recode(an_SOC3380$disc_polqc, "Rarement" = "Quelques fois")

an_SOC3380$disc_polqc <- factor(an_SOC3380$disc_polqc, levels = c("Jamais", 
                                                                  "Quelques fois", 
                                                                  "Plusieurs fois", 
                                                                  "Très souvent", 
                                                                  "Tout le temps"), ordered = TRUE)

an_SOC3380$disc_polqc <- as.numeric(an_SOC3380$disc_polqc, na.rm = TRUE)

##interp_polqc

an_SOC3380$interp_polqc <- factor(an_SOC3380$interp_polqc, levels = c("Pas interpelé(e)", 
                                                                      "Pas vraiment interpelé(e)", 
                                                                      "Plutôt interpelé(e)", 
                                                                      "Interpelé(e)", 
                                                                      "Très interpelé(e)"), ordered = TRUE)

an_SOC3380$interp_polqc <- as.numeric(an_SOC3380$interp_polqc, na.rm = TRUE)

##qc_val_fr

an_SOC3380$qc_val_fr <- factor(an_SOC3380$qc_val_fr, levels = mesure_lvl, ordered = TRUE)
an_SOC3380$qc_val_fr <- as.numeric(an_SOC3380$qc_val_fr, na.rm = TRUE)

##qc_val_dem

an_SOC3380$qc_val_dem <- factor(an_SOC3380$qc_val_dem, levels = mesure_lvl, ordered = TRUE)
an_SOC3380$qc_val_dem <- as.numeric(an_SOC3380$qc_val_dem, na.rm = TRUE)

##qc_val_equal

an_SOC3380$qc_val_equal <- factor(an_SOC3380$qc_val_equal, levels = mesure_lvl, ordered = TRUE)
an_SOC3380$qc_val_equal <- as.numeric(an_SOC3380$qc_val_equal, na.rm = TRUE)

##qc_val_med14

an_SOC3380$qc_val_med14 <- factor(an_SOC3380$qc_val_med14, levels = mesure_lvl, ordered = TRUE)
an_SOC3380$qc_val_med14 <- as.numeric(an_SOC3380$qc_val_med14, na.rm = TRUE)

##qc_val_relation

an_SOC3380$qc_val_relation <- factor(an_SOC3380$qc_val_relation, levels = mesure_lvl, ordered = TRUE)
an_SOC3380$qc_val_relation <- as.numeric(an_SOC3380$qc_val_relation, na.rm = TRUE)

##qc_val_laic

an_SOC3380$qc_val_laic <- factor(an_SOC3380$qc_val_laic, levels = mesure_lvl, ordered = TRUE)
an_SOC3380$qc_val_laic <- as.numeric(an_SOC3380$qc_val_laic, na.rm = TRUE)

##qc_val_droitlib

an_SOC3380$qc_val_droitlib <- factor(an_SOC3380$qc_val_droitlib, levels = mesure_lvl, ordered = TRUE)
an_SOC3380$qc_val_droitlib <- as.numeric(an_SOC3380$qc_val_droitlib, na.rm = TRUE)

# Contrôle

an_SOC3380$rs_time <- factor(an_SOC3380$rs_time, levels = c("Moins d'une heure", 
                                                            "Entre une et trois heures", 
                                                            "Entre trois et cinq heures", 
                                                            "Entre cinq et sept heures", 
                                                            "Plus de sept heures"), ordered = TRUE)

an_SOC3380$rs_time_n <- as.numeric(an_SOC3380$rs_time, na.rm = TRUE)

## âge

an_SOC3380$age <- factor(an_SOC3380$age, levels = c("Moins de 18 ans", 
                                                    "Entre 18 et 25 ans", 
                                                    "Entre 26 et 35 ans", 
                                                    "Entre 36 et 50 ans", 
                                                    "Plus de 50 ans"), ordered = TRUE)

an_SOC3380$age_n <- as.numeric(an_SOC3380$age, na.omit(NA))

v_sentimapp <- c("sentim_app", 
                 "chez_moi_qc", 
                 "feel_queb", 
                 "forget_orig", 
                 "percep_qc", 
                 "dev_qc", 
                 "agir_qc", 
                 "import_adapt", 
                 "iso_qc")

v_NV <- c("valeur_fr", 
          "valeur_dem", 
          "valeur_equal", 
          "valeur_med14", 
          "valeur_relation", 
          "valeur_laic", 
          "valeur_droitlib")

v_rsContact <- c("rs_disc_diff", 
                 "disc_maj_qc", 
                 "rs_renc_diff")

v_percepQC <- c("qc_val_fr", 
                "qc_val_dem", 
                "qc_val_equal", 
                "qc_val_med14", 
                "qc_val_relation", 
                "qc_val_laic", 
                "qc_val_droitlib")

accord_lvl <- c("Pas du tout d'accord", 
                "Plutôt pas d'accord", 
                "Plutôt d'accord", 
                "D'accord", 
                "Tout à fait d'accord")

an_SOC3380$rs_inf <- factor(an_SOC3380$rs_inf, levels = accord_lvl, ordered = TRUE)

an_SOC3380$rs_inf_n <- as.numeric(an_SOC3380$rs_inf, na.omit(NA))

an_SOC3380$ton_migr <- recode(an_SOC3380$ton_migr, "Très positive" = "Positive")

an_SOC3380$ton_migr <- recode(an_SOC3380$ton_migr, "Très négative" = "Négsative")

ton_migr_lvl <- c("Négative", 
                  "Plutôt négative", 
                  "Neutre", 
                  "Plutôt positive", 
                  "Positive")

an_SOC3380$ton_migr <- factor(an_SOC3380$ton_migr, levels = ton_migr_lvl, ordered = TRUE)
an_SOC3380$ton_migr_n <- as.numeric(an_SOC3380$ton_migr, na.omit(NA))

an_SOC3380$year_qc <- factor(an_SOC3380$year_qc, levels = c("1 an", 
                                                            "2 ans", 
                                                            "3 ans", 
                                                            "4 ans", 
                                                            "5 ans", 
                                                            "6 ans", 
                                                            "7 ans", 
                                                            "8 ans", 
                                                            "9 ans", 
                                                            "10 ans", 
                                                            "Plus de 10 ans"), ordered = TRUE)

an_SOC3380$year_qc_n <- as.numeric(an_SOC3380$year_qc, na.omit(NA))

v_score_ExpoDisc <- c("rs_time_n", 
                      "ton_migr_n", 
                      "rs_inf_n")

v_media_accueil <- c("Youtube", 
                     "Instagram", 
                     "TikTok", 
                     "Snapchat", 
                     "Pinterest", 
                     "Piscart", 
                     "Twitch", 
                     "Twitter", 
                     "Reddit", 
                     "Quora", 
                     "Teams", 
                     "Discord", 
                     "Stack Exchange", 
                     "Facebook", 
                     "Messenger", 
                     "Telegram", 
                     "LinkedIn", 
                     "Skype", 
                     "Viber", 
                     "imo", 
                     "Signal")

v_media_diasp <- c("Douyin", 
                   "Kuaishou", 
                   "Likee", 
                   "Tieba", 
                   "WeChat", 
                   "QQ", 
                   "Qzone", 
                   "Line", 
                   "WhatsApp")
```

```{r}
#création des indices

an_SOC3380 <-
  an_SOC3380 %>%
  rowwise() %>%
  mutate(score_sentimapp = (sum(c_across(all_of(v_sentimapp)), na.rm = TRUE) - length(v_sentimapp)) / (4*length(v_sentimapp))) %>%
  mutate(score_sentimapp = if (score_sentimapp < 0) NA_real_ else score_sentimapp) %>%
  ungroup()

an_SOC3380 <- 
  an_SOC3380 %>%
  rowwise() %>%
  mutate(score_rsContact = (sum(c_across(all_of(v_rsContact)), na.rm = TRUE) -length(v_rsContact)) / (4*length(v_rsContact))) %>%
  mutate(score_rsContact = if (score_rsContact < 0)  NA_real_ else score_rsContact) %>%
  ungroup()

an_SOC3380 <- 
  an_SOC3380 %>%
  rowwise() %>%
  mutate(score_NV = (sum(c_across(all_of(v_NV)), na.rm = TRUE) -length(v_NV)) / (2*length(v_NV))) %>%
  mutate(score_NV = if (score_NV < 0) NA_real_ else score_NV) %>%
  ungroup()

an_SOC3380 <- 
  an_SOC3380 %>%
  rowwise() %>%
  mutate(score_percepQC = (sum(c_across(all_of(v_percepQC)), na.rm = TRUE) -length(v_percepQC)) / (2*length(v_percepQC))) %>%
  mutate(score_percepQC = if (score_percepQC < 0) NA_real_ else score_percepQC) %>%
  ungroup()

an_SOC3380 <- 
  an_SOC3380 %>%
  rowwise() %>%
  mutate(score_ExpoDisc = (sum(c_across(all_of(v_score_ExpoDisc)), na.rm = TRUE) - length(v_score_ExpoDisc)) / (4*length(v_score_ExpoDisc))) %>%
  mutate(score_ExpoDisc = if_else(score_ExpoDisc < 0 | score_ExpoDisc == 0, NA_real_, score_ExpoDisc)) %>%
  ungroup()

an_SOC3380 <-
  an_SOC3380 %>%
  rowwise() %>%
  mutate(score_media_accueil = (sum(c_across(all_of(v_media_accueil)), na.rm = TRUE)) / nbr_rs)

an_SOC3380 <- 
  an_SOC3380 %>%
  rowwise() %>%
  mutate(score_media_diasp = (sum(c_across(all_of(v_media_diasp)), na.rm = TRUE)) / nbr_rs)
```

```{r}
an_SOC3380 <- an_SOC3380 %>%
  mutate(rs_partage_medias = if_else(str_detect(rs_util, "Youtube|Instagram|TikTok|Douyin|Snapchat|Kuaishou|Pinterest|Piscart|Likee|Twitch", negate = FALSE), 1, 0),
         rs_blogues = if_else(str_detect(rs_util, "Twitter|Weibo", negate = FALSE), 1, 0),
         rs_commun_online = if_else(str_detect(rs_util, "Reddit|Quora|Tieba|Teams|Discord|Stack Exchange", negate = FALSE), 1, 0),
         rs_SRS = if_else(str_detect(rs_util, "Facebook|Messenger|WhatsApp|WeChat|Telegram|QQ|Qzone|LinledIn|Skype|Viber|imo|Line|Signal", negate = FALSE), 1, 0))

an_SOC3380 <-
  an_SOC3380 %>%
  mutate(rs_blogues_f = case_when(
    rs_blogues == 0 ~ "Non",
    rs_blogues == 1 ~ "Oui"
  ))

an_SOC3380 <-
  an_SOC3380 %>%
  mutate(rs_commun_online_f = case_when(
    rs_commun_online == 0 ~ "Non",
    rs_commun_online == 1 ~ "Oui"
  ))

an_SOC3380 <-
  an_SOC3380 %>%
  mutate(rs_SRS_f = case_when(
    rs_SRS == 0 ~ "Non",
    rs_SRS == 1 ~ "Oui"
  ))

an_SOC3380 <-
  an_SOC3380 %>%
  mutate(rs_partage_medias_f = case_when(
    rs_partage_medias == 0 ~ "Non",
    rs_partage_medias == 1 ~ "Oui"
  ))

an_SOC3380 <- an_SOC3380 %>%
  mutate(rs_home = if_else(str_detect(rs_util,"Youtube|Instagram|TikTok|Snapchat|Pinterest|Piscart|Twitch|Twitter|Reddit|Quora|Teams|Discord|Stack Exchange|Facebook|Messenger|Telegram|LinkedIn|Skype|Viber|imo|Signal", negate = FALSE), 1, 0))

an_SOC3380 <- an_SOC3380 %>%
  mutate(rs_diasp = if_else(str_detect(rs_util, "Signal|Douyin|Kuaishou|Likee|Tieba|WeChat|QQ|Qzone|Line|WhatsApp|Messenger|Skype", negate = FALSE), 1, 0))

an_SOC3380 <-
  an_SOC3380 %>%
  mutate(rs_home_f = case_when(
    rs_home == 0 ~ "Non",
    rs_home == 1 ~ "Oui"
  ))

an_SOC3380 <-
  an_SOC3380 %>%
  mutate(rs_diasp_f = case_when(
    rs_diasp == 0 ~ "Non",
    rs_diasp == 1 ~ "Oui"
  ))

an_SOC3380 <-
  an_SOC3380 %>%
  mutate(year_qc_f = case_when(
    year_qc_n < 6 ~ "1 à 5 ans",
    year_qc_n == 6| year_qc_n == 7|year_qc_n == 8|year_qc_n == 9 ~ "6 à 9 ans",
    year_qc_n > 9 ~ "dix ans et plus"
  ))
```

```{r}
write.csv(an_SOC3380, file = "an_SOC3380")
```

### 2023-05-03 Expérimentations (Théa)/1. SOC3380 (data)/2. Analyse descriptive (code).Rmd
Expérimentations sur les données du cours SOC3380
Je me suis amusée à appliquer ce que j'ai appris sur les données de notre petite enquête dans le cours SOC3380.

---
title: "Analyse descriptive"
author: "Théa"
date: "2023-03-17"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

```{r}
#Chargement des packages

library(dplyr)
library(tidyverse)
library(summarytools)
library(readr)
library(questionr)
library(rcompanion)
```

```{r}
#Chargement de la base de données scriptée

an_SOC3380 <- read_csv("an_SOC3380")
```

# Un test de régression logistique appliqué à d'autres données

```{r}
#VD = rs_commun_online (les personnes parcourant des forums)
#VI = score_rsContact (l'indice de contact interculturel en ligne)
model1 <- glm(formula = rs_commun_online ~ score_rsContact, family = "binomial", data = an_SOC3380)
summary(model1)

ggplot(an_SOC3380 %>% filter(!is.na(rs_commun_online), !is.na(score_rsContact)), aes(x = score_rsContact, y = rs_commun_online)) + 
  geom_point() +
  stat_smooth(method = "glm", color = "blue", se = FALSE,
    method.args = list(family = binomial))
```

# Une analyse ANOVA appliquée à d'autres données

```{r}
#VI = rs_blogues_f (utilisation de blogs)
#VD = score_rsContact
#VC = year_qc (nombre d'années passées au Québec) & diplome (niveau d'éducation)
an_SOC3380 %>%
  filter(diplome != "Je ne souhaite pas répondre") %>%
  select(rs_blogues_f, score_rsContact, year_qc, diplome) %>%
  drop_na() %>%
  aov(score_rsContact ~ rs_blogues_f + year_qc + diplome, data = .) %>%
  summary()

ggplot(an_SOC3380 %>% filter(!is.na(score_rsContact), diplome != "Je ne souhaite pas répondre"), aes(x = rs_blogues_f, y = score_rsContact, fill = diplome)) + 
  geom_boxplot(na.rm = TRUE) +
  labs(x = "Utilisation de blogs", y = "Indice de contact interculturel", fill = "Diplôme")+
  facet_wrap(~ diplome, scales = "free_x") +
  scale_fill_brewer(name = "Diplôme", palette = "Set1")
```

### 2023-05-03 Expérimentations (Théa)/2. Expérimentations (Bookdown)/Bookdown/src/clean-up-database.R
Suppression des premières expériences
Suppression des fichiers sur lesquels j'ai expérimenté lundi, mardi et mercredi.

@@ -1,226 +1,0 @@
-#!/usr/bin/R
-
-# Recode columns
-# Assumes that questionnaire is imported when exported as CSV in English
-# using read_csv and named immigr_init
-immigr_init <-
-  immigr_init %>%
-  dplyr::rename(
-    "age" = "1 - Quel âge avez vous ?",
-    "year_qc" = "2 - Depuis combien d'années habitez vous au Québec ?",
-    "born_country" = "3a - Dans quel pays êtes vous né ?",
-    "country_15" = "3b - Dans quel pays avez vous vécu jusqu'à vos 15 ans ?",
-    "genre" = "4 - A quel genre vous identifiez vous ?",
-    "statut" = "5 - Quel est votre statut actuel au Canada ?",
-    "etat_civil" = "6 - Quel est votre état civil ?",
-    "diplome" = "7 - Quel est votre niveau de diplôme le plus élevé ?",
-    "revenu" = "8 - Dans quelle tranche se situe votre revenu actuel ?",
-    "lang_enf" = "9 - Quelle est la première langue que vous avez apprise à la maison pendant votre enfance et que vous comprenez encore aujourd'hui ?",
-    "lang_home" = "10 - Parmi ces langues, laquelle parlez vous le plus souvent à la maison ? Indiquez plus d'une langue seulement si elles sont parlées aussi souvent l'une que l'autre à la maison.",
-    "lang_home_autre" = "11 - Parlez vous d'autres langues à la maison ?",
-    "lang_trav" = "12 -  Parmi ces langues, laquelle parlez vous le plus souvent au travail ? Indiquez plus d'une langue seulement si elles sont parlées aussi souvent l'une que l'autre au travail.",
-    "lang_trav_autre" = "13 - Parlez vous d'autres langues au travail ?",
-    "rs_util" = "14 - Quel(s) réseaux sociaux utilisez vous régulièrement ?",
-    "rs_time" = "15 - Combien de temps passez vous par jour sur les réseaux sociaux ?",
-    "rs_disc_same" = "16 - Pour quelles raisons allez-vous sur les réseaux sociaux ? [Pour discuter avec des personnes qui proviennent de mon pays d'origine]",
-    "rs_disc_diff" = "16 - Pour quelles raisons allez-vous sur les réseaux sociaux ? [Pour discuter avec des personnes qui ne proviennent pas de mon pays d'origine]",
-    "rs_hab" = "16 - Pour quelles raisons allez-vous sur les réseaux sociaux ? [Par habitude]",
[…]

### 2023-05-03 Expérimentations (Théa)/2. Expérimentations (Bookdown)/Bookdown/src/load-packages.R

@@ -1,13 +1,0 @@
-#!/usr/bin/R
-
-library(dplyr)
-library(car)
-library(plyr)
-library(tidyverse)
-library(summarytools)
-library(readr)
-library(gplots)
-library(graphics)
-library(readxl)
-library(pander)
-library(extrafont)

### 2023-05-03 Expérimentations (Théa)/2. Expérimentations (Bookdown)/Bookdown/src/print-index.R

@@ -1,100 +1,0 @@
-#!/usr/bin/R
-
-index_names <- names(immigr_init)
-index_desc <- c(
- "1 - Quel âge avez vous ?",
- "2 - Depuis combien d'années habitez vous au Québec ?",
- "3a - Dans quel pays êtes vous né ?",
- "3b - Dans quel pays avez vous vécu jusqu'à vos 15 ans ?",
- "4 - A quel genre vous identifiez vous ?",
- "5 - Quel est votre statut actuel au Canada ?",
- "6 - Quel est votre état civil ?",
- "7 - Quel est votre niveau de diplôme le plus élevé ?",
- "8 - Dans quelle tranche se situe votre revenu actuel ?",
- "9 - Quelle est la première langue que vous avez apprise à la maison pendant votre enfance et que vous comprenez encore aujourd'hui ?",
- "10 - Parmi ces langues, laquelle parlez vous le plus souvent à la maison ?\n Indiquez plus d'une langue seulement si elles sont parlées aussi souvent l'une que l'autre à la maison.",
- "11 - Parlez vous d'autres langues à la maison ?",
- "12 - Parmi ces langues, laquelle parlez vous le plus souvent au travail ? Indiquez plus d'une langue seulement si elles sont parlées aussi souvent l'une que l'autre au travail.",
- "13 - Parlez vous d'autres langues au travail ?",
- "14 - Quel(s) réseaux sociaux utilisez vous régulièrement ?",
- "15 - Combien de temps passez vous par jour sur les réseaux sociaux ?",
- "16 - Pour quelles raisons allez-vous sur les réseaux sociaux ? [Pour discuter avec des personnes qui proviennent de mon pays d'origine]",
- "16 - Pour quelles raisons allez-vous sur les réseaux sociaux ? [Pour discuter avec des personnes qui ne proviennent pas de mon pays d'origine]",
- "16 - Pour quelles raisons allez-vous sur les réseaux sociaux ? [Par habitude]",
- "16 - Pour quelles raisons allez-vous sur les réseaux sociaux ? [Pour me divertir]",
- "16 - Pour quelles raisons allez-vous sur les réseaux sociaux ? [Pour passer le temps]",
- "16 - Pour quelles raisons allez-vous sur les réseaux sociaux ? [Pour m'informer]",
- "16 - Pour quelles raisons allez-vous sur les réseaux sociaux ? [Pour rencontrer des personnes de mon pays d'origine]",
[…]

### 2023-05-03 Tests/Tests.Rmd
Tests de régression logistique et d'analyse ANOVA
Je me suis amusée avec une autre base de données.

---
title: "Tests"
author: "Théa Carlier"
date: "2023-05-03"
output:
  html_document: default
  pdf_document: default
---

```{r}

# Chargement des librairies

library(readr)
library(dplyr)
library(tidyverse)
library(ggthemes)
```

```{r}

# Chargement de la base de données

an_test <- read_csv("an_test.csv")
```

# Analyse par régression logistique de la relation entre le niveau d'éducation et le revenu net

```{r}

# Transformation de la variable indépendante "Niveau d'éducation"

an_test <-
  an_test %>%
  filter(!is.na(CFATINC), !is.na(HLEV2G)) %>%
  mutate(HLEV2G_2 = case_when(
    HLEV2G == 1 ~ 0,
    HLEV2G == 2 ~ 0,
    HLEV2G == 3 ~ 0,
    HLEV2G == 4 ~ 1,
    HLEV2G == 6 ~ NA,
    HLEV2G == 9 ~ NA
  ))
```

```{r}

# Création du modèle de régression logistique

lrm1 <- glm(formula = HLEV2G_2 ~ CFATINC, family = "binomial", data = an_test)

# Affichage du modèle

summary(lrm1)
```

```{r}

# Visualisation du modèle

ggplot(an_test %>% filter(!is.na(HLEV2G_2)), aes(x = CFATINC, y = HLEV2G_2)) +
  geom_point() +
  stat_smooth(method = "glm", color = "red", se = FALSE,
              method.args = list(family = binomial)) +
  labs(
    title = "Salaire net selon le niveau d'éducation", 
    x = "Salaire net", 
    y = "Diplôme universitaire (0 = non ; 1 = oui)") +
  theme_few()
```

# Analyse ANOVA de la relation entre l'âge et le revenu net contrôlé par le niveau d'éducation

```{r}

# Transformation de la variable indépendante "Groupe d'âge"

an_test <-
  an_test %>%
  mutate(AGEGP_2 = case_when(
         AGEGP < 5 ~ NA,
         AGEGP >= 5 & AGEGP < 9 ~ 1,
         AGEGP >= 9 & AGEGP < 14 ~ 2,
         AGEGP >= 14 ~ 3
  ))
```

```{r}

# Modèle d'analyse ANOVA

an_test %>%
  select(AGEGP, CFATINC, HLEV2G_2) %>%
  drop_na() %>%
  aov(CFATINC ~ AGEGP + HLEV2G_2, data = .) %>%
  summary()
```

```{r}

# Visualisation du modèle d'analyse ANOVA

ggplot(an_test %>% filter(!is.na(HLEV2G_2), !is.na(AGEGP_2)), aes(x = as.factor(AGEGP_2), y = CFATINC, fill = as.factor(HLEV2G_2))) +
  geom_boxplot(na.rm = TRUE) +
  labs(
    title = "Salaire net selon le groupe d'âge, contrôlé par le niveau d'éducation", 
    x = "Salaire net", 
    y = "Groupe d'âge (1 = [18;34] ; 2 = [35;60] ; 3 = [60;99]") +
  facet_wrap(~ as.factor(HLEV2G_2), scales = "free_x") +
  scale_fill_brewer(name = "Diplôme universitaire (0 = non ; 1 = oui)", palette = "Set1") +
  theme_few()
```

### 2023-05-04 Tests/an_test.Rmd
Quelques paramétrages de test
Je commence à paramétrer ma méthode de travail pour les semaines à venir.

---
title: "df_scripted"
author: "Théa Carlier"
date: "2023-05-03"
output:
  html_document: default
  pdf_document: default
---

```{r}

# Chargement des librairies

library(readr)
library(dplyr)
library(tidyverse)
library(ggthemes)
```

```{r}

# Chargement de la base de données

df_scripted <- read_csv("df_scripted")
```

# Analyse par régression logistique de la relation entre le niveau d'éducation et le revenu net

```{r}

# Création du modèle de régression logistique

lrm1 <- glm(formula = HLEV2G_2 ~ CFATINC, family = "binomial", data = df_scripted)

# Affichage du modèle

summary(lrm1)
```

```{r}

# Visualisation du modèle

ggplot(df_scripted %>% filter(!is.na(HLEV2G_2)), aes(x = CFATINC, y = HLEV2G_2)) +
  geom_point() +
  stat_smooth(method = "glm", color = "red", se = FALSE,
              method.args = list(family = binomial)) +
  labs(
    title = "Salaire net selon le niveau d'éducation", 
    x = "Salaire net", 
    y = "Diplôme universitaire (0 = non ; 1 = oui)") +
  theme_few()
```

# Analyse ANOVA de la relation entre l'âge et le revenu net contrôlé par le niveau d'éducation

```{r}

# Modèle d'analyse ANOVA

df_scripted %>%
  select(AGEGP, CFATINC, HLEV2G_2) %>%
  drop_na() %>%
  aov(CFATINC ~ AGEGP + HLEV2G_2, data = .) %>%
  summary()
```

```{r}

# Visualisation du modèle d'analyse ANOVA

ggplot(df_scripted %>% filter(!is.na(HLEV2G_2), !is.na(AGEGP_2)), aes(x = as.factor(AGEGP_2), y = CFATINC, fill = as.factor(HLEV2G_2))) +
  geom_boxplot(na.rm = TRUE) +
  labs(
    title = "Salaire net selon le groupe d'âge, contrôlé par le niveau d'éducation", 
    x = "Salaire net", 
    y = "Groupe d'âge (1 = [18;34] ; 2 = [35;60] ; 3 = [60;99]") +
  facet_wrap(~ as.factor(HLEV2G_2), scales = "free_x") +
  scale_fill_brewer(name = "Diplôme universitaire (0 = non ; 1 = oui)", palette = "Set1") +
  theme_few()
```

### 2023-05-04 Tests/cr_test.Rmd
Quelques paramétrages de test
Je commence à paramétrer ma méthode de travail pour les semaines à venir.

---
title: "cr_test"
author: "Théa Carlier"
date: "2023-05-04"
output: html_document
---

```{r}

# Chargement des librairies

library(readr)
library(dplyr)
library(tidyverse)
```

```{r}
df_scripted <- read_csv("df_root.csv")
```

```{r}

# Transformation de la variable indépendante "Niveau d'éducation"

df_scripted <-
  df_scripted %>%
  filter(!is.na(CFATINC), !is.na(HLEV2G)) %>%
  mutate(HLEV2G_2 = case_when(
    HLEV2G == 1 ~ 0,
    HLEV2G == 2 ~ 0,
    HLEV2G == 3 ~ 0,
    HLEV2G == 4 ~ 1,
    HLEV2G == 6 ~ NA,
    HLEV2G == 9 ~ NA
  ))
```

```{r}

# Transformation de la variable indépendante "Groupe d'âge"

df_scripted <-
  df_scripted %>%
  mutate(AGEGP_2 = case_when(
         AGEGP < 5 ~ NA,
         AGEGP >= 5 & AGEGP < 9 ~ 1,
         AGEGP >= 9 & AGEGP < 14 ~ 2,
         AGEGP >= 14 ~ 3
  ))
```

```{r}
write.csv(df_scripted, file = "df_scripted")
```

### 2023-05-04 Tests/exercices.Rmd
Quelques exercices sur les boucles + intégration de git dans rstudio

---
title: "exercices"
author: "Théa Carlier"
date: "2023-05-04"
output: html_document
---

```{r}
library(tidyverse)
library(readr)
```

# Exemple

```{r}
df <- tibble(
  a = rnorm(10),
  b = rnorm(10),
  c = rnorm(10),
  d = rnorm(10),
)
```

```{r}
output <- vector("double", ncol(df)) # 1. output (création d'un vecteur)
for (i in seq_along(df)) {           # 2. sequence (dans lequel la boucle va opérer pour chaque élément)
  output[[i]] <- median(df[[i]])     # 3. body 
}
output
```

# Exercice

## 1 : Calculez la moyenne de chaque colonne dans la base de donnée *mtcars*

```{r}
mtcars <- read_csv("mtcars.csv")
```

```{r}
mtcars_2 <-
  mtcars %>%
  select(mpg, disp, hp, drat, wt, qsec)

output_2 <- vector("double", ncol(mtcars_2))
for (i in seq_along(mtcars_2)) {
  output_2[[i]] <- mean(mtcars_2[[i]])
}
output_2
```

## 2 : déterminez le type de chaque colonnes

```{r}
output_3 <- vector("double", ncol(mtcars))
for (i in seq_along(mtcars)) {
  output_3[[i]] <- class(mtcars[[i]])
}
output_3
```

## 3 : calculez le nombre de valeurs uniques dans chaque colonnes

```{r}
output_4 <- vector("double", ncol(mtcars))
for (i in seq_along(mtcars)) {
  output_4[[i]] <- n_distinct(mtcars[[i]])
}
output_4
```

### 2023-05-08 Tests/exercices.Rmd
exercices visualisation 1

@@
 
-# Exercice
+# Exercices (boucles)
 
-## 1 : Calculez la moyenne de chaque colonne dans la base de donnée *mtcars*
+## 1 : calculez la moyenne de chaque colonne dans la base de donnée *mtcars*
 
@@
 ```
+
+# Exemple (visualisation)
+
+## Graphique en nuage de points
+
+```{r}
+ggplot2::mpg
+```
+
+```{r}
+ggplot(data = mpg) +
+  geom_point(mapping = aes(x = displ, y = hwy))
+```
+
+# Exercices (visualisation)
+
+## 1 : run ggplot(data = mpg) que voyez-vous ?
+
+```{r}
+ggplot(data = mpg)
+```
+Je vois un rectangle blanc.
+
+## 2 : Combien de lignes et de colonnes contien mtcars ?
+
+234 lignes et 11 colonnes.
+
+## 3 : à quoi correspond la variable drv ?
+
+```{r}
+?mpg
+```
+
+"the type of drive train, where f = front-wheel drive, r = rear wheel drive, 4 = 4wd"
+
+Traduction : drv nous renseigne si le véhicule est de type traction, propulsion ou 4x4.
+
+## 4 : réalisez un graphique en nuage de points à partir des variables hwy et cwy
+
+```{r}
+ggplot(data = mpg) +
+  geom_point(mapping = aes(x = hwy, y = cyl))
+```
+
+## 5 : Réalisez un graphique a nuage de points avec les variables class et drv. Pourquoi cette visualisation n'est-elle pas pertinente ?
+
+```{r}
+ggplot(data = mpg) +
+  geom_point(mapping = aes(x = class, y = drv))
+```
+
+Parce que les deux variables à étudiées possèdent un niveau de mesure nominal.

### 2023-05-08 Tests/exercices.Rmd
exercices visualisation 2

@@
 
-## Graphique en nuage de points
+## open data
 
@@
 
+## simple scaterplot
+
 ```{r}
@@
   geom_point(mapping = aes(x = displ, y = hwy))
+```
+
+## color
+
+```{r}
+ggplot(data = mpg) +
+  geom_point(mapping = aes(x = displ, y = hwy, color = class))
+```
+
+## size
+
+```{r}
+ggplot(data = mpg) +
+  geom_point(mapping = aes(x = displ, y = hwy, size = class))
+```
+
+## alpha
+
+```{r}
+ggplot(data = mpg) +
+  geom_point(mapping = aes(x = displ, y = hwy, alpha = class))
+```
+
+## shape
+
+```{r}
+ggplot(data = mpg) +
+  geom_point(mapping = aes(x = displ, y = hwy, shape = class))
+
+# ATTENTION, cette esthétique est limitée à six modalités pour la variable de contrôle !
+```
+
+## multiple aesthetics
+
+```{r}
+ggplot(data = mpg) +
+  geom_point(mapping = aes(x = displ, y = hwy), color = "blue", size = 2, shape = 3)
+```
+
+## simple facet_wrap()
+
+```{r}
+ggplot(data = mpg) +
+  geom_point(mapping = aes(x = displ, y = hwy)) +
+  facet_wrap(~ class, nrow = 2)
+```
+
+## simple facet_grid()
+
+```{r}
+ggplot(data = mpg) +
+  geom_point(mapping = aes(x = displ, y = hwy)) +
+  facet_grid(. ~ cyl)
+```
+
+## double facet_grid()
+
+```{r}
+ggplot(data = mpg) +
+  geom_point(mapping = aes(x = displ, y = hwy)) +
+  facet_grid(drv ~ cyl)
+```
+
+## geom_smooth()
+
+```{r}
+ggplot(data = mpg) +
+  geom_smooth(mapping = aes(x = displ, y = hwy))
+```
+
+## multivaried geom_smooth
+
+```{r}
+ggplot(data = mpg) +
+  geom_smooth(mapping = aes(x = displ, y = hwy, linetype = drv))
+```
+
+## multivaried geom_smooth with color
+
+```{r}
+ggplot(data = mpg) +
+  geom_smooth(mapping = aes(x = displ, y = hwy, linetype = drv, color = drv))
+```
+
+## multiple geom_smooth
+
+```{r}
+ggplot(data = mpg) +
+  geom_smooth(mapping = aes(x = displ, y = hwy))
+
+  ggplot(data = mpg) +
+  geom_smooth(mapping = aes(x = displ, y = hwy, linetype = drv))
+
+  ggplot(data = mpg) +
+  geom_smooth(mapping = aes(x = displ, y = hwy, linetype = drv, color = drv), show.legend = FALSE)
+```
+
+## simple geom_smooth and geom point
+
+```{r}
+ggplot(data = mpg) +
+  geom_smooth(mapping = aes(x = displ, y = hwy)) +
+  geom_point(mapping = aes(x = displ, y = hwy, color = class))
+```
+
+## geom_smooth et geom_point multivariés et colorés
+
+```{r}
+ggplot(data = mpg) +
+  geom_smooth(mapping = aes(x = displ, y = hwy, linetype = drv, color = drv),se = FALSE) +
+  geom_point(mapping = aes(x = displ, y = hwy, color = drv))
+```
+
+## specific geom_smooth
+
+```{r}
+ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
+  geom_point(mapping = aes(color = class)) +
+  geom_smooth(data = filter(mpg, class == "subcompact"), se = FALSE, color = "purple")
 ```
@@
 
-## 2 : Combien de lignes et de colonnes contien mtcars ?
+## 2 : combien de lignes et de colonnes contien mtcars ?
 
@@
 
-## 5 : Réalisez un graphique a nuage de points avec les variables class et drv. Pourquoi cette visualisation n'est-elle pas pertinente ?
+## 5 : réalisez un graphique a nuage de points avec les variables class et drv. Pourquoi cette visualisation n'est-elle pas pertinente ?
 
@@
 Parce que les deux variables à étudiées possèdent un niveau de mesure nominal.
+
+## 6 : corrigez le code suivant
+
+```{r}
+ggplot(data = mpg) +
+  geom_point(mapping = aes(x = displ, y = hwy, color = "blue"))
+```
+
+Source de l'erreur : l'argument "color = "blue"" a été inséré en tant que variable de contrôle.
+Correction :
+
+```{r}
+ggplot(data = mpg) +
+  geom_point(mapping = aes(x = displ, y = hwy), color = "blue")
+```
+
+## 7 : dans mpg, quelles variables sont catégorielles et quelles variables sont continues ?
+
+```{r}
+ggplot2::mpg
+```
+
+"manufacturer" = character
+"model" = character
+"displ" = numeric
+"year" = integer
+"cyl" = integer
+"trans" = character
+"drv" = character
+"cty" = integer
+"hwy" = integer
+"fi" = character
+
+character = variable catégorielle alphabétique ou alphanumérique
+integer = variable numérique discrète
+numeric = variable numérique continue
+
+## 8 : assignez une taille, une forme et une couleur à un variable continue. Quelle différence observez-vous avec les variables factorielles ?
+
+```{r}
+ggplot(data = mpg) +
+  geom_point(mapping = aes(x = displ, y = hwy, color = year))
+```
+
+```{r}
+ggplot(data = mpg) +
+  geom_point(mapping = aes(x = displ, y = hwy, shape = year))
+```
+
+```{r}
+ggplot(data = mpg) +
+  geom_point(mapping = aes(x = displ, y = hwy, size = year))
+```
+
+La couleur donne un gradiant, la forme donne un message d'erreur et la taille donne une échelle.
+
+## 9 : que se passerait-il si l'on appliquait plusieurs esthétiques à une variable de contrôle ?
+
+```{r}
+ggplot(data = mpg) +
+  geom_point(mapping = aes(x = displ, y = hwy, color = year, size = year))
+```
+
+Elles se superposent.
+
+## 10 : que fait l'esthétique stroke ? Avec quelles formes fonctionne-t-elle ?
+
+```{r}
+?geom_point
+```
+
+```{r}
+ggplot(data = mpg) +
+  geom_point(mapping = aes(x = displ, y = hwy, color = year), stroke = 5, shape = 23)
+```
+
+Stroke ressemble à size, mais donne un rendu plus gras.
+
+## 11 : que se passe-t-il lorsque vous appliquez une esthétique à autre chose qu'une variable ?
+
+```{r}
+ggplot(data = mpg) +
+  geom_point(mapping = aes(x = displ, y = hwy, color = displ < 5))
+```
+
+Cela donne une couleur différente aux observations où displ < 5.
+
+## 12 : que se passe-t-il si vous facet une variable continue ?
+
+```{r}
+ggplot(data = mpg) +
+  geom_point(mapping = aes(x = displ, y = hwy)) +
+  facet_wrap(~ displ, nrow = 4)
+```
+
+On obtient un graphique multivarié dans lequel chaque valeur de displ devient un graphique.
+
+## 13 : que signifient les cellules vides and le graphique suivant ?
+
+```{r}
+ggplot2::mpg
+```
+
+```{r}
+?mpg
+```
+
+```{r}
+ggplot(data = mpg) +
+  geom_point(mapping = aes(x = displ, y = hwy)) +
+  facet_grid(drv ~ cyl)
+```
+
+Elles signifient qu'aucun véhicule à propulsion ne possède de moteur à 4 ou 5 cylindres.
+
+## 14 : expliquez les deux graphiques suivants et le rôle de "."
+
+```{r}
+ggplot(data = mpg) +
+  geom_point(mapping = aes(x = displ, y = hwy)) +
+  facet_grid(drv ~ .)
+
+ggplot(data = mpg) +
+  geom_point(mapping = aes(x = displ, y = hwy)) +
+  facet_grid(. ~ cyl)
+```
+
+Le point permet de produire des grids avec une variable. Sa position dans le code détermine si facet_grid() doit opérer en ligne ou en colonne.
+
+## 15 : quel avantages apporte le facet par rapport à une esthétique de couleur ?
+
+```{r}
+ggplot(data = mpg) +
+  geom_point(mapping = aes(x = displ, y = hwy)) +
+  facet_wrap(~ class, nrow = 2)
+```
+
+Facet permet d'isoler la relation dans des graphiques séparés en fonction d'une variable de contrôle catégorielle.
+Plus n est grand, plus le facet sera lisible par rapport à l'application d'une esthétique couleur sur la variable de contrôle.
+
+## 16 : lisez ?facet_wrap. Que fait ncol ? Quelles autres options sont possibles ?
+
+```{r}
+?facet_wrap
+```
+
+```{r}
+ggplot(data = mpg) +
+  geom_point(mapping = aes(x = displ, y = hwy)) +
+  facet_wrap(~ class, ncol = 2)
+```
+
+Ncol contrôle le nombre de colonnes affichées.
+
+```{r}
+ggplot(data = mpg) +
+  geom_point(mapping = aes(x = displ, y = hwy)) +
+  facet_wrap(~ class, nrow = 2, scales = "free_x")
+```
+
+```{r}
+ggplot(data = mpg) +
+  geom_point(mapping = aes(x = displ, y = hwy)) +
+  facet_wrap(~ class, ncol = 2, shrink = TRUE)
+```
+
+```{r}
+ggplot(data = mpg) +
+  geom_point(mapping = aes(x = displ, y = hwy)) +
+  facet_wrap(~ class, ncol = 2, as.table = FALSE)
+```
+
+```{r}
+ggplot(data = mpg) +
+  geom_point(mapping = aes(x = displ, y = hwy)) +
+  facet_wrap(~ class, ncol = 2, drop = TRUE)
+```
+
+```{r}
+ggplot(data = mpg) +
+  geom_point(mapping = aes(x = displ, y = hwy)) +
+  facet_wrap(~ class, ncol = 2, dir = "v")
+```
+
+```{r}
+ggplot(data = mpg) +
+  geom_point(mapping = aes(x = displ, y = hwy)) +
+  facet_wrap(~ class, ncol = 2, strip.position = "bottom")
+```
+
+## 17 : avec grid, pourquoi assigner la variable avec le plus de modalités en colonnes ?
+
+Pour rendre le graphique plus lisible.
+
+## 18 : quelles fonctions utiliseriez-vous pour produire un graphique en ligne, en boîte, en zones ou encore un histogramme ?
+
+```{r}
+ggplot2::mpg
+```
+
+```{r}
+ggplot(data = mpg) +
+  geom_line(mapping = aes(x = displ, y = hwy), color = "red")
+```
+
+```{r}
+ggplot(data = mpg) +
+  geom_boxplot(mapping = aes(x = class, y = displ, fill = drv))
+```
+
+```{r}
+ggplot(data = mpg, mapping = aes(x = displ, y = hwy, fill = class)) +
+  geom_area(mapping = NULL, position = "stack")
+```
+
+## 19 : à quoi ressemble ce graphique ?
+
+```{r}
+ggplot(
+  data = mpg,
+  mapping = aes(x = displ, y = hwy, color = drv)
+) +
+  geom_point() +
+  geom_smooth(se = FALSE)
+```
+
+Ma prédiction se révèle exacte.
+
+## 20 : que fait show.legend = FALSE ? Pourquoi l'utiliser ?
+
+Il retire la légende. On l'utilise pour rendre le graphique plus lisible (dans certains cas).
+
+## 21 : que fait geom_smooth ?
+
+Il créé une courbe similaire (dans l'idée) à une droite de régression linéaire, en mobilisant la moyenne des données brut grâce à des outils mathématiques avancés qui dépassent mes connaissances actuelles.
+
+## 22 : ces deux graphiques auront-ils l'air différent ? Pourquoi ?
+
+```{r}
+ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
+  geom_point() +
+  geom_smooth()
+
+ggplot() +
+  geom_point(
+    data = mpg,
+    mapping = aes(x = displ, y = hwy)
+  ) +
+  geom_smooth(
+    data = mpg, mapping = aes(x = displ, y = hwy)
+  )
+```
+
+Ma prédiction était exacte. Le premier code est simplifié.
+
+## 23 : reproduisez les six graphiques.
+
+```{r}
+ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
+  geom_point() +
+  geom_smooth(se = FALSE)
+```
+
+```{r}
+ggplot(data = mpg, mapping = aes(x = displ, y = hwy, linetype = drv)) +
+  geom_point() +
+  geom_smooth(se = FALSE, show.legend = FALSE)
+```
+
+```{r}
+ggplot(data = mpg, mapping = aes(x = displ, y = hwy, linetype = drv, color = drv)) +
+  geom_point() +
+  geom_smooth(se = FALSE, show.legend = FALSE)
+```
+
+```{r}
+ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) +
+  geom_point() +
+  geom_smooth(se = FALSE, show.legend = FALSE, data = filter(mpg, drv == "4"))
+```
+
+```{r}
+ggplot(data = mpg, mapping = aes(x = displ, y = hwy, linetype = drv, color = drv)) +
+  geom_point() +
+  geom_smooth(se = FALSE, show.legend = FALSE, color = "blue")
+```
+
+```{r}
+ggplot(data = mpg) +
+  geom_point(mapping = aes(x = displ, y = hwy, color = drv))
+```

### 2023-05-08 Tests/exercices.Rmd
exercices visualisation 3

@@
   geom_smooth(data = filter(mpg, class == "subcompact"), se = FALSE, color = "purple")
+```
+
+## geom_bar
+
+```{r}
+ggplot(data = diamonds) +
+  geom_bar(mapping = aes(x = cut, fill = cut))
+```
+
+## specific geom_bar
+
+```{r}
+demo <- tribble(
+  ~a, ~b,
+  "bar_1", 20,
+  "bar_2", 30,
+  "bar_3", 40
+)
+
+ggplot(data = demo) +
+  geom_bar(
+    mapping = aes(x = a, y = b), stat = "identity"
+  )
+```
+
+## proportion geom_bar
+
+```{r}
+ggplot(data = diamonds) +
+  geom_bar(mapping = aes(x = cut, y = after_stat(prop), group = 1))
+```
+
+## goated geom_bar
+
+```{r}
+ggplot(data = diamonds) +
+  stat_summary(mapping = aes(x = cut, y = depth),
+           fun.ymin = min,
+           fun.ymax = max,
+           fun.y = median
+           )
+```
+
+## multivaried barchart
+
+```{r}
+ggplot(data = diamonds) +
+  geom_bar(mapping = aes(x = cut, fill = clarity))
+```
+
+## multivaried stacked barchart
+
+```{r}
+ggplot(data = diamonds) +
+  geom_bar(mapping = aes(x = cut, fill = clarity,), position = "fill")
+```
+
+## multivaried dodged barchart
+
+```{r}
+ggplot(data = diamonds) +
+  geom_bar(mapping = aes(x = cut, fill = clarity,), position = "dodge")
+```
+
+# jittering scaterplot
+
+```{r}
+ggplot(data = mpg) +
+  geom_jitter(mapping = aes(x = displ, y = hwy))
 ```
@@
 
+```{r}
+ggplot(data = mpg, mapping = aes(x = mpg$displ)) +
+  geom_histogram(fill = "blue") +
+  geom_vline(xintercept = mean(mpg$displ))
+```
+
 ## 19 : à quoi ressemble ce graphique ?
@@
 ```
+
+## 24 : écrivez un code avec geom_bar permettant de visualiser le goated geom_bar.
+
+```{r}
+ggplot(data = diamonds) +
+  geom_col(mapping = aes(x = cut, y = depth, fill = cut))
+```
+
+## 25 : Quelle est la spécificité de geom_col ?
+
+```{r}
+?geom_col
+```
+
+Geom_col permet de faire des diagrammes en barres représentant deux variables.
+
+## 26 : Quel est le point commun entre les arguments entre stat et geom ?
+
+Je n'ai pas compris la question.
+
+## 27 : Quelles variables stat_smooth mobilise ?
+
+Différents types de moyennes.
+
+## 28 : Quel est le problème de ces deux graphiques ?
+
+```{r}
+ggplot(data = diamonds) +
+  geom_bar(mapping = aes(x = cut, y = after_stat(prop), group = 1)) +
+  geom_bar(mapping = aes(x = cut, y = after_stat(prop), group = 1))
+```
+
+Il faut assigner un groupe à la fonction after_stat, sinon le graphique est illisible.

### 2023-05-09 Tests/exercices.Rmd
exercices visualisation 4

@@
 
+## coord_flip
+
+```{r}
+ggplot(data = mpg, mapping = aes(x = class, y = hwy)) +
+  geom_boxplot() +
+  coord_flip()
+```
+
+## coord_polar
+
+```{r}
+bar <- ggplot(data = diamonds) +
+  geom_bar(
+    mapping = aes(x = cut, fill = cut),
+    show.legend = FALSE,
+    width = 1
+  ) +
+  theme(aspect.ratio = 1) +
+  labs(x = NULL, y = NULL)
+
+bar + coord_flip()
+bar + coord_polar()
+```
+
 # Exercices (visualisation)
@@
   geom_point(mapping = aes(x = displ, y = hwy, color = year))
-```
-
-```{r}
-ggplot(data = mpg) +
-  geom_point(mapping = aes(x = displ, y = hwy, shape = year))
 ```
@@
 
-## 25 : Quelle est la spécificité de geom_col ?
+## 25 : quelle est la spécificité de geom_col ?
 
@@
 
-## 26 : Quel est le point commun entre les arguments entre stat et geom ?
+## 26 : quel est le point commun entre les arguments entre stat et geom ?
 
@@
 
-## 27 : Quelles variables stat_smooth mobilise ?
+## 27 : quelles variables stat_smooth mobilise ?
 
@@
 
-## 28 : Quel est le problème de ces deux graphiques ?
+## 28 : quel est le problème de ces deux graphiques ?
 
@@
 Il faut assigner un groupe à la fonction after_stat, sinon le graphique est illisible.
+
+## 29 : comment améliorer ce graphique ?
+
+```{r}
+ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) +
+  geom_point()
+```
+
+Code revu :
+
+```{r}
+ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) +
+  geom_jitter()
+```
+
+## 30 : quels paramètres contrôlent le jittering ?
+
+Widht and height.
+
+## 31 : comparez geom_jitter et geom_count
+
+```{r}
+ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) +
+  geom_jitter()
+
+ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) +
+  geom_count()
+```
+
+## 32 : quel est l'ajustement de la position par défaut pour geom_boxplot ? Démontrez-le avec mpg
+
+```{r}
+ggplot(data = mpg, mapping = aes(x = class, y = displ)) +
+  geom_boxplot(fill = "pink") +
+  geom_jitter(width = 1)
+```
+
+## 33 : transformez un diagramme de barres en diagramme circulaire
+
+```{r}
+ggplot(data = diamonds, mapping = aes(x = cut, fill = cut)) +
+  geom_bar() +
+  coord_polar()
+```
+
+## 34 : que fait labs ?
+
+```{r}
+?labs
+```
+
+Modify axis, legend, and plot labels
+
+## 35 : que nous dit le graphique suivant sur la relation entre cty et hwy ?
+
+```{r}
+ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) +
+  geom_jitter() +
+  geom_abline() +
+  coord_fixed()
+```
+
+```{r}
+?geom_abline
+```

### 2023-05-09 Tests/1. exercices/trans.Rmd
exos trans 1
Fonctions filter() et select().

---
title: "trans"
author: "Thea"
date: "2023-05-09"
output: html_document
---

```{r}

library(tidyverse)

library(nycflights13)

```

# Exemples

## tybble

```{r}

nycflights13::flights

```

## filter

```{r}

(jan <- filter(flights, month == 1, day == 1))

```

## arrange

```{r}

arrange(flights, year, month, day)

```

## desc

```{r}

arrange(flights, desc(arr_delay))

```

## select (inclusive)

```{r}

select(flights, year : day)

```

## select (exclusive)

```{r}

select(flights, !(year : day))

```

## rename

```{r}

rename(flights, tail_num = tailnum)

```

## everything

```{r}

select(flights, air_time, distance, everything())

```

# Exercices

## 1 : affichez les vols retardés de deux heurs ou plus

```{r}

filter(flights, 
       arr_delay >= 2, 
       dest == "IAH" | origin == "HOU", 
       carrier == "UA" | carrier == "AA" | carrier == "DL",
       month %in% c(7, 8, 9),
       dep_delay <= 0,
       dep_time >= 0 & dep_time <= 600)

```

## 2 : simplifiez votre code avec between()

```{r}

?between

```

```{r}

filter(flights,
       arr_delay >= 2, 
       dest == "IAH" | origin == "HOU", 
       carrier == "UA" | carrier == "AA" | carrier == "DL",
       month %in% c(7, 8, 9),
       dep_delay <= 0,
       between(dep_time, 0, 600)
       )

```

## 3 : affichez les vols dont l'heure de départ est inconnue

```{r}

filter(flights, is.na(dep_time))

```

Ces colonnes représentent des vols annulés.

## 4 : question difficile à comprendre (voir manuel)

## 5 : classer les NA en tête de colonnes

```{r}

na <- arrange(flights, desc(is.na(dep_delay)))

```

## 6 : triez les vols par ordre décroissant de retard à l'arrivée et par ordre croissant d'heure du décollage

```{r}

arrange(flights, dep_time, desc(arr_delay))

```

## 7 : triez les vols selon leur vitesse

```{r}

arrange(flights, distance / air_time)

```

## 8 : quels vols ont parcourus la distance la plus grande et la plus courte ?

```{r}

arrange(flights, desc(distance))

```

```{r}

arrange(flights, distance)

```

## 9 : immaginez toutes les manières de séléctionner dep et arr

```{r}

# J'ai échoué !

select(flights, dep_time, dep_delay, arr_time, arr_delay)

```

## 10 : que se passe-t-il s'il on appelle plusieurs fois la même variable dans select() ?

La variable n’apparaît qu'une fois.

## 11 : quel est le rôle de la fonction one_of() ?

```{r}

vars <- c("year", "month", "day", "dep_delay", "arr_delay")

select(flights, one_of(vars))

```

## 12 : je n'ai pas compris la question

```{r}

?select

```

```{r}

select(flights, contains("time"))

```

### 2023-05-09 Tests/1. exercices/visua.Rmd
exos trans 1
Fonctions filter() et select().

--- 2023-05-09 Tests/exercices.Rmd
@@
 ---
-title: "exercices"
+title: "visualisation"
 author: "Théa Carlier"
@@
 library(tidyverse)
-library(readr)
 ```
 
-# Exemple
-
-```{r}
-df <- tibble(
-  a = rnorm(10),
-  b = rnorm(10),
-  c = rnorm(10),
-  d = rnorm(10),
-)
-```
-
-```{r}
-output <- vector("double", ncol(df)) # 1. output (création d'un vecteur)
-for (i in seq_along(df)) {           # 2. sequence (dans lequel la boucle va opérer pour chaque élément)
-  output[[i]] <- median(df[[i]])     # 3. body 
-}
-output
-```
-
-# Exercices (boucles)
-
-## 1 : calculez la moyenne de chaque colonne dans la base de donnée *mtcars*
-
-```{r}
-mtcars <- read_csv("mtcars.csv")
-```
-
-```{r}
-mtcars_2 <-
-  mtcars %>%
-  select(mpg, disp, hp, drat, wt, qsec)
-
-output_2 <- vector("double", ncol(mtcars_2))
-for (i in seq_along(mtcars_2)) {
-  output_2[[i]] <- mean(mtcars_2[[i]])
-}
-output_2
-```
-
-## 2 : déterminez le type de chaque colonnes
-
-```{r}
-output_3 <- vector("double", ncol(mtcars))
-for (i in seq_along(mtcars)) {
-  output_3[[i]] <- class(mtcars[[i]])
-}
-output_3
-```
-
-## 3 : calculez le nombre de valeurs uniques dans chaque colonnes
-
-```{r}
-output_4 <- vector("double", ncol(mtcars))
-for (i in seq_along(mtcars)) {
-  output_4[[i]] <- n_distinct(mtcars[[i]])
-}
-output_4
-```
-
-# Exemple (visualisation)
+# Exemples (visualisation)
 
@@
 
-# jittering scaterplot
+## jittering scaterplot
 
@@
 bar + coord_polar()
+```
+
+```{r}
+ggplot(data = diamonds) +
+  stat_count(
+    mapping = aes(x = cut, fill = cut), geom = "bar", position = "stack", show.legend = FALSE) +
+  coord_flip() +
+  facet_wrap(diamonds$color, nrow = 4, ) +
+  theme_few()
 ```

### 2023-05-11 Tests/1. tests/an_test.Rmd
rename 2. tests

--- 2023-05-04 Tests/an_test.Rmd
@@
     x = "Salaire net", 
-    y = "Diplôme universitaire (0 = non ; 1 = oui)") +
+    y = "Diplome universitaire (0 = non ; 1 = oui)") +
   theme_few()
@@
 df_scripted %>%
-  select(AGEGP, CFATINC, HLEV2G_2) %>%
+  select(AGEGP, 
+         CFATINC, 
+         HLEV2G_2) %>%
   drop_na() %>%

### 2023-05-17 Article-Reseaux sociaux_sentiment.Rmd
update

---
title: "Article:réseaux sociaux et sentiments d'appartenance au Québec"
author:
- Vissého Adjiwanou ^[corresponding author - adjiwanou.vissého@uqam.ca]
- Département de Sociologie, Université du Québec à Montréal (UQAM)
- Vincent Bertrand-Huneault (Département de démographie, Université de Montréal)
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  word_document:
    keep_md: yes
  html_document:
    toc: yes
    toc_depth: 3
  template: svm-latex-ms2.tex
  pdf_document: default
  citation_package: natbib
bibliography:
- covid_Nima.bib
- Covid_Canada.bib
- Covid_Family.bib
---

### 2023-05-18 Codes/Article - Sentiment d'appartenance.Rmd
Déplacement du projet d'un dossier à l'autre

---
title: "Article - Sentiment d'appartenance"
author: "Thea"
date: "2023-05-17"
output: pdf_document
---

```{r}
library(summarytools)
library(tidyverse)
library(ggthemes)
library(haven)
```

```{r}
Base_Trajipaq_Finale_Ok <- read_dta("~/Desktop/Base_Trajipaq_Finale_Ok.dta")
```

# Question de recherche : de quelles façons les discriminations que subissent les personnes immigrantes influencent leur sentiment d'appartenance au Québec ?

## Variable indépendante = Les discriminations que subissent les personnes immigrantes :

- dans l'emploi
- dans la recherche de logement
- dans la santé
- sur le lieu de formation

Type de variable = factoriel
Niveau de mesure = ordinal (échelle de Likert)

Réponses possibles :

01. Tout à fait d'accord
02. Plutôt d'accord
03. Plutôt pas d'accord
04. Pas du tout d'accord
98. Je ne sais pas
99. Je ne préfère pas répondre

## Variable dépendante = le sentiment d'appartenance au Québec :

- se sentir chez soi au Québec
- la nécessité d'oublier sa culture d'origine pour s'intégrer au Québec
- être vu(e) comme un(e) Québécois(e)
- se sentir Québécois(e)
- se sentir Canadien(ne)
- avoir un sentiment d'appartenance avec son pays d'origine

Type de variable = factoriel
Niveau de mesure = ordinal (échelle de Likert)

Réponses possibles :

01. Tout à fait d'accord
02. Plutôt d'accord
03. Plutôt pas d'accord
04. Pas du tout d'accord
98. Je ne sais pas
99. Je ne préfère pas répondre

# Présentation de la variable indépendante

## Q39A. Dites si vous êtes tout à fait d’accord, plutôt d’accord, plutôt pas d’accord ou pas du tout d’accord avec les propositions suivantes (participation identitaire et sentiment d’appartenance) : je me sens chez moi au Québec.

```{r}

Base_Trajipaq_Finale_Ok <-
  Base_Trajipaq_Finale_Ok %>%
  mutate(Q39A = case_when(
    Q39A == 1 ~ 4,
    Q39A == 2 ~ 3,
    Q39A == 3 ~ 2,
    Q39A == 4 ~ 1,
    Q39A == 98 ~ NA,
    Q39A == 99 ~ NA
  ))

```

```{r}

freq(Base_Trajipaq_Finale_Ok$Q39A)

```

Dans le tableau de fréquence ci-dessus, on observe que la distribution est asymétrique vers la gauche. Autrement dit, 87.01% des personnes interrogées sont tout à fait d'accord ou plutôt d'accord pour dire qu'elles se sentent chez elles au Québec. On remarque également que 1.79% des personnes interrogées ne savaient pas ou n'ont pas souhaité répondre à la question. Etant donné la faible proportion de ces réponses, celles-ci seront retirées de l'analyse.

```{r}

ggplot(data = Base_Trajipaq_Finale_Ok, mapping = aes(x = Q39A)) +
  geom_bar(fill = "pink",
           na.rm = TRUE
           ) +
  theme_few()

```

Dans le diagramme de barres ci-dessus, on observe l'asymétrie vers la gauche mentionnée précédemment.

## Q39B. Dites si vous êtes tout à fait d’accord, plutôt d’accord, plutôt pas d’accord ou pas du tout d’accord avec les propositions suivantes (participation identitaire et sentiment d’appartenance) : il faut faire oublier ses origines pour se faire accepter au Québec.

```{r}

Base_Trajipaq_Finale_Ok <-
  Base_Trajipaq_Finale_Ok %>%
  mutate(Q39B = case_when(
    Q39B == 1 ~ 4,
    Q39B == 2 ~ 3,
    Q39B == 3 ~ 2,
    Q39B == 4 ~ 1,
    Q39B == 98 ~ NA,
    Q39B == 99 ~ NA
  ))

```

```{r}

freq(Base_Trajipaq_Finale_Ok$Q39B)

```

Dans le tableau de fréquence ci-dessus, on observe que la distribution est asymétrique vers la droite. Autrement dit, 67.88% des personnes interrogées ne sont pas du tout d'accord ou plutôt pas d'accord pour dire qu'il faut faire oublier ses origines pour se faire accepter au Québec. On remarque également que 3.96% des personnes interrogées ne savaient pas ou n'ont pas souhaité répondre à la question. Etant donné la faible proportion de ces réponses, celles-ci seront retirées de l'analyse.

```{r}

ggplot(data = Base_Trajipaq_Finale_Ok, mapping = aes(x = Q39B)) +
  geom_bar(fill = "pink",
           na.rm = TRUE
           ) +
  theme_few()

```

Dans le diagramme de barres ci-dessus, on observe l'asymétrie vers la droite mentionnée précédemment.

## Q39C. Dites si vous êtes tout à fait d’accord, plutôt d’accord, plutôt pas d’accord ou pas du tout d’accord avec les propositions suivantes (participation identitaire et sentiment d’appartenance) : on me voit comme une Québécoise.

```{r}

Base_Trajipaq_Finale_Ok <-
  Base_Trajipaq_Finale_Ok %>%
  mutate(Q39C = case_when(
    Q39C == 1 ~ 4,
    Q39C == 2 ~ 3,
    Q39C == 3 ~ 2,
    Q39C == 4 ~ 1,
    Q39C == 98 ~ NA,
    Q39C == 99 ~ NA
  ))

```

```{r}

freq(Base_Trajipaq_Finale_Ok$Q39C)

```

Dans le tableau de fréquence ci-dessus, on observe que la distribution est bimodale (1 et 4). Autrement dit, cette question divise nos répondants. On remarque également que 5.63% des personnes interrogées ne savaient pas ou n'ont pas souhaité répondre à la question. Etant donné la faible proportion de ces réponses, celles-ci seront retirées de l'analyse.

```{r}

ggplot(data = Base_Trajipaq_Finale_Ok, mapping = aes(x = Q39C)) +
  geom_bar(fill = "pink",
           na.rm = TRUE
           ) +
  theme_few()

```

Dans le diagramme de barres ci-dessus, on observe les deux modes de la distribution décrits précédemment.

## Q39D. Dites si vous êtes tout à fait d’accord, plutôt d’accord, plutôt pas d’accord ou pas du tout d’accord avec les propositions suivantes (participation identitaire et sentiment d’appartenance) : je me sens Québécois(e).

```{r}

Base_Trajipaq_Finale_Ok <-
  Base_Trajipaq_Finale_Ok %>%
  mutate(Q39D = case_when(
    Q39D == 1 ~ 4,
    Q39D == 2 ~ 3,
    Q39D == 3 ~ 2,
    Q39D == 4 ~ 1,
    Q39D == 98 ~ NA,
    Q39D == 99 ~ NA
  ))

```

```{r}

freq(Base_Trajipaq_Finale_Ok$Q39D)

```

Dans le tableau de fréquence ci-dessus, on observe que la distribution est asymétrique vers la gauche. Autrement dit, 60.72% des personnes interrogées sont tout à fait d'accord ou plutôt d'accord pour dire qu'elles se sentent Québécois(e)s. On remarque également que 4.03% des personnes interrogées ne savaient pas ou n'ont pas souhaité répondre à la question. Etant donné la faible proportion de ces réponses, celles-ci seront retirées de l'analyse.

```{r}

ggplot(data = Base_Trajipaq_Finale_Ok, mapping = aes(x = Q39D)) +
  geom_bar(fill = "pink",
           na.rm = TRUE
           ) +
  theme_few()

```

Dans le diagramme de barres ci-dessus, on observe l'asymétrie vers la gauche mentionnée précédemment.

## Q39E. Dites si vous êtes tout à fait d’accord, plutôt d’accord, plutôt pas d’accord ou pas du tout d’accord avec les propositions suivantes (participation identitaire et sentiment d’appartenance) : je me sens Canadien(ne).

```{r}

Base_Trajipaq_Finale_Ok <-
  Base_Trajipaq_Finale_Ok %>%
  mutate(Q39E = case_when(
    Q39E == 1 ~ 4,
    Q39E == 2 ~ 3,
    Q39E == 3 ~ 2,
    Q39E == 4 ~ 1,
    Q39E == 98 ~ NA,
    Q39E == 99 ~ NA
  ))

```

```{r}

freq(Base_Trajipaq_Finale_Ok$Q39E)

```

Dans le tableau de fréquence ci-dessus, on observe que la distribution est asymétrique vers la gauche. Autrement dit, 79.98% des personnes interrogées sont tout à fait d'accord ou plutôt d'accord pour dire qu'elles se sentent Canadien(ne). On remarque également que 2.69% des personnes interrogées ne savaient pas ou n'ont pas souhaité répondre à la question. Etant donné la faible proportion de ces réponses, celles-ci seront retirées de l'analyse.

```{r}

ggplot(data = Base_Trajipaq_Finale_Ok, mapping = aes(x = Q39E)) +
  geom_bar(fill = "pink",
           na.rm = TRUE
           ) +
  theme_few()

```

Dans le diagramme de barres ci-dessus, on observe l'asymétrie vers la gauche mentionnée précédemment.

## Q39F. Dites si vous êtes tout à fait d’accord, plutôt d’accord, plutôt pas d’accord ou pas du tout d’accord avec les propositions suivantes (participation identitaire et sentiment d’appartenance) : j'ai un sentiment d'appartenance à mon pays d'origine.

```{r}

Base_Trajipaq_Finale_Ok <-
  Base_Trajipaq_Finale_Ok %>%
  mutate(Q39F = case_when(
    Q39F == 1 ~ 4,
    Q39F == 2 ~ 3,
    Q39F == 3 ~ 2,
    Q39F == 4 ~ 1,
    Q39F == 98 ~ NA,
    Q39F == 99 ~ NA
  ))

```

```{r}

freq(Base_Trajipaq_Finale_Ok$Q39F)

```

Dans le tableau de fréquence ci-dessus, on observe que la distribution est asymétrique vers la gauche. Autrement dit, 81.42% des personnes interrogées sont tout à fait d'accord ou plutôt d'accord pour dire qu'elles possèdent un sentiment d'appartenance envers leur pays d'origine. On remarque également que 2.69% des personnes interrogées ne savaient pas ou n'ont pas souhaité répondre à la question. Etant donné la faible proportion de ces réponses, celles-ci seront retirées de l'analyse.

```{r}

ggplot(data = Base_Trajipaq_Finale_Ok, mapping = aes(x = Q39F)) +
  geom_bar(fill = "pink",
           na.rm = TRUE
           ) +
  theme_few()

```

Dans le diagramme de barres ci-dessus, on observe l'asymétrie vers la gauche mentionnée précédemment.

## I1 : L'indice du sentiment d'appartenance au Québec

A la différence des variables précédentes, l'indice du sentiment d'appartenance est une variable numérique continue dont le niveau de mesure est ratio.

```{r}

v_I1 <- c("Q39A",
          "Q39B",
          "Q39C",
          "Q39D",
          "Q39E",
          "Q39F"
          )

Base_Trajipaq_Finale_Ok <-
  Base_Trajipaq_Finale_Ok %>%
  rowwise() %>%
  mutate(I1 = (sum(c_across(all_of(v_I1)), na.rm = TRUE) - length(v_I1)) / (3*length(v_I1))) %>%
  mutate(I1 = if (I1 < 0) NA_real_ else I1) %>%
  ungroup()

```

```{r}

descr(Base_Trajipaq_Finale_Ok$I1)

```

```{r}

ggplot(data = Base_Trajipaq_Finale_Ok, mapping = aes(x = I1)) +
  geom_histogram(na.rm = TRUE, fill = "pink", binwidth = 0.01) +
  theme_bw()

```

Le tableau descriptif et l'histogramme ci-dessous nous indiquent que la distribution est asymétrique vers la gauche.

### 2023-05-18 Codes/Article - Sentiment d'appartenance.Rmd
Début de la description des variables de contrôle

@@
 
+## Variables de contrôle = caractéristiques sociodémographiques :
+
+- sexe
+- statut matrimonial
+- pays d'origine
+- année d'obtention de la résidence permanente
+- année d'obtention de la citoyenneté canadienne
+- année d'arrivée au Québec
+- année depuis la première installation au Québec
+- statut d'immigration avant l'obtention de la résidence permanente
+- niveau d'éducation
+- religion
+- importance accordée à la religion
+- compétence : lire le français
+- compétence : parler français
+- compétence : rédiger en français
+- compétence : comprendre le français
+- niveau de connaissance en français lors de l'arrivée au Québec
+- compétence : lire l'anglais
+- compétence : parler anglais
+- compétence : rédiger en anglais
+- compétence : comprendre l'anglais
+- niveau de connaissance en anglais lors de l'arrivée au Québec
+- suivi de cours dédiés à la francisation après l'arrivée au Québec
+- état de santé
+- problèmes de santé avant l'installation au Québec
+- problèmes de santé au cours des 12 derniers mois
+- problèmes de santé il y a moins de 5 ans
+- problèmes de santé il y a plus de 10 ans
+- lieu de naissance de la mère
+- niveau d'éducation de la mère
+- lieu de naissance du père
+- niveau d'éducation du père
+
+# Présentation des variables de contrôle
+
+## Q10. Êtes-vous un homme, une femme ou autre ?
+
+01. Homme
+02. Femme
+97. Autre
+
+Type de variable = catégorielle
+Niveau de mesure = nominal
+
+```{r}
+
+Base_Trajipaq_Finale_Ok <-
+  Base_Trajipaq_Finale_Ok %>%
+  mutate(Q10_F = case_when(
+    Q10 == 1 ~ "Homme",
+    Q10 == 2 ~ "Femme",
+    Q10 == 97 ~ "Autre"
+  ))
+
+```
+
+```{r}
+
+freq(Base_Trajipaq_Finale_Ok$Q10_F)
+
+```
+
+```{r}
+
+ggplot(data = Base_Trajipaq_Finale_Ok, mapping = aes(x = Q10_F)) +
+  geom_bar(fill = "pink") +
+  theme_bw()
+
+```
+
+Dans le tableau de fréquence et le diagramme de barres ci-dessus, on observe que le sexe modal est "Femme".
+
+## STATUT. Êtes-vous actuellement... ?
+
+01. Célibataire/jamais marié(e) ou conjoint(e) de fait
+02. Marié(e)
+03. Conjoint(e) de fait/en union libre
+04. Séparé(e)
+05. Divorcé(e)
+06. Veuf(ve)
+
+Type de variable = catégorielle
+Niveau de mesure = nominal
+
+```{r}
+
+freq(Base_Trajipaq_Finale_Ok$STATUT)
+
+```
+
+```{r}
+
+ggplot(data = Base_Trajipaq_Finale_Ok, mapping = aes(x = STATUT)) +
+  geom_bar(fill = "pink", na.rm = TRUE) +
+  theme_bw()
+
+```
+
+Dans le tableau de fréquence et le diagramme de barres ci-dessus, on observe que le statut modal est "Marié(e)".
+
+## Q11. Dans quel pays êtes-vous né ?
+
+Pays. (liste déroulante)
+998. Je ne sais pas
+999. Je préfère ne pas répondre
+
+Type de variable = catégorielle
+Niveau de mesure = nominal
+
+## Q12. En quelle année avez-vous obtenu la résidence permanente ?
+
+Année
+9998. Je ne sais pas
+9999. Je préfère ne pas répondre
+
+Type de variable = catégorielle
+Niveau de mesure = nominale
+
+```{r}
+
+freq(Base_Trajipaq_Finale_Ok$Q11)
+
+```
+
+```{r}
+ggplot(data = Base_Trajipaq_Finale_Ok, mapping = aes(x = as.factor(Q11))) +
+  geom_bar(fill = "pink") +
+  theme_bw()
+```
+
+Dans le tableau de fréquence et le diagramme de barres ci-dessous, on peut observer que les pays d'origine modaux sont la France (26.685%) et l'Algérie (6.649%).
+
+## Q13. En quelle année est-ce que vous êtes devenu citoyen(ne) ?
+
+Année
+9998. Je ne sais pas
+9999. Je préfère ne pas répondre
+
+Type de variable = numérique continue
+Niveau de mesure = ratio
+
 # Présentation de la variable indépendante
@@
            ) +
-  theme_few()
+  theme_bw()
 

### 2023-05-22 Codes/Article - Sentiment d'appartenance.Rmd
Présentation de la variable dépendante.

@@
 Le tableau descriptif et l'histogramme ci-dessous nous indiquent que la distribution est asymétrique vers la gauche.
+Autrement dit, les répondants estiment en moyenne entretenir un sentiment d'appartenance relativement développé (0.59).
+
+# Présentation de la variable dépendante
+
+## Q40A.
+Au cours des 12 mois (avant la crise de la COVID-19), pensez-vous avoir subi des traitements inégalitaires ou des discriminations ?
+Un traitement inégalitaire ou une discrimination se repère si « vous avez été moins bien traité que les autres, sans raison valable » : dans l'emploi ?
+
+```{r}
+
+Base_Trajipaq_Finale_Ok <-
+  Base_Trajipaq_Finale_Ok %>%
+  mutate(Q40A = case_when(
+    Q40A == 1 ~ 3,
+    Q40A == 2 ~ 2,
+    Q40A == 3 ~ 1,
+    Q40A == 96 ~ NA,
+    Q40A == 98 ~ NA,
+    Q40A == 99 ~ NA
+  ))
+
+```
+
+```{r}
+
+freq(Base_Trajipaq_Finale_Ok$Q40A)
+
+```
+
+Dans le tableau de fréquence ci-dessus, on observe que la distribution est asymétrique vers la droite Autrement dit, 65.44% des personnes interrogées disent ne jamais avoir subit de discrimination dans l'emploi au cours des 12 derniers mois précédent la crise du covid 19.
+On remarque également que 9.47% des personnes interrogées ne savaient pas ou n'ont pas souhaité répondre à la question.
+Etant donné la faible proportion de ces réponses, celles-ci seront retirées de l'analyse.
+
+```{r}
+
+ggplot(data = Base_Trajipaq_Finale_Ok, mapping = aes(x = Q40A)) +
+  geom_bar(fill = "pink", na.rm = TRUE) +
+  theme_bw()
+
+```
+
+Dans le diagramme de barres ci-dessus, on observe l'asymétrie vers la droite mentionnée précédemment.
+
+## Q40B.
+Au cours des 12 mois (avant la crise de la COVID-19), pensez-vous avoir subi des traitements inégalitaires ou des discriminations ?
+Un traitement inégalitaire ou une discrimination se repère si « vous avez été moins bien traité que les autres, sans raison valable » : dans la recherche de logement ?
+
+```{r}
+
+Base_Trajipaq_Finale_Ok <-
+  Base_Trajipaq_Finale_Ok %>%
+  mutate(Q40B = case_when(
+    Q40B == 1 ~ 3,
+    Q40B == 2 ~ 2,
+    Q40B == 3 ~ 1,
+    Q40B == 96 ~ NA,
+    Q40B == 98 ~ NA,
+    Q40B == 99 ~ NA
+  ))
+
+```
+
+```{r}
+
+freq(Base_Trajipaq_Finale_Ok$Q40B)
+
+```
+
+Dans le tableau de fréquence ci-dessus, on observe que la distribution est asymétrique vers la droite Autrement dit, 79.64% des personnes interrogées disent ne jamais avoir subit de discrimination dans la recherche de logement au cours des 12 derniers mois précédent la crise du covid 19.
+On remarque également que 17.34% des personnes interrogées ne savaient pas ou n'ont pas souhaité répondre à la question.
+Etant donné la faible proportion de ces réponses, celles-ci seront retirées de l'analyse.
+
+```{r}
+
+ggplot(data = Base_Trajipaq_Finale_Ok, mapping = aes(x = Q40B)) +
+  geom_bar(fill = "pink", na.rm = TRUE) +
+  theme_bw()
+
+```
+
+Dans le diagramme de barres ci-dessus, on observe l'asymétrie vers la droite mentionnée précédemment.
+
+## Q40C.
+Au cours des 12 mois (avant la crise de la COVID-19), pensez-vous avoir subi des traitements inégalitaires ou des discriminations ?
+Un traitement inégalitaire ou une discrimination se repère si « vous avez été moins bien traité que les autres, sans raison valable » : dans la santé ?
+
+```{r}
+
+Base_Trajipaq_Finale_Ok <-
+  Base_Trajipaq_Finale_Ok %>%
+  mutate(Q40C = case_when(
+    Q40C == 1 ~ 3,
+    Q40C == 2 ~ 2,
+    Q40C == 3 ~ 1,
+    Q40C == 96 ~ NA,
+    Q40C == 98 ~ NA,
+    Q40C == 99 ~ NA
+  ))
+
+```
+
+```{r}
+
+freq(Base_Trajipaq_Finale_Ok$Q40C)
+
+```
+
+Dans le tableau de fréquence ci-dessus, on observe que la distribution est asymétrique vers la droite Autrement dit, 82.26% des personnes interrogées disent ne jamais avoir subit de discrimination dans la santé au cours des 12 derniers mois précédent la crise du covid 19.
+On remarque également que 10.17% des personnes interrogées ne savaient pas ou n'ont pas souhaité répondre à la question.
+Etant donné la faible proportion de ces réponses, celles-ci seront retirées de l'analyse.
+
+```{r}
+
+ggplot(data = Base_Trajipaq_Finale_Ok, mapping = aes(x = Q40C)) +
+  geom_bar(fill = "pink", na.rm = TRUE) +
+  theme_bw()
+
+```
+
+Dans le diagramme de barres ci-dessus, on observe l'asymétrie vers la droite mentionnée précédemment.
+
+## Q40D.
+Au cours des 12 mois (avant la crise de la COVID-19), pensez-vous avoir subi des traitements inégalitaires ou des discriminations ?
+Un traitement inégalitaire ou une discrimination se repère si « vous avez été moins bien traité que les autres, sans raison valable » : sur le lieu de formation ?
+
+```{r}
+
+Base_Trajipaq_Finale_Ok <-
+  Base_Trajipaq_Finale_Ok %>%
+  mutate(Q40D = case_when(
+    Q40D == 1 ~ 3,
+    Q40D == 2 ~ 2,
+    Q40D == 3 ~ 1,
+    Q40D == 96 ~ NA,
+    Q40D == 98 ~ NA,
+    Q40D == 99 ~ NA
+  ))
+
+```
+
+```{r}
+
+freq(Base_Trajipaq_Finale_Ok$Q40D)
+
+```
+
+Dans le tableau de fréquence ci-dessus, on observe que la distribution est asymétrique vers la droite Autrement dit, 81.40% des personnes interrogées disent ne jamais avoir subit de discrimination sur le lieu de formation au cours des 12 derniers mois précédent la crise du covid 19.
+On remarque également que 15.74% des personnes interrogées ne savaient pas ou n'ont pas souhaité répondre à la question.
+Etant donné la faible proportion de ces réponses, celles-ci seront retirées de l'analyse.
+
+```{r}
+
+ggplot(data = Base_Trajipaq_Finale_Ok, mapping = aes(x = Q40D)) +
+  geom_bar(fill = "pink", na.rm = TRUE) +
+  theme_bw()
+
+```
+
+Dans le diagramme de barres ci-dessus, on observe l'asymétrie vers la droite mentionnée précédemment.
+
+## I2 : L'indice de discrimination lors des 12 derniers mois précédent la crise du covid 19
+
+A la différence des variables précédentes, l'indice du sentiment d'appartenance est une variable numérique continue dont le niveau de mesure est ratio.
+
+```{r}
+
+v_I2 <- c("Q40A",
+          "Q40B",
+          "Q40C",
+          "Q39D"
+          )
+
+Base_Trajipaq_Finale_Ok <-
+  Base_Trajipaq_Finale_Ok %>%
+  rowwise() %>%
+  mutate(I2 = (sum(c_across(all_of(v_I2)), na.rm = TRUE) - length(v_I2)) / (2.25*length(v_I2))) %>%
+  mutate(I2 = if (I2 < 0) NA_real_ else I2) %>%
+  ungroup()
+
+```
+
+```{r}
+
+descr(Base_Trajipaq_Finale_Ok$I2)
+
+```
+
+```{r}
+
+ggplot(data = Base_Trajipaq_Finale_Ok, mapping = aes(x = I2)) +
+  geom_histogram(na.rm = TRUE, fill = "pink", binwidth = 0.01) +
+  theme_bw()
+
+```
+
+Le tableau descriptif et l'histogramme ci-dessous nous indiquent que la distribution est asymétrique vers la droite.
+Autrement dit, les répondants estiment en moyenne avoir subit une faible discrimination (0.28).

### 2023-05-22 Codes/Article - Sentiment d'appartenance.Rmd
Progression dans la présentation des variables socio-démographiques.

@@
 
-## Q12. En quelle année avez-vous obtenu la résidence permanente ?
-
-Année
-9998. Je ne sais pas
-9999. Je préfère ne pas répondre
-
-Type de variable = catégorielle
-Niveau de mesure = nominale
-
 ```{r}
@@
 
+## Q12. En quelle année avez-vous obtenu la résidence permanente ?
+
+Année
+9998. Je ne sais pas
+9999. Je préfère ne pas répondre
+
+Type de variable = catégorielle
+Niveau de mesure = nominale
+
 ## Q13. En quelle année est-ce que vous êtes devenu citoyen(ne) ?
@@
 Niveau de mesure = ratio
+
+## Q14I. Depuis quelle année vivez-vous au Québec ?
+
+Année
+9996. La même année que la résidence permanente
+9998. Je ne sais pas
+9999. Je préfère ne pas répondre
+
+Type de variable = numérique continue
+Niveau de mesure = ratio
+
+## Q14N. En quelle année avez-vous commencé à vivre au Québec pendant au moins un an ?
+
+Année
+9996. Depuis la naissance
+9998. Je ne sais pas
+9999. Je préfère ne pas répondre
+
+Type de variable = numérique continue
+Niveau de mesure = ratio
+
+## Q15I. Avant l’obtention de la résidence permanente, quel était votre statut d’immigration ?
+
+1. Travailleur temporaire
+02. Demandeur d’asile
+03. Étudiant étranger
+04. Touriste
+97. Autre, préciser
+98. Je ne sais pas
+99. Je préfère ne pas répondre
+
+Type de variable = catégorielle
+Niveau de mesure = nominal
+
+```{r}
+
+Base_Trajipaq_Finale_Ok <-
+  Base_Trajipaq_Finale_Ok %>%
+  mutate(Q15M1 = case_when(
+    Q15M1 == 1 ~ 1,
+    Q15M1 == 2 ~ 2,
+    Q15M1 == 3 ~ 3,
+    Q15M1 == 4 ~ 4,
+    Q15M1 == 97 ~ 5,
+    Q15M1 == 98 ~ NA,
+    Q15M1 == 99 ~ NA
+  ))
+
+```
+
+```{r}
+
+freq(Base_Trajipaq_Finale_Ok$Q15M1)
+
+```
+
+```{r}
+
+ggplot(data = Base_Trajipaq_Finale_Ok, mapping = aes(x = Q15M1)) +
+  geom_bar(fill = "pink", na.rm = TRUE) +
+  theme_bw()
+
+```
+
+Dans le tableau de fréquence et le diagramme de barres ci-dessus, on observe que la réponse modale est "travailleur temporaire".
+
+## Q16. Quel est votre plus haut diplôme d’études obtenu ?
+
+01. Moins que le secondaire
+02. Secondaire ou équivalent
+03. Collégial
+04. Formation professionnelle ou post-secondaire
+05. Université 1er cycle
+06. Université 2ème cycle et plus
+98. Je ne sais pas
+99. Je préfère ne pas répondre
+
+Type de variable = factorielle
+Niveau de mesure = ordinal
+
+```{r}
+
+Base_Trajipaq_Finale_Ok <-
+  Base_Trajipaq_Finale_Ok %>%
+  mutate(Q16 = case_when(
+    Q16 == 1 ~ 1,
+    Q16 == 2 ~ 2,
+    Q16 == 3 ~ 3,
+    Q16 == 4 ~ 4,
+    Q16 == 5 ~ 5,
+    Q16 == 6 ~ 6,
+    Q16 == 98 ~ NA,
+    Q16 == 99 ~ NA
+  ))
+
+```
+
+```{r}
+
+freq(Base_Trajipaq_Finale_Ok$Q16)
+
+```
+
+```{r}
+
+ggplot(data = Base_Trajipaq_Finale_Ok, mapping = aes(x = Q16)) +
+  geom_bar(fill = "pink", na.rm = TRUE) +
+  theme_bw()
+
+```
+
+Dans le tableau de fréquence et le diagramme de barres ci-dessus, on observe que la distribution du niveau d'éducation est asymétrique vers la gauche.
+Autrement dit, l'échantillon est composé à 67.31% de personnes titulaires d'un diplôme universitaire.
+
+## Q17. Actuellement, avez-vous une religion ?
+
+00. Sans religion
+01. Catholique
+02. Protestante
+03. Évangéliste
+04. Mormone
+05. Témoins de Jéhovah
+06. Orthodoxe
+07. Autres Chrétiens
+08. Musulmans
+09. Juif
+10. Bouddhiste
+11. Hindou
+12. Animiste
+97. Autre
+98. Je ne sais pas
+99. Je préfère ne pas répondre
 
@@
 Autrement dit, les répondants estiment en moyenne avoir subit une faible discrimination (0.28).
+Lorsque l'on regarde la distribution de chaque question composant l'indice, le mode correspond à l'absence de discrimination (1).
+Or, l'indice présente une moyenne non nulle (0.28 > 0), suggérant que peu de répondants ne subissent aucune discrimination.

### 2023-05-22 Codes/Article - Sentiment d'appartenance.Rmd
Progression dans la présentation des variables socio-démographiques.

@@
 99. Je préfère ne pas répondre
+
+Type de variable = catégorielle
+Niveau de mesure = nominal
+
+```{r}
+
+Base_Trajipaq_Finale_Ok <-
+  Base_Trajipaq_Finale_Ok %>%
+  mutate(Q17 = case_when(
+    Q17 == 0 ~ 0,
+    Q17 == 1 ~ 1,
+    Q17 == 2 ~ 2,
+    Q17 == 3 ~ 3,
+    Q17 == 4 ~ 4,
+    Q17 == 5 ~ 5,
+    Q17 == 6 ~ 6,
+    Q17 == 7 ~ 7,
+    Q17 == 8 ~ 8,
+    Q17 == 9 ~ 9,
+    Q17 == 10 ~ 10,
+    Q17 == 11 ~ 11,
+    Q17 == 12 ~ 12,
+    Q17 == 97 ~ 13,
+    Q17 == 98 ~ NA,
+    Q17 == 99 ~ NA
+  ))
+
+```
+
+```{r}
+
+freq(Base_Trajipaq_Finale_Ok$Q17)
+
+```
+
+```{r}
+
+ggplot(data = Base_Trajipaq_Finale_Ok, mapping = aes(x = Q17)) +
+  geom_bar(fill = "pink", na.rm = TRUE) +
+  theme_bw()
+
+```
+
+Le tableau de fréquence et le diagramme de barres ci-dessus nous montrent que les réponses modales sont "sans religion" et "catholique".
+On remarque également un nombre remarquable de réponses "orthodoxe" et "musulman".
+
+## Q18. Quelle importance accordez-vous aujourd’hui à la religion dans votre vie ?
+
+01. Pas du tout d’importance
+02. Un peu d’importance
+03. Assez d’importance
+04. Beaucoup d’importance
+98. Je ne sais pas
+99. Je préfère ne pas répondre
+
+Type de variable = factoriel
+Niveau de mesure = ordinal
+
+```{r}
+
+Base_Trajipaq_Finale_Ok <-
+  Base_Trajipaq_Finale_Ok %>%
+  mutate(Q18 = case_when(
+    Q18 == 1 ~ 1,
+    Q18 == 2 ~ 2,
+    Q18 == 3 ~ 3,
+    Q18 == 4 ~ 4,
+    Q18 == 98 ~ NA,
+    Q18 == 99 ~ NA
+  ))
+
+```
+
+```{r}
+
+freq(Base_Trajipaq_Finale_Ok$Q18)
+
+```
+
+```{r}
+
+ggplot(data = Base_Trajipaq_Finale_Ok, mapping = aes(x = Q18)) +
+  geom_bar(fill = "pink", na.rm = TRUE) +
+  theme_bw()
+
+```
+
+Sur le tableau de fréquence et le diagramme de barres ci-dessus, on observe que la distribution est asymétrique sur la droite.
+Autrement dit, 51.7% des répondants disent que la religion occupe une place dans leur vie.
+Néanmoins, le nombre élevé de réponse "je ne sais pas" (39.8%) réduit la validité de cette question.
+
+## Q20. Quand vous étiez enfant, quelles étaient la langue ou les langues que vous parliez le plus souvent, à l’âge de 5 ans et que vous comprenez toujours? (Plusieurs réponses possibles).
+
+01. Français
+02. Anglais
+97. Autre : préciser, possibilité d’en déclarer deux
+98. Je ne sais pas
+99. Je préfère ne pas répondre
+
+Type de variable = catégoriel
+Niveau de mesure = nominal
+
+```{r}
+
+Base_Trajipaq_Finale_Ok <-
+  Base_Trajipaq_Finale_Ok %>%
+  mutate(Q20M1 = case_when(
+    Q20M1 == 1 ~ 1,
+    Q20M1 == 2 ~ 2,
+    Q20M1 == 97 ~ 3,
+    Q20M1 == 98 ~ NA,
+    Q20M1 == 99 ~ NA
+  ))
+
+```
+
+```{r}
+freq(Base_Trajipaq_Finale_Ok$Q20M1)
+```
+
+```{r}
+
+ggplot(data = Base_Trajipaq_Finale_Ok, mapping = aes(x = Q20M1)) +
+  geom_bar(fill = "pink", na.rm = TRUE) +
+  theme_bw()
+
+```
+
+Dans le tableau de fréquence et le diagramme de barres ci-dessus, on observe que la réponse modale est "français". Autrement dit, le français constitue la langue maternelle de 63.84% des répondants.
+
+## I3. Score de compétence linguistique (français)
+
+Type de variable = factoriel
+Niveau de mesure = ordinal
+
+```{r}
+
+Base_Trajipaq_Finale_Ok <-
+  Base_Trajipaq_Finale_Ok %>%
+  mutate(Q21A = case_when(
+    Q21A == 1 ~ 1,
+    Q21A == 2 ~ 2,
+    Q21A == 3 ~ 3,
+    Q21A == 4 ~ 4,
+    Q21A == 98 ~ NA,
+    Q21A == 99 ~ NA
+  ))
+
+Base_Trajipaq_Finale_Ok <-
+  Base_Trajipaq_Finale_Ok %>%
+  mutate(Q21B = case_when(
+    Q21B == 1 ~ 1,
+    Q21B == 2 ~ 2,
+    Q21B == 3 ~ 3,
+    Q21B == 4 ~ 4,
+    Q21B == 98 ~ NA,
+    Q21B == 99 ~ NA
+  ))
+
+Base_Trajipaq_Finale_Ok <-
+  Base_Trajipaq_Finale_Ok %>%
+  mutate(Q21C = case_when(
+    Q21C == 1 ~ 1,
+    Q21C == 2 ~ 2,
+    Q21C == 3 ~ 3,
+    Q21C == 4 ~ 4,
+    Q21C == 98 ~ NA,
+    Q21C == 99 ~ NA
+  ))
+
+Base_Trajipaq_Finale_Ok <-
+  Base_Trajipaq_Finale_Ok %>%
+  mutate(Q21D = case_when(
+    Q21D == 1 ~ 1,
+    Q21D == 2 ~ 2,
+    Q21D == 3 ~ 3,
+    Q21D == 4 ~ 4,
+    Q21D == 98 ~ NA,
+    Q21D == 99 ~ NA
+  ))
+
+```
+
+```{r}
+
+v_I3 <- c("Q21A",
+          "Q21B",
+          "Q21C",
+          "Q21D"
+          )
+
+Base_Trajipaq_Finale_Ok <-
+  Base_Trajipaq_Finale_Ok %>%
+  rowwise() %>%
+  mutate(I3 = (sum(c_across(all_of(v_I3)), na.rm = TRUE) - length(v_I3)) / (3*length(v_I3))) %>%
+  mutate(I3 = if (I3 < 0) NA_real_ else I3) %>%
+  ungroup()
+
+```
+
+```{r}
+
+descr(Base_Trajipaq_Finale_Ok$I3)
+
+```
+
+```{r}
+
+ggplot(data = Base_Trajipaq_Finale_Ok, mapping = aes(x = I3)) +
+  geom_histogram(na.rm = TRUE, fill = "pink", binwidth = 0.01) +
+  theme_bw()
+
+```
 

### 2023-05-22 Codes/Article - Sentiment d'appartenance.Rmd
Fin de la présentation des variables.

@@
 - pays d'origine
-- année d'obtention de la résidence permanente
-- année d'obtention de la citoyenneté canadienne
-- année d'arrivée au Québec
-- année depuis la première installation au Québec
-- statut d'immigration avant l'obtention de la résidence permanente
+- statut d'immigration au Québec
+- statut d'immigration pré-résidence
 - niveau d'éducation
 - religion
-- importance accordée à la religion
-- compétence : lire le français
-- compétence : parler français
-- compétence : rédiger en français
-- compétence : comprendre le français
-- niveau de connaissance en français lors de l'arrivée au Québec
-- compétence : lire l'anglais
-- compétence : parler anglais
-- compétence : rédiger en anglais
-- compétence : comprendre l'anglais
-- niveau de connaissance en anglais lors de l'arrivée au Québec
-- suivi de cours dédiés à la francisation après l'arrivée au Québec
-- état de santé
-- problèmes de santé avant l'installation au Québec
-- problèmes de santé au cours des 12 derniers mois
-- problèmes de santé il y a moins de 5 ans
-- problèmes de santé il y a plus de 10 ans
-- lieu de naissance de la mère
+- compétence linguistique (français)
+- compétence linguistique (anglais)
+- état de santé actuel
+- état de santé depuis 12 mois
 - niveau d'éducation de la mère
-- lieu de naissance du père
 - niveau d'éducation du père
@@
 
-Dans le tableau de fréquence et le diagramme de barres ci-dessus, on observe que le sexe modal est "Femme".
+La réponse modale est "femme" (56.69%).
 
@@
 
-Dans le tableau de fréquence et le diagramme de barres ci-dessus, on observe que le statut modal est "Marié(e)".
+La réponse modale est "Marié(e)" (47.53%)
 
@@
 
-Dans le tableau de fréquence et le diagramme de barres ci-dessous, on peut observer que les pays d'origine modaux sont la France (26.685%) et l'Algérie (6.649%).
-
-## Q12. En quelle année avez-vous obtenu la résidence permanente ?
-
-Année
-9998. Je ne sais pas
-9999. Je préfère ne pas répondre
-
-Type de variable = catégorielle
-Niveau de mesure = nominale
-
-## Q13. En quelle année est-ce que vous êtes devenu citoyen(ne) ?
-
-Année
-9998. Je ne sais pas
-9999. Je préfère ne pas répondre
-
-Type de variable = numérique continue
-Niveau de mesure = ratio
-
-## Q14I. Depuis quelle année vivez-vous au Québec ?
-
-Année
-9996. La même année que la résidence permanente
-9998. Je ne sais pas
-9999. Je préfère ne pas répondre
-
-Type de variable = numérique continue
-Niveau de mesure = ratio
-
-## Q14N. En quelle année avez-vous commencé à vivre au Québec pendant au moins un an ?
-
-Année
-9996. Depuis la naissance
-9998. Je ne sais pas
-9999. Je préfère ne pas répondre
-
-Type de variable = numérique continue
-Niveau de mesure = ratio
-
-## Q15I. Avant l’obtention de la résidence permanente, quel était votre statut d’immigration ?
-
-1. Travailleur temporaire
+Les réponses modales sont "France" (26.685%) et "Algérie" (6.649%).
+
+## Q14N. Statut migratoire
+
+Type de variable = catégoriel
+Niveau de mesure = Nominal
+
+```{r}
+
+Base_Trajipaq_Finale_Ok <-
+  Base_Trajipaq_Finale_Ok %>%
+  mutate(Q14N = case_when(
+    Q14N < 2017 ~ 1,
+    Q14N == 9996 ~ 2,
+    Q14N == 9998 ~ NA,
+    Q14N == 9999 ~ NA
+  ))
+
+```
+
+```{r}
+
+freq(Base_Trajipaq_Finale_Ok$Q14N)
+
+```
+
+```{r}
+
+ggplot(data = Base_Trajipaq_Finale_Ok, mapping = aes(x = Q14N)) +
+  geom_bar(fill = "pink", na.rm = TRUE) +
+  theme_bw()
+
+```
+
+La réponse modale est "deuxième génération" (92.87%).
+
+## Q15I. Statut migratoire avant d'obtenir la résidence permanente
+
+01. Travailleur temporaire (visa de travail ou permis de
+travail)
 02. Demandeur d’asile
-03. Étudiant étranger
+03. Étudiant étranger (visa d’études)
 04. Touriste
-97. Autre, préciser
+97. Autre
 98. Je ne sais pas
@@
 
-Type de variable = catégorielle
+Type de variable = catégoriel
 Niveau de mesure = nominal
@@
 
-Dans le tableau de fréquence et le diagramme de barres ci-dessus, on observe que la réponse modale est "travailleur temporaire".
+La réponse modale est "travailleur temporaire" (57.78%).
 
@@
 
-Dans le tableau de fréquence et le diagramme de barres ci-dessus, on observe que la distribution du niveau d'éducation est asymétrique vers la gauche.
-Autrement dit, l'échantillon est composé à 67.31% de personnes titulaires d'un diplôme universitaire.
+La distribution du niveau d'éducation est asymétrique vers la gauche. 
+L'échantillon est composé à 67.31% de personnes titulaires d'un diplôme universitaire.
 
@@
 
-Le tableau de fréquence et le diagramme de barres ci-dessus nous montrent que les réponses modales sont "sans religion" et "catholique".
-On remarque également un nombre remarquable de réponses "orthodoxe" et "musulman".
-
-## Q18. Quelle importance accordez-vous aujourd’hui à la religion dans votre vie ?
-
-01. Pas du tout d’importance
-02. Un peu d’importance
-03. Assez d’importance
-04. Beaucoup d’importance
-98. Je ne sais pas
-99. Je préfère ne pas répondre
-
-Type de variable = factoriel
-Niveau de mesure = ordinal
-
-```{r}
-
-Base_Trajipaq_Finale_Ok <-
-  Base_Trajipaq_Finale_Ok %>%
-  mutate(Q18 = case_when(
-    Q18 == 1 ~ 1,
-    Q18 == 2 ~ 2,
-    Q18 == 3 ~ 3,
-    Q18 == 4 ~ 4,
-    Q18 == 98 ~ NA,
-    Q18 == 99 ~ NA
-  ))
-
-```
-
-```{r}
-
-freq(Base_Trajipaq_Finale_Ok$Q18)
-
-```
-
-```{r}
-
-ggplot(data = Base_Trajipaq_Finale_Ok, mapping = aes(x = Q18)) +
-  geom_bar(fill = "pink", na.rm = TRUE) +
-  theme_bw()
-
-```
-
-Sur le tableau de fréquence et le diagramme de barres ci-dessus, on observe que la distribution est asymétrique sur la droite.
-Autrement dit, 51.7% des répondants disent que la religion occupe une place dans leur vie.
-Néanmoins, le nombre élevé de réponse "je ne sais pas" (39.8%) réduit la validité de cette question.
+Les réponses modales sont "sans religion" (36.36) et "catholique" (32.21%). 
+On remarque également un nombre significatif de réponses "musulman" (13.83%).
 
@@
 
-Dans le tableau de fréquence et le diagramme de barres ci-dessus, on observe que la réponse modale est "français". Autrement dit, le français constitue la langue maternelle de 63.84% des répondants.
-
-## I3. Score de compétence linguistique (français)
-
-Type de variable = factoriel
-Niveau de mesure = ordinal
+La réponse modale est "français" (63.84%).
+
+## I3. Indice de la compétence linguistique (français)
+
+Type de variable = numeric
+Niveau de mesure = ratio
 
@@
 
+La distribution est asymétrique vers la gauche.
+L'indice moyen de maîtrise du français est 0.87.
+
+## I4. Indice de la compétence linguistique (anglais).
+
+Type de variable = numeric
+Niveau de mesure = ratio
+
+```{r}
+
+Base_Trajipaq_Finale_Ok <-
+  Base_Trajipaq_Finale_Ok %>%
+  mutate(Q22A = case_when(
+    Q22A == 1 ~ 1,
+    Q22A == 2 ~ 2,
+    Q22A == 3 ~ 3,
+    Q22A == 4 ~ 4,
+    Q22A == 98 ~ NA,
+    Q22A == 99 ~ NA
+  ))
+
+Base_Trajipaq_Finale_Ok <-
+  Base_Trajipaq_Finale_Ok %>%
+  mutate(Q22B = case_when(
+    Q22B == 1 ~ 1,
+    Q22B == 2 ~ 2,
+    Q22B == 3 ~ 3,
+    Q22B == 4 ~ 4,
+    Q22B == 98 ~ NA,
+    Q22B == 99 ~ NA
+  ))
+
+Base_Trajipaq_Finale_Ok <-
+  Base_Trajipaq_Finale_Ok %>%
+  mutate(Q22C = case_when(
+    Q22C == 1 ~ 1,
+    Q22C == 2 ~ 2,
+    Q22C == 3 ~ 3,
+    Q22C == 4 ~ 4,
+    Q22C == 98 ~ NA,
+    Q22C == 99 ~ NA
+  ))
+
+Base_Trajipaq_Finale_Ok <-
+  Base_Trajipaq_Finale_Ok %>%
+  mutate(Q22D = case_when(
+    Q22D == 1 ~ 1,
+    Q22D == 2 ~ 2,
+    Q22D == 3 ~ 3,
+    Q22D == 4 ~ 4,
+    Q22D == 98 ~ NA,
+    Q22D == 99 ~ NA
+  ))
+
+```
+
+```{r}
+
+v_I4 <- c("Q22A",
+          "Q22B",
+          "Q22C",
+          "Q22D"
+          )
+
+Base_Trajipaq_Finale_Ok <-
+  Base_Trajipaq_Finale_Ok %>%
+  rowwise() %>%
+  mutate(I4 = (sum(c_across(all_of(v_I4)), na.rm = TRUE) - length(v_I4)) / (3*length(v_I4))) %>%
+  mutate(I4 = if (I4 < 0) NA_real_ else I4) %>%
+  ungroup()
+
+```
+
+```{r}
+
+descr(Base_Trajipaq_Finale_Ok$I4)
+
+```
+
+```{r}
+
+ggplot(data = Base_Trajipaq_Finale_Ok, mapping = aes(x = I4)) +
+  geom_histogram(na.rm = TRUE, fill = "pink", binwidth = 0.01) +
+  theme_bw()
+
+```
+
+La distribution est asymétrique vers la gauche.
+L'indice moyen de maîtrise du français est 0.73.
+
+## Q23. Comment est votre état de santé en général ?
+
+01. Très bon
+02. Bon
+03. Moyen
+04. Mauvais
+05. Très mauvais
+98. Je ne sais pas
+99. Je préfère ne pas répondre
+
+Type de variable = factoriel
+Niveau de mesure = ordinal
+
+```{r}
+
+Base_Trajipaq_Finale_Ok <-
+  Base_Trajipaq_Finale_Ok %>%
+  mutate(Q23 = case_when(
+    Q23 == 1 ~ 5,
+    Q23 == 2 ~ 4,
+    Q23 == 3 ~ 3,
+    Q23 == 4 ~ 2,
+    Q23 == 5 ~ 1,
+    Q23 == 98 ~ NA,
+    Q23 == 99 ~ NA
+  ))
+
+```
+
+```{r}
+
+freq(Base_Trajipaq_Finale_Ok$Q23)
+
+```
+
+```{r}
+ggplot(data = Base_Trajipaq_Finale_Ok, mapping = aes(x = Q23)) +
+  geom_bar(fill = "pink", na.rm = TRUE) +
+  theme_bw()
+```
+
+La distribution est asymétrique vers la gauche. 
+98.47% des répondants ont un état de santé moyen ou bon.
+
+## I5. Indice de l'état de santé au cours des 12 derniers mois
+
+```{r}
+
+Base_Trajipaq_Finale_Ok <-
+  Base_Trajipaq_Finale_Ok %>%
+  mutate(Q24A = case_when(
+    Q24A == 1 ~ 1,
+    Q24A == 2 ~ 2,
+    Q24A == 96 ~ NA,
+    Q24A == 97 ~ NA,
+    Q24A == 98 ~ NA,
+    Q24A == 99 ~ NA
+  ))
+
+Base_Trajipaq_Finale_Ok <-
+  Base_Trajipaq_Finale_Ok %>%
+  mutate(Q24B = case_when(
+    Q24B == 1 ~ 1,
+    Q24B == 2 ~ 2,
+    Q24B == 96 ~ NA,
+    Q24B == 97 ~ NA,
+    Q24B == 98 ~ NA,
+    Q24B == 99 ~ NA
+  ))
+
+Base_Trajipaq_Finale_Ok <-
+  Base_Trajipaq_Finale_Ok %>%
+  mutate(Q24C = case_when(
+    Q24C == 1 ~ 1,
+    Q24C == 2 ~ 2,
+    Q24C == 96 ~ NA,
+    Q24C == 97 ~ NA,
+    Q24C == 98 ~ NA,
+    Q24C == 99 ~ NA
+  ))
+
+Base_Trajipaq_Finale_Ok <-
+  Base_Trajipaq_Finale_Ok %>%
+  mutate(Q24D = case_when(
+    Q24D == 1 ~ 1,
+    Q24D == 2 ~ 2,
+    Q24D == 96 ~ NA,
+    Q24D == 97 ~ NA,
+    Q24D == 98 ~ NA,
+    Q24D == 99 ~ NA
+  ))
+
+Base_Trajipaq_Finale_Ok <-
+  Base_Trajipaq_Finale_Ok %>%
+  mutate(Q24E = case_when(
+    Q24E == 1 ~ 1,
+    Q24E == 2 ~ 2,
+    Q24E == 96 ~ NA,
+    Q24E == 97 ~ NA,
+    Q24E == 98 ~ NA,
+    Q24E == 99 ~ NA
+  ))
+
+```
+
+```{r}
+
+v_I5 <- c("Q24A",
+          "Q24B",
+          "Q24C",
+          "Q24D",
+          "Q24E"
+          )
+
+Base_Trajipaq_Finale_Ok <-
+  Base_Trajipaq_Finale_Ok %>%
+  rowwise() %>%
+  mutate(I5 = (sum(c_across(all_of(v_I5)), na.rm = TRUE) - length(v_I5)) / (1*length(v_I5))) %>%
+  mutate(I5 = if (I5 < 0) NA_real_ else I5) %>%
+  ungroup()
+
+```
+
+```{r}
+
+descr(Base_Trajipaq_Finale_Ok$I5)
+
+```
+
+```{r}
+
+ggplot(data = Base_Trajipaq_Finale_Ok, mapping = aes(x = I5)) +
+  geom_histogram(na.rm = TRUE, fill = "pink", binwidth = 0.01) +
+  theme_bw()
+
+```
+
+La distribution est asymétrique vers la gauche. 
+L'indice moyen de l'état de santé au cours des 12 derniers mois est de 0.78.
+
+## Q26. Quel était le plus haut diplôme d’études obtenu par votre mère ?
+
+01. Moins que le secondaire
+02. Secondaire ou équivalent
+03. Collégial
+04. Formation professionnelle ou post-secondaire
+05. Université
+06. Inconnu
+98. Je ne sais pas
+99. Je préfère ne pas répondre
+
+Type de variable = factoriel
+Niveau de mesure = ordinal
+
+```{r}
+
+Base_Trajipaq_Finale_Ok <-
+  Base_Trajipaq_Finale_Ok %>%
+  mutate(Q26 = case_when(
+    Q26 == 1 ~ 1,
+    Q26 == 2 ~ 2,
+    Q26 == 3 ~ 3,
+    Q26 == 4 ~ 4,
+    Q26 == 5 ~ 5,
+    Q26 == 6 ~ NA,
+    Q26 == 98 ~ NA,
+    Q26 == 99 ~ NA
+  ))
+
+```
+
+```{r}
+
+freq(Base_Trajipaq_Finale_Ok$Q26)
+
+```
+
+```{r}
+
+ggplot(data = Base_Trajipaq_Finale_Ok, mapping = aes(x = Q26)) +
+  geom_bar(fill = "pink", na.rm = TRUE) +
+  theme_bw()
+
+```
+
+La valeur modale est "diplôme universitaire" (28.62%).
+
+## Q28. Quel était le plus haut diplôme d’études obtenu par votre père ?
+
+01. Moins que le secondaire
+02. Secondaire ou équivalent
+03. Collégial
+04. Formation professionnelle ou post-secondaire
+05. Université
+06. Inconnu
+98. Je ne sais pas
+99. Je préfère ne pas répondre
+
+Type de variable = factoriel
+Niveau de mesure = ordinal
+
+```{r}
+
+Base_Trajipaq_Finale_Ok <-
+  Base_Trajipaq_Finale_Ok %>%
+  mutate(Q28 = case_when(
+    Q28 == 1 ~ 1,
+    Q28 == 2 ~ 2,
+    Q28 == 3 ~ 3,
+    Q28 == 4 ~ 4,
+    Q28 == 5 ~ 5,
+    Q28 == 6 ~ NA,
+    Q28 == 98 ~ NA,
+    Q28 == 99 ~ NA
+  ))
+
+```
+
+```{r}
+
+freq(Base_Trajipaq_Finale_Ok$Q28)
+
+```
+
+```{r}
+
+ggplot(data = Base_Trajipaq_Finale_Ok, mapping = aes(x = Q28)) +
+  geom_bar(fill = "pink", na.rm = TRUE) +
+  theme_bw()
+
+```
+
+La valeur modale est "diplôme universitaire" (34.70%).
+
 # Présentation de la variable indépendante
@@
 
-Dans le tableau de fréquence ci-dessus, on observe que la distribution est asymétrique vers la gauche.
-Autrement dit, 87.01% des personnes interrogées sont tout à fait d'accord ou plutôt d'accord pour dire qu'elles se sentent chez elles au Québec.
-On remarque également que 1.79% des personnes interrogées ne savaient pas ou n'ont pas souhaité répondre à la question.
-Etant donné la faible proportion de ces réponses, celles-ci seront retirées de l'analyse.
-
 ```{r}
@@
 
-Dans le diagramme de barres ci-dessus, on observe l'asymétrie vers la gauche mentionnée précédemment.
+La distribution est asymétrique vers la gauche. 
+87.01% des personnes interrogées sont tout à fait d'accord ou plutôt d'accord pour dire qu'elles se sentent chez elles au Québec.
 
@@
 
-Dans le tableau de fréquence ci-dessus, on observe que la distribution est asymétrique vers la droite.
-Autrement dit, 67.88% des personnes interrogées ne sont pas du tout d'accord ou plutôt pas d'accord pour dire qu'il faut faire oublier ses origines pour se faire accepter au Québec.
-On remarque également que 3.96% des personnes interrogées ne savaient pas ou n'ont pas souhaité répondre à la question.
-Etant donné la faible proportion de ces réponses, celles-ci seront retirées de l'analyse.
-
 ```{r}
@@
 
-Dans le diagramme de barres ci-dessus, on observe l'asymétrie vers la droite mentionnée précédemment.
+La distribution est asymétrique vers la droite.
+67.88% des personnes interrogées ne sont pas du tout d'accord ou plutôt pas d'accord pour dire qu'il faut faire oublier ses origines pour se faire accepter au Québec.
 
@@
 
-Dans le tableau de fréquence ci-dessus, on observe que la distribution est bimodale (1 et 4).
-Autrement dit, cette question divise nos répondants.
-On remarque également que 5.63% des personnes interrogées ne savaient pas ou n'ont pas souhaité répondre à la question.
-Etant donné la faible proportion de ces réponses, celles-ci seront retirées de l'analyse.
-
 ```{r}
@@
 
-Dans le diagramme de barres ci-dessus, on observe les deux modes de la distribution décrits précédemment.
+Les réponses modales sont "tout à fait d'accord" (29.36%) et "pas du tout d'accord" (29.97%).
 
@@
 
-Dans le tableau de fréquence ci-dessus, on observe que la distribution est asymétrique vers la gauche.
-Autrement dit, 60.72% des personnes interrogées sont tout à fait d'accord ou plutôt d'accord pour dire qu'elles se sentent Québécois(e)s.
-On remarque également que 4.03% des personnes interrogées ne savaient pas ou n'ont pas souhaité répondre à la question.
-Etant donné la faible proportion de ces réponses, celles-ci seront retirées de l'analyse.
-
 ```{r}
@@
 
-Dans le diagramme de barres ci-dessus, on observe l'asymétrie vers la gauche mentionnée précédemment.
+La distribution est asymétrique vers la gauche. 
+60.72% des personnes interrogées sont tout à fait d'accord ou plutôt d'accord pour dire qu'elles se sentent Québécois(e)s.
 
@@
 
-Dans le tableau de fréquence ci-dessus, on observe que la distribution est asymétrique vers la gauche.
-Autrement dit, 79.98% des personnes interrogées sont tout à fait d'accord ou plutôt d'accord pour dire qu'elles se sentent Canadien(ne).
-On remarque également que 2.69% des personnes interrogées ne savaient pas ou n'ont pas souhaité répondre à la question.
-Etant donné la faible proportion de ces réponses, celles-ci seront retirées de l'analyse.
-
 ```{r}
@@
 
-Dans le diagramme de barres ci-dessus, on observe l'asymétrie vers la gauche mentionnée précédemment.
+La distribution est asymétrique vers la gauche. 
+79.98% des personnes interrogées sont tout à fait d'accord ou plutôt d'accord pour dire qu'elles se sentent Canadien(ne).
 
@@
 
-Dans le tableau de fréquence ci-dessus, on observe que la distribution est asymétrique vers la gauche.
-Autrement dit, 81.42% des personnes interrogées sont tout à fait d'accord ou plutôt d'accord pour dire qu'elles possèdent un sentiment d'appartenance envers leur pays d'origine.
-On remarque également que 2.69% des personnes interrogées ne savaient pas ou n'ont pas souhaité répondre à la question.
-Etant donné la faible proportion de ces réponses, celles-ci seront retirées de l'analyse.
-
 ```{r}
@@
 
-Dans le diagramme de barres ci-dessus, on observe l'asymétrie vers la gauche mentionnée précédemment.
+La distribution est asymétrique vers la gauche. 
+81.42% des personnes interrogées sont tout à fait d'accord ou plutôt d'accord pour dire qu'elles possèdent un sentiment d'appartenance envers leur pays d'origine.
 
@@
 
-Le tableau descriptif et l'histogramme ci-dessous nous indiquent que la distribution est asymétrique vers la gauche.
-Autrement dit, les répondants estiment en moyenne entretenir un sentiment d'appartenance relativement développé (0.59).
+La distribution est asymétrique vers la gauche. 
+L'indice moyen du sentiment d'appartenance au Québec est 0.59.
 
@@
 
-Dans le tableau de fréquence ci-dessus, on observe que la distribution est asymétrique vers la droite Autrement dit, 65.44% des personnes interrogées disent ne jamais avoir subit de discrimination dans l'emploi au cours des 12 derniers mois précédent la crise du covid 19.
-On remarque également que 9.47% des personnes interrogées ne savaient pas ou n'ont pas souhaité répondre à la question.
-Etant donné la faible proportion de ces réponses, celles-ci seront retirées de l'analyse.
-
 ```{r}
@@
 
-Dans le diagramme de barres ci-dessus, on observe l'asymétrie vers la droite mentionnée précédemment.
+La distribution est asymétrique vers la droite.
+65.44% des personnes interrogées disent ne jamais avoir subit de discrimination dans l'emploi au cours des 12 derniers mois précédent la crise du covid 19.
 
@@
 
-Dans le tableau de fréquence ci-dessus, on observe que la distribution est asymétrique vers la droite Autrement dit, 79.64% des personnes interrogées disent ne jamais avoir subit de discrimination dans la recherche de logement au cours des 12 derniers mois précédent la crise du covid 19.
-On remarque également que 17.34% des personnes interrogées ne savaient pas ou n'ont pas souhaité répondre à la question.
-Etant donné la faible proportion de ces réponses, celles-ci seront retirées de l'analyse.
-
 ```{r}
@@
 
-Dans le diagramme de barres ci-dessus, on observe l'asymétrie vers la droite mentionnée précédemment.
+La distribution est asymétrique vers la droite.
+79.64% des personnes interrogées disent ne jamais avoir subit de discrimination dans la recherche de logement au cours des 12 derniers mois précédent la crise du covid 19.
 
@@
 
-Dans le tableau de fréquence ci-dessus, on observe que la distribution est asymétrique vers la droite Autrement dit, 82.26% des personnes interrogées disent ne jamais avoir subit de discrimination dans la santé au cours des 12 derniers mois précédent la crise du covid 19.
-On remarque également que 10.17% des personnes interrogées ne savaient pas ou n'ont pas souhaité répondre à la question.
-Etant donné la faible proportion de ces réponses, celles-ci seront retirées de l'analyse.
-
 ```{r}
@@
 
-Dans le diagramme de barres ci-dessus, on observe l'asymétrie vers la droite mentionnée précédemment.
+La distribution est asymétrique vers la droite.
+82.26% des personnes interrogées disent ne jamais avoir subit de discrimination dans la santé au cours des 12 derniers mois précédent la crise du covid 19.
 
@@
 
-Dans le tableau de fréquence ci-dessus, on observe que la distribution est asymétrique vers la droite Autrement dit, 81.40% des personnes interrogées disent ne jamais avoir subit de discrimination sur le lieu de formation au cours des 12 derniers mois précédent la crise du covid 19.
-On remarque également que 15.74% des personnes interrogées ne savaient pas ou n'ont pas souhaité répondre à la question.
-Etant donné la faible proportion de ces réponses, celles-ci seront retirées de l'analyse.
-
 ```{r}
@@
 
-Dans le diagramme de barres ci-dessus, on observe l'asymétrie vers la droite mentionnée précédemment.
+La distribution est asymétrique vers la droite.
+81.40% des personnes interrogées disent ne jamais avoir subit de discrimination sur le lieu de formation au cours des 12 derniers mois précédent la crise du covid 19.
 
@@
 
-Le tableau descriptif et l'histogramme ci-dessous nous indiquent que la distribution est asymétrique vers la droite.
-Autrement dit, les répondants estiment en moyenne avoir subit une faible discrimination (0.28).
-Lorsque l'on regarde la distribution de chaque question composant l'indice, le mode correspond à l'absence de discrimination (1).
+La distribution est asymétrique vers la droite. 
+L'indice moyen de discrimination subie est 0.28.
+
+```{r}
+
+# Lorsque l'on regarde la distribution de chaque question composant l'indice, le mode correspond à l'absence de discrimination (1).
 Or, l'indice présente une moyenne non nulle (0.28 > 0), suggérant que peu de répondants ne subissent aucune discrimination.
+
+```

### 2023-05-22 Article/Visseho_al_sentiment_appartenance.Rmd

---
title: "Sentiment d'appartenance des immigrants"
author:
- Vissého Adjiwanou ^[corresponding author - adjiwanou.vissého@uqam.ca]
- Département de Sociologie, Université du Québec à Montréal (UQAM)
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  word_document:
    keep_md: yes
  html_document:
    toc: yes
    toc_depth: 3
  template: svm-latex-ms2.tex
  pdf_document: default
  citation_package: natbib
bibliography:
- covid_Nima.bib
- Covid_Canada.bib
- Covid_Family.bib
---
<!--
# Journal: family issues - SAGE
https://journals-sagepub-com.proxy.bibliotheques.uqam.ca/author-instructions/JFI

https://fishandwhistle.net/post/2020/getting-started-zotero-better-bibtex-rmarkdown/
https://www.zotero.org/styles?fields=sociology
-->

\newpage

# Résumé

Les immigrants sont souvent pris entre deux ou plusieurs cultures, celles de leur pays d'origine et celles de leur nouveau pays de résidence. Cet article essaie de comprendre comment ces sentiments sont associés à leur pays d'origine et aux éventuelles situations de discriminations vécues qu'ils vivent dans leur pays d'accueil

## Variables à utiliser

**Variables dépendantes**

- Crée la variable Sentiment d'appartenance qui prend les valeurs suivantes (selon les terminologies utilisées par Berry et Hou, 2017)

  - Exclusivement au pays de naissance (separated, 5%)
  - Exclusivement au pays d'accueil (assimilated, 15%)
  - Aux deux  (integrated, 75%)
  - Ni l'un ni l'autre (marginalised, 6%)

**Quelle est la distribution de cette nouvelle variable?**

Cette variable est crée par combinaison des deux variables suivantes:

  - Sentiment d'appartenance au Canada
  - Sentiment d'appartenance à leur pays d'origine

**Variables indépendantes**

- Origine de l'immigrant
  - Régions d'origine
  - Proximité du pays d'accueil par rapport au pays d'origine 
  - Niveau de stabilité dans le pays d'accueil ou d'origine au moment de l'enquête

- Discrimination perçue

# références

<!--La recherche montre que l’identification au pays d’origine et l’identification au pays d’accueil ne sont pas nécessairement et mutuellement exclusives (Berry, 1997; Zimmermann et al., 2007) et que les immigrants et leurs descendants se débattent sur la question de la combinaison de l’identification ethnique et nationale (Leszczensky, 2013; Phinney et al., 2006; Portes et Rumbaut, 2001; Verkuyten et Yildiz, 2007). Les immigrants qui arrivent dans un nouveau pays prennent une décision non seulement sur l’adaptation à la culture et à la société d’accueil, mais aussi sur le niveau d’attachement à la culture et à la société d’origine qu’ils veulent conserver. De ce fait, le modèle bipolaire unidimensionnel (Eisenstadt, 1954), suggère que les immigrants arrivent dans le pays d’accueil avec une seule identité, celle associée à leur pays d’origine, mais cette identité s’affaiblit et qu’au fil du temps, les immigrants acquièrent progressivement l’identité associée au pays d’accueil (Ngo et Li, 2016; Tartakovsky, 2013). On s’attend donc à ce que les immigrants libèrent progressivement les attachements antérieurs envers leur pays d’origine, et développent un sentiment d’appartenance au pays d’accueil (Amit et Bar-Lev, 2015). Cependant, des études indiquent que les immigrants n’abandonnent pas simplement leurs affinités culturelles avec leur pays d’origine mais les utilisent comme outils d’interprétation dans la construction de leur identité individuelle et collective dans la nouvelle société (Amit et Bar-Lev, 2015). À tout moment après leur arrivée dans le pays d’accueil, les immigrants peuvent également avoir une combinaison d’engagements envers la société d’accueil et envers la société d’origine ou adopter une identité biculturelle (Berry, 1997, 1994; Phinney et al., 2006; Verkuyten et Martinovic, 2012). Dès lors, les immigrants parviennent rarement à une identité nationale ou biculturelle et adoptent plus fréquemment une stratégie de séparation, c’est-à-dire, une grande valeur accordée au désir de conserver leur propre patrimoine culturel et, simultanément, peu d’intérêt à adopter la culture de la société d’accueil (Giuliani et al., 2018; Verkuyten et Martinovic, 2012).-->

- Leszczensky, L. (2013). Do national identification and interethnic friendships affect one another? A longitudinal test with adolescents of Turkish origin in Germany. Social Science Research, 42(3), 775‑788. https://doi.org/10.1016/j.ssresearch.2012.12.019

- Amit, K. et Bar-Lev, S. (2015). Immigrants’ Sense of Belonging to the Host Country: The Role of Life Satisfaction, Language Proficiency, and Religious Motives. Social Indicators Research, 124(3), 947‑961. https://doi.org/10.1007/s11205-014-0823-3

- Phinney, J. S., Berry, J. W., Vedder, P. et Liebkind, K. (2006). The Acculturation Experience: Attitudes, Identities and Behaviors of Immigrant Youth. Dans Immigrant youth in cultural transition: Acculturation, identity, and adaptation across national contexts (p. 71‑116). Lawrence Erlbaum Associates Publishers. https://doi.org/10.4324/9780415963619-4

- Verkuyten, M. et Yildiz, A. A. (2007). National (Dis)identification and Ethnic and Religious Identity: A Study Among Turkish-Dutch Muslims. Personality and Social Psychology Bulletin, 33(10), 1448‑1462. https://doi.org/10.1177/0146167207304276

- Ngo, H. et Li, H. (2016). Cultural Identity and Adaptation of Mainland Chinese Immigrants in Hong Kong. American Behavioral Scientist, 60(5), 730‑749. https://doi.org/10.1177/0002764216632837

- Giuliani, C., Tagliabue, S. et Regalia, C. (2018). Psychological Well-Being, Multiple Identities, and Discrimination Among First and Second Generation Immigrant Muslims. Europe’s Journal of Psychology, 14(1), 66‑87. https://doi.org/10.5964/ejop.v14i1.1434

- Phinney, J. S., Horenczyk, G., Liebkind, K. et Vedder, P. (2001). Ethnic Identity, Immigration, and Well-Being: An Interactional Perspective. Journal of Social Issues, 57(3), 493‑510. https://doi.org/10.1111/0022-4537.00225

- Phinney, J. S. et Ong, A. D. (2007). Conceptualization and measurement of ethnic identity: Current status and future directions. Journal of Counseling Psychology, 54, 271‑281. https://doi.org/10.1037/0022-0167.54.3.271

- Verkuyten, M. (2016). The Integration Paradox: Empiric Evidence From the Netherlands. American Behavioral Scientist, 60(5), 583‑596. https://doi.org/10.1177/0002764216632838

- Verkuyten, M. et Martinovic, B. (2012). Immigrants’ National Identification: Meanings, Determinants, and Consequences. Social Issues and Policy Review, 6(1), 82‑112. https://doi.org/10.1111/j.1751-2409.2011.01036.x

- Tartakovsky, E. (2013). Cultural identities of immigrants: Their formation in the pre-migration period and their transformation during adjustment to the host country. Immigration: Policies, Challenges and Impact, 231‑254.

- Zimmermann, L., Zimmermann, K. F. et Constant, A. (2007). Ethnic Self-Identification of First-Generation Immigrants. International Migration Review, 41(3), 769‑781. https://doi.org/10.1111/j.1747-7379.2007.00093.x

- Berry, J. W. (2006). Mutual attitudes among immigrants and ethnocultural groups in Canada. International Journal of Intercultural Relations, 30(6), 719‑734. https://doi.org/10.1016/j.ijintrel.2006.06.004

- Berry, J. W. (1997). Immigration, Acculturation, and Adaptation. Applied Psychology, 46(1), 5‑34. https://doi.org/10.1111/j.1464-0597.1997.tb01087.x

- Berry, J. W. (1994). Acculturation and Psychological Adaptation: An Overview. Dans Journeys into Cross-Cultural Psychology. (p. 129‑141). Swets & Zeitlinger Publishers.

- Berry, J. W. et Hou, F. (2017). Acculturation, discrimination and wellbeing among second generation of immigrants in Canada. International Journal of Intercultural Relations, 61, 29‑39. https://doi.org/10.1016/j.ijintrel.2017.08.003

- Berry, J. W. et Hou, F. (2021). Immigrant acculturation and wellbeing across generations and settlement contexts in Canada. International Review of Psychiatry, 33(1), 140‑153. https://doi.org/10.1080/09540261.2020.1750801

- Berry, J. W. et Sabatier, C. (2010). Acculturation, discrimination, and adaptation among second generation immigrant youth in Montreal and Paris. International Journal of Intercultural Relations, 34(3), 191‑207. https://doi.org/10.1016/j.ijintrel.2009.11.007

### 2023-05-22 Codes/Article - Sentiment d'appartenance.Rmd
Progression dans l'analyse.

@@
 - sexe
+- groupe d'âge
 - statut matrimonial
@@
 
+## Q8B A quel groupe d'âge appartenez-vous ?
+
+01. 0 – 19 ans
+02. 20 – 24 ans
+03. 25 – 29 ans
+04. 30 – 34 ans
+05. 35 – 39 ans
+06. 40 – 44 ans
+07. 45 – 49 ans
+08. 50 – 60 ans
+09. 61 ans et plus
+99. Je préfère ne pas répondre
+
+Type de variable = factoriel
+Niveau de mesure = ordinal
+
+```{r}
+
+freq(Base_Trajipaq_Finale_Ok$QAGE)
+
+```
+
+```{r}
+
+ggplot(data = Base_Trajipaq_Finale_Ok, mapping = aes(x = QAGE)) +
+  geom_bar(fill = "pink", na.rm = TRUE) +
+  theme_bw()
+
+```
+
+La distribution est symétrique.
+Le groupe d'âge modal est "35-39 ans" (24.63%).
+
 ## STATUT. Êtes-vous actuellement... ?
@@
 
-## Q14N. Statut migratoire
+## Q4. Êtes-vous né au Canada ?
 
 Type de variable = catégoriel
-Niveau de mesure = Nominal
-
-```{r}
-
-Base_Trajipaq_Finale_Ok <-
-  Base_Trajipaq_Finale_Ok %>%
-  mutate(Q14N = case_when(
-    Q14N < 2017 ~ 1,
-    Q14N == 9996 ~ 2,
-    Q14N == 9998 ~ NA,
-    Q14N == 9999 ~ NA
-  ))
-
-```
-
-```{r}
-
-freq(Base_Trajipaq_Finale_Ok$Q14N)
-
-```
-
-```{r}
-
-ggplot(data = Base_Trajipaq_Finale_Ok, mapping = aes(x = Q14N)) +
+Niveau de mesure = nominal
+
+```{r}
+
+freq(Base_Trajipaq_Finale_Ok$Q4)
+
+```
+
+```{r}
+
+ggplot(data = Base_Trajipaq_Finale_Ok, mapping = aes(x = Q4)) +
   geom_bar(fill = "pink", na.rm = TRUE) +
@@
 
-Les réponses modales sont "sans religion" (36.36) et "catholique" (32.21%). 
+Les réponses modales sont "sans religion" (36.36%) et "catholique" (32.21%). 
 On remarque également un nombre significatif de réponses "musulman" (13.83%).
@@
 
+## Résumé
+
+L'échantillon possède les caractéristiques socio-démographiques suivantes :
+
+- faible prépondérance de femmes (56.69%)
+- faible prépondérance de personnes âgées de 35 à 39 ans (24.63%)
+- forte prépondérance de personnes mariées (47.53%)
+- prépondérance de français (26.685) et d'algériens (6.649%)
+- forte prépondérance de personnes nées au Canada (71.21%)
+- forte prépondérance d'anciens travailleurs temporaires parmi les immigrés de première génération (57.79%)
+- forte prépondérance de personnes possédant un diplôme universitaire (67.31%)
+- prépondérance de personnes athées (36.36%), catholiques (32.21%) et musulmanes (13.83%)
+- prépondérance de personnes ayant pour langue maternelle le français (63.84%)
+- forte maîtrise moyenne du français (0.87)
+- forte maîtrise moyenne de l'anglais (0.73)
+- très forte prépondérance de personnes actuellement en bonne ou moyenne santé (98.47%)
+- bon état de santé moyen durant les 12 derniers mois (0.78)
+- prépondérance d'un diplôme universitaire chez la mère (28.62%)
+- prépondérance d'un diplôme universitaire chez la mère (34.70%)
+
 # Présentation de la variable indépendante
@@
 ```
+
+# Analyse
+
+## Modèle de régression
+
+```{r}
+
+reg_1 <- lm(Base_Trajipaq_Finale_Ok$I2 ~ Base_Trajipaq_Finale_Ok$I1 +
+              Base_Trajipaq_Finale_Ok$Q10 +
+              Base_Trajipaq_Finale_Ok$Q8B +
+              Base_Trajipaq_Finale_Ok$Q4 +
+              Base_Trajipaq_Finale_Ok$Q16 +
+              Base_Trajipaq_Finale_Ok$Q17 +
+              Base_Trajipaq_Finale_Ok$Q20 +
+              Base_Trajipaq_Finale_Ok$Q23
+            )
+summary(reg_1)
+
+```
+
+## Coefficient de corrélation
+
+```{r}
+
+cor.test(Base_Trajipaq_Finale_Ok$I2, Base_Trajipaq_Finale_Ok$I1)
+
+```
+
+## Visualisation
+
+```{r}
+
+vis <- ggplot(Base_Trajipaq_Finale_Ok %>%
+                filter(!is.na(I1), !is.na(I2)), aes(x = I2, y = I1)) +
+  geom_jitter(color = "pink") +
+  stat_smooth(color = "purple",se = FALSE) +
+  labs(title = "Discrimination et sentiment d'appartenance", 
+       x = "Discrimination subie au cours des 12 derniers mois", 
+       y = "Sentiment d'appartenance au Quebec") +
+  theme_bw()
+print(vis)
+
+```
+
+## Visualisation contrôlée : sexe
+
+```{r}
+
+vis_sex <- ggplot(Base_Trajipaq_Finale_Ok %>%
+                filter(!is.na(I1), !is.na(I2), !is.na(Q10)), aes(x = I2, y = I1, color = as.factor(Q10))) +
+  geom_jitter(color = "pink") +
+  stat_smooth(se = FALSE) +
+  scale_color_brewer(name = "Sexe", palette = 3) +
+  labs(title = "Discrimination et sentiment d'appartenance", 
+       x = "Discrimination subie au cours des 12 derniers mois", 
+       y = "Sentiment d'appartenance au Quebec") +
+  theme_bw()
+print(vis_sex)
+
+```
+
+## Visualisation contrôlée : âge
+
+```{r}
+
+Base_Trajipaq_Finale_Ok <-
+  Base_Trajipaq_Finale_Ok %>%
+  mutate(Q8A_F = case_when(
+    Q8A <= 1963 ~ ">=60",
+    Q8A >= 1963 & Q8A < 1968 ~ "[59;55]",
+    Q8A >= 1968 & Q8A < 1973 ~ "[54;50]",
+    Q8A >= 1973 & Q8A < 1978 ~ "[49;45]",
+    Q8A >= 1978 & Q8A < 1983 ~ "[44;35]",
+    Q8A >= 1983 & Q8A < 1988 ~ "[34;30]",
+    Q8A >= 1988 & Q8A < 1993 ~ "[29;25]",
+    Q8A >= 1993 & Q8A < 1998 ~ "[24;20]",
+    Q8A >= 1998 & Q8A < 2023 ~ "<19"
+  ))
+
+```
+
+```{r}
+
+vis_age <- ggplot(Base_Trajipaq_Finale_Ok %>%
+                filter(!is.na(I1), !is.na(I2), !is.na(Q8A_F)), aes(x = I2, y = I1, color = Q8A_F)) +
+  geom_jitter(color = "pink") +
+  stat_smooth(se = FALSE, span = 0.9) +
+  scale_color_brewer(name = "Age", palette = 3) +
+  labs(title = "Discrimination et sentiment d'appartenance", 
+       x = "Discrimination subie au cours des 12 derniers mois", 
+       y = "Sentiment d'appartenance au Quebec",
+       ) +
+  theme_bw()
+print(vis_age)
+
+```
+
+## Visualisation contrôlée : statut matrimonial
+
+```{r}
+
+Base_Trajipaq_Finale_Ok <-
+  Base_Trajipaq_Finale_Ok %>%
+  mutate(STATUT_F = case_when(
+    STATUT == 1 ~ "Celibataire",
+    STATUT == 2 ~ "Marie(e)",
+    STATUT == 3 ~ "Conjoint(e) de fait",
+    STATUT == 4 ~ "Separe(e)",
+    STATUT == 5 ~ "Divorce(e)",
+    STATUT == 6 ~ NA
+  ))
+
+```
+
+```{r}
+
+vis_born <- ggplot(Base_Trajipaq_Finale_Ok %>%
+                filter(!is.na(I1), !is.na(I2), !is.na(STATUT_F)), aes(x = I2, y = I1, color = STATUT_F)) +
+  geom_jitter(color = "pink") +
+  stat_smooth(se = FALSE, span = 0.9) +
+  scale_color_brewer(name = "Statut matrimonial", palette = 3) +
+  labs(title = "Discrimination et sentiment d'appartenance", 
+       x = "Discrimination subie au cours des 12 derniers mois", 
+       y = "Sentiment d'appartenance au Quebec") +
+  theme_bw()
+print(vis_born)
+
+```
+
+## Visualisation contrôlée : statut migratoire
+
+```{r}
+
+vis_born <- ggplot(Base_Trajipaq_Finale_Ok %>%
+                filter(!is.na(I1), !is.na(I2), !is.na(Q4)), aes(x = I2, y = I1, color = as.factor(Q4))) +
+  geom_jitter(color = "pink") +
+  stat_smooth(se = FALSE, span = 0.9) +
+  scale_color_brewer(name = "Ne au Canada", palette = 3) +
+  labs(title = "Discrimination et sentiment d'appartenance", 
+       x = "Discrimination subie au cours des 12 derniers mois", 
+       y = "Sentiment d'appartenance au Quebec") +
+  theme_bw()
+print(vis_born)
+
+```
+
+## Visualisation contrôlée : niveau d'éducation
+
+```{r}
+
+vis_educ <- ggplot(Base_Trajipaq_Finale_Ok %>%
+                filter(!is.na(I1), !is.na(I2), !is.na(Q16)), aes(x = I2, y = I1, color = as.factor(Q16))) +
+  geom_jitter(color = "pink") +
+  stat_smooth(se = FALSE, span = 1) +
+  scale_color_brewer(name = "Niveau d'education", palette = 3) +
+  labs(title = "Discrimination et sentiment d'appartenance", 
+       x = "Discrimination subie au cours des 12 derniers mois", 
+       y = "Sentiment d'appartenance au Quebec") +
+  theme_bw()
+print(vis_educ)
+
+```
+
+## Visualisation contrôlée : religion
+
+```{r}
+
+Base_Trajipaq_Finale_Ok <-
+  Base_Trajipaq_Finale_Ok %>%
+  mutate(Q17_F = case_when(
+    Q17 == 0 ~ "Sans religion",
+    Q17 == 1 ~ "Catholique",
+    Q17 == 8 ~ "Musulman",
+    Q17 != 0 & Q17 != 1 & Q17 != 8 ~ "Autre"
+  ))
+
+```
+
+```{r}
+
+vis_relig <- ggplot(Base_Trajipaq_Finale_Ok %>%
+                filter(!is.na(I1), !is.na(I2), !is.na(Q17_F)), aes(x = I2, y = I1, color = Q17_F)) +
+  geom_jitter(color = "pink") +
+  stat_smooth(se = FALSE, span = 0.9) +
+  scale_color_brewer(name = "Religion", palette = 3) +
+  labs(title = "Discrimination et sentiment d'appartenance", 
+       x = "Discrimination subie au cours des 12 derniers mois", 
+       y = "Sentiment d'appartenance au Quebec") +
+  theme_bw()
+print(vis_relig)
+
+```
+
+## Visualisation contrôlée : langue maternelle
+
+```{r}
+
+Base_Trajipaq_Finale_Ok <-
+  Base_Trajipaq_Finale_Ok %>%
+  mutate(Q20_F = case_when(
+    Q20M1 == 1 ~ "Francais",
+    Q20M1 == 2 ~ "Anglais",
+    Q20M1 == 3 ~ "Autre"
+  ))
+
+```
+
+```{r}
+
+vis_lang <- ggplot(Base_Trajipaq_Finale_Ok %>%
+                filter(!is.na(I1), !is.na(I2), !is.na(Q20_F)), aes(x = I2, y = I1, color = Q20_F)) +
+  geom_jitter(color = "pink") +
+  stat_smooth(se = FALSE, span = 0.8) +
+  scale_color_brewer(name = "Langue maternelle", palette = 3) +
+  labs(title = "Discrimination et sentiment d'appartenance", 
+       x = "Discrimination subie au cours des 12 derniers mois", 
+       y = "Sentiment d'appartenance au Quebec") +
+  theme_bw()
+print(vis_lang)
+
+```
+
+## Visualisation contrôlée : état de santé
+
+```{r}
+
+vis_health <- ggplot(Base_Trajipaq_Finale_Ok %>%
+                filter(!is.na(I1), !is.na(I2), !is.na(Q23)), aes(x = I2, y = I1, color = as.factor(Q23))) +
+  geom_jitter(color = "pink") +
+  stat_smooth(se = FALSE, span = 1.1) +
+  scale_color_brewer(name = "Etat de sante", palette = 3) +
+  labs(title = "Discrimination et sentiment d'appartenance", 
+       x = "Discrimination subie au cours des 12 derniers mois", 
+       y = "Sentiment d'appartenance au Quebec") +
+  theme_bw()
+print(vis_health)
+
+```

### 2023-05-22 Codes/Article---Sentiment-d-appartenance.tex
Visualisation LateX.

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Article - Sentiment d'appartenance},
  pdfauthor={Thea},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{Article - Sentiment d'appartenance}
\author{Thea}
\date{2023-05-17}

\begin{document}
\maketitle

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(summarytools)}
\FunctionTok{library}\NormalTok{(tidyverse)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## -- Attaching core tidyverse packages ------------------------ tidyverse 2.0.0 --
## v dplyr     1.1.2     v readr     2.1.4
## v forcats   1.0.0     v stringr   1.5.0
## v ggplot2   3.4.2     v tibble    3.2.1
## v lubridate 1.9.2     v tidyr     1.3.0
## v purrr     1.0.1     
## -- Conflicts ------------------------------------------ tidyverse_conflicts() --
## x dplyr::filter() masks stats::filter()
## x dplyr::lag()    masks stats::lag()
## x tibble::view()  masks summarytools::view()
## i Use the conflicted package (<http://conflicted.r-lib.org/>) to force all conflicts to become errors
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(ggthemes)}
\FunctionTok{library}\NormalTok{(haven)}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{Base\_Trajipaq\_Finale\_Ok }\OtherTok{\textless{}{-}} \FunctionTok{read\_dta}\NormalTok{(}\StringTok{"\textasciitilde{}/Desktop/Base\_Trajipaq\_Finale\_Ok.dta"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\hypertarget{question-de-recherche-de-quelles-fauxe7ons-les-discriminations-que-subissent-les-personnes-immigrantes-influencent-leur-sentiment-dappartenance-au-quuxe9bec}{%
\section{Question de recherche : de quelles façons les discriminations
que subissent les personnes immigrantes influencent leur sentiment
d'appartenance au Québec
?}\label{question-de-recherche-de-quelles-fauxe7ons-les-discriminations-que-subissent-les-personnes-immigrantes-influencent-leur-sentiment-dappartenance-au-quuxe9bec}}

\hypertarget{variable-induxe9pendante-les-discriminations-que-subissent-les-personnes-immigrantes}{%
\subsection{Variable indépendante = Les discriminations que subissent
les personnes immigrantes
:}\label{variable-induxe9pendante-les-discriminations-que-subissent-les-personnes-immigrantes}}

\begin{itemize}
\tightlist
\item
  dans l'emploi
\item
  dans la recherche de logement
\item
  dans la santé
\item
  sur le lieu de formation
\end{itemize}

Type de variable = factoriel Niveau de mesure = ordinal (échelle de
Likert)

Réponses possibles :

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Tout à fait d'accord
\item
  Plutôt d'accord
\item
  Plutôt pas d'accord
\item
  Pas du tout d'accord
\item
  Je ne sais pas
\item
  Je ne préfère pas répondre
\end{enumerate}

\hypertarget{variable-duxe9pendante-le-sentiment-dappartenance-au-quuxe9bec}{%
\subsection{Variable dépendante = le sentiment d'appartenance au Québec
:}\label{variable-duxe9pendante-le-sentiment-dappartenance-au-quuxe9bec}}

\begin{itemize}
\tightlist
\item
  se sentir chez soi au Québec
\item
  la nécessité d'oublier sa culture d'origine pour s'intégrer au Québec
\item
  être vu(e) comme un(e) Québécois(e)
\item
  se sentir Québécois(e)
\item
  se sentir Canadien(ne)
\item
  avoir un sentiment d'appartenance avec son pays d'origine
\end{itemize}

Type de variable = factoriel Niveau de mesure = ordinal (échelle de
Likert)

Réponses possibles :

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Tout à fait d'accord
\item
  Plutôt d'accord
\item
  Plutôt pas d'accord
\item
  Pas du tout d'accord
\item
  Je ne sais pas
\item
  Je ne préfère pas répondre
\end{enumerate}

\hypertarget{variables-de-contruxf4le-caractuxe9ristiques-socioduxe9mographiques}{%
\subsection{Variables de contrôle = caractéristiques sociodémographiques
:}\label{variables-de-contruxf4le-caractuxe9ristiques-socioduxe9mographiques}}

\begin{itemize}
\tightlist
\item
  sexe
\item
  groupe d'âge
\item
  statut matrimonial
\item
  pays d'origine
\item
  statut d'immigration au Québec
\item
  statut d'immigration pré-résidence
\item
  niveau d'éducation
\item
  religion
\item
  compétence linguistique (français)
\item
  compétence linguistique (anglais)
\item
  état de santé actuel
\item
  état de santé depuis 12 mois
\item
  niveau d'éducation de la mère
\item
  niveau d'éducation du père
\end{itemize}

\hypertarget{pruxe9sentation-des-variables-de-contruxf4le}{%
\section{Présentation des variables de
contrôle}\label{pruxe9sentation-des-variables-de-contruxf4le}}

\hypertarget{q10.-uxeates-vous-un-homme-une-femme-ou-autre}{%
\subsection{Q10. Êtes-vous un homme, une femme ou autre
?}\label{q10.-uxeates-vous-un-homme-une-femme-ou-autre}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Homme
\item
  Femme
\item
  Autre
\end{enumerate}

Type de variable = catégorielle Niveau de mesure = nominal

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{Base\_Trajipaq\_Finale\_Ok }\OtherTok{\textless{}{-}}
\NormalTok{  Base\_Trajipaq\_Finale\_Ok }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{Q10\_F =} \FunctionTok{case\_when}\NormalTok{(}
\NormalTok{    Q10 }\SpecialCharTok{==} \DecValTok{1} \SpecialCharTok{\textasciitilde{}} \StringTok{"Homme"}\NormalTok{,}
\NormalTok{    Q10 }\SpecialCharTok{==} \DecValTok{2} \SpecialCharTok{\textasciitilde{}} \StringTok{"Femme"}\NormalTok{,}
\NormalTok{    Q10 }\SpecialCharTok{==} \DecValTok{97} \SpecialCharTok{\textasciitilde{}} \StringTok{"Autre"}
\NormalTok{  ))}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{freq}\NormalTok{(Base\_Trajipaq\_Finale\_Ok}\SpecialCharTok{$}\NormalTok{Q10\_F)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Frequencies  
## Base_Trajipaq_Finale_Ok$Q10_F  
## Type: Character  
## 
##               Freq   % Valid   % Valid Cum.   % Total   % Total Cum.
## ----------- ------ --------- -------------- --------- --------------
##       Femme    886     56.69          56.69     56.69          56.69
##       Homme    677     43.31         100.00     43.31         100.00
##        <NA>      0                               0.00         100.00
##       Total   1563    100.00         100.00    100.00         100.00
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{ggplot}\NormalTok{(}\AttributeTok{data =}\NormalTok{ Base\_Trajipaq\_Finale\_Ok, }\AttributeTok{mapping =} \FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ Q10\_F)) }\SpecialCharTok{+}
  \FunctionTok{geom\_bar}\NormalTok{(}\AttributeTok{fill =} \StringTok{"pink"}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{theme\_bw}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\includegraphics{Article---Sentiment-d-appartenance_files/figure-latex/unnamed-chunk-5-1.pdf}

La réponse modale est ``femme'' (56.69\%).

\hypertarget{q8b-a-quel-groupe-duxe2ge-appartenez-vous}{%
\subsection{Q8B A quel groupe d'âge appartenez-vous
?}\label{q8b-a-quel-groupe-duxe2ge-appartenez-vous}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  0 -- 19 ans
\item
  20 -- 24 ans
\item
  25 -- 29 ans
\item
  30 -- 34 ans
\item
  35 -- 39 ans
\item
  40 -- 44 ans
\item
  45 -- 49 ans
\item
  50 -- 60 ans
\item
  61 ans et plus
\item
  Je préfère ne pas répondre
\end{enumerate}

Type de variable = factoriel Niveau de mesure = ordinal

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{freq}\NormalTok{(Base\_Trajipaq\_Finale\_Ok}\SpecialCharTok{$}\NormalTok{QAGE)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Frequencies  
## Base_Trajipaq_Finale_Ok$QAGE  
## Label: Vous appartenez <U+00E0> quel groupe d<U+2019><U+00E2>ge?  
## Type: Numeric  
## 
##               Freq   % Valid   % Valid Cum.   % Total   % Total Cum.
## ----------- ------ --------- -------------- --------- --------------
##           2     76      4.86           4.86      4.86           4.86
##           3    136      8.70          13.56      8.70          13.56
##           4    259     16.57          30.13     16.57          30.13
##           5    385     24.63          54.77     24.63          54.77
##           6    328     20.99          75.75     20.99          75.75
##           7    182     11.64          87.40     11.64          87.40
##           8    125      8.00          95.39      8.00          95.39
##           9     72      4.61         100.00      4.61         100.00
##        <NA>      0                               0.00         100.00
##       Total   1563    100.00         100.00    100.00         100.00
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{ggplot}\NormalTok{(}\AttributeTok{data =}\NormalTok{ Base\_Trajipaq\_Finale\_Ok, }\AttributeTok{mapping =} \FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ QAGE)) }\SpecialCharTok{+}
  \FunctionTok{geom\_bar}\NormalTok{(}\AttributeTok{fill =} \StringTok{"pink"}\NormalTok{, }\AttributeTok{na.rm =} \ConstantTok{TRUE}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{theme\_bw}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\includegraphics{Article---Sentiment-d-appartenance_files/figure-latex/unnamed-chunk-7-1.pdf}

La distribution est symétrique. Le groupe d'âge modal est ``35-39 ans''
(24.63\%).

\hypertarget{statut.-uxeates-vous-actuellement}{%
\subsection{STATUT. Êtes-vous actuellement\ldots{}
?}\label{statut.-uxeates-vous-actuellement}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Célibataire/jamais marié(e) ou conjoint(e) de fait
\item
  Marié(e)
\item
  Conjoint(e) de fait/en union libre
\item
  Séparé(e)
\item
  Divorcé(e)
\item
  Veuf(ve)
\end{enumerate}

Type de variable = catégorielle Niveau de mesure = nominal

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{freq}\NormalTok{(Base\_Trajipaq\_Finale\_Ok}\SpecialCharTok{$}\NormalTok{STATUT)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Frequencies  
## Base_Trajipaq_Finale_Ok$STATUT  
## Label: <U+00CA>tes-vous actuellement...?  
## Type: Numeric  
## 
##               Freq   % Valid   % Valid Cum.   % Total   % Total Cum.
## ----------- ------ --------- -------------- --------- --------------
##           1    363     24.26          24.26     23.22          23.22
##           2    711     47.53          71.79     45.49          68.71
##           3    331     22.13          93.92     21.18          89.89
##           4     39      2.61          96.52      2.50          92.39
##           5     46      3.07          99.60      2.94          95.33
##           6      6      0.40         100.00      0.38          95.71
##        <NA>     67                               4.29         100.00
##       Total   1563    100.00         100.00    100.00         100.00
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{ggplot}\NormalTok{(}\AttributeTok{data =}\NormalTok{ Base\_Trajipaq\_Finale\_Ok, }\AttributeTok{mapping =} \FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ STATUT)) }\SpecialCharTok{+}
  \FunctionTok{geom\_bar}\NormalTok{(}\AttributeTok{fill =} \StringTok{"pink"}\NormalTok{, }\AttributeTok{na.rm =} \ConstantTok{TRUE}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{theme\_bw}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\includegraphics{Article---Sentiment-d-appartenance_files/figure-latex/unnamed-chunk-9-1.pdf}

La réponse modale est ``Marié(e)'' (47.53\%)

\hypertarget{q11.-dans-quel-pays-uxeates-vous-nuxe9}{%
\subsection{Q11. Dans quel pays êtes-vous né
?}\label{q11.-dans-quel-pays-uxeates-vous-nuxe9}}

Pays. (liste déroulante) 998. Je ne sais pas 999. Je préfère ne pas
répondre

Type de variable = catégorielle Niveau de mesure = nominal

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{freq}\NormalTok{(Base\_Trajipaq\_Finale\_Ok}\SpecialCharTok{$}\NormalTok{Q11)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Frequencies  
## Base_Trajipaq_Finale_Ok$Q11  
## Label: Dans quel pays <U+00EA>tes-vous n<U+00E9>(e) ?Si vous ne vous en souvenez pas, inscrire : <U+00AB> Je  
## Type: Numeric  
## 
##               Freq   % Valid   % Valid Cum.   % Total   % Total Cum.
## ----------- ------ --------- -------------- --------- --------------
##           2      1     0.090          0.090     0.064          0.064
##           4      1     0.090          0.180     0.064          0.128
##           5     74     6.649          6.828     4.734          4.862
##           6      6     0.539          7.367     0.384          5.246
##          12      2     0.180          7.547     0.128          5.374
##          13      8     0.719          8.266     0.512          5.886
##          14      1     0.090          8.356     0.064          5.950
##          16      2     0.180          8.535     0.128          6.078
##          18      1     0.090          8.625     0.064          6.142
##          21      3     0.270          8.895     0.192          6.334
##          23      3     0.270          9.164     0.192          6.526
##          24     18     1.617         10.782     1.152          7.678
##          26      4     0.359         11.141     0.256          7.933
##          35     24     2.156         13.297     1.536          9.469
##          37     12     1.078         14.376     0.768         10.237
##          38      4     0.359         14.735     0.256         10.493
##          39      3     0.270         15.004     0.192         10.685
##          40      1     0.090         15.094     0.064         10.749
##          42      1     0.090         15.184     0.064         10.813
##          43     30     2.695         17.880     1.919         12.732
##          45      1     0.090         17.969     0.064         12.796
##          46      1     0.090         18.059     0.064         12.860
##          47     35     3.145         21.204     2.239         15.099
##          50      1     0.090         21.294     0.064         15.163
##          52     36     3.235         24.528     2.303         17.466
##          54      6     0.539         25.067     0.384         17.850
##          55      1     0.090         25.157     0.064         17.914
##          58      1     0.090         25.247     0.064         17.978
##          59      1     0.090         25.337     0.064         18.042
##          60     12     1.078         26.415     0.768         18.810
##          62     15     1.348         27.763     0.960         19.770
##          65      1     0.090         27.853     0.064         19.834
##          66      3     0.270         28.122     0.192         20.026
##          68      6     0.539         28.661     0.384         20.409
##          69      3     0.270         28.931     0.192         20.601
##          70      1     0.090         29.021     0.064         20.665
##          73      4     0.359         29.380     0.256         20.921
##          75     17     1.527         30.907     1.088         22.009
##          79      1     0.090         30.997     0.064         22.073
##          81    297    26.685         57.682    19.002         41.075
##          82      1     0.090         57.772     0.064         41.139
##          86      1     0.090         57.862     0.064         41.203
##          88      1     0.090         57.951     0.064         41.267
##          89      1     0.090         58.041     0.064         41.331
##          91      1     0.090         58.131     0.064         41.395
##          95      4     0.359         58.491     0.256         41.651
##         100     28     2.516         61.006     1.791         43.442
##         103     10     0.898         61.905     0.640         44.082
##         106      1     0.090         61.995     0.064         44.146
##         107     13     1.168         63.163     0.832         44.978
##         109      8     0.719         63.881     0.512         45.489
##         113      6     0.539         64.420     0.384         45.873
##         114      6     0.539         64.960     0.384         46.257
##         115      2     0.180         65.139     0.128         46.385
##         116      6     0.539         65.678     0.384         46.769
##         119      3     0.270         65.948     0.192         46.961
##         121      1     0.090         66.038     0.064         47.025
##         127      1     0.090         66.128     0.064         47.089
##         128     35     3.145         69.272     2.239         49.328
##         136      1     0.090         69.362     0.064         49.392
##         140      4     0.359         69.721     0.256         49.648
##         143     57     5.121         74.843     3.647         53.295
##         146      8     0.719         75.562     0.512         53.807
##         149     14     1.258         76.819     0.896         54.702
##         151     16     1.438         78.257     1.024         55.726
##         152      1     0.090         78.347     0.064         55.790
##         161      1     0.090         78.437     0.064         55.854
##         162      2     0.180         78.616     0.128         55.982
##         167      1     0.090         78.706     0.064         56.046
##         169      1     0.090         78.796     0.064         56.110
##         170      1     0.090         78.886     0.064         56.174
##         171      3     0.270         79.155     0.192         56.366
##         174      2     0.180         79.335     0.128         56.494
##         176      1     0.090         79.425     0.064         56.558
##         177     12     1.078         80.503     0.768         57.326
##         178      8     0.719         81.222     0.512         57.837
##         180      2     0.180         81.402     0.128         57.965
##         181      1     0.090         81.491     0.064         58.029
##         183      2     0.180         81.671     0.128         58.157
##         185      2     0.180         81.851     0.128         58.285
##         186     32     2.875         84.726     2.047         60.333
##         187      7     0.629         85.355     0.448         60.781
##         188     11     0.988         86.343     0.704         61.484
##         189      1     0.090         86.433     0.064         61.548
##         200      1     0.090         86.523     0.064         61.612
##         205      6     0.539         87.062     0.384         61.996
##         210      1     0.090         87.152     0.064         62.060
##         211      1     0.090         87.242     0.064         62.124
##         216      2     0.180         87.421     0.128         62.252
##         217      1     0.090         87.511     0.064         62.316
##         218      8     0.719         88.230     0.512         62.828
##         222     12     1.078         89.308     0.768         63.596
##         224      3     0.270         89.578     0.192         63.788
##         226      3     0.270         89.847     0.192         63.980
##         232      4     0.359         90.207     0.256         64.235
##         234      1     0.090         90.296     0.064         64.299
##         236     27     2.426         92.722     1.727         66.027
##         239      4     0.359         93.082     0.256         66.283
##         241     12     1.078         94.160     0.768         67.051
##         242      7     0.629         94.789     0.448         67.498
##         244     14     1.258         96.047     0.896         68.394
##         247     10     0.898         96.945     0.640         69.034
##         249      1     0.090         97.035     0.064         69.098
##         998      9     0.809         97.844     0.576         69.674
##         999     24     2.156        100.000     1.536         71.209
##        <NA>    450                             28.791        100.000
##       Total   1563   100.000        100.000   100.000        100.000
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{ggplot}\NormalTok{(}\AttributeTok{data =}\NormalTok{ Base\_Trajipaq\_Finale\_Ok, }\AttributeTok{mapping =} \FunctionTok{aes}\NormalTok{(}\AttributeTok{x =} \FunctionTok{as.factor}\NormalTok{(Q11))) }\SpecialCharTok{+}
  \FunctionTok{geom\_bar}\NormalTok{(}\AttributeTok{fill =} \StringTok{"pink"}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{theme\_bw}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\includegraphics{Article---Sentiment-d-appartenance_files/figure-latex/unnamed-chunk-11-1.pdf}

Les réponses modales sont ``France'' (26.685\%) et ``Algérie''
(6.649\%).

\hypertarget{q4.-uxeates-vous-nuxe9-au-canada}{%
\subsection{Q4. Êtes-vous né au Canada
?}\label{q4.-uxeates-vous-nuxe9-au-canada}}

Type de variable = catégoriel Niveau de mesure = nominal

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{freq}\NormalTok{(Base\_Trajipaq\_Finale\_Ok}\SpecialCharTok{$}\NormalTok{Q4)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Frequencies  
## Base_Trajipaq_Finale_Ok$Q4  
## Label: <U+00CA>tes-vous n<U+00E9>(e) au Canada ?  
## Type: Numeric  
## 
##               Freq   % Valid   % Valid Cum.   % Total   % Total Cum.
## ----------- ------ --------- -------------- --------- --------------
##           1   1113     71.21          71.21     71.21          71.21
##           2    450     28.79         100.00     28.79         100.00
##        <NA>      0                               0.00         100.00
##       Total   1563    100.00         100.00    100.00         100.00
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{ggplot}\NormalTok{(}\AttributeTok{data =}\NormalTok{ Base\_Trajipaq\_Finale\_Ok, }\AttributeTok{mapping =} \FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ Q4)) }\SpecialCharTok{+}
  \FunctionTok{geom\_bar}\NormalTok{(}\AttributeTok{fill =} \StringTok{"pink"}\NormalTok{, }\AttributeTok{na.rm =} \ConstantTok{TRUE}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{theme\_bw}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\includegraphics{Article---Sentiment-d-appartenance_files/figure-latex/unnamed-chunk-13-1.pdf}

La réponse modale est ``deuxième génération'' (92.87\%).

\hypertarget{q15i.-statut-migratoire-avant-dobtenir-la-ruxe9sidence-permanente}{%
\subsection{Q15I. Statut migratoire avant d'obtenir la résidence
permanente}\label{q15i.-statut-migratoire-avant-dobtenir-la-ruxe9sidence-permanente}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Travailleur temporaire (visa de travail ou permis de travail)
\item
  Demandeur d'asile
\item
  Étudiant étranger (visa d'études)
\item
  Touriste
\item
  Autre
\item
  Je ne sais pas
\item
  Je préfère ne pas répondre
\end{enumerate}

Type de variable = catégoriel Niveau de mesure = nominal

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{Base\_Trajipaq\_Finale\_Ok }\OtherTok{\textless{}{-}}
\NormalTok{  Base\_Trajipaq\_Finale\_Ok }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{Q15M1 =} \FunctionTok{case\_when}\NormalTok{(}
\NormalTok{    Q15M1 }\SpecialCharTok{==} \DecValTok{1} \SpecialCharTok{\textasciitilde{}} \DecValTok{1}\NormalTok{,}
\NormalTok{    Q15M1 }\SpecialCharTok{==} \DecValTok{2} \SpecialCharTok{\textasciitilde{}} \DecValTok{2}\NormalTok{,}
\NormalTok{    Q15M1 }\SpecialCharTok{==} \DecValTok{3} \SpecialCharTok{\textasciitilde{}} \DecValTok{3}\NormalTok{,}
\NormalTok{    Q15M1 }\SpecialCharTok{==} \DecValTok{4} \SpecialCharTok{\textasciitilde{}} \DecValTok{4}\NormalTok{,}
\NormalTok{    Q15M1 }\SpecialCharTok{==} \DecValTok{97} \SpecialCharTok{\textasciitilde{}} \DecValTok{5}\NormalTok{,}
\NormalTok{    Q15M1 }\SpecialCharTok{==} \DecValTok{98} \SpecialCharTok{\textasciitilde{}} \ConstantTok{NA}\NormalTok{,}
\NormalTok{    Q15M1 }\SpecialCharTok{==} \DecValTok{99} \SpecialCharTok{\textasciitilde{}} \ConstantTok{NA}
\NormalTok{  ))}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{freq}\NormalTok{(Base\_Trajipaq\_Finale\_Ok}\SpecialCharTok{$}\NormalTok{Q15M1)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Frequencies  
## Base_Trajipaq_Finale_Ok$Q15M1  
## Type: Numeric  
## 
##               Freq   % Valid   % Valid Cum.   % Total   % Total Cum.
## ----------- ------ --------- -------------- --------- --------------
##           1    241     57.79          57.79     15.42          15.42
##           2     14      3.36          61.15      0.90          16.31
##           3     82     19.66          80.82      5.25          21.56
##           4     35      8.39          89.21      2.24          23.80
##           5     45     10.79         100.00      2.88          26.68
##        <NA>   1146                              73.32         100.00
##       Total   1563    100.00         100.00    100.00         100.00
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{ggplot}\NormalTok{(}\AttributeTok{data =}\NormalTok{ Base\_Trajipaq\_Finale\_Ok, }\AttributeTok{mapping =} \FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ Q15M1)) }\SpecialCharTok{+}
  \FunctionTok{geom\_bar}\NormalTok{(}\AttributeTok{fill =} \StringTok{"pink"}\NormalTok{, }\AttributeTok{na.rm =} \ConstantTok{TRUE}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{theme\_bw}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\includegraphics{Article---Sentiment-d-appartenance_files/figure-latex/unnamed-chunk-16-1.pdf}

La réponse modale est ``travailleur temporaire'' (57.78\%).

\hypertarget{q16.-quel-est-votre-plus-haut-dipluxf4me-duxe9tudes-obtenu}{%
\subsection{Q16. Quel est votre plus haut diplôme d'études obtenu
?}\label{q16.-quel-est-votre-plus-haut-dipluxf4me-duxe9tudes-obtenu}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Moins que le secondaire
\item
  Secondaire ou équivalent
\item
  Collégial
\item
  Formation professionnelle ou post-secondaire
\item
  Université 1er cycle
\item
  Université 2ème cycle et plus
\item
  Je ne sais pas
\item
  Je préfère ne pas répondre
\end{enumerate}

Type de variable = factorielle Niveau de mesure = ordinal

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{Base\_Trajipaq\_Finale\_Ok }\OtherTok{\textless{}{-}}
\NormalTok{  Base\_Trajipaq\_Finale\_Ok }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{Q16 =} \FunctionTok{case\_when}\NormalTok{(}
\NormalTok{    Q16 }\SpecialCharTok{==} \DecValTok{1} \SpecialCharTok{\textasciitilde{}} \DecValTok{1}\NormalTok{,}
\NormalTok{    Q16 }\SpecialCharTok{==} \DecValTok{2} \SpecialCharTok{\textasciitilde{}} \DecValTok{2}\NormalTok{,}
\NormalTok{    Q16 }\SpecialCharTok{==} \DecValTok{3} \SpecialCharTok{\textasciitilde{}} \DecValTok{3}\NormalTok{,}
\NormalTok{    Q16 }\SpecialCharTok{==} \DecValTok{4} \SpecialCharTok{\textasciitilde{}} \DecValTok{4}\NormalTok{,}
\NormalTok{    Q16 }\SpecialCharTok{==} \DecValTok{5} \SpecialCharTok{\textasciitilde{}} \DecValTok{5}\NormalTok{,}
\NormalTok{    Q16 }\SpecialCharTok{==} \DecValTok{6} \SpecialCharTok{\textasciitilde{}} \DecValTok{6}\NormalTok{,}
\NormalTok{    Q16 }\SpecialCharTok{==} \DecValTok{98} \SpecialCharTok{\textasciitilde{}} \ConstantTok{NA}\NormalTok{,}
\NormalTok{    Q16 }\SpecialCharTok{==} \DecValTok{99} \SpecialCharTok{\textasciitilde{}} \ConstantTok{NA}
\NormalTok{  ))}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{freq}\NormalTok{(Base\_Trajipaq\_Finale\_Ok}\SpecialCharTok{$}\NormalTok{Q16)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Frequencies  
## Base_Trajipaq_Finale_Ok$Q16  
## Type: Numeric  
## 
##               Freq   % Valid   % Valid Cum.   % Total   % Total Cum.
## ----------- ------ --------- -------------- --------- --------------
##           1     11      0.71           0.71      0.70           0.70
##           2    101      6.51           7.22      6.46           7.17
##           3    254     16.38          23.60     16.25          23.42
##           4    141      9.09          32.69      9.02          32.44
##           5    520     33.53          66.22     33.27          65.71
##           6    524     33.78         100.00     33.53          99.23
##        <NA>     12                               0.77         100.00
##       Total   1563    100.00         100.00    100.00         100.00
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{ggplot}\NormalTok{(}\AttributeTok{data =}\NormalTok{ Base\_Trajipaq\_Finale\_Ok, }\AttributeTok{mapping =} \FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ Q16)) }\SpecialCharTok{+}
  \FunctionTok{geom\_bar}\NormalTok{(}\AttributeTok{fill =} \StringTok{"pink"}\NormalTok{, }\AttributeTok{na.rm =} \ConstantTok{TRUE}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{theme\_bw}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\includegraphics{Article---Sentiment-d-appartenance_files/figure-latex/unnamed-chunk-19-1.pdf}

La distribution du niveau d'éducation est asymétrique vers la gauche.
L'échantillon est composé à 67.31\% de personnes titulaires d'un diplôme
universitaire.

\hypertarget{q17.-actuellement-avez-vous-une-religion}{%
\subsection{Q17. Actuellement, avez-vous une religion
?}\label{q17.-actuellement-avez-vous-une-religion}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{-1}
\tightlist
\item
  Sans religion
\item
  Catholique
\item
  Protestante
\item
  Évangéliste
\item
  Mormone
\item
  Témoins de Jéhovah
\item
  Orthodoxe
\item
  Autres Chrétiens
\item
  Musulmans
\item
  Juif
\item
  Bouddhiste
\item
  Hindou
\item
  Animiste
\item
  Autre
\item
  Je ne sais pas
\item
  Je préfère ne pas répondre
\end{enumerate}

Type de variable = catégorielle Niveau de mesure = nominal

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{Base\_Trajipaq\_Finale\_Ok }\OtherTok{\textless{}{-}}
\NormalTok{  Base\_Trajipaq\_Finale\_Ok }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{Q17 =} \FunctionTok{case\_when}\NormalTok{(}
\NormalTok{    Q17 }\SpecialCharTok{==} \DecValTok{0} \SpecialCharTok{\textasciitilde{}} \DecValTok{0}\NormalTok{,}
\NormalTok{    Q17 }\SpecialCharTok{==} \DecValTok{1} \SpecialCharTok{\textasciitilde{}} \DecValTok{1}\NormalTok{,}
\NormalTok{    Q17 }\SpecialCharTok{==} \DecValTok{2} \SpecialCharTok{\textasciitilde{}} \DecValTok{2}\NormalTok{,}
\NormalTok{    Q17 }\SpecialCharTok{==} \DecValTok{3} \SpecialCharTok{\textasciitilde{}} \DecValTok{3}\NormalTok{,}
\NormalTok{    Q17 }\SpecialCharTok{==} \DecValTok{4} \SpecialCharTok{\textasciitilde{}} \DecValTok{4}\NormalTok{,}
\NormalTok{    Q17 }\SpecialCharTok{==} \DecValTok{5} \SpecialCharTok{\textasciitilde{}} \DecValTok{5}\NormalTok{,}
\NormalTok{    Q17 }\SpecialCharTok{==} \DecValTok{6} \SpecialCharTok{\textasciitilde{}} \DecValTok{6}\NormalTok{,}
\NormalTok{    Q17 }\SpecialCharTok{==} \DecValTok{7} \SpecialCharTok{\textasciitilde{}} \DecValTok{7}\NormalTok{,}
\NormalTok{    Q17 }\SpecialCharTok{==} \DecValTok{8} \SpecialCharTok{\textasciitilde{}} \DecValTok{8}\NormalTok{,}
\NormalTok{    Q17 }\SpecialCharTok{==} \DecValTok{9} \SpecialCharTok{\textasciitilde{}} \DecValTok{9}\NormalTok{,}
\NormalTok{    Q17 }\SpecialCharTok{==} \DecValTok{10} \SpecialCharTok{\textasciitilde{}} \DecValTok{10}\NormalTok{,}
\NormalTok{    Q17 }\SpecialCharTok{==} \DecValTok{11} \SpecialCharTok{\textasciitilde{}} \DecValTok{11}\NormalTok{,}
\NormalTok{    Q17 }\SpecialCharTok{==} \DecValTok{12} \SpecialCharTok{\textasciitilde{}} \DecValTok{12}\NormalTok{,}
\NormalTok{    Q17 }\SpecialCharTok{==} \DecValTok{97} \SpecialCharTok{\textasciitilde{}} \DecValTok{13}\NormalTok{,}
\NormalTok{    Q17 }\SpecialCharTok{==} \DecValTok{98} \SpecialCharTok{\textasciitilde{}} \ConstantTok{NA}\NormalTok{,}
\NormalTok{    Q17 }\SpecialCharTok{==} \DecValTok{99} \SpecialCharTok{\textasciitilde{}} \ConstantTok{NA}
\NormalTok{  ))}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{freq}\NormalTok{(Base\_Trajipaq\_Finale\_Ok}\SpecialCharTok{$}\NormalTok{Q17)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Frequencies  
## Base_Trajipaq_Finale_Ok$Q17  
## Type: Numeric  
## 
##               Freq   % Valid   % Valid Cum.   % Total   % Total Cum.
## ----------- ------ --------- -------------- --------- --------------
##           0    552     36.36          36.36     35.32          35.32
##           1    489     32.21          68.58     31.29          66.60
##           2     42      2.77          71.34      2.69          69.29
##           3     23      1.52          72.86      1.47          70.76
##           4      2      0.13          72.99      0.13          70.89
##           5      5      0.33          73.32      0.32          71.21
##           6     89      5.86          79.18      5.69          76.90
##           7     17      1.12          80.30      1.09          77.99
##           8    210     13.83          94.14     13.44          91.43
##           9     18      1.19          95.32      1.15          92.58
##          10     22      1.45          96.77      1.41          93.99
##          11     11      0.72          97.50      0.70          94.69
##          12      2      0.13          97.63      0.13          94.82
##          13     36      2.37         100.00      2.30          97.12
##        <NA>     45                               2.88         100.00
##       Total   1563    100.00         100.00    100.00         100.00
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{ggplot}\NormalTok{(}\AttributeTok{data =}\NormalTok{ Base\_Trajipaq\_Finale\_Ok, }\AttributeTok{mapping =} \FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ Q17)) }\SpecialCharTok{+}
  \FunctionTok{geom\_bar}\NormalTok{(}\AttributeTok{fill =} \StringTok{"pink"}\NormalTok{, }\AttributeTok{na.rm =} \ConstantTok{TRUE}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{theme\_bw}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\includegraphics{Article---Sentiment-d-appartenance_files/figure-latex/unnamed-chunk-22-1.pdf}

Les réponses modales sont ``sans religion'' (36.36\%) et ``catholique''
(32.21\%). On remarque également un nombre significatif de réponses
``musulman'' (13.83\%).

\hypertarget{q20.-quand-vous-uxe9tiez-enfant-quelles-uxe9taient-la-langue-ou-les-langues-que-vous-parliez-le-plus-souvent-uxe0-luxe2ge-de-5-ans-et-que-vous-comprenez-toujours-plusieurs-ruxe9ponses-possibles.}{%
\subsection{Q20. Quand vous étiez enfant, quelles étaient la langue ou
les langues que vous parliez le plus souvent, à l'âge de 5 ans et que
vous comprenez toujours? (Plusieurs réponses
possibles).}\label{q20.-quand-vous-uxe9tiez-enfant-quelles-uxe9taient-la-langue-ou-les-langues-que-vous-parliez-le-plus-souvent-uxe0-luxe2ge-de-5-ans-et-que-vous-comprenez-toujours-plusieurs-ruxe9ponses-possibles.}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Français
\item
  Anglais
\item
  Autre : préciser, possibilité d'en déclarer deux
\item
  Je ne sais pas
\item
  Je préfère ne pas répondre
\end{enumerate}

Type de variable = catégoriel Niveau de mesure = nominal

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{Base\_Trajipaq\_Finale\_Ok }\OtherTok{\textless{}{-}}
\NormalTok{  Base\_Trajipaq\_Finale\_Ok }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{Q20M1 =} \FunctionTok{case\_when}\NormalTok{(}
\NormalTok{    Q20M1 }\SpecialCharTok{==} \DecValTok{1} \SpecialCharTok{\textasciitilde{}} \DecValTok{1}\NormalTok{,}
\NormalTok{    Q20M1 }\SpecialCharTok{==} \DecValTok{2} \SpecialCharTok{\textasciitilde{}} \DecValTok{2}\NormalTok{,}
\NormalTok{    Q20M1 }\SpecialCharTok{==} \DecValTok{97} \SpecialCharTok{\textasciitilde{}} \DecValTok{3}\NormalTok{,}
\NormalTok{    Q20M1 }\SpecialCharTok{==} \DecValTok{98} \SpecialCharTok{\textasciitilde{}} \ConstantTok{NA}\NormalTok{,}
\NormalTok{    Q20M1 }\SpecialCharTok{==} \DecValTok{99} \SpecialCharTok{\textasciitilde{}} \ConstantTok{NA}
\NormalTok{  ))}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{freq}\NormalTok{(Base\_Trajipaq\_Finale\_Ok}\SpecialCharTok{$}\NormalTok{Q20M1)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Frequencies  
## Base_Trajipaq_Finale_Ok$Q20M1  
## Type: Numeric  
## 
##               Freq   % Valid   % Valid Cum.   % Total   % Total Cum.
## ----------- ------ --------- -------------- --------- --------------
##           1    987     63.84          63.84     63.15          63.15
##           2    129      8.34          72.19      8.25          71.40
##           3    430     27.81         100.00     27.51          98.91
##        <NA>     17                               1.09         100.00
##       Total   1563    100.00         100.00    100.00         100.00
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{ggplot}\NormalTok{(}\AttributeTok{data =}\NormalTok{ Base\_Trajipaq\_Finale\_Ok, }\AttributeTok{mapping =} \FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ Q20M1)) }\SpecialCharTok{+}
  \FunctionTok{geom\_bar}\NormalTok{(}\AttributeTok{fill =} \StringTok{"pink"}\NormalTok{, }\AttributeTok{na.rm =} \ConstantTok{TRUE}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{theme\_bw}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\includegraphics{Article---Sentiment-d-appartenance_files/figure-latex/unnamed-chunk-25-1.pdf}

La réponse modale est ``français'' (63.84\%).

\hypertarget{i3.-indice-de-la-compuxe9tence-linguistique-franuxe7ais}{%
\subsection{I3. Indice de la compétence linguistique
(français)}\label{i3.-indice-de-la-compuxe9tence-linguistique-franuxe7ais}}

Type de variable = numeric Niveau de mesure = ratio

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{Base\_Trajipaq\_Finale\_Ok }\OtherTok{\textless{}{-}}
\NormalTok{  Base\_Trajipaq\_Finale\_Ok }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{Q21A =} \FunctionTok{case\_when}\NormalTok{(}
\NormalTok{    Q21A }\SpecialCharTok{==} \DecValTok{1} \SpecialCharTok{\textasciitilde{}} \DecValTok{1}\NormalTok{,}
\NormalTok{    Q21A }\SpecialCharTok{==} \DecValTok{2} \SpecialCharTok{\textasciitilde{}} \DecValTok{2}\NormalTok{,}
\NormalTok{    Q21A }\SpecialCharTok{==} \DecValTok{3} \SpecialCharTok{\textasciitilde{}} \DecValTok{3}\NormalTok{,}
\NormalTok{    Q21A }\SpecialCharTok{==} \DecValTok{4} \SpecialCharTok{\textasciitilde{}} \DecValTok{4}\NormalTok{,}
\NormalTok{    Q21A }\SpecialCharTok{==} \DecValTok{98} \SpecialCharTok{\textasciitilde{}} \ConstantTok{NA}\NormalTok{,}
\NormalTok{    Q21A }\SpecialCharTok{==} \DecValTok{99} \SpecialCharTok{\textasciitilde{}} \ConstantTok{NA}
\NormalTok{  ))}

\NormalTok{Base\_Trajipaq\_Finale\_Ok }\OtherTok{\textless{}{-}}
\NormalTok{  Base\_Trajipaq\_Finale\_Ok }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{Q21B =} \FunctionTok{case\_when}\NormalTok{(}
\NormalTok{    Q21B }\SpecialCharTok{==} \DecValTok{1} \SpecialCharTok{\textasciitilde{}} \DecValTok{1}\NormalTok{,}
\NormalTok{    Q21B }\SpecialCharTok{==} \DecValTok{2} \SpecialCharTok{\textasciitilde{}} \DecValTok{2}\NormalTok{,}
\NormalTok{    Q21B }\SpecialCharTok{==} \DecValTok{3} \SpecialCharTok{\textasciitilde{}} \DecValTok{3}\NormalTok{,}
\NormalTok{    Q21B }\SpecialCharTok{==} \DecValTok{4} \SpecialCharTok{\textasciitilde{}} \DecValTok{4}\NormalTok{,}
\NormalTok{    Q21B }\SpecialCharTok{==} \DecValTok{98} \SpecialCharTok{\textasciitilde{}} \ConstantTok{NA}\NormalTok{,}
\NormalTok{    Q21B }\SpecialCharTok{==} \DecValTok{99} \SpecialCharTok{\textasciitilde{}} \ConstantTok{NA}
\NormalTok{  ))}

\NormalTok{Base\_Trajipaq\_Finale\_Ok }\OtherTok{\textless{}{-}}
\NormalTok{  Base\_Trajipaq\_Finale\_Ok }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{Q21C =} \FunctionTok{case\_when}\NormalTok{(}
\NormalTok{    Q21C }\SpecialCharTok{==} \DecValTok{1} \SpecialCharTok{\textasciitilde{}} \DecValTok{1}\NormalTok{,}
\NormalTok{    Q21C }\SpecialCharTok{==} \DecValTok{2} \SpecialCharTok{\textasciitilde{}} \DecValTok{2}\NormalTok{,}
\NormalTok{    Q21C }\SpecialCharTok{==} \DecValTok{3} \SpecialCharTok{\textasciitilde{}} \DecValTok{3}\NormalTok{,}
\NormalTok{    Q21C }\SpecialCharTok{==} \DecValTok{4} \SpecialCharTok{\textasciitilde{}} \DecValTok{4}\NormalTok{,}
\NormalTok{    Q21C }\SpecialCharTok{==} \DecValTok{98} \SpecialCharTok{\textasciitilde{}} \ConstantTok{NA}\NormalTok{,}
\NormalTok{    Q21C }\SpecialCharTok{==} \DecValTok{99} \SpecialCharTok{\textasciitilde{}} \ConstantTok{NA}
\NormalTok{  ))}

\NormalTok{Base\_Trajipaq\_Finale\_Ok }\OtherTok{\textless{}{-}}
\NormalTok{  Base\_Trajipaq\_Finale\_Ok }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{Q21D =} \FunctionTok{case\_when}\NormalTok{(}
\NormalTok{    Q21D }\SpecialCharTok{==} \DecValTok{1} \SpecialCharTok{\textasciitilde{}} \DecValTok{1}\NormalTok{,}
\NormalTok{    Q21D }\SpecialCharTok{==} \DecValTok{2} \SpecialCharTok{\textasciitilde{}} \DecValTok{2}\NormalTok{,}
\NormalTok{    Q21D }\SpecialCharTok{==} \DecValTok{3} \SpecialCharTok{\textasciitilde{}} \DecValTok{3}\NormalTok{,}
\NormalTok{    Q21D }\SpecialCharTok{==} \DecValTok{4} \SpecialCharTok{\textasciitilde{}} \DecValTok{4}\NormalTok{,}
\NormalTok{    Q21D }\SpecialCharTok{==} \DecValTok{98} \SpecialCharTok{\textasciitilde{}} \ConstantTok{NA}\NormalTok{,}
\NormalTok{    Q21D }\SpecialCharTok{==} \DecValTok{99} \SpecialCharTok{\textasciitilde{}} \ConstantTok{NA}
\NormalTok{  ))}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{v\_I3 }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}\StringTok{"Q21A"}\NormalTok{,}
          \StringTok{"Q21B"}\NormalTok{,}
          \StringTok{"Q21C"}\NormalTok{,}
          \StringTok{"Q21D"}
\NormalTok{          )}

\NormalTok{Base\_Trajipaq\_Finale\_Ok }\OtherTok{\textless{}{-}}
\NormalTok{  Base\_Trajipaq\_Finale\_Ok }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{rowwise}\NormalTok{() }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{I3 =}\NormalTok{ (}\FunctionTok{sum}\NormalTok{(}\FunctionTok{c\_across}\NormalTok{(}\FunctionTok{all\_of}\NormalTok{(v\_I3)), }\AttributeTok{na.rm =} \ConstantTok{TRUE}\NormalTok{) }\SpecialCharTok{{-}} \FunctionTok{length}\NormalTok{(v\_I3)) }\SpecialCharTok{/}\NormalTok{ (}\DecValTok{3}\SpecialCharTok{*}\FunctionTok{length}\NormalTok{(v\_I3))) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{I3 =} \ControlFlowTok{if}\NormalTok{ (I3 }\SpecialCharTok{\textless{}} \DecValTok{0}\NormalTok{) }\ConstantTok{NA\_real\_} \ControlFlowTok{else}\NormalTok{ I3) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{ungroup}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{descr}\NormalTok{(Base\_Trajipaq\_Finale\_Ok}\SpecialCharTok{$}\NormalTok{I3)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Descriptive Statistics  
## Base_Trajipaq_Finale_Ok$I3  
## N: 1563  
## 
##                          I3
## ----------------- ---------
##              Mean      0.87
##           Std.Dev      0.24
##               Min      0.00
##                Q1      0.83
##            Median      1.00
##                Q3      1.00
##               Max      1.00
##               MAD      0.00
##               IQR      0.17
##                CV      0.28
##          Skewness     -1.98
##       SE.Skewness      0.07
##          Kurtosis      3.23
##           N.Valid   1117.00
##         Pct.Valid     71.47
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{ggplot}\NormalTok{(}\AttributeTok{data =}\NormalTok{ Base\_Trajipaq\_Finale\_Ok, }\AttributeTok{mapping =} \FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ I3)) }\SpecialCharTok{+}
  \FunctionTok{geom\_histogram}\NormalTok{(}\AttributeTok{na.rm =} \ConstantTok{TRUE}\NormalTok{, }\AttributeTok{fill =} \StringTok{"pink"}\NormalTok{, }\AttributeTok{binwidth =} \FloatTok{0.01}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{theme\_bw}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\includegraphics{Article---Sentiment-d-appartenance_files/figure-latex/unnamed-chunk-29-1.pdf}

La distribution est asymétrique vers la gauche. L'indice moyen de
maîtrise du français est 0.87.

\hypertarget{i4.-indice-de-la-compuxe9tence-linguistique-anglais.}{%
\subsection{I4. Indice de la compétence linguistique
(anglais).}\label{i4.-indice-de-la-compuxe9tence-linguistique-anglais.}}

Type de variable = numeric Niveau de mesure = ratio

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{Base\_Trajipaq\_Finale\_Ok }\OtherTok{\textless{}{-}}
\NormalTok{  Base\_Trajipaq\_Finale\_Ok }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{Q22A =} \FunctionTok{case\_when}\NormalTok{(}
\NormalTok{    Q22A }\SpecialCharTok{==} \DecValTok{1} \SpecialCharTok{\textasciitilde{}} \DecValTok{1}\NormalTok{,}
\NormalTok{    Q22A }\SpecialCharTok{==} \DecValTok{2} \SpecialCharTok{\textasciitilde{}} \DecValTok{2}\NormalTok{,}
\NormalTok{    Q22A }\SpecialCharTok{==} \DecValTok{3} \SpecialCharTok{\textasciitilde{}} \DecValTok{3}\NormalTok{,}
\NormalTok{    Q22A }\SpecialCharTok{==} \DecValTok{4} \SpecialCharTok{\textasciitilde{}} \DecValTok{4}\NormalTok{,}
\NormalTok{    Q22A }\SpecialCharTok{==} \DecValTok{98} \SpecialCharTok{\textasciitilde{}} \ConstantTok{NA}\NormalTok{,}
\NormalTok{    Q22A }\SpecialCharTok{==} \DecValTok{99} \SpecialCharTok{\textasciitilde{}} \ConstantTok{NA}
\NormalTok{  ))}

\NormalTok{Base\_Trajipaq\_Finale\_Ok }\OtherTok{\textless{}{-}}
\NormalTok{  Base\_Trajipaq\_Finale\_Ok }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{Q22B =} \FunctionTok{case\_when}\NormalTok{(}
\NormalTok{    Q22B }\SpecialCharTok{==} \DecValTok{1} \SpecialCharTok{\textasciitilde{}} \DecValTok{1}\NormalTok{,}
\NormalTok{    Q22B }\SpecialCharTok{==} \DecValTok{2} \SpecialCharTok{\textasciitilde{}} \DecValTok{2}\NormalTok{,}
\NormalTok{    Q22B }\SpecialCharTok{==} \DecValTok{3} \SpecialCharTok{\textasciitilde{}} \DecValTok{3}\NormalTok{,}
\NormalTok{    Q22B }\SpecialCharTok{==} \DecValTok{4} \SpecialCharTok{\textasciitilde{}} \DecValTok{4}\NormalTok{,}
\NormalTok{    Q22B }\SpecialCharTok{==} \DecValTok{98} \SpecialCharTok{\textasciitilde{}} \ConstantTok{NA}\NormalTok{,}
\NormalTok{    Q22B }\SpecialCharTok{==} \DecValTok{99} \SpecialCharTok{\textasciitilde{}} \ConstantTok{NA}
\NormalTok{  ))}

\NormalTok{Base\_Trajipaq\_Finale\_Ok }\OtherTok{\textless{}{-}}
\NormalTok{  Base\_Trajipaq\_Finale\_Ok }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{Q22C =} \FunctionTok{case\_when}\NormalTok{(}
\NormalTok{    Q22C }\SpecialCharTok{==} \DecValTok{1} \SpecialCharTok{\textasciitilde{}} \DecValTok{1}\NormalTok{,}
\NormalTok{    Q22C }\SpecialCharTok{==} \DecValTok{2} \SpecialCharTok{\textasciitilde{}} \DecValTok{2}\NormalTok{,}
\NormalTok{    Q22C }\SpecialCharTok{==} \DecValTok{3} \SpecialCharTok{\textasciitilde{}} \DecValTok{3}\NormalTok{,}
\NormalTok{    Q22C }\SpecialCharTok{==} \DecValTok{4} \SpecialCharTok{\textasciitilde{}} \DecValTok{4}\NormalTok{,}
\NormalTok{    Q22C }\SpecialCharTok{==} \DecValTok{98} \SpecialCharTok{\textasciitilde{}} \ConstantTok{NA}\NormalTok{,}
\NormalTok{    Q22C }\SpecialCharTok{==} \DecValTok{99} \SpecialCharTok{\textasciitilde{}} \ConstantTok{NA}
\NormalTok{  ))}

\NormalTok{Base\_Trajipaq\_Finale\_Ok }\OtherTok{\textless{}{-}}
\NormalTok{  Base\_Trajipaq\_Finale\_Ok }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{Q22D =} \FunctionTok{case\_when}\NormalTok{(}
\NormalTok{    Q22D }\SpecialCharTok{==} \DecValTok{1} \SpecialCharTok{\textasciitilde{}} \DecValTok{1}\NormalTok{,}
\NormalTok{    Q22D }\SpecialCharTok{==} \DecValTok{2} \SpecialCharTok{\textasciitilde{}} \DecValTok{2}\NormalTok{,}
\NormalTok{    Q22D }\SpecialCharTok{==} \DecValTok{3} \SpecialCharTok{\textasciitilde{}} \DecValTok{3}\NormalTok{,}
\NormalTok{    Q22D }\SpecialCharTok{==} \DecValTok{4} \SpecialCharTok{\textasciitilde{}} \DecValTok{4}\NormalTok{,}
\NormalTok{    Q22D }\SpecialCharTok{==} \DecValTok{98} \SpecialCharTok{\textasciitilde{}} \ConstantTok{NA}\NormalTok{,}
\NormalTok{    Q22D }\SpecialCharTok{==} \DecValTok{99} \SpecialCharTok{\textasciitilde{}} \ConstantTok{NA}
\NormalTok{  ))}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{v\_I4 }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}\StringTok{"Q22A"}\NormalTok{,}
          \StringTok{"Q22B"}\NormalTok{,}
          \StringTok{"Q22C"}\NormalTok{,}
          \StringTok{"Q22D"}
\NormalTok{          )}

\NormalTok{Base\_Trajipaq\_Finale\_Ok }\OtherTok{\textless{}{-}}
\NormalTok{  Base\_Trajipaq\_Finale\_Ok }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{rowwise}\NormalTok{() }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{I4 =}\NormalTok{ (}\FunctionTok{sum}\NormalTok{(}\FunctionTok{c\_across}\NormalTok{(}\FunctionTok{all\_of}\NormalTok{(v\_I4)), }\AttributeTok{na.rm =} \ConstantTok{TRUE}\NormalTok{) }\SpecialCharTok{{-}} \FunctionTok{length}\NormalTok{(v\_I4)) }\SpecialCharTok{/}\NormalTok{ (}\DecValTok{3}\SpecialCharTok{*}\FunctionTok{length}\NormalTok{(v\_I4))) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{I4 =} \ControlFlowTok{if}\NormalTok{ (I4 }\SpecialCharTok{\textless{}} \DecValTok{0}\NormalTok{) }\ConstantTok{NA\_real\_} \ControlFlowTok{else}\NormalTok{ I4) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{ungroup}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{descr}\NormalTok{(Base\_Trajipaq\_Finale\_Ok}\SpecialCharTok{$}\NormalTok{I4)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Descriptive Statistics  
## Base_Trajipaq_Finale_Ok$I4  
## N: 1563  
## 
##                          I4
## ----------------- ---------
##              Mean      0.73
##           Std.Dev      0.29
##               Min      0.00
##                Q1      0.58
##            Median      0.75
##                Q3      1.00
##               Max      1.00
##               MAD      0.37
##               IQR      0.42
##                CV      0.39
##          Skewness     -0.83
##       SE.Skewness      0.07
##          Kurtosis     -0.21
##           N.Valid   1106.00
##         Pct.Valid     70.76
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{ggplot}\NormalTok{(}\AttributeTok{data =}\NormalTok{ Base\_Trajipaq\_Finale\_Ok, }\AttributeTok{mapping =} \FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ I4)) }\SpecialCharTok{+}
  \FunctionTok{geom\_histogram}\NormalTok{(}\AttributeTok{na.rm =} \ConstantTok{TRUE}\NormalTok{, }\AttributeTok{fill =} \StringTok{"pink"}\NormalTok{, }\AttributeTok{binwidth =} \FloatTok{0.01}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{theme\_bw}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\includegraphics{Article---Sentiment-d-appartenance_files/figure-latex/unnamed-chunk-33-1.pdf}

La distribution est asymétrique vers la gauche. L'indice moyen de
maîtrise du français est 0.73.

\hypertarget{q23.-comment-est-votre-uxe9tat-de-santuxe9-en-guxe9nuxe9ral}{%
\subsection{Q23. Comment est votre état de santé en général
?}\label{q23.-comment-est-votre-uxe9tat-de-santuxe9-en-guxe9nuxe9ral}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Très bon
\item
  Bon
\item
  Moyen
\item
  Mauvais
\item
  Très mauvais
\item
  Je ne sais pas
\item
  Je préfère ne pas répondre
\end{enumerate}

Type de variable = factoriel Niveau de mesure = ordinal

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{Base\_Trajipaq\_Finale\_Ok }\OtherTok{\textless{}{-}}
\NormalTok{  Base\_Trajipaq\_Finale\_Ok }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{Q23 =} \FunctionTok{case\_when}\NormalTok{(}
\NormalTok{    Q23 }\SpecialCharTok{==} \DecValTok{1} \SpecialCharTok{\textasciitilde{}} \DecValTok{5}\NormalTok{,}
\NormalTok{    Q23 }\SpecialCharTok{==} \DecValTok{2} \SpecialCharTok{\textasciitilde{}} \DecValTok{4}\NormalTok{,}
\NormalTok{    Q23 }\SpecialCharTok{==} \DecValTok{3} \SpecialCharTok{\textasciitilde{}} \DecValTok{3}\NormalTok{,}
\NormalTok{    Q23 }\SpecialCharTok{==} \DecValTok{4} \SpecialCharTok{\textasciitilde{}} \DecValTok{2}\NormalTok{,}
\NormalTok{    Q23 }\SpecialCharTok{==} \DecValTok{5} \SpecialCharTok{\textasciitilde{}} \DecValTok{1}\NormalTok{,}
\NormalTok{    Q23 }\SpecialCharTok{==} \DecValTok{98} \SpecialCharTok{\textasciitilde{}} \ConstantTok{NA}\NormalTok{,}
\NormalTok{    Q23 }\SpecialCharTok{==} \DecValTok{99} \SpecialCharTok{\textasciitilde{}} \ConstantTok{NA}
\NormalTok{  ))}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{freq}\NormalTok{(Base\_Trajipaq\_Finale\_Ok}\SpecialCharTok{$}\NormalTok{Q23)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Frequencies  
## Base_Trajipaq_Finale_Ok$Q23  
## Type: Numeric  
## 
##               Freq   % Valid   % Valid Cum.   % Total   % Total Cum.
## ----------- ------ --------- -------------- --------- --------------
##           2      4      0.28           0.28      0.26           0.26
##           3     18      1.24           1.52      1.15           1.41
##           4    578     39.94          41.47     36.98          38.39
##           5    847     58.53         100.00     54.19          92.58
##        <NA>    116                               7.42         100.00
##       Total   1563    100.00         100.00    100.00         100.00
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{ggplot}\NormalTok{(}\AttributeTok{data =}\NormalTok{ Base\_Trajipaq\_Finale\_Ok, }\AttributeTok{mapping =} \FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ Q23)) }\SpecialCharTok{+}
  \FunctionTok{geom\_bar}\NormalTok{(}\AttributeTok{fill =} \StringTok{"pink"}\NormalTok{, }\AttributeTok{na.rm =} \ConstantTok{TRUE}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{theme\_bw}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\includegraphics{Article---Sentiment-d-appartenance_files/figure-latex/unnamed-chunk-36-1.pdf}

La distribution est asymétrique vers la gauche. 98.47\% des répondants
ont un état de santé moyen ou bon.

\hypertarget{i5.-indice-de-luxe9tat-de-santuxe9-au-cours-des-12-derniers-mois}{%
\subsection{I5. Indice de l'état de santé au cours des 12 derniers
mois}\label{i5.-indice-de-luxe9tat-de-santuxe9-au-cours-des-12-derniers-mois}}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{Base\_Trajipaq\_Finale\_Ok }\OtherTok{\textless{}{-}}
\NormalTok{  Base\_Trajipaq\_Finale\_Ok }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{Q24A =} \FunctionTok{case\_when}\NormalTok{(}
\NormalTok{    Q24A }\SpecialCharTok{==} \DecValTok{1} \SpecialCharTok{\textasciitilde{}} \DecValTok{1}\NormalTok{,}
\NormalTok{    Q24A }\SpecialCharTok{==} \DecValTok{2} \SpecialCharTok{\textasciitilde{}} \DecValTok{2}\NormalTok{,}
\NormalTok{    Q24A }\SpecialCharTok{==} \DecValTok{96} \SpecialCharTok{\textasciitilde{}} \ConstantTok{NA}\NormalTok{,}
\NormalTok{    Q24A }\SpecialCharTok{==} \DecValTok{97} \SpecialCharTok{\textasciitilde{}} \ConstantTok{NA}\NormalTok{,}
\NormalTok{    Q24A }\SpecialCharTok{==} \DecValTok{98} \SpecialCharTok{\textasciitilde{}} \ConstantTok{NA}\NormalTok{,}
\NormalTok{    Q24A }\SpecialCharTok{==} \DecValTok{99} \SpecialCharTok{\textasciitilde{}} \ConstantTok{NA}
\NormalTok{  ))}

\NormalTok{Base\_Trajipaq\_Finale\_Ok }\OtherTok{\textless{}{-}}
\NormalTok{  Base\_Trajipaq\_Finale\_Ok }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{Q24B =} \FunctionTok{case\_when}\NormalTok{(}
\NormalTok{    Q24B }\SpecialCharTok{==} \DecValTok{1} \SpecialCharTok{\textasciitilde{}} \DecValTok{1}\NormalTok{,}
\NormalTok{    Q24B }\SpecialCharTok{==} \DecValTok{2} \SpecialCharTok{\textasciitilde{}} \DecValTok{2}\NormalTok{,}
\NormalTok{    Q24B }\SpecialCharTok{==} \DecValTok{96} \SpecialCharTok{\textasciitilde{}} \ConstantTok{NA}\NormalTok{,}
\NormalTok{    Q24B }\SpecialCharTok{==} \DecValTok{97} \SpecialCharTok{\textasciitilde{}} \ConstantTok{NA}\NormalTok{,}
\NormalTok{    Q24B }\SpecialCharTok{==} \DecValTok{98} \SpecialCharTok{\textasciitilde{}} \ConstantTok{NA}\NormalTok{,}
\NormalTok{    Q24B }\SpecialCharTok{==} \DecValTok{99} \SpecialCharTok{\textasciitilde{}} \ConstantTok{NA}
\NormalTok{  ))}

\NormalTok{Base\_Trajipaq\_Finale\_Ok }\OtherTok{\textless{}{-}}
\NormalTok{  Base\_Trajipaq\_Finale\_Ok }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{Q24C =} \FunctionTok{case\_when}\NormalTok{(}
\NormalTok{    Q24C }\SpecialCharTok{==} \DecValTok{1} \SpecialCharTok{\textasciitilde{}} \DecValTok{1}\NormalTok{,}
\NormalTok{    Q24C }\SpecialCharTok{==} \DecValTok{2} \SpecialCharTok{\textasciitilde{}} \DecValTok{2}\NormalTok{,}
\NormalTok{    Q24C }\SpecialCharTok{==} \DecValTok{96} \SpecialCharTok{\textasciitilde{}} \ConstantTok{NA}\NormalTok{,}
\NormalTok{    Q24C }\SpecialCharTok{==} \DecValTok{97} \SpecialCharTok{\textasciitilde{}} \ConstantTok{NA}\NormalTok{,}
\NormalTok{    Q24C }\SpecialCharTok{==} \DecValTok{98} \SpecialCharTok{\textasciitilde{}} \ConstantTok{NA}\NormalTok{,}
\NormalTok{    Q24C }\SpecialCharTok{==} \DecValTok{99} \SpecialCharTok{\textasciitilde{}} \ConstantTok{NA}
\NormalTok{  ))}

\NormalTok{Base\_Trajipaq\_Finale\_Ok }\OtherTok{\textless{}{-}}
\NormalTok{  Base\_Trajipaq\_Finale\_Ok }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{Q24D =} \FunctionTok{case\_when}\NormalTok{(}
\NormalTok{    Q24D }\SpecialCharTok{==} \DecValTok{1} \SpecialCharTok{\textasciitilde{}} \DecValTok{1}\NormalTok{,}
\NormalTok{    Q24D }\SpecialCharTok{==} \DecValTok{2} \SpecialCharTok{\textasciitilde{}} \DecValTok{2}\NormalTok{,}
\NormalTok{    Q24D }\SpecialCharTok{==} \DecValTok{96} \SpecialCharTok{\textasciitilde{}} \ConstantTok{NA}\NormalTok{,}
\NormalTok{    Q24D }\SpecialCharTok{==} \DecValTok{97} \SpecialCharTok{\textasciitilde{}} \ConstantTok{NA}\NormalTok{,}
\NormalTok{    Q24D }\SpecialCharTok{==} \DecValTok{98} \SpecialCharTok{\textasciitilde{}} \ConstantTok{NA}\NormalTok{,}
\NormalTok{    Q24D }\SpecialCharTok{==} \DecValTok{99} \SpecialCharTok{\textasciitilde{}} \ConstantTok{NA}
\NormalTok{  ))}

\NormalTok{Base\_Trajipaq\_Finale\_Ok }\OtherTok{\textless{}{-}}
\NormalTok{  Base\_Trajipaq\_Finale\_Ok }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{Q24E =} \FunctionTok{case\_when}\NormalTok{(}
\NormalTok{    Q24E }\SpecialCharTok{==} \DecValTok{1} \SpecialCharTok{\textasciitilde{}} \DecValTok{1}\NormalTok{,}
\NormalTok{    Q24E }\SpecialCharTok{==} \DecValTok{2} \SpecialCharTok{\textasciitilde{}} \DecValTok{2}\NormalTok{,}
\NormalTok{    Q24E }\SpecialCharTok{==} \DecValTok{96} \SpecialCharTok{\textasciitilde{}} \ConstantTok{NA}\NormalTok{,}
\NormalTok{    Q24E }\SpecialCharTok{==} \DecValTok{97} \SpecialCharTok{\textasciitilde{}} \ConstantTok{NA}\NormalTok{,}
\NormalTok{    Q24E }\SpecialCharTok{==} \DecValTok{98} \SpecialCharTok{\textasciitilde{}} \ConstantTok{NA}\NormalTok{,}
\NormalTok{    Q24E }\SpecialCharTok{==} \DecValTok{99} \SpecialCharTok{\textasciitilde{}} \ConstantTok{NA}
\NormalTok{  ))}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{v\_I5 }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}\StringTok{"Q24A"}\NormalTok{,}
          \StringTok{"Q24B"}\NormalTok{,}
          \StringTok{"Q24C"}\NormalTok{,}
          \StringTok{"Q24D"}\NormalTok{,}
          \StringTok{"Q24E"}
\NormalTok{          )}

\NormalTok{Base\_Trajipaq\_Finale\_Ok }\OtherTok{\textless{}{-}}
\NormalTok{  Base\_Trajipaq\_Finale\_Ok }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{rowwise}\NormalTok{() }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{I5 =}\NormalTok{ (}\FunctionTok{sum}\NormalTok{(}\FunctionTok{c\_across}\NormalTok{(}\FunctionTok{all\_of}\NormalTok{(v\_I5)), }\AttributeTok{na.rm =} \ConstantTok{TRUE}\NormalTok{) }\SpecialCharTok{{-}} \FunctionTok{length}\NormalTok{(v\_I5)) }\SpecialCharTok{/}\NormalTok{ (}\DecValTok{1}\SpecialCharTok{*}\FunctionTok{length}\NormalTok{(v\_I5))) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{I5 =} \ControlFlowTok{if}\NormalTok{ (I5 }\SpecialCharTok{\textless{}} \DecValTok{0}\NormalTok{) }\ConstantTok{NA\_real\_} \ControlFlowTok{else}\NormalTok{ I5) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{ungroup}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{descr}\NormalTok{(Base\_Trajipaq\_Finale\_Ok}\SpecialCharTok{$}\NormalTok{I5)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Descriptive Statistics  
## Base_Trajipaq_Finale_Ok$I5  
## N: 1563  
## 
##                          I5
## ----------------- ---------
##              Mean      0.78
##           Std.Dev      0.29
##               Min      0.00
##                Q1      0.60
##            Median      1.00
##                Q3      1.00
##               Max      1.00
##               MAD      0.00
##               IQR      0.40
##                CV      0.37
##          Skewness     -1.05
##       SE.Skewness      0.06
##          Kurtosis      0.12
##           N.Valid   1430.00
##         Pct.Valid     91.49
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{ggplot}\NormalTok{(}\AttributeTok{data =}\NormalTok{ Base\_Trajipaq\_Finale\_Ok, }\AttributeTok{mapping =} \FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ I5)) }\SpecialCharTok{+}
  \FunctionTok{geom\_histogram}\NormalTok{(}\AttributeTok{na.rm =} \ConstantTok{TRUE}\NormalTok{, }\AttributeTok{fill =} \StringTok{"pink"}\NormalTok{, }\AttributeTok{binwidth =} \FloatTok{0.01}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{theme\_bw}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\includegraphics{Article---Sentiment-d-appartenance_files/figure-latex/unnamed-chunk-40-1.pdf}

La distribution est asymétrique vers la gauche. L'indice moyen de l'état
de santé au cours des 12 derniers mois est de 0.78.

\hypertarget{q26.-quel-uxe9tait-le-plus-haut-dipluxf4me-duxe9tudes-obtenu-par-votre-muxe8re}{%
\subsection{Q26. Quel était le plus haut diplôme d'études obtenu par
votre mère
?}\label{q26.-quel-uxe9tait-le-plus-haut-dipluxf4me-duxe9tudes-obtenu-par-votre-muxe8re}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Moins que le secondaire
\item
  Secondaire ou équivalent
\item
  Collégial
\item
  Formation professionnelle ou post-secondaire
\item
  Université
\item
  Inconnu
\item
  Je ne sais pas
\item
  Je préfère ne pas répondre
\end{enumerate}

Type de variable = factoriel Niveau de mesure = ordinal

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{Base\_Trajipaq\_Finale\_Ok }\OtherTok{\textless{}{-}}
\NormalTok{  Base\_Trajipaq\_Finale\_Ok }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{Q26 =} \FunctionTok{case\_when}\NormalTok{(}
\NormalTok{    Q26 }\SpecialCharTok{==} \DecValTok{1} \SpecialCharTok{\textasciitilde{}} \DecValTok{1}\NormalTok{,}
\NormalTok{    Q26 }\SpecialCharTok{==} \DecValTok{2} \SpecialCharTok{\textasciitilde{}} \DecValTok{2}\NormalTok{,}
\NormalTok{    Q26 }\SpecialCharTok{==} \DecValTok{3} \SpecialCharTok{\textasciitilde{}} \DecValTok{3}\NormalTok{,}
\NormalTok{    Q26 }\SpecialCharTok{==} \DecValTok{4} \SpecialCharTok{\textasciitilde{}} \DecValTok{4}\NormalTok{,}
\NormalTok{    Q26 }\SpecialCharTok{==} \DecValTok{5} \SpecialCharTok{\textasciitilde{}} \DecValTok{5}\NormalTok{,}
\NormalTok{    Q26 }\SpecialCharTok{==} \DecValTok{6} \SpecialCharTok{\textasciitilde{}} \ConstantTok{NA}\NormalTok{,}
\NormalTok{    Q26 }\SpecialCharTok{==} \DecValTok{98} \SpecialCharTok{\textasciitilde{}} \ConstantTok{NA}\NormalTok{,}
\NormalTok{    Q26 }\SpecialCharTok{==} \DecValTok{99} \SpecialCharTok{\textasciitilde{}} \ConstantTok{NA}
\NormalTok{  ))}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{freq}\NormalTok{(Base\_Trajipaq\_Finale\_Ok}\SpecialCharTok{$}\NormalTok{Q26)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Frequencies  
## Base_Trajipaq_Finale_Ok$Q26  
## Type: Numeric  
## 
##               Freq   % Valid   % Valid Cum.   % Total   % Total Cum.
## ----------- ------ --------- -------------- --------- --------------
##           1    238     16.49          16.49     15.23          15.23
##           2    355     24.60          41.09     22.71          37.94
##           3    241     16.70          57.80     15.42          53.36
##           4    196     13.58          71.38     12.54          65.90
##           5    413     28.62         100.00     26.42          92.32
##        <NA>    120                               7.68         100.00
##       Total   1563    100.00         100.00    100.00         100.00
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{ggplot}\NormalTok{(}\AttributeTok{data =}\NormalTok{ Base\_Trajipaq\_Finale\_Ok, }\AttributeTok{mapping =} \FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ Q26)) }\SpecialCharTok{+}
  \FunctionTok{geom\_bar}\NormalTok{(}\AttributeTok{fill =} \StringTok{"pink"}\NormalTok{, }\AttributeTok{na.rm =} \ConstantTok{TRUE}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{theme\_bw}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\includegraphics{Article---Sentiment-d-appartenance_files/figure-latex/unnamed-chunk-43-1.pdf}

La valeur modale est ``diplôme universitaire'' (28.62\%).

\hypertarget{q28.-quel-uxe9tait-le-plus-haut-dipluxf4me-duxe9tudes-obtenu-par-votre-puxe8re}{%
\subsection{Q28. Quel était le plus haut diplôme d'études obtenu par
votre père
?}\label{q28.-quel-uxe9tait-le-plus-haut-dipluxf4me-duxe9tudes-obtenu-par-votre-puxe8re}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Moins que le secondaire
\item
  Secondaire ou équivalent
\item
  Collégial
\item
  Formation professionnelle ou post-secondaire
\item
  Université
\item
  Inconnu
\item
  Je ne sais pas
\item
  Je préfère ne pas répondre
\end{enumerate}

Type de variable = factoriel Niveau de mesure = ordinal

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{Base\_Trajipaq\_Finale\_Ok }\OtherTok{\textless{}{-}}
\NormalTok{  Base\_Trajipaq\_Finale\_Ok }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{Q28 =} \FunctionTok{case\_when}\NormalTok{(}
\NormalTok{    Q28 }\SpecialCharTok{==} \DecValTok{1} \SpecialCharTok{\textasciitilde{}} \DecValTok{1}\NormalTok{,}
\NormalTok{    Q28 }\SpecialCharTok{==} \DecValTok{2} \SpecialCharTok{\textasciitilde{}} \DecValTok{2}\NormalTok{,}
\NormalTok{    Q28 }\SpecialCharTok{==} \DecValTok{3} \SpecialCharTok{\textasciitilde{}} \DecValTok{3}\NormalTok{,}
\NormalTok{    Q28 }\SpecialCharTok{==} \DecValTok{4} \SpecialCharTok{\textasciitilde{}} \DecValTok{4}\NormalTok{,}
\NormalTok{    Q28 }\SpecialCharTok{==} \DecValTok{5} \SpecialCharTok{\textasciitilde{}} \DecValTok{5}\NormalTok{,}
\NormalTok{    Q28 }\SpecialCharTok{==} \DecValTok{6} \SpecialCharTok{\textasciitilde{}} \ConstantTok{NA}\NormalTok{,}
\NormalTok{    Q28 }\SpecialCharTok{==} \DecValTok{98} \SpecialCharTok{\textasciitilde{}} \ConstantTok{NA}\NormalTok{,}
\NormalTok{    Q28 }\SpecialCharTok{==} \DecValTok{99} \SpecialCharTok{\textasciitilde{}} \ConstantTok{NA}
\NormalTok{  ))}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{freq}\NormalTok{(Base\_Trajipaq\_Finale\_Ok}\SpecialCharTok{$}\NormalTok{Q28)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Frequencies  
## Base_Trajipaq_Finale_Ok$Q28  
## Type: Numeric  
## 
##               Freq   % Valid   % Valid Cum.   % Total   % Total Cum.
## ----------- ------ --------- -------------- --------- --------------
##           1    206     14.56          14.56     13.18          13.18
##           2    288     20.35          34.91     18.43          31.61
##           3    201     14.20          49.12     12.86          44.47
##           4    229     16.18          65.30     14.65          59.12
##           5    491     34.70         100.00     31.41          90.53
##        <NA>    148                               9.47         100.00
##       Total   1563    100.00         100.00    100.00         100.00
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{ggplot}\NormalTok{(}\AttributeTok{data =}\NormalTok{ Base\_Trajipaq\_Finale\_Ok, }\AttributeTok{mapping =} \FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ Q28)) }\SpecialCharTok{+}
  \FunctionTok{geom\_bar}\NormalTok{(}\AttributeTok{fill =} \StringTok{"pink"}\NormalTok{, }\AttributeTok{na.rm =} \ConstantTok{TRUE}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{theme\_bw}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\includegraphics{Article---Sentiment-d-appartenance_files/figure-latex/unnamed-chunk-46-1.pdf}

La valeur modale est ``diplôme universitaire'' (34.70\%).

\hypertarget{ruxe9sumuxe9}{%
\subsection{Résumé}\label{ruxe9sumuxe9}}

L'échantillon possède les caractéristiques socio-démographiques
suivantes :

\begin{itemize}
\tightlist
\item
  faible prépondérance de femmes (56.69\%)
\item
  faible prépondérance de personnes âgées de 35 à 39 ans (24.63\%)
\item
  forte prépondérance de personnes mariées (47.53\%)
\item
  prépondérance de français (26.685) et d'algériens (6.649\%)
\item
  forte prépondérance de personnes nées au Canada (71.21\%)
\item
  forte prépondérance d'anciens travailleurs temporaires parmi les
  immigrés de première génération (57.79\%)
\item
  forte prépondérance de personnes possédant un diplôme universitaire
  (67.31\%)
\item
  prépondérance de personnes athées (36.36\%), catholiques (32.21\%) et
  musulmanes (13.83\%)
\item
  prépondérance de personnes ayant pour langue maternelle le français
  (63.84\%)
\item
  forte maîtrise moyenne du français (0.87)
\item
  forte maîtrise moyenne de l'anglais (0.73)
\item
  très forte prépondérance de personnes actuellement en bonne ou moyenne
  santé (98.47\%)
\item
  bon état de santé moyen durant les 12 derniers mois (0.78)
\item
  prépondérance d'un diplôme universitaire chez la mère (28.62\%)
\item
  prépondérance d'un diplôme universitaire chez la mère (34.70\%)
\end{itemize}

\hypertarget{pruxe9sentation-de-la-variable-induxe9pendante}{%
\section{Présentation de la variable
indépendante}\label{pruxe9sentation-de-la-variable-induxe9pendante}}

\hypertarget{q39a.-dites-si-vous-uxeates-tout-uxe0-fait-daccord-plutuxf4t-daccord-plutuxf4t-pas-daccord-ou-pas-du-tout-daccord-avec-les-propositions-suivantes-participation-identitaire-et-sentiment-dappartenance-je-me-sens-chez-moi-au-quuxe9bec.}{%
\subsection{Q39A. Dites si vous êtes tout à fait d'accord, plutôt
d'accord, plutôt pas d'accord ou pas du tout d'accord avec les
propositions suivantes (participation identitaire et sentiment
d'appartenance) : je me sens chez moi au
Québec.}\label{q39a.-dites-si-vous-uxeates-tout-uxe0-fait-daccord-plutuxf4t-daccord-plutuxf4t-pas-daccord-ou-pas-du-tout-daccord-avec-les-propositions-suivantes-participation-identitaire-et-sentiment-dappartenance-je-me-sens-chez-moi-au-quuxe9bec.}}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{Base\_Trajipaq\_Finale\_Ok }\OtherTok{\textless{}{-}}
\NormalTok{  Base\_Trajipaq\_Finale\_Ok }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{Q39A =} \FunctionTok{case\_when}\NormalTok{(}
\NormalTok{    Q39A }\SpecialCharTok{==} \DecValTok{1} \SpecialCharTok{\textasciitilde{}} \DecValTok{4}\NormalTok{,}
\NormalTok{    Q39A }\SpecialCharTok{==} \DecValTok{2} \SpecialCharTok{\textasciitilde{}} \DecValTok{3}\NormalTok{,}
\NormalTok{    Q39A }\SpecialCharTok{==} \DecValTok{3} \SpecialCharTok{\textasciitilde{}} \DecValTok{2}\NormalTok{,}
\NormalTok{    Q39A }\SpecialCharTok{==} \DecValTok{4} \SpecialCharTok{\textasciitilde{}} \DecValTok{1}\NormalTok{,}
\NormalTok{    Q39A }\SpecialCharTok{==} \DecValTok{98} \SpecialCharTok{\textasciitilde{}} \ConstantTok{NA}\NormalTok{,}
\NormalTok{    Q39A }\SpecialCharTok{==} \DecValTok{99} \SpecialCharTok{\textasciitilde{}} \ConstantTok{NA}
\NormalTok{  ))}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{freq}\NormalTok{(Base\_Trajipaq\_Finale\_Ok}\SpecialCharTok{$}\NormalTok{Q39A)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Frequencies  
## Base_Trajipaq_Finale_Ok$Q39A  
## Type: Numeric  
## 
##               Freq   % Valid   % Valid Cum.   % Total   % Total Cum.
## ----------- ------ --------- -------------- --------- --------------
##           1     59      3.84           3.84      3.77           3.77
##           2    116      7.56          11.40      7.42          11.20
##           3    513     33.42          44.82     32.82          44.02
##           4    847     55.18         100.00     54.19          98.21
##        <NA>     28                               1.79         100.00
##       Total   1563    100.00         100.00    100.00         100.00
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{ggplot}\NormalTok{(}\AttributeTok{data =}\NormalTok{ Base\_Trajipaq\_Finale\_Ok, }\AttributeTok{mapping =} \FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ Q39A)) }\SpecialCharTok{+}
  \FunctionTok{geom\_bar}\NormalTok{(}\AttributeTok{fill =} \StringTok{"pink"}\NormalTok{,}
           \AttributeTok{na.rm =} \ConstantTok{TRUE}
\NormalTok{           ) }\SpecialCharTok{+}
  \FunctionTok{theme\_bw}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\includegraphics{Article---Sentiment-d-appartenance_files/figure-latex/unnamed-chunk-49-1.pdf}

La distribution est asymétrique vers la gauche. 87.01\% des personnes
interrogées sont tout à fait d'accord ou plutôt d'accord pour dire
qu'elles se sentent chez elles au Québec.

\hypertarget{q39b.-dites-si-vous-uxeates-tout-uxe0-fait-daccord-plutuxf4t-daccord-plutuxf4t-pas-daccord-ou-pas-du-tout-daccord-avec-les-propositions-suivantes-participation-identitaire-et-sentiment-dappartenance-il-faut-faire-oublier-ses-origines-pour-se-faire-accepter-au-quuxe9bec.}{%
\subsection{Q39B. Dites si vous êtes tout à fait d'accord, plutôt
d'accord, plutôt pas d'accord ou pas du tout d'accord avec les
propositions suivantes (participation identitaire et sentiment
d'appartenance) : il faut faire oublier ses origines pour se faire
accepter au
Québec.}\label{q39b.-dites-si-vous-uxeates-tout-uxe0-fait-daccord-plutuxf4t-daccord-plutuxf4t-pas-daccord-ou-pas-du-tout-daccord-avec-les-propositions-suivantes-participation-identitaire-et-sentiment-dappartenance-il-faut-faire-oublier-ses-origines-pour-se-faire-accepter-au-quuxe9bec.}}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{Base\_Trajipaq\_Finale\_Ok }\OtherTok{\textless{}{-}}
\NormalTok{  Base\_Trajipaq\_Finale\_Ok }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{Q39B =} \FunctionTok{case\_when}\NormalTok{(}
\NormalTok{    Q39B }\SpecialCharTok{==} \DecValTok{1} \SpecialCharTok{\textasciitilde{}} \DecValTok{4}\NormalTok{,}
\NormalTok{    Q39B }\SpecialCharTok{==} \DecValTok{2} \SpecialCharTok{\textasciitilde{}} \DecValTok{3}\NormalTok{,}
\NormalTok{    Q39B }\SpecialCharTok{==} \DecValTok{3} \SpecialCharTok{\textasciitilde{}} \DecValTok{2}\NormalTok{,}
\NormalTok{    Q39B }\SpecialCharTok{==} \DecValTok{4} \SpecialCharTok{\textasciitilde{}} \DecValTok{1}\NormalTok{,}
\NormalTok{    Q39B }\SpecialCharTok{==} \DecValTok{98} \SpecialCharTok{\textasciitilde{}} \ConstantTok{NA}\NormalTok{,}
\NormalTok{    Q39B }\SpecialCharTok{==} \DecValTok{99} \SpecialCharTok{\textasciitilde{}} \ConstantTok{NA}
\NormalTok{  ))}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{freq}\NormalTok{(Base\_Trajipaq\_Finale\_Ok}\SpecialCharTok{$}\NormalTok{Q39B)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Frequencies  
## Base_Trajipaq_Finale_Ok$Q39B  
## Type: Numeric  
## 
##               Freq   % Valid   % Valid Cum.   % Total   % Total Cum.
## ----------- ------ --------- -------------- --------- --------------
##           1    671     44.70          44.70     42.93          42.93
##           2    390     25.98          70.69     24.95          67.88
##           3    298     19.85          90.54     19.07          86.95
##           4    142      9.46         100.00      9.09          96.03
##        <NA>     62                               3.97         100.00
##       Total   1563    100.00         100.00    100.00         100.00
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{ggplot}\NormalTok{(}\AttributeTok{data =}\NormalTok{ Base\_Trajipaq\_Finale\_Ok, }\AttributeTok{mapping =} \FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ Q39B)) }\SpecialCharTok{+}
  \FunctionTok{geom\_bar}\NormalTok{(}\AttributeTok{fill =} \StringTok{"pink"}\NormalTok{,}
           \AttributeTok{na.rm =} \ConstantTok{TRUE}
\NormalTok{           ) }\SpecialCharTok{+}
  \FunctionTok{theme\_bw}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\includegraphics{Article---Sentiment-d-appartenance_files/figure-latex/unnamed-chunk-52-1.pdf}

La distribution est asymétrique vers la droite. 67.88\% des personnes
interrogées ne sont pas du tout d'accord ou plutôt pas d'accord pour
dire qu'il faut faire oublier ses origines pour se faire accepter au
Québec.

\hypertarget{q39c.-dites-si-vous-uxeates-tout-uxe0-fait-daccord-plutuxf4t-daccord-plutuxf4t-pas-daccord-ou-pas-du-tout-daccord-avec-les-propositions-suivantes-participation-identitaire-et-sentiment-dappartenance-on-me-voit-comme-une-quuxe9buxe9coise.}{%
\subsection{Q39C. Dites si vous êtes tout à fait d'accord, plutôt
d'accord, plutôt pas d'accord ou pas du tout d'accord avec les
propositions suivantes (participation identitaire et sentiment
d'appartenance) : on me voit comme une
Québécoise.}\label{q39c.-dites-si-vous-uxeates-tout-uxe0-fait-daccord-plutuxf4t-daccord-plutuxf4t-pas-daccord-ou-pas-du-tout-daccord-avec-les-propositions-suivantes-participation-identitaire-et-sentiment-dappartenance-on-me-voit-comme-une-quuxe9buxe9coise.}}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{Base\_Trajipaq\_Finale\_Ok }\OtherTok{\textless{}{-}}
\NormalTok{  Base\_Trajipaq\_Finale\_Ok }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{Q39C =} \FunctionTok{case\_when}\NormalTok{(}
\NormalTok{    Q39C }\SpecialCharTok{==} \DecValTok{1} \SpecialCharTok{\textasciitilde{}} \DecValTok{4}\NormalTok{,}
\NormalTok{    Q39C }\SpecialCharTok{==} \DecValTok{2} \SpecialCharTok{\textasciitilde{}} \DecValTok{3}\NormalTok{,}
\NormalTok{    Q39C }\SpecialCharTok{==} \DecValTok{3} \SpecialCharTok{\textasciitilde{}} \DecValTok{2}\NormalTok{,}
\NormalTok{    Q39C }\SpecialCharTok{==} \DecValTok{4} \SpecialCharTok{\textasciitilde{}} \DecValTok{1}\NormalTok{,}
\NormalTok{    Q39C }\SpecialCharTok{==} \DecValTok{98} \SpecialCharTok{\textasciitilde{}} \ConstantTok{NA}\NormalTok{,}
\NormalTok{    Q39C }\SpecialCharTok{==} \DecValTok{99} \SpecialCharTok{\textasciitilde{}} \ConstantTok{NA}
\NormalTok{  ))}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{freq}\NormalTok{(Base\_Trajipaq\_Finale\_Ok}\SpecialCharTok{$}\NormalTok{Q39C)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Frequencies  
## Base_Trajipaq_Finale_Ok$Q39C  
## Type: Numeric  
## 
##               Freq   % Valid   % Valid Cum.   % Total   % Total Cum.
## ----------- ------ --------- -------------- --------- --------------
##           1    442     29.97          29.97     28.28          28.28
##           2    341     23.12          53.08     21.82          50.10
##           3    259     17.56          70.64     16.57          66.67
##           4    433     29.36         100.00     27.70          94.37
##        <NA>     88                               5.63         100.00
##       Total   1563    100.00         100.00    100.00         100.00
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{ggplot}\NormalTok{(}\AttributeTok{data =}\NormalTok{ Base\_Trajipaq\_Finale\_Ok, }\AttributeTok{mapping =} \FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ Q39C)) }\SpecialCharTok{+}
  \FunctionTok{geom\_bar}\NormalTok{(}\AttributeTok{fill =} \StringTok{"pink"}\NormalTok{,}
           \AttributeTok{na.rm =} \ConstantTok{TRUE}
\NormalTok{           ) }\SpecialCharTok{+}
  \FunctionTok{theme\_bw}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\includegraphics{Article---Sentiment-d-appartenance_files/figure-latex/unnamed-chunk-55-1.pdf}

Les réponses modales sont ``tout à fait d'accord'' (29.36\%) et ``pas du
tout d'accord'' (29.97\%).

\hypertarget{q39d.-dites-si-vous-uxeates-tout-uxe0-fait-daccord-plutuxf4t-daccord-plutuxf4t-pas-daccord-ou-pas-du-tout-daccord-avec-les-propositions-suivantes-participation-identitaire-et-sentiment-dappartenance-je-me-sens-quuxe9buxe9coise.}{%
\subsection{Q39D. Dites si vous êtes tout à fait d'accord, plutôt
d'accord, plutôt pas d'accord ou pas du tout d'accord avec les
propositions suivantes (participation identitaire et sentiment
d'appartenance) : je me sens
Québécois(e).}\label{q39d.-dites-si-vous-uxeates-tout-uxe0-fait-daccord-plutuxf4t-daccord-plutuxf4t-pas-daccord-ou-pas-du-tout-daccord-avec-les-propositions-suivantes-participation-identitaire-et-sentiment-dappartenance-je-me-sens-quuxe9buxe9coise.}}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{Base\_Trajipaq\_Finale\_Ok }\OtherTok{\textless{}{-}}
\NormalTok{  Base\_Trajipaq\_Finale\_Ok }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{Q39D =} \FunctionTok{case\_when}\NormalTok{(}
\NormalTok{    Q39D }\SpecialCharTok{==} \DecValTok{1} \SpecialCharTok{\textasciitilde{}} \DecValTok{4}\NormalTok{,}
\NormalTok{    Q39D }\SpecialCharTok{==} \DecValTok{2} \SpecialCharTok{\textasciitilde{}} \DecValTok{3}\NormalTok{,}
\NormalTok{    Q39D }\SpecialCharTok{==} \DecValTok{3} \SpecialCharTok{\textasciitilde{}} \DecValTok{2}\NormalTok{,}
\NormalTok{    Q39D }\SpecialCharTok{==} \DecValTok{4} \SpecialCharTok{\textasciitilde{}} \DecValTok{1}\NormalTok{,}
\NormalTok{    Q39D }\SpecialCharTok{==} \DecValTok{98} \SpecialCharTok{\textasciitilde{}} \ConstantTok{NA}\NormalTok{,}
\NormalTok{    Q39D }\SpecialCharTok{==} \DecValTok{99} \SpecialCharTok{\textasciitilde{}} \ConstantTok{NA}
\NormalTok{  ))}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{freq}\NormalTok{(Base\_Trajipaq\_Finale\_Ok}\SpecialCharTok{$}\NormalTok{Q39D)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Frequencies  
## Base_Trajipaq_Finale_Ok$Q39D  
## Type: Numeric  
## 
##               Freq   % Valid   % Valid Cum.   % Total   % Total Cum.
## ----------- ------ --------- -------------- --------- --------------
##           1    259     17.27          17.27     16.57          16.57
##           2    292     19.47          36.73     18.68          35.25
##           3    406     27.07          63.80     25.98          61.23
##           4    543     36.20         100.00     34.74          95.97
##        <NA>     63                               4.03         100.00
##       Total   1563    100.00         100.00    100.00         100.00
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{ggplot}\NormalTok{(}\AttributeTok{data =}\NormalTok{ Base\_Trajipaq\_Finale\_Ok, }\AttributeTok{mapping =} \FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ Q39D)) }\SpecialCharTok{+}
  \FunctionTok{geom\_bar}\NormalTok{(}\AttributeTok{fill =} \StringTok{"pink"}\NormalTok{,}
           \AttributeTok{na.rm =} \ConstantTok{TRUE}
\NormalTok{           ) }\SpecialCharTok{+}
  \FunctionTok{theme\_bw}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\includegraphics{Article---Sentiment-d-appartenance_files/figure-latex/unnamed-chunk-58-1.pdf}

La distribution est asymétrique vers la gauche. 60.72\% des personnes
interrogées sont tout à fait d'accord ou plutôt d'accord pour dire
qu'elles se sentent Québécois(e)s.

\hypertarget{q39e.-dites-si-vous-uxeates-tout-uxe0-fait-daccord-plutuxf4t-daccord-plutuxf4t-pas-daccord-ou-pas-du-tout-daccord-avec-les-propositions-suivantes-participation-identitaire-et-sentiment-dappartenance-je-me-sens-canadienne.}{%
\subsection{Q39E. Dites si vous êtes tout à fait d'accord, plutôt
d'accord, plutôt pas d'accord ou pas du tout d'accord avec les
propositions suivantes (participation identitaire et sentiment
d'appartenance) : je me sens
Canadien(ne).}\label{q39e.-dites-si-vous-uxeates-tout-uxe0-fait-daccord-plutuxf4t-daccord-plutuxf4t-pas-daccord-ou-pas-du-tout-daccord-avec-les-propositions-suivantes-participation-identitaire-et-sentiment-dappartenance-je-me-sens-canadienne.}}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{Base\_Trajipaq\_Finale\_Ok }\OtherTok{\textless{}{-}}
\NormalTok{  Base\_Trajipaq\_Finale\_Ok }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{Q39E =} \FunctionTok{case\_when}\NormalTok{(}
\NormalTok{    Q39E }\SpecialCharTok{==} \DecValTok{1} \SpecialCharTok{\textasciitilde{}} \DecValTok{4}\NormalTok{,}
\NormalTok{    Q39E }\SpecialCharTok{==} \DecValTok{2} \SpecialCharTok{\textasciitilde{}} \DecValTok{3}\NormalTok{,}
\NormalTok{    Q39E }\SpecialCharTok{==} \DecValTok{3} \SpecialCharTok{\textasciitilde{}} \DecValTok{2}\NormalTok{,}
\NormalTok{    Q39E }\SpecialCharTok{==} \DecValTok{4} \SpecialCharTok{\textasciitilde{}} \DecValTok{1}\NormalTok{,}
\NormalTok{    Q39E }\SpecialCharTok{==} \DecValTok{98} \SpecialCharTok{\textasciitilde{}} \ConstantTok{NA}\NormalTok{,}
\NormalTok{    Q39E }\SpecialCharTok{==} \DecValTok{99} \SpecialCharTok{\textasciitilde{}} \ConstantTok{NA}
\NormalTok{  ))}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{freq}\NormalTok{(Base\_Trajipaq\_Finale\_Ok}\SpecialCharTok{$}\NormalTok{Q39E)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Frequencies  
## Base_Trajipaq_Finale_Ok$Q39E  
## Type: Numeric  
## 
##               Freq   % Valid   % Valid Cum.   % Total   % Total Cum.
## ----------- ------ --------- -------------- --------- --------------
##           1    100      6.57           6.57      6.40           6.40
##           2    171     11.24          17.82     10.94          17.34
##           3    550     36.16          53.98     35.19          52.53
##           4    700     46.02         100.00     44.79          97.31
##        <NA>     42                               2.69         100.00
##       Total   1563    100.00         100.00    100.00         100.00
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{ggplot}\NormalTok{(}\AttributeTok{data =}\NormalTok{ Base\_Trajipaq\_Finale\_Ok, }\AttributeTok{mapping =} \FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ Q39E)) }\SpecialCharTok{+}
  \FunctionTok{geom\_bar}\NormalTok{(}\AttributeTok{fill =} \StringTok{"pink"}\NormalTok{,}
           \AttributeTok{na.rm =} \ConstantTok{TRUE}
\NormalTok{           ) }\SpecialCharTok{+}
  \FunctionTok{theme\_bw}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\includegraphics{Article---Sentiment-d-appartenance_files/figure-latex/unnamed-chunk-61-1.pdf}

La distribution est asymétrique vers la gauche. 79.98\% des personnes
interrogées sont tout à fait d'accord ou plutôt d'accord pour dire
qu'elles se sentent Canadien(ne).

\hypertarget{q39f.-dites-si-vous-uxeates-tout-uxe0-fait-daccord-plutuxf4t-daccord-plutuxf4t-pas-daccord-ou-pas-du-tout-daccord-avec-les-propositions-suivantes-participation-identitaire-et-sentiment-dappartenance-jai-un-sentiment-dappartenance-uxe0-mon-pays-dorigine.}{%
\subsection{Q39F. Dites si vous êtes tout à fait d'accord, plutôt
d'accord, plutôt pas d'accord ou pas du tout d'accord avec les
propositions suivantes (participation identitaire et sentiment
d'appartenance) : j'ai un sentiment d'appartenance à mon pays
d'origine.}\label{q39f.-dites-si-vous-uxeates-tout-uxe0-fait-daccord-plutuxf4t-daccord-plutuxf4t-pas-daccord-ou-pas-du-tout-daccord-avec-les-propositions-suivantes-participation-identitaire-et-sentiment-dappartenance-jai-un-sentiment-dappartenance-uxe0-mon-pays-dorigine.}}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{Base\_Trajipaq\_Finale\_Ok }\OtherTok{\textless{}{-}}
\NormalTok{  Base\_Trajipaq\_Finale\_Ok }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{Q39F =} \FunctionTok{case\_when}\NormalTok{(}
\NormalTok{    Q39F }\SpecialCharTok{==} \DecValTok{1} \SpecialCharTok{\textasciitilde{}} \DecValTok{4}\NormalTok{,}
\NormalTok{    Q39F }\SpecialCharTok{==} \DecValTok{2} \SpecialCharTok{\textasciitilde{}} \DecValTok{3}\NormalTok{,}
\NormalTok{    Q39F }\SpecialCharTok{==} \DecValTok{3} \SpecialCharTok{\textasciitilde{}} \DecValTok{2}\NormalTok{,}
\NormalTok{    Q39F }\SpecialCharTok{==} \DecValTok{4} \SpecialCharTok{\textasciitilde{}} \DecValTok{1}\NormalTok{,}
\NormalTok{    Q39F }\SpecialCharTok{==} \DecValTok{98} \SpecialCharTok{\textasciitilde{}} \ConstantTok{NA}\NormalTok{,}
\NormalTok{    Q39F }\SpecialCharTok{==} \DecValTok{99} \SpecialCharTok{\textasciitilde{}} \ConstantTok{NA}
\NormalTok{  ))}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{freq}\NormalTok{(Base\_Trajipaq\_Finale\_Ok}\SpecialCharTok{$}\NormalTok{Q39F)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Frequencies  
## Base_Trajipaq_Finale_Ok$Q39F  
## Type: Numeric  
## 
##               Freq   % Valid   % Valid Cum.   % Total   % Total Cum.
## ----------- ------ --------- -------------- --------- --------------
##           1     82      5.44           5.44      5.25           5.25
##           2    198     13.14          18.58     12.67          17.91
##           3    550     36.50          55.08     35.19          53.10
##           4    677     44.92         100.00     43.31          96.42
##        <NA>     56                               3.58         100.00
##       Total   1563    100.00         100.00    100.00         100.00
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{ggplot}\NormalTok{(}\AttributeTok{data =}\NormalTok{ Base\_Trajipaq\_Finale\_Ok, }\AttributeTok{mapping =} \FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ Q39F)) }\SpecialCharTok{+}
  \FunctionTok{geom\_bar}\NormalTok{(}\AttributeTok{fill =} \StringTok{"pink"}\NormalTok{,}
           \AttributeTok{na.rm =} \ConstantTok{TRUE}
\NormalTok{           ) }\SpecialCharTok{+}
  \FunctionTok{theme\_bw}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\includegraphics{Article---Sentiment-d-appartenance_files/figure-latex/unnamed-chunk-64-1.pdf}

La distribution est asymétrique vers la gauche. 81.42\% des personnes
interrogées sont tout à fait d'accord ou plutôt d'accord pour dire
qu'elles possèdent un sentiment d'appartenance envers leur pays
d'origine.

\hypertarget{i1-lindice-du-sentiment-dappartenance-au-quuxe9bec}{%
\subsection{I1 : L'indice du sentiment d'appartenance au
Québec}\label{i1-lindice-du-sentiment-dappartenance-au-quuxe9bec}}

A la différence des variables précédentes, l'indice du sentiment
d'appartenance est une variable numérique continue dont le niveau de
mesure est ratio.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{v\_I1 }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}\StringTok{"Q39A"}\NormalTok{,}
          \StringTok{"Q39B"}\NormalTok{,}
          \StringTok{"Q39C"}\NormalTok{,}
          \StringTok{"Q39D"}\NormalTok{,}
          \StringTok{"Q39E"}\NormalTok{,}
          \StringTok{"Q39F"}
\NormalTok{          )}

\NormalTok{Base\_Trajipaq\_Finale\_Ok }\OtherTok{\textless{}{-}}
\NormalTok{  Base\_Trajipaq\_Finale\_Ok }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{rowwise}\NormalTok{() }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{I1 =}\NormalTok{ (}\FunctionTok{sum}\NormalTok{(}\FunctionTok{c\_across}\NormalTok{(}\FunctionTok{all\_of}\NormalTok{(v\_I1)), }\AttributeTok{na.rm =} \ConstantTok{TRUE}\NormalTok{) }\SpecialCharTok{{-}} \FunctionTok{length}\NormalTok{(v\_I1)) }\SpecialCharTok{/}\NormalTok{ (}\DecValTok{3}\SpecialCharTok{*}\FunctionTok{length}\NormalTok{(v\_I1))) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{I1 =} \ControlFlowTok{if}\NormalTok{ (I1 }\SpecialCharTok{\textless{}} \DecValTok{0}\NormalTok{) }\ConstantTok{NA\_real\_} \ControlFlowTok{else}\NormalTok{ I1) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{ungroup}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{descr}\NormalTok{(Base\_Trajipaq\_Finale\_Ok}\SpecialCharTok{$}\NormalTok{I1)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Descriptive Statistics  
## Base_Trajipaq_Finale_Ok$I1  
## N: 1563  
## 
##                          I1
## ----------------- ---------
##              Mean      0.59
##           Std.Dev      0.21
##               Min      0.00
##                Q1      0.44
##            Median      0.61
##                Q3      0.72
##               Max      1.00
##               MAD      0.25
##               IQR      0.28
##                CV      0.35
##          Skewness     -0.30
##       SE.Skewness      0.06
##          Kurtosis     -0.41
##           N.Valid   1550.00
##         Pct.Valid     99.17
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{ggplot}\NormalTok{(}\AttributeTok{data =}\NormalTok{ Base\_Trajipaq\_Finale\_Ok, }\AttributeTok{mapping =} \FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ I1)) }\SpecialCharTok{+}
  \FunctionTok{geom\_histogram}\NormalTok{(}\AttributeTok{na.rm =} \ConstantTok{TRUE}\NormalTok{, }\AttributeTok{fill =} \StringTok{"pink"}\NormalTok{, }\AttributeTok{binwidth =} \FloatTok{0.01}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{theme\_bw}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\includegraphics{Article---Sentiment-d-appartenance_files/figure-latex/unnamed-chunk-67-1.pdf}

La distribution est asymétrique vers la gauche. L'indice moyen du
sentiment d'appartenance au Québec est 0.59.

\hypertarget{pruxe9sentation-de-la-variable-duxe9pendante}{%
\section{Présentation de la variable
dépendante}\label{pruxe9sentation-de-la-variable-duxe9pendante}}

\hypertarget{q40a.-au-cours-des-12-mois-avant-la-crise-de-la-covid-19-pensez-vous-avoir-subi-des-traitements-inuxe9galitaires-ou-des-discriminations-un-traitement-inuxe9galitaire-ou-une-discrimination-se-repuxe8re-si-vous-avez-uxe9tuxe9-moins-bien-traituxe9-que-les-autres-sans-raison-valable-dans-lemploi}{%
\subsection{Q40A. Au cours des 12 mois (avant la crise de la COVID-19),
pensez-vous avoir subi des traitements inégalitaires ou des
discriminations ? Un traitement inégalitaire ou une discrimination se
repère si « vous avez été moins bien traité que les autres, sans raison
valable » : dans l'emploi
?}\label{q40a.-au-cours-des-12-mois-avant-la-crise-de-la-covid-19-pensez-vous-avoir-subi-des-traitements-inuxe9galitaires-ou-des-discriminations-un-traitement-inuxe9galitaire-ou-une-discrimination-se-repuxe8re-si-vous-avez-uxe9tuxe9-moins-bien-traituxe9-que-les-autres-sans-raison-valable-dans-lemploi}}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{Base\_Trajipaq\_Finale\_Ok }\OtherTok{\textless{}{-}}
\NormalTok{  Base\_Trajipaq\_Finale\_Ok }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{Q40A =} \FunctionTok{case\_when}\NormalTok{(}
\NormalTok{    Q40A }\SpecialCharTok{==} \DecValTok{1} \SpecialCharTok{\textasciitilde{}} \DecValTok{3}\NormalTok{,}
\NormalTok{    Q40A }\SpecialCharTok{==} \DecValTok{2} \SpecialCharTok{\textasciitilde{}} \DecValTok{2}\NormalTok{,}
\NormalTok{    Q40A }\SpecialCharTok{==} \DecValTok{3} \SpecialCharTok{\textasciitilde{}} \DecValTok{1}\NormalTok{,}
\NormalTok{    Q40A }\SpecialCharTok{==} \DecValTok{96} \SpecialCharTok{\textasciitilde{}} \ConstantTok{NA}\NormalTok{,}
\NormalTok{    Q40A }\SpecialCharTok{==} \DecValTok{98} \SpecialCharTok{\textasciitilde{}} \ConstantTok{NA}\NormalTok{,}
\NormalTok{    Q40A }\SpecialCharTok{==} \DecValTok{99} \SpecialCharTok{\textasciitilde{}} \ConstantTok{NA}
\NormalTok{  ))}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{freq}\NormalTok{(Base\_Trajipaq\_Finale\_Ok}\SpecialCharTok{$}\NormalTok{Q40A)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Frequencies  
## Base_Trajipaq_Finale_Ok$Q40A  
## Type: Numeric  
## 
##               Freq   % Valid   % Valid Cum.   % Total   % Total Cum.
## ----------- ------ --------- -------------- --------- --------------
##           1    926     65.44          65.44     59.25          59.25
##           2    383     27.07          92.51     24.50          83.75
##           3    106      7.49         100.00      6.78          90.53
##        <NA>    148                               9.47         100.00
##       Total   1563    100.00         100.00    100.00         100.00
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{ggplot}\NormalTok{(}\AttributeTok{data =}\NormalTok{ Base\_Trajipaq\_Finale\_Ok, }\AttributeTok{mapping =} \FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ Q40A)) }\SpecialCharTok{+}
  \FunctionTok{geom\_bar}\NormalTok{(}\AttributeTok{fill =} \StringTok{"pink"}\NormalTok{, }\AttributeTok{na.rm =} \ConstantTok{TRUE}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{theme\_bw}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\includegraphics{Article---Sentiment-d-appartenance_files/figure-latex/unnamed-chunk-70-1.pdf}

La distribution est asymétrique vers la droite. 65.44\% des personnes
interrogées disent ne jamais avoir subit de discrimination dans l'emploi
au cours des 12 derniers mois précédent la crise du covid 19.

\hypertarget{q40b.-au-cours-des-12-mois-avant-la-crise-de-la-covid-19-pensez-vous-avoir-subi-des-traitements-inuxe9galitaires-ou-des-discriminations-un-traitement-inuxe9galitaire-ou-une-discrimination-se-repuxe8re-si-vous-avez-uxe9tuxe9-moins-bien-traituxe9-que-les-autres-sans-raison-valable-dans-la-recherche-de-logement}{%
\subsection{Q40B. Au cours des 12 mois (avant la crise de la COVID-19),
pensez-vous avoir subi des traitements inégalitaires ou des
discriminations ? Un traitement inégalitaire ou une discrimination se
repère si « vous avez été moins bien traité que les autres, sans raison
valable » : dans la recherche de logement
?}\label{q40b.-au-cours-des-12-mois-avant-la-crise-de-la-covid-19-pensez-vous-avoir-subi-des-traitements-inuxe9galitaires-ou-des-discriminations-un-traitement-inuxe9galitaire-ou-une-discrimination-se-repuxe8re-si-vous-avez-uxe9tuxe9-moins-bien-traituxe9-que-les-autres-sans-raison-valable-dans-la-recherche-de-logement}}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{Base\_Trajipaq\_Finale\_Ok }\OtherTok{\textless{}{-}}
\NormalTok{  Base\_Trajipaq\_Finale\_Ok }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{Q40B =} \FunctionTok{case\_when}\NormalTok{(}
\NormalTok{    Q40B }\SpecialCharTok{==} \DecValTok{1} \SpecialCharTok{\textasciitilde{}} \DecValTok{3}\NormalTok{,}
\NormalTok{    Q40B }\SpecialCharTok{==} \DecValTok{2} \SpecialCharTok{\textasciitilde{}} \DecValTok{2}\NormalTok{,}
\NormalTok{    Q40B }\SpecialCharTok{==} \DecValTok{3} \SpecialCharTok{\textasciitilde{}} \DecValTok{1}\NormalTok{,}
\NormalTok{    Q40B }\SpecialCharTok{==} \DecValTok{96} \SpecialCharTok{\textasciitilde{}} \ConstantTok{NA}\NormalTok{,}
\NormalTok{    Q40B }\SpecialCharTok{==} \DecValTok{98} \SpecialCharTok{\textasciitilde{}} \ConstantTok{NA}\NormalTok{,}
\NormalTok{    Q40B }\SpecialCharTok{==} \DecValTok{99} \SpecialCharTok{\textasciitilde{}} \ConstantTok{NA}
\NormalTok{  ))}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{freq}\NormalTok{(Base\_Trajipaq\_Finale\_Ok}\SpecialCharTok{$}\NormalTok{Q40B)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Frequencies  
## Base_Trajipaq_Finale_Ok$Q40B  
## Type: Numeric  
## 
##               Freq   % Valid   % Valid Cum.   % Total   % Total Cum.
## ----------- ------ --------- -------------- --------- --------------
##           1   1029     79.64          79.64     65.83          65.83
##           2    195     15.09          94.74     12.48          78.31
##           3     68      5.26         100.00      4.35          82.66
##        <NA>    271                              17.34         100.00
##       Total   1563    100.00         100.00    100.00         100.00
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{ggplot}\NormalTok{(}\AttributeTok{data =}\NormalTok{ Base\_Trajipaq\_Finale\_Ok, }\AttributeTok{mapping =} \FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ Q40B)) }\SpecialCharTok{+}
  \FunctionTok{geom\_bar}\NormalTok{(}\AttributeTok{fill =} \StringTok{"pink"}\NormalTok{, }\AttributeTok{na.rm =} \ConstantTok{TRUE}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{theme\_bw}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\includegraphics{Article---Sentiment-d-appartenance_files/figure-latex/unnamed-chunk-73-1.pdf}

La distribution est asymétrique vers la droite. 79.64\% des personnes
interrogées disent ne jamais avoir subit de discrimination dans la
recherche de logement au cours des 12 derniers mois précédent la crise
du covid 19.

\hypertarget{q40c.-au-cours-des-12-mois-avant-la-crise-de-la-covid-19-pensez-vous-avoir-subi-des-traitements-inuxe9galitaires-ou-des-discriminations-un-traitement-inuxe9galitaire-ou-une-discrimination-se-repuxe8re-si-vous-avez-uxe9tuxe9-moins-bien-traituxe9-que-les-autres-sans-raison-valable-dans-la-santuxe9}{%
\subsection{Q40C. Au cours des 12 mois (avant la crise de la COVID-19),
pensez-vous avoir subi des traitements inégalitaires ou des
discriminations ? Un traitement inégalitaire ou une discrimination se
repère si « vous avez été moins bien traité que les autres, sans raison
valable » : dans la santé
?}\label{q40c.-au-cours-des-12-mois-avant-la-crise-de-la-covid-19-pensez-vous-avoir-subi-des-traitements-inuxe9galitaires-ou-des-discriminations-un-traitement-inuxe9galitaire-ou-une-discrimination-se-repuxe8re-si-vous-avez-uxe9tuxe9-moins-bien-traituxe9-que-les-autres-sans-raison-valable-dans-la-santuxe9}}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{Base\_Trajipaq\_Finale\_Ok }\OtherTok{\textless{}{-}}
\NormalTok{  Base\_Trajipaq\_Finale\_Ok }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{Q40C =} \FunctionTok{case\_when}\NormalTok{(}
\NormalTok{    Q40C }\SpecialCharTok{==} \DecValTok{1} \SpecialCharTok{\textasciitilde{}} \DecValTok{3}\NormalTok{,}
\NormalTok{    Q40C }\SpecialCharTok{==} \DecValTok{2} \SpecialCharTok{\textasciitilde{}} \DecValTok{2}\NormalTok{,}
\NormalTok{    Q40C }\SpecialCharTok{==} \DecValTok{3} \SpecialCharTok{\textasciitilde{}} \DecValTok{1}\NormalTok{,}
\NormalTok{    Q40C }\SpecialCharTok{==} \DecValTok{96} \SpecialCharTok{\textasciitilde{}} \ConstantTok{NA}\NormalTok{,}
\NormalTok{    Q40C }\SpecialCharTok{==} \DecValTok{98} \SpecialCharTok{\textasciitilde{}} \ConstantTok{NA}\NormalTok{,}
\NormalTok{    Q40C }\SpecialCharTok{==} \DecValTok{99} \SpecialCharTok{\textasciitilde{}} \ConstantTok{NA}
\NormalTok{  ))}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{freq}\NormalTok{(Base\_Trajipaq\_Finale\_Ok}\SpecialCharTok{$}\NormalTok{Q40C)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Frequencies  
## Base_Trajipaq_Finale_Ok$Q40C  
## Type: Numeric  
## 
##               Freq   % Valid   % Valid Cum.   % Total   % Total Cum.
## ----------- ------ --------- -------------- --------- --------------
##           1   1155     82.26          82.26     73.90          73.90
##           2    184     13.11          95.37     11.77          85.67
##           3     65      4.63         100.00      4.16          89.83
##        <NA>    159                              10.17         100.00
##       Total   1563    100.00         100.00    100.00         100.00
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{ggplot}\NormalTok{(}\AttributeTok{data =}\NormalTok{ Base\_Trajipaq\_Finale\_Ok, }\AttributeTok{mapping =} \FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ Q40C)) }\SpecialCharTok{+}
  \FunctionTok{geom\_bar}\NormalTok{(}\AttributeTok{fill =} \StringTok{"pink"}\NormalTok{, }\AttributeTok{na.rm =} \ConstantTok{TRUE}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{theme\_bw}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\includegraphics{Article---Sentiment-d-appartenance_files/figure-latex/unnamed-chunk-76-1.pdf}

La distribution est asymétrique vers la droite. 82.26\% des personnes
interrogées disent ne jamais avoir subit de discrimination dans la santé
au cours des 12 derniers mois précédent la crise du covid 19.

\hypertarget{q40d.-au-cours-des-12-mois-avant-la-crise-de-la-covid-19-pensez-vous-avoir-subi-des-traitements-inuxe9galitaires-ou-des-discriminations-un-traitement-inuxe9galitaire-ou-une-discrimination-se-repuxe8re-si-vous-avez-uxe9tuxe9-moins-bien-traituxe9-que-les-autres-sans-raison-valable-sur-le-lieu-de-formation}{%
\subsection{Q40D. Au cours des 12 mois (avant la crise de la COVID-19),
pensez-vous avoir subi des traitements inégalitaires ou des
discriminations ? Un traitement inégalitaire ou une discrimination se
repère si « vous avez été moins bien traité que les autres, sans raison
valable » : sur le lieu de formation
?}\label{q40d.-au-cours-des-12-mois-avant-la-crise-de-la-covid-19-pensez-vous-avoir-subi-des-traitements-inuxe9galitaires-ou-des-discriminations-un-traitement-inuxe9galitaire-ou-une-discrimination-se-repuxe8re-si-vous-avez-uxe9tuxe9-moins-bien-traituxe9-que-les-autres-sans-raison-valable-sur-le-lieu-de-formation}}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{Base\_Trajipaq\_Finale\_Ok }\OtherTok{\textless{}{-}}
\NormalTok{  Base\_Trajipaq\_Finale\_Ok }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{Q40D =} \FunctionTok{case\_when}\NormalTok{(}
\NormalTok{    Q40D }\SpecialCharTok{==} \DecValTok{1} \SpecialCharTok{\textasciitilde{}} \DecValTok{3}\NormalTok{,}
\NormalTok{    Q40D }\SpecialCharTok{==} \DecValTok{2} \SpecialCharTok{\textasciitilde{}} \DecValTok{2}\NormalTok{,}
\NormalTok{    Q40D }\SpecialCharTok{==} \DecValTok{3} \SpecialCharTok{\textasciitilde{}} \DecValTok{1}\NormalTok{,}
\NormalTok{    Q40D }\SpecialCharTok{==} \DecValTok{96} \SpecialCharTok{\textasciitilde{}} \ConstantTok{NA}\NormalTok{,}
\NormalTok{    Q40D }\SpecialCharTok{==} \DecValTok{98} \SpecialCharTok{\textasciitilde{}} \ConstantTok{NA}\NormalTok{,}
\NormalTok{    Q40D }\SpecialCharTok{==} \DecValTok{99} \SpecialCharTok{\textasciitilde{}} \ConstantTok{NA}
\NormalTok{  ))}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{freq}\NormalTok{(Base\_Trajipaq\_Finale\_Ok}\SpecialCharTok{$}\NormalTok{Q40D)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Frequencies  
## Base_Trajipaq_Finale_Ok$Q40D  
## Type: Numeric  
## 
##               Freq   % Valid   % Valid Cum.   % Total   % Total Cum.
## ----------- ------ --------- -------------- --------- --------------
##           1   1072     81.40          81.40     68.59          68.59
##           2    194     14.73          96.13     12.41          81.00
##           3     51      3.87         100.00      3.26          84.26
##        <NA>    246                              15.74         100.00
##       Total   1563    100.00         100.00    100.00         100.00
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{ggplot}\NormalTok{(}\AttributeTok{data =}\NormalTok{ Base\_Trajipaq\_Finale\_Ok, }\AttributeTok{mapping =} \FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ Q40D)) }\SpecialCharTok{+}
  \FunctionTok{geom\_bar}\NormalTok{(}\AttributeTok{fill =} \StringTok{"pink"}\NormalTok{, }\AttributeTok{na.rm =} \ConstantTok{TRUE}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{theme\_bw}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\includegraphics{Article---Sentiment-d-appartenance_files/figure-latex/unnamed-chunk-79-1.pdf}

La distribution est asymétrique vers la droite. 81.40\% des personnes
interrogées disent ne jamais avoir subit de discrimination sur le lieu
de formation au cours des 12 derniers mois précédent la crise du covid
19.

\hypertarget{i2-lindice-de-discrimination-lors-des-12-derniers-mois-pruxe9cuxe9dent-la-crise-du-covid-19}{%
\subsection{I2 : L'indice de discrimination lors des 12 derniers mois
précédent la crise du covid
19}\label{i2-lindice-de-discrimination-lors-des-12-derniers-mois-pruxe9cuxe9dent-la-crise-du-covid-19}}

A la différence des variables précédentes, l'indice du sentiment
d'appartenance est une variable numérique continue dont le niveau de
mesure est ratio.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{v\_I2 }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}\StringTok{"Q40A"}\NormalTok{,}
          \StringTok{"Q40B"}\NormalTok{,}
          \StringTok{"Q40C"}\NormalTok{,}
          \StringTok{"Q39D"}
\NormalTok{          )}

\NormalTok{Base\_Trajipaq\_Finale\_Ok }\OtherTok{\textless{}{-}}
\NormalTok{  Base\_Trajipaq\_Finale\_Ok }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{rowwise}\NormalTok{() }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{I2 =}\NormalTok{ (}\FunctionTok{sum}\NormalTok{(}\FunctionTok{c\_across}\NormalTok{(}\FunctionTok{all\_of}\NormalTok{(v\_I2)), }\AttributeTok{na.rm =} \ConstantTok{TRUE}\NormalTok{) }\SpecialCharTok{{-}} \FunctionTok{length}\NormalTok{(v\_I2)) }\SpecialCharTok{/}\NormalTok{ (}\FloatTok{2.25}\SpecialCharTok{*}\FunctionTok{length}\NormalTok{(v\_I2))) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{I2 =} \ControlFlowTok{if}\NormalTok{ (I2 }\SpecialCharTok{\textless{}} \DecValTok{0}\NormalTok{) }\ConstantTok{NA\_real\_} \ControlFlowTok{else}\NormalTok{ I2) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{ungroup}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{descr}\NormalTok{(Base\_Trajipaq\_Finale\_Ok}\SpecialCharTok{$}\NormalTok{I2)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Descriptive Statistics  
## Base_Trajipaq_Finale_Ok$I2  
## N: 1563  
## 
##                          I2
## ----------------- ---------
##              Mean      0.27
##           Std.Dev      0.17
##               Min      0.00
##                Q1      0.11
##            Median      0.22
##                Q3      0.33
##               Max      1.00
##               MAD      0.16
##               IQR      0.22
##                CV      0.64
##          Skewness      0.94
##       SE.Skewness      0.06
##          Kurtosis      2.17
##           N.Valid   1440.00
##         Pct.Valid     92.13
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{ggplot}\NormalTok{(}\AttributeTok{data =}\NormalTok{ Base\_Trajipaq\_Finale\_Ok, }\AttributeTok{mapping =} \FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ I2)) }\SpecialCharTok{+}
  \FunctionTok{geom\_histogram}\NormalTok{(}\AttributeTok{na.rm =} \ConstantTok{TRUE}\NormalTok{, }\AttributeTok{fill =} \StringTok{"pink"}\NormalTok{, }\AttributeTok{binwidth =} \FloatTok{0.01}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{theme\_bw}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\includegraphics{Article---Sentiment-d-appartenance_files/figure-latex/unnamed-chunk-82-1.pdf}

La distribution est asymétrique vers la droite. L'indice moyen de
discrimination subie est 0.28.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# Lorsque l\textquotesingle{}on regarde la distribution de chaque question composant l\textquotesingle{}indice, le mode correspond à l\textquotesingle{}absence de discrimination (1). Or, l\textquotesingle{}indice présente une moyenne non nulle (0.28 \textgreater{} 0), suggérant que peu de répondants ne subissent aucune discrimination.}
\end{Highlighting}
\end{Shaded}

\hypertarget{analyse}{%
\section{Analyse}\label{analyse}}

\hypertarget{moduxe8le-de-ruxe9gression}{%
\subsection{Modèle de régression}\label{moduxe8le-de-ruxe9gression}}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{reg\_1 }\OtherTok{\textless{}{-}} \FunctionTok{lm}\NormalTok{(Base\_Trajipaq\_Finale\_Ok}\SpecialCharTok{$}\NormalTok{I2 }\SpecialCharTok{\textasciitilde{}}\NormalTok{ Base\_Trajipaq\_Finale\_Ok}\SpecialCharTok{$}\NormalTok{I1 }\SpecialCharTok{+}
\NormalTok{              Base\_Trajipaq\_Finale\_Ok}\SpecialCharTok{$}\NormalTok{Q10 }\SpecialCharTok{+}
\NormalTok{              Base\_Trajipaq\_Finale\_Ok}\SpecialCharTok{$}\NormalTok{Q8B }\SpecialCharTok{+}
\NormalTok{              Base\_Trajipaq\_Finale\_Ok}\SpecialCharTok{$}\NormalTok{Q4 }\SpecialCharTok{+}
\NormalTok{              Base\_Trajipaq\_Finale\_Ok}\SpecialCharTok{$}\NormalTok{Q16 }\SpecialCharTok{+}
\NormalTok{              Base\_Trajipaq\_Finale\_Ok}\SpecialCharTok{$}\NormalTok{Q17 }\SpecialCharTok{+}
\NormalTok{              Base\_Trajipaq\_Finale\_Ok}\SpecialCharTok{$}\NormalTok{Q20 }\SpecialCharTok{+}
\NormalTok{              Base\_Trajipaq\_Finale\_Ok}\SpecialCharTok{$}\NormalTok{Q23}
\NormalTok{            )}
\FunctionTok{summary}\NormalTok{(reg\_1)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## 
## Call:
## lm(formula = Base_Trajipaq_Finale_Ok$I2 ~ Base_Trajipaq_Finale_Ok$I1 + 
##     Base_Trajipaq_Finale_Ok$Q10 + Base_Trajipaq_Finale_Ok$Q8B + 
##     Base_Trajipaq_Finale_Ok$Q4 + Base_Trajipaq_Finale_Ok$Q16 + 
##     Base_Trajipaq_Finale_Ok$Q17 + Base_Trajipaq_Finale_Ok$Q20 + 
##     Base_Trajipaq_Finale_Ok$Q23)
## 
## Residuals:
##      Min       1Q   Median       3Q      Max 
## -0.26036 -0.10944  0.00757  0.06785  0.62439 
## 
## Coefficients:
##                                   Estimate Std. Error t value Pr(>|t|)    
## (Intercept)                       0.112370   0.290556   0.387    0.701    
## Base_Trajipaq_Finale_Ok$I1        0.665029   0.129454   5.137 6.81e-06 ***
## Base_Trajipaq_Finale_Ok$Q10       0.044990   0.058790   0.765    0.448    
## Base_Trajipaq_Finale_Ok$Q8B      -0.006314   0.015015  -0.421    0.676    
## Base_Trajipaq_Finale_Ok$Q4       -0.073409   0.075565  -0.971    0.337    
## Base_Trajipaq_Finale_Ok$Q16       0.020841   0.021606   0.965    0.340    
## Base_Trajipaq_Finale_Ok$Q17      -0.004538   0.007530  -0.603    0.550    
## Base_Trajipaq_Finale_Ok$Q2001 02  0.021275   0.089699   0.237    0.814    
## Base_Trajipaq_Finale_Ok$Q2001 97  0.042064   0.143345   0.293    0.771    
## Base_Trajipaq_Finale_Ok$Q2002     0.098238   0.090287   1.088    0.283    
## Base_Trajipaq_Finale_Ok$Q2002 97  0.010253   0.132639   0.077    0.939    
## Base_Trajipaq_Finale_Ok$Q2097    -0.017889   0.058807  -0.304    0.762    
## Base_Trajipaq_Finale_Ok$Q2099     0.112975   0.106832   1.057    0.296    
## Base_Trajipaq_Finale_Ok$Q23      -0.052585   0.044384  -1.185    0.243    
## ---
## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
## 
## Residual standard error: 0.1709 on 42 degrees of freedom
##   (1507 observations deleted due to missingness)
## Multiple R-squared:  0.4171, Adjusted R-squared:  0.2367 
## F-statistic: 2.312 on 13 and 42 DF,  p-value: 0.02026
\end{verbatim}

\hypertarget{coefficient-de-corruxe9lation}{%
\subsection{Coefficient de
corrélation}\label{coefficient-de-corruxe9lation}}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{cor.test}\NormalTok{(Base\_Trajipaq\_Finale\_Ok}\SpecialCharTok{$}\NormalTok{I2, Base\_Trajipaq\_Finale\_Ok}\SpecialCharTok{$}\NormalTok{I1)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## 
##  Pearson's product-moment correlation
## 
## data:  Base_Trajipaq_Finale_Ok$I2 and Base_Trajipaq_Finale_Ok$I1
## t = 17.57, df = 1436, p-value < 2.2e-16
## alternative hypothesis: true correlation is not equal to 0
## 95 percent confidence interval:
##  0.3771562 0.4622894
## sample estimates:
##       cor 
## 0.4206484
\end{verbatim}

\hypertarget{visualisation}{%
\subsection{Visualisation}\label{visualisation}}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{vis }\OtherTok{\textless{}{-}} \FunctionTok{ggplot}\NormalTok{(Base\_Trajipaq\_Finale\_Ok }\SpecialCharTok{\%\textgreater{}\%}
                \FunctionTok{filter}\NormalTok{(}\SpecialCharTok{!}\FunctionTok{is.na}\NormalTok{(I1), }\SpecialCharTok{!}\FunctionTok{is.na}\NormalTok{(I2)), }\FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ I2, }\AttributeTok{y =}\NormalTok{ I1)) }\SpecialCharTok{+}
  \FunctionTok{geom\_jitter}\NormalTok{(}\AttributeTok{color =} \StringTok{"pink"}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{stat\_smooth}\NormalTok{(}\AttributeTok{color =} \StringTok{"purple"}\NormalTok{,}\AttributeTok{se =} \ConstantTok{FALSE}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{labs}\NormalTok{(}\AttributeTok{title =} \StringTok{"Discrimination et sentiment d\textquotesingle{}appartenance"}\NormalTok{, }
       \AttributeTok{x =} \StringTok{"Discrimination subie au cours des 12 derniers mois"}\NormalTok{, }
       \AttributeTok{y =} \StringTok{"Sentiment d\textquotesingle{}appartenance au Quebec"}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{theme\_bw}\NormalTok{()}
\FunctionTok{print}\NormalTok{(vis)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## `geom_smooth()` using method = 'gam' and formula = 'y ~ s(x, bs = "cs")'
\end{verbatim}

\includegraphics{Article---Sentiment-d-appartenance_files/figure-latex/unnamed-chunk-86-1.pdf}

\hypertarget{visualisation-contruxf4luxe9e-sexe}{%
\subsection{Visualisation contrôlée :
sexe}\label{visualisation-contruxf4luxe9e-sexe}}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{vis\_sex }\OtherTok{\textless{}{-}} \FunctionTok{ggplot}\NormalTok{(Base\_Trajipaq\_Finale\_Ok }\SpecialCharTok{\%\textgreater{}\%}
                \FunctionTok{filter}\NormalTok{(}\SpecialCharTok{!}\FunctionTok{is.na}\NormalTok{(I1), }\SpecialCharTok{!}\FunctionTok{is.na}\NormalTok{(I2), }\SpecialCharTok{!}\FunctionTok{is.na}\NormalTok{(Q10)), }\FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ I2, }\AttributeTok{y =}\NormalTok{ I1, }\AttributeTok{color =} \FunctionTok{as.factor}\NormalTok{(Q10))) }\SpecialCharTok{+}
  \FunctionTok{geom\_jitter}\NormalTok{(}\AttributeTok{color =} \StringTok{"pink"}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{stat\_smooth}\NormalTok{(}\AttributeTok{se =} \ConstantTok{FALSE}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{scale\_color\_brewer}\NormalTok{(}\AttributeTok{name =} \StringTok{"Sexe"}\NormalTok{, }\AttributeTok{palette =} \DecValTok{3}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{labs}\NormalTok{(}\AttributeTok{title =} \StringTok{"Discrimination et sentiment d\textquotesingle{}appartenance"}\NormalTok{, }
       \AttributeTok{x =} \StringTok{"Discrimination subie au cours des 12 derniers mois"}\NormalTok{, }
       \AttributeTok{y =} \StringTok{"Sentiment d\textquotesingle{}appartenance au Quebec"}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{theme\_bw}\NormalTok{()}
\FunctionTok{print}\NormalTok{(vis\_sex)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## `geom_smooth()` using method = 'loess' and formula = 'y ~ x'
\end{verbatim}

\includegraphics{Article---Sentiment-d-appartenance_files/figure-latex/unnamed-chunk-87-1.pdf}

\hypertarget{visualisation-contruxf4luxe9e-uxe2ge}{%
\subsection{Visualisation contrôlée :
âge}\label{visualisation-contruxf4luxe9e-uxe2ge}}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{Base\_Trajipaq\_Finale\_Ok }\OtherTok{\textless{}{-}}
\NormalTok{  Base\_Trajipaq\_Finale\_Ok }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{Q8A\_F =} \FunctionTok{case\_when}\NormalTok{(}
\NormalTok{    Q8A }\SpecialCharTok{\textless{}=} \DecValTok{1963} \SpecialCharTok{\textasciitilde{}} \StringTok{"\textgreater{}=60"}\NormalTok{,}
\NormalTok{    Q8A }\SpecialCharTok{\textgreater{}=} \DecValTok{1963} \SpecialCharTok{\&}\NormalTok{ Q8A }\SpecialCharTok{\textless{}} \DecValTok{1968} \SpecialCharTok{\textasciitilde{}} \StringTok{"[59;55]"}\NormalTok{,}
\NormalTok{    Q8A }\SpecialCharTok{\textgreater{}=} \DecValTok{1968} \SpecialCharTok{\&}\NormalTok{ Q8A }\SpecialCharTok{\textless{}} \DecValTok{1973} \SpecialCharTok{\textasciitilde{}} \StringTok{"[54;50]"}\NormalTok{,}
\NormalTok{    Q8A }\SpecialCharTok{\textgreater{}=} \DecValTok{1973} \SpecialCharTok{\&}\NormalTok{ Q8A }\SpecialCharTok{\textless{}} \DecValTok{1978} \SpecialCharTok{\textasciitilde{}} \StringTok{"[49;45]"}\NormalTok{,}
\NormalTok{    Q8A }\SpecialCharTok{\textgreater{}=} \DecValTok{1978} \SpecialCharTok{\&}\NormalTok{ Q8A }\SpecialCharTok{\textless{}} \DecValTok{1983} \SpecialCharTok{\textasciitilde{}} \StringTok{"[44;35]"}\NormalTok{,}
\NormalTok{    Q8A }\SpecialCharTok{\textgreater{}=} \DecValTok{1983} \SpecialCharTok{\&}\NormalTok{ Q8A }\SpecialCharTok{\textless{}} \DecValTok{1988} \SpecialCharTok{\textasciitilde{}} \StringTok{"[34;30]"}\NormalTok{,}
\NormalTok{    Q8A }\SpecialCharTok{\textgreater{}=} \DecValTok{1988} \SpecialCharTok{\&}\NormalTok{ Q8A }\SpecialCharTok{\textless{}} \DecValTok{1993} \SpecialCharTok{\textasciitilde{}} \StringTok{"[29;25]"}\NormalTok{,}
\NormalTok{    Q8A }\SpecialCharTok{\textgreater{}=} \DecValTok{1993} \SpecialCharTok{\&}\NormalTok{ Q8A }\SpecialCharTok{\textless{}} \DecValTok{1998} \SpecialCharTok{\textasciitilde{}} \StringTok{"[24;20]"}\NormalTok{,}
\NormalTok{    Q8A }\SpecialCharTok{\textgreater{}=} \DecValTok{1998} \SpecialCharTok{\&}\NormalTok{ Q8A }\SpecialCharTok{\textless{}} \DecValTok{2023} \SpecialCharTok{\textasciitilde{}} \StringTok{"\textless{}19"}
\NormalTok{  ))}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{vis\_age }\OtherTok{\textless{}{-}} \FunctionTok{ggplot}\NormalTok{(Base\_Trajipaq\_Finale\_Ok }\SpecialCharTok{\%\textgreater{}\%}
                \FunctionTok{filter}\NormalTok{(}\SpecialCharTok{!}\FunctionTok{is.na}\NormalTok{(I1), }\SpecialCharTok{!}\FunctionTok{is.na}\NormalTok{(I2), }\SpecialCharTok{!}\FunctionTok{is.na}\NormalTok{(Q8A\_F)), }\FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ I2, }\AttributeTok{y =}\NormalTok{ I1, }\AttributeTok{color =}\NormalTok{ Q8A\_F)) }\SpecialCharTok{+}
  \FunctionTok{geom\_jitter}\NormalTok{(}\AttributeTok{color =} \StringTok{"pink"}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{stat\_smooth}\NormalTok{(}\AttributeTok{se =} \ConstantTok{FALSE}\NormalTok{, }\AttributeTok{span =} \FloatTok{0.9}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{scale\_color\_brewer}\NormalTok{(}\AttributeTok{name =} \StringTok{"Age"}\NormalTok{, }\AttributeTok{palette =} \DecValTok{3}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{labs}\NormalTok{(}\AttributeTok{title =} \StringTok{"Discrimination et sentiment d\textquotesingle{}appartenance"}\NormalTok{, }
       \AttributeTok{x =} \StringTok{"Discrimination subie au cours des 12 derniers mois"}\NormalTok{, }
       \AttributeTok{y =} \StringTok{"Sentiment d\textquotesingle{}appartenance au Quebec"}\NormalTok{,}
\NormalTok{       ) }\SpecialCharTok{+}
  \FunctionTok{theme\_bw}\NormalTok{()}
\FunctionTok{print}\NormalTok{(vis\_age)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## `geom_smooth()` using method = 'loess' and formula = 'y ~ x'
\end{verbatim}

\includegraphics{Article---Sentiment-d-appartenance_files/figure-latex/unnamed-chunk-89-1.pdf}

\hypertarget{visualisation-contruxf4luxe9e-statut-matrimonial}{%
\subsection{Visualisation contrôlée : statut
matrimonial}\label{visualisation-contruxf4luxe9e-statut-matrimonial}}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{Base\_Trajipaq\_Finale\_Ok }\OtherTok{\textless{}{-}}
\NormalTok{  Base\_Trajipaq\_Finale\_Ok }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{STATUT\_F =} \FunctionTok{case\_when}\NormalTok{(}
\NormalTok{    STATUT }\SpecialCharTok{==} \DecValTok{1} \SpecialCharTok{\textasciitilde{}} \StringTok{"Celibataire"}\NormalTok{,}
\NormalTok{    STATUT }\SpecialCharTok{==} \DecValTok{2} \SpecialCharTok{\textasciitilde{}} \StringTok{"Marie(e)"}\NormalTok{,}
\NormalTok{    STATUT }\SpecialCharTok{==} \DecValTok{3} \SpecialCharTok{\textasciitilde{}} \StringTok{"Conjoint(e) de fait"}\NormalTok{,}
\NormalTok{    STATUT }\SpecialCharTok{==} \DecValTok{4} \SpecialCharTok{\textasciitilde{}} \StringTok{"Separe(e)"}\NormalTok{,}
\NormalTok{    STATUT }\SpecialCharTok{==} \DecValTok{5} \SpecialCharTok{\textasciitilde{}} \StringTok{"Divorce(e)"}\NormalTok{,}
\NormalTok{    STATUT }\SpecialCharTok{==} \DecValTok{6} \SpecialCharTok{\textasciitilde{}} \ConstantTok{NA}
\NormalTok{  ))}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{vis\_born }\OtherTok{\textless{}{-}} \FunctionTok{ggplot}\NormalTok{(Base\_Trajipaq\_Finale\_Ok }\SpecialCharTok{\%\textgreater{}\%}
                \FunctionTok{filter}\NormalTok{(}\SpecialCharTok{!}\FunctionTok{is.na}\NormalTok{(I1), }\SpecialCharTok{!}\FunctionTok{is.na}\NormalTok{(I2), }\SpecialCharTok{!}\FunctionTok{is.na}\NormalTok{(STATUT\_F)), }\FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ I2, }\AttributeTok{y =}\NormalTok{ I1, }\AttributeTok{color =}\NormalTok{ STATUT\_F)) }\SpecialCharTok{+}
  \FunctionTok{geom\_jitter}\NormalTok{(}\AttributeTok{color =} \StringTok{"pink"}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{stat\_smooth}\NormalTok{(}\AttributeTok{se =} \ConstantTok{FALSE}\NormalTok{, }\AttributeTok{span =} \FloatTok{0.9}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{scale\_color\_brewer}\NormalTok{(}\AttributeTok{name =} \StringTok{"Statut matrimonial"}\NormalTok{, }\AttributeTok{palette =} \DecValTok{3}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{labs}\NormalTok{(}\AttributeTok{title =} \StringTok{"Discrimination et sentiment d\textquotesingle{}appartenance"}\NormalTok{, }
       \AttributeTok{x =} \StringTok{"Discrimination subie au cours des 12 derniers mois"}\NormalTok{, }
       \AttributeTok{y =} \StringTok{"Sentiment d\textquotesingle{}appartenance au Quebec"}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{theme\_bw}\NormalTok{()}
\FunctionTok{print}\NormalTok{(vis\_born)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## `geom_smooth()` using method = 'loess' and formula = 'y ~ x'
\end{verbatim}

\includegraphics{Article---Sentiment-d-appartenance_files/figure-latex/unnamed-chunk-91-1.pdf}

\hypertarget{visualisation-contruxf4luxe9e-statut-migratoire}{%
\subsection{Visualisation contrôlée : statut
migratoire}\label{visualisation-contruxf4luxe9e-statut-migratoire}}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{vis\_born }\OtherTok{\textless{}{-}} \FunctionTok{ggplot}\NormalTok{(Base\_Trajipaq\_Finale\_Ok }\SpecialCharTok{\%\textgreater{}\%}
                \FunctionTok{filter}\NormalTok{(}\SpecialCharTok{!}\FunctionTok{is.na}\NormalTok{(I1), }\SpecialCharTok{!}\FunctionTok{is.na}\NormalTok{(I2), }\SpecialCharTok{!}\FunctionTok{is.na}\NormalTok{(Q4)), }\FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ I2, }\AttributeTok{y =}\NormalTok{ I1, }\AttributeTok{color =} \FunctionTok{as.factor}\NormalTok{(Q4))) }\SpecialCharTok{+}
  \FunctionTok{geom\_jitter}\NormalTok{(}\AttributeTok{color =} \StringTok{"pink"}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{stat\_smooth}\NormalTok{(}\AttributeTok{se =} \ConstantTok{FALSE}\NormalTok{, }\AttributeTok{span =} \FloatTok{0.9}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{scale\_color\_brewer}\NormalTok{(}\AttributeTok{name =} \StringTok{"Ne au Canada"}\NormalTok{, }\AttributeTok{palette =} \DecValTok{3}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{labs}\NormalTok{(}\AttributeTok{title =} \StringTok{"Discrimination et sentiment d\textquotesingle{}appartenance"}\NormalTok{, }
       \AttributeTok{x =} \StringTok{"Discrimination subie au cours des 12 derniers mois"}\NormalTok{, }
       \AttributeTok{y =} \StringTok{"Sentiment d\textquotesingle{}appartenance au Quebec"}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{theme\_bw}\NormalTok{()}
\FunctionTok{print}\NormalTok{(vis\_born)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## `geom_smooth()` using method = 'loess' and formula = 'y ~ x'
\end{verbatim}

\includegraphics{Article---Sentiment-d-appartenance_files/figure-latex/unnamed-chunk-92-1.pdf}

\hypertarget{visualisation-contruxf4luxe9e-niveau-duxe9ducation}{%
\subsection{Visualisation contrôlée : niveau
d'éducation}\label{visualisation-contruxf4luxe9e-niveau-duxe9ducation}}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{vis\_educ }\OtherTok{\textless{}{-}} \FunctionTok{ggplot}\NormalTok{(Base\_Trajipaq\_Finale\_Ok }\SpecialCharTok{\%\textgreater{}\%}
                \FunctionTok{filter}\NormalTok{(}\SpecialCharTok{!}\FunctionTok{is.na}\NormalTok{(I1), }\SpecialCharTok{!}\FunctionTok{is.na}\NormalTok{(I2), }\SpecialCharTok{!}\FunctionTok{is.na}\NormalTok{(Q16)), }\FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ I2, }\AttributeTok{y =}\NormalTok{ I1, }\AttributeTok{color =} \FunctionTok{as.factor}\NormalTok{(Q16))) }\SpecialCharTok{+}
  \FunctionTok{geom\_jitter}\NormalTok{(}\AttributeTok{color =} \StringTok{"pink"}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{stat\_smooth}\NormalTok{(}\AttributeTok{se =} \ConstantTok{FALSE}\NormalTok{, }\AttributeTok{span =} \DecValTok{1}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{scale\_color\_brewer}\NormalTok{(}\AttributeTok{name =} \StringTok{"Niveau d\textquotesingle{}education"}\NormalTok{, }\AttributeTok{palette =} \DecValTok{3}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{labs}\NormalTok{(}\AttributeTok{title =} \StringTok{"Discrimination et sentiment d\textquotesingle{}appartenance"}\NormalTok{, }
       \AttributeTok{x =} \StringTok{"Discrimination subie au cours des 12 derniers mois"}\NormalTok{, }
       \AttributeTok{y =} \StringTok{"Sentiment d\textquotesingle{}appartenance au Quebec"}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{theme\_bw}\NormalTok{()}
\FunctionTok{print}\NormalTok{(vis\_educ)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## `geom_smooth()` using method = 'loess' and formula = 'y ~ x'
\end{verbatim}

\includegraphics{Article---Sentiment-d-appartenance_files/figure-latex/unnamed-chunk-93-1.pdf}

\hypertarget{visualisation-contruxf4luxe9e-religion}{%
\subsection{Visualisation contrôlée :
religion}\label{visualisation-contruxf4luxe9e-religion}}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{Base\_Trajipaq\_Finale\_Ok }\OtherTok{\textless{}{-}}
\NormalTok{  Base\_Trajipaq\_Finale\_Ok }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{Q17\_F =} \FunctionTok{case\_when}\NormalTok{(}
\NormalTok{    Q17 }\SpecialCharTok{==} \DecValTok{0} \SpecialCharTok{\textasciitilde{}} \StringTok{"Sans religion"}\NormalTok{,}
\NormalTok{    Q17 }\SpecialCharTok{==} \DecValTok{1} \SpecialCharTok{\textasciitilde{}} \StringTok{"Catholique"}\NormalTok{,}
\NormalTok{    Q17 }\SpecialCharTok{==} \DecValTok{8} \SpecialCharTok{\textasciitilde{}} \StringTok{"Musulman"}\NormalTok{,}
\NormalTok{    Q17 }\SpecialCharTok{!=} \DecValTok{0} \SpecialCharTok{\&}\NormalTok{ Q17 }\SpecialCharTok{!=} \DecValTok{1} \SpecialCharTok{\&}\NormalTok{ Q17 }\SpecialCharTok{!=} \DecValTok{8} \SpecialCharTok{\textasciitilde{}} \StringTok{"Autre"}
\NormalTok{  ))}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{vis\_relig }\OtherTok{\textless{}{-}} \FunctionTok{ggplot}\NormalTok{(Base\_Trajipaq\_Finale\_Ok }\SpecialCharTok{\%\textgreater{}\%}
                \FunctionTok{filter}\NormalTok{(}\SpecialCharTok{!}\FunctionTok{is.na}\NormalTok{(I1), }\SpecialCharTok{!}\FunctionTok{is.na}\NormalTok{(I2), }\SpecialCharTok{!}\FunctionTok{is.na}\NormalTok{(Q17\_F)), }\FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ I2, }\AttributeTok{y =}\NormalTok{ I1, }\AttributeTok{color =}\NormalTok{ Q17\_F)) }\SpecialCharTok{+}
  \FunctionTok{geom\_jitter}\NormalTok{(}\AttributeTok{color =} \StringTok{"pink"}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{stat\_smooth}\NormalTok{(}\AttributeTok{se =} \ConstantTok{FALSE}\NormalTok{, }\AttributeTok{span =} \FloatTok{0.9}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{scale\_color\_brewer}\NormalTok{(}\AttributeTok{name =} \StringTok{"Religion"}\NormalTok{, }\AttributeTok{palette =} \DecValTok{3}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{labs}\NormalTok{(}\AttributeTok{title =} \StringTok{"Discrimination et sentiment d\textquotesingle{}appartenance"}\NormalTok{, }
       \AttributeTok{x =} \StringTok{"Discrimination subie au cours des 12 derniers mois"}\NormalTok{, }
       \AttributeTok{y =} \StringTok{"Sentiment d\textquotesingle{}appartenance au Quebec"}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{theme\_bw}\NormalTok{()}
\FunctionTok{print}\NormalTok{(vis\_relig)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## `geom_smooth()` using method = 'loess' and formula = 'y ~ x'
\end{verbatim}

\includegraphics{Article---Sentiment-d-appartenance_files/figure-latex/unnamed-chunk-95-1.pdf}

\hypertarget{visualisation-contruxf4luxe9e-langue-maternelle}{%
\subsection{Visualisation contrôlée : langue
maternelle}\label{visualisation-contruxf4luxe9e-langue-maternelle}}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{Base\_Trajipaq\_Finale\_Ok }\OtherTok{\textless{}{-}}
\NormalTok{  Base\_Trajipaq\_Finale\_Ok }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{Q20\_F =} \FunctionTok{case\_when}\NormalTok{(}
\NormalTok{    Q20M1 }\SpecialCharTok{==} \DecValTok{1} \SpecialCharTok{\textasciitilde{}} \StringTok{"Francais"}\NormalTok{,}
\NormalTok{    Q20M1 }\SpecialCharTok{==} \DecValTok{2} \SpecialCharTok{\textasciitilde{}} \StringTok{"Anglais"}\NormalTok{,}
\NormalTok{    Q20M1 }\SpecialCharTok{==} \DecValTok{3} \SpecialCharTok{\textasciitilde{}} \StringTok{"Autre"}
\NormalTok{  ))}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{vis\_lang }\OtherTok{\textless{}{-}} \FunctionTok{ggplot}\NormalTok{(Base\_Trajipaq\_Finale\_Ok }\SpecialCharTok{\%\textgreater{}\%}
                \FunctionTok{filter}\NormalTok{(}\SpecialCharTok{!}\FunctionTok{is.na}\NormalTok{(I1), }\SpecialCharTok{!}\FunctionTok{is.na}\NormalTok{(I2), }\SpecialCharTok{!}\FunctionTok{is.na}\NormalTok{(Q20\_F)), }\FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ I2, }\AttributeTok{y =}\NormalTok{ I1, }\AttributeTok{color =}\NormalTok{ Q20\_F)) }\SpecialCharTok{+}
  \FunctionTok{geom\_jitter}\NormalTok{(}\AttributeTok{color =} \StringTok{"pink"}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{stat\_smooth}\NormalTok{(}\AttributeTok{se =} \ConstantTok{FALSE}\NormalTok{, }\AttributeTok{span =} \FloatTok{0.8}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{scale\_color\_brewer}\NormalTok{(}\AttributeTok{name =} \StringTok{"Langue maternelle"}\NormalTok{, }\AttributeTok{palette =} \DecValTok{3}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{labs}\NormalTok{(}\AttributeTok{title =} \StringTok{"Discrimination et sentiment d\textquotesingle{}appartenance"}\NormalTok{, }
       \AttributeTok{x =} \StringTok{"Discrimination subie au cours des 12 derniers mois"}\NormalTok{, }
       \AttributeTok{y =} \StringTok{"Sentiment d\textquotesingle{}appartenance au Quebec"}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{theme\_bw}\NormalTok{()}
\FunctionTok{print}\NormalTok{(vis\_lang)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## `geom_smooth()` using method = 'loess' and formula = 'y ~ x'
\end{verbatim}

\includegraphics{Article---Sentiment-d-appartenance_files/figure-latex/unnamed-chunk-97-1.pdf}

\hypertarget{visualisation-contruxf4luxe9e-uxe9tat-de-santuxe9}{%
\subsection{Visualisation contrôlée : état de
santé}\label{visualisation-contruxf4luxe9e-uxe9tat-de-santuxe9}}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{vis\_health }\OtherTok{\textless{}{-}} \FunctionTok{ggplot}\NormalTok{(Base\_Trajipaq\_Finale\_Ok }\SpecialCharTok{\%\textgreater{}\%}
                \FunctionTok{filter}\NormalTok{(}\SpecialCharTok{!}\FunctionTok{is.na}\NormalTok{(I1), }\SpecialCharTok{!}\FunctionTok{is.na}\NormalTok{(I2), }\SpecialCharTok{!}\FunctionTok{is.na}\NormalTok{(Q23)), }\FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ I2, }\AttributeTok{y =}\NormalTok{ I1, }\AttributeTok{color =} \FunctionTok{as.factor}\NormalTok{(Q23))) }\SpecialCharTok{+}
  \FunctionTok{geom\_jitter}\NormalTok{(}\AttributeTok{color =} \StringTok{"pink"}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{stat\_smooth}\NormalTok{(}\AttributeTok{se =} \ConstantTok{FALSE}\NormalTok{, }\AttributeTok{span =} \FloatTok{1.1}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{scale\_color\_brewer}\NormalTok{(}\AttributeTok{name =} \StringTok{"Etat de sante"}\NormalTok{, }\AttributeTok{palette =} \DecValTok{3}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{labs}\NormalTok{(}\AttributeTok{title =} \StringTok{"Discrimination et sentiment d\textquotesingle{}appartenance"}\NormalTok{, }
       \AttributeTok{x =} \StringTok{"Discrimination subie au cours des 12 derniers mois"}\NormalTok{, }
       \AttributeTok{y =} \StringTok{"Sentiment d\textquotesingle{}appartenance au Quebec"}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{theme\_bw}\NormalTok{()}
\FunctionTok{print}\NormalTok{(vis\_health)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## `geom_smooth()` using method = 'loess' and formula = 'y ~ x'
\end{verbatim}

\includegraphics{Article---Sentiment-d-appartenance_files/figure-latex/unnamed-chunk-98-1.pdf}

\end{document}

### 2023-05-22 Codes/Article - Sentiment d'appartenance.Rmd
Visualisation HTML.

@@
 date: "2023-05-17"
-output: pdf_document
+output:
+  html_document:
+    df_print: paged
 ---
@@
 ```{r}
-Base_Trajipaq_Finale_Ok <- read_dta("~/Desktop/Base_Trajipaq_Finale_Ok.dta")
+Base_Trajipaq_Finale_Ok <- read_dta("C:/Users/Théo CARLIER/OneDrive/Bureau/Base_Trajipaq_Finale_Ok.dta")
 ```

### 2023-05-22 Codes/Article---Sentiment-d-appartenance.tex
Visualisation HTML.

@@
 \FunctionTok{library}\NormalTok{(summarytools)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+## Warning: le package 'summarytools' a été compilé avec la version R 4.2.3
+\end{verbatim}
+
+\begin{Shaded}
+\begin{Highlighting}[]
 \FunctionTok{library}\NormalTok{(tidyverse)}
@@
 \end{Shaded}
+
+\begin{verbatim}
+## Warning: le package 'tidyverse' a été compilé avec la version R 4.2.3
+\end{verbatim}
+
+\begin{verbatim}
+## Warning: le package 'ggplot2' a été compilé avec la version R 4.2.3
+\end{verbatim}
+
+\begin{verbatim}
+## Warning: le package 'tibble' a été compilé avec la version R 4.2.3
+\end{verbatim}
+
+\begin{verbatim}
+## Warning: le package 'tidyr' a été compilé avec la version R 4.2.3
+\end{verbatim}
+
+\begin{verbatim}
+## Warning: le package 'readr' a été compilé avec la version R 4.2.3
+\end{verbatim}
+
+\begin{verbatim}
+## Warning: le package 'dplyr' a été compilé avec la version R 4.2.3
+\end{verbatim}
+
+\begin{verbatim}
+## Warning: le package 'forcats' a été compilé avec la version R 4.2.3
+\end{verbatim}
+
+\begin{verbatim}
+## Warning: le package 'lubridate' a été compilé avec la version R 4.2.3
+\end{verbatim}
 
@@
 ## x tibble::view()  masks summarytools::view()
-## i Use the conflicted package (<http://conflicted.r-lib.org/>) to force all conflicts to become errors
+## i Use the ]8;;http://conflicted.r-lib.org/conflicted package]8;; to force all conflicts to become errors
 \end{verbatim}
@@
 \FunctionTok{library}\NormalTok{(ggthemes)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+## Warning: le package 'ggthemes' a été compilé avec la version R 4.2.3
+\end{verbatim}
+
+\begin{Shaded}
+\begin{Highlighting}[]
 \FunctionTok{library}\NormalTok{(haven)}
@@
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{Base\_Trajipaq\_Finale\_Ok }\OtherTok{\textless{}{-}} \FunctionTok{read\_dta}\NormalTok{(}\StringTok{"\textasciitilde{}/Desktop/Base\_Trajipaq\_Finale\_Ok.dta"}\NormalTok{)}
+\begin{verbatim}
+## Warning: le package 'haven' a été compilé avec la version R 4.2.3
+\end{verbatim}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{Base\_Trajipaq\_Finale\_Ok }\OtherTok{\textless{}{-}} \FunctionTok{read\_dta}\NormalTok{(}\StringTok{"C:/Users/Théo CARLIER/OneDrive/Bureau/Base\_Trajipaq\_Finale\_Ok.dta"}\NormalTok{)}
 \end{Highlighting}
@@
 ## Base_Trajipaq_Finale_Ok$QAGE  
-## Label: Vous appartenez <U+00E0> quel groupe d<U+2019><U+00E2>ge?  
+## Label: Vous appartenez à quel groupe d’âge?  
 ## Type: Numeric  
@@
 ## Base_Trajipaq_Finale_Ok$STATUT  
-## Label: <U+00CA>tes-vous actuellement...?  
+## Label: Êtes-vous actuellement...?  
 ## Type: Numeric  
@@
 ## Base_Trajipaq_Finale_Ok$Q11  
-## Label: Dans quel pays <U+00EA>tes-vous n<U+00E9>(e) ?Si vous ne vous en souvenez pas, inscrire : <U+00AB> Je  
+## Label: Dans quel pays êtes-vous né(e) ?Si vous ne vous en souvenez pas, inscrire : « Je  
 ## Type: Numeric  
@@
 ## Base_Trajipaq_Finale_Ok$Q4  
-## Label: <U+00CA>tes-vous n<U+00E9>(e) au Canada ?  
+## Label: Êtes-vous né(e) au Canada ?  
 ## Type: Numeric  
@@
 ## Residual standard error: 0.1709 on 42 degrees of freedom
-##   (1507 observations deleted due to missingness)
+##   (1507 observations effacées parce que manquantes)
 ## Multiple R-squared:  0.4171, Adjusted R-squared:  0.2367 

### 2023-05-23 Codes/Article - Sentiment d'appartenance.Rmd
Progression dans l'analyse.

--- 2023-05-22 Codes/Article - Sentiment d'appartenance.Rmd
@@
 
+## Tableau croisé
+
+```{r}
+
+Base_Trajipaq_Finale_Ok <-
+  Base_Trajipaq_Finale_Ok %>%
+  mutate(I1_F = case_when(
+    I1 < 0.25 ~ "Faible",
+    I1 >= 0.25 & I1 < 0.5 ~ "Moyen",
+    I1 >= 0.5 & I1 < 0.75 ~ "Fort",
+    I1 >= 0.75 & I1 <= 1 ~ "Tres fort",
+  ))
+
+Base_Trajipaq_Finale_Ok$I1_F <- factor(Base_Trajipaq_Finale_Ok$I1_F, levels = c("Tres fort",
+                                                                                "Fort",
+                                                                                "Moyen",
+                                                                                "Faible"
+                                                                                ), ordered = TRUE)
+
+Base_Trajipaq_Finale_Ok <-
+  Base_Trajipaq_Finale_Ok %>%
+  mutate(I2_F = case_when(
+    I2 < 0.25 ~ "Faible",
+    I2 >= 0.25 & I2 < 0.5 ~ "Moyen",
+    I2 >= 0.5 & I2 < 0.75 ~ "Fort",
+    I2 >= 0.75 & I2 <= 1 ~ "Tres fort",
+  ))
+
+Base_Trajipaq_Finale_Ok$I2_F <- factor(Base_Trajipaq_Finale_Ok$I2_F, levels = c("Tres fort",
+                                                                                "Fort",
+                                                                                "Moyen",
+                                                                                "Faible"
+                                                                                ), ordered = TRUE)
+
+```
+
+```{r}
+
+ctable(x = Base_Trajipaq_Finale_Ok$I2_F, y = Base_Trajipaq_Finale_Ok$I1_F)
+
+```
+
+```{r}
+
+ggplot(data = Base_Trajipaq_Finale_Ok, mapping = aes(x = I2_F, fill = I1_F)) +
+  geom_bar(position = "dodge", na.rm = TRUE) +
+  theme_bw()
+
+```
+
+## Test d'association
+
+```{r}
+
+chisq.test(x = Base_Trajipaq_Finale_Ok$I2_F, y = Base_Trajipaq_Finale_Ok$I1_F)
+
+```
+
 ## Modèle de régression
@@
   geom_jitter(color = "pink") +
-  stat_smooth(color = "purple",se = FALSE) +
+  stat_smooth(method = "lm", formula = "y ~ x", color = "purple",se = FALSE) +
   labs(title = "Discrimination et sentiment d'appartenance", 
@@
   geom_jitter(color = "pink") +
-  stat_smooth(se = FALSE) +
+  stat_smooth(method = "lm", formula = "y ~ x", se = FALSE) +
   scale_color_brewer(name = "Sexe", palette = 3) +
@@
   geom_jitter(color = "pink") +
-  stat_smooth(se = FALSE, span = 0.9) +
+  stat_smooth(method = "lm", formula = "y ~ x", se = FALSE) +
   scale_color_brewer(name = "Age", palette = 3) +
@@
     STATUT == 5 ~ "Divorce(e)",
-    STATUT == 6 ~ NA
+    STATUT == 6 ~ "Veuf(ve)"
   ))
@@
   geom_jitter(color = "pink") +
-  stat_smooth(se = FALSE, span = 0.9) +
+  stat_smooth(method = "lm", formula = "y ~ x", se = FALSE) +
   scale_color_brewer(name = "Statut matrimonial", palette = 3) +
@@
   geom_jitter(color = "pink") +
-  stat_smooth(se = FALSE, span = 0.9) +
+  stat_smooth(method = "lm", formula = "y ~ x", se = FALSE) +
   scale_color_brewer(name = "Ne au Canada", palette = 3) +
@@
   geom_jitter(color = "pink") +
-  stat_smooth(se = FALSE, span = 1) +
+  stat_smooth(method = "lm", formula = "y ~ x", se = FALSE) +
   scale_color_brewer(name = "Niveau d'education", palette = 3) +
@@
   geom_jitter(color = "pink") +
-  stat_smooth(se = FALSE, span = 0.9) +
+  stat_smooth(method = "lm", formula = "y ~ x", se = FALSE) +
   scale_color_brewer(name = "Religion", palette = 3) +
@@
   geom_jitter(color = "pink") +
-  stat_smooth(se = FALSE, span = 0.8) +
+  stat_smooth(method = "lm", formula = "y ~ x", se = FALSE) +
   scale_color_brewer(name = "Langue maternelle", palette = 3) +
@@
   geom_jitter(color = "pink") +
-  stat_smooth(se = FALSE, span = 1.1) +
+  stat_smooth(method = "lm", formula = "y ~ x", se = FALSE) +
   scale_color_brewer(name = "Etat de sante", palette = 3) +

### 2023-05-23 Codes/Article - Sentiment d'appartenance.Rmd
Arrangement du document de travail.

--- 2023-05-22 Codes/Article - Sentiment d'appartenance.Rmd
@@
 ```{r}
-Base_Trajipaq_Finale_Ok <- read_dta("C:/Users/Théo CARLIER/OneDrive/Bureau/Base_Trajipaq_Finale_Ok.dta")
+Base_Trajipaq_Finale_Ok <- read_dta("~/Desktop/Base_Trajipaq_Finale_Ok.dta")
 ```
@@
 
-# Présentation de la variable indépendante
+# VD1 : sentiment d'appartenance au Québec
 
@@
 
-## I1 : L'indice du sentiment d'appartenance au Québec
+## VD1 : L'indice du sentiment d'appartenance au Québec
 
@@
 
-v_I1 <- c("Q39A",
+v_VD1 <- c("Q39A",
           "Q39B",
@@
   rowwise() %>%
-  mutate(I1 = (sum(c_across(all_of(v_I1)), na.rm = TRUE) - length(v_I1)) / (3*length(v_I1))) %>%
-  mutate(I1 = if (I1 < 0) NA_real_ else I1) %>%
+  mutate(VD1 = (sum(c_across(all_of(v_VD1)), na.rm = TRUE) - length(v_VD1)) / (3*length(v_VD1))) %>%
+  mutate(VD1 = if (VD1 < 0) NA_real_ else VD1) %>%
   ungroup()
@@
 
-descr(Base_Trajipaq_Finale_Ok$I1)
-
-```
-
-```{r}
-
-ggplot(data = Base_Trajipaq_Finale_Ok, mapping = aes(x = I1)) +
+descr(Base_Trajipaq_Finale_Ok$VD1)
+
+```
+
+```{r}
+
+ggplot(data = Base_Trajipaq_Finale_Ok, mapping = aes(x = VD1)) +
   geom_histogram(na.rm = TRUE, fill = "pink", binwidth = 0.01) +
@@
 
-# Présentation de la variable dépendante
+# VI1 : discrimination subie au cours des 12 derniers mois
 
@@
 
-## I2 : L'indice de discrimination lors des 12 derniers mois précédent la crise du covid 19
+## VI1 : L'indice de discrimination lors des 12 derniers mois précédent la crise du covid 19
 
@@
 
-v_I2 <- c("Q40A",
+v_VI1 <- c("Q40A",
           "Q40B",
@@
   rowwise() %>%
-  mutate(I2 = (sum(c_across(all_of(v_I2)), na.rm = TRUE) - length(v_I2)) / (2.25*length(v_I2))) %>%
-  mutate(I2 = if (I2 < 0) NA_real_ else I2) %>%
+  mutate(VI1 = (sum(c_across(all_of(v_VI1)), na.rm = TRUE) - length(v_VI1)) / (2.25*length(v_VI1))) %>%
+  mutate(VI1 = if (VI1 < 0) NA_real_ else VI1) %>%
   ungroup()
@@
 
-descr(Base_Trajipaq_Finale_Ok$I2)
-
-```
-
-```{r}
-
-ggplot(data = Base_Trajipaq_Finale_Ok, mapping = aes(x = I2)) +
+descr(Base_Trajipaq_Finale_Ok$VI1)
+
+```
+
+```{r}
+
+ggplot(data = Base_Trajipaq_Finale_Ok, mapping = aes(x = VI1)) +
   geom_histogram(na.rm = TRUE, fill = "pink", binwidth = 0.01) +
@@
 
-## Modèle de régression
-
-```{r}
-
-reg_1 <- lm(Base_Trajipaq_Finale_Ok$I2 ~ Base_Trajipaq_Finale_Ok$I1 +
-              Base_Trajipaq_Finale_Ok$Q10 +
-              Base_Trajipaq_Finale_Ok$Q8B +
-              Base_Trajipaq_Finale_Ok$Q4 +
-              Base_Trajipaq_Finale_Ok$Q16 +
-              Base_Trajipaq_Finale_Ok$Q17 +
-              Base_Trajipaq_Finale_Ok$Q20 +
-              Base_Trajipaq_Finale_Ok$Q23
+## VI1 + VD1
+
+### Tableau croisé
+
+```{r}
+
+Base_Trajipaq_Finale_Ok <-
+  Base_Trajipaq_Finale_Ok %>%
+  mutate(VD1_F = case_when(
+    VD1 < 0.25 ~ "Faible",
+    VD1 >= 0.25 & VD1 < 0.5 ~ "Moyen",
+    VD1 >= 0.5 & VD1 < 0.75 ~ "Fort",
+    VD1 >= 0.75 & VD1 <= 1 ~ "Tres fort",
+  ))
+
+Base_Trajipaq_Finale_Ok$VD1_F <- factor(Base_Trajipaq_Finale_Ok$VD1_F, levels = c("Tres fort",
+                                                                                "Fort",
+                                                                                "Moyen",
+                                                                                "Faible"
+                                                                                ), ordered = TRUE)
+
+Base_Trajipaq_Finale_Ok <-
+  Base_Trajipaq_Finale_Ok %>%
+  mutate(VI1_F = case_when(
+    VI1 < 0.25 ~ "Faible",
+    VI1 >= 0.25 & VI1 < 0.5 ~ "Moyen",
+    VI1 >= 0.5 & VI1 < 0.75 ~ "Fort",
+    VI1 >= 0.75 & VI1 <= 1 ~ "Tres fort",
+  ))
+
+Base_Trajipaq_Finale_Ok$VI1_F <- factor(Base_Trajipaq_Finale_Ok$VI1_F, levels = c("Tres fort",
+                                                                                "Fort",
+                                                                                "Moyen",
+                                                                                "Faible"
+                                                                                ), ordered = TRUE)
+
+```
+
+```{r}
+
+ctable(x = Base_Trajipaq_Finale_Ok$VI1_F, y = Base_Trajipaq_Finale_Ok$VD1_F)
+
+```
+
+```{r}
+
+ggplot(data = Base_Trajipaq_Finale_Ok, mapping = aes(x = VI1_F, fill = VD1_F)) +
+  geom_bar(position = "dodge", na.rm = TRUE) +
+  theme_bw()
+
+```
+
+### Test d'association
+
+```{r}
+
+chisq.test(x = Base_Trajipaq_Finale_Ok$VI1_F, y = Base_Trajipaq_Finale_Ok$VD1_F)
+
+```
+
+### Modèle de régression
+
+```{r}
+
+reg_1 <- lm(Base_Trajipaq_Finale_Ok$VI1 ~ Base_Trajipaq_Finale_Ok$VD1 +
+              Base_Trajipaq_Finale_Ok$Q10 +     # Sexe
+              Base_Trajipaq_Finale_Ok$Q8B +     # Âge
+              Base_Trajipaq_Finale_Ok$STATUT +  # Statut matrimonial
+              Base_Trajipaq_Finale_Ok$Q4 +      # Lieu de naissance
+              Base_Trajipaq_Finale_Ok$Q16 +     # Niveau d'éducation
+              Base_Trajipaq_Finale_Ok$Q17 +     # Religion
+              Base_Trajipaq_Finale_Ok$Q20 +     # Langue maternelle
+              Base_Trajipaq_Finale_Ok$Q23       # Etat de santé
             )
@@
 
-## Coefficient de corrélation
-
-```{r}
-
-cor.test(Base_Trajipaq_Finale_Ok$I2, Base_Trajipaq_Finale_Ok$I1)
-
-```
-
-## Visualisation
+### Coefficient de corrélation
+
+```{r}
+
+cor.test(Base_Trajipaq_Finale_Ok$VI1, Base_Trajipaq_Finale_Ok$VD1)
+
+```
+
+### Visualisation
 
@@
 vis <- ggplot(Base_Trajipaq_Finale_Ok %>%
-                filter(!is.na(I1), !is.na(I2)), aes(x = I2, y = I1)) +
+                filter(!is.na(VD1), !is.na(VI1)), aes(x = VI1, y = VD1)) +
   geom_jitter(color = "pink") +
-  stat_smooth(color = "purple",se = FALSE) +
+  stat_smooth(method = "lm", formula = "y ~ x", color = "purple",se = FALSE) +
   labs(title = "Discrimination et sentiment d'appartenance", 
@@
 
-## Visualisation contrôlée : sexe
+#### Sexe
 
@@
 vis_sex <- ggplot(Base_Trajipaq_Finale_Ok %>%
-                filter(!is.na(I1), !is.na(I2), !is.na(Q10)), aes(x = I2, y = I1, color = as.factor(Q10))) +
+                filter(!is.na(VD1), !is.na(VI1), !is.na(Q10)), aes(x = VI1, y = VD1, color = as.factor(Q10))) +
   geom_jitter(color = "pink") +
-  stat_smooth(se = FALSE) +
+  stat_smooth(method = "lm", formula = "y ~ x", se = FALSE) +
   scale_color_brewer(name = "Sexe", palette = 3) +
@@
 
-## Visualisation contrôlée : âge
+#### Âge
 
@@
 vis_age <- ggplot(Base_Trajipaq_Finale_Ok %>%
-                filter(!is.na(I1), !is.na(I2), !is.na(Q8A_F)), aes(x = I2, y = I1, color = Q8A_F)) +
+                filter(!is.na(VD1), !is.na(VI1), !is.na(Q8A_F)), aes(x = VI1, y = VD1, color = Q8A_F)) +
   geom_jitter(color = "pink") +
-  stat_smooth(se = FALSE, span = 0.9) +
+  stat_smooth(method = "lm", formula = "y ~ x", se = FALSE) +
   scale_color_brewer(name = "Age", palette = 3) +
@@
 
-## Visualisation contrôlée : statut matrimonial
+#### Statut matrimonial
 
@@
 vis_born <- ggplot(Base_Trajipaq_Finale_Ok %>%
-                filter(!is.na(I1), !is.na(I2), !is.na(STATUT_F)), aes(x = I2, y = I1, color = STATUT_F)) +
+                filter(!is.na(VD1), !is.na(VI1), !is.na(STATUT_F)), aes(x = VI1, y = VD1, color = STATUT_F)) +
   geom_jitter(color = "pink") +
-  stat_smooth(se = FALSE, span = 0.9) +
+  stat_smooth(method = "lm", formula = "y ~ x", se = FALSE) +
   scale_color_brewer(name = "Statut matrimonial", palette = 3) +
@@
 
-## Visualisation contrôlée : statut migratoire
+#### Lieu de naissance
 
@@
 vis_born <- ggplot(Base_Trajipaq_Finale_Ok %>%
-                filter(!is.na(I1), !is.na(I2), !is.na(Q4)), aes(x = I2, y = I1, color = as.factor(Q4))) +
+                filter(!is.na(VD1), !is.na(VI1), !is.na(Q4)), aes(x = VI1, y = VD1, color = as.factor(Q4))) +
   geom_jitter(color = "pink") +
-  stat_smooth(se = FALSE, span = 0.9) +
+  stat_smooth(method = "lm", formula = "y ~ x", se = FALSE) +
   scale_color_brewer(name = "Ne au Canada", palette = 3) +
@@
 
-## Visualisation contrôlée : niveau d'éducation
+#### Niveau d'éducation
 
@@
 vis_educ <- ggplot(Base_Trajipaq_Finale_Ok %>%
-                filter(!is.na(I1), !is.na(I2), !is.na(Q16)), aes(x = I2, y = I1, color = as.factor(Q16))) +
+                filter(!is.na(VD1), !is.na(VI1), !is.na(Q16)), aes(x = VI1, y = VD1, color = as.factor(Q16))) +
   geom_jitter(color = "pink") +
-  stat_smooth(se = FALSE, span = 1) +
+  stat_smooth(method = "lm", formula = "y ~ x", se = FALSE) +
   scale_color_brewer(name = "Niveau d'education", palette = 3) +
@@
 
-## Visualisation contrôlée : religion
+#### Religion
 
@@
 vis_relig <- ggplot(Base_Trajipaq_Finale_Ok %>%
-                filter(!is.na(I1), !is.na(I2), !is.na(Q17_F)), aes(x = I2, y = I1, color = Q17_F)) +
+                filter(!is.na(VD1), !is.na(VI1), !is.na(Q17_F)), aes(x = VI1, y = VD1, color = Q17_F)) +
   geom_jitter(color = "pink") +
-  stat_smooth(se = FALSE, span = 0.9) +
+  stat_smooth(method = "lm", formula = "y ~ x", se = FALSE) +
   scale_color_brewer(name = "Religion", palette = 3) +
@@
 
-## Visualisation contrôlée : langue maternelle
+#### Langue maternelle
 
@@
 vis_lang <- ggplot(Base_Trajipaq_Finale_Ok %>%
-                filter(!is.na(I1), !is.na(I2), !is.na(Q20_F)), aes(x = I2, y = I1, color = Q20_F)) +
+                filter(!is.na(VD1), !is.na(VI1), !is.na(Q20_F)), aes(x = VI1, y = VD1, color = Q20_F)) +
   geom_jitter(color = "pink") +
-  stat_smooth(se = FALSE, span = 0.8) +
+  stat_smooth(method = "lm", formula = "y ~ x", se = FALSE) +
   scale_color_brewer(name = "Langue maternelle", palette = 3) +
@@
 
-## Visualisation contrôlée : état de santé
+#### Etat de santé
 
@@
 vis_health <- ggplot(Base_Trajipaq_Finale_Ok %>%
-                filter(!is.na(I1), !is.na(I2), !is.na(Q23)), aes(x = I2, y = I1, color = as.factor(Q23))) +
+                filter(!is.na(VD1), !is.na(VI1), !is.na(Q23)), aes(x = VI1, y = VD1, color = as.factor(Q23))) +
   geom_jitter(color = "pink") +
-  stat_smooth(se = FALSE, span = 1.1) +
+  stat_smooth(method = "lm", formula = "y ~ x", se = FALSE) +
   scale_color_brewer(name = "Etat de sante", palette = 3) +

### 2023-05-23 Article/Visseho_al_sentiment_appartenance.Rmd
update

@@
 
+1. Sentiment d'appartenance au Québec (VD1) (indice)
+
+2. Double sentiment d'appartenance (pays origine/accueil) VD2
+
 - Crée la variable Sentiment d'appartenance qui prend les valeurs suivantes (selon les terminologies utilisées par Berry et Hou, 2017)
@@
 
-  - Sentiment d'appartenance au Canada
-  - Sentiment d'appartenance à leur pays d'origine
+  - Sentiment d'appartenance au Canada Q39E (recodé à oui/non)
+  - Sentiment d'appartenance à leur pays d'origine Q39F (recodé à oui/non)
+
+                    Q39E
+  Q39F              Oui       non
+            Oui     intégré   assimilé
+            non     séparé    marginal
+
+3. double appartenance (Québec/Canada) VD3
+
+  - Exclusivement au Canada
+  - Exclusivement au Québec
+  - Aux deux 
+  - Ni l'un ni l'autre
+
+Utilisé les variables 
+  - Q39D: Je me sens Québécois
+  - Q39E: Je me sens Canadien
 
@@
 
-- Origine de l'immigrant
-  - Régions d'origine
-  - Proximité du pays d'accueil par rapport au pays d'origine 
-  - Niveau de stabilité dans le pays d'accueil ou d'origine au moment de l'enquête
+**- Origine de l'immigrant**
 
-- Discrimination perçue
+1.  - Régions d'origine VI1
+    - Tableau croisé entre region d'origine et nouvelle variable
+
+2.  - Proximité du pays d'accueil par rapport au pays d'origine  VI2
+    Comment on mesure cette proximité: 
+      - region d'origine en deux catégories (Europe-Amerique/Autres)
+      - minorité visible / non minorité visible
+
+3.  - Niveau de stabilité dans le pays d'accueil ou d'origine au moment de l'enquête VI3
+
+**- Discrimination perçue** VI4
+
+**Statistiques à sortir**
+
+1. Statistique univarié (fréquences) de 
+  - VD1 (moyenne et écart type)
+  - VD2
+  - VD3
+
+2. Statistique univariés (fréquence) de
+- VI1
+- VI2
+- VI3
+- VI4 (moyenne et écart-type)
+
+3. tableau croisé entre
+
+  - VD1 (recodé) et (VI1 à VI3)
+  - VD2 et (VI1 à VI3)
+  - VD3 et (VI1 à VI3)
+
+4. Avec discrimation (le transformer en en ordinal)  
+  - jamais, parfois, souvent 
+  - faible, modéré, fort
+
+  refaire les tableaux croisés entre les VD avec VI4-recodé
+
+5. Entre VD1 et VI4, on peut avoir 
+    - régression linéraire
+
+## Question de causalité
+
+Discrimination ----------> Sentiment d'appartenance
+
+ceux qui disent avoir subi une discrimination ont un plus faible sentiment d'appartenance
+
+nombre de pompiers ------> dommage à la propriété
 

### 2023-05-24 Article/Visseho_al_sentiment_appartenance.Rmd
Réorganisation du travail.

--- 2023-05-22 Article/Visseho_al_sentiment_appartenance.Rmd
@@
 ---
+
+```{=html}
 <!--
@@
 -->
-
+```
 \newpage
@@
 Les immigrants sont souvent pris entre deux ou plusieurs cultures, celles de leur pays d'origine et celles de leur nouveau pays de résidence.
-Cet article essaie de comprendre comment ces sentiments sont associés à leur pays d'origine et aux éventuelles situations de discriminations vécues qu'ils vivent dans leur pays d'accueil
+Cet article essaie de comprendre comment ces sentiments sont associés à leur pays d'origine et aux éventuelles situations de discriminations vécues qu'ils vivent dans leur pays d'accueil.
 
-## Variables à utiliser
+## Variables
 
-**Variables dépendantes**
+### Variables dépendantes
 
-- Crée la variable Sentiment d'appartenance qui prend les valeurs suivantes (selon les terminologies utilisées par Berry et Hou, 2017)
+#### VD1. Sentiment d'appartenance au Québec
 
-  - Exclusivement au pays de naissance (separated, 5%)
-  - Exclusivement au pays d'accueil (assimilated, 15%)
-  - Aux deux  (integrated, 75%)
-  - Ni l'un ni l'autre (marginalised, 6%)
+-   Q39A (facteur/ordinal)
 
-**Quelle est la distribution de cette nouvelle variable?**
+-   Q39B (facteur/ordinal)
 
-Cette variable est crée par combinaison des deux variables suivantes:
+-   Q39C (facteur/ordinal)
 
-  - Sentiment d'appartenance au Canada
-  - Sentiment d'appartenance à leur pays d'origine
+-   Q39D (facteur/ordinal)
 
-**Variables indépendantes**
+-   VD1 (numérique/ratio)
 
-- Origine de l'immigrant
-  - Régions d'origine
-  - Proximité du pays d'accueil par rapport au pays d'origine 
-  - Niveau de stabilité dans le pays d'accueil ou d'origine au moment de l'enquête
+#### VD2. Double sentiment d'appartenance (Canada/Pays d'origine)
 
-- Discrimination perçue
+-   Q39E (facteur/nominal)
+
+-   Q39F (facteur/nominal)
+
+-   VD2 (facteur/nominal)
+
+    |                    | **Canada** | **Oui**  | **Non**     |
+    |--------------------|------------|----------|-------------|
+    | **Pays d'origine** |            |          |             |
+    | **Oui**            |            | Intégré  | Séparé      |
+    | **Non**            |            | Assimilé | Marginalisé |
+
+#### VD3. Double sentiment d'appartenance (Canada/Québec)
+
+-   Q39D (facteur/nominal)
+
+-   Q39E (facteur/nominal)
+
+-   VD3 (facteur/nominal)
+
+### Variables indépendantes
+
+#### VI1. Région d'origine
+
+-   Q11 (facteur/nominal)
+
+-   VI1 (facteur/nominal)
+
+#### VI2. Distance culturelle avec le Canada
+
+-   ??? (numérique/ratio)
+
+#### VI3. Stabilité dans le pays d'origine
+
+-   ??? (numérique/ratio)
+
+#### VI4. Discrimination subie au cours des 12 derniers mois
+
+-   Q40A (facteur/ordinal)
+
+-   Q40B (facteur/ordinal)
+
+-   Q40C (facteur/ordinal)
+
+-   Q40D (facteur/ordinal)
+
+-   VI4 (numérique/ratio)
+
+## Analyses
+
+### Variables dépendantes
+
+-   VD1 (moyenne/écart-type)
+
+-   VD2 (fréquence/mode)
+
+-   VD3 (fréquence/mode)
+
+### Variables indépendantes
+
+-   VI1 (fréquence/mode)
+
+-   VI2 (moyenne/écart-type)
+
+-   VI3 (moyenne/écart-type)
+
+-   VI4 (moyenne/écart-type)
+
+### Tableaux croisés
+
+-   VD1 / (VI1 : VI3)
+
+-   VD2 / (VI1 : VI3)
+
+-   VD3 / (VI1 : VI3)
+
+### Régression linéaire
+
+-   VD1 / VI4
+
+## Causalité éventuelle
+
+Discrimination --\> Sentiment d'appartenance
 
@@
 Social Science Research, 42(3), 775‑788.
-https://doi.org/10.1016/j.ssresearch.2012.12.019
+<https://doi.org/10.1016/j.ssresearch.2012.12.019>
 
@@
 (2015).
-Immigrants’ Sense of Belonging to the Host Country: The Role of Life Satisfaction, Language Proficiency, and Religious Motives.
+Immigrants' Sense of Belonging to the Host Country: The Role of Life Satisfaction, Language Proficiency, and Religious Motives.
 Social Indicators Research, 124(3), 947‑961.
-https://doi.org/10.1007/s11205-014-0823-3
+<https://doi.org/10.1007/s11205-014-0823-3>
 
@@
 Lawrence Erlbaum Associates Publishers.
-https://doi.org/10.4324/9780415963619-4
+<https://doi.org/10.4324/9780415963619-4>
 
@@
 Personality and Social Psychology Bulletin, 33(10), 1448‑1462.
-https://doi.org/10.1177/0146167207304276
+<https://doi.org/10.1177/0146167207304276>
 
-- Ngo, H. et Li, H. (2016). Cultural Identity and Adaptation of Mainland Chinese Immigrants in Hong Kong. American Behavioral Scientist, 60(5), 730‑749. https://doi.org/10.1177/0002764216632837
+-   Ngo, H. et Li, H. (2016). Cultural Identity and Adaptation of Mainland Chinese Immigrants in Hong Kong. American Behavioral Scientist, 60(5), 730‑749. <https://doi.org/10.1177/0002764216632837>
 
@@
 Psychological Well-Being, Multiple Identities, and Discrimination Among First and Second Generation Immigrant Muslims.
-Europe’s Journal of Psychology, 14(1), 66‑87.
-https://doi.org/10.5964/ejop.v14i1.1434
+Europe's Journal of Psychology, 14(1), 66‑87.
+<https://doi.org/10.5964/ejop.v14i1.1434>
 
@@
 Journal of Social Issues, 57(3), 493‑510.
-https://doi.org/10.1111/0022-4537.00225
+<https://doi.org/10.1111/0022-4537.00225>
 
@@
 Journal of Counseling Psychology, 54, 271‑281.
-https://doi.org/10.1037/0022-0167.54.3.271
+<https://doi.org/10.1037/0022-0167.54.3.271>
 
-- Verkuyten, M. (2016). The Integration Paradox: Empiric Evidence From the Netherlands. American Behavioral Scientist, 60(5), 583‑596. https://doi.org/10.1177/0002764216632838
+-   Verkuyten, M. (2016). The Integration Paradox: Empiric Evidence From the Netherlands. American Behavioral Scientist, 60(5), 583‑596. <https://doi.org/10.1177/0002764216632838>
 
@@
 (2012).
-Immigrants’ National Identification: Meanings, Determinants, and Consequences.
+Immigrants' National Identification: Meanings, Determinants, and Consequences.
 Social Issues and Policy Review, 6(1), 82‑112.
-https://doi.org/10.1111/j.1751-2409.2011.01036.x
+<https://doi.org/10.1111/j.1751-2409.2011.01036.x>
 
@@
 International Migration Review, 41(3), 769‑781.
-https://doi.org/10.1111/j.1747-7379.2007.00093.x
+<https://doi.org/10.1111/j.1747-7379.2007.00093.x>
 
@@
 International Journal of Intercultural Relations, 30(6), 719‑734.
-https://doi.org/10.1016/j.ijintrel.2006.06.004
+<https://doi.org/10.1016/j.ijintrel.2006.06.004>
 
-- Berry, J. W. (1997). Immigration, Acculturation, and Adaptation. Applied Psychology, 46(1), 5‑34. https://doi.org/10.1111/j.1464-0597.1997.tb01087.x
+-   Berry, J. W. (1997). Immigration, Acculturation, and Adaptation. Applied Psychology, 46(1), 5‑34. <https://doi.org/10.1111/j.1464-0597.1997.tb01087.x>
 
@@
 International Journal of Intercultural Relations, 61, 29‑39.
-https://doi.org/10.1016/j.ijintrel.2017.08.003
+<https://doi.org/10.1016/j.ijintrel.2017.08.003>
 
@@
 International Review of Psychiatry, 33(1), 140‑153.
-https://doi.org/10.1080/09540261.2020.1750801
+<https://doi.org/10.1080/09540261.2020.1750801>
 
@@
 International Journal of Intercultural Relations, 34(3), 191‑207.
-https://doi.org/10.1016/j.ijintrel.2009.11.007
+<https://doi.org/10.1016/j.ijintrel.2009.11.007>

### 2023-05-24 Codes/Sentiment_an1.Rmd
Réorganisation du travail.

---
title: "Sentiment_an1"
author: "Thea"
date: "2023-05-24"
output: html_document
---

### 2023-05-24 Codes/Sentiment_an1.Rmd
Révision complète du travail.

@@
 ---
+
+```{r}
+library(summarytools)
+library(tidyverse)
+library(ggthemes)
+library(haven)
+```
+
+```{r}
+Base_Trajipaq_Finale_Ok <- read_dta("~/Desktop/Base_Trajipaq_Finale_Ok.dta")
+```
+
+# Présentation des variables
+
+## Variables de contrôle
+
+### Sexe
+
+```{r}
+
+# Q10 = êtes-vous un homme, une femme ou autre (précisez) ?
+
+Base_Trajipaq_Finale_Ok <-
+  Base_Trajipaq_Finale_Ok %>%
+  mutate(SEXE = case_when(
+    Q10 == 1 ~ "Homme",
+    Q10 == 2 ~ "Femme",
+    Q10 == 97 ~ "Autre"
+  ))
+
+```
+
+```{r}
+
+freq(Base_Trajipaq_Finale_Ok$SEXE)
+
+```
+
+```{r}
+
+ggplot(data = Base_Trajipaq_Finale_Ok, mapping = aes(x = SEXE)) +
+  geom_bar(fill = "pink") +
+  labs(
+    title = "Distribution du sexe",
+    x = "Sexe"
+       ) +
+  coord_flip() +
+  theme_bw()
+
+```
+
+### Âge
+
+```{r}
+
+# QAGE = à quel groupe d'âge appartenez-vous ?
+
+Base_Trajipaq_Finale_Ok <-
+  Base_Trajipaq_Finale_Ok %>%
+  mutate(AGE = case_when(
+    QAGE == 1 ~ "[0;19] ans",
+    QAGE == 2 ~ "[20;24] ans",
+    QAGE == 3 ~ "[25;29] ans",
+    QAGE == 4 ~ "[30;34] ans",
+    QAGE == 5 ~ "[35;39] ans",
+    QAGE == 6 ~ "[40;44] ans",
+    QAGE == 7 ~ "[45;49] ans",
+    QAGE == 8 ~ "[50;60] ans",
+    QAGE == 9 ~ "[61;99] ans",
+    QAGE == 99 ~ NA
+  ))
+
+Base_Trajipaq_Finale_Ok$AGE <- factor(Base_Trajipaq_Finale_Ok$AGE,levels = c(
+  "[0;19] ans",
+  "[20;24] ans",
+  "[25;29] ans",
+  "[30;34] ans",
+  "[35;39] ans",
+  "[40;44] ans",
+  "[45;49] ans",
+  "[50;60] ans",
+  "[61;99] ans",
+  NA
+), ordered = TRUE)
+
+```
+
+```{r}
+
+freq(Base_Trajipaq_Finale_Ok$AGE)
+
+```
+
+```{r}
+
+ggplot(data = Base_Trajipaq_Finale_Ok, mapping = aes(x = AGE)) +
+  geom_bar(fill = "pink") +
+  labs(
+    title = "Distribution de l'age par groupes",
+    x = "Age"
+       ) +
+  coord_flip() +
+  theme_bw()
+
+```
+
+### Education
+
+```{r}
+
+Base_Trajipaq_Finale_Ok <-
+  Base_Trajipaq_Finale_Ok %>%
+  mutate(EDUCATION = case_when(
+    Q26 == 1 ~ "Elementaire",
+    Q26 == 2 ~ "Secondaire",
+    Q26 == 3 ~ "Collegial",
+    Q26 == 4 ~ "Post secondaire",
+    Q26 == 5 ~ "Universitaire",
+    Q26 == 6 ~ NA,
+    Q26 == 98 ~ NA,
+    Q26 == 99 ~ NA
+  ))
+
+Base_Trajipaq_Finale_Ok$EDUCATION <- factor(Base_Trajipaq_Finale_Ok$EDUCATION,levels = c(
+  "Elementaire",
+  "Secondaire",
+  "Post secondaire",
+  "Collegial",
+  "Universitaire",
+  NA
+), ordered = TRUE)
+
+```
+
+```{r}
+
+freq(Base_Trajipaq_Finale_Ok$EDUCATION)
+
+```
+
+```{r}
+
+ggplot(subset(Base_Trajipaq_Finale_Ok, !is.na(EDUCATION)), aes(x = EDUCATION)) +
+  geom_bar(fill = "pink") +
+  labs(
+    title = "Distribution du niveau d'education",
+    x = "Niveau d'education"
+       ) +
+  coord_flip() +
+  theme_bw()
+
+```
+
+### Origine
+
+```{r}
+
+Base_Trajipaq_Finale_Ok <-
+  Base_Trajipaq_Finale_Ok %>%
+  mutate(ORIGINE = case_when(
+    region_naissance == "ANMO" ~ "Afrique du Nord et Moyen Orient",
+    region_naissance == "ACAS" ~ "Amerique centrale et Amerique su Sud",
+    region_naissance == "Afrique subsaharienne" ~ "Afrique subsaharienne",
+    region_naissance == "Asie" ~ "Asie",
+    region_naissance == "Autres" ~ "Autres",
+    region_naissance == "Canada" ~ "Canada",
+    region_naissance == "Caraibes" ~ "Caraibes",
+    region_naissance == "Europe" ~ "Europe",
+    region_naissance == "France" ~ "France",
+    region_naissance == "NSP" ~ NA,
+  ))
+
+```
+
+```{r}
+
+freq(Base_Trajipaq_Finale_Ok$ORIGINE)
+
+```
+
+```{r}
+
+ggplot(subset(Base_Trajipaq_Finale_Ok, !is.na(ORIGINE)), aes(x = ORIGINE)) +
+  geom_bar(fill = "pink") +
+  labs(
+    title = "Distribution de l'origine par region",
+    x = "Region"
+       ) +
+  coord_flip() +
+  theme_bw()
+
+```
+
+### Revenu
+
+```{r}
+
+Base_Trajipaq_Finale_Ok <-
+  Base_Trajipaq_Finale_Ok %>%
+  mutate(REVENU = case_when(
+    Q100 == 1 ~ "[0 ; 19 999]",
+    Q100 == 2 ~ "[0 ; 19 999]",
+    Q100 == 3 ~ "[20 000 ; 39 999]",
+    Q100 == 4 ~ "[20 000 ; 39 999]",
+    Q100 == 5 ~ "[40 000 ; 59 999]",
+    Q100 == 6 ~ "[40 000 ; 59 999]",
+    Q100 == 7 ~ "[60 000 ; 79 999]",
+    Q100 == 8 ~ "[60 000 ; 79 999]",
+    Q100 == 9 ~ "[80 000 ; 99 999]",
+    Q100 == 10 ~ "[90 000 ; 99 999]",
+    Q100 == 11 ~ "[100 000 ; 149 999]",
+    Q100 == 12 ~ "[150 000 et plus]",
+    Q100 == 98 ~ NA,
+    Q100 == 99 ~ NA
+  ))
+
+Base_Trajipaq_Finale_Ok$REVENU <- factor(Base_Trajipaq_Finale_Ok$REVENU,levels = c(
+  "[0 ; 19 999]",
+  "[20 000 ; 39 999]",
+  "[40 000 ; 59 999]",
+  "[60 000 ; 79 999]",
+  "[80 000 ; 99 999]",
+  "[100 000 ; 149 999]",
+  "[150 000 et plus]",
+  NA
+), ordered = TRUE)
+
+```
+
+```{r}
+
+freq(Base_Trajipaq_Finale_Ok$REVENU)
+
+```
+
+```{r}
+
+ggplot(subset(Base_Trajipaq_Finale_Ok, !is.na(REVENU)), aes(x = REVENU)) +
+  geom_bar(fill = "pink") +
+  labs(
+    title = "Distribution du revenu par groupes",
+    x = "Revenu"
+       ) +
+  coord_flip() +
+  theme_bw()
+
+```
+
+### Langue maternelle
+
+```{r}
+
+Base_Trajipaq_Finale_Ok <-
+  Base_Trajipaq_Finale_Ok %>%
+  mutate(LANGUE = case_when(
+    Q20M1 == 1 ~ "Francais",
+    Q20M1 == 2 ~ "Anglais",
+    Q20M1 == 97 ~ "Autre",
+    Q20M1 == 98 ~ NA,
+    Q20M1 == 99 ~ NA
+  ))
+
+```
+
+```{r}
+
+freq(Base_Trajipaq_Finale_Ok$LANGUE)
+
+```
+
+```{r}
+
+ggplot(subset(Base_Trajipaq_Finale_Ok, !is.na(LANGUE)), aes(x = LANGUE)) +
+  geom_bar(fill = "pink") +
+  labs(
+    title = "Distribution de la langue maternelle",
+    x = "Langue maternelle"
+       ) +
+  coord_flip() +
+  theme_bw()
+
+```
+
+### Religion
+
+```{r}
+
+Base_Trajipaq_Finale_Ok <-
+  Base_Trajipaq_Finale_Ok %>%
+  mutate(RELIGION = case_when(
+    Q17 == 0 ~ "Sans religion",
+    Q17 == 1 ~ "Catholique",
+    Q17 == 2 ~ "Protestante",
+    Q17 == 3 ~ "Autres Chretiens",
+    Q17 == 4 ~ "Autres Chretiens",
+    Q17 == 5 ~ "Autres Chretiens",
+    Q17 == 6 ~ "Orthodoxe",
+    Q17 == 7 ~ "Autres Chretiens",
+    Q17 == 8 ~ "Musulmans",
+    Q17 == 9 ~ "Autre",
+    Q17 == 10 ~ "Autre",
+    Q17 == 11 ~ "Autre",
+    Q17 == 12 ~ "Autre",
+    Q17 == 97 ~ "Autre",
+    Q17 == 98 ~ NA,
+    Q17 == 99 ~ NA
+  ))
+
+```
+
+```{r}
+
+freq(Base_Trajipaq_Finale_Ok$RELIGION)
+
+```
+
+```{r}
+
+ggplot(subset(Base_Trajipaq_Finale_Ok, !is.na(RELIGION)), aes(x = RELIGION)) +
+  geom_bar(fill = "pink") +
+  labs(
+    title = "Distribution de la religion",
+    x = "Religion"
+       ) +
+  coord_flip() +
+  theme_bw()
+
+```
+
+### Statut matrimonial
+
+```{r}
+
+Base_Trajipaq_Finale_Ok <-
+  Base_Trajipaq_Finale_Ok %>%
+  mutate(MATRIMONIAL = case_when(
+    STATUT == 1 ~ "Celibataire",
+    STATUT == 2 ~ "Marie(e)",
+    STATUT == 3 ~ "Conjoint(e) de fait",
+    STATUT == 4 ~ "Separe(e)",
+    STATUT == 5 ~ "Divorce(e)",
+    STATUT == 6 ~ "Veuf(ve)"
+  ))
+
+```
+
+```{r}
+
+freq(Base_Trajipaq_Finale_Ok$MATRIMONIAL)
+
+```
+
+```{r}
+
+ggplot(subset(Base_Trajipaq_Finale_Ok, !is.na(MATRIMONIAL)), aes(x = MATRIMONIAL)) +
+  geom_bar(fill = "pink") +
+  labs(
+    title = "Distribution du statut matrimonial",
+    x = "Statut matrimonial"
+       ) +
+  coord_flip() +
+  theme_bw()
+
+```
+
+## Variables dépendantes
+
+### VD1 : sentiment d'appartenance au Québec
+
+```{r}
+
+Base_Trajipaq_Finale_Ok <-
+  Base_Trajipaq_Finale_Ok %>%
+  mutate(Q39A = case_when(
+    Q39A == 1 ~ 4,
+    Q39A == 2 ~ 3,
+    Q39A == 3 ~ 2,
+    Q39A == 4 ~ 1,
+    Q39A == 98 ~ NA,
+    Q39A == 99 ~ NA
+  ))
+
+Base_Trajipaq_Finale_Ok <-
+  Base_Trajipaq_Finale_Ok %>%
+  mutate(Q39B = case_when(
+    Q39B == 1 ~ 4,
+    Q39B == 2 ~ 3,
+    Q39B == 3 ~ 2,
+    Q39B == 4 ~ 1,
+    Q39B == 98 ~ NA,
+    Q39B == 99 ~ NA
+  ))
+
+Base_Trajipaq_Finale_Ok <-
+  Base_Trajipaq_Finale_Ok %>%
+  mutate(Q39C = case_when(
+    Q39C == 1 ~ 4,
+    Q39C == 2 ~ 3,
+    Q39C == 3 ~ 2,
+    Q39C == 4 ~ 1,
+    Q39C == 98 ~ NA,
+    Q39C == 99 ~ NA
+  ))
+
+Base_Trajipaq_Finale_Ok <-
+  Base_Trajipaq_Finale_Ok %>%
+  mutate(Q39D = case_when(
+    Q39D == 1 ~ 4,
+    Q39D == 2 ~ 3,
+    Q39D == 3 ~ 2,
+    Q39D == 4 ~ 1,
+    Q39D == 98 ~ NA,
+    Q39D == 99 ~ NA
+  ))
+
+```
+
+```{r}
+
+v_VD1 <- c(
+  "Q39A",
+  "Q39B",
+  "Q39C",
+  "Q39D"
+)
+
+Base_Trajipaq_Finale_Ok <-
+  Base_Trajipaq_Finale_Ok %>%
+  rowwise() %>%
+  mutate(VD1 = (sum(c_across(all_of(v_VD1)), na.rm = TRUE) - length(v_VD1)) / (3*length(v_VD1))) %>%
+  mutate(VD1 = if (VD1 < 0) NA_real_ else VD1) %>%
+  ungroup()
+
+```
+
+```{r}
+
+descr(Base_Trajipaq_Finale_Ok$VD1)
+
+```
+
+```{r}
+
+ggplot(data = Base_Trajipaq_Finale_Ok, mapping = aes(x = VD1)) +
+  geom_histogram(na.rm = TRUE, fill = "pink", binwidth = 0.01) +
+  labs(
+    title = "Distribution du sentiment d'appartenance au Quebec",
+    x = "Indice du sentiment d'appartenance au Quebec"
+       ) +
+  theme_bw()
+
+```
+
+### VD2 : sentiment d'appartenance (Canada)
+
+```{r}
+
+Base_Trajipaq_Finale_Ok <-
+  Base_Trajipaq_Finale_Ok %>%
+  mutate(VD2 = case_when(
+    Q39E == 1 ~ "Oui",
+    Q39E == 2 ~ "Oui",
+    Q39E == 3 ~ "Non",
+    Q39E == 4 ~ "Non",
+    Q39E == 98 ~ NA,
+    Q39E == 99 ~ NA
+  ))
+
+```
+
+```{r}
+
+freq(Base_Trajipaq_Finale_Ok$VD2)
+
+```
+
+```{r}
+
+ggplot(subset(Base_Trajipaq_Finale_Ok, !is.na(VD2)), aes(x = VD2)) +
+  geom_bar(fill = "pink") +
+  labs(
+    title = "Distribution du sentiment d'appartenance au Canada",
+    x = "Sentiment d'appartenance au Canada"
+       ) +
+  coord_flip() +
+  theme_bw()
+
+```
+
+### VD3 : sentiment d'appartenance au pays d'origine
+
+```{r}
+
+Base_Trajipaq_Finale_Ok <-
+  Base_Trajipaq_Finale_Ok %>%
+  mutate(VD3 = case_when(
+    Q39F == 1 ~ "Oui",
+    Q39F == 2 ~ "Oui",
+    Q39F == 3 ~ "Non",
+    Q39F == 4 ~ "Non",
+    Q39F == 98 ~ NA,
+    Q39F == 99 ~ NA
+  ))
+
+```
+
+```{r}
+
+freq(Base_Trajipaq_Finale_Ok$VD3)
+
+```
+
+```{r}
+
+ggplot(subset(Base_Trajipaq_Finale_Ok, !is.na(VD3)), aes(x = VD3)) +
+  geom_bar(fill = "pink") +
+  labs(
+    title = "Distribution du sentiment d'appartenance au pays d'origine",
+    x = "Sentiment d'appartenance au pays d'origine"
+       ) +
+  coord_flip() +
+  theme_bw()
+
+```
+
+## Variables indépendantes
+
+### VI4 : discrimination subie au cours des 12 derniers mois
+
+```{r}
+
+Base_Trajipaq_Finale_Ok <-
+  Base_Trajipaq_Finale_Ok %>%
+  mutate(Q40A = case_when(
+    Q40A == 1 ~ 3,
+    Q40A == 2 ~ 2,
+    Q40A == 3 ~ 1,
+    Q40A == 96 ~ NA,
+    Q40A == 98 ~ NA,
+    Q40A == 99 ~ NA
+  ))
+
+Base_Trajipaq_Finale_Ok <-
+  Base_Trajipaq_Finale_Ok %>%
+  mutate(Q40B = case_when(
+    Q40B == 1 ~ 3,
+    Q40B == 2 ~ 2,
+    Q40B == 3 ~ 1,
+    Q40B == 96 ~ NA,
+    Q40B == 98 ~ NA,
+    Q40B == 99 ~ NA
+  ))
+
+Base_Trajipaq_Finale_Ok <-
+  Base_Trajipaq_Finale_Ok %>%
+  mutate(Q40C = case_when(
+    Q40C == 1 ~ 3,
+    Q40C == 2 ~ 2,
+    Q40C == 3 ~ 1,
+    Q40C == 96 ~ NA,
+    Q40C == 98 ~ NA,
+    Q40C == 99 ~ NA
+  ))
+
+Base_Trajipaq_Finale_Ok <-
+  Base_Trajipaq_Finale_Ok %>%
+  mutate(Q40D = case_when(
+    Q40D == 1 ~ 3,
+    Q40D == 2 ~ 2,
+    Q40D == 3 ~ 1,
+    Q40D == 96 ~ NA,
+    Q40D == 98 ~ NA,
+    Q40D == 99 ~ NA
+  ))
+
+```
+
+```{r}
+
+v_VI4 <- c(
+  "Q40A",
+  "Q40B",
+  "Q40C",
+  "Q39D"
+  )
+
+Base_Trajipaq_Finale_Ok <-
+  Base_Trajipaq_Finale_Ok %>%
+  rowwise() %>%
+  mutate(VI4 = (sum(c_across(all_of(v_VI4)), na.rm = TRUE) - length(v_VI4)) / (2.25*length(v_VI4))) %>%
+  mutate(VI4 = if (VI4 < 0) NA_real_ else VI4) %>%
+  ungroup()
+
+```
+
+```{r}
+
+descr(Base_Trajipaq_Finale_Ok$VI4)
+
+```
+
+```{r}
+
+ggplot(data = Base_Trajipaq_Finale_Ok, mapping = aes(x = VI4)) +
+  geom_histogram(na.rm = TRUE, fill = "pink", binwidth = 0.01) +
+  labs(
+    title = "Distribution de la discrimination subie au cours des 12 derniers mois",
+    x = "Indice de la discrimination subie au cours des 12 derniers mois"
+       ) +
+  theme_bw()
+
+```
+
+# Analyse
+
+## H1 : VD1 entretient une relation négative avec VI4
+
+### Modèle de régression
+
+```{r}
+
+reg_1 <- lm(Base_Trajipaq_Finale_Ok$VI4 ~ Base_Trajipaq_Finale_Ok$VD1 +
+              Base_Trajipaq_Finale_Ok$SEXE +
+              Base_Trajipaq_Finale_Ok$AGE +
+              Base_Trajipaq_Finale_Ok$EDUCATION +
+              Base_Trajipaq_Finale_Ok$ORIGINE +
+              Base_Trajipaq_Finale_Ok$REVENU +
+              Base_Trajipaq_Finale_Ok$LANGUE +
+              Base_Trajipaq_Finale_Ok$RELIGION +
+              Base_Trajipaq_Finale_Ok$MATRIMONIAL
+            )
+summary(reg_1)
+
+```
+
+### Coefficient de corrélation
+
+```{r}
+
+cor.test(Base_Trajipaq_Finale_Ok$VI4, Base_Trajipaq_Finale_Ok$VD1)
+
+```
+
+### Visualisation
+
+```{r}
+
+vis_H1 <- ggplot(Base_Trajipaq_Finale_Ok %>%
+                filter(!is.na(VD1), !is.na(VI4), !is.na(SEXE)), aes(x = VI4, y = VD1)) +
+  geom_jitter(color = "pink") +
+  stat_smooth(method = "lm", 
+              formula = "y ~ x", 
+              se = FALSE,
+              color = "purple") +
+  labs(title = "Sentiment d'appartenance et discrimination", 
+       x = "Discrimination subie au cours des 12 derniers mois", 
+       y = "Sentiment d'appartenance au Quebec") +
+  theme_bw()
+print(vis_H1)
+
+```
+
+#### Sexe
+
+```{r}
+
+vis_sex <- ggplot(Base_Trajipaq_Finale_Ok %>%
+                filter(!is.na(VD1), !is.na(VI4), !is.na(SEXE)), aes(x = VI4, y = VD1, color = SEXE)) +
+  geom_jitter(color = "pink") +
+  stat_smooth(method = "lm", formula = "y ~ x", se = FALSE) +
+  scale_color_brewer(name = "Sexe", palette = 1) +
+  labs(title = "Sentiment d'appartenance et discrimination en fonction du sexe", 
+       x = "Discrimination subie au cours des 12 derniers mois", 
+       y = "Sentiment d'appartenance au Quebec") +
+  theme_bw()
+print(vis_sex)
+
+```
+
+#### Age
+
+```{r}
+
+vis_sex <- ggplot(Base_Trajipaq_Finale_Ok %>%
+                filter(!is.na(VD1), !is.na(VI4), !is.na(AGE)), aes(x = VI4, y = VD1, color = as.factor(AGE))) +
+  geom_jitter(color = "pink") +
+  stat_smooth(method = "lm", formula = "y ~ x", se = FALSE) +
+  scale_color_brewer(name = "Age", palette = 1) +
+  labs(title = "Sentiment d'appartenance et discrimination en fonction de l'age", 
+       x = "Discrimination subie au cours des 12 derniers mois", 
+       y = "Sentiment d'appartenance au Quebec") +
+  theme_bw()
+print(vis_sex)
+
+```
+
+#### Education
+
+```{r}
+
+vis_sex <- ggplot(Base_Trajipaq_Finale_Ok %>%
+                filter(!is.na(VD1), !is.na(VI4), !is.na(EDUCATION)), aes(x = VI4, y = VD1, color = as.factor(EDUCATION))) +
+  geom_jitter(color = "pink") +
+  stat_smooth(method = "lm", formula = "y ~ x", se = FALSE) +
+  scale_color_brewer(name = "Niveau d'education", palette = 1) +
+  labs(title = "Sentiment d'appartenance et discrimination en fonction de l'education", 
+       x = "Discrimination subie au cours des 12 derniers mois", 
+       y = "Sentiment d'appartenance au Quebec") +
+  theme_bw()
+print(vis_sex)
+
+```
+
+#### Origine
+
+```{r}
+
+vis_sex <- ggplot(Base_Trajipaq_Finale_Ok %>%
+                filter(!is.na(VD1), !is.na(VI4), !is.na(ORIGINE)), aes(x = VI4, y = VD1, color = as.factor(ORIGINE))) +
+  geom_jitter(color = "pink") +
+  stat_smooth(method = "lm", formula = "y ~ x", se = FALSE) +
+  scale_color_brewer(name = "Region d'origine", palette = 1) +
+  labs(title = "Sentiment d'appartenance et discrimination en fonction de l'origine", 
+       x = "Discrimination subie au cours des 12 derniers mois", 
+       y = "Sentiment d'appartenance au Quebec") +
+  theme_bw()
+print(vis_sex)
+
+```
+
+#### Revenu
+
+```{r}
+
+vis_sex <- ggplot(Base_Trajipaq_Finale_Ok %>%
+                filter(!is.na(VD1), !is.na(VI4), !is.na(REVENU)), aes(x = VI4, y = VD1, color = as.factor(REVENU))) +
+  geom_jitter(color = "pink") +
+  stat_smooth(method = "lm", formula = "y ~ x", se = FALSE) +
+  scale_color_brewer(name = "Revenu", palette = 1) +
+  labs(title = "Sentiment d'appartenance et discrimination en fonction du revenu", 
+       x = "Discrimination subie au cours des 12 derniers mois", 
+       y = "Sentiment d'appartenance au Quebec") +
+  theme_bw()
+print(vis_sex)
+
+```
+
+#### Langue maternelle
+
+```{r}
+
+vis_sex <- ggplot(Base_Trajipaq_Finale_Ok %>%
+                filter(!is.na(VD1), !is.na(VI4), !is.na(LANGUE)), aes(x = VI4, y = VD1, color = as.factor(LANGUE))) +
+  geom_jitter(color = "pink") +
+  stat_smooth(method = "lm", formula = "y ~ x", se = FALSE) +
+  scale_color_brewer(name = "Langue maternelle", palette = 1) +
+  labs(title = "Sentiment d'appartenance et discrimination en fonction de la langue maternelle", 
+       x = "Discrimination subie au cours des 12 derniers mois", 
+       y = "Sentiment d'appartenance au Quebec") +
+  theme_bw()
+print(vis_sex)
+
+```
+
+### Religion
+
+```{r}
+
+vis_sex <- ggplot(Base_Trajipaq_Finale_Ok %>%
+                filter(!is.na(VD1), !is.na(VI4), !is.na(RELIGION)), aes(x = VI4, y = VD1, color = as.factor(RELIGION))) +
+  geom_jitter(color = "pink") +
+  stat_smooth(method = "lm", formula = "y ~ x", se = FALSE) +
+  scale_color_brewer(name = "Religion", palette = 1) +
+  labs(title = "Sentiment d'appartenance et discrimination en fonction de la religion", 
+       x = "Discrimination subie au cours des 12 derniers mois", 
+       y = "Sentiment d'appartenance au Quebec") +
+  theme_bw()
+print(vis_sex)
+
+```
+
+### Statut matrimonial
+
+```{r}
+
+vis_sex <- ggplot(Base_Trajipaq_Finale_Ok %>%
+                filter(!is.na(VD1), !is.na(VI4), !is.na(MATRIMONIAL)), aes(x = VI4, y = VD1, color = as.factor(MATRIMONIAL))) +
+  geom_jitter(color = "pink") +
+  stat_smooth(method = "lm", formula = "y ~ x", se = FALSE) +
+  scale_color_brewer(name = "Statut matrimonial", palette = 1) +
+  labs(title = "Sentiment d'appartenance et discrimination en fonction du statut matrimonial", 
+       x = "Discrimination subie au cours des 12 derniers mois", 
+       y = "Sentiment d'appartenance au Quebec") +
+  theme_bw()
+print(vis_sex)
+
+```
+
+## H2 : VD2 entretient une relation négative avec VI4
+
+### Modèle de régression
+
+```{r}
+
+Base_Trajipaq_Finale_Ok <-
+  Base_Trajipaq_Finale_Ok %>%
+  mutate(VD2_N = case_when(
+    VD2 == "Non" ~ 0,
+    VD2 == "Oui" ~ 1
+  ))
+
+```
+
+```{r}
+
+reg_2 <- glm(formula = VD2_N ~ VI4, family = "binomial", data = Base_Trajipaq_Finale_Ok)
+
+summary(reg_2)
+
+```
+
+### Coefficient de corrélation
+
+```{r}
+
+cor.test(Base_Trajipaq_Finale_Ok$VD2_N, Base_Trajipaq_Finale_Ok$VI4)
+
+```
+
+### Visualisation
+
+```{r}
+
+ggplot(Base_Trajipaq_Finale_Ok %>% filter(!is.na(VD2_N), !is.na(VI4)), aes(x = VI4, y = VD2_N)) +
+  stat_smooth(method = "glm",
+              formula = "y ~ x",
+              color = "pink", 
+              se = FALSE,
+              method.args = list(family = binomial)) +
+  labs(
+    title = "Sentiment d'appartenance et discrimination", 
+    x = "Discrimination subie au cours des 12 derniers mois", 
+    y = "Sentiment d'appartenance au Canada") +
+  theme_bw()
+
+```
+
+#### Sexe
+
+```{r}
+
+ggplot(Base_Trajipaq_Finale_Ok %>% filter(!is.na(VD2_N), !is.na(VI4)), aes(x = VI4, y = VD2_N)) +
+  stat_smooth(method = "glm",
+              formula = "y ~ x",
+              color = "pink", 
+              se = FALSE,
+              method.args = list(family = binomial)) +
+  labs(
+    title = "Sentiment d'appartenance et discrimination en fonction du sexe", 
+    x = "Discrimination subie au cours des 12 derniers mois", 
+    y = "Sentiment d'appartenance au Canada") +
+  facet_wrap(~ SEXE, nrow = 2) +
+  theme_bw()
+
+```
+
+#### Age
+
+```{r}
+
+ggplot(Base_Trajipaq_Finale_Ok %>% filter(!is.na(VD2_N), !is.na(VI4)), aes(x = VI4, y = VD2_N)) +
+  stat_smooth(method = "glm",
+              formula = "y ~ x",
+              color = "pink", 
+              se = FALSE,
+              method.args = list(family = binomial)) +
+  labs(
+    title = "Sentiment d'appartenance et discrimination en fonction de l'age", 
+    x = "Discrimination subie au cours des 12 derniers mois", 
+    y = "Sentiment d'appartenance au Canada") +
+  facet_wrap(~ AGE, nrow = 3) +
+  theme_bw()
+
+```
+
+#### Education
+
+```{r}
+
+ggplot(Base_Trajipaq_Finale_Ok %>% filter(!is.na(VD2_N), !is.na(VI4), !is.na(EDUCATION)), aes(x = VI4, y = VD2_N)) +
+  stat_smooth(method = "glm",
+              formula = "y ~ x",
+              color = "pink", 
+              se = FALSE,
+              method.args = list(family = binomial)) +
+  labs(
+    title = "Sentiment d'appartenance et discrimination en fonction de l'education", 
+    x = "Discrimination subie au cours des 12 derniers mois", 
+    y = "Sentiment d'appartenance au Canada") +
+  facet_wrap(~ EDUCATION, nrow = 3) +
+  theme_bw()
+
+```
+
+#### Origine
+
+```{r}
+
+ggplot(Base_Trajipaq_Finale_Ok %>% filter(!is.na(VD2_N), !is.na(VI4), !is.na(ORIGINE)), aes(x = VI4, y = VD2_N)) +
+  stat_smooth(method = "glm",
+              formula = "y ~ x",
+              color = "pink", 
+              se = FALSE,
+              method.args = list(family = binomial)) +
+  labs(
+    title = "Sentiment d'appartenance et discrimination en fonction de l'origine", 
+    x = "Discrimination subie au cours des 12 derniers mois", 
+    y = "Sentiment d'appartenance au Canada") +
+  facet_wrap(~ ORIGINE, nrow = 3) +
+  theme_bw()
+
+```
+
+#### Revenu
+
+```{r}
+
+ggplot(Base_Trajipaq_Finale_Ok %>% filter(!is.na(VD2_N), !is.na(VI4), !is.na(REVENU)), aes(x = VI4, y = VD2_N)) +
+  stat_smooth(method = "glm",
+              formula = "y ~ x",
+              color = "pink", 
+              se = FALSE,
+              method.args = list(family = binomial)) +
+  labs(
+    title = "Sentiment d'appartenance et discrimination en fonction du revenu", 
+    x = "Discrimination subie au cours des 12 derniers mois", 
+    y = "Sentiment d'appartenance au Canada") +
+  facet_wrap(~ REVENU, nrow = 3) +
+  theme_bw()
+
+```
+
+#### Langue maternelle
+
+```{r}
+
+ggplot(Base_Trajipaq_Finale_Ok %>% filter(!is.na(VD2_N), !is.na(VI4), !is.na(LANGUE)), aes(x = VI4, y = VD2_N)) +
+  stat_smooth(method = "glm",
+              formula = "y ~ x",
+              color = "pink", 
+              se = FALSE,
+              method.args = list(family = binomial)) +
+  labs(
+    title = "Sentiment d'appartenance et discrimination en fonction du revenu", 
+    x = "Discrimination subie au cours des 12 derniers mois", 
+    y = "Sentiment d'appartenance au Canada") +
+  facet_wrap(~ LANGUE, nrow = 2) +
+  theme_bw()
+
+```
+
+#### Religion
+
+```{r}
+
+ggplot(Base_Trajipaq_Finale_Ok %>% filter(!is.na(VD2_N), !is.na(VI4), !is.na(RELIGION)), aes(x = VI4, y = VD2_N)) +
+  stat_smooth(method = "glm",
+              formula = "y ~ x",
+              color = "pink", 
+              se = FALSE,
+              method.args = list(family = binomial)) +
+  labs(
+    title = "Sentiment d'appartenance et discrimination en fonction de la religion", 
+    x = "Discrimination subie au cours des 12 derniers mois", 
+    y = "Sentiment d'appartenance au Canada") +
+  facet_wrap(~ RELIGION, nrow = 3) +
+  theme_bw()
+
+```
+
+#### Statut matrimonial
+
+```{r}
+
+ggplot(Base_Trajipaq_Finale_Ok %>% filter(!is.na(VD2_N), !is.na(VI4), !is.na(MATRIMONIAL)), aes(x = VI4, y = VD2_N)) +
+  stat_smooth(method = "glm",
+              formula = "y ~ x",
+              color = "pink", 
+              se = FALSE,
+              method.args = list(family = binomial)) +
+  labs(
+    title = "Sentiment d'appartenance et discrimination en fonction du statut matrimonial", 
+    x = "Discrimination subie au cours des 12 derniers mois", 
+    y = "Sentiment d'appartenance au Canada") +
+  facet_wrap(~ MATRIMONIAL, nrow = 3) +
+  theme_bw()
+
+```
+
+## H3 : VD3 entretient une relation négative avec VI4
+
+### Modèle de régression
+
+```{r}
+
+Base_Trajipaq_Finale_Ok <-
+  Base_Trajipaq_Finale_Ok %>%
+  mutate(VD3_N = case_when(
+    VD3 == "Non" ~ 0,
+    VD3 == "Oui" ~ 1
+  ))
+
+```
+
+```{r}
+
+reg_3 <- glm(formula = VD3_N ~ VI4, family = "binomial", data = Base_Trajipaq_Finale_Ok)
+
+summary(reg_3)
+
+```
+
+## H4 : VD1 entretien une relation avec VD3 (en tant que variable indépendante)
+
+```{r}
+
+Base_Trajipaq_Finale_Ok <-
+  Base_Trajipaq_Finale_Ok %>%
+  mutate(VD1_F = case_when(
+    VD1 < 0.5 ~ "Non",
+    VD1 >= 0.5 ~ "Oui"
+  ))
+
+```
+
+```{r}
+
+ctable(Base_Trajipaq_Finale_Ok$VD3, Base_Trajipaq_Finale_Ok$VD1_F)
+
+```
+
+```{r}
+
+chisq.test(Base_Trajipaq_Finale_Ok$VD3, Base_Trajipaq_Finale_Ok$VD1_F)
+
+```

### 2023-05-24 Codes/Sentiment_an1.Rmd
Progression générale dans le travail.

@@
   mutate(AGE = case_when(
-    QAGE == 1 ~ "[0;19] ans",
-    QAGE == 2 ~ "[20;24] ans",
-    QAGE == 3 ~ "[25;29] ans",
-    QAGE == 4 ~ "[30;34] ans",
-    QAGE == 5 ~ "[35;39] ans",
-    QAGE == 6 ~ "[40;44] ans",
-    QAGE == 7 ~ "[45;49] ans",
-    QAGE == 8 ~ "[50;60] ans",
-    QAGE == 9 ~ "[61;99] ans",
+    QAGE == 1 ~ "[0;19]",
+    QAGE == 2 ~ "[20;24]",
+    QAGE == 3 ~ "[25;29]",
+    QAGE == 4 ~ "[30;34]",
+    QAGE == 5 ~ "[35;39]",
+    QAGE == 6 ~ "[40;44]",
+    QAGE == 7 ~ "[45;49]",
+    QAGE == 8 ~ "[50;60]",
+    QAGE == 9 ~ "[61;99]",
     QAGE == 99 ~ NA
@@
 Base_Trajipaq_Finale_Ok$AGE <- factor(Base_Trajipaq_Finale_Ok$AGE,levels = c(
-  "[0;19] ans",
-  "[20;24] ans",
-  "[25;29] ans",
-  "[30;34] ans",
-  "[35;39] ans",
-  "[40;44] ans",
-  "[45;49] ans",
-  "[50;60] ans",
-  "[61;99] ans",
+  "[0;19]",
+  "[20;24]",
+  "[25;29]",
+  "[30;34]",
+  "[35;39]",
+  "[40;44]",
+  "[45;49]",
+  "[50;60]",
+  "[61;99]",
   NA
@@
     title = "Distribution de l'age par groupes",
-    x = "Age"
+    x = "Age en annees"
        ) +
@@
 
+# Le modèle de régression traite les variables d'une manière inattendue.
+
 ```
@@
 
+# La relation est - de manière inattendue - positive.
+
 ```
@@
 
-vis_sex <- ggplot(Base_Trajipaq_Finale_Ok %>%
+vis_SEXE <- ggplot(Base_Trajipaq_Finale_Ok %>%
                 filter(!is.na(VD1), !is.na(VI4), !is.na(SEXE)), aes(x = VI4, y = VD1, color = SEXE)) +
@@
   theme_bw()
-print(vis_sex)
+print(vis_SEXE)
+
+# La relation semble plus forte chez les femmes avec cette méthode de visualisation.
+
+```
+
+```{r}
+
+vis_SEXE_2 <- ggplot(Base_Trajipaq_Finale_Ok %>%
+                filter(!is.na(VD1), !is.na(VI4), !is.na(SEXE)), aes(x = VI4, y = VD1)) +
+  geom_jitter(color = "pink") +
+  stat_smooth(method = "lm", 
+              formula = "y ~ x", 
+              se = FALSE,
+              color = "purple") +
+  labs(title = "Sentiment d'appartenance et discrimination en fonction du sexe", 
+       x = "Discrimination subie au cours des 12 derniers mois", 
+       y = "Sentiment d'appartenance au Quebec") +
+  facet_wrap(~ SEXE, nrow = 2) +
+  theme_bw()
+print(vis_SEXE_2)
+
+# La relation ne semble pas être affectée par le sexe avec cette méthode de visualisation.
 
@@
 
-vis_sex <- ggplot(Base_Trajipaq_Finale_Ok %>%
-                filter(!is.na(VD1), !is.na(VI4), !is.na(AGE)), aes(x = VI4, y = VD1, color = as.factor(AGE))) +
+vis_AGE <- ggplot(Base_Trajipaq_Finale_Ok %>%
+                filter(!is.na(VD1), !is.na(VI4), !is.na(AGE)), aes(x = VI4, y = VD1, color = AGE)) +
   geom_jitter(color = "pink") +
@@
   theme_bw()
-print(vis_sex)
-
-```
-
+print(vis_AGE)
+
+# La relation semble affectée par l'âge avec cette méthode de visualisation.
+
+```
+
+```{r}
+
+vis_AGE_2 <- ggplot(Base_Trajipaq_Finale_Ok %>%
+                filter(!is.na(VD1), !is.na(VI4), !is.na(AGE)), aes(x = VI4, y = VD1)) +
+  geom_jitter(color = "pink") +
+  stat_smooth(method = "lm", 
+              formula = "y ~ x", 
+              se = FALSE,
+              color = "purple") +
+  labs(title = "Sentiment d'appartenance et discrimination en fonction de l'age", 
+       x = "Discrimination subie au cours des 12 derniers mois", 
+       y = "Sentiment d'appartenance au Quebec") +
+  facet_wrap(~ AGE, nrow = 3) +
+  theme_bw()
+print(vis_AGE_2)
+
+# La relation semble également être affectée par l'âge avec cette méthode de visualisation.
+
+```
 #### Education
@@
 
-vis_sex <- ggplot(Base_Trajipaq_Finale_Ok %>%
-                filter(!is.na(VD1), !is.na(VI4), !is.na(EDUCATION)), aes(x = VI4, y = VD1, color = as.factor(EDUCATION))) +
+vis_EDUCATION <- ggplot(Base_Trajipaq_Finale_Ok %>%
+                filter(!is.na(VD1), !is.na(VI4), !is.na(EDUCATION)), aes(x = VI4, y = VD1, color = EDUCATION)) +
   geom_jitter(color = "pink") +
@@
   theme_bw()
-print(vis_sex)
+print(vis_EDUCATION)
+
+# La relation semble être affectée par le niveau d'éducation avec cette méthode de visualisation.
+
+```
+
+```{r}
+
+vis_EDUCATION_2 <- ggplot(Base_Trajipaq_Finale_Ok %>%
+                filter(!is.na(VD1), !is.na(VI4), !is.na(EDUCATION)), aes(x = VI4, y = VD1)) +
+  geom_jitter(color = "pink") +
+  stat_smooth(method = "lm", 
+              formula = "y ~ x", 
+              se = FALSE,
+              color = "purple") +
+  labs(title = "Sentiment d'appartenance et discrimination en fonction de l'education", 
+       x = "Discrimination subie au cours des 12 derniers mois", 
+       y = "Sentiment d'appartenance au Quebec") +
+  facet_wrap(~ EDUCATION, nrow = 3) +
+  theme_bw()
+print(vis_EDUCATION_2)
+
+# La relation ne semble pas être affectée par le niveau d'éducation avec cette méthode de visualisation.
 
@@
 
-vis_sex <- ggplot(Base_Trajipaq_Finale_Ok %>%
-                filter(!is.na(VD1), !is.na(VI4), !is.na(ORIGINE)), aes(x = VI4, y = VD1, color = as.factor(ORIGINE))) +
+vis_ORIGINE <- ggplot(Base_Trajipaq_Finale_Ok %>%
+                filter(!is.na(VD1), !is.na(VI4), !is.na(ORIGINE)), aes(x = VI4, y = VD1, color = ORIGINE)) +
   geom_jitter(color = "pink") +
@@
   theme_bw()
-print(vis_sex)
+print(vis_ORIGINE)
+
+# La relation semble être affectée par la région d'origine avec cette méthode de visualisation.
+
+```
+
+```{r}
+
+vis_ORIGINE_2 <- ggplot(Base_Trajipaq_Finale_Ok %>%
+                filter(!is.na(VD1), !is.na(VI4), !is.na(ORIGINE)), aes(x = VI4, y = VD1)) +
+  geom_jitter(color = "pink") +
+  stat_smooth(method = "lm", 
+              formula = "y ~ x", 
+              se = FALSE,
+              color = "purple") +
+  labs(title = "Sentiment d'appartenance et discrimination en fonction de l'origine", 
+       x = "Discrimination subie au cours des 12 derniers mois", 
+       y = "Sentiment d'appartenance au Quebec") +
+  facet_wrap(~ ORIGINE, nrow = 3) +
+  theme_bw()
+print(vis_ORIGINE_2)
+
+# La relation semble également être affectée avec cette méthode de visualisation.
 
@@
 
-vis_sex <- ggplot(Base_Trajipaq_Finale_Ok %>%
-                filter(!is.na(VD1), !is.na(VI4), !is.na(REVENU)), aes(x = VI4, y = VD1, color = as.factor(REVENU))) +
+vis_REVENU <- ggplot(Base_Trajipaq_Finale_Ok %>%
+                filter(!is.na(VD1), !is.na(VI4), !is.na(REVENU)), aes(x = VI4, y = VD1, color = REVENU)) +
   geom_jitter(color = "pink") +
@@
   theme_bw()
-print(vis_sex)
+print(vis_REVENU)
+
+# La relation semble être affectée par le revenu avec cette méthode de visualisation.
+
+```
+
+```{r}
+
+vis_REVENU_2 <- ggplot(Base_Trajipaq_Finale_Ok %>%
+                filter(!is.na(VD1), !is.na(VI4), !is.na(REVENU)), aes(x = VI4, y = VD1)) +
+  geom_jitter(color = "pink") +
+  stat_smooth(method = "lm", 
+              formula = "y ~ x", 
+              se = FALSE,
+              color = "purple") +
+  labs(title = "Sentiment d'appartenance et discrimination en fonction du revenu", 
+       x = "Discrimination subie au cours des 12 derniers mois", 
+       y = "Sentiment d'appartenance au Quebec") +
+  facet_wrap(~ REVENU, nrow = 3) +
+  theme_bw()
+print(vis_REVENU_2)
+
+# La relation semble également être affectée par le revenu avec cette méthode de visualisation.
 
@@
 
-vis_sex <- ggplot(Base_Trajipaq_Finale_Ok %>%
-                filter(!is.na(VD1), !is.na(VI4), !is.na(LANGUE)), aes(x = VI4, y = VD1, color = as.factor(LANGUE))) +
+vis_LANGUE <- ggplot(Base_Trajipaq_Finale_Ok %>%
+                filter(!is.na(VD1), !is.na(VI4), !is.na(LANGUE)), aes(x = VI4, y = VD1, color = LANGUE)) +
   geom_jitter(color = "pink") +
@@
   theme_bw()
-print(vis_sex)
+print(vis_LANGUE)
+
+# La relation semble être affectée par la langue maternelle avec cette méthode de visualisation.
+
+```
+
+```{r}
+
+vis_LANGUE_2 <- ggplot(Base_Trajipaq_Finale_Ok %>%
+                filter(!is.na(VD1), !is.na(VI4), !is.na(LANGUE)), aes(x = VI4, y = VD1)) +
+  geom_jitter(color = "pink") +
+  stat_smooth(method = "lm", 
+              formula = "y ~ x", 
+              se = FALSE,
+              color = "purple") +
+  labs(title = "Sentiment d'appartenance et discrimination en fonction de la langue maternelle", 
+       x = "Discrimination subie au cours des 12 derniers mois", 
+       y = "Sentiment d'appartenance au Quebec") +
+  facet_wrap(~ LANGUE, nrow = 2) +
+  theme_bw()
+print(vis_LANGUE_2)
+
+# La relation semble être également affectée par la langue maternelle avec cette méthode de visualisation.
 
@@
 
-vis_sex <- ggplot(Base_Trajipaq_Finale_Ok %>%
-                filter(!is.na(VD1), !is.na(VI4), !is.na(RELIGION)), aes(x = VI4, y = VD1, color = as.factor(RELIGION))) +
+vis_RELIGION <- ggplot(Base_Trajipaq_Finale_Ok %>%
+                filter(!is.na(VD1), !is.na(VI4), !is.na(RELIGION)), aes(x = VI4, y = VD1, color = RELIGION)) +
   geom_jitter(color = "pink") +
@@
   theme_bw()
-print(vis_sex)
+print(vis_RELIGION)
+
+# La relation semble être affectée par la religion avec cette méthode de visualisation.
+
+```
+
+```{r}
+
+vis_RELIGION_2 <- ggplot(Base_Trajipaq_Finale_Ok %>%
+                filter(!is.na(VD1), !is.na(VI4), !is.na(RELIGION)), aes(x = VI4, y = VD1)) +
+  geom_jitter(color = "pink") +
+  stat_smooth(method = "lm", 
+              formula = "y ~ x", 
+              se = FALSE,
+              color = "purple") +
+  labs(title = "Sentiment d'appartenance et discrimination en fonction de la religion", 
+       x = "Discrimination subie au cours des 12 derniers mois", 
+       y = "Sentiment d'appartenance au Quebec") +
+  facet_wrap(~ RELIGION, nrow = 3) +
+  theme_bw()
+print(vis_RELIGION_2)
+
+# La relation semble être également affectée par la religion avec cette méthode de visualisation.
 
@@
 
-vis_sex <- ggplot(Base_Trajipaq_Finale_Ok %>%
-                filter(!is.na(VD1), !is.na(VI4), !is.na(MATRIMONIAL)), aes(x = VI4, y = VD1, color = as.factor(MATRIMONIAL))) +
+vis_MATRIMONIAL <- ggplot(Base_Trajipaq_Finale_Ok %>%
+                filter(!is.na(VD1), !is.na(VI4), !is.na(MATRIMONIAL)), aes(x = VI4, y = VD1, color = MATRIMONIAL)) +
   geom_jitter(color = "pink") +
@@
   theme_bw()
-print(vis_sex)
+print(vis_MATRIMONIAL)
+
+# La relation semble être affectée par le statut matrimonial avec cette méthode de visualisation.
+
+```
+
+```{r}
+
+vis_MATRIMONIAL_2 <- ggplot(Base_Trajipaq_Finale_Ok %>%
+                filter(!is.na(VD1), !is.na(VI4), !is.na(MATRIMONIAL)), aes(x = VI4, y = VD1)) +
+  geom_jitter(color = "pink") +
+  stat_smooth(method = "lm", 
+              formula = "y ~ x", 
+              se = FALSE,
+              color = "purple") +
+  labs(title = "Sentiment d'appartenance et discrimination en fonction du statut matrimonial", 
+       x = "Discrimination subie au cours des 12 derniers mois", 
+       y = "Sentiment d'appartenance au Quebec") +
+  facet_wrap(~ MATRIMONIAL, nrow = 3) +
+  theme_bw()
+print(vis_MATRIMONIAL_2)
+
+# La relation semble être également affectée par le statut matrimonial avec cette méthode de visualisation.
 
@@
 
+# La relation semble être affectée par le sexe.
+
 ```
@@
 
+# La relation semble être affectée par l'âge.
+
 ```
@@
 
+# La relation semble être affectée par le niveau d'éducation.
+
 ```
@@
 
+# La relation semble être affectée par le pays d'origine.
+
 ```
@@
 
+# La relation semble être affectée par le revenu.
+
 ```
@@
 
+# La relation semble être affectée par la langue maternelle.
+
 ```
@@
 
+# La relation semble être affectée par la religion.
+
 ```
@@
   theme_bw()
+
+# La relation semble être affectée par le statut matrimonial.
 
@@
 
+# La relation n'est pas significative avec une chance d'erreur alpha égale à 5%.
+
 ```
@@
 
-```
+# La relation n'est pas significative avec une chance d'erreur alpha égale à 5%.
+
+```

### 2023-05-24 Tests/1. tests/Sentiment_test.Rmd
Progression générale dans le travail.

--- 2023-05-23 Codes/Article - Sentiment d'appartenance.Rmd
@@
 
-## Q10. Êtes-vous un homme, une femme ou autre ?
+## Sexe
+
+### Q10. Êtes-vous un homme, une femme ou autre ?
 
@@
 
-## Q8B A quel groupe d'âge appartenez-vous ?
+## Âge
+
+### Q8B A quel groupe d'âge appartenez-vous ?
 
@@
 
-## STATUT. Êtes-vous actuellement... ?
+## Statut matrimonial
+
+### STATUT. Êtes-vous actuellement... ?
 
@@
 
-## Q11. Dans quel pays êtes-vous né ?
+## Pays d'origine
+
+### Q11. Dans quel pays êtes-vous né ?
 
@@
 
-## Q4. Êtes-vous né au Canada ?
+## Lieu de naissance
+
+### Q4. Êtes-vous né au Canada ?
 
@@
 
-## Q15I. Statut migratoire avant d'obtenir la résidence permanente
+## Statut migratoire
+
+### Q15I. Statut migratoire avant d'obtenir la résidence permanente
 
@@
 
-## Q16. Quel est votre plus haut diplôme d’études obtenu ?
+## Niveau d'éducation
+
+### Q16. Quel est votre plus haut diplôme d’études obtenu ?
 
@@
 
-## Q17. Actuellement, avez-vous une religion ?
+## Religion
+
+### Q17. Actuellement, avez-vous une religion ?
 
@@
 
-## Q20. Quand vous étiez enfant, quelles étaient la langue ou les langues que vous parliez le plus souvent, à l’âge de 5 ans et que vous comprenez toujours? (Plusieurs réponses possibles).
+## Langue maternelle
+
+### Q20. Quand vous étiez enfant, quelles étaient la langue ou les langues que vous parliez le plus souvent, à l’âge de 5 ans et que vous comprenez toujours? (Plusieurs réponses possibles).
 
@@
 
-## I3. Indice de la compétence linguistique (français)
+## Français
+
+### I3. Indice de la compétence linguistique (français)
 
@@
 
-## I4. Indice de la compétence linguistique (anglais).
+## Anglais
+
+### I4. Indice de la compétence linguistique (anglais).
 
@@
 
-## Q23. Comment est votre état de santé en général ?
+## Etat de santé
+
+### Q23. Comment est votre état de santé en général ?
 
@@
 
-## I5. Indice de l'état de santé au cours des 12 derniers mois
+## Santé (12 derniers mois)
+
+### I5. Indice de l'état de santé au cours des 12 derniers mois
 
@@
 
-## Q26. Quel était le plus haut diplôme d’études obtenu par votre mère ?
+## Niveau d'éducation (mère)
+
+### Q26. Quel était le plus haut diplôme d’études obtenu par votre mère ?
 
@@
 
-## Q28. Quel était le plus haut diplôme d’études obtenu par votre père ?
+## Niveau d'éducation (père)
+
+### Q28. Quel était le plus haut diplôme d’études obtenu par votre père ?
 
@@
 
-# VD1 : sentiment d'appartenance au Québec
-
-## Q39A.
-Dites si vous êtes tout à fait d’accord, plutôt d’accord, plutôt pas d’accord ou pas du tout d’accord avec les propositions suivantes (participation identitaire et sentiment d’appartenance) : je me sens chez moi au Québec.
+# VD1 : sentiment d'appartenance (Québec)
 
@@
 
+```{r}
+
+v_VD1 <- c("Q39A",
+          "Q39B",
+          "Q39C",
+          "Q39D"
+)
+
+Base_Trajipaq_Finale_Ok <-
+  Base_Trajipaq_Finale_Ok %>%
+  rowwise() %>%
+  mutate(VD1 = (sum(c_across(all_of(v_VD1)), na.rm = TRUE) - length(v_VD1)) / (3*length(v_VD1))) %>%
+  mutate(VD1 = if (VD1 < 0) NA_real_ else VD1) %>%
+  ungroup()
+
+```
+
+```{r}
+
+descr(Base_Trajipaq_Finale_Ok$VD1)
+
+```
+
+```{r}
+
+ggplot(data = Base_Trajipaq_Finale_Ok, mapping = aes(x = VD1)) +
+  geom_histogram(na.rm = TRUE, fill = "pink", binwidth = 0.01) +
+  theme_bw()
+
+```
+
+La distribution est asymétrique vers la gauche. 
+L'indice moyen du sentiment d'appartenance au Québec est 0.53.
+
+# VD2 : sentiment d'appartenance (Québec, Canada et pays d'origine)
+
+## Q39A.
+Dites si vous êtes tout à fait d’accord, plutôt d’accord, plutôt pas d’accord ou pas du tout d’accord avec les propositions suivantes (participation identitaire et sentiment d’appartenance) : je me sens chez moi au Québec.
+
+```{r}
+
+Base_Trajipaq_Finale_Ok <-
+  Base_Trajipaq_Finale_Ok %>%
+  mutate(Q39A = case_when(
+    Q39A == 1 ~ 4,
+    Q39A == 2 ~ 3,
+    Q39A == 3 ~ 2,
+    Q39A == 4 ~ 1,
+    Q39A == 98 ~ NA,
+    Q39A == 99 ~ NA
+  ))
+
+```
+
+```{r}
+
+freq(Base_Trajipaq_Finale_Ok$Q39A)
+
+```
+
+```{r}
+
+ggplot(data = Base_Trajipaq_Finale_Ok, mapping = aes(x = Q39A)) +
+  geom_bar(fill = "pink",
+           na.rm = TRUE
+           ) +
+  theme_bw()
+
+```
+
+La distribution est asymétrique vers la gauche. 
+87.01% des personnes interrogées sont tout à fait d'accord ou plutôt d'accord pour dire qu'elles se sentent chez elles au Québec.
+
+## Q39B.
+Dites si vous êtes tout à fait d’accord, plutôt d’accord, plutôt pas d’accord ou pas du tout d’accord avec les propositions suivantes (participation identitaire et sentiment d’appartenance) : il faut faire oublier ses origines pour se faire accepter au Québec.
+
+```{r}
+
+Base_Trajipaq_Finale_Ok <-
+  Base_Trajipaq_Finale_Ok %>%
+  mutate(Q39B = case_when(
+    Q39B == 1 ~ 4,
+    Q39B == 2 ~ 3,
+    Q39B == 3 ~ 2,
+    Q39B == 4 ~ 1,
+    Q39B == 98 ~ NA,
+    Q39B == 99 ~ NA
+  ))
+
+```
+
+```{r}
+
+freq(Base_Trajipaq_Finale_Ok$Q39B)
+
+```
+
+```{r}
+
+ggplot(data = Base_Trajipaq_Finale_Ok, mapping = aes(x = Q39B)) +
+  geom_bar(fill = "pink",
+           na.rm = TRUE
+           ) +
+  theme_bw()
+
+```
+
+La distribution est asymétrique vers la droite.
+67.88% des personnes interrogées ne sont pas du tout d'accord ou plutôt pas d'accord pour dire qu'il faut faire oublier ses origines pour se faire accepter au Québec.
+
+## Q39C.
+Dites si vous êtes tout à fait d’accord, plutôt d’accord, plutôt pas d’accord ou pas du tout d’accord avec les propositions suivantes (participation identitaire et sentiment d’appartenance) : on me voit comme une Québécoise.
+
+```{r}
+
+Base_Trajipaq_Finale_Ok <-
+  Base_Trajipaq_Finale_Ok %>%
+  mutate(Q39C = case_when(
+    Q39C == 1 ~ 4,
+    Q39C == 2 ~ 3,
+    Q39C == 3 ~ 2,
+    Q39C == 4 ~ 1,
+    Q39C == 98 ~ NA,
+    Q39C == 99 ~ NA
+  ))
+
+```
+
+```{r}
+
+freq(Base_Trajipaq_Finale_Ok$Q39C)
+
+```
+
+```{r}
+
+ggplot(data = Base_Trajipaq_Finale_Ok, mapping = aes(x = Q39C)) +
+  geom_bar(fill = "pink",
+           na.rm = TRUE
+           ) +
+  theme_bw()
+
+```
+
+Les réponses modales sont "tout à fait d'accord" (29.36%) et "pas du tout d'accord" (29.97%).
+
+## Q39D.
+Dites si vous êtes tout à fait d’accord, plutôt d’accord, plutôt pas d’accord ou pas du tout d’accord avec les propositions suivantes (participation identitaire et sentiment d’appartenance) : je me sens Québécois(e).
+
+```{r}
+
+Base_Trajipaq_Finale_Ok <-
+  Base_Trajipaq_Finale_Ok %>%
+  mutate(Q39D = case_when(
+    Q39D == 1 ~ 4,
+    Q39D == 2 ~ 3,
+    Q39D == 3 ~ 2,
+    Q39D == 4 ~ 1,
+    Q39D == 98 ~ NA,
+    Q39D == 99 ~ NA
+  ))
+
+```
+
+```{r}
+
+freq(Base_Trajipaq_Finale_Ok$Q39D)
+
+```
+
+```{r}
+
+ggplot(data = Base_Trajipaq_Finale_Ok, mapping = aes(x = Q39D)) +
+  geom_bar(fill = "pink",
+           na.rm = TRUE
+           ) +
+  theme_bw()
+
+```
+
+La distribution est asymétrique vers la gauche. 
+60.72% des personnes interrogées sont tout à fait d'accord ou plutôt d'accord pour dire qu'elles se sentent Québécois(e)s.
+
 ## Q39E.
@@
 
-## VD1 : L'indice du sentiment d'appartenance au Québec
-
 A la différence des variables précédentes, l'indice du sentiment d'appartenance est une variable numérique continue dont le niveau de mesure est ratio.
@@
 
-v_VD1 <- c("Q39A",
+v_VD2 <- c("Q39A",
           "Q39B",
@@
   rowwise() %>%
-  mutate(VD1 = (sum(c_across(all_of(v_VD1)), na.rm = TRUE) - length(v_VD1)) / (3*length(v_VD1))) %>%
-  mutate(VD1 = if (VD1 < 0) NA_real_ else VD1) %>%
+  mutate(VD2 = (sum(c_across(all_of(v_VD2)), na.rm = TRUE) - length(v_VD2)) / (3*length(v_VD2))) %>%
+  mutate(VD2 = if (VD2 < 0) NA_real_ else VD2) %>%
   ungroup()
@@
 
-descr(Base_Trajipaq_Finale_Ok$VD1)
-
-```
-
-```{r}
-
-ggplot(data = Base_Trajipaq_Finale_Ok, mapping = aes(x = VD1)) +
+descr(Base_Trajipaq_Finale_Ok$VD2)
+
+```
+
+```{r}
+
+ggplot(data = Base_Trajipaq_Finale_Ok, mapping = aes(x = VD2)) +
+  geom_histogram(na.rm = TRUE, fill = "pink", binwidth = 0.01) +
+  theme_bw()
+
+```
+
+La distribution est asymétrique vers la gauche. 
+L'indice moyen du sentiment d'appartenance au Québec est 0.59.
+A la différence des variables précédentes, l'indice du sentiment d'appartenance est une variable numérique continue dont le niveau de mesure est ratio.
+
+```{r}
+
+v_VD2 <- c("Q39A",
+          "Q39B",
+          "Q39C",
+          "Q39D",
+          "Q39E"
+          )
+
+Base_Trajipaq_Finale_Ok <-
+  Base_Trajipaq_Finale_Ok %>%
+  rowwise() %>%
+  mutate(VD2 = (sum(c_across(all_of(v_VD2)), na.rm = TRUE) - length(v_VD2)) / (3*length(v_VD2))) %>%
+  mutate(VD2 = if (VD2 < 0) NA_real_ else VD2) %>%
+  ungroup()
+
+```
+
+```{r}
+
+descr(Base_Trajipaq_Finale_Ok$VD2)
+
+```
+
+```{r}
+
+ggplot(data = Base_Trajipaq_Finale_Ok, mapping = aes(x = VD2)) +
   geom_histogram(na.rm = TRUE, fill = "pink", binwidth = 0.01) +
@@
 
-## VI1 + VD1
+## Q39E + Q39F
+
+```{r}
+
+Base_Trajipaq_Finale_Ok <-
+  Base_Trajipaq_Finale_Ok %>%
+  mutate(Q39E_N = case_when(
+    Q39E == 1 | Q39E == 2 ~ "Non",
+    Q39E == 3 | Q39E == 4 ~ "Oui"
+  ))
+
+Base_Trajipaq_Finale_Ok <-
+  Base_Trajipaq_Finale_Ok %>%
+  mutate(Q39F_N = case_when(
+    Q39E == 1 | Q39E == 2 ~ "Non",
+    Q39E == 3 | Q39E == 4 ~ "Oui"
+  ))
+
+```
+
+```{r}
+
+ctable(x = Base_Trajipaq_Finale_Ok$Q39E_N, # Sentiment d'appartenance au Canada
+       y =  Base_Trajipaq_Finale_Ok$Q39F_N # Sentiment d’appartenance au pays d'origine
+       )
+
+```
+
+## VI2 + VD2
 
@@
   Base_Trajipaq_Finale_Ok %>%
-  mutate(VD1_F = case_when(
-    VD1 < 0.25 ~ "Faible",
-    VD1 >= 0.25 & VD1 < 0.5 ~ "Moyen",
-    VD1 >= 0.5 & VD1 < 0.75 ~ "Fort",
-    VD1 >= 0.75 & VD1 <= 1 ~ "Tres fort",
-  ))
-
-Base_Trajipaq_Finale_Ok$VD1_F <- factor(Base_Trajipaq_Finale_Ok$VD1_F, levels = c("Tres fort",
+  mutate(VD2_F = case_when(
+    VD2 < 0.25 ~ "Faible",
+    VD2 >= 0.25 & VD2 < 0.5 ~ "Moyen",
+    VD2 >= 0.5 & VD2 < 0.75 ~ "Fort",
+    VD2 >= 0.75 & VD2 <= 1 ~ "Tres fort",
+  ))
+
+Base_Trajipaq_Finale_Ok$VD2_F <- factor(Base_Trajipaq_Finale_Ok$VD2_F, levels = c("Tres fort",
                                                                                 "Fort",
@@
 
-ctable(x = Base_Trajipaq_Finale_Ok$VI1_F, y = Base_Trajipaq_Finale_Ok$VD1_F)
-
-```
-
-```{r}
-
-ggplot(data = Base_Trajipaq_Finale_Ok, mapping = aes(x = VI1_F, fill = VD1_F)) +
+ctable(x = Base_Trajipaq_Finale_Ok$VI1_F, y = Base_Trajipaq_Finale_Ok$VD2_F)
+
+```
+
+```{r}
+
+ggplot(data = Base_Trajipaq_Finale_Ok, mapping = aes(x = VI1_F, fill = VD2_F)) +
   geom_bar(position = "dodge", na.rm = TRUE) +
@@
 
-chisq.test(x = Base_Trajipaq_Finale_Ok$VI1_F, y = Base_Trajipaq_Finale_Ok$VD1_F)
+chisq.test(x = Base_Trajipaq_Finale_Ok$VI1_F, y = Base_Trajipaq_Finale_Ok$VD2_F)
 
@@
 
-reg_1 <- lm(Base_Trajipaq_Finale_Ok$VI1 ~ Base_Trajipaq_Finale_Ok$VD1 +
+reg_1 <- lm(Base_Trajipaq_Finale_Ok$VI1 ~ Base_Trajipaq_Finale_Ok$VD2 +
               Base_Trajipaq_Finale_Ok$Q10 +     # Sexe
@@
 
-cor.test(Base_Trajipaq_Finale_Ok$VI1, Base_Trajipaq_Finale_Ok$VD1)
+cor.test(Base_Trajipaq_Finale_Ok$VI1, Base_Trajipaq_Finale_Ok$VD2)
 
@@
 vis <- ggplot(Base_Trajipaq_Finale_Ok %>%
-                filter(!is.na(VD1), !is.na(VI1)), aes(x = VI1, y = VD1)) +
+                filter(!is.na(VD2), !is.na(VI1)), aes(x = VI1, y = VD2)) +
   geom_jitter(color = "pink") +
@@
 vis_sex <- ggplot(Base_Trajipaq_Finale_Ok %>%
-                filter(!is.na(VD1), !is.na(VI1), !is.na(Q10)), aes(x = VI1, y = VD1, color = as.factor(Q10))) +
+                filter(!is.na(VD2), !is.na(VI1), !is.na(Q10)), aes(x = VI1, y = VD2, color = as.factor(Q10))) +
   geom_jitter(color = "pink") +
@@
 vis_age <- ggplot(Base_Trajipaq_Finale_Ok %>%
-                filter(!is.na(VD1), !is.na(VI1), !is.na(Q8A_F)), aes(x = VI1, y = VD1, color = Q8A_F)) +
+                filter(!is.na(VD2), !is.na(VI1), !is.na(Q8A_F)), aes(x = VI1, y = VD2, color = Q8A_F)) +
   geom_jitter(color = "pink") +
@@
 vis_born <- ggplot(Base_Trajipaq_Finale_Ok %>%
-                filter(!is.na(VD1), !is.na(VI1), !is.na(STATUT_F)), aes(x = VI1, y = VD1, color = STATUT_F)) +
+                filter(!is.na(VD2), !is.na(VI1), !is.na(STATUT_F)), aes(x = VI1, y = VD2, color = STATUT_F)) +
   geom_jitter(color = "pink") +
@@
 vis_born <- ggplot(Base_Trajipaq_Finale_Ok %>%
-                filter(!is.na(VD1), !is.na(VI1), !is.na(Q4)), aes(x = VI1, y = VD1, color = as.factor(Q4))) +
+                filter(!is.na(VD2), !is.na(VI1), !is.na(Q4)), aes(x = VI1, y = VD2, color = as.factor(Q4))) +
   geom_jitter(color = "pink") +
@@
 vis_educ <- ggplot(Base_Trajipaq_Finale_Ok %>%
-                filter(!is.na(VD1), !is.na(VI1), !is.na(Q16)), aes(x = VI1, y = VD1, color = as.factor(Q16))) +
+                filter(!is.na(VD2), !is.na(VI1), !is.na(Q16)), aes(x = VI1, y = VD2, color = as.factor(Q16))) +
   geom_jitter(color = "pink") +
@@
 vis_relig <- ggplot(Base_Trajipaq_Finale_Ok %>%
-                filter(!is.na(VD1), !is.na(VI1), !is.na(Q17_F)), aes(x = VI1, y = VD1, color = Q17_F)) +
+                filter(!is.na(VD2), !is.na(VI1), !is.na(Q17_F)), aes(x = VI1, y = VD2, color = Q17_F)) +
   geom_jitter(color = "pink") +
@@
 vis_lang <- ggplot(Base_Trajipaq_Finale_Ok %>%
-                filter(!is.na(VD1), !is.na(VI1), !is.na(Q20_F)), aes(x = VI1, y = VD1, color = Q20_F)) +
+                filter(!is.na(VD2), !is.na(VI1), !is.na(Q20_F)), aes(x = VI1, y = VD2, color = Q20_F)) +
   geom_jitter(color = "pink") +
@@
 vis_health <- ggplot(Base_Trajipaq_Finale_Ok %>%
-                filter(!is.na(VD1), !is.na(VI1), !is.na(Q23)), aes(x = VI1, y = VD1, color = as.factor(Q23))) +
+                filter(!is.na(VD2), !is.na(VI1), !is.na(Q23)), aes(x = VI1, y = VD2, color = as.factor(Q23))) +
   geom_jitter(color = "pink") +

### 2023-05-25 Codes/Sentiment_an1.Rmd
Ajout de la variable "année d'arrivée au Québec".

@@
   coord_flip() +
+  theme_bw()
+
+```
+
+### Année d'arrivée au Québec
+
+```{r}
+
+Base_Trajipaq_Finale_Ok <-
+  Base_Trajipaq_Finale_Ok %>%
+  mutate(Q12 = case_when(
+    Q12 == 9998 ~ NA_real_,
+    Q12 == 9999 ~ NA_real_,
+    TRUE ~ as.numeric(Q12)
+  ))
+
+Base_Trajipaq_Finale_Ok <- 
+  Base_Trajipaq_Finale_Ok %>%
+  mutate(ARRIVEE = case_when(
+    as.numeric(Q14I) == 9996 ~ as.numeric(Q12),
+    as.numeric(Q14I) == 9998 ~ NA_real_,
+    as.numeric(Q14I) == 9999 ~ NA_real_,
+    TRUE ~ as.numeric(Q14I)
+  ))
+
+```
+
+```{r}
+
+descr(Base_Trajipaq_Finale_Ok$ARRIVEE)
+
+```
+
+```{r}
+
+ggplot(data = Base_Trajipaq_Finale_Ok, mapping = aes(x = ARRIVEE)) +
+  geom_boxplot(na.rm = TRUE,
+               fill = "pink") +
+  labs(title = "Distribution de l'annee d'arrivee au Quebec",
+       x = "Annee d'arrivee au Quebec") +
   theme_bw()
@@
               Base_Trajipaq_Finale_Ok$EDUCATION +
+              Base_Trajipaq_Finale_Ok$ARRIVEE +
               Base_Trajipaq_Finale_Ok$ORIGINE +
@@
 
+#### Année d'arrivée au Québec
+
+```{r}
+
+Base_Trajipaq_Finale_Ok <-
+  Base_Trajipaq_Finale_Ok %>%
+  mutate(ARRIVEE_C = case_when(
+    ARRIVEE >= 1990 & ARRIVEE < 2000 ~ "[1990;1999]",
+    ARRIVEE >= 2000 & ARRIVEE < 2005 ~ "[2000;2004]",
+    ARRIVEE >= 2005 & ARRIVEE < 2010 ~ "[2005;2009]",
+    ARRIVEE >= 2010 & ARRIVEE < 2015 ~ "[2010;2014]",
+    ARRIVEE >= 2015 & ARRIVEE < 2020 ~ "[2015;2020]",
+  ))
+
+```
+
+```{r}
+
+vis_ARRIVEE <- ggplot(Base_Trajipaq_Finale_Ok %>%
+                filter(!is.na(VD1), !is.na(VI4), !is.na(ARRIVEE_C)), aes(x = VI4, y = VD1, color = as.factor(ARRIVEE_C))) +
+  geom_jitter(color = "pink") +
+  stat_smooth(method = "lm", formula = "y ~ x", se = FALSE) +
+  scale_color_brewer(name = "Annee d'arrivee au Quebec", palette = 1) +
+  labs(title = "Sentiment d'appartenance et discrimination en fonction de l'annee d'arrivee", 
+       x = "Discrimination subie au cours des 12 derniers mois", 
+       y = "Sentiment d'appartenance au Quebec") +
+  theme_bw()
+print(vis_ARRIVEE)
+
+# La relation semble être affectée par l'année d'arrivée au Québec avec cette méthode de visualisation.
+
+```
+
+```{r}
+
+vis_ARRIVEE_2 <- ggplot(Base_Trajipaq_Finale_Ok %>%
+                filter(!is.na(VD1), !is.na(VI4), !is.na(ARRIVEE_C)), aes(x = VI4, y = VD1)) +
+  geom_jitter(color = "pink") +
+  stat_smooth(method = "lm", 
+              formula = "y ~ x", 
+              se = FALSE,
+              color = "purple") +
+  labs(title = "Sentiment d'appartenance et discrimination en fonction de l'annee d'arrivee", 
+       x = "Discrimination subie au cours des 12 derniers mois", 
+       y = "Sentiment d'appartenance au Quebec") +
+  facet_wrap(~ ARRIVEE_C, nrow = 3) +
+  theme_bw()
+print(vis_ARRIVEE_2)
+
+# La relation semble également affectée par l'année d'arrivée au Québec avec cette méthode de visualisation.
+
+```
+
 #### Origine
@@
 
+#### Année d'arrivée au Québec
+
+```{r}
+
+ggplot(Base_Trajipaq_Finale_Ok %>% filter(!is.na(VD2_N), !is.na(VI4), !is.na(ARRIVEE_C)), aes(x = VI4, y = VD2_N)) +
+  stat_smooth(method = "glm",
+              formula = "y ~ x",
+              color = "pink", 
+              se = FALSE,
+              method.args = list(family = binomial)) +
+  labs(
+    title = "Sentiment d'appartenance et discrimination en fonction de l'annee d'arivee", 
+    x = "Discrimination subie au cours des 12 derniers mois", 
+    y = "Sentiment d'appartenance au Canada") +
+  facet_wrap(~ ARRIVEE_C, nrow = 3) +
+  theme_bw()
+
+# La relation semble être affectée par le niveau d'éducation.
+
+```
+
 #### Origine

### 2023-05-29 Article-Reseaux sociaux_sentiment.Rmd
Préparation du début de la rédaction.

@@
 ---
+
+# Résumé
+
+```{r}
+
+# Partie à rédiger en dernier.
+
+```
+
+# Introduction
+
+```{r}
+
+# Partie à rédiger après la discussion des résultats.
+
+# Question de recherche = comment les sentiments d'appartenances sont ils associés au pays d'origine et aux situations de discriminations vécues par les immigrants établis au Québec.
+
+```
+
+# Revue de littérature
+
+```{r}
+
+# Partie à rédiger en priorité.
+
+```
+
+## Intégration culturelle
+
+## Sentiments d'appartenance
+
+## Situations de discriminations vécues par les immigrants
+
+## Sentiments d'appartenance et discriminations
+
+# Méthodologie
+
+```{r}
+
+# Partie à rédiger en priorité.
+
+```
+
+## Présentation de l'échantillon
+
+## Présentation des variables
+
+```{r}
+
+# VD1 = indice du sentiment d'appartenance au Québec
+
+# VD2 = sentiment d'appartenance au Canada
+
+# VI1 = région d'origine
+
+# VI2 = indice de la distance culturelle du pays d'origine avec le Canada
+
+# VI3 = indice de la stabilité politique du pays d'origine par rapport à celle du Canada
+
+# VI4 = indice des situations de discriminations subies
+
+```
+
+## Hypothèses de recherche
+
+```{r}
+
+# H1 = la région d'origine des immigrants (VI1) joue un rôle dans le développement du sentiment d'appartenance au Québec (VD1)
+
+# H2 = la région d'origine des immigrants (VI1) joue un rôle dans le développement du sentiment d'appartenance au Canada (VD2)
+
+# H3 = la distance culturelle du pays d'origine des immigrants avec le Canada (VI2) joue un rôle dans le développement du sentiment d'appartenance au Québec (VD1)
+
+# H4 = la distance culturelle du pays d'origine des immigrants avec le Canada (VI2) joue un rôle dans le développement du sentiment d'appartenance au Canada (VD2)
+
+# H5 = la stabilité politique du pays d'origine des immigrants par rapport à celle du Canada (VI3) joue un rôle dans le développement du sentiment d'appartenance au Québec (VD1)
+
+# H6 = la stabilité politique du pays d'origine des immigrants par rapport à celle du Canada (VI3) joue un rôle dans le développement du sentiment d'appartenance au Canada (VD2)
+
+# H7 = les situations de discriminations vécues par les immigrants (VI4) réduisent le développement du sentiment d'appartenance au Québec (VD1)
+
+# H8 = les situations de discriminations vécues par les immigrants (VI4) réduisent le développement du sentiment d'appartenance au Canada (VD2)
+
+```
+
+# Résultats
+
+```{r}
+
+# Partie à rédiger après la revue de littérature et la méthodologie.
+
+```
+
+## Présentation des résultats liés à la première hypothèse
+
+```{r}
+
+# Trouver un titre évocateur après la rédaction.
+
+```
+
+## Présentation des résultats liés à la deuxième hypothèse
+
+```{r}
+
+# Trouver un titre évocateur après la rédaction.
+
+```
+
+## Présentation des résultats liés à la troisième hypothèse
+
+```{r}
+
+# Trouver un titre évocateur après la rédaction.
+
+```
+
+# Discussion
+
+```{r}
+
+# Partie à rédiger après la présentation des résultats.
+
+```
+
+## Discussion des résultats liés à la première hypothèse
+
+```{r}
+
+# Trouver un titre évocateur après la rédaction.
+
+```
+
+## Discussion des résultats liés à la deuxième hypothèse
+
+```{r}
+
+# Trouver un titre évocateur après la rédaction.
+
+```
+
+## Discussion des résultats liés à la troisième hypothèse
+
+```{r}
+
+# Trouver un titre évocateur après la rédaction.
+
+```
+
+## Discussion des limites de l'étude
+
+# Conclusion
+
+```{r}
+
+# Partie à réaliser après la discussion des résultats
+
+```
+
+# Bibliographie
+
+```{r}
+
+# Partie générée automatiquement, mais qui doit tout de même être relue à la fin du travail de rédaction.
+
+```

### 2023-05-29 Article/Visseho_al_sentiment_appartenance.Rmd
Suppression du dossier "Tests" et modification du fichier "Vissého_al_sentiment_appartenance.Rmd".
Préparation du début de la rédaction.

@@
 Les immigrants sont souvent pris entre deux ou plusieurs cultures, celles de leur pays d'origine et celles de leur nouveau pays de résidence.
-Cet article essaie de comprendre comment ces sentiments sont associés à leur pays d'origine et aux éventuelles situations de discriminations vécues qu'ils vivent dans leur pays d'accueil.
+Cet article essaie de comprendre comment ces sentiments sont associés à leur pays d'origine et aux éventuelles situations de discriminations qu'ils vivent dans leur pays d'accueil.
 
@@
 
-#### VD2. Double sentiment d'appartenance (Canada/Pays d'origine)
+#### VD2. Sentiment d'appartenance au Canada
 
 -   Q39E (facteur/nominal)
-
--   Q39F (facteur/nominal)
-
--   VD2 (facteur/nominal)
 
@@
 
-#### VD3. Double sentiment d'appartenance (Canada/Québec)
+#### VD3. Sentiment d'appartenance au pays d'origine
 
--   Q39D (facteur/nominal)
-
--   Q39E (facteur/nominal)
+-   Q39F (facteur/nominal)
 

### 2023-05-30 Codes/Sentiment_an1.Rmd
ajout Q14I
Q14I
Update

@@
 ```{r}
+
 library(summarytools)
@@
 library(haven)
-```
-
-```{r}
+
+```
+
+```{r}
+
 Base_Trajipaq_Finale_Ok <- read_dta("~/Desktop/Base_Trajipaq_Finale_Ok.dta")
+
 ```
@@
 
+### Type d'intégration culturelle
+
+```{r}
+
+Base_Trajipaq_Finale_Ok <-
+  Base_Trajipaq_Finale_Ok %>%
+  mutate(INTEGRATION = case_when(
+    VD1_F == "Non" & VD3 == "Non" ~ "Marginal",
+    VD1_F == "Oui" & VD3 == "Non" ~ "Assimile",
+    VD1_F == "Non" & VD3 == "Oui" ~ "Separe",
+    VD1_F == "Oui" & VD3 == "Oui" ~ "Integre"
+  ))
+
+```
+
+```{r}
+
+freq(Base_Trajipaq_Finale_Ok$INTEGRATION)
+
+```
+
+```{r}
+
+ggplot(subset(Base_Trajipaq_Finale_Ok, !is.na(INTEGRATION)), aes(x = INTEGRATION)) +
+  geom_bar(fill = "pink") +
+  labs(
+    title = "Distribution des types d'intégration",
+    x = "Types d'intégration au Quebec"
+       ) +
+  coord_flip() +
+  theme_bw()
+
+```
+
 ## Variables dépendantes
@@
               Base_Trajipaq_Finale_Ok$RELIGION +
-              Base_Trajipaq_Finale_Ok$MATRIMONIAL
+              Base_Trajipaq_Finale_Ok$MATRIMONIAL +
+              Base_Trajipaq_Finale_Ok$INTEGRATION
             )
@@
 
+### Type d'intégration
+
+```{r}
+
+vis_INTEGRATION <- ggplot(Base_Trajipaq_Finale_Ok %>%
+                filter(!is.na(VD1), !is.na(VI4), !is.na(INTEGRATION)), aes(x = VI4, y = VD1, color = INTEGRATION)) +
+  geom_jitter(color = "pink") +
+  stat_smooth(method = "lm", formula = "y ~ x", se = FALSE) +
+  scale_color_brewer(name = "Type d'integration", palette = 1) +
+  labs(title = "Sentiment d'appartenance et discrimination en fonction du statut matrimonial", 
+       x = "Discrimination subie au cours des 12 derniers mois", 
+       y = "Sentiment d'appartenance au Quebec") +
+  theme_bw()
+print(vis_INTEGRATION)
+
+# La relation semble être affectée par le type d'intégration avec cette méthode de visualisation.
+
+```
+
+```{r}
+
+vis_INTEGRATION_2 <- ggplot(Base_Trajipaq_Finale_Ok %>%
+                filter(!is.na(VD1), !is.na(VI4), !is.na(INTEGRATION)), aes(x = VI4, y = VD1)) +
+  geom_jitter(color = "pink") +
+  stat_smooth(method = "lm", 
+              formula = "y ~ x", 
+              se = FALSE,
+              color = "purple") +
+  labs(title = "Sentiment d'appartenance et discrimination en fonction du type d'integration", 
+       x = "Discrimination subie au cours des 12 derniers mois", 
+       y = "Sentiment d'appartenance au Quebec") +
+  facet_wrap(~ INTEGRATION, nrow = 3) +
+  theme_bw()
+print(vis_INTEGRATION_2)
+
+# La relation semble être également affectée par le type d'intégration avec cette méthode de visualisation.
+
+```
+
 ### Statut matrimonial

### 2023-05-30 Codes/Sentiment_an_2.Rmd
Update

---
title: "Sentiment d'appartenance des immigrants"
author: "Thea"
date: "2023-05-24"
output: html_document
---

# Résumé

Les immigrants sont souvent pris entre deux ou plusieurs cultures, celles de leur pays d'origine et celles de leur nouveau pays de résidence. Cet article essaie de comprendre comment ces sentiments sont associés à leur pays d'origine et aux éventuelles situations de discriminations qu'ils vivent dans leur pays d'accueil.

# Création des variables

```{r}

library(summarytools)
library(tidyverse)
library(ggthemes)
library(haven)

trajipaq <- read_dta("../Data/Base_Trajipaq_Finale_Ok.dta")

```

### Sexe

```{r}

# Q10 = êtes-vous un homme, une femme ou autre (précisez) ?

trajipaq <-
  trajipaq %>%
  mutate(sexe = case_when(
    Q10 == 1 ~ "Homme",
    Q10 == 2 ~ "Femme",
    Q10 == 97 ~ "Autre"
  ))

freq(trajipaq$sexe)

```

### Âge

```{r}

# Qage = à quel groupe d'âge appartenez-vous ?

freq(trajipaq$QAGE)

trajipaq <-
  trajipaq %>%
  mutate(age = case_when(
    QAGE == 1 ~ "[0;19]",
    QAGE == 2 ~ "[20;24]",
    QAGE == 3 ~ "[25;29]",
    QAGE == 4 ~ "[30;34]",
    QAGE == 5 ~ "[35;39]",
    QAGE == 6 ~ "[40;44]",
    QAGE == 7 ~ "[45;49]",
    QAGE == 8 ~ "[50;60]",
    QAGE == 9 ~ "[61;99]",
    QAGE == 99 ~ NA
  ))

trajipaq$age <- factor(trajipaq$age,levels = c(
  "[0;19]",
  "[20;24]",
  "[25;29]",
  "[30;34]",
  "[35;39]",
  "[40;44]",
  "[45;49]",
  "[50;60]",
  "[61;99]",
  NA
), ordered = TRUE)

freq(trajipaq$age)

```

### Education

```{r}

trajipaq <-
  trajipaq %>%
  mutate(EDUCATION = factor(case_when(
    Q26 == 1 ~ "Elementaire",
    Q26 == 2 ~ "Secondaire",
    Q26 == 3 ~ "Collegial",
    Q26 == 4 ~ "Post secondaire",
    Q26 == 5 ~ "Universitaire",
    Q26 == 6 ~ NA,   ## Q26 >= 6 ~ NA
    Q26 == 98 ~ NA,
    Q26 == 99 ~ NA
  ), levels = c("Elementaire", "Secondaire", "Post secondaire", "Collegial", "Universitaire")))

```

### Année d'arrivée au Québec

```{r}

trajipaq <-
  trajipaq %>%
  mutate(Q12 = case_when(
    Q12 == 9998 ~ NA_real_,
    Q12 == 9999 ~ NA_real_,
    TRUE ~ as.numeric(Q12)
  ))

trajipaq <- 
  trajipaq %>%
  mutate(ARRIVEE = case_when(
    as.numeric(Q14I) == 9996 ~ as.numeric(Q12),
    as.numeric(Q14I) == 9998 ~ NA_real_,
    as.numeric(Q14I) == 9999 ~ NA_real_,
    TRUE ~ as.numeric(Q14I)
  ))

```

```{r}

descr(trajipaq$ARRIVEE)

```

```{r}

ggplot(data = trajipaq, mapping = aes(x = ARRIVEE)) +
  geom_boxplot(na.rm = TRUE,
               fill = "pink") +
  labs(title = "Distribution de l'annee d'arrivee au Quebec",
       x = "Annee d'arrivee au Quebec") +
  theme_bw()

```

### Origine

```{r}

trajipaq <-
  trajipaq %>%
  mutate(ORIGINE = case_when(
    region_naissance == "ANMO" ~ "Afrique du Nord et Moyen Orient",
    region_naissance == "ACAS" ~ "Amerique centrale et Amerique su Sud",
    region_naissance == "Afrique subsaharienne" ~ "Afrique subsaharienne",
    region_naissance == "Asie" ~ "Asie",
    region_naissance == "Autres" ~ "Autres",
    region_naissance == "Canada" ~ "Canada",
    region_naissance == "Caraibes" ~ "Caraibes",
    region_naissance == "Europe" ~ "Europe",
    region_naissance == "France" ~ "France",
    region_naissance == "NSP" ~ NA,
  ))

```

```{r}

freq(trajipaq$ORIGINE)

```

```{r}

ggplot(subset(trajipaq, !is.na(ORIGINE)), aes(x = ORIGINE)) +
  geom_bar(fill = "pink") +
  labs(
    title = "Distribution de l'origine par region",
    x = "Region"
       ) +
  coord_flip() +
  theme_bw()

```

### Revenu

```{r}

trajipaq <-
  trajipaq %>%
  mutate(REVENU = case_when(
    Q100 == 1 ~ "[0 ; 19 999]",
    Q100 == 2 ~ "[0 ; 19 999]",
    Q100 == 3 ~ "[20 000 ; 39 999]",
    Q100 == 4 ~ "[20 000 ; 39 999]",
    Q100 == 5 ~ "[40 000 ; 59 999]",
    Q100 == 6 ~ "[40 000 ; 59 999]",
    Q100 == 7 ~ "[60 000 ; 79 999]",
    Q100 == 8 ~ "[60 000 ; 79 999]",
    Q100 == 9 ~ "[80 000 ; 99 999]",
    Q100 == 10 ~ "[90 000 ; 99 999]",
    Q100 == 11 ~ "[100 000 ; 149 999]",
    Q100 == 12 ~ "[150 000 et plus]",
    Q100 == 98 ~ NA,
    Q100 == 99 ~ NA
  ))

trajipaq$REVENU <- factor(trajipaq$REVENU,levels = c(
  "[0 ; 19 999]",
  "[20 000 ; 39 999]",
  "[40 000 ; 59 999]",
  "[60 000 ; 79 999]",
  "[80 000 ; 99 999]",
  "[100 000 ; 149 999]",
  "[150 000 et plus]",
  NA
), ordered = TRUE)

```

```{r}

freq(trajipaq$REVENU)

```

```{r}

ggplot(subset(trajipaq, !is.na(REVENU)), aes(x = REVENU)) +
  geom_bar(fill = "pink") +
  labs(
    title = "Distribution du revenu par groupes",
    x = "Revenu"
       ) +
  coord_flip() +
  theme_bw()

```

### Langue maternelle

```{r}

trajipaq <-
  trajipaq %>%
  mutate(LANGUE = case_when(
    Q20M1 == 1 ~ "Francais",
    Q20M1 == 2 ~ "Anglais",
    Q20M1 == 97 ~ "Autre",
    Q20M1 == 98 ~ NA,
    Q20M1 == 99 ~ NA
  ))

```

```{r}

freq(trajipaq$LANGUE)

```

```{r}

ggplot(subset(trajipaq, !is.na(LANGUE)), aes(x = LANGUE)) +
  geom_bar(fill = "pink") +
  labs(
    title = "Distribution de la langue maternelle",
    x = "Langue maternelle"
       ) +
  coord_flip() +
  theme_bw()

```

### Religion

```{r}

trajipaq <-
  trajipaq %>%
  mutate(RELIGION = case_when(
    Q17 == 0 ~ "Sans religion",
    Q17 == 1 ~ "Catholique",
    Q17 == 2 ~ "Protestante",
    Q17 == 3 ~ "Autres Chretiens",
    Q17 == 4 ~ "Autres Chretiens",
    Q17 == 5 ~ "Autres Chretiens",
    Q17 == 6 ~ "Orthodoxe",
    Q17 == 7 ~ "Autres Chretiens",
    Q17 == 8 ~ "Musulmans",
    Q17 == 9 ~ "Autre",
    Q17 == 10 ~ "Autre",
    Q17 == 11 ~ "Autre",
    Q17 == 12 ~ "Autre",
    Q17 == 97 ~ "Autre",
    Q17 == 98 ~ NA,
    Q17 == 99 ~ NA
  ))

```

```{r}

freq(trajipaq$RELIGION)

```

```{r}

ggplot(subset(trajipaq, !is.na(RELIGION)), aes(x = RELIGION)) +
  geom_bar(fill = "pink") +
  labs(
    title = "Distribution de la religion",
    x = "Religion"
       ) +
  coord_flip() +
  theme_bw()

```

### Statut matrimonial

```{r}

trajipaq <-
  trajipaq %>%
  mutate(MATRIMONIAL = case_when(
    STATUT == 1 ~ "Celibataire",
    STATUT == 2 ~ "Marie(e)",
    STATUT == 3 ~ "Conjoint(e) de fait",
    STATUT == 4 ~ "Separe(e)",
    STATUT == 5 ~ "Divorce(e)",
    STATUT == 6 ~ "Veuf(ve)"
  ))

```

```{r}

freq(trajipaq$MATRIMONIAL)

```

```{r}

ggplot(subset(trajipaq, !is.na(MATRIMONIAL)), aes(x = MATRIMONIAL)) +
  geom_bar(fill = "pink") +
  labs(
    title = "Distribution du statut matrimonial",
    x = "Statut matrimonial"
       ) +
  coord_flip() +
  theme_bw()

```

## Variables dépendantes

### VD1 : sentiment d'appartenance au Québec

```{r}

freq(trajipaq$Q39A)

trajipaq <-
  trajipaq %>%
  mutate(Q39A = case_when(
    Q39A == 1 ~ 4,
    Q39A == 2 ~ 3,
    Q39A == 3 ~ 2,
    Q39A == 4 ~ 1,
    Q39A == 98 ~ NA,
    Q39A == 99 ~ NA
  ))

trajipaq <-
  trajipaq %>%
  mutate(Q39B = case_when(
    Q39B == 1 ~ 4,
    Q39B == 2 ~ 3,
    Q39B == 3 ~ 2,
    Q39B == 4 ~ 1,
    Q39B == 98 ~ NA,
    Q39B == 99 ~ NA
  ))

trajipaq <-
  trajipaq %>%
  mutate(Q39C = case_when(
    Q39C == 1 ~ 4,
    Q39C == 2 ~ 3,
    Q39C == 3 ~ 2,
    Q39C == 4 ~ 1,
    Q39C == 98 ~ NA,
    Q39C == 99 ~ NA
  ))

trajipaq <-
  trajipaq %>%
  mutate(Q39D = case_when(
    Q39D == 1 ~ 4,
    Q39D == 2 ~ 3,
    Q39D == 3 ~ 2,
    Q39D == 4 ~ 1,
    Q39D == 98 ~ NA,
    Q39D == 99 ~ NA
  ))

trajipaq <-
  trajipaq %>% 
  mutate(Q39_new_A = Q39A,
         Q39_new_B = Q39B,
         Q39_new_C = Q39C,
         Q39_new_D = Q39D) %>% 
  mutate_at(
    vars(starts_with("Q39_new")),
    funs(case_when(
      . == 1 ~ 4,
      . == 2 ~ 3,
      . == 3 ~ 2,
      . == 4 ~ 1,
      . == 98 ~ NA,
      . == 99 ~ NA
    ))
  )

ctable(trajipaq$Q39A, trajipaq$Q39_new_A, "no")

```

```{r}

v_VD1 <- c(
  "Q39_new_A",
  "Q39_new_B",
  "Q39_new_C",
  "Q39_new_D"
)

nbre_lvl <- 4 # nombre de réponses possibles

trajipaq <-
  trajipaq %>%
  rowwise() %>%
  mutate(VD1 = (sum(c_across(all_of(v_VD1)), na.rm = TRUE) - length(v_VD1)) / ((nbre_lvl - 1)*length(v_VD1))) %>%
  mutate(VD1 = if (VD1 < 0) NA_real_ else VD1) %>%
  ungroup()

trajipaq <-
  trajipaq %>% 
  rowwise() %>% 
  mutate (nbre_rep_rep = sum(!is.na(across(starts_with("Q39_new")))))

trajipaq <-
  trajipaq %>%
  rowwise() %>%
  mutate(sumVD1 = sum(c_across(all_of(v_VD1)), na.rm = TRUE)) %>% 
  mutate(VD1_new = (sumVD1 - nbre_rep_rep) / ((nbre_lvl - 1)*nbre_rep_rep)) %>%
  mutate(VD1_new = if (nbre_rep_rep == 0) NA_real_ else VD1_new) %>%
  ungroup()

descr(trajipaq$VD1_new)

```

```{r}

ggplot(data = trajipaq, mapping = aes(x = VD1_new)) +
  geom_histogram(na.rm = TRUE, fill = "pink", binwidth = 0.01) +
  labs(
    title = "Distribution du sentiment d'appartenance au Quebec",
    x = "Indice du sentiment d'appartenance au Quebec"
       ) +
  theme_bw()

```

### VD2 : sentiment d'appartenance (Canada)

```{r}

trajipaq <-
  trajipaq %>%
  mutate(VD2 = case_when(
    Q39E == 1 ~ "Oui",
    Q39E == 2 ~ "Oui",
    Q39E == 3 ~ "Non",
    Q39E == 4 ~ "Non",
    Q39E == 98 ~ NA,
    Q39E == 99 ~ NA
  ))

```

```{r}

freq(trajipaq$VD2)

```

```{r}

ggplot(subset(trajipaq, !is.na(VD2)), aes(x = VD2)) +
  geom_bar(fill = "pink") +
  labs(
    title = "Distribution du sentiment d'appartenance au Canada",
    x = "Sentiment d'appartenance au Canada"
       ) +
  coord_flip() +
  theme_bw()

```

### VD3 : sentiment d'appartenance au pays d'origine

```{r}

trajipaq <-
  trajipaq %>%
  mutate(VD3 = case_when(
    Q39F == 1 ~ "Oui",
    Q39F == 2 ~ "Oui",
    Q39F == 3 ~ "Non",
    Q39F == 4 ~ "Non",
    Q39F == 98 ~ NA,
    Q39F == 99 ~ NA
  ))

freq(trajipaq$VD3)

```

```{r}

ggplot(subset(trajipaq, !is.na(VD3)), aes(x = VD3)) +
  geom_bar(fill = "pink") +
  labs(
    title = "Distribution du sentiment d'appartenance au pays d'origine",
    x = "Sentiment d'appartenance au pays d'origine"
       ) +
  coord_flip() +
  theme_bw()

```

### VD4 : sentiment d'appartenance au Canada comparativement au pays d'origine

```{r}

trajipaq <-
  trajipaq %>% 
  mutate(VD4= case_when(
    VD2 == "Oui" & VD3 == "Oui" ~ "Intégré",
    VD2 == "Oui" & VD3 == "Non" ~ "Assimilé",
    VD2 == "Non" & VD3 == "Oui" ~ "Séparé",
    VD2 == "Non" & VD3 == "Non" ~ "Marginalisé"
  ))

freq(trajipaq$VD4)

```

### VD5 : sentiment d'appartenance au Canada comparativement au Québec

Q39E: Je me sens Canadien et 
Q39D: Je me sens Québecois

```{r}

trajipaq <-
  trajipaq %>% 
  mutate(Q39Douinon = case_when(
         Q39D <= 2 ~ "Oui", 
         Q39D == 3 | Q39D == 4 ~ "Non", 
         Q39D >= 97 ~ NA_character_))

ctable(trajipaq$Q39D, trajipaq$Q39Douinon, "no")

trajipaq <-
  trajipaq %>% 
  mutate(VD5= factor(case_when(
    VD2 == "Oui" & Q39Douinon == "Oui" ~ 1,
    VD2 == "Oui" & Q39Douinon == "Non" ~ 2,
    VD2 == "Non" & Q39Douinon == "Oui" ~ 3,
    VD2 == "Non" & Q39Douinon == "Non" ~ 4
  ),
  labels = c("Canada et Québec", "Canada seule", "Québec seule", "Aucun"),
  levels = c(1, 2, 3, 4)))

freq(trajipaq$VD5)

```

## Graphiques

```{r}

```

## Variables indépendantes

### VI4 : discrimination subie au cours des 12 derniers mois

```{r}

trajipaq <-
  trajipaq %>%
  mutate(Q40A = case_when(
    Q40A == 1 ~ 3,
    Q40A == 2 ~ 2,
    Q40A == 3 ~ 1,
    Q40A == 96 ~ NA,
    Q40A == 98 ~ NA,
    Q40A == 99 ~ NA
  ))

trajipaq <-
  trajipaq %>%
  mutate(Q40B = case_when(
    Q40B == 1 ~ 3,
    Q40B == 2 ~ 2,
    Q40B == 3 ~ 1,
    Q40B == 96 ~ NA,
    Q40B == 98 ~ NA,
    Q40B == 99 ~ NA
  ))

trajipaq <-
  trajipaq %>%
  mutate(Q40C = case_when(
    Q40C == 1 ~ 3,
    Q40C == 2 ~ 2,
    Q40C == 3 ~ 1,
    Q40C == 96 ~ NA,
    Q40C == 98 ~ NA,
    Q40C == 99 ~ NA
  ))

trajipaq <-
  trajipaq %>%
  mutate(Q40D = case_when(
    Q40D == 1 ~ 3,
    Q40D == 2 ~ 2,
    Q40D == 3 ~ 1,
    Q40D == 96 ~ NA,
    Q40D == 98 ~ NA,
    Q40D == 99 ~ NA
  ))

```

```{r}

v_VI4 <- c(
  "Q40A",
  "Q40B",
  "Q40C",
  "Q39D"
  )

trajipaq <-
  trajipaq %>%
  rowwise() %>%
  mutate(VI4 = (sum(c_across(all_of(v_VI4)), na.rm = TRUE) - length(v_VI4)) / (2.25*length(v_VI4))) %>%
  mutate(VI4 = if (VI4 < 0) NA_real_ else VI4) %>%
  ungroup()

```

```{r}

descr(trajipaq$VI4)

```

```{r}

ggplot(data = trajipaq, mapping = aes(x = VI4)) +
  geom_histogram(na.rm = TRUE, fill = "pink", binwidth = 0.01) +
  labs(
    title = "Distribution de la discrimination subie au cours des 12 derniers mois",
    x = "Indice de la discrimination subie au cours des 12 derniers mois"
       ) +
  theme_bw()

```

# Analyse

## H1 : VD1 entretient une relation négative avec VI4

### Modèle de régression

```{r}

reg_1 <- lm(trajipaq$VI4 ~ trajipaq$VD1 +
              trajipaq$sexe +
              trajipaq$age +
              trajipaq$EDUCATION +
              trajipaq$ARRIVEE +
              trajipaq$ORIGINE +
              trajipaq$REVENU +
              trajipaq$LANGUE +
              trajipaq$RELIGION +
              trajipaq$MATRIMONIAL
            )
summary(reg_1)

# Le modèle de régression traite les variables d'une manière inattendue.

```

### Coefficient de corrélation

```{r}

cor.test(trajipaq$VI4, trajipaq$VD1)

```

### Visualisation

```{r}

vis_H1 <- ggplot(trajipaq %>%
                filter(!is.na(VD1), !is.na(VI4), !is.na(sexe)), aes(x = VI4, y = VD1)) +
  geom_jitter(color = "pink") +
  stat_smooth(method = "lm", 
              formula = "y ~ x", 
              se = FALSE,
              color = "purple") +
  labs(title = "Sentiment d'appartenance et discrimination", 
       x = "Discrimination subie au cours des 12 derniers mois", 
       y = "Sentiment d'appartenance au Quebec") +
  theme_bw()
print(vis_H1)

# La relation est - de manière inattendue - positive.

```

#### Sexe

```{r}

vis_sexe <- ggplot(trajipaq %>%
                filter(!is.na(VD1), !is.na(VI4), !is.na(sexe)), aes(x = VI4, y = VD1, color = sexe)) +
  geom_jitter(color = "pink") +
  stat_smooth(method = "lm", formula = "y ~ x", se = FALSE) +
  scale_color_brewer(name = "Sexe", palette = 1) +
  labs(title = "Sentiment d'appartenance et discrimination en fonction du sexe", 
       x = "Discrimination subie au cours des 12 derniers mois", 
       y = "Sentiment d'appartenance au Quebec") +
  theme_bw()
print(vis_sexe)

# La relation semble plus forte chez les femmes avec cette méthode de visualisation.

```

```{r}

vis_sexe_2 <- ggplot(trajipaq %>%
                filter(!is.na(VD1), !is.na(VI4), !is.na(sexe)), aes(x = VI4, y = VD1)) +
  geom_jitter(color = "pink") +
  stat_smooth(method = "lm", 
              formula = "y ~ x", 
              se = FALSE,
              color = "purple") +
  labs(title = "Sentiment d'appartenance et discrimination en fonction du sexe", 
       x = "Discrimination subie au cours des 12 derniers mois", 
       y = "Sentiment d'appartenance au Quebec") +
  facet_wrap(~ sexe, nrow = 2) +
  theme_bw()
print(vis_sexe_2)

# La relation ne semble pas être affectée par le sexe avec cette méthode de visualisation.

```

#### Age

```{r}

vis_age <- ggplot(trajipaq %>%
                filter(!is.na(VD1), !is.na(VI4), !is.na(age)), aes(x = VI4, y = VD1, color = age)) +
  geom_jitter(color = "pink") +
  stat_smooth(method = "lm", formula = "y ~ x", se = FALSE) +
  scale_color_brewer(name = "Age", palette = 1) +
  labs(title = "Sentiment d'appartenance et discrimination en fonction de l'age", 
       x = "Discrimination subie au cours des 12 derniers mois", 
       y = "Sentiment d'appartenance au Quebec") +
  theme_bw()
print(vis_age)

# La relation semble affectée par l'âge avec cette méthode de visualisation.

```

```{r}

vis_age_2 <- ggplot(trajipaq %>%
                filter(!is.na(VD1), !is.na(VI4), !is.na(age)), aes(x = VI4, y = VD1)) +
  geom_jitter(color = "pink") +
  stat_smooth(method = "lm", 
              formula = "y ~ x", 
              se = FALSE,
              color = "purple") +
  labs(title = "Sentiment d'appartenance et discrimination en fonction de l'age", 
       x = "Discrimination subie au cours des 12 derniers mois", 
       y = "Sentiment d'appartenance au Quebec") +
  facet_wrap(~ age, nrow = 3) +
  theme_bw()
print(vis_age_2)

# La relation semble également être affectée par l'âge avec cette méthode de visualisation.

```
#### Education

```{r}

vis_EDUCATION <- ggplot(trajipaq %>%
                filter(!is.na(VD1), !is.na(VI4), !is.na(EDUCATION)), aes(x = VI4, y = VD1, color = EDUCATION)) +
  geom_jitter(color = "pink") +
  stat_smooth(method = "lm", formula = "y ~ x", se = FALSE) +
  scale_color_brewer(name = "Niveau d'education", palette = 1) +
  labs(title = "Sentiment d'appartenance et discrimination en fonction de l'education", 
       x = "Discrimination subie au cours des 12 derniers mois", 
       y = "Sentiment d'appartenance au Quebec") +
  theme_bw()
print(vis_EDUCATION)

# La relation semble être affectée par le niveau d'éducation avec cette méthode de visualisation.

```

```{r}

vis_EDUCATION_2 <- ggplot(trajipaq %>%
                filter(!is.na(VD1), !is.na(VI4), !is.na(EDUCATION)), aes(x = VI4, y = VD1)) +
  geom_jitter(color = "pink") +
  stat_smooth(method = "lm", 
              formula = "y ~ x", 
              se = FALSE,
              color = "purple") +
  labs(title = "Sentiment d'appartenance et discrimination en fonction de l'education", 
       x = "Discrimination subie au cours des 12 derniers mois", 
       y = "Sentiment d'appartenance au Quebec") +
  facet_wrap(~ EDUCATION, nrow = 3) +
  theme_bw()
print(vis_EDUCATION_2)

# La relation ne semble pas être affectée par le niveau d'éducation avec cette méthode de visualisation.

```

#### Année d'arrivée au Québec

```{r}

trajipaq <-
  trajipaq %>%
  mutate(ARRIVEE_C = case_when(
    ARRIVEE >= 1990 & ARRIVEE < 2000 ~ "[1990;1999]",
    ARRIVEE >= 2000 & ARRIVEE < 2005 ~ "[2000;2004]",
    ARRIVEE >= 2005 & ARRIVEE < 2010 ~ "[2005;2009]",
    ARRIVEE >= 2010 & ARRIVEE < 2015 ~ "[2010;2014]",
    ARRIVEE >= 2015 & ARRIVEE < 2020 ~ "[2015;2020]",
  ))

```

```{r}

vis_ARRIVEE <- ggplot(trajipaq %>%
                filter(!is.na(VD1), !is.na(VI4), !is.na(ARRIVEE_C)), aes(x = VI4, y = VD1, color = as.factor(ARRIVEE_C))) +
  geom_jitter(color = "pink") +
  stat_smooth(method = "lm", formula = "y ~ x", se = FALSE) +
  scale_color_brewer(name = "Annee d'arrivee au Quebec", palette = 1) +
  labs(title = "Sentiment d'appartenance et discrimination en fonction de l'annee d'arrivee", 
       x = "Discrimination subie au cours des 12 derniers mois", 
       y = "Sentiment d'appartenance au Quebec") +
  theme_bw()
print(vis_ARRIVEE)

# La relation semble être affectée par l'année d'arrivée au Québec avec cette méthode de visualisation.

```

```{r}

vis_ARRIVEE_2 <- ggplot(trajipaq %>%
                filter(!is.na(VD1), !is.na(VI4), !is.na(ARRIVEE_C)), aes(x = VI4, y = VD1)) +
  geom_jitter(color = "pink") +
  stat_smooth(method = "lm", 
              formula = "y ~ x", 
              se = FALSE,
              color = "purple") +
  labs(title = "Sentiment d'appartenance et discrimination en fonction de l'annee d'arrivee", 
       x = "Discrimination subie au cours des 12 derniers mois", 
       y = "Sentiment d'appartenance au Quebec") +
  facet_wrap(~ ARRIVEE_C, nrow = 3) +
  theme_bw()
print(vis_ARRIVEE_2)

# La relation semble également affectée par l'année d'arrivée au Québec avec cette méthode de visualisation.

```

#### Origine

```{r}

vis_ORIGINE <- ggplot(trajipaq %>%
                filter(!is.na(VD1), !is.na(VI4), !is.na(ORIGINE)), aes(x = VI4, y = VD1, color = ORIGINE)) +
  geom_jitter(color = "pink") +
  stat_smooth(method = "lm", formula = "y ~ x", se = FALSE) +
  scale_color_brewer(name = "Region d'origine", palette = 1) +
  labs(title = "Sentiment d'appartenance et discrimination en fonction de l'origine", 
       x = "Discrimination subie au cours des 12 derniers mois", 
       y = "Sentiment d'appartenance au Quebec") +
  theme_bw()
print(vis_ORIGINE)

# La relation semble être affectée par la région d'origine avec cette méthode de visualisation.

```

```{r}

vis_ORIGINE_2 <- ggplot(trajipaq %>%
                filter(!is.na(VD1), !is.na(VI4), !is.na(ORIGINE)), aes(x = VI4, y = VD1)) +
  geom_jitter(color = "pink") +
  stat_smooth(method = "lm", 
              formula = "y ~ x", 
              se = FALSE,
              color = "purple") +
  labs(title = "Sentiment d'appartenance et discrimination en fonction de l'origine", 
       x = "Discrimination subie au cours des 12 derniers mois", 
       y = "Sentiment d'appartenance au Quebec") +
  facet_wrap(~ ORIGINE, nrow = 3) +
  theme_bw()
print(vis_ORIGINE_2)

# La relation semble également être affectée avec cette méthode de visualisation.

```

#### Revenu

```{r}

vis_REVENU <- ggplot(trajipaq %>%
                filter(!is.na(VD1), !is.na(VI4), !is.na(REVENU)), aes(x = VI4, y = VD1, color = REVENU)) +
  geom_jitter(color = "pink") +
  stat_smooth(method = "lm", formula = "y ~ x", se = FALSE) +
  scale_color_brewer(name = "Revenu", palette = 1) +
  labs(title = "Sentiment d'appartenance et discrimination en fonction du revenu", 
       x = "Discrimination subie au cours des 12 derniers mois", 
       y = "Sentiment d'appartenance au Quebec") +
  theme_bw()
print(vis_REVENU)

# La relation semble être affectée par le revenu avec cette méthode de visualisation.

```

```{r}

vis_REVENU_2 <- ggplot(trajipaq %>%
                filter(!is.na(VD1), !is.na(VI4), !is.na(REVENU)), aes(x = VI4, y = VD1)) +
  geom_jitter(color = "pink") +
  stat_smooth(method = "lm", 
              formula = "y ~ x", 
              se = FALSE,
              color = "purple") +
  labs(title = "Sentiment d'appartenance et discrimination en fonction du revenu", 
       x = "Discrimination subie au cours des 12 derniers mois", 
       y = "Sentiment d'appartenance au Quebec") +
  facet_wrap(~ REVENU, nrow = 3) +
  theme_bw()
print(vis_REVENU_2)

# La relation semble également être affectée par le revenu avec cette méthode de visualisation.

```

#### Langue maternelle

```{r}

vis_LANGUE <- ggplot(trajipaq %>%
                filter(!is.na(VD1), !is.na(VI4), !is.na(LANGUE)), aes(x = VI4, y = VD1, color = LANGUE)) +
  geom_jitter(color = "pink") +
  stat_smooth(method = "lm", formula = "y ~ x", se = FALSE) +
  scale_color_brewer(name = "Langue maternelle", palette = 1) +
  labs(title = "Sentiment d'appartenance et discrimination en fonction de la langue maternelle", 
       x = "Discrimination subie au cours des 12 derniers mois", 
       y = "Sentiment d'appartenance au Quebec") +
  theme_bw()
print(vis_LANGUE)

# La relation semble être affectée par la langue maternelle avec cette méthode de visualisation.

```

```{r}

vis_LANGUE_2 <- ggplot(trajipaq %>%
                filter(!is.na(VD1), !is.na(VI4), !is.na(LANGUE)), aes(x = VI4, y = VD1)) +
  geom_jitter(color = "pink") +
  stat_smooth(method = "lm", 
              formula = "y ~ x", 
              se = FALSE,
              color = "purple") +
  labs(title = "Sentiment d'appartenance et discrimination en fonction de la langue maternelle", 
       x = "Discrimination subie au cours des 12 derniers mois", 
       y = "Sentiment d'appartenance au Quebec") +
  facet_wrap(~ LANGUE, nrow = 2) +
  theme_bw()
print(vis_LANGUE_2)

# La relation semble être également affectée par la langue maternelle avec cette méthode de visualisation.

```

### Religion

```{r}

vis_RELIGION <- ggplot(trajipaq %>%
                filter(!is.na(VD1), !is.na(VI4), !is.na(RELIGION)), aes(x = VI4, y = VD1, color = RELIGION)) +
  geom_jitter(color = "pink") +
  stat_smooth(method = "lm", formula = "y ~ x", se = FALSE) +
  scale_color_brewer(name = "Religion", palette = 1) +
  labs(title = "Sentiment d'appartenance et discrimination en fonction de la religion", 
       x = "Discrimination subie au cours des 12 derniers mois", 
       y = "Sentiment d'appartenance au Quebec") +
  theme_bw()
print(vis_RELIGION)

# La relation semble être affectée par la religion avec cette méthode de visualisation.

```

```{r}

vis_RELIGION_2 <- ggplot(trajipaq %>%
                filter(!is.na(VD1), !is.na(VI4), !is.na(RELIGION)), aes(x = VI4, y = VD1)) +
  geom_jitter(color = "pink") +
  stat_smooth(method = "lm", 
              formula = "y ~ x", 
              se = FALSE,
              color = "purple") +
  labs(title = "Sentiment d'appartenance et discrimination en fonction de la religion", 
       x = "Discrimination subie au cours des 12 derniers mois", 
       y = "Sentiment d'appartenance au Quebec") +
  facet_wrap(~ RELIGION, nrow = 3) +
  theme_bw()
print(vis_RELIGION_2)

# La relation semble être également affectée par la religion avec cette méthode de visualisation.

```

### Statut matrimonial

```{r}

vis_MATRIMONIAL <- ggplot(trajipaq %>%
                filter(!is.na(VD1), !is.na(VI4), !is.na(MATRIMONIAL)), aes(x = VI4, y = VD1, color = MATRIMONIAL)) +
  geom_jitter(color = "pink") +
  stat_smooth(method = "lm", formula = "y ~ x", se = FALSE) +
  scale_color_brewer(name = "Statut matrimonial", palette = 1) +
  labs(title = "Sentiment d'appartenance et discrimination en fonction du statut matrimonial", 
       x = "Discrimination subie au cours des 12 derniers mois", 
       y = "Sentiment d'appartenance au Quebec") +
  theme_bw()
print(vis_MATRIMONIAL)

# La relation semble être affectée par le statut matrimonial avec cette méthode de visualisation.

```

```{r}

vis_MATRIMONIAL_2 <- ggplot(trajipaq %>%
                filter(!is.na(VD1), !is.na(VI4), !is.na(MATRIMONIAL)), aes(x = VI4, y = VD1)) +
  geom_jitter(color = "pink") +
  stat_smooth(method = "lm", 
              formula = "y ~ x", 
              se = FALSE,
              color = "purple") +
  labs(title = "Sentiment d'appartenance et discrimination en fonction du statut matrimonial", 
       x = "Discrimination subie au cours des 12 derniers mois", 
       y = "Sentiment d'appartenance au Quebec") +
  facet_wrap(~ MATRIMONIAL, nrow = 3) +
  theme_bw()
print(vis_MATRIMONIAL_2)

# La relation semble être également affectée par le statut matrimonial avec cette méthode de visualisation.

```

## H2 : VD2 entretient une relation négative avec VI4

### Modèle de régression

```{r}

trajipaq <-
  trajipaq %>%
  mutate(VD2_N = case_when(
    VD2 == "Non" ~ 0,
    VD2 == "Oui" ~ 1
  ))

```

```{r}

reg_2 <- glm(formula = VD2_N ~ VI4, family = "binomial", data = trajipaq)

summary(reg_2)

```

### Coefficient de corrélation

```{r}

cor.test(trajipaq$VD2_N, trajipaq$VI4)

```

### Visualisation

```{r}

ggplot(trajipaq %>% filter(!is.na(VD2_N), !is.na(VI4)), aes(x = VI4, y = VD2_N)) +
  stat_smooth(method = "glm",
              formula = "y ~ x",
              color = "pink", 
              se = FALSE,
              method.args = list(family = binomial)) +
  labs(
    title = "Sentiment d'appartenance et discrimination", 
    x = "Discrimination subie au cours des 12 derniers mois", 
    y = "Sentiment d'appartenance au Canada") +
  theme_bw()

# La relation est - de manière inattendue - positive.

```

#### Sexe

```{r}

ggplot(trajipaq %>% filter(!is.na(VD2_N), !is.na(VI4)), aes(x = VI4, y = VD2_N)) +
  stat_smooth(method = "glm",
              formula = "y ~ x",
              color = "pink", 
              se = FALSE,
              method.args = list(family = binomial)) +
  labs(
    title = "Sentiment d'appartenance et discrimination en fonction du sexe", 
    x = "Discrimination subie au cours des 12 derniers mois", 
    y = "Sentiment d'appartenance au Canada") +
  facet_wrap(~ sexe, nrow = 2) +
  theme_bw()

# La relation semble être affectée par le sexe.

```

#### Age

```{r}

ggplot(trajipaq %>% filter(!is.na(VD2_N), !is.na(VI4)), aes(x = VI4, y = VD2_N)) +
  stat_smooth(method = "glm",
              formula = "y ~ x",
              color = "pink", 
              se = FALSE,
              method.args = list(family = binomial)) +
  labs(
    title = "Sentiment d'appartenance et discrimination en fonction de l'age", 
    x = "Discrimination subie au cours des 12 derniers mois", 
    y = "Sentiment d'appartenance au Canada") +
  facet_wrap(~ age, nrow = 3) +
  theme_bw()

# La relation semble être affectée par l'âge.

```

#### Education

```{r}

ggplot(trajipaq %>% filter(!is.na(VD2_N), !is.na(VI4), !is.na(EDUCATION)), aes(x = VI4, y = VD2_N)) +
  stat_smooth(method = "glm",
              formula = "y ~ x",
              color = "pink", 
              se = FALSE,
              method.args = list(family = binomial)) +
  labs(
    title = "Sentiment d'appartenance et discrimination en fonction de l'education", 
    x = "Discrimination subie au cours des 12 derniers mois", 
    y = "Sentiment d'appartenance au Canada") +
  facet_wrap(~ EDUCATION, nrow = 3) +
  theme_bw()

# La relation semble être affectée par le niveau d'éducation.

```

#### Année d'arrivée au Québec

```{r}

ggplot(trajipaq %>% filter(!is.na(VD2_N), !is.na(VI4), !is.na(ARRIVEE_C)), aes(x = VI4, y = VD2_N)) +
  stat_smooth(method = "glm",
              formula = "y ~ x",
              color = "pink", 
              se = FALSE,
              method.args = list(family = binomial)) +
  labs(
    title = "Sentiment d'appartenance et discrimination en fonction de l'annee d'arivee", 
    x = "Discrimination subie au cours des 12 derniers mois", 
    y = "Sentiment d'appartenance au Canada") +
  facet_wrap(~ ARRIVEE_C, nrow = 3) +
  theme_bw()

# La relation semble être affectée par le niveau d'éducation.

```

#### Origine

```{r}

ggplot(trajipaq %>% filter(!is.na(VD2_N), !is.na(VI4), !is.na(ORIGINE)), aes(x = VI4, y = VD2_N)) +
  stat_smooth(method = "glm",
              formula = "y ~ x",
              color = "pink", 
              se = FALSE,
              method.args = list(family = binomial)) +
  labs(
    title = "Sentiment d'appartenance et discrimination en fonction de l'origine", 
    x = "Discrimination subie au cours des 12 derniers mois", 
    y = "Sentiment d'appartenance au Canada") +
  facet_wrap(~ ORIGINE, nrow = 3) +
  theme_bw()

# La relation semble être affectée par le pays d'origine.

```

#### Revenu

```{r}

ggplot(trajipaq %>% filter(!is.na(VD2_N), !is.na(VI4), !is.na(REVENU)), aes(x = VI4, y = VD2_N)) +
  stat_smooth(method = "glm",
              formula = "y ~ x",
              color = "pink", 
              se = FALSE,
              method.args = list(family = binomial)) +
  labs(
    title = "Sentiment d'appartenance et discrimination en fonction du revenu", 
    x = "Discrimination subie au cours des 12 derniers mois", 
    y = "Sentiment d'appartenance au Canada") +
  facet_wrap(~ REVENU, nrow = 3) +
  theme_bw()

# La relation semble être affectée par le revenu.

```

#### Langue maternelle

```{r}

ggplot(trajipaq %>% filter(!is.na(VD2_N), !is.na(VI4), !is.na(LANGUE)), aes(x = VI4, y = VD2_N)) +
  stat_smooth(method = "glm",
              formula = "y ~ x",
              color = "pink", 
              se = FALSE,
              method.args = list(family = binomial)) +
  labs(
    title = "Sentiment d'appartenance et discrimination en fonction du revenu", 
    x = "Discrimination subie au cours des 12 derniers mois", 
    y = "Sentiment d'appartenance au Canada") +
  facet_wrap(~ LANGUE, nrow = 2) +
  theme_bw()

# La relation semble être affectée par la langue maternelle.

```

#### Religion

```{r}

ggplot(trajipaq %>% filter(!is.na(VD2_N), !is.na(VI4), !is.na(RELIGION)), aes(x = VI4, y = VD2_N)) +
  stat_smooth(method = "glm",
              formula = "y ~ x",
              color = "pink", 
              se = FALSE,
              method.args = list(family = binomial)) +
  labs(
    title = "Sentiment d'appartenance et discrimination en fonction de la religion", 
    x = "Discrimination subie au cours des 12 derniers mois", 
    y = "Sentiment d'appartenance au Canada") +
  facet_wrap(~ RELIGION, nrow = 3) +
  theme_bw()

# La relation semble être affectée par la religion.

```

#### Statut matrimonial

```{r}

ggplot(trajipaq %>% filter(!is.na(VD2_N), !is.na(VI4), !is.na(MATRIMONIAL)), aes(x = VI4, y = VD2_N)) +
  stat_smooth(method = "glm",
              formula = "y ~ x",
              color = "pink", 
              se = FALSE,
              method.args = list(family = binomial)) +
  labs(
    title = "Sentiment d'appartenance et discrimination en fonction du statut matrimonial", 
    x = "Discrimination subie au cours des 12 derniers mois", 
    y = "Sentiment d'appartenance au Canada") +
  facet_wrap(~ MATRIMONIAL, nrow = 3) +
  theme_bw()

# La relation semble être affectée par le statut matrimonial.

```

## H3 : VD3 entretient une relation négative avec VI4

### Modèle de régression

```{r}

trajipaq <-
  trajipaq %>%
  mutate(VD3_N = case_when(
    VD3 == "Non" ~ 0,
    VD3 == "Oui" ~ 1
  ))

```

```{r}

reg_3 <- glm(formula = VD3_N ~ VI4, family = "binomial", data = trajipaq)

summary(reg_3)

# La relation n'est pas significative avec une chance d'erreur alpha égale à 5%.

```

## H4 : VD1 entretien une relation avec VD3 (en tant que variable indépendante)

```{r}

trajipaq <-
  trajipaq %>%
  mutate(VD1_F = case_when(
    VD1 < 0.5 ~ "Non",
    VD1 >= 0.5 ~ "Oui"
  ))

```

```{r}

ctable(trajipaq$VD3, trajipaq$VD1_F)

```

```{r}

chisq.test(trajipaq$VD3, trajipaq$VD1_F)

# La relation n'est pas significative avec une chance d'erreur alpha égale à 5%.

```

### 2023-05-30 Codes/Sentiment_an_2.Rmd
update

@@
 ```{r}
-
 library(summarytools)
@@
 
-### Sexe
+- Sexe
 
@@
 
-### Âge
+- Âge
 
@@
 
-### Education
+- Education
 
@@
 
+```
+
+```{r}
+
 freq(trajipaq$VD3)
@@
   theme_bw()
-
-```
-
-### VD4 : sentiment d'appartenance au Canada comparativement au pays d'origine
-
-```{r}
-
-trajipaq <-
-  trajipaq %>% 
-  mutate(VD4= case_when(
-    VD2 == "Oui" & VD3 == "Oui" ~ "Intégré",
-    VD2 == "Oui" & VD3 == "Non" ~ "Assimilé",
-    VD2 == "Non" & VD3 == "Oui" ~ "Séparé",
-    VD2 == "Non" & VD3 == "Non" ~ "Marginalisé"
-  ))
-
-freq(trajipaq$VD4)
-
-```
-
-### VD5 : sentiment d'appartenance au Canada comparativement au Québec
-
-Q39E: Je me sens Canadien et 
-Q39D: Je me sens Québecois
-
-```{r}
-
-trajipaq <-
-  trajipaq %>% 
-  mutate(Q39Douinon = case_when(
-         Q39D <= 2 ~ "Oui", 
-         Q39D == 3 | Q39D == 4 ~ "Non", 
-         Q39D >= 97 ~ NA_character_))
-
-ctable(trajipaq$Q39D, trajipaq$Q39Douinon, "no")
-
-trajipaq <-
-  trajipaq %>% 
-  mutate(VD5= factor(case_when(
-    VD2 == "Oui" & Q39Douinon == "Oui" ~ 1,
-    VD2 == "Oui" & Q39Douinon == "Non" ~ 2,
-    VD2 == "Non" & Q39Douinon == "Oui" ~ 3,
-    VD2 == "Non" & Q39Douinon == "Non" ~ 4
-  ),
-  labels = c("Canada et Québec", "Canada seule", "Québec seule", "Aucun"),
-  levels = c(1, 2, 3, 4)))
-
-freq(trajipaq$VD5)
-
-```
-
-## Graphiques
-
-```{r}
 

### 2023-05-30 Article-Reseaux sociaux_sentiment.Rmd

@@
 
-# Question de recherche = comment les sentiments d'appartenances sont ils associés au pays d'origine et aux situations de discriminations vécues par les immigrants établis au Québec.
+# Question de recherche = comment les sentiments d'appartenances sont ils associés au pays d'origine et aux situations de discriminations vécues par les immigrants établis au Québec ?
 

### 2023-05-31 1. Trajipaq_sentiment/1. Article/Visseho_al_sentiment_appartenance.Rmd
Réorganisation de l'espace de travail.

--- 2023-05-29 Article/Visseho_al_sentiment_appartenance.Rmd
@@
 
-Les immigrants sont souvent pris entre deux ou plusieurs cultures, celles de leur pays d'origine et celles de leur nouveau pays de résidence.
+```{r}
+
+# Partie à rédiger en dernier.
+
+# Les immigrants sont souvent pris entre deux ou plusieurs cultures, celles de leur pays d'origine et celles de leur nouveau pays de résidence.
 Cet article essaie de comprendre comment ces sentiments sont associés à leur pays d'origine et aux éventuelles situations de discriminations qu'ils vivent dans leur pays d'accueil.
 
-## Variables
+```
 
-### Variables dépendantes
+# Introduction
 
-#### VD1. Sentiment d'appartenance au Québec
+```{r}
 
--   Q39A (facteur/ordinal)
+# Partie à rédiger après la discussion des résultats.
 
--   Q39B (facteur/ordinal)
+# Question de recherche = comment les sentiments d'appartenances sont ils associés au pays d'origine et aux situations de discriminations vécues par les immigrants établis au Québec.
 
--   Q39C (facteur/ordinal)
+```
 
--   Q39D (facteur/ordinal)
+# Revue de littérature
 
--   VD1 (numérique/ratio)
+```{r}
 
-#### VD2. Sentiment d'appartenance au Canada
+# Partie à rédiger en priorité.
 
--   Q39E (facteur/nominal)
-
-    |                    | **Canada** | **Oui**  | **Non**     |
-    |--------------------|------------|----------|-------------|
-    | **Pays d'origine** |            |          |             |
-    | **Oui**            |            | Intégré  | Séparé      |
-    | **Non**            |            | Assimilé | Marginalisé |
-
-#### VD3. Sentiment d'appartenance au pays d'origine
-
--   Q39F (facteur/nominal)
-
--   VD3 (facteur/nominal)
-
-### Variables indépendantes
-
-#### VI1. Région d'origine
-
--   Q11 (facteur/nominal)
-
--   VI1 (facteur/nominal)
-
-#### VI2. Distance culturelle avec le Canada
-
--   ??? (numérique/ratio)
-
-#### VI3. Stabilité dans le pays d'origine
-
--   ??? (numérique/ratio)
-
-#### VI4. Discrimination subie au cours des 12 derniers mois
-
--   Q40A (facteur/ordinal)
-
--   Q40B (facteur/ordinal)
-
--   Q40C (facteur/ordinal)
-
--   Q40D (facteur/ordinal)
-
--   VI4 (numérique/ratio)
-
-## Analyses
-
-### Variables dépendantes
-
--   VD1 (moyenne/écart-type)
-
--   VD2 (fréquence/mode)
-
--   VD3 (fréquence/mode)
-
-### Variables indépendantes
-
--   VI1 (fréquence/mode)
-
--   VI2 (moyenne/écart-type)
-
--   VI3 (moyenne/écart-type)
-
--   VI4 (moyenne/écart-type)
-
-### Tableaux croisés
-
--   VD1 / (VI1 : VI3)
-
--   VD2 / (VI1 : VI3)
-
--   VD3 / (VI1 : VI3)
-
-### Régression linéaire
-
--   VD1 / VI4
-
-## Causalité éventuelle
-
-Discrimination --\> Sentiment d'appartenance
-
-# références
+```
 
@@
 Dès lors, les immigrants parviennent rarement à une identité nationale ou biculturelle et adoptent plus fréquemment une stratégie de séparation, c’est-à-dire, une grande valeur accordée au désir de conserver leur propre patrimoine culturel et, simultanément, peu d’intérêt à adopter la culture de la société d’accueil (Giuliani et al., 2018; Verkuyten et Martinovic, 2012).-->
+
+# Intégration culturelle
+
+# Sentiments d'appartenance et discriminations
+
+## Sentiments d'appartenance
+
+## Situations de discriminations vécues par les immigrants
+
+# Méthodologie
+
+```{r}
+
+# Partie à rédiger en priorité.
+
+```
+
+## Présentation de l'échantillon
+
+## Présentation des variables
+
+```{r}
+
+# VD1 = indice du sentiment d'appartenance au Québec
+
+# VD2 = sentiment d'appartenance au Canada
+
+# VI1 = région d'origine
+
+# VI2 = indice de la distance culturelle du pays d'origine avec le Canada
+
+# VI3 = indice de la stabilité politique du pays d'origine par rapport à celle du Canada
+
+# VI4 = indice des situations de discriminations subies
+
+```
+
+## Hypothèses de recherche
+
+```{r}
+
+# H1 = la région d'origine des immigrants (VI1) joue un rôle dans le développement du sentiment d'appartenance au Québec (VD1)
+
+# H2 = la région d'origine des immigrants (VI1) joue un rôle dans le développement du sentiment d'appartenance au Canada (VD2)
+
+# H3 = la distance culturelle du pays d'origine des immigrants avec le Canada (VI2) joue un rôle dans le développement du sentiment d'appartenance au Québec (VD1)
+
+# H4 = la distance culturelle du pays d'origine des immigrants avec le Canada (VI2) joue un rôle dans le développement du sentiment d'appartenance au Canada (VD2)
+
+# H5 = la stabilité politique du pays d'origine des immigrants par rapport à celle du Canada (VI3) joue un rôle dans le développement du sentiment d'appartenance au Québec (VD1)
+
+# H6 = la stabilité politique du pays d'origine des immigrants par rapport à celle du Canada (VI3) joue un rôle dans le développement du sentiment d'appartenance au Canada (VD2)
+
+# H7 = les situations de discriminations vécues par les immigrants (VI4) réduisent le développement du sentiment d'appartenance au Québec (VD1)
+
+# H8 = les situations de discriminations vécues par les immigrants (VI4) réduisent le développement du sentiment d'appartenance au Canada (VD2)
+
+```
+
+# Résultats
+
+```{r}
+
+# Partie à rédiger après la revue de littérature et la méthodologie.
+
+```
+
+## Présentation des résultats liés à la première hypothèse
+
+```{r}
+
+# Trouver un titre évocateur après la rédaction.
+
+```
+
+## Présentation des résultats liés à la deuxième hypothèse
+
+```{r}
+
+# Trouver un titre évocateur après la rédaction.
+
+```
+
+## Présentation des résultats liés à la troisième hypothèse
+
+```{r}
+
+# Trouver un titre évocateur après la rédaction.
+
+```
+
+# Discussion
+
+```{r}
+
+# Partie à rédiger après la présentation des résultats.
+
+```
+
+## Discussion des résultats liés à la première hypothèse
+
+```{r}
+
+# Trouver un titre évocateur après la rédaction.
+
+```
+
+## Discussion des résultats liés à la deuxième hypothèse
+
+```{r}
+
+# Trouver un titre évocateur après la rédaction.
+
+```
+
+## Discussion des résultats liés à la troisième hypothèse
+
+```{r}
+
+# Trouver un titre évocateur après la rédaction.
+
+```
+
+## Discussion des limites de l'étude
+
+# Conclusion
+
+```{r}
+
+# Partie à réaliser après la discussion des résultats
+
+```
+
+# Bibliographie
+
+```{r}
+
+# Partie générée automatiquement, mais qui doit tout de même être relue à la fin du travail de rédaction.
+
+```
 
@@
 <https://doi.org/10.1016/j.ijintrel.2009.11.007>
+
+<!-- ## Variables
+
+### Variables dépendantes
+
+#### VD1. Sentiment d'appartenance au Québec
+
+-   Q39A (facteur/ordinal)
+
+-   Q39B (facteur/ordinal)
+
+-   Q39C (facteur/ordinal)
+
+-   Q39D (facteur/ordinal)
+
+-   VD1 (numérique/ratio)
+
+#### VD2. Sentiment d'appartenance au Canada
+
+-   Q39E (facteur/nominal)
+
+    |                    | **Canada** | **Oui**  | **Non**     |
+    |--------------------|------------|----------|-------------|
+    | **Pays d'origine** |            |          |             |
+    | **Oui**            |            | Intégré  | Séparé      |
+    | **Non**            |            | Assimilé | Marginalisé |
+
+#### VD3. Sentiment d'appartenance au pays d'origine
+
+-   Q39F (facteur/nominal)
+
+-   VD3 (facteur/nominal)
+
+### Variables indépendantes
+
+#### VI1. Région d'origine
+
+-   Q11 (facteur/nominal)
+
+-   VI1 (facteur/nominal)
+
+#### VI2. Distance culturelle avec le Canada
+
+-   ??? (numérique/ratio)
+
+#### VI3. Stabilité dans le pays d'origine
+
+-   ??? (numérique/ratio)
+
+#### VI4. Discrimination subie au cours des 12 derniers mois
+
+-   Q40A (facteur/ordinal)
+
+-   Q40B (facteur/ordinal)
+
+-   Q40C (facteur/ordinal)
+
+-   Q40D (facteur/ordinal)
+
+-   VI4 (numérique/ratio)
+
+## Analyses
+
+### Variables dépendantes
+
+-   VD1 (moyenne/écart-type)
+
+-   VD2 (fréquence/mode)
+
+-   VD3 (fréquence/mode)
+
+### Variables indépendantes
+
+-   VI1 (fréquence/mode)
+
+-   VI2 (moyenne/écart-type)
+
+-   VI3 (moyenne/écart-type)
+
+-   VI4 (moyenne/écart-type)
+
+### Tableaux croisés
+
+-   VD1 / (VI1 : VI3)
+
+-   VD2 / (VI1 : VI3)
+
+-   VD3 / (VI1 : VI3)
+
+### Régression linéaire
+
+-   VD1 / VI4
+
+## Causalité éventuelle
+
+Discrimination --\> Sentiment d'appartenance -->

### 2023-05-31 1. Trajipaq_sentiment/2. Codes/Trajipaq_sentiment_an.Rmd
Réorganisation de l'espace de travail.

---
title: "Trajipaq_sentiment_an"
author: "Thea"
date: "2023-05-31"
output: html_document
---

```{r}

library(summarytools)
library(tidyverse)
library(ggthemes)
library(haven)

trajipaq <- read_csv("../3. Data/Trajipaq_sentiment_cr.csv")

```

# Analyse

## H1 : VD1 entretient une relation négative avec VI4

### Modèle de régression

```{r}

reg_1 <- lm(trajipaq$VI4 ~ trajipaq$VD1 +
              trajipaq$sexe +
              trajipaq$age +
              trajipaq$education +
              trajipaq$annee_qc +
              trajipaq$origine +
              trajipaq$revenu +
              trajipaq$langue_mat +
              trajipaq$religion +
              trajipaq$statut_mat
            )
summary(reg_1)

# Le modèle de régression traite les variables d'une manière inattendue.

```

### Coefficient de corrélation

```{r}

cor.test(trajipaq$VI4, trajipaq$VD1)

```

### Visualisation

```{r}

vis_H1 <- ggplot(trajipaq %>%
                filter(!is.na(VD1), !is.na(VI4), !is.na(sexe)), aes(x = VI4, y = VD1)) +
  geom_jitter(color = "pink") +
  stat_smooth(method = "lm", 
              formula = "y ~ x", 
              se = FALSE,
              color = "purple") +
  labs(title = "Sentiment d'appartenance et discrimination", 
       x = "Perception de la Perception de la discrimination subie au cours des 12 derniers mois", 
       y = "Sentiment d'appartenance au Quebec") +
  theme_bw()
print(vis_H1)

```

#### Sexe

```{r}

vis_sexe <- ggplot(trajipaq %>%
                filter(!is.na(VD1), !is.na(VI4), !is.na(sexe)), aes(x = VI4, y = VD1, color = sexe)) +
  geom_jitter(color = "pink") +
  stat_smooth(method = "lm", formula = "y ~ x", se = FALSE) +
  scale_color_brewer(name = "Sexe", palette = 1) +
  labs(title = "Sentiment d'appartenance et discrimination en fonction du sexe", 
       x = "Perception de la discrimination subie au cours des 12 derniers mois", 
       y = "Sentiment d'appartenance au Quebec") +
  theme_bw()
print(vis_sexe)

vis_sexe_2 <- ggplot(trajipaq %>%
                filter(!is.na(VD1), !is.na(VI4), !is.na(sexe)), aes(x = VI4, y = VD1)) +
  geom_jitter(color = "pink") +
  stat_smooth(method = "lm", 
              formula = "y ~ x", 
              se = FALSE,
              color = "purple") +
  labs(title = "Sentiment d'appartenance et discrimination en fonction du sexe", 
       x = "Perception de la discrimination subie au cours des 12 derniers mois", 
       y = "Sentiment d'appartenance au Quebec") +
  facet_wrap(~ sexe, nrow = 2) +
  theme_bw()
print(vis_sexe_2)

```

#### Age

```{r}

vis_age <- ggplot(trajipaq %>%
                filter(!is.na(VD1), !is.na(VI4), !is.na(age)), aes(x = VI4, y = VD1, color = age)) +
  geom_jitter(color = "pink") +
  stat_smooth(method = "lm", formula = "y ~ x", se = FALSE) +
  scale_color_brewer(name = "Age", palette = 1) +
  labs(title = "Sentiment d'appartenance et discrimination en fonction de l'age", 
       x = "Perception de la discrimination subie au cours des 12 derniers mois", 
       y = "Sentiment d'appartenance au Quebec") +
  theme_bw()
print(vis_age)

vis_age_2 <- ggplot(trajipaq %>%
                filter(!is.na(VD1), !is.na(VI4), !is.na(age)), aes(x = VI4, y = VD1)) +
  geom_jitter(color = "pink") +
  stat_smooth(method = "lm", 
              formula = "y ~ x", 
              se = FALSE,
              color = "purple") +
  labs(title = "Sentiment d'appartenance et discrimination en fonction de l'age", 
       x = "Perception de la discrimination subie au cours des 12 derniers mois", 
       y = "Sentiment d'appartenance au Quebec") +
  facet_wrap(~ age, nrow = 3) +
  theme_bw()
print(vis_age_2)

```
#### education

```{r}

vis_education <- ggplot(trajipaq %>%
                filter(!is.na(VD1), !is.na(VI4), !is.na(education)), aes(x = VI4, y = VD1, color = education)) +
  geom_jitter(color = "pink") +
  stat_smooth(method = "lm", formula = "y ~ x", se = FALSE) +
  scale_color_brewer(name = "Niveau d'education", palette = 1) +
  labs(title = "Sentiment d'appartenance et discrimination en fonction de l'education", 
       x = "Perception de la discrimination subie au cours des 12 derniers mois", 
       y = "Sentiment d'appartenance au Quebec") +
  theme_bw()
print(vis_education)

vis_education_2 <- ggplot(trajipaq %>%
                filter(!is.na(VD1), !is.na(VI4), !is.na(education)), aes(x = VI4, y = VD1)) +
  geom_jitter(color = "pink") +
  stat_smooth(method = "lm", 
              formula = "y ~ x", 
              se = FALSE,
              color = "purple") +
  labs(title = "Sentiment d'appartenance et discrimination en fonction de l'education", 
       x = "Perception de la discrimination subie au cours des 12 derniers mois", 
       y = "Sentiment d'appartenance au Quebec") +
  facet_wrap(~ education, nrow = 3) +
  theme_bw()
print(vis_education_2)

```

#### Année d'arrivée au Québec

```{r}

trajipaq <-
  trajipaq %>%
  mutate(annee_qc_f = case_when(
    annee_qc >= 1990 & annee_qc < 2000 ~ "[1990;1999]",
    annee_qc >= 2000 & annee_qc < 2005 ~ "[2000;2004]",
    annee_qc >= 2005 & annee_qc < 2010 ~ "[2005;2009]",
    annee_qc >= 2010 & annee_qc < 2015 ~ "[2010;2014]",
    annee_qc >= 2015 & annee_qc < 2020 ~ "[2015;2020]",
  ))

vis_annee_qc <- ggplot(trajipaq %>%
                filter(!is.na(VD1), !is.na(VI4), !is.na(annee_qc_f)), aes(x = VI4, y = VD1, color = as.factor(annee_qc_f))) +
  geom_jitter(color = "pink") +
  stat_smooth(method = "lm", formula = "y ~ x", se = FALSE) +
  scale_color_brewer(name = "Annee d'annee_qc au Quebec", palette = 1) +
  labs(title = "Sentiment d'appartenance et discrimination en fonction de l'annee d'annee_qc", 
       x = "Perception de la discrimination subie au cours des 12 derniers mois", 
       y = "Sentiment d'appartenance au Quebec") +
  theme_bw()
print(vis_annee_qc)

vis_annee_qc_2 <- ggplot(trajipaq %>%
                filter(!is.na(VD1), !is.na(VI4), !is.na(annee_qc_f)), aes(x = VI4, y = VD1)) +
  geom_jitter(color = "pink") +
  stat_smooth(method = "lm", 
              formula = "y ~ x", 
              se = FALSE,
              color = "purple") +
  labs(title = "Sentiment d'appartenance et discrimination en fonction de l'annee d'annee_qc", 
       x = "Perception de la discrimination subie au cours des 12 derniers mois", 
       y = "Sentiment d'appartenance au Quebec") +
  facet_wrap(~ annee_qc_f, nrow = 3) +
  theme_bw()
print(vis_annee_qc_2)

```

#### origine

```{r}

vis_origine <- ggplot(trajipaq %>%
                filter(!is.na(VD1), !is.na(VI4), !is.na(origine)), aes(x = VI4, y = VD1, color = origine)) +
  geom_jitter(color = "pink") +
  stat_smooth(method = "lm", formula = "y ~ x", se = FALSE) +
  scale_color_brewer(name = "Region d'origine", palette = 1) +
  labs(title = "Sentiment d'appartenance et discrimination en fonction de l'origine", 
       x = "Perception de la discrimination subie au cours des 12 derniers mois", 
       y = "Sentiment d'appartenance au Quebec") +
  theme_bw()
print(vis_origine)

vis_origine_2 <- ggplot(trajipaq %>%
                filter(!is.na(VD1), !is.na(VI4), !is.na(origine)), aes(x = VI4, y = VD1)) +
  geom_jitter(color = "pink") +
  stat_smooth(method = "lm", 
              formula = "y ~ x", 
              se = FALSE,
              color = "purple") +
  labs(title = "Sentiment d'appartenance et discrimination en fonction de l'origine", 
       x = "Perception de la discrimination subie au cours des 12 derniers mois", 
       y = "Sentiment d'appartenance au Quebec") +
  facet_wrap(~ origine, nrow = 3) +
  theme_bw()
print(vis_origine_2)

```

#### revenu

```{r}

vis_revenu <- ggplot(trajipaq %>%
                filter(!is.na(VD1), !is.na(VI4), !is.na(revenu)), aes(x = VI4, y = VD1, color = revenu)) +
  geom_jitter(color = "pink") +
  stat_smooth(method = "lm", formula = "y ~ x", se = FALSE) +
  scale_color_brewer(name = "revenu", palette = 1) +
  labs(title = "Sentiment d'appartenance et discrimination en fonction du revenu", 
       x = "Perception de la discrimination subie au cours des 12 derniers mois", 
       y = "Sentiment d'appartenance au Quebec") +
  theme_bw()
print(vis_revenu)

vis_revenu_2 <- ggplot(trajipaq %>%
                filter(!is.na(VD1), !is.na(VI4), !is.na(revenu)), aes(x = VI4, y = VD1)) +
  geom_jitter(color = "pink") +
  stat_smooth(method = "lm", 
              formula = "y ~ x", 
              se = FALSE,
              color = "purple") +
  labs(title = "Sentiment d'appartenance et discrimination en fonction du revenu", 
       x = "Perception de la discrimination subie au cours des 12 derniers mois", 
       y = "Sentiment d'appartenance au Quebec") +
  facet_wrap(~ revenu, nrow = 3) +
  theme_bw()
print(vis_revenu_2)

```

#### langue_mat maternelle

```{r}

vis_langue_mat <- ggplot(trajipaq %>%
                filter(!is.na(VD1), !is.na(VI4), !is.na(langue_mat)), aes(x = VI4, y = VD1, color = langue_mat)) +
  geom_jitter(color = "pink") +
  stat_smooth(method = "lm", formula = "y ~ x", se = FALSE) +
  scale_color_brewer(name = "langue_mat maternelle", palette = 1) +
  labs(title = "Sentiment d'appartenance et discrimination en fonction de la langue maternelle", 
       x = "Perception de la discrimination subie au cours des 12 derniers mois", 
       y = "Sentiment d'appartenance au Quebec") +
  theme_bw()
print(vis_langue_mat)

vis_langue_mat_2 <- ggplot(trajipaq %>%
                filter(!is.na(VD1), !is.na(VI4), !is.na(langue_mat)), aes(x = VI4, y = VD1)) +
  geom_jitter(color = "pink") +
  stat_smooth(method = "lm", 
              formula = "y ~ x", 
              se = FALSE,
              color = "purple") +
  labs(title = "Sentiment d'appartenance et discrimination en fonction de la langue maternelle", 
       x = "Perception de la discrimination subie au cours des 12 derniers mois", 
       y = "Sentiment d'appartenance au Quebec") +
  facet_wrap(~ langue_mat, nrow = 2) +
  theme_bw()
print(vis_langue_mat_2)

```

### religion

```{r}

vis_religion <- ggplot(trajipaq %>%
                filter(!is.na(VD1), !is.na(VI4), !is.na(religion)), aes(x = VI4, y = VD1, color = religion)) +
  geom_jitter(color = "pink") +
  stat_smooth(method = "lm", formula = "y ~ x", se = FALSE) +
  scale_color_brewer(name = "religion", palette = 1) +
  labs(title = "Sentiment d'appartenance et discrimination en fonction de la religion", 
       x = "Perception de la discrimination subie au cours des 12 derniers mois", 
       y = "Sentiment d'appartenance au Quebec") +
  theme_bw()
print(vis_religion)

vis_religion_2 <- ggplot(trajipaq %>%
                filter(!is.na(VD1), !is.na(VI4), !is.na(religion)), aes(x = VI4, y = VD1)) +
  geom_jitter(color = "pink") +
  stat_smooth(method = "lm", 
              formula = "y ~ x", 
              se = FALSE,
              color = "purple") +
  labs(title = "Sentiment d'appartenance et discrimination en fonction de la religion", 
       x = "Perception de la discrimination subie au cours des 12 derniers mois", 
       y = "Sentiment d'appartenance au Quebec") +
  facet_wrap(~ religion, nrow = 3) +
  theme_bw()
print(vis_religion_2)

```

### Statut statut_mat

```{r}

vis_statut_mat <- ggplot(trajipaq %>%
                filter(!is.na(VD1), !is.na(VI4), !is.na(statut_mat)), aes(x = VI4, y = VD1, color = statut_mat)) +
  geom_jitter(color = "pink") +
  stat_smooth(method = "lm", formula = "y ~ x", se = FALSE) +
  scale_color_brewer(name = "Statut statut_mat", palette = 1) +
  labs(title = "Sentiment d'appartenance et discrimination en fonction du statut statut_mat", 
       x = "Perception de la discrimination subie au cours des 12 derniers mois", 
       y = "Sentiment d'appartenance au Quebec") +
  theme_bw()
print(vis_statut_mat)

vis_statut_mat_2 <- ggplot(trajipaq %>%
                filter(!is.na(VD1), !is.na(VI4), !is.na(statut_mat)), aes(x = VI4, y = VD1)) +
  geom_jitter(color = "pink") +
  stat_smooth(method = "lm", 
              formula = "y ~ x", 
              se = FALSE,
              color = "purple") +
  labs(title = "Sentiment d'appartenance et discrimination en fonction du statut statut_mat", 
       x = "Perception de la discrimination subie au cours des 12 derniers mois", 
       y = "Sentiment d'appartenance au Quebec") +
  facet_wrap(~ statut_mat, nrow = 3) +
  theme_bw()
print(vis_statut_mat_2)

```

## H2 : VD2 entretient une relation négative avec VI4

### Modèle de régression

```{r}

trajipaq <-
  trajipaq %>%
  mutate(VD2_n = case_when(
    VD2 == "Non" ~ 0,
    VD2 == "Oui" ~ 1
  ))

reg_2 <- glm(formula = VD2_n ~ VI4, family = "binomial", data = trajipaq)

summary(reg_2)

```

### Coefficient de corrélation

```{r}

cor.test(trajipaq$VD2_n, trajipaq$VI4)

```

### Visualisation

```{r}

ggplot(trajipaq %>% filter(!is.na(VD2_n), !is.na(VI4)), aes(x = VI4, y = VD2_n)) +
  stat_smooth(method = "glm",
              formula = "y ~ x",
              color = "pink", 
              se = FALSE,
              method.args = list(family = binomial)) +
  labs(
    title = "Sentiment d'appartenance et discrimination", 
    x = "Perception de la discrimination subie au cours des 12 derniers mois", 
    y = "Sentiment d'appartenance au Canada") +
  theme_bw()

```

#### Sexe

```{r}

ggplot(trajipaq %>% filter(!is.na(VD2_n), !is.na(VI4)), aes(x = VI4, y = VD2_n)) +
  stat_smooth(method = "glm",
              formula = "y ~ x",
              color = "pink", 
              se = FALSE,
              method.args = list(family = binomial)) +
  labs(
    title = "Sentiment d'appartenance et discrimination en fonction du sexe", 
    x = "Perception de la discrimination subie au cours des 12 derniers mois", 
    y = "Sentiment d'appartenance au Canada") +
  facet_wrap(~ sexe, nrow = 2) +
  theme_bw()

```

#### Age

```{r}

ggplot(trajipaq %>% filter(!is.na(VD2_n), !is.na(VI4)), aes(x = VI4, y = VD2_n)) +
  stat_smooth(method = "glm",
              formula = "y ~ x",
              color = "pink", 
              se = FALSE,
              method.args = list(family = binomial)) +
  labs(
    title = "Sentiment d'appartenance et discrimination en fonction de l'age", 
    x = "Perception de la discrimination subie au cours des 12 derniers mois", 
    y = "Sentiment d'appartenance au Canada") +
  facet_wrap(~ age, nrow = 3) +
  theme_bw()

```

#### education

```{r}

ggplot(trajipaq %>% filter(!is.na(VD2_n), !is.na(VI4), !is.na(education)), aes(x = VI4, y = VD2_n)) +
  stat_smooth(method = "glm",
              formula = "y ~ x",
              color = "pink", 
              se = FALSE,
              method.args = list(family = binomial)) +
  labs(
    title = "Sentiment d'appartenance et discrimination en fonction de l'education", 
    x = "Perception de la discrimination subie au cours des 12 derniers mois", 
    y = "Sentiment d'appartenance au Canada") +
  facet_wrap(~ education, nrow = 3) +
  theme_bw()

```

#### Année d'arrivée au Québec

```{r}

ggplot(trajipaq %>% filter(!is.na(VD2_n), !is.na(VI4), !is.na(annee_qc_f)), aes(x = VI4, y = VD2_n)) +
  stat_smooth(method = "glm",
              formula = "y ~ x",
              color = "pink", 
              se = FALSE,
              method.args = list(family = binomial)) +
  labs(
    title = "Sentiment d'appartenance et discrimination en fonction de l'annee d'arivee", 
    x = "Perception de la discrimination subie au cours des 12 derniers mois", 
    y = "Sentiment d'appartenance au Canada") +
  facet_wrap(~ annee_qc_f, nrow = 3) +
  theme_bw()

```

#### origine

```{r}

ggplot(trajipaq %>% filter(!is.na(VD2_n), !is.na(VI4), !is.na(origine)), aes(x = VI4, y = VD2_n)) +
  stat_smooth(method = "glm",
              formula = "y ~ x",
              color = "pink", 
              se = FALSE,
              method.args = list(family = binomial)) +
  labs(
    title = "Sentiment d'appartenance et discrimination en fonction de l'origine", 
    x = "Perception de la discrimination subie au cours des 12 derniers mois", 
    y = "Sentiment d'appartenance au Canada") +
  facet_wrap(~ origine, nrow = 3) +
  theme_bw()

```

#### revenu

```{r}

ggplot(trajipaq %>% filter(!is.na(VD2_n), !is.na(VI4), !is.na(revenu)), aes(x = VI4, y = VD2_n)) +
  stat_smooth(method = "glm",
              formula = "y ~ x",
              color = "pink", 
              se = FALSE,
              method.args = list(family = binomial)) +
  labs(
    title = "Sentiment d'appartenance et discrimination en fonction du revenu", 
    x = "Perception de la discrimination subie au cours des 12 derniers mois", 
    y = "Sentiment d'appartenance au Canada") +
  facet_wrap(~ revenu, nrow = 3) +
  theme_bw()

```

#### langue_mat maternelle

```{r}

ggplot(trajipaq %>% filter(!is.na(VD2_n), !is.na(VI4), !is.na(langue_mat)), aes(x = VI4, y = VD2_n)) +
  stat_smooth(method = "glm",
              formula = "y ~ x",
              color = "pink", 
              se = FALSE,
              method.args = list(family = binomial)) +
  labs(
    title = "Sentiment d'appartenance et discrimination en fonction du revenu", 
    x = "Perception de la discrimination subie au cours des 12 derniers mois", 
    y = "Sentiment d'appartenance au Canada") +
  facet_wrap(~ langue_mat, nrow = 2) +
  theme_bw()

```

#### religion

```{r}

ggplot(trajipaq %>% filter(!is.na(VD2_n), !is.na(VI4), !is.na(religion)), aes(x = VI4, y = VD2_n)) +
  stat_smooth(method = "glm",
              formula = "y ~ x",
              color = "pink", 
              se = FALSE,
              method.args = list(family = binomial)) +
  labs(
    title = "Sentiment d'appartenance et discrimination en fonction de la religion", 
    x = "Perception de la discrimination subie au cours des 12 derniers mois", 
    y = "Sentiment d'appartenance au Canada") +
  facet_wrap(~ religion, nrow = 3) +
  theme_bw()

```

#### Statut statut_mat

```{r}

ggplot(trajipaq %>% filter(!is.na(VD2_n), !is.na(VI4), !is.na(statut_mat)), aes(x = VI4, y = VD2_n)) +
  stat_smooth(method = "glm",
              formula = "y ~ x",
              color = "pink", 
              se = FALSE,
              method.args = list(family = binomial)) +
  labs(
    title = "Sentiment d'appartenance et discrimination en fonction du statut statut_mat", 
    x = "Perception de la discrimination subie au cours des 12 derniers mois", 
    y = "Sentiment d'appartenance au Canada") +
  facet_wrap(~ statut_mat, nrow = 3) +
  theme_bw()

```

## H3 : VD3 entretient une relation négative avec VI4

### Modèle de régression

```{r}

trajipaq <-
  trajipaq %>%
  mutate(VD3_n = case_when(
    VD3 == "Non" ~ 0,
    VD3 == "Oui" ~ 1
  ))

reg_3 <- glm(formula = VD3_n ~ VI4, family = "binomial", data = trajipaq)

summary(reg_3)

```

## H4 : VD1 entretien une relation avec VD3 (en tant que variable indépendante)

```{r}

trajipaq <-
  trajipaq %>%
  mutate(VD1_f = case_when(
    VD1 < 0.5 ~ "Non",
    VD1 >= 0.5 ~ "Oui"
  ))

ctable(trajipaq$VD3, trajipaq$VD1_f)

```

```{r}

chisq.test(trajipaq$VD3, trajipaq$VD1_f)

```

### 2023-05-31 1. Trajipaq_sentiment/2. Codes/Trajipaq_sentiment_cr.Rmd
Réorganisation de l'espace de travail.

---
title: "Trajipaq_sentiment_cr"
author: "Thea"
date: "2023-05-24"
output: html_document
---

# Résumé

Les immigrants sont souvent pris entre deux ou plusieurs cultures, celles de leur pays d'origine et celles de leur nouveau pays de résidence. Cet article essaie de comprendre comment ces sentiments sont associés à leur pays d'origine et aux éventuelles situations de discriminations qu'ils vivent dans leur pays d'accueil.

# Création des variables

## Variables de contrôle

```{r}

library(summarytools)
library(tidyverse)
library(ggthemes)
library(haven)

trajipaq <- read_dta("../3. Data/Base_Trajipaq_Finale_Ok.dta")

```

### Sexe

```{r}

# Q10 = êtes-vous un homme, une femme ou autre (précisez) ?

trajipaq <-
  trajipaq %>%
  mutate(sexe = case_when( # Création de la variable "sexe".
    Q10 == 1 ~ "Homme",
    Q10 == 2 ~ "Femme",
    Q10 == 97 ~ "Autre"
  ))

freq(trajipaq$sexe)

```

### VC1 : groupe d'âge

```{r}

# Qage = à quel groupe d'âge appartenez-vous ?

trajipaq <-
  trajipaq %>%
  mutate(age = factor(case_when( # création de la variable "age".
    QAGE == 1 ~ "[0;19]",
    QAGE == 2 ~ "[20;24]",
    QAGE == 3 ~ "[25;29]",
    QAGE == 4 ~ "[30;34]",
    QAGE == 5 ~ "[35;39]",
    QAGE == 6 ~ "[40;44]",
    QAGE == 7 ~ "[45;49]",
    QAGE == 8 ~ "[50;60]",
    QAGE == 9 ~ "[61;99]",
    QAGE == 99 ~ NA
  ), levels = c( # Ordonnancement de la variable "age".
  "[0;19]",
  "[20;24]",
  "[25;29]",
  "[30;34]",
  "[35;39]",
  "[40;44]",
  "[45;49]",
  "[50;60]",
  "[61;99]",
  NA)))

freq(trajipaq$age)

```

### VC2 : niveau d'éducation

```{r}

trajipaq <-
  trajipaq %>%
  mutate(education = factor(case_when( # Création de la variable "education".
    Q16 == 1 ~ "Elementaire",
    Q16 == 2 ~ "Secondaire",
    Q16 == 3 ~ "Collegial",
    Q16 == 4 ~ "Post secondaire",
    Q16 == 5 ~ "Universitaire",
    Q26 >= 6 ~ NA
  ), levels = c( # Ordonnancement de la variable "education".
    "Elementaire", 
    "Secondaire", 
    "Post secondaire", 
    "Collegial", 
    "Universitaire")))

freq(trajipaq$education)

```

### VC3 : année d'arrivée au Québec

```{r}

# Q12 et Q14I = en quelle année êtes-vous arrivé au Québec ?

trajipaq <-
  trajipaq %>%
  mutate(Q12 = case_when( # Élimination des données manquantes pour Q12.
    Q12 == 9998 ~ NA_real_,
    Q12 == 9999 ~ NA_real_,
    TRUE ~ as.numeric(Q12)
  )) %>%
  mutate(annee_qc = case_when( # Création de la variable "annee_qc".
    as.numeric(Q14I) == 9996 ~ as.numeric(Q12), # Fusion des questions Q12 et Q14I.
    as.numeric(Q14I) == 9998 ~ NA_real_,
    as.numeric(Q14I) == 9999 ~ NA_real_,
    TRUE ~ as.numeric(Q14I)
  ))

descr(trajipaq$annee_qc)

```

### VC4 : région d'origine

```{r}

# region_naissance = de quelle région du monde êtes-vous originaire ?

trajipaq <-
  trajipaq %>%
  mutate(origine = case_when( # Création de la variable "region".
    region_naissance == "ANMO" ~ "Afrique du Nord et Moyen Orient",
    region_naissance == "ACAS" ~ "Amerique centrale et Amerique su Sud",
    region_naissance == "Afrique subsaharienne" ~ "Afrique subsaharienne",
    region_naissance == "Asie" ~ "Asie",
    region_naissance == "Autres" ~ "Autres",
    region_naissance == "Canada" ~ "Canada",
    region_naissance == "Caraibes" ~ "Caraibes",
    region_naissance == "Europe" ~ "Europe",
    region_naissance == "France" ~ "France",
    region_naissance == "NSP" ~ NA,
  ))

freq(trajipaq$origine)

```

### VC5 : revenu annuel brut sur l'année 2019

```{r}

# Q100 = Au meilleur de votre connaissance, à combien estimez–vous le revenu total que vous gagnez (aviez gagné), provenant de toutes les sources, avant impôts et autres déductions, au cours de l’année se terminant le 31 décembre 2019 ?

trajipaq <-
  trajipaq %>%
  mutate(revenu = factor(case_when( # Création de la variable "revenu".
    Q100 == 1 ~ "[0 ; 19 999]",
    Q100 == 2 ~ "[0 ; 19 999]",
    Q100 == 3 ~ "[20 000 ; 39 999]",
    Q100 == 4 ~ "[20 000 ; 39 999]",
    Q100 == 5 ~ "[40 000 ; 59 999]",
    Q100 == 6 ~ "[40 000 ; 59 999]",
    Q100 == 7 ~ "[60 000 ; 79 999]",
    Q100 == 8 ~ "[60 000 ; 79 999]",
    Q100 == 9 ~ "[80 000 ; 99 999]",
    Q100 == 10 ~ "[90 000 ; 99 999]",
    Q100 == 11 ~ "[100 000 ; 149 999]",
    Q100 == 12 ~ "[150 000 et plus]",
    Q100 == 98 ~ NA,
    Q100 == 99 ~ NA
  ), levels = c(
  "[0 ; 19 999]",
  "[20 000 ; 39 999]",
  "[40 000 ; 59 999]",
  "[60 000 ; 79 999]",
  "[80 000 ; 99 999]",
  "[100 000 ; 149 999]",
  "[150 000 et plus]",
  NA)))

freq(trajipaq$revenu)

```

### VC6 : langue maternelle

```{r}

# Q20 = Quand vous étiez enfant, quelles étaient la langue_mat ou les langue_mats que vous parliez le plus souvent, à l’âge de 5 ans et que vous comprenez toujours ?

trajipaq <-
  trajipaq %>%
  mutate(langue_mat = case_when( # Création de la variable "langue_mat_mat".
    Q20M1 == 1 ~ "Francais",
    Q20M1 == 2 ~ "Anglais",
    Q20M1 == 97 ~ "Autre",
    Q20M1 >= 98 ~ NA
  ))

freq(trajipaq$langue_mat)

```

### VC7 : confession religieuse

```{r}

# Q17 = Actuellement, avez-vous une religion ?

trajipaq <-
  trajipaq %>%
  mutate(religion = case_when( # Création de la variable "religion".
    Q17 == 0 ~ "Sans religion",
    Q17 == 1 ~ "Catholique",
    Q17 == 2 ~ "Protestante",
    Q17 == 3 ~ "Autres Chretiens",
    Q17 == 4 ~ "Autres Chretiens",
    Q17 == 5 ~ "Autres Chretiens",
    Q17 == 6 ~ "Orthodoxe",
    Q17 == 7 ~ "Autres Chretiens",
    Q17 == 8 ~ "Musulmans",
    Q17 == 9 ~ "Autre",
    Q17 == 10 ~ "Autre",
    Q17 == 11 ~ "Autre",
    Q17 == 12 ~ "Autre",
    Q17 == 97 ~ "Autre",
    Q17 >= 98 ~ NA
  ))

freq(trajipaq$religion)

```

### VC8 : statut matrimonial

```{r}

# STATUT = actuellement, vous êtes... ?

trajipaq <-
  trajipaq %>%
  mutate(statut_mat = case_when( # Création de la variable "statut_mat".
    STATUT == 1 ~ "Celibataire",
    STATUT == 2 ~ "Marie(e)",
    STATUT == 3 ~ "Conjoint(e) de fait",
    STATUT == 4 ~ "Separe(e)",
    STATUT == 5 ~ "Divorce(e)",
    STATUT == 6 ~ "Veuf(ve)"
  ))

freq(trajipaq$statut_mat)

```

## Variables dépendantes

### VD1 : sentiment d'appartenance au Québec

```{r}

# Préparation des variables en vue de créer le premier indice.

trajipaq <-
  trajipaq %>% 
  mutate(Q39_new_A = Q39A, # Création de nouvelles variables.
         Q39_new_B = Q39B,
         Q39_new_C = Q39C,
         Q39_new_D = Q39D) %>% 
  mutate_at(
    vars(starts_with("Q39_new")),
    funs(case_when( # Ordonnancement des nouvelles variables.
      . == 1 ~ 4,
      . == 2 ~ 3,
      . == 3 ~ 2,
      . == 4 ~ 1,
      . == 98 ~ NA,
      . == 99 ~ NA
    ))
  )

v_VD1 <- c( # Création d'un vecteur regroupant les nouvelles variables.
  "Q39_new_A",
  "Q39_new_B",
  "Q39_new_C",
  "Q39_new_D"
)

nbre_lvl <- 4 # Création d'un vecteur regroupant le nombre de choix possibles dans les questions posées.

trajipaq <-
  trajipaq %>% 
  rowwise() %>% 
  mutate (nbre_rep_rep = sum(!is.na(across(starts_with("Q39_new"))))) # Création d'un vecteur comptant les données invalides.

# Création de l'indice du sentiment d'appartenance au Québec.

trajipaq <-
  trajipaq %>%
  rowwise() %>%
  mutate(sumVD1 = sum(c_across(all_of(v_VD1)), na.rm = TRUE)) %>%
  mutate(VD1 = (sumVD1 - nbre_rep_rep) / ((nbre_lvl - 1)*nbre_rep_rep)) %>%
  mutate(VD1 = if (nbre_rep_rep == 0) NA_real_ else VD1) %>%
  ungroup()

descr(trajipaq$VD1)

```

```{r}

ggplot(data = trajipaq, mapping = aes(x = VD1)) +
  geom_histogram(na.rm = TRUE, fill = "pink", binwidth = 0.01) +
  labs(
    title = "Distribution du sentiment d'appartenance au Quebec",
    x = "Indice du sentiment d'appartenance au Quebec"
       ) +
  theme_bw()

```

### VD2 : sentiment d'appartenance (Canada)

```{r}

# Q39E = Je me sens canadienne.

trajipaq <-
  trajipaq %>%
  mutate(VD2 = case_when( # Création de la variable "VD2".
    Q39E == 1 ~ "Oui",
    Q39E == 2 ~ "Oui",
    Q39E == 3 ~ "Non",
    Q39E == 4 ~ "Non",
    Q39E == 98 ~ NA,
    Q39E == 99 ~ NA
  ))

freq(trajipaq$VD2)

```

### VD3 : sentiment d'appartenance au pays d'origine

```{r}

# Q39F = ’ai un sentiment d’appartenance à mon pays de naissance.

trajipaq <-
  trajipaq %>%
  mutate(VD3 = case_when( # Création de la variable "VD3".
    Q39F == 1 ~ "Oui",
    Q39F == 2 ~ "Oui",
    Q39F == 3 ~ "Non",
    Q39F == 4 ~ "Non",
    Q39F >= 98 ~ NA
  ))

freq(trajipaq$VD3)

```

### VD4 : sentiment d'appartenance au Canada comparativement au pays d'origine

```{r}

trajipaq <-
  trajipaq %>% 
  mutate(VD4 = case_when( # Création de la variable "VD4".
    VD2 == "Oui" & VD3 == "Oui" ~ "Integre",
    VD2 == "Oui" & VD3 == "Non" ~ "Assimile",
    VD2 == "Non" & VD3 == "Oui" ~ "Separe",
    VD2 == "Non" & VD3 == "Non" ~ "Marginalise"
  ))

freq(trajipaq$VD4)

```

### VD5 : sentiment d'appartenance au Canada comparativement au Québec

```{r}

trajipaq <-
  trajipaq %>% 
  mutate(Q39D_new_2 = case_when( # Création d'une nouvelle variable.
         Q39D <= 2 ~ "Oui", 
         Q39D == 3 | Q39D == 4 ~ "Non", 
         Q39D >= 97 ~ NA_character_)) %>% 
  mutate(VD5 = factor(case_when( # Création de la variable "VD5".
    VD2 == "Oui" & Q39D_new_2 == "Oui" ~ 1,
    VD2 == "Oui" & Q39D_new_2 == "Non" ~ 2,
    VD2 == "Non" & Q39D_new_2 == "Oui" ~ 3,
    VD2 == "Non" & Q39D_new_2 == "Non" ~ 4
  ),
  labels = c( # Création du nom des catégories.
    "Canada et Quebec", 
    "Canada seule", 
    "Quebec seule", 
    "Aucun"),
  levels = c( # Ordonnancement des catégories.
    1, 
    2, 
    3, 
    4)))

freq(trajipaq$VD5)

```

## Variables indépendantes

### VI4 : perception des discriminations subies au cours des 12 derniers mois

```{r}

# Préparation des variables en vue de créer le deuxième indice.

trajipaq <-
  trajipaq %>% 
  mutate(Q40_new_A = Q40A, # Création de nouvelles variables.
         Q40_new_B = Q40B,
         Q40_new_C = Q40C,
         Q40_new_D = Q40D) %>% 
  mutate_at(
    vars(starts_with("Q40_new")),
    funs(case_when( # Ordonnancement des nouvelles variables.
      . == 1 ~ 3,
      . == 2 ~ 2,
      . == 3 ~ 1,
      . == 4 ~ 1,
      . >= 96 ~ NA
    ))
  )

v_VI4 <- c( # Création d'un vecteur regroupant les nouvelles variables.
  "Q40_new_A",
  "Q40_new_B",
  "Q40_new_C",
  "Q40_new_D"
)

nbre_lvl_2 <- 3 # Création d'un vecteur regroupant le nombre de choix possibles dans les questions posées.

trajipaq <-
  trajipaq %>% 
  rowwise() %>% 
  mutate (nbre_rep_rep_2 = sum(!is.na(across(starts_with("Q40_new"))))) # Création d'un vecteur comptant les données invalides.

# Création de l'indice de la perception des discriminations subies au cours des 12 derniers mois.

trajipaq <-
  trajipaq %>%
  rowwise() %>%
  mutate(sumVI4 = sum(c_across(all_of(v_VI4)), na.rm = TRUE)) %>%
  mutate(VI4 = (sumVI4 - nbre_rep_rep_2) / ((nbre_lvl_2 - 1)*nbre_rep_rep_2)) %>%
  mutate(VI4 = if (nbre_rep_rep_2 == 0) NA_real_ else VI4) %>%
  ungroup()

descr(trajipaq$VI4)

```

```{r}

ggplot(data = trajipaq, mapping = aes(x = VI4)) +
  geom_histogram(na.rm = TRUE, fill = "pink", binwidth = 0.01) +
  labs(
    title = "Distribution de la discrimination subie au cours des 12 derniers mois",
    x = "Indice de la discrimination subie au cours des 12 derniers mois"
       ) +
  theme_bw()

```

# Création de la nouvelle base de données

```{r}

Trajipaq_sentiment_cr <-
  trajipaq %>%
  select("age", # Création de la base de données dédiée à l'analyse.
         "education",
         "sexe",
         "annee_qc",
         "origine",
         "revenu",
         "langue_mat",
         "religion",
         "statut_mat",
         "VD1",
         "VD2",
         "VD3",
         "VD4",
         "VD5",
         "VI4")

write.csv(Trajipaq_sentiment_cr, "../3. Data/Trajipaq_sentiment_cr.csv") # Ecriture de la nouvelle base de données.

```

### 2023-05-31 2. Trajipaq_covid/2. Codes/Trajipaq_covid_cr1.Rmd
Réorganisation de l'espace de travail.

---
title: "Familial consequences of the Covid-19 in Québec: Comparison between immigrants and natives"
subtitle: "Conséquences familiales de la Covid-19 au Québec: comparaison immigrants - natifs"
author:
- Vissého Adjiwanou ^[corresponding author - adjiwanou.vissého@uqam.ca]
- Département de Sociologie, Université du Québec à Montréal (UQAM)
- Solène Lardoux (Département de démographie, Université de Montréal)
- Nima Zahedinameghi (Département de Sociologie, UQAM)
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  word_document: default
  html_document: 
    toc: true
    toc_depth: 3
  template: svm-latex-ms2.tex 
  pdf_document: default
  citation_package: natbib
bibliography: Covid_Canada.bib
---

## Ouverture de la base de données

```{r, echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
rm(list = ls())

if (!require("pacman")) install.packages("pacman")
pacman::p_load(tidyverse, # for tidyverse 
               ggpubr, # for arranging ggplots   
               ggthemes, # for fancy ggplot themes
               here, # for reproducibility 
               patchwork, # for easy ggarrange
               ggsci, # for pubs 
               fastDummies, # to create dummy variables fast
            #   readtext, # for reading text
            #   quanteda, # for text preprocessing 
            #   data.table, # for fast data manipulation
            #   stm, # for structural topic modeling
             #  future, # for parallel and distributed computing      
             #  purrr, # for functional programming 
            #   keyATM, # keyATM
            #   latex2exp,
               summarytools,
               hrbrthemes, 
            #   rvest,
               wesanderson,
            #   tidytext, 
            #   textstem,
            #   stringr, 
            #   SnowballC, 
            #   googleLanguageR,
            #   wordcloud, 
            #   wordcloud2, 
            #   webshot, 
            #   htmlwidgets, 
            #   reshape2, 
            #   topicmodels,
               lubridate,
            gtsummary,
            haven
             #  cli,
             #  cld2
            )
#library(devtools)
#install_github("hemken/Statamarkdown")   # Ne fonctionne pas
#install.packages("https://www.ssc.wisc.edu/~hemken/Stataworkshops/Stata%20and%20R%20Markdown/Statamarkdown_0.5.5.tar.gz", type="source", repos=NULL)

#library(Statamarkdown)

# for publication-friendly theme 
theme_set(theme_pubr())

## My preferred ggplot2 plotting theme (optional)
theme_set(hrbrthemes::theme_ipsum())

# For keyword based topic modeling (development version)
#devtools::install_github("keyATM/keyATM", ref = "package_dev", force = TRUE)

```

## Ouverture de la base de données

```{r}

#trajipaq <- read_sav("../Data/Trajipaq_Total_19032021_Final.sav") # Ma dernière version - probleme
#trajipaq <- read_sav("../Data/Trajipaq Total 19032021.sav")    # Ma dernière version - probleme
#trajipaq <- read_sav("../Data/Trajipaq_Total_17032021_final.sav") # dernier version chez Georges - probleme
trajipaq <- readxl::read_xlsx("../Data/Trajipaq Total 19032021.xlsx")

#write_dta(trajipaq, "Trajipaq_Total_19032021_Final.dta")

```

```{stata}

#use "Trajipaq_Total_19032021_Final.dta", clear
#ta

#sysuse auto
#summarize
```

# Création des variables

## Variables indépendantes:

### Statut d'immigrants

```{r}

# Statut d'immigration

freq(trajipaq$Q99)

trajipaq <-
  trajipaq %>% 
  mutate(statut_imm = factor(if_else(Q4 == 1, "Immigrant", "Non-immigrant")))

freq(trajipaq$statut_imm)

# Région d'origine

freq(trajipaq$REGIX)

# Durée de résidence

freq(trajipaq$Q14I)
freq(trajipaq$REGIO)

trajipaq <-
  trajipaq %>% 
  mutate(duree_res1 = if_else(Q14I == 9996, Q12, Q14I),
         duree_res = factor(case_when(
           is.na(duree_res1) ~ 1,
           duree_res1 >= 2016 & duree_res1 <= 2021 ~ 2,
           duree_res1 >= 2011 & duree_res1 <= 2015 ~ 3,
           duree_res1 >= 1965 & duree_res1 <= 2010 ~ 4), labels = c("Born in Canada", "Less than 5 years", "5-10", "More than 10 years")))
freq(trajipaq$duree_res)

# Lieu de résidence

```

- Minorité visible

```{r}

freq(trajipaq$Q99O)
pmv <- c("Caucasian",  "Caucasien European", "Européen", "Européen, cocasien", "Européenne", "Européens",  "Italienne", "Je suis Ukrainienne", "Jewish", "Mes racines sont européenes mais je suis née en Argentine et grandit là-bas.", "Nord-Américaine",  "Québécoise", "Russe", "Russian", "jewish", " white and lebanese")

trajipaq <-
  trajipaq %>% 
  mutate(mnv = if_else(is.na(Q99O), NA_real_, 
                       if_else(Q99O %in% pmv, 1, 2)))
freq(trajipaq$mnv)

with(trajipaq,
     ctable(Q99M1, mnv, useNA = "ifany", "no"))

# Minorité visible

freq(trajipaq$Q99M1)

trajipaq <-
  trajipaq %>% 
  mutate(minorite_vis = factor(case_when(
    mnv == 1 ~ 1,
    mnv == 2 ~ 2,
    Q99M1 == 0 ~ 1,
    Q99M1 >= 1 & Q99M1 <= 12 ~ 2,
    Q99M1 == 98 | Q99M1 == 99 ~ 3),
    #labels = c("Pas minorité visible", "Minorité visible", "Non réponse"
    labels = c("Not a visible minority", "Visible minority", "No response")))

trajipaq <-
  trajipaq %>% 
  mutate(minorite_vis = replace_na(minorite_vis, "No response"))

freq(trajipaq$minorite_vis)

?recode

```

- Statut d'immigration et minorité visible

```{r}

ctable(trajipaq$statut_imm, trajipaq$minorite_vis)

trajipaq <-
  trajipaq %>% 
  mutate(imm_minorite = factor(case_when(
    statut_imm == "Immigrant" & minorite_vis == "Not a visible minority" ~ "Non-racialised immigrant",
    statut_imm == "Immigrant" & minorite_vis == "Visible minority" ~ "Racialised immigrant",
    statut_imm == "Non-immigrant" & minorite_vis == "Not a visible minority" ~ "Majority group",
    TRUE ~ "Others")))

ctable(trajipaq$imm_minorite, trajipaq$minorite_vis, "no")
ctable(trajipaq$imm_minorite, trajipaq$statut_imm, "no")

```

- durée de Résidence

```{r}

freq(trajipaq$Q14I)

trajipaq <- trajipaq %>% 
  mutate(length_stay_t = if_else(Q14I == 9996, Q12, Q14I),
         length_stay = factor(case_when(
           is.na(length_stay_t) ~ 4,
           length_stay_t >= 9998 ~ NA_real_,
           length_stay_t >= 2016 ~ 1, 
           length_stay_t >= 2011 & length_stay_t <=2015 ~ 2,
           length_stay_t >= 1965 & length_stay_t <= 2010 ~ 3),
           labels = c("Less than 5 years", "5-10 years", "More than 10 years", "Born here")))

trajipaq <- trajipaq %>% 
  mutate(length_stay_t = if_else(Q14I == 9996, Q12, Q14I),
         length_stay_2 = factor(case_when(
           is.na(length_stay_t) ~ 3,
           length_stay_t >= 9998 ~ NA_real_,
           length_stay_t >= 2016 ~ 1, 
           length_stay_t >= 1965 & length_stay_t <=2015 ~ 2),
           labels = c("Less than 5 years", "More than 5 years", "Born here")))

freq(trajipaq$length_stay_2)

```

### 2. Caractéristiques socio-démographiques

```{r}

freq(trajipaq$Q79)

trajipaq <-
  trajipaq %>% 
  mutate(sexe = factor(if_else(Q10 == 1, "Male", "Female")),
         gpage = factor(case_when(
           QAGE <= 4 ~ 1,
           QAGE >= 5 & QAGE <= 6 ~ 2,
           QAGE >= 7 ~ 3), labels = c("20-34", "35-44", "45-60")),
         niv_etude = factor(case_when(
           Q16 <= 2 ~ 1,
           Q16 >=3 & Q16 <= 4 ~ 2,
           Q16 >=5 & Q16 <= 6 ~ 3,
           Q16 >90 ~ NA_real_), 
           #labels = c("Secondaire ou moins", "Collegial ou fp", "Université")),
         labels = c("Secondary or less", "Collegial or professional training", "University")),
         statut_mat = factor(case_when(
           STATUT == 1 ~ 1,
           STATUT == 2 ~ 2,
           STATUT == 3 ~ 3,
           STATUT >=4 & STATUT <= 6 ~ 4), 
           #labels = c("Célibataire", "Marié", "Conjoint de fait", "Séparé, divorcé, veuf.ve")),
         labels = c("Single", "Maried", "Living together", "Separed, Divorced, Widow")),
         statut_emploi = factor(case_when(
           Q79 == 1 | Q79 == 10 ~ 1,
           Q79 == 4 | Q79 == 9 ~ 2,
           Q79 == 2 | Q79 == 3 | Q79 == 97 ~ 3, 
           Q79 >= 5 & Q79 <= 8 ~ 3,
           Q79 >=98 ~ NA_real_), 
           #labels = c("En emploi", "En chômage", "Autres")))
           labels = c("Employed", "Unemployed", "Others")))

## Réligion

# Vérification

with(trajipaq,
     ctable(Q10, sexe, "no"))

with(trajipaq,
     ctable(QAGE, gpage, "no"))

with(trajipaq,
     ctable(Q16, niv_etude, "no"))

with(trajipaq,
     ctable(STATUT, statut_mat, "no"))

with(trajipaq,
     ctable(Q79, statut_emploi, "no"))

```

- Présence d'enfants dans le ménage

* Enfants de moins de 4 ans vivant avec la personne (3 categories)

<!--   Code Stata de Houlio
ge kid_4 = .
replace kid_4 = 1 if Q74_1 < 4 & Q76_1 <= 2
replace kid_4 = 2 if Q74_1 <4 & Q76_1 >= 3
replace kid_4 = 2 if Q74_1 >= 4 & Q74_1 <=40
replace kid_4 = 0 if Q74_1 >= 48
replace kid_4 = . if Q74_1 == 48 | Q74_1 == 99

la def kid_4 1"Oui" 2"Non" 0"Pas d'enfants" ,modify
la var kid_4 "Présence enfant de moins de 4 ans"	
la val kid_4 kid_4
ta kid_4, m
ta kid_4
ta kid_4[iw = poids_final]
browse Q74_1 Q72 Q76_1 kid_4*/
***************************************************
* Enfants de moins de 4 ans vivant avec la personne (2 categories)

ge kid_4a = .
replace kid_4a = 1 if Q74_1 < 4 & Q76_1 <= 2
replace kid_4a = 0 if Q74_1 <4 & Q76_1 >= 3
replace kid_4a = 0 if Q74_1 >= 4 & Q74_1 <=40
replace kid_4a = 0 if Q74_1 >= 48
replace kid_4a = . if Q74_1 == 48 | Q74_1 == 99

la def kid_4a 1"Yes" 0"No" ,replace
la var kid_4a "Children under 4 years in household"	
la val kid_4a kid_4a
ta kid_4a, m
ta kid_4a
-->

## Variables dépendantes

### 1. Exposition à la maladie

```{r}

# Un membre de la famille a continué à travailler en dehors du domicile (personnel essentiel)
oui_non <- c("non", "oui", "non reponse")
oui_non_2 <- c("non", "oui", "Non reponse", "Non reponse")

freq(trajipaq$COV8)
trajipaq <-
  trajipaq %>% 
  mutate(e_hors_dom = factor(case_when(
    COV8 == 1 ~ 1,
    COV8 == 2 ~ 0,
    COV8 == 4 ~ 4,
    COV8 == 98 ~ NA_real_), labels = oui_non))

freq(trajipaq$e_hors_dom)

# Un membre de la famille est un fournisseur de soins de santé / premier répondant fournissant des soins directs

freq(trajipaq$COV9)
trajipaq <-
  trajipaq %>% 
  mutate(e_four_soin = factor(case_when(
    COV9 == 1 ~ 1,
    COV9 == 2 ~ 0,
    COV9 == 4 ~ 4,
    COV9 == 98 ~ NA_real_), labels = oui_non))

freq(trajipaq$e_four_soin)

# Nous nous sommes mis en quarantaine en raison de voyages ou d'une exposition possible

freq(trajipaq$COV14)
trajipaq <-
  trajipaq %>% 
  mutate(e_quarantaine = factor(case_when(
    COV14 == 1 ~ 1,
    COV14 == 2 ~ 0,
    COV14 == 4 ~ 4,
    COV14 == 98 ~ NA_real_), labels = oui_non))

freq(trajipaq$e_quarantaine)

# Un membre de la famille a été exposé à quelqu'un avec COVID-19? Si "Oui", Veuillez préciser qui?  (par exemple moi-même, mon enfant, mon conjoint, mes parents, etc.)

freq(trajipaq$COV21) 
trajipaq <-
  trajipaq %>% 
  mutate(e_expose = factor(case_when(
    COV21 == 1 ~ 1,
    COV21 == 2 ~ 0,
    COV21 == 98 | COV21 == 99  ~ 4), labels = oui_non))

freq(trajipaq$e_expose)

# Un membre de la famille a présenté des symptômes ou a reçu un diagnostic de COVID-19? Si "Oui", Veuillez préciser qui?  (par exemple moi-même, mon enfant, mon conjoint, mes parents, etc.)

freq(trajipaq$COV22)
trajipaq <-
  trajipaq %>% 
  mutate(e_symptome = factor(case_when(
    COV22 == 1 ~ 1,
    COV22 == 2 ~ 0,
    COV22 == 98 | COV22 == 99  ~ 4), labels = oui_non))

freq(trajipaq$e_symptome)

# Un membre de la famille a été hospitalisé pour COVID-19? Si "Oui", Veuillez préciser qui?  (par exemple moi-même, mon enfant, mon conjoint, mes parents, etc.)

freq(trajipaq$COV23)
trajipaq <-
  trajipaq %>% 
  mutate(e_hospitalise = factor(case_when(
    COV23 == 1 ~ 1,
    COV23 == 2 ~ 0,
    COV23 == 98 | COV23 == 99  ~ 4), labels = oui_non))

freq(trajipaq$e_hospitalise)

# Un membre de la famille était dans l'unité de soins intensifs (USI) pour COVID-19? Si "Oui", Veuillez préciser qui?  (par exemple moi-même, mon enfant, mon conjoint, mes parents, etc.)

freq(trajipaq$COV24)
trajipaq <-
  trajipaq %>% 
  mutate(e_intensif = factor(case_when(
    COV24 == 1 ~ 1,
    COV24 == 2 ~ 0,
    COV24 == 98 | COV24 == 99  ~ 4), labels = oui_non))

freq(trajipaq$e_intensif)

# Un membre de la famille est décédé de COVID-19? Si "Oui", Veuillez préciser qui?  (par exemple moi-même, mon enfant, mon conjoint, mes parents, etc.)

freq(trajipaq$COV25)  
trajipaq <-
  trajipaq %>% 
  mutate(e_decede = factor(case_when(
    COV25 == 1 ~ 1,
    COV25 == 2 ~ 0,
    COV25 == 98 | COV25 == 99  ~ 4), labels = oui_non))

freq(trajipaq$e_intensif)

#exposition <- c(e_hors_dom, e_four_soin, e_quarantaine, e_expose, e_symptome, e_hospitalise, e_intensif, e_decede)

```

### 2. Variables sur les conséquences de la maladie

---- Famille -----

- Notre famille vivait séparément pour des raisons de santé, de sécurité ou d'emploi
- Quelqu'un a emménagé (ou est revenu) dans notre maison
- Nous avons dû quitter notre domicile
- Nous avons eu de la difficulté à obtenir de la nourriture
- Nous avons eu du mal à obtenir des médicaments
- Nous avons eu de la difficulté à obtenir des soins de santé quand nous en avions besoin
- Notre revenu familial a diminué
- Nous avons perdu l'assurance maladie / prestations
- Nous avons manqué un évènement familial important ou il a été annulé (par exemple, mariage, remise des diplômes, naissance, funérailles, voyages (y compris les vacances), autres
- Nous avons eu de la difficulté à obtenir d'autres éléments essentiels?

```{r}
freq(trajipaq$COV5)  
trajipaq <-
  trajipaq %>% 
  mutate(f_separe = factor(case_when(
    COV5 == 1 ~ 1,
    COV5 == 2 ~ 0,
    COV5 == 98 ~ 4), labels = oui_non),
    f_emmenage = factor(case_when(
      COV6 == 1 ~ 1,
      COV6 == 2 ~ 0,
      COV6 == 98 ~ 4), labels = oui_non),
    f_quitte_dom = factor(case_when(
      COV7 == 1 ~ 1,
      COV7 == 2 ~ 0,
      COV7 == 98 ~ 4), labels = oui_non),
    f_nourriture = factor(case_when(
      COV10 == 1 ~ 1,
      COV10 == 2 ~ 0,
      COV10 == 98 ~ 4), labels = oui_non),
    f_medicaments = factor(case_when(
      COV11 == 1 ~ 1,
      COV11 == 2 ~ 0,
      COV11 == 98 ~ 4), labels = oui_non),
    f_sante = factor(case_when(
      COV12 == 1 ~ 1,
      COV12 == 2 ~ 0,
      COV12 == 98 ~ 4), labels = oui_non),
    f_revenu = factor(case_when(
      COV15 == 1 ~ 1,
      COV15 == 2 ~ 0,
      COV15 == 98 ~ 4), labels = oui_non),
    f_assurance = factor(case_when(
      COV19 == 1 ~ 1,
      COV19 == 2 ~ 0,
      COV19 == 98 ~ 4), labels = oui_non),
    f_evenement = factor(case_when(
      COV20 == 1 ~ 1,
      COV20 == 2 ~ 0,
      COV20 == 98 ~ 4), labels = oui_non),
    f_essentiel = factor(case_when(
      COV13 == 1 ~ 1,
      COV13 == 2 ~ 0,
      COV13 == 98 ~ 4), labels = oui_non),
          )
freq(trajipaq$f_emmenage)

#famille <- c("COV5", "COV6") COV5 - COV12
#for (fa in famille) {
#  trajipaq <-
#  trajipaq %>% 
#  mutate(paste(f_,fa)  = factor(case_when(
#    COV25 == 1 ~ 1,
#    COV25 == 2 ~ 0,
#    COV25 == 98 ~ 4), labels = oui_non))
#}

freq(trajipaq$COV19)

```

------- Emploi -------

- Un membre de la famille a dû réduire ses heures de travail? Si "Oui", Veuillez préciser qui? (par exemple moi-même, mon enfant, mon conjoint, mes parents, etc.)

- Un membre de la famille devait cesser de travailler (s'attendre à être rappelé)? Si "Oui", Veuillez préciser qui?  (par exemple moi-même, mon enfant, mon conjoint, mes parents, etc.)

- Un membre de la famille a perdu son emploi de façon permanente? Si "Oui", Veuillez préciser qui?  (par exemple moi-même, mon enfant, mon conjoint, mes parents, etc.)

```{r}

freq(trajipaq$COV18)
trajipaq <-
  trajipaq %>% 
  mutate(f_reduire = factor(case_when(
    COV16 == 1 ~ 1,
    COV16 == 2 ~ 0,
    COV16 == 98 | 99 ~ 4), labels = oui_non),
    f_cesser = factor(case_when(
      COV17 == 1 ~ 1,
      COV17 == 2 ~ 0,
      COV17 == 98 | 99 ~ 4), labels = oui_non),
    f_perdu = factor(case_when(
      COV18 == 1 ~ 1,
      COV18 == 2 ~ 0,
      COV18 == 98 | 99 ~ 4), labels = oui_non))

freq(trajipaq$f_reduire)

```

--- Entente dans la famille ------
La COVID-19 peut avoir de nombreux impacts sur vous et votre vie de famille. En général, comment la pandémie de COVID-19 a-t-elle affecté chacun des éléments suivants?

- COV26. Le rôle de parent
- Comment les membres de la famille s'entendent
- Capacité de prendre soin de votre enfant avec une maladie/condition
- Capacité de prendre soin d'autres enfants de votre famille
- Capacité de prendre soin des personnes âgées ou des personnes handicapées de votre famille

```{r}

#label2 <- c("Beaucoup amélioré", "Un peu amélioré", "Aucun changement", "Un peu dégradé", "Beaucoup dégradé", "Ne s'applique pas")

label2 <- c("Much improved", "A little improved", "No change", "A little degraded", "Much degraded", "Not applicable")

freq(trajipaq$COV26)
 trajipaq <-
  trajipaq %>% 
  mutate(f_role_parent = factor(case_when(
    COV26 == 0 ~ 1,
    COV26 == 1 ~ 2,
    COV26 == 2 ~ 3,
    COV26 == 3 ~ 4,
    COV26 == 4 ~ 5,
    COV26 == 96 ~ NA_real_), levels = c(1:6),  labels = label2),
    f_entende = factor(case_when(
      COV27 == 0 ~ 1,
      COV27 == 1 ~ 2,
      COV27 == 2 ~ 3,
      COV27 == 3 ~ 4,
      COV27 == 4 ~ 5,
      COV27 == 96 ~ NA_real_), levels = c(1:6),  labels = label2),
    f_enfant_maladie = factor(case_when(
      COV28 == 0 ~ 1,
      COV28 == 1 ~ 2,
      COV28 == 2 ~ 3,
      COV28 == 3 ~ 4,
      COV28 == 4 ~ 5,
      COV28 == 96 ~ NA_real_), levels = c(1:6),  labels = label2),
    f_autre_enfant = factor(case_when(
      COV29 == 0 ~ 1,
      COV29 == 1 ~ 2,
      COV29 == 2 ~ 3,
      COV29 == 3 ~ 4,
      COV29 == 4 ~ 5,
      COV29 == 96 ~ NA_real_), levels = c(1:6),  labels = label2),
    f_pers_agees = factor(case_when(
      COV30 == 0 ~ 1,
      COV30 == 1 ~ 2,
      COV30 == 2 ~ 3,
      COV30 == 3 ~ 4,
      COV30 == 4 ~ 5,
      COV30 == 96 ~ NA_real_), levels = c(1:6),  labels = label2))

freq(trajipaq$f_role_parent)
ctable(trajipaq$f_role_parent,trajipaq$COV26,  "no")

```

--- Bien-être ------

La COVID-19 peut avoir de nombreux impacts sur vous et votre vie de famille. En général, comment la pandémie de COVID-19 a-t-elle affecté chacun des éléments suivants?

- Votre bien-être physique - exercice
- Votre bien-être physique - manger
- Votre bien-être physique - dormir
- Votre bien-être émotionnel - anxiété
- Votre bien-être émotionnel - humeur

Dans l'ensemble, dans quelle mesure avez-vous ressenti de la détresse liée à la COVID-19? (échelle de 1 à 10)

```{r}

freq(trajipaq$COV31)
trajipaq <-
  trajipaq %>% 
  mutate(be_exercice = factor(case_when(
    COV31 == 0 ~ 1,
    COV31 == 1 ~ 2,
    COV31 == 2 ~ 3,
    COV31 == 3 ~ 4,
    COV31 == 4 ~ 5,
    COV31 == 96 ~ NA_real_), levels = c(1:6),  labels = label2),
    be_manger = factor(case_when(
      COV32 == 0 ~ 1,
      COV32 == 1 ~ 2,
      COV32 == 2 ~ 3,
      COV32 == 3 ~ 4,
      COV32 == 4 ~ 5,
      COV32 == 96 ~ NA_real_), levels = c(1:6),  labels = label2),
    be_dormir = factor(case_when(
      COV33 == 0 ~ 1,
      COV33 == 1 ~ 2,
      COV33 == 2 ~ 3,
      COV33 == 3 ~ 4,
      COV33 == 4 ~ 5,
      COV33 == 96 ~ NA_real_), levels = c(1:6),  labels = label2),
    be_anxiete = factor(case_when(
      COV34 == 0 ~ 1,
      COV34 == 1 ~ 2,
      COV34 == 2 ~ 3,
      COV34 == 3 ~ 4,
      COV34 == 4 ~ 5,
      COV34 == 96 ~ NA_real_), levels = c(1:6),  labels = label2),
    be_humeur = factor(case_when(
      COV35 == 0 ~ 1,
      COV35 == 1 ~ 2,
      COV35 == 2 ~ 3,
      COV35 == 3 ~ 4,
      COV35 == 4 ~ 5,
      COV35 == 96 ~ NA_real_), levels = c(1:6),  labels = label2))

```

### 3. Variables sur l'aide du gouvernement

La troisième variable dépendante cerne l'appui du gouvernement provincial et fédéral aux familles. Elle cherche à savoir si les familles ont eu à bénéficier de ces soutiens.

------
- Avez-vous bénéficié de l'aide temporaire aux travailleurs du gouvernement du Québec (PATT) ?

- Avez-vous bénéficié de la prestation d'urgence canadienne du gouvernement fédéral (PCU) ?

```{r}

freq(trajipaq$COV39)

#label3 <- c("Non", "Oui", "Non reponse")
label3 <- c("Non", "Oui")

freq(trajipaq$COV38)

trajipaq <-
  trajipaq %>% 
  mutate(patt = factor(case_when(
    COV38 == 1 ~ 1,
    COV38 == 2 ~ 0,
    COV38 == 0 ~ NA_real_,
    COV38 == 98 ~ NA_real_,
    COV38 == 99 ~ NA_real_,
    is.na(COV38) ~ NA_real_), labels = label3),
    pcu = factor(case_when(
    COV39 == 1 ~ 1,
    COV39 == 2 ~ 0,
    COV39 == 0 ~ NA_real_,
    COV39 == 98 ~ NA_real_,
    COV39 == 99 ~ NA_real_), labels = label3))

freq(trajipaq$patt)
freq(trajipaq$COV38)

with(trajipaq,  
  ctable(patt, COV38, "no"))

with(trajipaq,  
  ctable(pcu, COV39, "no"))

```

## Création des variables dépendante unique combinant plusieurs variables

- Exposition

```{r}

# création de la variable dépendante unique d'exposition

expo <-
  trajipaq %>% 
  select(e_hors_dom, e_four_soin, e_quarantaine, e_expose, e_symptome, e_hospitalise, e_intensif, e_decede, imm_minorite)

expo_freq <- expo %>% 
  tbl_summary(by = imm_minorite)

expo_freq

freq(trajipaq$e_hors_dom)

trajipaq <-
  trajipaq %>% 
  mutate(expo = factor(case_when(
    e_hors_dom == "oui" | e_four_soin == "oui" | e_quarantaine == "oui" | e_expose == "oui" | e_symptome == "oui" | e_hospitalise == "oui" | e_intensif == "oui" | e_decede == "oui" ~ 1,
    e_hors_dom == "non response" & e_four_soin == "non response" & e_quarantaine == "non response" & e_expose == "non response" & e_symptome == "non response" & e_hospitalise == "non response" & e_intensif == "non response" & e_decede == "non response" ~ NA_real_,
    is.na(e_hors_dom) & is.na(e_four_soin) & is.na(e_quarantaine) & is.na(e_expose) & is.na(e_symptome) &  is.na(e_hospitalise) & is.na(e_intensif) & is.na(e_decede) ~ NA_real_,
    TRUE ~ 0), labels = c("No", "At least one factor")))

freq(trajipaq$expo)

```

### Échantillon

```{r}

trajipaq_covid <-
  trajipaq %>% 
  select(CaseId, Poidstot1, Poidstot2, statut_imm, duree_res, minorite_vis, imm_minorite, sexe, gpage, niv_etude, statut_mat, statut_emploi, e_hors_dom, e_four_soin, e_quarantaine, e_expose, e_symptome, e_hospitalise, e_intensif, e_decede, f_separe, f_emmenage, f_quitte_dom, f_nourriture, f_medicaments, f_sante, f_revenu, f_assurance, f_evenement, f_essentiel, f_reduire, f_cesser, f_perdu, f_role_parent, f_entende, f_enfant_maladie, f_autre_enfant, f_pers_agees, be_exercice, be_manger, be_dormir, be_anxiete, be_humeur, patt, pcu, expo, length_stay, length_stay_2)

write_rds(trajipaq_covid, "../Data/trajipaq_covid.RData")

#trajipaq_cov <- read_rds("trajipaq_covid.RData")

```

### 2023-05-31 2. Trajipaq_covid/2. Codes/Trajipaq_covid_reg1.Rmd
Réorganisation de l'espace de travail.

---
title: "Familial consequences of the Covid-19 in Québec: Comparison between immigrants and natives"
subtitle: "Conséquences familiales de la Covid-19 au Québec: comparaison immigrants - natifs"
author:
- Vissého Adjiwanou ^[corresponding author - adjiwanou.vissého@uqam.ca]
- Département de Sociologie, Université du Québec à Montréal (UQAM)
- Solène Lardoux (Département de démographie, Université de Montréal)
- Nima Zahedinameghi (Département de Sociologie, UQAM)
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document: 
    toc: true
    toc_depth: 3
  word_document: default
  template: svm-latex-ms2.tex 
  pdf_document: default
  citation_package: natbib
bibliography: Covid_Canada.bib
---

<!--
Ce fichier est pour les modèles de régression

-->

```{r, echo=FALSE, warning=FALSE}

rm(list = ls())

library(tidyverse)
library(summarytools)
library(srvyr)
library(survey)
library(ggpubr)
library(writexl)
library(stargazer)
library(modelsummary)

trajipaq_covid <- read_rds("../Data/trajipaq_covid.RData")

```

## Recodage des variables

## Pondération

```{r}

#trajipaq_nm <- na.omit(trajipaq_covid)
#trajipaq.design <- svydesign(id = ~ CaseId, weights = ~ Poidstot1, data = trajipaq_nm)
trajipaq.design <- svydesign(id = ~ CaseId, weights = ~ Poidstot1, data = trajipaq_covid)

```

## Modèle de régression logistique

### Exposition

```{r}

#var_ind <- c("statut_imm + duree_res + minorite_vis + sexe + gpage + niv_etude + statut_mat + statut_emploi")
#var_ind

#e_hors_dom + e_four_soin + e_quarantaine + e_expose + e_symptome + e_hospitalise + e_intensif + e_decede

#freq(trajipaq_covid$e_four_soin)

#expo_reg1 <- glm(formula = e_hors_dom ~ statut_imm, data = trajipaq_covid, family = "binomial")
#summary(expo_reg1)

#expo_reg2 <- glm(formula = e_hors_dom ~ statut_imm + duree_res + minorite_vis + 
#                   sexe + gpage + niv_etude + statut_mat + statut_emploi, 
#                 data = trajipaq_covid, family = "binomial")
#summary(expo_reg2)

#exp(cbind(OR = coef(expo_reg2), confint(expo_reg2)))

#stargazer(expo_reg1, expo_reg2, apply.coef = exp, apply.se = exp, align = TRUE, type = "text")

expo <-
  trajipaq_covid %>% 
  dplyr::select(e_hors_dom, e_four_soin, e_quarantaine, e_expose, e_symptome, e_hospitalise, e_intensif, e_decede)

dfSummary(expo, 
          plain.ascii  = FALSE, 
          style        = "grid", 
          graph.magnif = 0.75, 
          valid.col    = FALSE,
          tmp.img.dir  = "/tmp")

dep1 <- list("e_hors_dom~", "e_four_soin~", "e_quarantaine~", "e_expose~",
            "e_symptome~", "e_hospitalise~", "e_intensif~", "e_decede~") # list of unique dependent variables with ~ 
indep1 <- list("relevel(statut_imm, 2)")  # list of first unique independent variables 
#indep2<-list("statut_imm", "sexe", "gpage", "niv_etude", "statut_mat", "statut_emploi") # list of second unique independent variables
indep2<-list("relevel(statut_imm, 2) + relevel(sexe, 2) + gpage + niv_etude + statut_mat + statut_emploi") #Map(function(x,y,z) lm(as.formula(paste(x,paste(list(y,z),collapse="+"))),data=trajipaq_covid),dep,indep1,indep2)
#indep3<-list("gpage")

model1 <- Map(function(x,y) glm(as.formula(paste(x,paste(list(y),collapse="+"))),data=trajipaq_covid, family="binomial"), dep1, indep1)

model2 <- Map(function(x, y) glm(as.formula(paste(x,paste(list(y),collapse="+"))),data=trajipaq_covid, family="binomial"), dep1, indep2)

stargazer(model1, ci = TRUE, apply.coef = exp, apply.ci = exp, align = TRUE, type = "text")
stargazer(model2, ci = TRUE, apply.coef = exp, apply.ci = exp, align = TRUE, type = "text")

#apply.se
#table_model1 <- modelsummary(model1, statistic = "conf.int")
#table_model1

```

### Consequences 1

```{r}

consequences <-
  trajipaq_covid %>% 
  dplyr::select(f_separe, f_emmenage, f_quitte_dom, f_nourriture, f_medicaments, f_sante, f_revenu, f_assurance, f_evenement, f_essentiel)

dfSummary(consequences, 
          plain.ascii  = FALSE, 
          style        = "grid", 
          graph.magnif = 0.75, 
          valid.col    = FALSE,
          tmp.img.dir  = "/tmp")

dep2 <- list("f_separe~", "f_emmenage~", "f_quitte_dom~", "f_nourriture~", "f_medicaments~", "f_sante~", "f_revenu~", "f_assurance~", "f_evenement~", "f_essentiel~")

model3 <- Map(function(x,y) glm(as.formula(paste(x,paste(list(y),collapse="+"))),data=trajipaq_covid, family="binomial"), dep2, indep1) 
stargazer(model3, ci = TRUE, apply.coef = exp, apply.ci = exp, align = TRUE, type = "text")

model4 <- Map(function(x, y) glm(as.formula(paste(x,paste(list(y),collapse="+"))),data=trajipaq_covid, family="binomial"), dep2, indep2) 
stargazer(model4, ci = TRUE, apply.coef = exp, apply.ci = exp, align = TRUE, type = "text")

```

### Emploi

```{r}

emploi <-
  trajipaq_covid %>% 
  dplyr::select(f_reduire, f_cesser, f_perdu)

dfSummary(emploi, 
          plain.ascii  = FALSE, 
          style        = "grid", 
          graph.magnif = 0.75, 
          valid.col    = FALSE,
          tmp.img.dir  = "/tmp")

dep3 <- list("f_reduire~", "f_cesser~", "f_perdu~")

model5 <- Map(function(x,y) glm(as.formula(paste(x,paste(list(y),collapse="+"))),data=trajipaq_covid, family="binomial"), dep3, indep1) 
stargazer(model5, ci = TRUE, apply.coef = exp, apply.ci = exp, align = TRUE, type = "text")

model6 <- Map(function(x, y) glm(as.formula(paste(x,paste(list(y),collapse="+"))),data=trajipaq_covid, family="binomial"), dep3, indep2) 
stargazer(model6, ci = TRUE, apply.coef = exp, apply.ci = exp, align = TRUE, type = "text")

```

### Entente dans la famille

```{r}

library(MASS)

entente <-
  trajipaq_covid %>% 
  dplyr::select(f_role_parent, f_entende, f_enfant_maladie, f_autre_enfant, f_pers_agees)

dfSummary(entente, 
          plain.ascii  = FALSE, 
          style        = "grid", 
          graph.magnif = 0.75, 
          valid.col    = FALSE,
          tmp.img.dir  = "/tmp")

dep4 <- list("f_role_parent~", "f_entende~", "f_enfant_maladie~", "f_autre_enfant~", "f_pers_agees~")

model7 <- Map(function(x,y) polr(as.formula(paste(x,paste(list(y),collapse="+"))), method = c("logistic"), Hess = TRUE, data=trajipaq_covid), dep4, indep1) 
stargazer(model7, ci = TRUE, apply.coef = exp, apply.ci = exp, align = TRUE, type = "text")

model8 <- Map(function(x, y) polr(as.formula(paste(x,paste(list(y),collapse="+"))), method = c("logistic"), Hess = TRUE, data=trajipaq_covid), dep4, indep2) 
stargazer(model8, ci = TRUE, apply.coef = exp, apply.ci = exp, type = "text")

```

### Bien-être

```{r}

bienetre <-
  trajipaq_covid %>% 
  dplyr::select(be_exercice, be_manger, be_dormir, be_anxiete, be_humeur)

dfSummary(bienetre, 
          plain.ascii  = FALSE, 
          style        = "grid", 
          graph.magnif = 0.75, 
          valid.col    = FALSE,
          tmp.img.dir  = "/tmp")

freq(trajipaq_covid$be_exercice)

dep5 <- list("be_exercice~", "be_manger~", "be_dormir~", "be_anxiete~", "be_humeur~")

model9 <- Map(function(x,y) polr(as.formula(paste(x,paste(list(y),collapse="+"))), method = c("logistic"), Hess = TRUE, data=trajipaq_covid), dep5, indep1) 
stargazer(model9, ci = TRUE, apply.coef = exp, apply.ci = exp, align = TRUE, type = "text")

model10 <- Map(function(x, y) polr(as.formula(paste(x,paste(list(y),collapse="+"))), method = c("logistic"), Hess = TRUE, data=trajipaq_covid), dep5, indep2) 
stargazer(model10, ci = TRUE, apply.coef = exp, apply.ci = exp, align = TRUE, type = "text")

```

### Aide reçue

```{r}

aide <-
  trajipaq_covid %>% 
  dplyr::select(patt, pcu)

dfSummary(aide, 
          plain.ascii  = FALSE, 
          style        = "grid", 
          graph.magnif = 0.75, 
          valid.col    = FALSE,
          tmp.img.dir  = "/tmp")

dep6 <- list("patt~", "pcu~")

model11 <- Map(function(x,y) glm(as.formula(paste(x,paste(list(y),collapse="+"))),data=trajipaq_covid, family="binomial"), dep6, indep1) 
stargazer(model11, ci = TRUE, apply.coef = exp, apply.ci = exp, align = TRUE, type = "text")

model12 <- Map(function(x, y) glm(as.formula(paste(x,paste(list(y),collapse="+"))),data=trajipaq_covid, family="binomial"), dep6, indep2) 
stargazer(model12, ci = TRUE, apply.coef = exp, apply.ci = exp, align = TRUE, type = "text")

```

### 2023-05-31 2. Trajipaq_covid/2. Codes/Trajipaq_covid_reg2.Rmd
Réorganisation de l'espace de travail.

---
title: "Familial consequences of the Covid-19 in Québec: Comparison between immigrants and natives"
subtitle: "Conséquences familiales de la Covid-19 au Québec: comparaison immigrants - natifs"
author:
- Vissého Adjiwanou ^[corresponding author - adjiwanou.vissého@uqam.ca]
- Département de Sociologie, Université du Québec à Montréal (UQAM)
- Solène Lardoux (Département de démographie, Université de Montréal)
- Nima Zahedinameghi (Département de Sociologie, UQAM)
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document: 
    toc: true
    toc_depth: 3
  word_document: default
  template: svm-latex-ms2.tex 
  pdf_document: default
  citation_package: natbib
bibliography: Covid_Canada.bib
---

<!--
Ce fichier est pour les modèles de régression
Reg2
Modèle avec la variable imm_minorite
-->

```{r, echo=FALSE, warning=FALSE}

rm(list = ls())

library(tidyverse)
library(summarytools)
library(srvyr)
library(survey)
library(ggpubr)
library(writexl)
library(stargazer)
library(modelsummary)
library(labelled)

trajipaq_covid <- read_rds("../Data/trajipaq_covid.RData")

labeled_vars <- sapply(trajipaq_covid, function(x) class(x) == "labelled")
trajipaq_covid[, labeled_vars] <- lapply(trajipaq_covid[, labeled_vars], as_factor)

# Base de données excluant others
trajipaq_covid <-
  trajipaq_covid %>% 
  filter(imm_minorite != "Others")

freq(trajipaq_covid$imm_minorite)
freq(trajipaq_covid$statut_mat)

freq(trajipaq_covid$length_stay_2)

```

## Recodage des variables

## Pondération

```{r}

#trajipaq_nm <- na.omit(trajipaq_covid)
#trajipaq.design <- svydesign(id = ~ CaseId, weights = ~ Poidstot1, data = trajipaq_nm)
trajipaq.design <- svydesign(id = ~ CaseId, weights = ~ Poidstot1, data = trajipaq_covid)

```

## Modèle de régression logistique

### Les différents modèles

```{r}

#expo <-
#  trajipaq_covid %>% 
#  dplyr::select(e_hors_dom, e_four_soin, e_quarantaine, e_expose, e_symptome, e_hospitalise, e_intensif, e_decede)

#dfSummary(expo, 
#          plain.ascii  = FALSE, 
#          style        = "grid", 
#          graph.magnif = 0.75, 
#          valid.col    = FALSE,
#          tmp.img.dir  = "/tmp")

freq(trajipaq_covid$statut_imm)
freq(trajipaq_covid$imm_minorite)

dep1 <- list("e_hors_dom~", "e_four_soin~", "e_quarantaine~", "e_expose~",
            "e_symptome~", "e_hospitalise~", "e_intensif~", "e_decede~")

dep2 <- list("expo~")

indep1 <- list("relevel(imm_minorite, 3)")

indep2 <- list("relevel(imm_minorite, 3) + relevel(sexe, 2) + length_stay_2 + gpage + niv_etude + statut_mat + statut_emploi")

indep2_bis <- list("relevel(imm_minorite, 3) + relevel(sexe, 2) + length_stay_2 + gpage + niv_etude + statut_mat")   # For emploi

# Change of the reference category

indep3 <- list("relevel(imm_minorite, 2)")

indep4 <- list("relevel(imm_minorite, 2) + relevel(sexe, 2) + length_stay_2 + gpage + niv_etude + statut_mat + statut_emploi") 
indep4_bis <- list("relevel(imm_minorite, 2) + relevel(sexe, 2) + length_stay_2 + gpage + niv_etude + statut_mat")

```

## Different models

- Model with dependent vars 1

model1.1 : only one independent variable (imm_minorite), reference : majority
model1.2 : control all independent variable (imm_minorite), reference : majority
model1.3 : only one independent variable (imm_minorite), reference : immigrant not visible
model1.4 : control all independent variable (imm_minorite), reference : immigrant not visible

- Modele unique

model_u1.1 : only one independent variable (imm_minorite), reference : majority
model_u1.2 : control all independent variable (imm_minorite), reference : majority
model_u1.3 : only one independent variable (imm_minorite), reference : immigrant not visible
model_u1.4 : control all independent variable (imm_minorite), reference : immigrant not visible

model1.2 : only one independent variable (imm_minorite), reference : immigrant not visible

## Distribution of the dependent variables

```{r}

freq(trajipaq_covid$f_role_parent)

```

## Model for all dependent variables except income and entente dans la famille

```{r}

model <- function(dep, indep){
  Map(function(x,y) glm(as.formula(paste(x,paste(list(y),collapse="+"))),data=trajipaq_covid, family="binomial"), dep, indep)
}

expo <- list("e_hors_dom~", "e_four_soin~", "e_quarantaine~", "e_expose~",
            "e_symptome~", "e_hospitalise~", "e_intensif~", "e_decede~")

ressources <- list("f_nourriture~", "f_medicaments~", "f_sante~", "f_assurance~",
                   "f_essentiel~")

ability_b <- list("f_separe~", "f_evenement~")

ability_o <- list("f_role_parent~", "f_entende~", "f_enfant_maladie~", "f_autre_enfant~", "f_pers_agees~")

#f_role_parent, f_entende, f_enfant_maladie, f_autre_enfant, f_pers_agees)

#cons <- list("f_separe~", "f_emmenage~", "f_quitte_dom~", "f_nourriture~", "f_medicaments~", "f_sante~", "f_assurance~", "f_evenement~", "f_essentiel~")

empl <- list("f_reduire~", "f_cesser~", "f_perdu~", "f_revenu~")

aide <- list("patt~", "pcu~")

independent <- list(indep1, indep2, indep3, indep4)
dependent <- list(expo, ressources, ability_b, aide)

model_global <- NULL
model_global <- list()

for (i in 1:length(dependent)) {
  for (j in 1:length(independent)) {
    model_i_j <- model(dependent[[i]], independent[[j]])
    model_global[[paste0("Model", i, j)]] <- model_i_j 
  }
}

cov_lab1 <- c("Non-racialised immigrant", "Racialised immigrant", "Female", "More than 5 years", "Born here", "35-44",
               "45-60", "Collegial or professional training", "University", "Maried", "Living together",
               "Separed, Divorced, Widow", "Unemployed", "Others")

cov_lab2 <- c("Majority group", "Racialised immigrant", "Female", "More than 5 years", "Born here", "35-44",
               "45-60", "Collegial or professional training", "University", "Maried", "Living together",
               "Separed, Divorced, Widow", "Unemployed", "Others")

#More than 5 years
# Result for exposition

dep_lab1 <- c("Work outside", "Health provider", "Quarantine", "Exposed", "Symptoms", "Hospitalised", "Intensive care", "Has dead")

stargazer(model_global[4], ci = TRUE, align = TRUE, type = "text", 
          dep.var.labels = c("Exposition"), column.labels = dep_lab1)

stargazer(model_global[1], ci = TRUE, align = TRUE, type = "text", covariate.labels = cov_lab1,
          dep.var.labels = c("Exposition"), column.labels = dep_lab1)

# stargazer(model_global[1], ci = TRUE, apply.coef = exp, apply.ci = exp, align = TRUE, type = "text", covariate.labels = cov_lab, dep.var.labels = c("Exposition"), column.labels = dep_lab1)

stargazer(model_global[2], ci = TRUE, align = TRUE, type = "text", covariate.labels = cov_lab1,
          dep.var.labels = c("Exposition"), column.labels = dep_lab1)

stargazer(model_global[3], ci = TRUE, align = TRUE, type = "text", covariate.labels = cov_lab2,
          dep.var.labels = c("Exposition"), column.labels = dep_lab1)

stargazer(model_global[4], ci = TRUE, align = TRUE, type = "text", covariate.labels = cov_lab2,
          dep.var.labels = c("Exposition"), column.labels = dep_lab1)

# On getting ressource

dep_lab2 <- c("Food", "Medicine", "Health care", "Family insurrance", "Other essentials")

stargazer(model_global[5], ci = TRUE, align = TRUE, type = "text", covariate.labels = cov_lab1,
          dep.var.labels = c("Difficulty getting..."), column.labels = dep_lab2)

stargazer(model_global[6], ci = TRUE, align = TRUE, type = "text", covariate.labels = cov_lab1,
          dep.var.labels = c("Difficulty getting..."), column.labels = dep_lab2)

stargazer(model_global[7], ci = TRUE, align = TRUE, type = "text", covariate.labels = cov_lab2,
          dep.var.labels = c("Difficulty getting..."), column.labels = dep_lab2)

stargazer(model_global[8], ci = TRUE, align = TRUE, type = "text", covariate.labels = cov_lab2,
          dep.var.labels = c("Difficulty getting..."), column.labels = dep_lab2)

# On ability to perform usual duties

dep_lab3 <- c("Separated", "Important event")

stargazer(model_global[9], ci = TRUE, align = TRUE, type = "text", covariate.labels = cov_lab1,
          dep.var.labels = c("Performing usual duties"), column.labels = dep_lab3)

stargazer(model_global[10], ci = TRUE, align = TRUE, type = "text", covariate.labels = cov_lab1,
          dep.var.labels = c("Performing usual duties"), column.labels = dep_lab3)

stargazer(model_global[11], ci = TRUE, align = TRUE, type = "text", covariate.labels = cov_lab2,
          dep.var.labels = c("Performing usual duties"), column.labels = dep_lab3)

stargazer(model_global[12], ci = TRUE, align = TRUE, type = "text", covariate.labels = cov_lab2,
          dep.var.labels = c("Performing usual duties"), column.labels = dep_lab3)

# Model for government assistance

dep_lab4 <- c("PATT", "PCU")

stargazer(model_global[13], ci = TRUE, align = TRUE, type = "text", covariate.labels = cov_lab1,
          dep.var.labels = c("Government assistance"), column.labels = dep_lab4)

stargazer(model_global[14], ci = TRUE, align = TRUE, type = "text", covariate.labels = cov_lab1,
          dep.var.labels = c("Government assistance"), column.labels = dep_lab4)

stargazer(model_global[15], ci = TRUE, align = TRUE, type = "text", covariate.labels = cov_lab2,
          dep.var.labels = c("Government assistance"), column.labels = dep_lab4)

stargazer(model_global[16], ci = TRUE, align = TRUE, type = "text", covariate.labels = cov_lab2,
          dep.var.labels = c("Government assistance"), column.labels = dep_lab4)

```

## Model for assistance with appropriate data

```{r}

freq(trajipaq_covid$f_revenu)

trajipaq_covid_aide <- trajipaq_covid %>% 
  filter(f_reduire == "oui" | f_cesser == "oui" | f_perdu == "oui" | f_revenu == "oui")

# 723 individus remplissent cette condition

ctable(trajipaq_covid_aide$imm_minorite, trajipaq_covid_aide$patt)

independent <- list(indep1, indep2, indep3, indep4)
aide <- list("patt~", "pcu~")

# New dependent variable

ctable(trajipaq_covid_aide$pcu, trajipaq_covid_aide$patt)

trajipaq_covid_aide <- trajipaq_covid_aide %>% 
  mutate(assistance = case_when(
    patt == "Oui" | pcu == "Oui" ~ 1,
    patt == "Non" & pcu == "Non" ~ 0,
    TRUE ~ NA_real_

  ))

freq(trajipaq_covid_aide$assistance)

# Model

aide1 <- list("assistance~")

model <- function(dep, indep){
  Map(function(x,y) glm(as.formula(paste(x,paste(list(y),collapse="+"))),data=trajipaq_covid_aide, family="binomial"), dep, indep)
}

model_aide <- NULL
model_aide <- list()

for (i in 1:length(independent)) {
  model_i <- model(aide1, independent[[i]])
  model_aide[[paste0("Model", i)]] <- model_i 
}

cov_lab1 <- c("Non-racialised immigrant", "Racialised immigrant", "Female", "More than 5 years", "Born here", "35-44",
               "45-60", "Collegial or professional training", "University", "Maried", "Living together",
               "Separed, Divorced, Widow", "Unemployed", "Others")

cov_lab2 <- c("Majority group", "Racialised immigrant", "Female", "More than 5 years", "Born here", "35-44",
               "45-60", "Collegial or professional training", "University", "Maried", "Living together",
               "Separed, Divorced, Widow", "Unemployed", "Others")

#dep_lab4 <- c("PATT", "PCU")
dep_lab4 <- c("Assistance received")

stargazer(model_aide[1], ci = TRUE, align = TRUE, type = "text", covariate.labels = cov_lab1,
          dep.var.labels = c("Government assistance"), column.labels = dep_lab4)

stargazer(model_aide[2], ci = TRUE, align = TRUE, type = "text", covariate.labels = cov_lab1,
          dep.var.labels = c("Government assistance"), column.labels = dep_lab4)

stargazer(model_aide[3], ci = TRUE, align = TRUE, type = "text", covariate.labels = cov_lab2,
          dep.var.labels = c("Government assistance"), column.labels = dep_lab4)

stargazer(model_aide[4], ci = TRUE, align = TRUE, type = "text", covariate.labels = cov_lab2,
          dep.var.labels = c("Government assistance"), column.labels = dep_lab4)

```

## model for emploi

### Emploi

```{r}

independent_bis <- list(indep1, indep2_bis, indep3, indep4_bis)

dep1 <- list("f_reduire~", "f_cesser~", "f_perdu~", "f_revenu~")

model_emp <- NULL
model_emp <- list()

for (i in 1:length(independent)) {
  model_i <- model(dep1, independent_bis[[i]])
  model_emp[[paste0("Model", i)]] <- model_i 
}

cov_lab1 <- c("Non-racialised immigrant", "Racialised immigrant", "Female", "More than 5 years", "Born here", "35-44",
               "45-60", "Collegial or professional training", "University", "Maried", "Living together",
               "Separed, Divorced, Widow")

cov_lab2 <- c("Majority group", "Racialised immigrant", "Female", "More than 5 years", "Born here", "35-44",
               "45-60", "Collegial or professional training", "University", "Maried", "Living together",
               "Separed, Divorced, Widow")

#, "Unemployed", "Others"
dep_lab <- c("Limit", "Stop", "Lost", "Income reduced")

stargazer(model_emp[1], ci = TRUE, align = TRUE, type = "text", covariate.labels = cov_lab1,
          dep.var.labels = c("Consequences on employment and income"), column.labels = dep_lab)

stargazer(model_emp[2], ci = TRUE, align = TRUE, type = "text", covariate.labels = cov_lab1,
          dep.var.labels = c("Consequences on employment and income"), column.labels = dep_lab)

stargazer(model_emp[3], ci = TRUE, align = TRUE, type = "text", covariate.labels = cov_lab2,
          dep.var.labels = c("Consequences on employment and income"), column.labels = dep_lab)

stargazer(model_emp[4], ci = TRUE, align = TRUE, type = "text", covariate.labels = cov_lab2,
          dep.var.labels = c("Consequences on employment and income"), column.labels = dep_lab)

```

### Entente dans la famille - on the ability to perform duties

```{r}

library(MASS)

dep1 <- list("f_role_parent~", "f_entende~", "f_enfant_maladie~", "f_autre_enfant~", "f_pers_agees~")

model <- function(dep, indep){
  Map(function(x,y) polr(as.formula(paste(x,paste(list(y),collapse="+"))), method = c("logistic"), Hess = TRUE, data=trajipaq_covid), dep, indep) }

independent <- list(indep1, indep2, indep3, indep4)

model_ent <- NULL
model_ent <- list()

for (i in 1:length(independent)) {
  model_i <- model(dep1, independent[[i]])
  model_ent[[paste0("Model", i)]] <- model_i 
}

#cov_lab1 <- c("Immigrant and a visible minority", "Immigrant not a visible minority", "Others", "Female", "35-44",
#               "45-60", "Collegial or professional training", "University", "Maried", "Living together",
#               "Separed, Divorced, Widow", "Unemployed", "Others")

dep_lab <- c("Parenting", "Get along", "Care for illness child", "Care for other child", "Care for elderly or disabled")

cov_lab1 <- c("Non-racialised immigrant", "Racialised immigrant", "Female", "More than 5 years", "Born here", "35-44",
               "45-60", "Collegial or professional training", "University", "Maried", "Living together",
               "Separed, Divorced, Widow", "Unemployed", "Others")

cov_lab2 <- c("Majority group", "Racialised immigrant", "Female", "More than 5 years", "Born here", "35-44",
               "45-60", "Collegial or professional training", "University", "Maried", "Living together",
               "Separed, Divorced, Widow", "Unemployed", "Others")

stargazer(model_ent[1], ci = TRUE, align = TRUE, type = "text", covariate.labels = cov_lab1,
          dep.var.labels = c("Ability to ..."), column.labels = dep_lab)

stargazer(model_ent[2], ci = TRUE, align = TRUE, type = "text", covariate.labels = cov_lab1,
          dep.var.labels = c("Ability to ..."), column.labels = dep_lab)

stargazer(model_ent[3], ci = TRUE, align = TRUE, type = "text", covariate.labels = cov_lab2,
          dep.var.labels = c("Ability to ..."), column.labels = dep_lab)

stargazer(model_ent[4], ci = TRUE, align = TRUE, type = "text", covariate.labels = cov_lab2,
          dep.var.labels = c("Ability to ..."), column.labels = dep_lab)

```

<!--

## Interaction model

```{r}

model <- function(dep, indep){
  Map(function(x,y) glm(as.formula(paste(x,paste(list(y),collapse="+"))),data=trajipaq_covid, family="binomial"), dep, indep)
}

independent <- list(indep1, indep2, indep3, indep4)

model_expo <- NULL
model_expo <- list()

for (i in 1:length(independent)) {
  model_i <- model(dep1, independent[[i]])
  model_expo[[paste0("Model", i)]] <- model_i 
}

cov_lab <- c("Immigrant and a visible minority", "Immigrant not a visible minority", "Others", "Female", "35-44",
               "45-60", "Collegial or professional training", "University", "Maried", "Living together",
               "Separed, Divorced, Widow", "Unemployed", "Others")

stargazer(model_expo[1], ci = TRUE, apply.coef = exp, apply.ci = exp, align = TRUE, type = "text",
          covariate.labels = cov_lab)
stargazer(model_expo[2], ci = TRUE, apply.coef = exp, apply.ci = exp, align = TRUE, type = "text",
          covariate.labels = cov_lab)
stargazer(model_expo[3], ci = TRUE, apply.coef = exp, apply.ci = exp, align = TRUE, type = "text")
stargazer(model_expo[4], ci = TRUE, apply.coef = exp, apply.ci = exp, align = TRUE, type = "text")

# Modele unique

cov_lab_u <- c("Immigrant and a visible minority", "Immigrant not a visible minority", "Others", "Female", "35-44",
               "45-60", "Collegial or professional training", "University", "Maried", "Living together",
               "Separed, Divorced, Widow", "Unemployed", "Others", "Immigrant and a visible minority", "Majority group",
               "Others")

model_expo_u <- NULL
model_expo_u <- list()

for (i in 1:4) {
  model_i <- model(dep2, independent[[i]])
  model_expo_u[[paste0("Model", i)]] <- model_i 
}

stargazer(model_expo_u[1:4], apply.coef = exp, ci = TRUE, apply.ci = exp, align = TRUE, type = "text",
          covariate.labels = cov_lab_u)

```

### Consequences 1

```{r}

consequences <-
  trajipaq_covid %>% 
  dplyr::select(f_separe, f_emmenage, f_quitte_dom, f_nourriture, f_medicaments, f_sante, f_revenu, f_assurance, f_evenement, f_essentiel)

dfSummary(consequences, 
          plain.ascii  = FALSE, 
          style        = "grid", 
          graph.magnif = 0.75, 
          valid.col    = FALSE,
          tmp.img.dir  = "/tmp")

dep1 <- list("f_separe~", "f_emmenage~", "f_quitte_dom~", "f_nourriture~", "f_medicaments~", "f_sante~", "f_revenu~", "f_assurance~", "f_evenement~", "f_essentiel~")

model_cons <- NULL
model_cons <- list()

for (i in 1:length(independent)) {
  model_i <- model(dep1, independent[[i]])
  model_cons[[paste0("Model", i)]] <- model_i 
}

cov_lab <- c("Immigrant and a visible minority", "Immigrant not a visible minority", "Others", "Female", "35-44",
               "45-60", "Collegial or professional training", "University", "Maried", "Living together",
               "Separed, Divorced, Widow", "Unemployed", "Others")

stargazer(model_cons[1], ci = TRUE, apply.coef = exp, apply.ci = exp, align = TRUE, type = "text",
          covariate.labels = cov_lab)
stargazer(model_cons[2], ci = TRUE, apply.coef = exp, apply.ci = exp, align = TRUE, type = "text",
          covariate.labels = cov_lab)
stargazer(model_cons[3], ci = TRUE, apply.coef = exp, apply.ci = exp, align = TRUE, type = "text")
stargazer(model_cons[4], ci = TRUE, apply.coef = exp, apply.ci = exp, align = TRUE, type = "text")

#model3 <- Map(function(x,y) glm(as.formula(paste(x,paste(list(y),collapse="+"))),data=trajipaq_covid, family="binomial"), dep2, indep1)

#stargazer(model3, ci = TRUE, apply.coef = exp, apply.ci = exp, align = TRUE, type = "text")

#model4 <- Map(function(x, y) glm(as.formula(paste(x,paste(list(y),collapse="+"))),data=trajipaq_covid, family="binomial"), dep2, indep2) 
#stargazer(model4, ci = TRUE, apply.coef = exp, apply.ci = exp, align = TRUE, type = "text")

```

### Emploi

```{r}

emploi <-
  trajipaq_covid %>% 
  dplyr::select(f_reduire, f_cesser, f_perdu)

dfSummary(emploi, 
          plain.ascii  = FALSE, 
          style        = "grid", 
          graph.magnif = 0.75, 
          valid.col    = FALSE,
          tmp.img.dir  = "/tmp")

dep1 <- list("f_reduire~", "f_cesser~", "f_perdu~")

model_emp <- NULL
model_emp <- list()

for (i in 1:length(independent)) {
  model_i <- model(dep1, independent[[i]])
  model_emp[[paste0("Model", i)]] <- model_i 
}

cov_lab <- c("Immigrant and a visible minority", "Immigrant not a visible minority", "Others", "Female", "35-44",
               "45-60", "Collegial or professional training", "University", "Maried", "Living together",
               "Separed, Divorced, Widow", "Unemployed", "Others")
dep_lab <- c("Limit", "Stop", "Lost")

stargazer(model_emp[1], ci = TRUE, apply.coef = exp, apply.ci = exp, align = TRUE, type = "text",
          covariate.labels = cov_lab, dep.var.labels = c("Consequences on employment"), column.labels = dep_lab)
stargazer(model_emp[2], ci = TRUE, apply.coef = exp, apply.ci = exp, align = TRUE, type = "text",
          covariate.labels = cov_lab)
stargazer(model_emp[3], ci = TRUE, apply.coef = exp, apply.ci = exp, align = TRUE, type = "text")
stargazer(model_emp[4], ci = TRUE, apply.coef = exp, apply.ci = exp, align = TRUE, type = "text")

model5 <- Map(function(x,y) glm(as.formula(paste(x,paste(list(y),collapse="+"))),data=trajipaq_covid, family="binomial"), dep3, indep1) 
stargazer(model5, ci = TRUE, apply.coef = exp, apply.ci = exp, align = TRUE, type = "text")

model6 <- Map(function(x, y) glm(as.formula(paste(x,paste(list(y),collapse="+"))),data=trajipaq_covid, family="binomial"), dep3, indep2) 
stargazer(model6, ci = TRUE, apply.coef = exp, apply.ci = exp, align = TRUE, type = "text")

```

### Bien-être

```{r}

bienetre <-
  trajipaq_covid %>% 
  dplyr::select(be_exercice, be_manger, be_dormir, be_anxiete, be_humeur)

dfSummary(bienetre, 
          plain.ascii  = FALSE, 
          style        = "grid", 
          graph.magnif = 0.75, 
          valid.col    = FALSE,
          tmp.img.dir  = "/tmp")

freq(trajipaq_covid$be_exercice)

dep5 <- list("be_exercice~", "be_manger~", "be_dormir~", "be_anxiete~", "be_humeur~")

model9 <- Map(function(x,y) polr(as.formula(paste(x,paste(list(y),collapse="+"))), method = c("logistic"), Hess = TRUE, data=trajipaq_covid), dep5, indep1) 
stargazer(model9, ci = TRUE, apply.coef = exp, apply.ci = exp, align = TRUE, type = "text")

model10 <- Map(function(x, y) polr(as.formula(paste(x,paste(list(y),collapse="+"))), method = c("logistic"), Hess = TRUE, data=trajipaq_covid), dep5, indep2) 
stargazer(model10, ci = TRUE, apply.coef = exp, apply.ci = exp, align = TRUE, type = "text")

```

### Aide reçue

```{r}

aide <-
  trajipaq_covid %>% 
  dplyr::select(patt, pcu)

dfSummary(aide, 
          plain.ascii  = FALSE, 
          style        = "grid", 
          graph.magnif = 0.75, 
          valid.col    = FALSE,
          tmp.img.dir  = "/tmp")

dep6 <- list("patt~", "pcu~")

model11 <- Map(function(x,y) glm(as.formula(paste(x,paste(list(y),collapse="+"))),data=trajipaq_covid, family="binomial"), dep6, indep1) 
stargazer(model11, ci = TRUE, apply.coef = exp, apply.ci = exp, align = TRUE, type = "text")

model12 <- Map(function(x, y) glm(as.formula(paste(x,paste(list(y),collapse="+"))),data=trajipaq_covid, family="binomial"), dep6, indep2) 
stargazer(model12, ci = TRUE, apply.coef = exp, apply.ci = exp, align = TRUE, type = "text")

```

-->

### 2023-06-03 1. Trajipaq_sentiment/2. Codes/Trajipaq_sentiment_cr.Rmd
Knit HTML des codes.
update

@@
 trajipaq <- read_dta("../3. Data/Base_Trajipaq_Finale_Ok.dta")
+wgidataset <- read_dta("../3. Data/wgidataset.dta")
 
@@
 
+# WGI
+
+```{r}
+
+trajipaq <- as_factor(trajipaq, only_labelled = TRUE)
+
+wgidataset <-
+  wgidataset %>%
+  mutate(countryname = recode(countryname, 
+                              "South Africa" = "Afrique du Sud, République d'",
+                              "Albania" = "Albanie",
+                              "Algeria" = "Algérie",
+                              "Germany" = "Allemagne",
+                              "Saudi Arabia" = "Arabie saoudite",
+                              "Argentina" = "Argentine",
+                              "Armenia" = "Arménie",
+                              "Australia" = "Australie",
+                              "Belarus" = "Bélarus",
+                              "Belgium" = "Belgique",
+                              "Benin" = "Bénin",
+                              "Brazil" = "Brésil",
+                              "Bulgaria" = "Bulgarie",
+                              "Cape Verde" = "Cabo Verde",
+                              "Cambodia" = "Cambodge",
+                              "Cameroon" = "Cameroun",
+                              "Central African Republic" = "Centrafricaine, République",
+                              "Chile" = "Chili",
+                              "China" = "Chine",
+                              "Jordan" = "Cisjordanie et bande de Gaza (Palestine)",
+                              "Colombia" = "Colombie",
+                              "Congo, Dem. Rep." = "Congo, République démocratique du", # ajouter modif sur trajipaq CR 1 wrong
+                              "Korea, Rep." = "Corée du sud",
+                              "Dominica" = "Dominicaine, République",
+                              "Egypt, Arab Rep." = "Égypte",
+                              "United Arab Emirates" = "Émirats arabes unis",
+                              "Spain" = "Espagne",
+                              "United States" = "États-Unis",
+                              "Fiji" = "Fidji",
+                              "Greece" = "Grèce",
+                              "Grenada" = "Grenade",
+                              "Haiti" = "Haïti",
+                              "Hong Kong SAR, China" = "Hong Kong",
+                              "India" = "Inde",
+                              "Iran, Islamic Rep." = "Iran",
+                              "Israel" = "Israël",
+                              "Italy" = "Italie",
+                              "Jamaica" = "Jamaïque",
+                              "Japan" = "Japon",
+                              "Latvia" = "Lettonie",
+                              "Lebanon" = "Liban",
+                              "Moroco" = "Maroc",
+                              "Mauritius" = "Maurice",
+                              "Mexico" = "Mexique",
+                              "Nigeria" = "Nigéria",
+                              "New Zealand" = "Nouvelle-Zélande",
+                              "Uganda" = "Ouganda",
+                              "Uzbekistan" = "Ouzbékistan",
+                              "Papua New Guinea" = "Papouasie-Nouvelle-Guinée",
+                              "Netherlands" = "Pays-Bas",
+                              "Peru" = "Pérou",
+                              "Poland" = "Pologne",
+                              "Romania" = "Roumanie",
+                              "United Kingdom" = "Royaume-Uni",
+                              "Russian Federation" = "Russie, Fédération de",
+                              "St. Vincent and the Grenadines" = "Saint-Vincent-et-les Grenadines",
+                              "Senegal" = "Sénégal",
+                              "Singapore" = "Singapour",
+                              "Slovak Republic" = "Slovaquie",
+                              "Sweden" = "Suède",
+                              "Switzerland" = "Suisse",
+                              "Syria" = "Syrie",
+                              "Taiwan, China" = "Taïwan",
+                              "Chad" = "Tchad",
+                              "Tunisia" = "Tunisie",
+                              "Venezuela, RB" = "Venezuela",
+                              "Vietnam" = "Viet Nam",
+                              "Yemen, Rep." = "Yémen"
+                              ))
+
+trajipaq <- merge(trajipaq, wgidataset, by.x = c("annee_qc", "Q11"),
+                by.y = c("year", "countryname"), all.x = TRUE)
+
+```
+
 # Création de la nouvelle base de données
@@
          "statut_mat",
+         "Q11",
          "VD1",
@@
          "VD5",
-         "VI4")
+         "VI4",
+         "vae",
+         "vas",
+         "van",
+         "var",
+         "val",
+         "vau",
+         "pve",
+         "pvs",
+         "pvn",
+         "pvr",
+         "pvl",
+         "pvu",
+         "gee",
+         "ges",
+         "gen",
+         "ger",
+         "gel",
+         "geu",
+         "rqe",
+         "rqs",
+         "rqn",
+         "rqr",
+         "rql",
+         "rqu",
+         "rle",
+         "rls",
+         "rln",
+         "rlr",
+         "rll",
+         "rlu",
+         "cce",
+         "ccs",
+         "ccn",
+         "ccr",
+         "ccl",
+         "ccu")
 

### 2023-06-05 1. Trajipaq_sentiment/2. Codes/Trajipaq_sentiment_cr.Rmd
update

--- 2023-05-31 1. Trajipaq_sentiment/2. Codes/Trajipaq_sentiment_cr.Rmd
@@
 
+# WGI
+
+```{r}
+
+trajipaq <- as_factor(trajipaq, only_labelled = TRUE)
+
+wgidataset <-
+  wgidataset %>%
+  mutate(countryname = recode(countryname, 
+                              "South Africa" = "Afrique du Sud, République d'",
+                              "Albania" = "Albanie",
+                              "Algeria" = "Algérie",
+
+                              "Germany" = "Allemagne",
+                              "Saudi Arabia" = "Arabie saoudite",
+                              "Argentina" = "Argentine",
+                              "Armenia" = "Arménie",
+                              "Australia" = "Australie",
+                              "Belarus" = "Bélarus",
+                              "Belgium" = "Belgique",
+                              "Benin" = "Bénin",
+                              "Brazil" = "Brésil",
+                              "Bulgaria" = "Bulgarie",
+                              "Cape Verde" = "Cabo Verde",
+                              "Cambodia" = "Cambodge",
+                              "Cameroon" = "Cameroun",
+                              "Central African Republic" = "Centrafricaine, République",
+                              "Chile" = "Chili",
+                              "China" = "Chine",
+                              "Jordan" = "Cisjordanie et bande de Gaza (Palestine)",
+                              "Colombia" = "Colombie",
+                              "Congo, Dem. Rep." = "Congo, République démocratique du", # ajouter modif sur trajipaq CR 1 wrong
+                              "Korea, Rep." = "Corée du sud",
+                              "Dominica" = "Dominicaine, République",
+                              "Egypt, Arab Rep." = "Égypte",
+                              "United Arab Emirates" = "Émirats arabes unis",
+                              "Spain" = "Espagne",
+                              "United States" = "États-Unis",
+                              "Fiji" = "Fidji",
+                              "Greece" = "Grèce",
+                              "Grenada" = "Grenade",
+                              "Haiti" = "Haïti",
+                              "Hong Kong SAR, China" = "Hong Kong",
+                              "India" = "Inde",
+                              "Iran, Islamic Rep." = "Iran",
+                              "Israel" = "Israël",
+                              "Italy" = "Italie",
+                              "Jamaica" = "Jamaïque",
+                              "Japan" = "Japon",
+                              "Latvia" = "Lettonie",
+                              "Lebanon" = "Liban",
+                              "Moroco" = "Maroc",
+                              "Mauritius" = "Maurice",
+                              "Mexico" = "Mexique",
+                              "Nigeria" = "Nigéria",
+                              "New Zealand" = "Nouvelle-Zélande",
+                              "Uganda" = "Ouganda",
+                              "Uzbekistan" = "Ouzbékistan",
+                              "Papua New Guinea" = "Papouasie-Nouvelle-Guinée",
+                              "Netherlands" = "Pays-Bas",
+                              "Peru" = "Pérou",
+                              "Poland" = "Pologne",
+                              "Romania" = "Roumanie",
+                              "United Kingdom" = "Royaume-Uni",
+                              "Russian Federation" = "Russie, Fédération de",
+                              "St. Vincent and the Grenadines" = "Saint-Vincent-et-les Grenadines",
+                              "Senegal" = "Sénégal",
+                              "Singapore" = "Singapour",
+                              "Slovak Republic" = "Slovaquie",
+                              "Sweden" = "Suède",
+                              "Switzerland" = "Suisse",
+                              "Syria" = "Syrie",
+                              "Taiwan, China" = "Taïwan",
+                              "Chad" = "Tchad",
+                              "Tunisia" = "Tunisie",
+                              "Venezuela, RB" = "Venezuela",
+                              "Vietnam" = "Viet Nam",
+                              "Yemen, Rep." = "Yémen"
+                              ))
+
+trajipaq <- merge(trajipaq, wgidataset, by.x = c("annee_qc", "Q11"),
+                by.y = c("year", "countryname"), all.x = TRUE)
+
+```
+
 # Création de la nouvelle base de données

### 2023-06-06 1. Trajipaq_sentiment/2. Codes/Trajipaq_sentiment_an.Rmd
Knit HTML des codes.
update
update WGI

@@
 trajipaq <- read_csv("../3. Data/Trajipaq_sentiment_cr.csv")
+view(trajipaq)
 
@@
 ```
+
 #### education
@@
 ```
+# WGI X Sentimapp
+
+```{r}
+reg_4 <- glm(formula = WGI ~ VD1, data = trajipaq)
+
+summary(reg_4)
+
+cor.test(trajipaq$WGI, trajipaq$VD1)
+
+ggplot(trajipaq %>% filter(!is.na(WGI), !is.na(VD1)), aes(x = WGI, y = VD1)) +
+  stat_smooth(method = "glm",
+              formula = "y ~ x",
+              color = "pink", 
+              se = FALSE) +
+  labs(
+    title = "Sentiment d'appartenance et indice de gouvernementalité du pays d'origine", 
+    x = "indice de gouvernementalité du pays d'origine au moment de l'immigration", 
+    y = "Sentiment d'appartenance au Canada") +
+  theme_bw()
+```
 

### 2023-06-06 1. Trajipaq_sentiment/2. Codes/Trajipaq_sentiment_cr.Rmd
update
update WGI

--- 2023-06-03 1. Trajipaq_sentiment/2. Codes/Trajipaq_sentiment_cr.Rmd
@@
 
+WGI_V <- c("vae", "pve", "gee", "rqe", "rle", "cce")
+
+trajipaq <-
+  trajipaq %>%
+  mutate(n_vae = vae,
+         n_pve = pve,
+         n_gee = gee,
+         n_rqe = rqe,
+         n_rle = rle,
+         n_cce = cce)
+
+WGI_n <- c("n_vae", "n_pve", "n_gee", "n_rqe", "n_rle", "n_cce")
+
+trajipaq <-
+  trajipaq %>% 
+  rowwise() %>% 
+  mutate (nbre_rep_WGI = sum(!is.na(across(starts_with("n_")))))
+
+trajipaq <-
+  trajipaq %>%
+  rowwise() %>%
+  mutate(sumWGI = sum(c_across(all_of(WGI_n)), na.rm = TRUE))%>%
+  mutate(WGI = (sumWGI / (nbre_rep_WGI))) %>%
+  mutate(WGI = if (nbre_rep_WGI == 0) NA_real_ else WGI) %>%
+  ungroup()
+
 ```
@@
          "ccl",
-         "ccu")
+         "ccu",
+         "WGI")
 

### 2023-06-07 1. Trajipaq_sentiment/1. Article/Visseho_al_sentiment_appartenance.Rmd
MàJ : revue de littérature.

@@
 
-## Situations de discriminations vécues par les immigrants
+## Discrimination perçue par les immigrants
+
+Dans une étude sur l'acculturation, la discrimination et l'adaptation des jeunes immigrants de deuxième génération, John W.
+Berry et al.
+observent que la discrimination perçue entretient une relation positive avec la diversité ethnoculturelle présente dans le pays d'accueil [-@berryImmigrantYouthCultural2006].
+Selon eux, la discrimination perçue constitue un prédicteur de la stratégie d'acculturation adoptée par les jeunes immigrants, ainsi qu'un indicateur de leur capacité à s'adapter dans leur pays d'accueil [@berryImmigrantYouthCultural2006].
+Leurs travaux révèlent notamment que les jeunes immigrants qui adoptent une stratégie d'intégration sont ceux qui déclarent subir le moins de discrimination, suivi en ordre croissant par ceux qui adoptent une stratégie d'assimilation, de séparation et de marginalisation [-@berryImmigrantYouthCultural2006].
+Ils montrent également que la discrimination perçue entretient une relation négative avec l'adaptation socioculturelle des jeunes immigrants [-@berryImmigrantYouthCultural2006].
+
+Dans une étude similaire comparant les contextes parisien et montréalais, Berry et Collette Sabatier remarquent que la discrimination collective perçue par les jeunes immigrants de deuxième génération varie en fonction de l'origine nationale, ethnique et culturelle du groupe étudié [-@berryAcculturationDiscriminationAdaptation2010, p.
+200].
+Ainsi, ils révèlent que ce sont les minorités les plus visibles (i.e, celles qui ont été historiquement racialisées) qui déclarent subir le plus de discrimination [-@berryAcculturationDiscriminationAdaptation2010].
+Dans le même temps, ils observent que la distance culturelle entre le pays d'origine des jeunes immigrants de deuxième génération et le pays d'accueil entretient une relation positive avec la discrimination perçue [-@berryAcculturationDiscriminationAdaptation2010].
+
+Selon Berry et Sabatier, la discrimination perçue par les jeunes immigrants de deuxième génération varie en fonction du pays d'accueil.
+À Montréal, ils observent que le fait d'entretenir une appartenance avec le pays d'origine tout en s'ouvrant à la culture du pays d'accueil (i.e, adopter une stratégie acculturative d'intégration) n'est pas associé positivement à la discrimination perçue [-@berryAcculturationDiscriminationAdaptation2010].
+Ils expliquent cette spécificité par l'influence du modèle multiculturaliste au Canada (et interculturaliste au Québec) depuis le début des années 1970.
+Selon eux, le sentiment d'appartenance au Québec des jeunes immigrants de deuxième génération entretient une relation négative avec la discrimination perçue.
+En outre, selon une étude antérieure réalisée par Berry et située elle aussi au Canada : lorsqu'un immigrant possède un lien durable et positif avec son pays d'origine, il tend à s'ouvrir plus facilement à la culture de la société d'accueil [-@berryMutualAttitudesImmigrants2006].
+
+Dans une étude sur l'identité, le bien-être et la discrimination des immigrants musulmans de seconde génération établis en Italie, Giuliani Cristina et al.
+observent que la discrimination perçue par ces derniers entretient une relation positive avec la montée en puissance de l'islamophobie en Occident depuis le début des années 2000 [-@giulianiPsychologicalWellbeingMultiple2018].
+Selon elle, l'identification religieuse des immigrants musulmans établis dans des sociétés historiquement chrétiennes constitue un point saillant de leur vie quotidienne [-@giulianiPsychologicalWellbeingMultiple2018].
+Conséquemment, en contexte islamophobe, ils observent que la discrimination perçue par les immigrants de seconde génération entretient une relation positive avec le sentiment d'appartenance religieux, ainsi qu'une relation négative avec le sentiment d'appartenance au pays d'accueil [-@giulianiPsychologicalWellbeingMultiple2018].
+
+Leur travail met en lumière la variation de la discrimination perçue par les immigrants de première et deuxième génération.
+Selon Giuliani et al.
+les immigrants de deuxième génération sont plus sensibles que leur parent face à la discrimination pour plusieurs raisons.
+Premièrement, les immigrants musulmans de deuxième génération grandissent dans leur pays d'accueil en tant que membres d'une minorité visible.
+Conséquemment, ils sont plus exposés que leurs parents à la discrimination [-@giulianiPsychologicalWellbeingMultiple2018].
+Deuxièmement, leur langue maternelle est celle du pays d'accueil.
+Ainsi, contrairement à leurs parents, ils interagissent plus fréquemment avec d'autres membres de la société d'accueil et sont plus à même de détecter la discrimination dans les paroles d'autrui [-@giulianiPsychologicalWellbeingMultiple2018].
+Enfin troisièmement, les immigrants musulmans de première génération conservent un lien émotionnel, social et culturel avec leur pays d'origine.
+Celui-ci leur permet de tamponner l'exposition à la discrimination dans le pays d'accueil.
+Au contraire, leurs enfants réalisent un arbitrage permanent entre la culture du pays d'origine et celle du pays d'accueil, ce qui les rend plus vulnérables face à la discrimination [-@giulianiPsychologicalWellbeingMultiple2018].
+
+Enfin, le travail de Giuliani et al.
+révèle également que les immigrants musulmans de première génération associent la discrimination qu'ils subissent à leur statut d'immigrant, tandis que les immigrants musulmans de deuxième génération associent cette même discrimination à leur identité ethnique et religieuse.
+Ainsi, les immigrants de deuxième génération tendent à être plus désenchantés que leurs parents vis-à-vis de leurs perspectives d'avenir dans le pays d'accueil [-@giulianiPsychologicalWellbeingMultiple2018].
 
@@
 
+```{=html}
 <!-- ## Variables
@@
 Discrimination --\> Sentiment d'appartenance -->
+```

### 2023-06-19 1. Trajipaq_sentiment/2. Codes/Trajipaq_sentiment_cr.Rmd
merge wgi data set full

@@
 ```{r}
-
 trajipaq <- as_factor(trajipaq, only_labelled = TRUE)
@@
 
+trajipaq$WGI_2021 <- merge(trajipaq, wgidataset[wgidataset$year == 2021, ],
+                           by.x = c("Q11"),
+                           by.y = c("countryname"),
+                           all.x = TRUE)
+```
+
+```{r}
 WGI_V <- c("vae", "pve", "gee", "rqe", "rle", "cce")
@@
          "ccu",
-         "WGI")
+         "WGI",
+         "WGI_2021")
 

### 2023-06-20 1. Trajipaq_sentiment/2. Codes/Trajipaq_sentiment_an.Rmd
final wgi merge

--- 2023-05-31 1. Trajipaq_sentiment/2. Codes/Trajipaq_sentiment_an.Rmd
@@
 ```
+# WGI X Sentimapp
+
+```{r}
+reg_4 <- glm(formula = vae_arrival ~ VD1, data = trajipaq)
+
+summary(reg_4)
+
+cor.test(trajipaq$vae_arrival, trajipaq$VD1)
+
+ggplot(trajipaq %>% filter(!is.na(vae_arrival), !is.na(VD1)), aes(x = vae_arrival, y = VD1)) +
+  stat_smooth(method = "glm",
+              formula = "y ~ x",
+              color = "pink", 
+              se = FALSE) +
+
+  labs(
+    title = "Sentiment d'appartenance et indice de gouvernementalité du pays d'origine", 
+    x = "indice de gouvernementalité du pays d'origine au moment de l'immigration", 
+    y = "Sentiment d'appartenance au Canada") +
+  theme_bw()
+```
 

### 2023-06-20 1. Trajipaq_sentiment/2. Codes/Trajipaq_sentiment_cr.Rmd
final wgi merge

--- 2023-06-06 1. Trajipaq_sentiment/2. Codes/Trajipaq_sentiment_cr.Rmd
@@
 
-WGI_V <- c("vae", "pve", "gee", "rqe", "rle", "cce")
+wgidataset_2021 <- filter(wgidataset, year == 2021)
 
-trajipaq <-
-  trajipaq %>%
-  mutate(n_vae = vae,
-         n_pve = pve,
-         n_gee = gee,
-         n_rqe = rqe,
-         n_rle = rle,
-         n_cce = cce)
+trajipaq <- trajipaq %>%
+  left_join(wgidataset_2021, by = c("Q11" = "countryname"))
 
-WGI_n <- c("n_vae", "n_pve", "n_gee", "n_rqe", "n_rle", "n_cce")
-
-trajipaq <-
-  trajipaq %>% 
-  rowwise() %>% 
-  mutate (nbre_rep_WGI = sum(!is.na(across(starts_with("n_")))))
-
-trajipaq <-
-  trajipaq %>%
-  rowwise() %>%
-  mutate(sumWGI = sum(c_across(all_of(WGI_n)), na.rm = TRUE))%>%
-  mutate(WGI = (sumWGI / (nbre_rep_WGI))) %>%
-  mutate(WGI = if (nbre_rep_WGI == 0) NA_real_ else WGI) %>%
-  ungroup()
+trajipaq <- trajipaq %>%
+  rename_with(~sub("\\.y$", "_2021", .), ends_with(".y"))%>%
+   rename_with(~sub("\\.x$", "_arrival", .), ends_with(".x"))
 
@@
          "VI4",
-         "vae",
-         "vas",
-         "van",
-         "var",
-         "val",
-         "vau",
-         "pve",
-         "pvs",
-         "pvn",
-         "pvr",
-         "pvl",
-         "pvu",
-         "gee",
-         "ges",
-         "gen",
-         "ger",
-         "gel",
-         "geu",
-         "rqe",
-         "rqs",
-         "rqn",
-         "rqr",
-         "rql",
-         "rqu",
-         "rle",
-         "rls",
-         "rln",
-         "rlr",
-         "rll",
-         "rlu",
-         "cce",
-         "ccs",
-         "ccn",
-         "ccr",
-         "ccl",
-         "ccu",
-         "WGI")
+         "vae_arrival",
+         "vas_arrival",
+         "van_arrival",
+         "var_arrival",
+         "val_arrival",
+         "vau_arrival",
+         "pve_arrival",
+         "pvs_arrival",
+         "pvn_arrival",
+         "pvr_arrival",
+         "pvl_arrival",
+         "pvu_arrival",
+         "gee_arrival",
+         "ges_arrival",
+         "gen_arrival",
+         "ger_arrival",
+         "gel_arrival",
+         "geu_arrival",
+         "rqe_arrival",
+         "rqs_arrival",
+         "rqn_arrival",
+         "rqr_arrival",
+         "rql_arrival",
+         "rqu_arrival",
+         "rle_arrival",
+         "rls_arrival",
+         "rln_arrival",
+         "rlr_arrival",
+         "rll_arrival",
+         "rlu_arrival",
+         "cce_arrival",
+         "ccs_arrival",
+         "ccn_arrival",
+         "ccr_arrival",
+         "ccl_arrival",
+         "ccu_arrival",
+         "vae_2021",
+         "vas_2021",
+         "van_2021",
+         "var_2021",
+         "val_2021",
+         "vau_2021",
+         "pve_2021",
+         "pvs_2021",
+         "pvn_2021",
+         "pvr_2021",
+         "pvl_2021",
+         "pvu_2021",
+         "gee_2021",
+         "ges_2021",
+         "gen_2021",
+         "ger_2021",
+         "gel_2021",
+         "geu_2021",
+         "rqe_2021",
+         "rqs_2021",
+         "rqn_2021",
+         "rqr_2021",
+         "rql_2021",
+         "rqu_2021",
+         "rle_2021",
+         "rls_2021",
+         "rln_2021",
+         "rlr_2021",
+         "rll_2021",
+         "rlu_2021",
+         "cce_2021",
+         "ccs_2021",
+         "ccn_2021",
+         "ccr_2021",
+         "ccl_2021",
+         "ccu_2021")
 

### 2023-06-20 1. Trajipaq_sentiment/2. Codes/Trajipaq_sentiment_an.Rmd
update WGI

@@
 ```{r}
-reg_4 <- glm(formula = vae_arrival ~ VD1, data = trajipaq)
+reg_4 <- glm(formula = vae_arrival ~ VD1 + VD2 + VD4 + VD5, data = trajipaq) # voice and acountability
+reg_5 <- glm(formula = pve_arrival ~ VD1 + VD2 + VD4 + VD5, data = trajipaq) # Political stability No Violence
+reg_6 <- glm(formula = gee_arrival ~ VD1 + VD2 + VD4 + VD5, data = trajipaq) # Government Effectiveness
+reg_7 <- glm(formula = rqe_arrival ~ VD1 + VD2 + VD4 + VD5, data = trajipaq) # Regulatory quality
+reg_8 <- glm(formula = rle_arrival ~ VD1 + VD2 + VD4 + VD5, data = trajipaq) # Rule fo law
+reg_9 <- glm(formula = cce_arrival ~ VD1 + VD2 + VD4 + VD5, data = trajipaq) # Control of corruption
 
 summary(reg_4)
+summary(reg_5)
+summary(reg_6)
+summary(reg_7)
+summary(reg_8)
+summary(reg_9)
 
-cor.test(trajipaq$vae_arrival, trajipaq$VD1)
+cor.test(trajipaq$vae_arrival, trajipaq$VD2)
 

### 2023-06-20 1. Trajipaq_sentiment/2. Codes/Trajipaq_sentiment_cr.Rmd
update WGI

@@
 wgidataset_2021 <- filter(wgidataset, year == 2021)
+wgidataset_canada <- filter(wgidataset, countryname == "Canada")
 
 trajipaq <- trajipaq %>%
-  left_join(wgidataset_2021, by = c("Q11" = "countryname"))
+  left_join(wgidataset_2021, by = c("Q11" = "countryname"))%>%
+  left_join(wgidataset_canada, by = c("annee_qc" = "year"))
+
+trajipaq <- trajipaq %>%
 

### 2023-06-21 1. Trajipaq_sentiment/2. Codes/Trajipaq_sentiment_an.Rmd
WGI update

--- 2023-06-20 1. Trajipaq_sentiment/2. Codes/Trajipaq_sentiment_an.Rmd
@@
 ```{r}
-reg_4 <- glm(formula = vae_arrival ~ VD1, data = trajipaq)
+reg_4 <- glm(formula = vae_arrival ~ VD1 + VD2 + VD4 + VD5, data = trajipaq) # voice and acountability
+  reg_4.1 <- glm(formula = vae_arrival_diff ~ VD1 + VD2 + VD3 + VD4, data = trajipaq)
+reg_5 <- glm(formula = pve_arrival ~ VD1 + VD2 + VD4 + VD5, data = trajipaq) # Political stability No Violence
+reg_6 <- glm(formula = gee_arrival ~ VD1 + VD2 + VD4 + VD5, data = trajipaq) # Government Effectiveness
+reg_7 <- glm(formula = rqe_arrival ~ VD1 + VD2 + VD4 + VD5, data = trajipaq) # Regulatory quality
+reg_8 <- glm(formula = rle_arrival ~ VD1 + VD2 + VD4 + VD5, data = trajipaq) # Rule fo law
+reg_9 <- glm(formula = cce_arrival ~ VD1 + VD2 + VD4 + VD5, data = trajipaq) # Control of corruption
 
 summary(reg_4)
+summary(reg_5)
+summary(reg_6)
+summary(reg_7)
+summary(reg_8)
+summary(reg_9)
+summary(reg_4.1)
 
-cor.test(trajipaq$vae_arrival, trajipaq$VD1)
+cor.test(trajipaq$vae_arrival, trajipaq$VD2)
 

### 2023-06-21 1. Trajipaq_sentiment/2. Codes/Trajipaq_sentiment_cr.Rmd
WGI update

@@
 
-# Q39F = ’ai un sentiment d’appartenance à mon pays de naissance.
+# Q39F = j’ai un sentiment d’appartenance à mon pays de naissance.
 
@@
 trajipaq <- trajipaq %>%
-
-trajipaq <- trajipaq %>%
   rename_with(~sub("\\.y$", "_2021", .), ends_with(".y"))%>%
@@
 
+trajipaq <- rename(trajipaq, "vae_can" = "vae")
+
+trajipaq <- trajipaq %>%
+  rowwise()%>%
+  mutate(vae_arrival_diff = vae_can - vae_arrival)
 ```
@@
          "vae_arrival",
-         "vas_arrival",
-         "van_arrival",
-         "var_arrival",
-         "val_arrival",
-         "vau_arrival",
          "pve_arrival",
-         "pvs_arrival",
-         "pvn_arrival",
-         "pvr_arrival",
-         "pvl_arrival",
-         "pvu_arrival",
          "gee_arrival",
-         "ges_arrival",
-         "gen_arrival",
-         "ger_arrival",
-         "gel_arrival",
-         "geu_arrival",
          "rqe_arrival",
-         "rqs_arrival",
-         "rqn_arrival",
-         "rqr_arrival",
-         "rql_arrival",
-         "rqu_arrival",
          "rle_arrival",
-         "rls_arrival",
-         "rln_arrival",
-         "rlr_arrival",
-         "rll_arrival",
-         "rlu_arrival",
          "cce_arrival",
-         "ccs_arrival",
-         "ccn_arrival",
-         "ccr_arrival",
-         "ccl_arrival",
-         "ccu_arrival",
          "vae_2021",
-         "vas_2021",
-         "van_2021",
-         "var_2021",
-         "val_2021",
-         "vau_2021",
          "pve_2021",
-         "pvs_2021",
-         "pvn_2021",
-         "pvr_2021",
-         "pvl_2021",
-         "pvu_2021",
          "gee_2021",
-         "ges_2021",
-         "gen_2021",
-         "ger_2021",
-         "gel_2021",
-         "geu_2021",
          "rqe_2021",
-         "rqs_2021",
-         "rqn_2021",
-         "rqr_2021",
-         "rql_2021",
-         "rqu_2021",
          "rle_2021",
-         "rls_2021",
-         "rln_2021",
-         "rlr_2021",
-         "rll_2021",
-         "rlu_2021",
          "cce_2021",
-         "ccs_2021",
-         "ccn_2021",
-         "ccr_2021",
-         "ccl_2021",
-         "ccu_2021")
+         "vae_can",
+         "vae_arrival_diff")
 

### 2023-06-21 1. Trajipaq_sentiment/2. Codes/Trajipaq_sentiment_an.Rmd
WGI update

@@
 ggplot(trajipaq %>% filter(!is.na(vae_arrival), !is.na(VD1)), aes(x = vae_arrival, y = VD1)) +
+  geom_smooth(method = "glm",
+              formula = "y ~ x",
+              color = "pink", 
+              se = FALSE) +
+  labs(
+    title = "Sentiment d'appartenance au Québec et indice de gouvernementalité du pays d'origine, au moment de l'immigration", 
+    x = "indice de gouvernementalité du pays d'origine au moment de l'immigration", 
+    y = "Sentiment d'appartenance au Québec") +
+  theme_bw()
+
+ggplot(trajipaq %>% filter(!is.na(vae_arrival), !is.na(VD2)), aes(x = vae_arrival, y = VD1)) +
   stat_smooth(method = "glm",
@@
               se = FALSE) +
-
   labs(
-    title = "Sentiment d'appartenance et indice de gouvernementalité du pays d'origine", 
+    title = "Sentiment d'appartenance au Canada et indice de gouvernementalité du pays d'origine, au moment de l'immigration", 
     x = "indice de gouvernementalité du pays d'origine au moment de l'immigration", 

### 2023-06-22 1. Trajipaq_sentiment/1. Article/Visseho_al_sentiment_appartenance.Rmd
update

@@
 
-# Les immigrants sont souvent pris entre deux ou plusieurs cultures, celles de leur pays d'origine et celles de leur nouveau pays de résidence.
-Cet article essaie de comprendre comment ces sentiments sont associés à leur pays d'origine et aux éventuelles situations de discriminations qu'ils vivent dans leur pays d'accueil.
+# Les immigrants sont souvent pris entre deux ou plusieurs cultures, celles de leur pays d'origine et celles de leur pays d’accueil.
+Alors que plusieurs études ont regardé la manière dont les conditions dans le pays d’accueil (notamment la discrimination) affectent leur sentiment d’appartenance, très peu ont analysé l’influence de la situation du pays d’origine.
+Cet article éssaie de rémedier à cela en combinant les données issues de l’enquêtes biographique sur les Trajectoires de participation individuelles des immigrants, des minorités ethnoculturelles, et le reste de la population québécoise (TrajIPaQ) menée en 2020 (Lardoux et al., 2022) avec les données sur la situation politique des pays d’origine des immigrants obtenues au moment de la collecte.
+L’enquête TrajIPaQ a touché 1000 immigrants au Québec et a collecté des informations sur leur situation familiale, sociale, éducative et professionnelle au cours des 15 dernières années depuis leur installation.
+.
 

### 2023-07-28 1. Trajipaq_sentiment/2. Codes/Trajipaq_sentiment_cr.Rmd
ajustement VD1
same

@@
   mutate(Q39_new_A = Q39A, # Création de nouvelles variables.
-         Q39_new_B = Q39B,
          Q39_new_C = Q39C,
@@
   "Q39_new_A",
-  "Q39_new_B",
   "Q39_new_C",
@@
 trajipaq <- rename(trajipaq, "vae_can" = "vae")
+trajipaq <- rename(trajipaq, "pve_can" = "pve")
+trajipaq <- rename(trajipaq, "gee_can" = "gee")
+trajipaq <- rename(trajipaq, "rqe_can" = "rqe")
+trajipaq <- rename(trajipaq, "rle_can" = "rle")
+trajipaq <- rename(trajipaq, "cce_can" = "cce")
 
-trajipaq <- trajipaq %>%
-  rowwise()%>%
-  mutate(vae_arrival_diff = vae_can - vae_arrival)
+trajipaq <- trajipaq %>% # création des WGI diff (can-home)
+  rowwise() %>%
+  mutate(vae_arrival_diff = vae_can - vae_arrival) %>%
+  mutate(pve_arrival_diff = pve_can - pve_arrival) %>%
+  mutate(gee_arrival_diff = gee_can - gee_arrival) %>%
+  mutate(rqe_arrival_diff = rqe_can - rqe_arrival) %>%
+  mutate(rle_arrival_diff = rle_can - rle_arrival) %>%
+  mutate(cce_arrival_diff = cce_can - cce_arrival)
+
+trajipaq <- trajipaq %>% #enlever les répondants d'origine canadienne (977 restants)
+  filter(origine != "Canada")
+
+freq(trajipaq$origine)
+
 ```
@@
          "vae_can",
-         "vae_arrival_diff")
+         "pve_can",
+         "gee_can",
+         "rqe_can",
+         "rle_can",
+         "cce_can",
+         "vae_arrival_diff",
+         "pve_arrival_diff",
+         "gee_arrival_diff",
+         "rqe_arrival_diff",
+         "rle_arrival_diff",
+         "cce_arrival_diff"
+
+         )
 

### 2023-07-28 1. Trajipaq_sentiment/2. Codes/Trajipaq_sentiment_an.Rmd
Après rencontre 0728

@@
 library(haven)
+library(stargazer)
+library(gtsumamry)
 
@@
 ```
-# WGI X Sentimapp
+# WGI X Canada
 
 ```{r}
-reg_4 <- glm(formula = vae_arrival ~ VD1 + VD2 + VD4 + VD5, data = trajipaq) # voice and acountability
-  reg_4.1 <- glm(formula = vae_arrival_diff ~ VD1 + VD2 + VD3 + VD4, data = trajipaq)
-reg_5 <- glm(formula = pve_arrival ~ VD1 + VD2 + VD4 + VD5, data = trajipaq) # Political stability No Violence
-reg_6 <- glm(formula = gee_arrival ~ VD1 + VD2 + VD4 + VD5, data = trajipaq) # Government Effectiveness
-reg_7 <- glm(formula = rqe_arrival ~ VD1 + VD2 + VD4 + VD5, data = trajipaq) # Regulatory quality
-reg_8 <- glm(formula = rle_arrival ~ VD1 + VD2 + VD4 + VD5, data = trajipaq) # Rule fo law
-reg_9 <- glm(formula = cce_arrival ~ VD1 + VD2 + VD4 + VD5, data = trajipaq) # Control of corruption
+#Au moment de l'immigration
 
-summary(reg_4)
-summary(reg_5)
-summary(reg_6)
-summary(reg_7)
-summary(reg_8)
-summary(reg_9)
-summary(reg_4.1)
+reg_WGI_can1 <- glm(formula = VD2_dico ~ vae_arrival, data = trajipaq, family = "binomial") # VD2 dichotomique
+reg_WGI_can1_ctrl <- glm(formula = VD2_dico ~ vae_arrival  + age + education + sexe + origine + revenu + langue_mat + religion + statut_mat , data = trajipaq, family = "binomial") #VD2 contrôlée par CSD
 
+summary(reg_WGI_can1)
+summary(reg_WGI_can1_ctrl)
+
+# Au moment de l'enquête (like)
+
+reg_WGI_can2 <- glm(formula = VD2_dico ~ vae_2021 , data = trajipaq, family = "binomial") # VD2 non regroupée
+reg_WGI_can2_ctrl <- glm(formula = VD2_dico ~ vae_2021  + age + education + sexe + origine + revenu + langue_mat + religion + statut_mat , data = trajipaq, family = "binomial")#VD2 contrôlée par CSD
+
+summary(reg_WGI_can2)
+summary(reg_WGI_can2_ctrl)
+
+# À l'arrivée et au moment de l'enquête
+
+reg_WGI_can3 <- glm(formula = VD2_dico ~ vae_2021 + vae_arrival , data = trajipaq, family = "binomial") # VD2 non regroupée
+reg_WGI_can3_ctrl <- glm(formula = VD2_dico ~ vae_2021 + vae_arrival + age + education + sexe + origine + revenu + langue_mat + religion + statut_mat , data = trajipaq, family = "binomial")# VD2 contrôlée par CSD
+
+summary(reg_WGI_can3)
+summary(reg_WGI_can3_ctrl)
+
+freq(trajipaq$VD2_long)
+
+```
+
+# WGI X Québec
+
+```{r}
+# Au moment de l'immigration
+
+reg_WGI_qc1 <- lm(formula = VD1  ~ vae_arrival, data = trajipaq) # VD1
+reg_WGI_qc1_ctrl <- lm(formula = VD1 ~ vae_arrival  + age + education + sexe + origine + revenu + langue_mat + religion + statut_mat , data = trajipaq) #VD1 contrôlée par CSD
+
+summary(reg_WGI_qc1)
+summary(reg_WGI_qc1_ctrl)
+
+# Au moment de l'enquête
+
+reg_WGI_qc2 <- lm(formula = VD1 ~ vae_2021 , data = trajipaq) # VD1
+reg_WGI_qc2_ctrl <- lm(formula = VD1 ~ vae_2021  + age + education + sexe + origine + revenu + langue_mat + religion + statut_mat , data = trajipaq) # VD1 contrôlée par CSD
+
+summary(reg_WGI_qc2)
+summary(reg_WGI_qc2_ctrl)
+
+# À l'arrivée et au moment de l'enquête
+
+reg_WGI_qc3 <- lm(formula = VD1 ~ vae_2021 + vae_arrival , data = trajipaq) # VD1
+reg_WGI_qc3_ctrl <- lm(formula = VD1 ~ vae_2021 + vae_arrival + age + education + sexe + origine + revenu + langue_mat + religion + statut_mat , data = trajipaq)# VD1 contrôlée par CSD
+
+summary(reg_WGI_qc3)
+summary(reg_WGI_qc3_ctrl)
+```
+Résumé des résultats Can + QC
+```{r}
+stargazer(reg_WGI_can2, reg_WGI_qc2, reg_WGI_can2_ctrl, reg_WGI_qc2_ctrl, title="Results",align=TRUE, type = "text")
+
+```
+
+```{r}
 cor.test(trajipaq$vae_arrival, trajipaq$VD2)

### 2023-07-28 1. Trajipaq_sentiment/2. Codes/Trajipaq_sentiment_cr.Rmd
Après rencontre 0728

@@
 freq(trajipaq$age)
+
+trajipaq <-
+  trajipaq %>%
+  mutate(age_small = factor(case_when( # création de la variable "age".
+    QAGE == 1 ~ "[0;29]",
+    QAGE == 2 ~ "[0;29]",
+    QAGE == 3 ~ "[0;29]",
+    QAGE == 4 ~ "[30;34]",
+    QAGE == 5 ~ "[35;39]",
+    QAGE == 6 ~ "[40;44]",
+    QAGE == 7 ~ "[45;49]",
+    QAGE == 8 ~ "[50;99]",
+    QAGE == 9 ~ "[50;99]",
+    QAGE == 99 ~ NA
+  ), levels = c( # Ordonnancement de la variable "age".
+  "[0;29]",
+  "[30;34]",
+  "[35;39]",
+  "[40;44]",
+  "[45;49]",
+  "[50;99]",
+  NA)))
 
@@
   trajipaq %>%
-  mutate(VD2 = case_when( # Création de la variable "VD2".
-    Q39E == 1 ~ "Oui",
-    Q39E == 2 ~ "Oui",
-    Q39E == 3 ~ "Non",
-    Q39E == 4 ~ "Non",
+  mutate(VD2_long = case_when( # Création de la variable "VD2".
+    Q39E == 1 ~ 1,
+    Q39E == 2 ~ 2,
+    Q39E == 3 ~ 3,
+    Q39E == 4 ~ 4,
     Q39E == 98 ~ NA,
@@
 
-freq(trajipaq$VD2)
+trajipaq <-
+  trajipaq %>%
+  mutate(VD2_dico = case_when( # Création de la variable "VD2".
+    Q39E == 1 ~ 1,
+    Q39E == 2 ~ 1,
+    Q39E == 3 ~ 0,
+    Q39E == 4 ~ 0,
+    Q39E == 98 ~ NA_real_,
+    Q39E == 99 ~ NA_real_
+  ))
+
+freq(trajipaq$VD2_long)
+freq(trajipaq$VD2_dico)
 
@@
   mutate(VD4 = case_when( # Création de la variable "VD4".
-    VD2 == "Oui" & VD3 == "Oui" ~ "Integre",
-    VD2 == "Oui" & VD3 == "Non" ~ "Assimile",
-    VD2 == "Non" & VD3 == "Oui" ~ "Separe",
-    VD2 == "Non" & VD3 == "Non" ~ "Marginalise"
+    VD2_dico == "Oui" & VD3 == "Oui" ~ "Integre",
+    VD2_dico == "Oui" & VD3 == "Non" ~ "Assimile",
+    VD2_dico == "Non" & VD3 == "Oui" ~ "Separe",
+    VD2_dico == "Non" & VD3 == "Non" ~ "Marginalise"
   ))
@@
   mutate(VD5 = factor(case_when( # Création de la variable "VD5".
-    VD2 == "Oui" & Q39D_new_2 == "Oui" ~ 1,
-    VD2 == "Oui" & Q39D_new_2 == "Non" ~ 2,
-    VD2 == "Non" & Q39D_new_2 == "Oui" ~ 3,
-    VD2 == "Non" & Q39D_new_2 == "Non" ~ 4
+    VD2_dico == "Oui" & Q39D_new_2 == "Oui" ~ 1,
+    VD2_dico == "Oui" & Q39D_new_2 == "Non" ~ 2,
+    VD2_dico == "Non" & Q39D_new_2 == "Oui" ~ 3,
+    VD2_dico == "Non" & Q39D_new_2 == "Non" ~ 4
   ),
@@
          "VD1",
-         "VD2",
+         "VD2_dico",
+         "VD2_long",
          "VD3",

### 2023-07-30 1. Trajipaq_sentiment/2. Codes/Trajipaq_sentiment_cr.Rmd
recode CSD

@@
   trajipaq %>%
-  mutate(age = factor(case_when( # création de la variable "age".
+  mutate(age_full = factor(case_when( # création de la variable "age".
     QAGE == 1 ~ "[0;19]",
@@
 
-freq(trajipaq$age)
+freq(trajipaq$age_full)
 
@@
   trajipaq %>%
-  mutate(age_small = factor(case_when( # création de la variable "age".
+  mutate(age = factor(case_when( # création de la variable "age".
     QAGE == 1 ~ "[0;29]",
@@
   mutate(education = factor(case_when( # Création de la variable "education".
-    Q16 == 1 ~ "Elementaire",
-    Q16 == 2 ~ "Secondaire",
-    Q16 == 3 ~ "Collegial",
+    Q16 == 1 ~ "Secondaire et moins",
+    Q16 == 2 ~ "Secondaire et moins",
+    Q16 == 3 ~ "Post secondaire",
     Q16 == 4 ~ "Post secondaire",
@@
   ), levels = c( # Ordonnancement de la variable "education".
-    "Elementaire", 
-    "Secondaire", 
+    "Secondaire et moins", 
     "Post secondaire",
-    "Collegial", 
     "Universitaire")))
@@
     region_naissance == "Canada" ~ "Canada",
-    region_naissance == "Caraibes" ~ "Caraibes",
+    region_naissance == "Caraibes" ~ "Autres",
     region_naissance == "Europe" ~ "Europe",
@@
   mutate(revenu = factor(case_when( # Création de la variable "revenu".
-    Q100 == 1 ~ "[0 ; 19 999]",
-    Q100 == 2 ~ "[0 ; 19 999]",
-    Q100 == 3 ~ "[20 000 ; 39 999]",
-    Q100 == 4 ~ "[20 000 ; 39 999]",
-    Q100 == 5 ~ "[40 000 ; 59 999]",
-    Q100 == 6 ~ "[40 000 ; 59 999]",
-    Q100 == 7 ~ "[60 000 ; 79 999]",
-    Q100 == 8 ~ "[60 000 ; 79 999]",
+    Q100 == 1 ~ "[0 ; 39 999]",
+    Q100 == 2 ~ "[0 ; 39 999]",
+    Q100 == 3 ~ "[0 ; 39 999]",
+    Q100 == 4 ~ "[0 ; 39 999]",
+    Q100 == 5 ~ "[40 000 ; 99 999]",
+    Q100 == 6 ~ "[40 000 ; 99 999]",
+    Q100 == 7 ~ "[60 000 ; 99 999]",
+    Q100 == 8 ~ "[60 000 ; 99 999]",
     Q100 == 9 ~ "[80 000 ; 99 999]",
-    Q100 == 10 ~ "[90 000 ; 99 999]",
-    Q100 == 11 ~ "[100 000 ; 149 999]",
-    Q100 == 12 ~ "[150 000 et plus]",
+    Q100 == 10 ~ "[100 000 et plus]",
+    Q100 == 11 ~ "[100 000 et plus]",
+    Q100 == 12 ~ "[100 000 et plus]",
     Q100 == 98 ~ NA,
@@
   ), levels = c(
-  "[0 ; 19 999]",
-  "[20 000 ; 39 999]",
-  "[40 000 ; 59 999]",
-  "[60 000 ; 79 999]",
-  "[80 000 ; 99 999]",
-  "[100 000 ; 149 999]",
-  "[150 000 et plus]",
+  "[0 ; 39 999]",
+  "[40 000 ; 39 999]",
+  "[60 000 ; 99 999]",
+  "[100 000 et plus]",
   NA)))
@@
     Q17 == 1 ~ "Catholique",
-    Q17 == 2 ~ "Protestante",
-    Q17 == 3 ~ "Autres Chretiens",
-    Q17 == 4 ~ "Autres Chretiens",
-    Q17 == 5 ~ "Autres Chretiens",
-    Q17 == 6 ~ "Orthodoxe",
-    Q17 == 7 ~ "Autres Chretiens",
+    Q17 == 2 ~ "Autres",
+    Q17 == 3 ~ "Autres",
+    Q17 == 4 ~ "Autres",
+    Q17 == 5 ~ "Autres",
+    Q17 == 6 ~ "Autres",
+    Q17 == 7 ~ "Autres",
     Q17 == 8 ~ "Musulmans",
-    Q17 == 9 ~ "Autre",
-    Q17 == 10 ~ "Autre",
-    Q17 == 11 ~ "Autre",
-    Q17 == 12 ~ "Autre",
-    Q17 == 97 ~ "Autre",
+    Q17 == 9 ~ "Autres",
+    Q17 == 10 ~ "Autres",
+    Q17 == 11 ~ "Autres",
+    Q17 == 12 ~ "Autres",
+    Q17 == 97 ~ "Autres",
     Q17 >= 98 ~ NA

### 2023-07-30 1. Trajipaq_sentiment/2. Codes/Trajipaq_sentiment_cr.Rmd
update CSD

@@
   mutate(revenu = factor(case_when( # Création de la variable "revenu".
-    Q100 == 1 ~ "[0 ; 39 999]",
-    Q100 == 2 ~ "[0 ; 39 999]",
-    Q100 == 3 ~ "[0 ; 39 999]",
-    Q100 == 4 ~ "[0 ; 39 999]",
-    Q100 == 5 ~ "[40 000 ; 99 999]",
-    Q100 == 6 ~ "[40 000 ; 99 999]",
-    Q100 == 7 ~ "[60 000 ; 99 999]",
-    Q100 == 8 ~ "[60 000 ; 99 999]",
+    Q100 == 1 ~ "[0 ; 19 999]",
+    Q100 == 2 ~ "[0 ; 19 999]",
+    Q100 == 3 ~ "[20 000 ; 29 999]",
+    Q100 == 4 ~ "[30 000 ; 39 999]",
+    Q100 == 5 ~ "[40 000 ; 49 999]",
+    Q100 == 6 ~ "[50 000 ; 59 999]",
+    Q100 == 7 ~ "[60 000 ; 69 999]",
+    Q100 == 8 ~ "[70 000 ; 79 999]",
     Q100 == 9 ~ "[80 000 ; 99 999]",
-    Q100 == 10 ~ "[100 000 et plus]",
+    Q100 == 10 ~ "[80 000 ; 99 999]",
     Q100 == 11 ~ "[100 000 et plus]",
@@
   ), levels = c(
-  "[0 ; 39 999]",
-  "[40 000 ; 39 999]",
-  "[60 000 ; 99 999]",
+  "[0 ; 19 999]",
+  "[20 000 ; 29 999]",
+  "[30 000 ; 39 999]",
+  "[40 000 ; 49 999]",
+  "[50 000 ; 59 999]",
+  "[60 000 ; 69 999]",
+  "[70 000 ; 79 999]",
+  "[80 000 ; 99 999]",
   "[100 000 et plus]",

### 2023-07-30 1. Trajipaq_sentiment/2. Codes/Trajipaq_sentiment_cr.Rmd
update CSD

--- 2023-07-30 1. Trajipaq_sentiment/2. Codes/Trajipaq_sentiment_cr.Rmd
@@
   mutate(revenu = factor(case_when( # Création de la variable "revenu".
-    Q100 == 1 ~ "[0 ; 39 999]",
-    Q100 == 2 ~ "[0 ; 39 999]",
-    Q100 == 3 ~ "[0 ; 39 999]",
-    Q100 == 4 ~ "[0 ; 39 999]",
-    Q100 == 5 ~ "[40 000 ; 99 999]",
-    Q100 == 6 ~ "[40 000 ; 99 999]",
-    Q100 == 7 ~ "[60 000 ; 99 999]",
-    Q100 == 8 ~ "[60 000 ; 99 999]",
+    Q100 == 1 ~ "[0 ; 19 999]",
+    Q100 == 2 ~ "[0 ; 19 999]",
+    Q100 == 3 ~ "[20 000 ; 39 999]",
+    Q100 == 4 ~ "[20 000 ; 39 999]",
+    Q100 == 5 ~ "[40 000 ; 49 999]",
+    Q100 == 6 ~ "[50 000 ; 59 999]",
+    Q100 == 7 ~ "[60 000 ; 79 999]",
+    Q100 == 8 ~ "[60 000 ; 79 999]",
     Q100 == 9 ~ "[80 000 ; 99 999]",
-    Q100 == 10 ~ "[100 000 et plus]",
+    Q100 == 10 ~ "[80 000 ; 99 999]",
     Q100 == 11 ~ "[100 000 et plus]",
@@
   ), levels = c(
-  "[0 ; 39 999]",
-  "[40 000 ; 39 999]",
-  "[60 000 ; 99 999]",
+  "[0 ; 19 999]",
+  "[20 000 ; 39 999]",
+  "[40 000 ; 49 999]",
+  "[50 000 ; 59 999]",
+  "[60 000 ; 79 999]",
+  "[80 000 ; 99 999]",
   "[100 000 et plus]",

### 2023-08-06 1. Trajipaq_sentiment/2. Codes/Trajipaq_sentiment_cr.Rmd
update wgi

@@
     Q16 == 2 ~ "Secondaire et moins",
-    Q16 == 3 ~ "Post secondaire",
-    Q16 == 4 ~ "Post secondaire",
-    Q16 == 5 ~ "Universitaire",
-    Q26 >= 6 ~ NA
+    Q16 == 3 ~ "Collegial",
+    Q16 == 4 ~ "Formation professionnelle ou postsecondaire",
+    Q16 == 5 ~ "Universitaire 1er cycle",
+    Q16 == 6 ~ "Universitaire 2e cycle et plus"
   ), levels = c( # Ordonnancement de la variable "education".
     "Secondaire et moins",
-    "Post secondaire",
-    "Universitaire")))
+    "Collegial",
+    "Formation professionnelle ou postsecondaire",
+    "Universitaire 1er cycle",
+    "Universitaire 2e cycle et plus")))
 
@@
     STATUT == 3 ~ "Conjoint(e) de fait",
-    STATUT == 4 ~ "Separe(e)",
-    STATUT == 5 ~ "Divorce(e)",
-    STATUT == 6 ~ "Veuf(ve)"
+    STATUT == 4 ~ "Separe(e)/divorce(e)/veuf(ve)",
+    STATUT == 5 ~ "Separe(e)/divorce(e)/veuf(ve)",
+    STATUT == 6 ~ "Separe(e)/divorce(e)/veuf(ve)"
   ))
@@
 
+V_WGI_origine_arrival <- c("vae_arrival", "pve_arrival", "gee_arrival", "rqe_arrival", "rle_arrival", "cce_arrival")
+
+trajipaq <- trajipaq %>%
+  rowwise()%>%
+  mutate(WGI_origine_arrival = sum(c_across(all_of(V_WGI_origine_arrival)), na.rm = TRUE)/length(V_WGI_origine_arrival))
+
+trajipaq <- trajipaq %>%
+  rowwise()%>%
+  mutate(WGI_origine_arrival = sum(c_across(all_of(V_WGI_origine_arrival)), na.rm = TRUE)/length(V_WGI_origine_arrival))
+
 trajipaq <- trajipaq %>% #enlever les répondants d'origine canadienne (977 restants)

### 2023-08-10 1. Trajipaq_sentiment/2. Codes/Trajipaq_sentiment_cr.Rmd
WGI update

@@
 wgidataset <- read_dta("../3. Data/wgidataset.dta")
+pays <- read_csv("../3. Data/pays.txt")
 
@@
 
-### VC4 : région d'origine
-
-```{r}
-
-# region_naissance = de quelle région du monde êtes-vous originaire ?
-
-trajipaq <-
-  trajipaq %>%
-  mutate(origine = case_when( # Création de la variable "region".
-    region_naissance == "ANMO" ~ "Afrique du Nord et Moyen Orient",
-    region_naissance == "ACAS" ~ "Amerique centrale et Amerique su Sud",
-    region_naissance == "Afrique subsaharienne" ~ "Afrique subsaharienne",
-    region_naissance == "Asie" ~ "Asie",
-    region_naissance == "Autres" ~ "Autres",
-    region_naissance == "Canada" ~ "Canada",
-    region_naissance == "Caraibes" ~ "Autres",
-    region_naissance == "Europe" ~ "Europe",
-    region_naissance == "France" ~ "France",
-    region_naissance == "NSP" ~ NA,
-  ))
-
-freq(trajipaq$origine)
-
-```
-
 ### VC5 : revenu annuel brut sur l'année 2019
@@
 
-### VD1 : sentiment d'appartenance au Québec
+### VD1 : sentiment d'appartenance (Canada)
 
 ```{r}
+# Q39E = Je me sens canadienne.
 
+trajipaq <-
+  trajipaq %>%
+  mutate(sentimapp_can_long = case_when( # Création de la variable "VD2".
+    Q39E == 1 ~ 1,
+    Q39E == 2 ~ 2,
+    Q39E == 3 ~ 3,
+    Q39E == 4 ~ 4,
+    Q39E == 98 ~ NA,
+    Q39E == 99 ~ NA
+  ))
+
+trajipaq <-
+  trajipaq %>%
+  mutate(sentimapp_can_dico = case_when( # Création de la variable "VD2".
+    Q39E == 1 ~ 1,
+    Q39E == 2 ~ 1,
+    Q39E == 3 ~ 0,
+    Q39E == 4 ~ 0,
+    Q39E == 98 ~ NA_real_,
+    Q39E == 99 ~ NA_real_
+  ))
+
+freq(trajipaq$sentimapp_can_long)
+freq(trajipaq$sentimapp_can_dico)
+```
+
+### VD2 : sentiment d'appartenance au Québec
+
+```{r}
 # Préparation des variables en vue de créer le premier indice.
@@
 
-v_VD1 <- c( # Création d'un vecteur regroupant les nouvelles variables.
+v_sentimapp_QC <- c( # Création d'un vecteur regroupant les nouvelles variables.
   "Q39_new_A",
@@
   rowwise() %>%
-  mutate(sumVD1 = sum(c_across(all_of(v_VD1)), na.rm = TRUE)) %>%
-  mutate(VD1 = (sumVD1 - nbre_rep_rep) / ((nbre_lvl - 1)*nbre_rep_rep)) %>%
-  mutate(VD1 = if (nbre_rep_rep == 0) NA_real_ else VD1) %>%
+  mutate(sumVD1 = sum(c_across(all_of(v_sentimapp_QC)), na.rm = TRUE)) %>%
+  mutate(sentimapp_QC = (sumVD1 - nbre_rep_rep) / ((nbre_lvl - 1)*nbre_rep_rep)) %>%
+  mutate(sentimapp_QC = if (nbre_rep_rep == 0) NA_real_ else sentimapp_QC) %>%
   ungroup()
 
-descr(trajipaq$VD1)
+descr(trajipaq$sentimapp_QC)
 
@@
 
-ggplot(data = trajipaq, mapping = aes(x = VD1)) +
+ggplot(data = trajipaq, mapping = aes(x = sentimapp_QC)) +
   geom_histogram(na.rm = TRUE, fill = "pink", binwidth = 0.01) +
@@
   theme_bw()
-
-```
-
-### VD2 : sentiment d'appartenance (Canada)
-
-```{r}
-
-# Q39E = Je me sens canadienne.
-
-trajipaq <-
-  trajipaq %>%
-  mutate(VD2_long = case_when( # Création de la variable "VD2".
-    Q39E == 1 ~ 1,
-    Q39E == 2 ~ 2,
-    Q39E == 3 ~ 3,
-    Q39E == 4 ~ 4,
-    Q39E == 98 ~ NA,
-    Q39E == 99 ~ NA
-  ))
-
-trajipaq <-
-  trajipaq %>%
-  mutate(VD2_dico = case_when( # Création de la variable "VD2".
-    Q39E == 1 ~ 1,
-    Q39E == 2 ~ 1,
-    Q39E == 3 ~ 0,
-    Q39E == 4 ~ 0,
-    Q39E == 98 ~ NA_real_,
-    Q39E == 99 ~ NA_real_
-  ))
-
-freq(trajipaq$VD2_long)
-freq(trajipaq$VD2_dico)
 
@@
   trajipaq %>%
-  mutate(VD3 = case_when( # Création de la variable "VD3".
-    Q39F == 1 ~ "Oui",
-    Q39F == 2 ~ "Oui",
-    Q39F == 3 ~ "Non",
-    Q39F == 4 ~ "Non",
+  mutate(sentimapp_origine_dico = case_when( # Création de la variable "VD3".
+    Q39F == 1 ~ 1,
+    Q39F == 2 ~ 1,
+    Q39F == 3 ~ 0,
+    Q39F == 4 ~ 0,
     Q39F >= 98 ~ NA
@@
 
-freq(trajipaq$VD3)
+freq(trajipaq$sentimapp_origine_dico)
 
@@
   trajipaq %>% 
-  mutate(VD4 = case_when( # Création de la variable "VD4".
-    VD2_dico == "Oui" & VD3 == "Oui" ~ "Integre",
-    VD2_dico == "Oui" & VD3 == "Non" ~ "Assimile",
-    VD2_dico == "Non" & VD3 == "Oui" ~ "Separe",
-    VD2_dico == "Non" & VD3 == "Non" ~ "Marginalise"
+  mutate(sentimapp_can_origine = case_when( # Création de la variable "VD4".
+    sentimapp_can_dico == 1 & sentimapp_origine_dico == 1 ~ "Integre",
+    sentimapp_can_dico == 1 & sentimapp_origine_dico == 0 ~ "Assimile",
+    sentimapp_can_dico == 0 & sentimapp_origine_dico == 1 ~ "Separe",
+    sentimapp_can_dico == 0 & sentimapp_origine_dico == 0 ~ "Marginalise"
   ))
 
-freq(trajipaq$VD4)
+freq(trajipaq$sentimapp_can_origine)
 
@@
 
-trajipaq <-
-  trajipaq %>% 
-  mutate(Q39D_new_2 = case_when( # Création d'une nouvelle variable.
-         Q39D <= 2 ~ "Oui", 
-         Q39D == 3 | Q39D == 4 ~ "Non", 
-         Q39D >= 97 ~ NA_character_)) %>% 
-  mutate(VD5 = factor(case_when( # Création de la variable "VD5".
-    VD2_dico == "Oui" & Q39D_new_2 == "Oui" ~ 1,
-    VD2_dico == "Oui" & Q39D_new_2 == "Non" ~ 2,
-    VD2_dico == "Non" & Q39D_new_2 == "Oui" ~ 3,
-    VD2_dico == "Non" & Q39D_new_2 == "Non" ~ 4
-  ),
-  labels = c( # Création du nom des catégories.
-    "Canada et Quebec", 
-    "Canada seule", 
-    "Quebec seule", 
-    "Aucun"),
-  levels = c( # Ordonnancement des catégories.
-    1, 
-    2, 
-    3, 
-    4)))
+#trajipaq <-
+#  trajipaq %>% 
+#  mutate(Q39D_new_2 = case_when( # Création d'une nouvelle variable.
+ #        Q39D <= 2 ~ "Oui", 
+  #       Q39D == 3 | Q39D == 4 ~ "Non", 
+   #      Q39D >= 97 ~ NA_character_)) %>% 
+#  mutate(VD5 = factor(case_when( # Création de la variable "VD5".
+ #   VD2_dico == "Oui" & Q39D_new_2 == "Oui" ~ 1,
+  #  VD2_dico == "Oui" & Q39D_new_2 == "Non" ~ 2,
+   # VD2_dico == "Non" & Q39D_new_2 == "Oui" ~ 3,
+    #VD2_dico == "Non" & Q39D_new_2 == "Non" ~ 4
+#  ),
+ # labels = c( # Création du nom des catégories.
+  #  "Canada et Quebec", 
+   # "Canada seule", 
+    #"Quebec seule", 
+    #"Aucun"),
+  #levels = c( # Ordonnancement des catégories.
+   # 1, 
+    #2, 
+    #3, 
+    #4)))
 
-freq(trajipaq$VD5)
+#freq(trajipaq$VD5)
 
@@
   mutate(sumVI4 = sum(c_across(all_of(v_VI4)), na.rm = TRUE)) %>%
-  mutate(VI4 = (sumVI4 - nbre_rep_rep_2) / ((nbre_lvl_2 - 1)*nbre_rep_rep_2)) %>%
-  mutate(VI4 = if (nbre_rep_rep_2 == 0) NA_real_ else VI4) %>%
+  mutate(discrim_score = (sumVI4 - nbre_rep_rep_2) / ((nbre_lvl_2 - 1)*nbre_rep_rep_2)) %>%
+  mutate(discrim_score = if (nbre_rep_rep_2 == 0) NA_real_ else discrim_score) %>%
   ungroup()
 
-descr(trajipaq$VI4)
+descr(trajipaq$discrim_score)
 
@@
 
-ggplot(data = trajipaq, mapping = aes(x = VI4)) +
+ggplot(data = trajipaq, mapping = aes(x = discrim_score)) +
   geom_histogram(na.rm = TRUE, fill = "pink", binwidth = 0.01) +
@@
 ```
+### VC4 : région d'origine
 
+```{r}
+
+# region_naissance = de quelle région du monde êtes-vous originaire ?
+trajipaq <- as_factor(trajipaq, only_labelled = TRUE)
+
+trajipaq <- merge(trajipaq, pays, by.x = "Q11",
+                by.y = "Country Name", all.x = TRUE)
+
+trajipaq <-
+  trajipaq %>%
+  mutate(origine = Region)
+
+trajipaq <-
+  trajipaq %>%
+  mutate(origine = case_when( # Création de la variable "region".
+    Subregion == "Amérique du Nord" ~ "Amérique du Nord",
+    Subregion == "Amérique latine et Caraïbes" ~ "Amérique latine et Caraïbes",
+    Subregion == "Afrique subsaharienne" ~ "Afrique subsaharienne",
+    Subregion == "Afrique du Nord" ~ "Afrique du Nord",
+    Subregion == "Afrique subsaharienne" ~ "Afrique subsaharienne",
+    TRUE ~ origine
+  ))
+
+freq(trajipaq$origine)
+
+```
 # WGI
@@
 ```{r}
-trajipaq <- as_factor(trajipaq, only_labelled = TRUE)
-
 wgidataset <-
@@
 trajipaq <- trajipaq %>% #enlever les répondants d'origine canadienne (977 restants)
-  filter(origine != "Canada")
-
-freq(trajipaq$origine)
+  filter(Q11 != "Canada")
 
@@
          "Q11",
-         "VD1",
-         "VD2_dico",
-         "VD2_long",
-         "VD3",
-         "VD4",
-         "VD5",
-         "VI4",
+         "sentimapp_QC",
+         "sentimapp_can_dico",
+         "sentimapp_can_long",
+         "sentimapp_origine_dico",
+         "sentimapp_can_origine",
+         "discrim_score",
          "vae_arrival",

### 2023-08-10 1. Trajipaq_sentiment/2. Codes/Trajipaq_sentiment_cr.Rmd
WGI update

@@
          "cce_arrival_diff"
-
          )

### 2023-08-10 1. Trajipaq_sentiment/2. Codes/Trajipaq_sentiment_cr.Rmd
update pays/origine
update origine

@@
 
+pays <-
+  pays %>%
+    mutate(Country = recode(Country,
+                            "Russie" = "Russie, Fédération de",
+                            "Congo (Kinshasa)" = "Congo, République démocratique du",
+                            "Congo (Brazzaville)" = "Congo, République du",
+                            "République dominicaine" = "Dominicaine, République",
+                            "République centrafricaine" = "Centrafricaine, République",
+                            "Afrique du Sud" = "Afrique du Sud, République d'"
+  ))
+
 trajipaq <- merge(trajipaq, pays, by.x = "Q11",
-                by.y = "Country Name", all.x = TRUE)
-
+                by.y = "Country", all.x = TRUE)
 trajipaq <-

### 2023-08-10 1. Trajipaq_sentiment/2. Codes/Trajipaq_sentiment_cr.Rmd
update WGI

@@
 
-trajipaq <- merge(trajipaq, wgidataset, by.x = c("annee_qc", "Q11"),
+trajipaq <- merge(trajipaq, wgidataset, by.x = c("annee_qc", "Q11"), # Origine à l'arrivée
                 by.y = c("year", "countryname"), all.x = TRUE)
 
-wgidataset_2021 <- filter(wgidataset, year == 2021)
-wgidataset_canada <- filter(wgidataset, countryname == "Canada")
+wgidataset_origin_2021 <- filter(wgidataset, year == 2021) # Origine en 2021
+wgidataset_canada <- filter(wgidataset, countryname == "Canada") #Canada à l'arrivée
+wgidataset_canada_2021 <- filter(wgidataset, countryname == "Canada" & year == 2021)
 
 trajipaq <- trajipaq %>%
-  left_join(wgidataset_2021, by = c("Q11" = "countryname"))%>%
+  left_join(wgidataset_origin_2021, by = c("Q11" = "countryname"))%>%
   left_join(wgidataset_canada, by = c("annee_qc" = "year"))
@@
 trajipaq <- trajipaq %>%
-  rename_with(~sub("\\.y$", "_2021", .), ends_with(".y"))%>%
-  rename_with(~sub("\\.x$", "_arrival", .), ends_with(".x"))
+  rename_with(~sub("\\.y$", "_2021_origin", .), ends_with(".y"))%>%
+  rename_with(~sub("\\.x$", "_arrival_canada", .), ends_with(".x"))
 

### 2023-08-11 1. Trajipaq_sentiment/2. Codes/Trajipaq_sentiment_cr.Rmd
wgi, presque fini

@@
 ```
-# WGI
+## WGI
 
@@
                               ))
+```
+
+### WGI origin/canada et arrival/2021
+
+```{r}
+wgi_origin_2021 <- filter(wgidataset, year == 2021) # Origine en 2021
+wgi_can <- filter(wgidataset, countryname == "Canada") #Canada à l'arrivée
+wgi_can_2021 <- filter(wgidataset, countryname == "Canada" & year == 2021)
+  col_can_2021 <- c("vae", "pve", "gee", "rqe", "rle", "cce")
 
@@
 
-wgidataset_origin_2021 <- filter(wgidataset, year == 2021) # Origine en 2021
-wgidataset_canada <- filter(wgidataset, countryname == "Canada") #Canada à l'arrivée
-wgidataset_canada_2021 <- filter(wgidataset, countryname == "Canada" & year == 2021)
+trajipaq <- trajipaq %>%
+  left_join(wgi_origin_2021, by = c("Q11" = "countryname"), suffix = c("", "_origin_arrival")) %>%
+  left_join(wgi_can, by = c("annee_qc" = "year"), suffix = c("", "_origin_2021")) %>%
+  left_join(wgi_can_2021, select(wgi_can_2021, all_of(col_can_2021)), 
+            by = character(0), suffix = c("", "_can_2021"))%>%
+  mutate(vae_can_arrival = vae,
+         pve_can_arrival = pve,
+         gee_can_arrival = gee,
+         rqe_can_arrival = rqe,
+         rle_can_arrival = rle,
+         cce_can_arrival = cce) %>%
+  select(-vae, -pve, -gee, -rqe, -rle, -cce)
+
+```
+
+### Score WGI
+
+```{r}
+V_origin_arrival <- c("vae_origin_arrival", "pve_origin_arrival", "gee_origin_arrival", "rqe_origin_arrival", "rle_origin_arrival", "cce_origin_arrival")
+V_origin_2021 <- c("vae_origin_2021", "pve_origin_2021", "gee_origin_2021", "rqe_origin_2021", "rle_origin_2021", "cce_origin_2021")
+V_can_arrival <- c("vae_can_arrival", "pve_can_arrival", "gee_can_arrival", "rqe_can_arrival", "rle_can_arrival", "cce_can_arrival")
+V_can_2021 <- c("vae_can_2021", "pve_can_2021", "gee_can_2021", "rqe_can_2021", "rle_can_2021", "cce_can_2021")
 
 trajipaq <- trajipaq %>%
-  left_join(wgidataset_origin_2021, by = c("Q11" = "countryname"))%>%
-  left_join(wgidataset_canada, by = c("annee_qc" = "year"))
+  rowwise()%>%
+  mutate(wgi_origin_arrival_score = sum(c_across(all_of(V_origin_arrival)), na.rm = TRUE)/length(V_origin_arrival)) %>%
+  mutate(wgi_origin_2021_score = sum(c_across(all_of(V_origin_2021)), na.rm = TRUE)/length(V_origin_2021)) %>%
+  mutate(wgi_can_arrival_score = sum(c_across(all_of(V_can_arrival)), na.rm = TRUE)/length(V_can_arrival)) %>%
+  mutate(wgi_can_2021_score = sum(c_across(all_of(V_can_2021)), na.rm = TRUE)/length(V_can_2021))
 
-trajipaq <- trajipaq %>%
-  rename_with(~sub("\\.y$", "_2021_origin", .), ends_with(".y"))%>%
-  rename_with(~sub("\\.x$", "_arrival_canada", .), ends_with(".x"))
-
-trajipaq <- rename(trajipaq, "vae_can" = "vae")
-trajipaq <- rename(trajipaq, "pve_can" = "pve")
-trajipaq <- rename(trajipaq, "gee_can" = "gee")
-trajipaq <- rename(trajipaq, "rqe_can" = "rqe")
-trajipaq <- rename(trajipaq, "rle_can" = "rle")
-trajipaq <- rename(trajipaq, "cce_can" = "cce")
+```
 
@@
 
-V_WGI_origine_arrival <- c("vae_arrival", "pve_arrival", "gee_arrival", "rqe_arrival", "rle_arrival", "cce_arrival")
-
-trajipaq <- trajipaq %>%
-  rowwise()%>%
-  mutate(WGI_origine_arrival = sum(c_across(all_of(V_WGI_origine_arrival)), na.rm = TRUE)/length(V_WGI_origine_arrival))
-
-trajipaq <- trajipaq %>%
-  rowwise()%>%
-  mutate(WGI_origine_arrival = sum(c_across(all_of(V_WGI_origine_arrival)), na.rm = TRUE)/length(V_WGI_origine_arrival))
-
-trajipaq <- trajipaq %>% #enlever les répondants d'origine canadienne (977 restants)
-  filter(Q11 != "Canada")
-
-```
-
 # Création de la nouvelle base de données
@@
   trajipaq %>%
+  filter(Q11 != "Canada")%>%
   select("age", # Création de la base de données dédiée à l'analyse.
@@
          "discrim_score",
-         "vae_arrival",
-         "pve_arrival",
-         "gee_arrival",
-         "rqe_arrival",
-         "rle_arrival",
-         "cce_arrival",
-         "vae_2021",
-         "pve_2021",
-         "gee_2021",
-         "rqe_2021",
-         "rle_2021",
-         "cce_2021",
-         "vae_can",
-         "pve_can",
-         "gee_can",
-         "rqe_can",
-         "rle_can",
-         "cce_can",
-         "vae_arrival_diff",
-         "pve_arrival_diff",
-         "gee_arrival_diff",
-         "rqe_arrival_diff",
-         "rle_arrival_diff",
-         "cce_arrival_diff"
+         starts_with("vae"),
+         starts_with("pve"),
+         starts_with("gee"),
+         starts_with("rqe"),
+         starts_with("rle"),
+         starts_with("cce"),
+         ends_with("score")
          )

### 2023-08-12 1. Trajipaq_sentiment/2. Codes/Trajipaq_sentiment_an.Rmd
recode CSD
brain about to blow up. Break needed

@@
 library(stargazer)
-library(gtsumamry)
+library(gtsummary)
 
@@
 
-reg_WGI_can1 <- glm(formula = VD2_dico ~ vae_arrival, data = trajipaq, family = "binomial") # VD2 dichotomique
-reg_WGI_can1_ctrl <- glm(formula = VD2_dico ~ vae_arrival  + age + education + sexe + origine + revenu + langue_mat + religion + statut_mat , data = trajipaq, family = "binomial") #VD2 contrôlée par CSD
+reg_WGI_can_arrival_dico <- glm(formula = sentimapp_can_dico ~ wgi_origin_arrival_score, data = trajipaq, family = "binomial") # VD2 dichotomique
+reg_WGI_can_arrival_dico_ctrl <- glm(formula = sentimapp_can_dico ~ wgi_origin_arrival_score  + age + education + sexe + origine + revenu + langue_mat + religion + statut_mat , data = trajipaq, family = "binomial") #VD2 contrôlée par CSD
 
-summary(reg_WGI_can1)
-summary(reg_WGI_can1_ctrl)
+reg_WGI_can_arrival_long <- lm(formula = sentimapp_can_long ~ wgi_origin_arrival_score, data = trajipaq) # VD2 dichotomique
+reg_WGI_can_arrival_long_ctrl <- lm(formula = sentimapp_can_long ~ wgi_origin_arrival_score  + age + education + sexe + origine + revenu + langue_mat + religion + statut_mat , data = trajipaq) #VD2 contrôlée par CSD
+
+summary(reg_WGI_can_arrival_dico) # p = 0.000116 ***
+summary(reg_WGI_can_arrival_dico_ctrl) # p = 0.0203 *
+summary(reg_WGI_can_arrival_long) # p = 8.7e-07 ***
+summary(reg_WGI_can_arrival_long_ctrl) # p = 0.00678 **
 
@@
 
-reg_WGI_can2 <- glm(formula = VD2_dico ~ vae_2021 , data = trajipaq, family = "binomial") # VD2 non regroupée
-reg_WGI_can2_ctrl <- glm(formula = VD2_dico ~ vae_2021  + age + education + sexe + origine + revenu + langue_mat + religion + statut_mat , data = trajipaq, family = "binomial")#VD2 contrôlée par CSD
+reg_WGI_can_2021_dico <- glm(formula = sentimapp_can_dico ~ wgi_origin_2021_score , data = trajipaq, family = "binomial") ## VD2 dichotomique
+reg_WGI_can_2021_dico_ctrl <- glm(formula = sentimapp_can_dico ~ wgi_origin_2021_score  + age + education + sexe + origine + revenu + langue_mat + religion + statut_mat , data = trajipaq, family = "binomial")#VD2 contrôlée par CSD
 
-summary(reg_WGI_can2)
-summary(reg_WGI_can2_ctrl)
+reg_WGI_can_2021_long <- lm(formula = sentimapp_can_long ~ wgi_origin_2021_score, data = trajipaq) # VD2 dichotomique
+reg_WGI_can_2021_long_ctrl <- lm(formula = sentimapp_can_long ~ wgi_origin_2021_score  + age + education + sexe + origine + revenu + langue_mat + religion + statut_mat , data = trajipaq) #VD2 contrôlée par CSD
+
+summary(reg_WGI_can_2021_dico) # 2.38e-05 ***
+summary(reg_WGI_can_2021_dico_ctrl) # p = 0.03426 *
+summary(reg_WGI_can_2021_long) # p = 5.5e-08 ***
+summary(reg_WGI_can_2021_long_ctrl) # p = 0.0124 *
+```
+
+```{r}
+stargazer(
+  reg_WGI_can_2021_long,
+  reg_WGI_can_2021_long_ctrl,
+  reg_WGI_can_arrival_long,
+  reg_WGI_can_arrival_long_ctrl,
+  title = "Results", 
+  align = TRUE, 
+  type = "text"
+)
+
+stargazer(
+  reg_WGI_can_2021_dico,
+  reg_WGI_can_2021_dico_ctrl,
+  title = "Results", 
+  align = TRUE, 
+  type = "text"
+)
+
+stargazer(
+  reg_WGI_can_2021_long,
+  reg_WGI_can_2021_long_ctrl,
+  title = "Results", 
+  align = TRUE, 
+  type = "text"
+)
+
+stargazer(
+  reg_WGI_can_2021_dico,
+  reg_WGI_can_2021_long_ctrl,
+  title = "Results", 
+  align = TRUE, 
+  type = "text"
+)
+
+stargazer(
+  reg_WGI_can_2021_long,
+  reg_WGI_can_2021_dico_ctrl,
+  title = "Results", 
+  align = TRUE, 
+  type = "text"
+)
+
+stargazer(
+  reg_WGI_can_2021_long,
+  reg_WGI_can_2021_long_ctrl,
+  reg_WGI_can_2021_dico,
+  title = "Results", 
+  align = TRUE, 
+  type = "text"
+)
+
+stargazer(
+  reg_WGI_can_2021_long,
+  reg_WGI_can_2021_long_ctrl,
+  reg_WGI_can_2021_dico_ctrl,
+  title = "Results", 
+  align = TRUE, 
+  type = "text"
+)
+
+stargazer(
+  reg_WGI_can_2021_long,
+  reg_WGI_can_2021_long_ctrl,
+  title = "Results", 
+  align = TRUE, 
+  type = "text"
+)
+
+stargazer(
+  reg_WGI_can_2021_long,
+  reg_WGI_can_2021_long_ctrl,
+  title = "Results", 
+  align = TRUE, 
+  type = "text"
+)
+
+stargazer(
+  reg_WGI_can_2021_dico,
+  reg_WGI_can_2021_dico_ctrl,
+  reg_WGI_can_2021_long,
+  reg_WGI_can_2021_long_ctrl,
+  title = "Results", 
+  align = TRUE, 
+  type = "text"
+)
+
+ # reg_WGI_can_2021_dico,
+ # reg_WGI_can_2021_dico_ctrl,
+
+freq(trajipaq$sentimapp_can_dico)
+freq(trajipaq$wgi_origin_arrival_score)
+freq(trajipaq$sentimapp_can_long)
+freq(trajipaq$wgi_origin_2021_score)
+
+trajipaq$age <- as.factor(trajipaq$age)
+trajipaq$education <- as.factor(trajipaq$education)
+trajipaq$sexe <- as.factor(trajipaq$sexe)
+trajipaq$origine <- as.factor(trajipaq$origine)
+trajipaq$revenu <- as.factor(trajipaq$revenu)
+trajipaq$langue_mat <- as.factor(trajipaq$langue_mat)
+trajipaq$religion <- as.factor(trajipaq$religion)
+trajipaq$statut_mat <- as.factor(trajipaq$statut_mat)
+
+```
 
@@
 
-reg_WGI_can3 <- glm(formula = VD2_dico ~ vae_2021 + vae_arrival , data = trajipaq, family = "binomial") # VD2 non regroupée
-reg_WGI_can3_ctrl <- glm(formula = VD2_dico ~ vae_2021 + vae_arrival + age + education + sexe + origine + revenu + langue_mat + religion + statut_mat , data = trajipaq, family = "binomial")# VD2 contrôlée par CSD
+reg_WGI_can_both_dico <- glm(formula = sentimapp_can_dico ~ wgi_origin_2021_score + wgi_origin_arrival_score , data = trajipaq, family = "binomial") # VD2 non regroupée
+reg_WGI_can_both_dico_ctrl <- glm(formula = sentimapp_can_dico ~ wgi_origin_2021_score + wgi_origin_arrival_score + age + education + sexe + origine + revenu + langue_mat + religion + statut_mat , data = trajipaq, family = "binomial")# VD2 contrôlée par CSD
 
-summary(reg_WGI_can3)
-summary(reg_WGI_can3_ctrl)
+summary(reg_WGI_can_both_dico)
+summary(reg_WGI_can_both_dico_ctrl)
 
@@
 
-```
-
-# WGI X Québec
+## WGI X Québec
 
@@
 
-reg_WGI_qc1 <- lm(formula = VD1  ~ vae_arrival, data = trajipaq) # VD1
+reg__qc1 <- lm(formula = VD1  ~ vae_arrival, data = trajipaq) # VD1
 reg_WGI_qc1_ctrl <- lm(formula = VD1 ~ vae_arrival  + age + education + sexe + origine + revenu + langue_mat + religion + statut_mat , data = trajipaq) #VD1 contrôlée par CSD
@@
 ```
-Résumé des résultats Can + QC
+
+## WGI X can_qc
+
 ```{r}
+reg_WGI_can_origin <- lm(formula = sentimapp_can_origin_N ~ wgi_origin_arrival_score, data = trajipaq, family = "binomial") # VD2 dichotomique
+
+reg_WGI_can_origin_ctrl <- lm(formula = sentimapp_can_origin_N ~ wgi_origin_arrival_score  + age + education + sexe + origine + revenu + langue_mat + religion + statut_mat , data = trajipaq, family = "binomial") #VD2 contrôlée par CSD
+
+summary(reg_WGI_can_origin)
+summary(reg_WGI_can_origin_ctrl)
+
 stargazer(reg_WGI_can2, reg_WGI_qc2, reg_WGI_can2_ctrl, reg_WGI_qc2_ctrl, title="Results",align=TRUE, type = "text")

### 2023-08-12 1. Trajipaq_sentiment/2. Codes/Trajipaq_sentiment_cr.Rmd
brain about to blow up. Break needed

@@
   trajipaq %>% 
-  mutate(sentimapp_can_origine = case_when( # Création de la variable "VD4".
+  mutate(sentimapp_can_origin = case_when( # Création de la variable "VD4".
     sentimapp_can_dico == 1 & sentimapp_origine_dico == 1 ~ "Integre",
@@
 
-freq(trajipaq$sentimapp_can_origine)
+trajipaq <-
+  trajipaq %>% 
+  mutate(sentimapp_can_origin_N = case_when( # Création de la variable "VD4".
+    sentimapp_can_dico == 1 & sentimapp_origine_dico == 1 ~ "1",
+    sentimapp_can_dico == 1 & sentimapp_origine_dico == 0 ~ "2",
+    sentimapp_can_dico == 0 & sentimapp_origine_dico == 1 ~ "3",
+    sentimapp_can_dico == 0 & sentimapp_origine_dico == 0 ~ "4"
+  ))
+freq(trajipaq$sentimapp_can_origin)
 
@@
 trajipaq <- trajipaq %>%
-  left_join(wgi_origin_2021, by = c("Q11" = "countryname"), suffix = c("", "_origin_arrival")) %>%
-  left_join(wgi_can, by = c("annee_qc" = "year"), suffix = c("", "_origin_2021")) %>%
+  left_join(wgi_origin_2021, by = c("Q11" = "countryname"), suffix = c("", "_origin_2021")) %>%
+  left_join(wgi_can, by = c("annee_qc" = "year"), suffix = c("", "_can_arrival")) %>%
   left_join(wgi_can_2021, select(wgi_can_2021, all_of(col_can_2021)), 
             by = character(0), suffix = c("", "_can_2021"))%>%
-  mutate(vae_can_arrival = vae,
-         pve_can_arrival = pve,
-         gee_can_arrival = gee,
-         rqe_can_arrival = rqe,
-         rle_can_arrival = rle,
-         cce_can_arrival = cce) %>%
+  mutate(vae_origin_arrival = vae,
+         pve_origin_arrival = pve,
+         gee_origin_arrival = gee,
+         rqe_origin_arrival = rqe,
+         rle_origin_arrival = rle,
+         cce_origin_arrival = cce) %>%
   select(-vae, -pve, -gee, -rqe, -rle, -cce)
@@
   mutate(wgi_origin_arrival_score = sum(c_across(all_of(V_origin_arrival)), na.rm = TRUE)/length(V_origin_arrival)) %>%
+  mutate(wgi_origin_arrival_score = if (wgi_origin_arrival_score == 0) NA_real_ else wgi_origin_arrival_score) %>%
   mutate(wgi_origin_2021_score = sum(c_across(all_of(V_origin_2021)), na.rm = TRUE)/length(V_origin_2021)) %>%
+  mutate(wgi_origin_2021_score = if (wgi_origin_2021_score == 0) NA_real_ else wgi_origin_2021_score) %>%
   mutate(wgi_can_arrival_score = sum(c_across(all_of(V_can_arrival)), na.rm = TRUE)/length(V_can_arrival)) %>%
-  mutate(wgi_can_2021_score = sum(c_across(all_of(V_can_2021)), na.rm = TRUE)/length(V_can_2021))
+  mutate(wgi_can_arrival_score = if (wgi_can_arrival_score == 0) NA_real_ else wgi_can_arrival_score) %>%
+  mutate(wgi_can_2021_score = sum(c_across(all_of(V_can_2021)), na.rm = TRUE)/length(V_can_2021)) %>%
+  mutate(wgi_can_2021_score = if (wgi_can_2021_score == 0) NA_real_ else wgi_can_2021_score)
 
@@
          "sentimapp_origine_dico",
-         "sentimapp_can_origine",
+         "sentimapp_can_origin",
+         "sentimapp_can_origin_N",
          "discrim_score",

### 2023-08-12 1. Trajipaq_sentiment/2. Codes/Trajipaq_sentiment_an.Rmd
again

@@
   reg_WGI_can_arrival_long,
-  reg_WGI_can_arrival_long_ctrl,
   title = "Results", 

### 2023-08-12 1. Trajipaq_sentiment/2. Codes/Trajipaq_sentiment_an.Rmd
wtv

@@
 
-reg_WGI_can_2021_dico <- glm(formula = sentimapp_can_dico ~ wgi_origin_2021_score , data = trajipaq, family = "binomial") ## VD2 dichotomique
+reg_WGI_can_2021_dico <- glm(formula = sentimapp_can_dico ~ wgi_origin_2021_score , data = trajipaq, family = "binomial", na.action = na.pass) ## VD2 dichotomique
 reg_WGI_can_2021_dico_ctrl <- glm(formula = sentimapp_can_dico ~ wgi_origin_2021_score  + age + education + sexe + origine + revenu + langue_mat + religion + statut_mat , data = trajipaq, family = "binomial")#VD2 contrôlée par CSD
@@
 stargazer(
-  reg_WGI_can_2021_long,
-  reg_WGI_can_2021_long_ctrl,
-  reg_WGI_can_arrival_long,
-  title = "Results", 
-  align = TRUE, 
-  type = "text"
-)
-
-stargazer(
-  reg_WGI_can_2021_dico,
-  reg_WGI_can_2021_dico_ctrl,
-  title = "Results", 
-  align = TRUE, 
-  type = "text"
-)
-
-stargazer(
-  reg_WGI_can_2021_long,
-  reg_WGI_can_2021_long_ctrl,
-  title = "Results", 
-  align = TRUE, 
-  type = "text"
-)
-
-stargazer(
-  reg_WGI_can_2021_dico,
-  reg_WGI_can_2021_long_ctrl,
-  title = "Results", 
-  align = TRUE, 
-  type = "text"
-)
-
-stargazer(
-  reg_WGI_can_2021_long,
-  reg_WGI_can_2021_dico_ctrl,
-  title = "Results", 
-  align = TRUE, 
-  type = "text"
-)
-
-stargazer(
-  reg_WGI_can_2021_long,
-  reg_WGI_can_2021_long_ctrl,
-  reg_WGI_can_2021_dico,
-  title = "Results", 
-  align = TRUE, 
-  type = "text"
-)
-
-stargazer(
-  reg_WGI_can_2021_long,
-  reg_WGI_can_2021_long_ctrl,
-  reg_WGI_can_2021_dico_ctrl,
-  title = "Results", 
-  align = TRUE, 
-  type = "text"
-)
-
-stargazer(
-  reg_WGI_can_2021_long,
-  reg_WGI_can_2021_long_ctrl,
-  title = "Results", 
-  align = TRUE, 
-  type = "text"
-)
-
-stargazer(
-  reg_WGI_can_2021_long,
-  reg_WGI_can_2021_long_ctrl,
-  title = "Results", 
-  align = TRUE, 
-  type = "text"
-)
-
-stargazer(
   reg_WGI_can_2021_dico,
@@
   reg_WGI_can_2021_long,
-  reg_WGI_can_2021_long_ctrl,
   title = "Results", 

### 2023-08-15 1. Trajipaq_sentiment/2. Codes/Trajipaq_sentiment_an.Rmd
update

@@
 summary(reg_WGI_can_arrival_dico) # p = 0.000116 ***
+  # Plus le score WGI du pays d'origine au moment de l'immigration est élevé, plus le sentiment d'appartenance au Canada est faible. Relation négative.
+
 summary(reg_WGI_can_arrival_dico_ctrl) # p = 0.0203 *
+  # aucune variable de contrôle significative (age [50;99] p = 0.0894)
+
 summary(reg_WGI_can_arrival_long) # p = 8.7e-07 ***
+  # Plus le score WGI du pays d'origine au moment del'immigration est élevé, plus le sentiment d'appartenance au Canada est élevé. Relation positive
+
 summary(reg_WGI_can_arrival_long_ctrl) # p = 0.00678 **
+  # aucune variable de contrôle significative
 
@@
 
-reg_WGI_can_2021_dico <- glm(formula = sentimapp_can_dico ~ wgi_origin_2021_score , data = trajipaq, family = "binomial", na.action = na.pass) ## VD2 dichotomique
+reg_WGI_can_2021_dico <- glm(formula = sentimapp_can_dico ~ wgi_origin_2021_score , data = trajipaq, family = "binomial") ## VD2 dichotomique
 reg_WGI_can_2021_dico_ctrl <- glm(formula = sentimapp_can_dico ~ wgi_origin_2021_score  + age + education + sexe + origine + revenu + langue_mat + religion + statut_mat , data = trajipaq, family = "binomial")#VD2 contrôlée par CSD
@@
 summary(reg_WGI_can_2021_dico) # 2.38e-05 ***
+  # Plus le score WGI du pays d'origine au moment de l'enquête est élevé, plus le sentiment d'appartenance au Canada est faible. Relation négative.
+
 summary(reg_WGI_can_2021_dico_ctrl) # p = 0.03426 *
+  # Seul l'âge [50;99] estr significative p = 0.04311
+
 summary(reg_WGI_can_2021_long) # p = 5.5e-08 ***
+  # Plus le score WGI du pays d'origine au moment de l'enquête est élevé, plus le sentiment d'appartenance au Canada est élevé. Relation positive.
+
 summary(reg_WGI_can_2021_long_ctrl) # p = 0.0124 *
+  # Seul l'âge [50;99] est significative p = 0.0210 et Origin (Océanie p = 0.0369)
+```
+
+## Visualisation
+
+```{r}
+
+ggplot(trajipaq %>% filter(!is.na(wgi_origin_arrival_score), !is.na(sentimapp_can_dico)), aes(x = wgi_origin_arrival_score, y = sentimapp_can_dico)) +
+  stat_smooth(method = "glm",
+              formula = "y ~ x",
+              color = "pink", 
+              se = FALSE,
+              method.args = list(family = binomial)) +
+  labs(
+    title = "Sentiment d'appartenance au Canada (dico) selon le score \nde gouvernementalité du pays d'origine au moment de l'immigraition", 
+    x = "Score de gouvernementalité du pays d'origine", 
+    y = "Sentiment d'appartenance au Canada") +
+  theme_bw()
+
+ggplot(trajipaq %>% filter(!is.na(wgi_origin_arrival_score), !is.na(sentimapp_can_long)), aes(x = wgi_origin_arrival_score, y = sentimapp_can_long)) +
+  stat_smooth(method = "lm",
+              formula = "y ~ x",
+              color = "pink", 
+              se = FALSE) +
+  labs(
+    title = "Sentiment d'appartenance au Canada (long) selon le score \nde gouvernementalité du pays d'origine au moment de l'immigraition", 
+    x = "Score de gouvernementalité du pays d'origine", 
+    y = "Sentiment d'appartenance au Canada") +
+  theme_bw()
+
+ggplot(trajipaq %>% filter(!is.na(wgi_origin_2021_score), !is.na(sentimapp_can_dico)), aes(x = wgi_origin_2021_score, y = sentimapp_can_dico)) +
+  stat_smooth(method = "glm",
+              formula = "y ~ x",
+              color = "pink", 
+              se = FALSE) +
+  labs(
+    title = "Sentiment d'appartenance au Canada (long) selon le score \nde gouvernementalité du pays d'origine au moment de l'immigraition", 
+    x = "Score de gouvernementalité du pays d'origine", 
+    y = "Sentiment d'appartenance au Canada") +
+  theme_bw()
+
+ggplot(trajipaq %>% filter(!is.na(wgi_origin_2021_score), !is.na(sentimapp_can_long)), aes(x = wgi_origin_2021_score, y = sentimapp_can_long)) +
+  stat_smooth(method = "lm",
+              formula = "y ~ x",
+              color = "pink", 
+              se = FALSE) +
+  labs(
+    title = "Sentiment d'appartenance au Canada (long) selon le score \nde gouvernementalité du pays d'origine au moment de l'immigraition", 
+    x = "Score de gouvernementalité du pays d'origine", 
+    y = "Sentiment d'appartenance au Canada") +
+  theme_bw()
+
 ```

### 2023-08-15 1. Trajipaq_sentiment/2. Codes/Trajipaq_sentiment_cr.Rmd
ajout WGI diff

@@
 
+trajipaq <- trajipaq %>%
+  rowwise() %>%
+  mutate(wgi_arrival_diff = wgi_origin_arrival_score - wgi_can_arrival_score) %>%
+  mutate(wgi_2021_diff = wgi_origin_2021_score - wgi_can_2021_score)
+
 ```
-
-trajipaq <- trajipaq %>% # création des WGI diff (can-home)
-  rowwise() %>%
-  mutate(vae_arrival_diff = vae_can - vae_arrival) %>%
-  mutate(pve_arrival_diff = pve_can - pve_arrival) %>%
-  mutate(gee_arrival_diff = gee_can - gee_arrival) %>%
-  mutate(rqe_arrival_diff = rqe_can - rqe_arrival) %>%
-  mutate(rle_arrival_diff = rle_can - rle_arrival) %>%
-  mutate(cce_arrival_diff = cce_can - cce_arrival)
 
@@
          starts_with("cce"),
-         ends_with("score")
+         ends_with("score"),
+         ends_with("diff")
          )

### 2023-08-15 1. Trajipaq_sentiment/2. Codes/Trajipaq_sentiment_an.Rmd
update

@@
 
+reg_WGI_arrival_diff_dico <- glm(formula = sentimapp_can_dico ~ wgi_arrival_diff, data = trajipaq, family = "binomial")
+reg_WGI_arrival_diff_dico_ctrl <- glm(formula = sentimapp_can_dico ~ wgi_arrival_diff  + age + education + sexe + origine + revenu + langue_mat + religion + statut_mat , data = trajipaq, family = "binomial")
+
+reg_WGI_arrival_diff_long <- lm(formula = sentimapp_can_long ~ wgi_arrival_diff, data = trajipaq)
+reg_WGI_arrival_diff_long_ctrl <- lm(formula = sentimapp_can_long ~ wgi_arrival_diff  + age + education + sexe + origine + revenu + langue_mat + religion + statut_mat , data = trajipaq)
+
 summary(reg_WGI_can_arrival_dico) # p = 0.000116 ***
@@
 
+summary(reg_WGI_arrival_diff_dico)
+summary(reg_WGI_arrival_diff_dico_ctrl)
+summary(reg_WGI_arrival_diff_long)
+summary(reg_WGI_arrival_diff_long_ctrl)
+
 # Au moment de l'enquête (like)
@@
 reg_WGI_can_2021_long_ctrl <- lm(formula = sentimapp_can_long ~ wgi_origin_2021_score  + age + education + sexe + origine + revenu + langue_mat + religion + statut_mat , data = trajipaq) #VD2 contrôlée par CSD
+
+reg_WGI_2021_diff_dico <- glm(formula = sentimapp_can_dico ~ wgi_2021_diff, data = trajipaq, family = "binomial")
+reg_WGI_2021_diff_dico_ctrl <- glm(formula = sentimapp_can_dico ~ wgi_2021_diff  + age + education + sexe + origine + revenu + langue_mat + religion + statut_mat , data = trajipaq, family = "binomial")
+
+reg_WGI_2021_diff_long <- lm(formula = sentimapp_can_long ~ wgi_2021_diff, data = trajipaq)
+reg_WGI_2021_diff_long_ctrl <- lm(formula = sentimapp_can_long ~ wgi_2021_diff  + age + education + sexe + origine + revenu + langue_mat + religion + statut_mat , data = trajipaq)
 
@@
   # Seul l'âge [50;99] est significative p = 0.0210 et Origin (Océanie p = 0.0369)
+
+summary(reg_WGI_2021_diff_dico)
+summary(reg_WGI_2021_diff_dico_ctrl)
+summary(reg_WGI_2021_diff_long)
+summary(reg_WGI_2021_diff_long_ctrl)
+
 ```
@@
   labs(
-    title = "Sentiment d'appartenance au Canada (dico) selon le score \nde gouvernementalité du pays d'origine au moment de l'immigraition", 
+    title = "Sentiment d'appartenance au Canada (dico) selon le score \nde gouvernementalité du pays d'origine au moment de l'immigration", 
     x = "Score de gouvernementalité du pays d'origine", 
@@
   labs(
-    title = "Sentiment d'appartenance au Canada (long) selon le score \nde gouvernementalité du pays d'origine au moment de l'immigraition", 
+    title = "Sentiment d'appartenance au Canada (long) selon le score \nde gouvernementalité du pays d'origine au moment de l'immigration", 
     x = "Score de gouvernementalité du pays d'origine", 
@@
   labs(
-    title = "Sentiment d'appartenance au Canada (long) selon le score \nde gouvernementalité du pays d'origine au moment de l'immigraition", 
+    title = "Sentiment d'appartenance au Canada (dico) selon le score \nde gouvernementalité du pays d'origine au moment de l'enquête", 
     x = "Score de gouvernementalité du pays d'origine", 
@@
   labs(
-    title = "Sentiment d'appartenance au Canada (long) selon le score \nde gouvernementalité du pays d'origine au moment de l'immigraition", 
+    title = "Sentiment d'appartenance au Canada (long) selon le score \nde gouvernementalité du pays d'origine au moment de l'enquête",
     x = "Score de gouvernementalité du pays d'origine", 
+    y = "Sentiment d'appartenance au Canada") +
+  theme_bw()
+
+ggplot(trajipaq %>% filter(!is.na(wgi_2021_diff), !is.na(sentimapp_can_dico)), aes(x = wgi_arrival_difffre, y = sentimapp_can_dico)) +
+  stat_smooth(method = "glm",
+              formula = "y ~ x",
+              color = "pink", 
+              se = FALSE) +
+  labs(
+    title = "Sentiment d'appartenance au Canada (dico) selon la différence de score \nde gouvernementalité du pays d'origine et du Canada au moment de l'immigraition", 
+    x = "Écart du score de gouvernementalité du pays d'origine et du Canada", 
+    y = "Sentiment d'appartenance au Canada") +
+  theme_bw()
+
+ggplot(trajipaq %>% filter(!is.na(wgi_arrival_diff), !is.na(sentimapp_can_long)), aes(x = wgi_2021_diff, y = sentimapp_can_long)) +
+  stat_smooth(method = "lm",
+              formula = "y ~ x",
+              color = "pink", 
+              se = FALSE) +
+  labs(
+    title = "Sentiment d'appartenance au Canada (dico) selon la différence de score \nde gouvernementalité du pays d'origine et du Canada au moment de l'enquête", 
+    x = "Écart du score de gouvernementalité du pays d'origine et du Canada", 
     y = "Sentiment d'appartenance au Canada") +

### 2023-08-15 1. Trajipaq_sentiment/2. Codes/Trajipaq_sentiment_an.Rmd
update

--- 2023-08-12 1. Trajipaq_sentiment/2. Codes/Trajipaq_sentiment_an.Rmd
@@
 ```
+
+# Uniformisation de la DB (NA)
+
+```{r}
+
+v_CSD <- c("age", "education", "revenu", "sexe", "origine", "langue_mat", "statut_mat", "religion")
+
+trajipaq <- trajipaq %>%
+  filter(rowSums(!is.na(across(all_of(v_CSD)))) == length(v_CSD))
+```
+
 # WGI X Canada
@@
 
-reg_WGI_can_arrival_dico <- glm(formula = sentimapp_can_dico ~ wgi_origin_arrival_score, data = trajipaq, family = "binomial") # VD2 dichotomique
-reg_WGI_can_arrival_dico_ctrl <- glm(formula = sentimapp_can_dico ~ wgi_origin_arrival_score  + age + education + sexe + origine + revenu + langue_mat + religion + statut_mat , data = trajipaq, family = "binomial") #VD2 contrôlée par CSD
+reg_WGI_can_arrival <- glm(formula = sentimapp_can_dico ~ wgi_origin_arrival_score, data = trajipaq, family = "binomial") # VD2 dichotomique
+reg_WGI_can_arrival_ctrl <- glm(formula = sentimapp_can_dico ~ wgi_origin_arrival_score  + age + education + sexe + origine + revenu + langue_mat + religion + statut_mat , data = trajipaq, family = "binomial") #VD2 contrôlée par CSD
 
-reg_WGI_can_arrival_long <- lm(formula = sentimapp_can_long ~ wgi_origin_arrival_score, data = trajipaq) # VD2 dichotomique
-reg_WGI_can_arrival_long_ctrl <- lm(formula = sentimapp_can_long ~ wgi_origin_arrival_score  + age + education + sexe + origine + revenu + langue_mat + religion + statut_mat , data = trajipaq) #VD2 contrôlée par CSD
+reg_WGI_arrival_diff <- glm(formula = sentimapp_can_dico ~ wgi_arrival_diff, data = trajipaq, family = "binomial")
+reg_WGI_arrival_diff_ctrl <- glm(formula = sentimapp_can_dico ~ wgi_arrival_diff  + age + education + sexe + origine + revenu + langue_mat + religion + statut_mat , data = trajipaq, family = "binomial")
 
-summary(reg_WGI_can_arrival_dico) # p = 0.000116 ***
-summary(reg_WGI_can_arrival_dico_ctrl) # p = 0.0203 *
-summary(reg_WGI_can_arrival_long) # p = 8.7e-07 ***
-summary(reg_WGI_can_arrival_long_ctrl) # p = 0.00678 **
+summary(reg_WGI_can_arrival) # p = 0.000116 ***
+  # Plus le score WGI du pays d'origine au moment de l'immigration est élevé, plus le sentiment d'appartenance au Canada est faible. Relation négative.
+
+summary(reg_WGI_can_arrival_ctrl) # p = 0.0203 *
+  # aucune variable de contrôle significative (age [50;99] p = 0.0894)
+
+summary(reg_WGI_arrival_diff)
+summary(reg_WGI_arrival_diff_ctrl)
 
@@
 
-reg_WGI_can_2021_dico <- glm(formula = sentimapp_can_dico ~ wgi_origin_2021_score , data = trajipaq, family = "binomial", na.action = na.pass) ## VD2 dichotomique
-reg_WGI_can_2021_dico_ctrl <- glm(formula = sentimapp_can_dico ~ wgi_origin_2021_score  + age + education + sexe + origine + revenu + langue_mat + religion + statut_mat , data = trajipaq, family = "binomial")#VD2 contrôlée par CSD
+reg_WGI_can_2021 <- glm(formula = sentimapp_can_dico ~ wgi_origin_2021_score , data = trajipaq, family = "binomial") ## VD2 dichotomique
+reg_WGI_can_2021_ctrl <- glm(formula = sentimapp_can_dico ~ wgi_origin_2021_score  + age + education + sexe + origine + revenu + langue_mat + religion + statut_mat , data = trajipaq, family = "binomial")#VD2 contrôlée par CSD
 
-reg_WGI_can_2021_long <- lm(formula = sentimapp_can_long ~ wgi_origin_2021_score, data = trajipaq) # VD2 dichotomique
-reg_WGI_can_2021_long_ctrl <- lm(formula = sentimapp_can_long ~ wgi_origin_2021_score  + age + education + sexe + origine + revenu + langue_mat + religion + statut_mat , data = trajipaq) #VD2 contrôlée par CSD
+reg_WGI_2021_diff <- glm(formula = sentimapp_can_dico ~ wgi_2021_diff, data = trajipaq, family = "binomial")
+reg_WGI_2021_diff_ctrl <- glm(formula = sentimapp_can_dico ~ wgi_2021_diff  + age + education + sexe + origine + revenu + langue_mat + religion + statut_mat , data = trajipaq, family = "binomial")
 
-summary(reg_WGI_can_2021_dico) # 2.38e-05 ***
-summary(reg_WGI_can_2021_dico_ctrl) # p = 0.03426 *
-summary(reg_WGI_can_2021_long) # p = 5.5e-08 ***
-summary(reg_WGI_can_2021_long_ctrl) # p = 0.0124 *
+summary(reg_WGI_can_2021) # 2.38e-05 ***
+  # Plus le score WGI du pays d'origine au moment de l'enquête est élevé, plus le sentiment d'appartenance au Canada est faible. Relation négative.
+
+summary(reg_WGI_can_2021_ctrl) # p = 0.03426 *
+  # Seul l'âge [50;99] estr significative p = 0.04311
+
+summary(reg_WGI_2021_diff)
+summary(reg_WGI_2021_diff_ctrl)
+```
+
+## Visualisation
+
+```{r}
+
+ggplot(trajipaq %>% filter(!is.na(wgi_origin_arrival_score), !is.na(sentimapp_can)), aes(x = wgi_origin_arrival_score, y = sentimapp_can_dico)) +
+  stat_smooth(method = "glm",
+              formula = "y ~ x",
+              color = "pink", 
+              se = FALSE,
+              method.args = list(family = binomial)) +
+  labs(
+    title = "Sentiment d'appartenance au Canada (dico) selon le score \nde gouvernementalité du pays d'origine au moment de l'immigration", 
+    x = "Score de gouvernementalité du pays d'origine", 
+    y = "Sentiment d'appartenance au Canada") +
+  theme_bw()
+
+ggplot(trajipaq %>% filter(!is.na(wgi_origin_2021_score), !is.na(sentimapp_can)), aes(x = wgi_origin_2021_score, y = sentimapp_can_dico)) +
+  stat_smooth(method = "glm",
+              formula = "y ~ x",
+              color = "pink", 
+              se = FALSE) +
+  labs(
+    title = "Sentiment d'appartenance au Canada (dico) selon le score \nde gouvernementalité du pays d'origine au moment de l'enquête", 
+    x = "Score de gouvernementalité du pays d'origine", 
+    y = "Sentiment d'appartenance au Canada") +
+  theme_bw()
+
+ggplot(trajipaq %>% filter(!is.na(wgi_2021_diff), !is.na(sentimapp_can)), aes(x = wgi_arrival_diff, y = sentimapp_can_dico)) +
+  stat_smooth(method = "glm",
+              formula = "y ~ x",
+              color = "pink", 
+              se = FALSE) +
+  labs(
+    title = "Sentiment d'appartenance au Canada (dico) selon la différence de score \nde gouvernementalité du pays d'origine et du Canada au moment de l'immigraition", 
+    x = "Écart du score de gouvernementalité du pays d'origine et du Canada", 
+    y = "Sentiment d'appartenance au Canada") +
+  theme_bw()
+
 ```
@@
 stargazer(
-  reg_WGI_can_2021_dico,
-  reg_WGI_can_2021_dico_ctrl,
-  reg_WGI_can_2021_long,
+  reg_WGI_can_2021,
+  reg_WGI_can_2021_ctrl,
+  title = "Results", 
+  align = TRUE, 
+  type = "text"
+)
+
+stargazer(
+  reg_WGI_can_arrival,
+  reg_WGI_can_arrival_ctrl,
+  title = "Results", 
+  align = TRUE, 
+  type = "text"
+)
+
+ stargazer(
+  reg_WGI_arrival_diff,
+  reg_WGI_arrival_diff_ctrl,
+  title = "Results", 
+  align = TRUE, 
+  type = "text"
+)
+
+ stargazer(
+  reg_WGI_2021_diff,
+  reg_WGI_2021_diff_ctrl,
   title = "Results", 
@@
 
-freq(trajipaq$sentimapp_can_dico)
+freq(trajipaq$sentimapp_can)
 freq(trajipaq$wgi_origin_arrival_score)
@@
 ## WGI X Québec
-
-```{r}
-# Au moment de l'immigration
-
-reg__qc1 <- lm(formula = VD1  ~ vae_arrival, data = trajipaq) # VD1
-reg_WGI_qc1_ctrl <- lm(formula = VD1 ~ vae_arrival  + age + education + sexe + origine + revenu + langue_mat + religion + statut_mat , data = trajipaq) #VD1 contrôlée par CSD
-
-summary(reg_WGI_qc1)
-summary(reg_WGI_qc1_ctrl)
-
-# Au moment de l'enquête
-
-reg_WGI_qc2 <- lm(formula = VD1 ~ vae_2021 , data = trajipaq) # VD1
-reg_WGI_qc2_ctrl <- lm(formula = VD1 ~ vae_2021  + age + education + sexe + origine + revenu + langue_mat + religion + statut_mat , data = trajipaq) # VD1 contrôlée par CSD
-
-summary(reg_WGI_qc2)
-summary(reg_WGI_qc2_ctrl)
-
-# À l'arrivée et au moment de l'enquête
-
-reg_WGI_qc3 <- lm(formula = VD1 ~ vae_2021 + vae_arrival , data = trajipaq) # VD1
-reg_WGI_qc3_ctrl <- lm(formula = VD1 ~ vae_2021 + vae_arrival + age + education + sexe + origine + revenu + langue_mat + religion + statut_mat , data = trajipaq)# VD1 contrôlée par CSD
-
-summary(reg_WGI_qc3)
-summary(reg_WGI_qc3_ctrl)
-```
-
-## WGI X can_qc
-
-```{r}
-reg_WGI_can_origin <- lm(formula = sentimapp_can_origin_N ~ wgi_origin_arrival_score, data = trajipaq, family = "binomial") # VD2 dichotomique
-
-reg_WGI_can_origin_ctrl <- lm(formula = sentimapp_can_origin_N ~ wgi_origin_arrival_score  + age + education + sexe + origine + revenu + langue_mat + religion + statut_mat , data = trajipaq, family = "binomial") #VD2 contrôlée par CSD
-
-summary(reg_WGI_can_origin)
-summary(reg_WGI_can_origin_ctrl)
-
-stargazer(reg_WGI_can2, reg_WGI_qc2, reg_WGI_can2_ctrl, reg_WGI_qc2_ctrl, title="Results",align=TRUE, type = "text")
-
-```
-
-```{r}
-cor.test(trajipaq$vae_arrival, trajipaq$VD2)
-
-ggplot(trajipaq %>% filter(!is.na(vae_arrival), !is.na(VD1)), aes(x = vae_arrival, y = VD1)) +
-  geom_smooth(method = "glm",
-              formula = "y ~ x",
-              color = "pink", 
-              se = FALSE) +
-  labs(
-    title = "Sentiment d'appartenance au Québec et indice de gouvernementalité du pays d'origine, au moment de l'immigration", 
-    x = "indice de gouvernementalité du pays d'origine au moment de l'immigration", 
-    y = "Sentiment d'appartenance au Québec") +
-  theme_bw()
-
-ggplot(trajipaq %>% filter(!is.na(vae_arrival), !is.na(VD2)), aes(x = vae_arrival, y = VD1)) +
-  stat_smooth(method = "glm",
-              formula = "y ~ x",
-              color = "pink", 
-              se = FALSE) +
-  labs(
-    title = "Sentiment d'appartenance au Canada et indice de gouvernementalité du pays d'origine, au moment de l'immigration", 
-    x = "indice de gouvernementalité du pays d'origine au moment de l'immigration", 
-    y = "Sentiment d'appartenance au Canada") +
-  theme_bw()
-```
-
-```{r}
-
-chisq.test(trajipaq$VD3, trajipaq$VD1_f)
-
-```

### 2023-08-16 1. Trajipaq_sentiment/2. Codes/Trajipaq_sentiment_an.Rmd
update

@@
 
-v_CSD <- c("age", "education", "revenu", "sexe", "origine", "langue_mat", "statut_mat", "religion")
+v_noNA <- c("age", "education", "revenu", "sexe", "origine", "langue_mat", "statut_mat", "religion", "sentimapp_can_dico", "wgi_origin_arrival_score", "wgi_arrival_diff", "wgi_origin_2021_score", "wgi_2021_diff")
 
 trajipaq <- trajipaq %>%
-  filter(rowSums(!is.na(across(all_of(v_CSD)))) == length(v_CSD))
+  filter(rowSums(!is.na(across(all_of(v_noNA)))) == length(v_noNA))
 ```
@@
  stargazer(
+  reg_WGI_can_arrival,
+  reg_WGI_can_arrival_ctrl,
+  reg_WGI_arrival_diff,
+  reg_WGI_arrival_diff_ctrl,
+  reg_WGI_can_2021,
+  reg_WGI_can_2021_ctrl,
+  reg_WGI_2021_diff,
+  reg_WGI_2021_diff,
+  reg_WGI_2021_diff_ctrl,
+  title = "Results", 
+  align = TRUE, 
+  type = "text"
+)
+
+stargazer(
   reg_WGI_can_2021,

### 2023-08-16 1. Trajipaq_sentiment/2. Codes/Trajipaq_sentiment_an.Rmd
update

@@
 
-reg_WGI_can_2021 <- glm(formula = sentimapp_can_dico ~ wgi_origin_2021_score , data = trajipaq, family = "binomial") ## VD2 dichotomique
+reg_WGI_can_2021 <- glm(formula = sentimapp_can_dico ~ wgi_origin_2021_score, data = trajipaq, family = "binomial") ## VD2 dichotomique
 reg_WGI_can_2021_ctrl <- glm(formula = sentimapp_can_dico ~ wgi_origin_2021_score  + age + education + sexe + origine + revenu + langue_mat + religion + statut_mat , data = trajipaq, family = "binomial")#VD2 contrôlée par CSD
@@
   reg_WGI_can_2021_ctrl,
-  reg_WGI_2021_diff,
   reg_WGI_2021_diff,

### 2023-08-16 1. Trajipaq_sentiment/2. Codes/Trajipaq_sentiment_WGI.Rmd
craétion Trajipaq_sentiment_WGI

---
title: "Trajipaq_sentiment_WGI"
author: "Vincent Bertyrand-Huneault"
date: "2023-08-16"
output: pdf_document
---

```{r}
library(summarytools)
library(tidyverse)
library(ggthemes)
library(haven)
library(stargazer)
library(gtsummary)

trajipaq <- read_csv("../3. Data/Trajipaq_sentiment_cr.csv")
#view(trajipaq)
```

# Uniformisation de la DB (NA)
```{r}
v_noNA <- c("age", "education", "revenu", "sexe", "origine", "langue_mat", "statut_mat", "religion", "sentimapp_can", "wgi_origin_arrival_score", "wgi_arrival_diff", "wgi_origin_2021_score", "wgi_2021_diff")

trajipaq <- trajipaq %>%
  filter(rowSums(!is.na(across(all_of(v_noNA)))) == length(v_noNA))
```

# Régressions logistiques
```{r}
#Au moment de l'immigration

reg1 <- glm(formula = sentimapp_can ~ wgi_origin_arrival_score, data = trajipaq, family = binomial)
reg2 <- glm(formula = sentimapp_can ~ wgi_origin_arrival_score  + age + education + sexe + origine + revenu + langue_mat + religion + statut_mat , data = trajipaq, family = binomial)
reg3 <- glm(formula = sentimapp_can ~ wgi_arrival_diff, data = trajipaq, family = binomial)
reg4 <- glm(formula = sentimapp_can ~ wgi_arrival_diff  + age + education + sexe + origine + revenu + langue_mat + religion + statut_mat , data = trajipaq, family = binomial)

# Au moment de l'enquête (like)

reg5 <- glm(formula = sentimapp_can ~ wgi_origin_2021_score, data = trajipaq, family = binomial)
reg6 <- glm(formula = sentimapp_can ~ wgi_origin_2021_score  + age + education + sexe + origine + revenu + langue_mat + religion + statut_mat , data = trajipaq, family = binomial)
reg7 <- glm(formula = sentimapp_can ~ wgi_2021_diff, data = trajipaq, family = binomial)
reg8 <- glm(formula = sentimapp_can ~ wgi_2021_diff  + age + education + sexe + origine + revenu + langue_mat + religion + statut_mat , data = trajipaq, family = binomial)
""
reg_order <- c("wgi_origin_arrival_score", "wgi_arrival_diff", "wgi_origin_2021_score", "wgi_2021_diff")

stargazer(
  reg1,
  reg2,
  reg3,
  reg4,
  reg5,
  reg6,
  reg7,
  reg8,
  title = "Results", 
  align = TRUE, 
  type = "text",
  order = reg_order
)
```

# Visualisations

## Origin_arrival

```{r}
ggplot(trajipaq, aes(x = wgi_origin_arrival_score, y = sentimapp_can)) +
  stat_smooth(method = "glm",
              formula = "y ~ x",
              color = "pink", 
              se = FALSE,
              method.args = list(family = binomial)) +
  labs(
    title = "Sentiment d'appartenance au Canada selon le score \nde gouvernementalité du pays d'origine au moment de l'immigration", 
    x = "Score de gouvernementalité du pays d'origine", 
    y = "Sentiment d'appartenance au Canada") +
  theme_bw()
```

## Origin_2021
```{r}
ggplot(trajipaq %>% filter(!is.na(wgi_origin_2021_score), !is.na(sentimapp_can)), aes(x = wgi_origin_2021_score, y = sentimapp_can)) +
  stat_smooth(method = "glm",
              formula = "y ~ x",
              color = "pink", 
              se = FALSE) +
  labs(
    title = "Sentiment d'appartenance au Canada selon le score \nde gouvernementalité du pays d'origine au moment de l'enquête", 
    x = "Score de gouvernementalité du pays d'origine", 
    y = "Sentiment d'appartenance au Canada") +
  theme_bw()
```

##2021_diff
```{r}
ggplot(trajipaq %>% filter(!is.na(wgi_2021_diff), !is.na(sentimapp_can)), aes(x = wgi_2021_diff, y = sentimapp_can)) +
  stat_smooth(method = "glm",
              formula = "y ~ x",
              color = "pink", 
              se = FALSE) +
  labs(
    title = "Sentiment d'appartenance au Canada selon la différence de score \nde gouvernementalité du pays d'origine et du Canada au moment de l'immigraition", 
    x = "Écart du score de gouvernementalité du pays d'origine et du Canada", 
    y = "Sentiment d'appartenance au Canada") +
  theme_bw()

```

## Avrrival_diff
```{r}
ggplot(trajipaq %>% filter(!is.na(wgi_arrival_diff), !is.na(sentimapp_can)), aes(x = wgi_arrival_diff, y = sentimapp_can)) +
  stat_smooth(method = "glm",
              formula = "y ~ x",
              color = "pink", 
              se = FALSE) +
  labs(
    title = "Sentiment d'appartenance au Canada selon la différence de score \nde gouvernementalité du pays d'origine et du Canada au moment de l'immigraition", 
    x = "Écart du score de gouvernementalité du pays d'origine et du Canada", 
    y = "Sentiment d'appartenance au Canada") +
  theme_bw()
```

### 2023-08-16 1. Trajipaq_sentiment/2. Codes/Trajipaq_sentiment_an.Rmd
craétion Trajipaq_sentiment_WGI

@@
 
-reg_WGI_can_arrival <- glm(formula = sentimapp_can_dico ~ wgi_origin_arrival_score, data = trajipaq, family = "binomial") # VD2 dichotomique
-reg_WGI_can_arrival_ctrl <- glm(formula = sentimapp_can_dico ~ wgi_origin_arrival_score  + age + education + sexe + origine + revenu + langue_mat + religion + statut_mat , data = trajipaq, family = "binomial") #VD2 contrôlée par CSD
-
-reg_WGI_arrival_diff <- glm(formula = sentimapp_can_dico ~ wgi_arrival_diff, data = trajipaq, family = "binomial")
-reg_WGI_arrival_diff_ctrl <- glm(formula = sentimapp_can_dico ~ wgi_arrival_diff  + age + education + sexe + origine + revenu + langue_mat + religion + statut_mat , data = trajipaq, family = "binomial")
-
-summary(reg_WGI_can_arrival) # p = 0.000116 ***
-  # Plus le score WGI du pays d'origine au moment de l'immigration est élevé, plus le sentiment d'appartenance au Canada est faible. Relation négative.
-
-summary(reg_WGI_can_arrival_ctrl) # p = 0.0203 *
-  # aucune variable de contrôle significative (age [50;99] p = 0.0894)
-
-summary(reg_WGI_arrival_diff)
-summary(reg_WGI_arrival_diff_ctrl)
+reg1 <- glm(formula = sentimapp_can_dico ~ wgi_origin_arrival_score, data = trajipaq, family = binomial)
+reg2 <- glm(formula = sentimapp_can_dico ~ wgi_origin_arrival_score  + age + education + sexe + origine + revenu + langue_mat + religion + statut_mat , data = trajipaq, family = binomial)
+reg3 <- glm(formula = sentimapp_can_dico ~ wgi_arrival_diff, data = trajipaq, family = binomial)
+reg4 <- glm(formula = sentimapp_can_dico ~ wgi_arrival_diff  + age + education + sexe + origine + revenu + langue_mat + religion + statut_mat , data = trajipaq, family = binomial)
 
@@
 
-reg_WGI_can_2021 <- glm(formula = sentimapp_can_dico ~ wgi_origin_2021_score, data = trajipaq, family = "binomial") ## VD2 dichotomique
-reg_WGI_can_2021_ctrl <- glm(formula = sentimapp_can_dico ~ wgi_origin_2021_score  + age + education + sexe + origine + revenu + langue_mat + religion + statut_mat , data = trajipaq, family = "binomial")#VD2 contrôlée par CSD
+reg5 <- glm(formula = sentimapp_can_dico ~ wgi_origin_2021_score, data = trajipaq, family = binomial)
+reg6 <- glm(formula = sentimapp_can_dico ~ wgi_origin_2021_score  + age + education + sexe + origine + revenu + langue_mat + religion + statut_mat , data = trajipaq, family = binomial)
+reg7 <- glm(formula = sentimapp_can_dico ~ wgi_2021_diff, data = trajipaq, family = binomial)
+reg8 <- glm(formula = sentimapp_can_dico ~ wgi_2021_diff  + age + education + sexe + origine + revenu + langue_mat + religion + statut_mat , data = trajipaq, family = binomial)
 
-reg_WGI_2021_diff <- glm(formula = sentimapp_can_dico ~ wgi_2021_diff, data = trajipaq, family = "binomial")
-reg_WGI_2021_diff_ctrl <- glm(formula = sentimapp_can_dico ~ wgi_2021_diff  + age + education + sexe + origine + revenu + langue_mat + religion + statut_mat , data = trajipaq, family = "binomial")
+stargazer(
+  reg1,
+  reg2,
+  reg3,
+  reg4,
+  reg5,
+  reg6,
+  reg7,
+  reg8,
+  title = "Results", 
+  align = TRUE, 
+  type = "text"
+)
 
-summary(reg_WGI_can_2021) # 2.38e-05 ***
-  # Plus le score WGI du pays d'origine au moment de l'enquête est élevé, plus le sentiment d'appartenance au Canada est faible. Relation négative.
-
-summary(reg_WGI_can_2021_ctrl) # p = 0.03426 *
-  # Seul l'âge [50;99] estr significative p = 0.04311
-
-summary(reg_WGI_2021_diff)
-summary(reg_WGI_2021_diff_ctrl)
 ```
@@
 
-```{r}
- stargazer(
-  reg_WGI_can_arrival,
-  reg_WGI_can_arrival_ctrl,
-  reg_WGI_arrival_diff,
-  reg_WGI_arrival_diff_ctrl,
-  reg_WGI_can_2021,
-  reg_WGI_can_2021_ctrl,
-  reg_WGI_2021_diff,
-  reg_WGI_2021_diff_ctrl,
-  title = "Results", 
-  align = TRUE, 
-  type = "text"
-)
+summary(reg_WGI_can_arrival) # p = 0.000116 ***
+  # Plus le score WGI du pays d'origine au moment de l'immigration est élevé, plus le sentiment d'appartenance au Canada est faible. Relation négative.
 
-stargazer(
-  reg_WGI_can_2021,
-  reg_WGI_can_2021_ctrl,
-  title = "Results", 
-  align = TRUE, 
-  type = "text"
-)
+summary(reg_WGI_can_arrival_ctrl) # p = 0.0203 *
+  # aucune variable de contrôle significative (age [50;99] p = 0.0894)
 
-stargazer(
-  reg_WGI_can_arrival,
-  reg_WGI_can_arrival_ctrl,
-  title = "Results", 
-  align = TRUE, 
-  type = "text"
-)
+summary(reg_WGI_can_2021) # 2.38e-05 ***
+  # Plus le score WGI du pays d'origine au moment de l'enquête est élevé, plus le sentiment d'appartenance au Canada est faible. Relation négative.
 
- stargazer(
-  reg_WGI_arrival_diff,
-  reg_WGI_arrival_diff_ctrl,
-  title = "Results", 
-  align = TRUE, 
-  type = "text"
-)
-
- stargazer(
-  reg_WGI_2021_diff,
-  reg_WGI_2021_diff_ctrl,
-  title = "Results", 
-  align = TRUE, 
-  type = "text"
-)
-
- # reg_WGI_can_2021_dico,
- # reg_WGI_can_2021_dico_ctrl,
-
-freq(trajipaq$sentimapp_can)
-freq(trajipaq$wgi_origin_arrival_score)
-freq(trajipaq$sentimapp_can_long)
-freq(trajipaq$wgi_origin_2021_score)
+summary(reg_WGI_can_2021_ctrl) # p = 0.03426 *
+  # Seul l'âge [50;99] estr significative p = 0.04311
 

### 2023-08-16 1. Trajipaq_sentiment/2. Codes/Trajipaq_sentiment_cr.Rmd
update
craétion Trajipaq_sentiment_WGI

@@
   mutate(sentimapp_can_long = case_when( # Création de la variable "VD2".
-    Q39E == 1 ~ 1,
-    Q39E == 2 ~ 2,
-    Q39E == 3 ~ 3,
-    Q39E == 4 ~ 4,
+    Q39E == 1 ~ 4,
+    Q39E == 2 ~ 3,
+    Q39E == 3 ~ 2,
+    Q39E == 4 ~ 1,
     Q39E == 98 ~ NA,
@@
   trajipaq %>%
-  mutate(sentimapp_can_dico = case_when( # Création de la variable "VD2".
+  mutate(sentimapp_can = case_when( # Création de la variable "VD2".
     Q39E == 1 ~ 1,
@@
 freq(trajipaq$sentimapp_can_long)
-freq(trajipaq$sentimapp_can_dico)
+freq(trajipaq$sentimapp_can)
 ```
@@
   mutate(sentimapp_can_origin = case_when( # Création de la variable "VD4".
-    sentimapp_can_dico == 1 & sentimapp_origine_dico == 1 ~ "Integre",
-    sentimapp_can_dico == 1 & sentimapp_origine_dico == 0 ~ "Assimile",
-    sentimapp_can_dico == 0 & sentimapp_origine_dico == 1 ~ "Separe",
-    sentimapp_can_dico == 0 & sentimapp_origine_dico == 0 ~ "Marginalise"
+    sentimapp_can == 1 & sentimapp_can == 1 ~ "Integre",
+    sentimapp_can == 1 & sentimapp_can == 0 ~ "Assimile",
+    sentimapp_can == 0 & sentimapp_can == 1 ~ "Separe",
+    sentimapp_can == 0 & sentimapp_can == 0 ~ "Marginalise"
   ))
@@
   mutate(sentimapp_can_origin_N = case_when( # Création de la variable "VD4".
-    sentimapp_can_dico == 1 & sentimapp_origine_dico == 1 ~ "1",
-    sentimapp_can_dico == 1 & sentimapp_origine_dico == 0 ~ "2",
-    sentimapp_can_dico == 0 & sentimapp_origine_dico == 1 ~ "3",
-    sentimapp_can_dico == 0 & sentimapp_origine_dico == 0 ~ "4"
+    sentimapp_can == 1 & sentimapp_can == 1 ~ "1",
+    sentimapp_can == 1 & sentimapp_can == 0 ~ "2",
+    sentimapp_can == 0 & sentimapp_can == 1 ~ "3",
+    sentimapp_can == 0 & sentimapp_can == 0 ~ "4"
   ))
@@
 ```{r}
+
+trajipaq$age <- as.factor(trajipaq$age)
+trajipaq$education <- as.factor(trajipaq$education)
+trajipaq$sexe <- as.factor(trajipaq$sexe)
+trajipaq$origine <- as.factor(trajipaq$origine)
+trajipaq$revenu <- as.factor(trajipaq$revenu)
+trajipaq$langue_mat <- as.factor(trajipaq$langue_mat)
+trajipaq$religion <- as.factor(trajipaq$religion)
+trajipaq$statut_mat <- as.factor(trajipaq$statut_mat)
+
 wgi_origin_2021 <- filter(wgidataset, year == 2021) # Origine en 2021
@@
 
+trajipaq <- trajipaq %>%
+  mutate(wgi_arrival_diff = wgi_origin_arrival_score - wgi_can_arrival_score) %>%
+  mutate(wgi_2021_diff = wgi_origin_2021_score - wgi_can_2021_score)
+
 ```
@@
          "sentimapp_QC",
-         "sentimapp_can_dico",
+         "sentimapp_can",
          "sentimapp_can_long",
